### 2015-07-14 writeup.tex
add latex.

\documentclass[11pt]{article}

\begin{document}

This is my first LaTeX document.

\end{document}

### 2015-07-14 writeup.tex
Add tex

@@
 
-This is my first LaTeX document.
+\section{Design}
+My program is a state machine which uses getopt to parse through commands. When it reads the "q" option, it puts itself in APPEND mode. This causes it to iterate through
+the further arguments to the program and parse them as input files. When it reads the "x" option, it puts itself in EXTRACT mode. This causes it to read the arfile argument
+and create files from that archive. Etc etc etc.
+
+\section{Worklog}
+July 11th 10:00am - 11:00am: Read through assignment details, coded skeleton and getopt stuff. \\
+
+July 11th 8:00pm - 9:00pm: Got getopt stuff to work, began with code to read in files and create an archive file. \\
+
+July 12th 8:00am - 11:00am: Got program to succesfully read in files and save them to a custom archive file (NOT the ar-specified archive file format) \\
+
+July 13th 2:00pm - 4:00pm: Got program to succesfully read in files and save them to an archive file somewhat reminiscent of the ar-specified format. \\
+
+July 13th 7:30pm - 12:00am: Got program to succesfully read in files, save them to the ar-specified format, and print their contents kind of. Still lots to do. \\
+
+\section{Challenges}
+It was difficult for me to figure out how the ar-specified file format worked, until I thought to Google it. I was surprised that there was a good wikipedia page about it. \\
+
+\section{Questions}
+Point of assignment: To get us used to the level of difficulty of programming assignments, to give an introduction to system programming and file IO. \\
+Correctness testing: By running the ar command and checking for exact duplication of behavior. \\
+Learn: Start earlier! Also lots about IO and string manipulation.\\
 

### 2015-07-19 A2/myar.c
Add file deletion feature.

@@ -531,80 +531,73 @@ int main(int argc, char** argv) {
 				exit(EXIT_FAILURE);
 			}
 			
-			char* ar_ptr = strchr(ar_buf, '\n') + 1;
-			char* ar_start = ar_ptr;
-					
-			char* ar_end = ar_buf + ar_size;
+			char temp_arfile[strlen(ar_buf)];
+			strcpy(temp_arfile, "!<arch>\n");
 			
-				while(ar_ptr < ar_end) 
-				{	
-					//printf("ar_ptr offset: %ld\t", ar_ptr - ar_buf);
-					//printf("ar_ptr pointing to char: %d, %c\t", (int) *ar_ptr, *ar_ptr);
-					
-					/*if (*ar_ptr == '\n')
-					{
-						ar_ptr++;
-					}*/
-					
-					int to_delete = 0;
-					int to_continue = 0;
+			char* ar_ptr = ar_buf + 8; // Start at first entry description
+			char* ar_end = ar_buf + ar_size;
+			int to_add;
+			
+			while(ar_ptr < ar_end) 
+			{	
+				to_add = 1;
+            
+				int entry_name_str_len = strchr(&ar_ptr[0], ' ') - &ar_ptr[0];
+				if (entry_name_str_len > 16)	entry_name_str_len = 16;
+				char entry_name[entry_name_str_len + 1];
+				strncpy(entry_name, ar_ptr, entry_name_str_len);
+				entry_name[entry_name_str_len] = '\0';
+            
+				int entry_size_str_len = strchr(&ar_ptr[48], ' ') - &ar_ptr[48];
+				if (entry_size_str_len > 10)	entry_size_str_len = 10;
+				char entry_size_str[entry_size_str_len + 1];
+				strncpy(entry_size_str, &ar_ptr[48], entry_size_str_len);
+				entry_size_str[entry_size_str_len] = '\0';
+				int entry_size = atoi(entry_size_str);
 				
-					char* entry_start = strchr(ar_ptr, '\n') + 1;
-
-					int entry_name_str_len = strchr(&ar_ptr[0], ' ') - &ar_ptr[0];
-					if (entry_name_str_len > 16)	entry_name_str_len = 16;
-					char entry_name[entry_name_str_len + 1];
-					strncpy(entry_name, &ar_ptr[0], entry_name_str_len);
-					entry_name[entry_name_str_len] = '\0';
-
-					int entry_size_str_len = strchr(&ar_ptr[48], ' ') - &ar_ptr[48];
[…]

### 2015-07-19 A2/myar.c
Add directory appending, for real this time.

@@ -9,6 +9,7 @@
 #include <sys/types.h>
 #include <time.h>
 #include <utime.h>
+#include <dirent.h>
 
 #define MAX_ENTRY_NUMBER_SIZE 100
 #define OUT_DESCRIPTION_MAX_SIZE 200
@@ -602,8 +603,145 @@ int main(int argc, char** argv) {
 		}
 		
 		case ADD:
-		{
-			printf("adding all fiels in directory...\n");
+		{	
+			DIR *dir;
+			struct dirent *dir_struct;
+			
+			dir = opendir(".");
+			int ar_fd;
+			struct stat test;
+			
+			if (stat(ar_file, &test) != 0) 
+			{
+				if ((ar_fd = open(ar_file, O_WRONLY | O_CREAT | O_EXCL, S_IRUSR | S_IWUSR | S_IRGRP | S_IROTH)) == -1)
+				{
+					perror("exclusive open archive file: ");
+					exit(EXIT_FAILURE);
+				}
+				
+				char init_str[] = "!<arch>\n";
+				
+				if (write(ar_fd, init_str, sizeof(char) * strlen(init_str)) == -1)
+				{
+					perror("write to arfile: ");
+					exit(EXIT_FAILURE);
+				}
+				
+				printf("myar: creating archive %s\n", ar_file);
+				if (close(ar_fd) == -1)
+				{
+					perror("close arfile: ");
+					exit(EXIT_FAILURE);
+				}
+			}
+			
+			if ((ar_fd = open(ar_file, O_WRONLY | O_APPEND)) == -1)
+			{
+				perror("open archive file: ");
+				exit(EXIT_FAILURE);
+			}
+			
+			while ((dir_struct = readdir(dir)) != NULL) {
+				if (dir_struct->d_type == DT_REG && strcmp(dir_struct->d_name, ar_file) != 0 && strcmp(dir_struct->d_name, "myar") != 0) {
+					printf("file: %s\n", dir_struct->d_name);
+					
+					char* in_file = dir_struct->d_name;
+					int in_fd;
+
+					if ((in_fd = open(in_file, O_RDONLY)) == -1)
+					{
+						perror("open input file: ");
+						exit(EXIT_FAILURE);
+					}
+
+					int in_size;
+					if ((in_size = lseek(in_fd, 0, SEEK_END)) == -1 || lseek(in_fd, 0, SEEK_SET) == -1)
+					{
+						perror("lseek input file: ");
[…]

### 2015-07-19 A2/writeup.tex
Add changes to writeup.tex.

--- 2015-07-14 writeup.tex
@@
 the further arguments to the program and parse them as input files. When it reads the "x" option, it puts itself in EXTRACT mode. This causes it to read the arfile argument
-and create files from that archive. Etc etc etc.
+and create files from that archive. Etc etc etc. \\
 
@@
 
+July 17th 3:00pm - 1:00am: Fixed bugs in outputting archive file properly (\n-padding odd outputs). Spent a long time debugging delete option problem. \\
+
+July 18th 4:00pm-8:00pm: Fixed delete option problem. Checked for errors, fixed minor ones. Finished worklog. \\
+
 \section{Challenges}
@@
 
+I also spent a considerable amount of time on the delete code. I was having problems with my pointer references and knowing where I was in the file. I spent an overly long
+amount of time trying to fix a bug, then eventually rewrote the entire section of code in about 20 minutes. I should have done that sooner. I think that for me, it is much 
+easier to rewrite code that involves strings and pointers than to debug existing code. \\
+
 \section{Questions}
-Point of assignment: To get us used to the level of difficulty of programming assignments, to give an introduction to system programming and file IO. \\
-Correctness testing: By running the ar command and checking for exact duplication of behavior. \\
-Learn: Start earlier! Also lots about IO and string manipulation.\\
+Point of assignment: To get us used to the level of difficulty of programming assignments, to give an introduction to system programming and file IO. To show how complicated
+even simple Linux tools are. \\
+Correctness testing: By running the ar command and checking for exact duplication of behavior. Also by checking to see if output from myar could be run with ar and 
+vice versa. I ran most of my tests with simple 3-character .txt files. The program works with more complicated text files; I can't imagine it working differently for larger
+files than 3-character ones. \\
+Learn: I learned next time I need to start earlier. Also I learned lots about IO and string manipulation.\\
 

### 2015-07-19 A2/myar.c
Fix arfile output permissions.

@@ -120,7 +120,7 @@ int main(int argc, char** argv) {
 			
 			if (stat(ar_file, &test) != 0) 
 			{
-				if ((ar_fd = open(ar_file, O_WRONLY | O_CREAT | O_EXCL, S_IRUSR | S_IWUSR | S_IRGRP | S_IROTH)) == -1)
+				if ((ar_fd = open(ar_file, O_WRONLY | O_CREAT | O_EXCL, S_IRUSR | S_IWUSR | S_IRGRP | S_IWGRP | S_IWOTH | S_IROTH)) == -1)
 				{
 					perror("exclusive open archive file: ");
 					exit(EXIT_FAILURE);
@@ -613,7 +613,7 @@ int main(int argc, char** argv) {
 			
 			if (stat(ar_file, &test) != 0) 
 			{
-				if ((ar_fd = open(ar_file, O_WRONLY | O_CREAT | O_EXCL, S_IRUSR | S_IWUSR | S_IRGRP | S_IROTH)) == -1)
+				if ((ar_fd = open(ar_file, O_WRONLY | O_CREAT | O_EXCL, S_IRUSR | S_IWUSR | S_IRGRP | S_IWGRP | S_IWOTH| S_IROTH)) == -1)
 				{
 					perror("exclusive open archive file: ");
 					exit(EXIT_FAILURE);

### 2015-07-20 A3/latex/writeup.tex
Add command command execution and beggining of shell string interpretation.

\documentclass[11pt]{article}

\begin{document}

\section{Design}

\section{Worklog}

\section{Challenges}

\section{Questions}
\textbf{Point of assignment}: To go over how the command shell works. To teach how

\textbf{Correctness testing}:

\textbf{Learn}:

\end{document}

### 2015-07-20 A3/mysh.c

@@ -1,0 +1,135 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <unistd.h>
+#include <sys/stat.h>
+#include <sys/types.h>
+
+int mysystem(char** command_string, int num_words);
+char* get_command_path(char* command);
+void run_command_str(char** arr_words, int num_words, FILE* out_file);
+
+int main(int argc, char** argv) {
+	// First, implement the system() library call
+	
+	mysystem(&argv[1], argc-1);
+	return 0;
+}
+
+int mysystem(char** command_string, int num_args) {
+	
+	run_command_str(command_string, num_args, stdout);
+	
+	return 0;
+}
+
+char* get_command_path(char* command) {
+	char* path_env_ptr = getenv("PATH");
+	char* end_path_ptr;
+	char* path;
+	
+	while ((end_path_ptr = strchr(path_env_ptr, ':')) != NULL)
+	{
+		char possible_path[end_path_ptr - path_env_ptr + 1 + strlen(command) + 1];
+		snprintf(possible_path, end_path_ptr - path_env_ptr + 1, "%s", path_env_ptr);
+		
+		strcat(possible_path, "/");
+		strcat(possible_path, command);
+		
+		if (access(possible_path, F_OK) == 0) {
+			path = malloc(strlen(possible_path));
+			strcpy(path, possible_path);
+			return path;
+		}
+		
+		path_env_ptr = end_path_ptr + 1;
+	}
+	
+	fprintf(stderr, "Error. Could not find command %s in path.\n", command);
+	exit(EXIT_FAILURE); 
+}
+
+
+void run_command_str(char** arr_words, int num_words, FILE* out_file) {
+	
+	char* command_file = get_command_path(arr_words[0]);
+	
+	char* params[num_words - 1 + 2];		// num_words minus command name plus command file and \0
+	params[0] = command_file;
+	params[num_words] = NULL;
+	
+	int c;
+	for(c = 1; c < num_words; c++) {
+		params[c] = arr_words[c];
+	}
+	
+	int f_val = fork();
+	switch(f_val) {
+		case 0:
+		{
+			
+			if (execvp(command_file, params) == -1) {
+				perror("execvp: ");
+				exit(EXIT_FAILURE);
+			}
+			exit(EXIT_SUCCESS);
+			break;
+		}
+		case -1:
+		{
+			perror("fork: ");
+			exit(EXIT_FAILURE);
[…]

### 2015-07-21 A3/mysh.c
Add command parser.

@@ -1,3 +1,4 @@
+#include <fcntl.h>
 #include <stdio.h>
 #include <stdlib.h>
 #include <string.h>
@@ -5,21 +6,78 @@
 #include <sys/stat.h>
 #include <sys/types.h>
 
-int mysystem(char** command_string, int num_words);
+enum MODES {
+	REDIR_OUT_APP,
+	REDIR_OUT_OVW,
+	REDIR_IN,
+	REDIR_IN_HERE,
+	PIPE
+};
+
+int mysystem(char* command_string);
 char* get_command_path(char* command);
-void run_command_str(char** arr_words, int num_words, FILE* out_file);
+void run_command_str(char** arr_words, int num_words, int out_fd);
+void command_string_slice(char* command_string, char*** argument_string_ptr, int* num_args);
 
 int main(int argc, char** argv) {
 	// First, implement the system() library call
 	
-	mysystem(&argv[1], argc-1);
+	mysystem(argv[1]);
 	return 0;
 }
 
-int mysystem(char** command_string, int num_args) {
+int mysystem(char* command_string) {
+	int cmd_start = 0;
+	int cmd_current = 0;
 	
-	run_command_str(command_string, num_args, stdout);
+	char** argument_string;
+	int num_args;
+	printf("command_string: |%s|\n", command_string);
+	command_string_slice(command_string, &argument_string, &num_args);
 	
+	int i;
+	for(i=0; i<num_args; i++)
+	{
+		printf("arg: |%s|\n", argument_string[i]);
+	}
+	
+	while (cmd_current < num_args)
+	{
+		printf("cmd_current: |%s|\n", argument_string[cmd_current]);
+		if (cmd_current == num_args-1)
+		{
+			
+			run_command_str(&argument_string[cmd_start], cmd_current - cmd_start + 1, STDOUT_FILENO);
+			cmd_current++;
+			
+		} 
+		
+		else if (argument_string[cmd_current][0] == '>') {
+			
+			// Lookahead, Open (TRUNC) new file, run command with file as stdout
+			char* outfile = argument_string[cmd_current + 1];
+			int out_fd;
+			if ((out_fd = open(outfile, O_WRONLY | O_CREAT | O_TRUNC, 0666)) == -1)
+			{
+				perror("open out file: ");
+				exit(EXIT_FAILURE);
+			}
+			
[…]

### 2015-07-21 A3/mysh.c
Add mostly bug-free piping.

@@ -16,7 +16,7 @@ enum MODES {
 
 int mysystem(char* command_string);
 char* get_command_path(char* command);
-void run_command_str(char** arr_words, int num_words, int out_fd);
+void run_command_str(char** arr_words, int num_words, int out_fd, int in_fd);
 void command_string_slice(char* command_string, char*** argument_string_ptr, int* num_args);
 
 int main(int argc, char** argv) {
@@ -30,29 +30,43 @@ int mysystem(char* command_string) {
 	int cmd_start = 0;
 	int cmd_current = 0;
 	
+	int piping = 0;
+	int pipe_fds[2];
+	
 	char** argument_string;
 	int num_args;
-	printf("command_string: |%s|\n", command_string);
+
 	command_string_slice(command_string, &argument_string, &num_args);
-	
-	int i;
-	for(i=0; i<num_args; i++)
-	{
-		printf("arg: |%s|\n", argument_string[i]);
-	}
 	
 	while (cmd_current < num_args)
 	{
-		printf("cmd_current: |%s|\n", argument_string[cmd_current]);
 		if (cmd_current == num_args-1)
 		{
+			if (piping)
+			{
+				run_command_str(&argument_string[cmd_start], cmd_current - cmd_start + 1, STDOUT_FILENO, pipe_fds[0]);
+				piping = 0;
+				if (close(pipe_fds[0]) == -1)
+				{
+					perror("close read pipe: ");
+					exit(EXIT_FAILURE);
+				}
+				if (close(pipe_fds[1]) == -1)
+				{
+					perror("close write pipe: ");
+					exit(EXIT_FAILURE);
+				}
+				cmd_current++;
+			}
 			
-			run_command_str(&argument_string[cmd_start], cmd_current - cmd_start + 1, STDOUT_FILENO);
-			cmd_current++;
-			
+			else 
+			{
+				run_command_str(&argument_string[cmd_start], cmd_current - cmd_start + 1, STDOUT_FILENO, -1);
+				cmd_current++;
+			}
 		} 
 		
-		else if (argument_string[cmd_current][0] == '>') {
+		else if (strcmp(argument_string[cmd_current], ">") == 0) {
 			
 			// Lookahead, Open (TRUNC) new file, run command with file as stdout
 			char* outfile = argument_string[cmd_current + 1];
[…]

### 2015-07-21 A3/mysh.c
Get all redirections to work.

@@ -79,7 +79,7 @@ int mysystem(char* command_string) {
 			
 			if (piping)
 			{
-				run_command_str(&argument_string[cmd_start], cmd_current - cmd_start + 1, out_fd, pipe_fds[0]);
+				run_command_str(&argument_string[cmd_start], cmd_current - cmd_start, out_fd, pipe_fds[0]);
 				piping = 0;
 				if (close(pipe_fds[0]) == -1)
 				{
@@ -120,7 +120,7 @@ int mysystem(char* command_string) {
 			
 			if (piping)
 			{
-				run_command_str(&argument_string[cmd_start], cmd_current - cmd_start + 1, out_fd, pipe_fds[0]);
+				run_command_str(&argument_string[cmd_start], cmd_current - cmd_start, out_fd, pipe_fds[0]);
 				piping = 0;
 				if (close(pipe_fds[0]) == -1)
 				{
@@ -151,40 +151,28 @@ int mysystem(char* command_string) {
 		else if (strcmp(argument_string[cmd_current], "<") == 0) {
 			
 			// Lookahead, Open (TRUNC) new file, run command with file as stdout
-			printf("this needs to be worked on. Also, get * to work.\n");
-			char* outfile = argument_string[cmd_current + 1];
-			int out_fd;
-			if ((out_fd = open(outfile, O_WRONLY | O_APPEND)) == -1)
+			char* infile = argument_string[cmd_current + 1];
+			int in_fd;
+			if ((in_fd = open(infile, O_RDONLY)) == -1)
 			{
-				perror("open out file: ");
+				perror("open in file: ");
 				exit(EXIT_FAILURE);
 			}
 			
 			if (piping)
 			{
-				run_command_str(&argument_string[cmd_start], cmd_current - cmd_start + 1, out_fd, pipe_fds[0]);
-				piping = 0;
-				if (close(pipe_fds[0]) == -1)
-				{
-					perror("close read pipe: ");
-					exit(EXIT_FAILURE);
-				}
-				if (close(pipe_fds[1]) == -1)
-				{
-					perror("close write pipe: ");
-					exit(EXIT_FAILURE);
-				}
-				cmd_current += 2;
+				printf("Error. Attempting to pipe and redirect. Exiting.\n");
+				exit(EXIT_FAILURE);
 			}
 			
 			else {
[…]

### 2015-07-21 A3/mysh.c
Made program shell-y! Now make respond to SIGINT/SIGQUIT...

@@ -21,8 +21,25 @@ void command_string_slice(char* command_string, char*** argument_string_ptr, int
 
 int main(int argc, char** argv) {
 	// First, implement the system() library call
+	int BUF_SIZE = 100000;
+	char buf[BUF_SIZE];
+	int read_bytes;
 	
-	mysystem(argv[1]);
+	while(1)
+	{
+		printf("prompt: ");
+		fflush(stdout);
+		if ((read_bytes = read(STDOUT_FILENO, buf, BUF_SIZE)) == -1)
+		{
+			perror("Couldn't read from stdout: ");
+			exit(EXIT_FAILURE);
+		}
+		buf[read_bytes-1] = '\0';
+		if (read_bytes > 1) {
+			mysystem(buf);
+		}
+	}
+	
 	return 0;
 }
 
@@ -49,12 +66,12 @@ int mysystem(char* command_string) {
 				if (close(pipe_fds[0]) == -1)
 				{
 					perror("close read pipe: ");
-					exit(EXIT_FAILURE);
+					//exit(EXIT_FAILURE);
 				}
 				if (close(pipe_fds[1]) == -1)
 				{
 					perror("close write pipe: ");
-					exit(EXIT_FAILURE);
+					//exit(EXIT_FAILURE);
 				}
 				cmd_current++;
 			}
@@ -74,7 +91,7 @@ int mysystem(char* command_string) {
 			if ((out_fd = open(outfile, O_WRONLY | O_CREAT | O_TRUNC, 0666)) == -1)
 			{
 				perror("open out file: ");
-				exit(EXIT_FAILURE);
+				//exit(EXIT_FAILURE);
 			}
 			
 			if (piping)
@@ -84,12 +101,12 @@ int mysystem(char* command_string) {
 				if (close(pipe_fds[0]) == -1)
 				{
 					perror("close read pipe: ");
-					exit(EXIT_FAILURE);
+					//exit(EXIT_FAILURE);
 				}
 				if (close(pipe_fds[1]) == -1)
 				{
 					perror("close write pipe: ");
-					exit(EXIT_FAILURE);
+					//exit(EXIT_FAILURE);
 				}
 				cmd_current += 2;
 			}
@@ -102,7 +119,7 @@ int mysystem(char* command_string) {
 			if (close(out_fd) == -1)
 			{
 				perror("close out file: ");
-				exit(EXIT_FAILURE);
+				//exit(EXIT_FAILURE);
 			}
 			
[…]

### 2015-07-21 A3/mysh.c
Add signal handling.

@@ -3,6 +3,7 @@
 #include <stdlib.h>
 #include <string.h>
 #include <unistd.h>
+#include <signal.h>
 #include <sys/stat.h>
 #include <sys/types.h>
 
@@ -19,11 +20,18 @@ char* get_command_path(char* command);
 void run_command_str(char** arr_words, int num_words, int out_fd, int in_fd);
 void command_string_slice(char* command_string, char*** argument_string_ptr, int* num_args);
 
+void signore(int sig) {
+	;
+}
+
 int main(int argc, char** argv) {
 	// First, implement the system() library call
 	int BUF_SIZE = 100000;
 	char buf[BUF_SIZE];
 	int read_bytes;
+	
+	signal(SIGINT, signore);
+	signal(SIGQUIT, signore);
 	
 	while(1)
 	{

### 2015-07-21 A3/mysh.c
Add final details. Revert here if you run into trouble, this is a submittable copy.

@@ -6,6 +6,7 @@
 #include <signal.h>
 #include <sys/stat.h>
 #include <sys/types.h>
+#include <sys/wait.h>
 
 enum MODES {
 	REDIR_OUT_APP,
@@ -33,9 +34,12 @@ int main(int argc, char** argv) {
 	signal(SIGINT, signore);
 	signal(SIGQUIT, signore);
 	
+	char pwdbuf[2048];
+	
 	while(1)
 	{
-		printf("prompt: ");
+		getcwd(pwdbuf, 2048);
+		printf("%s: ", pwdbuf);
 		fflush(stdout);
 		if ((read_bytes = read(STDOUT_FILENO, buf, BUF_SIZE)) == -1)
 		{
@@ -43,7 +47,10 @@ int main(int argc, char** argv) {
 			exit(EXIT_FAILURE);
 		}
 		buf[read_bytes-1] = '\0';
-		if (read_bytes > 1) {
+		if (strcmp(buf, "exit") == 0){
+			printf("Exiting shell.\n");
+			exit(EXIT_SUCCESS);
+		} else if (read_bytes > 1) {
 			mysystem(buf);
 		}
 	}
@@ -54,6 +61,8 @@ int main(int argc, char** argv) {
 int mysystem(char* command_string) {
 	int cmd_start = 0;
 	int cmd_current = 0;
+	
+	int saved_stderr = dup(STDERR_FILENO);
 	
 	int piping = 0;
 	int pipe_fds[2];
@@ -81,6 +90,11 @@ int mysystem(char* command_string) {
 					perror("close write pipe: ");
 					//exit(EXIT_FAILURE);
 				}
+				if (dup2(saved_stderr, STDERR_FILENO) == -1) 
+				{
+					perror("restoring file pointer of stderr: ");
+				}
+				saved_stderr = dup(STDERR_FILENO);
 				cmd_current++;
 			}
 			
@@ -158,6 +172,11 @@ int mysystem(char* command_string) {
 					//exit(EXIT_FAILURE);
 				}
 				cmd_current += 2;
+				if (dup2(saved_stderr, STDERR_FILENO) == -1) 
+				{
+					perror("restoring file pointer of stderr: ");
+				}
+				saved_stderr = dup(STDERR_FILENO);
 			}
 			
 			else {
@@ -304,7 +323,7 @@ void run_command_str(char** arr_words, int num_words, int out_fd, int in_fd) {
 		case 0:
 		{
 			
[…]

### 2015-07-21 A3/latex/writeup.tex
Add writeup.tex mark 1.

@@
 \section{Design}
+My shell implementation is named mysh. It is implemented as a shell loop which listens for lines delimited by the enter key. It parses lines from the user, 
+separates them into space-delimited symbols, groups the symbol list into commands separated by redirection operators (|, >, >>, <), and uses the fork and execve
+system calls to perform each command. The program sets STDIN and STDOUT of each child process to accomplish Linux redirection and piping.\\
+The shell portion of the program sets signal handlers to ignore SIGINT and SIGQUIT system calls. It also listens for the \"exit\" keyword, which prompts the 
+program to exit. Every step of the main loop (iterating by while(1)) the program reads the current working directory, prints it, flushes STDOUT, then listens for
+user input. When the enter key is pressed, it reads everything from STDIN, lops off the final newline and sends the input to a symbol parser which separates the
+command into an array of individual symbols. \\
+These symbols are then separated into subcommands. The first symbol in every subcommand is the name of the Linux program to run. The \${PATH} variable is searched
+through and the first path which contains a viable program file is passed to execve(), along with a formatted list of the rest of the options in the subcommand.
+If this subcommand is piping to another subcommand, being piped to by another subcommand, outputting to a file or getting input from a file, mysh sets its 
+STDOUT and STDIN values temporarily to the necessary file to accomplish the pipe or redirection. It then returns the streams to their original file descriptors.
+Both of these operations are done via dup2(). \\
+Once execve has run all the subcommands in the command, control returns to the shell loop, which restarts. \\
+This program does NOT implement the stream redirection operator x>&y.
 
 \section{Worklog}
+Sunday July 19th 12:00pm - 4:00pm: Explore exec wrapper functions, build temporary program which executes argv input via execv and fork. \\
+Monday July 20th 7:00am - 8:10am: Implement beginnings of subcommand parsing, supporting only redirection operators. \\
+Monday July 20th 9:30am - 12:00pm: Iron out bugs in redirection operator code, begin on piping. \\
+Monday July 20th 1:00pm - 6:30pm: Finish piping stuff. \\
+Monday July 20th 6:30pm - 9:00pm: Get piping and all other redirection to work fairly seamlessly. \\
+Monday July 20th 9:00pm - 10:00pm: Get shell portion of code to work. Correctly listen for user input and send lines to parser. \\
+Monday July 20th 10:00pm - 11:15pm: Finishing touches. pwd on each new line. Swap out execv for execve very easily. \\
+Monday July 20th 11:15pm: Work on worklog. \\
 
 \section{Challenges}
+This assignment was more open than the archive assignment, so it was challenging thinking of all the things a shell needed to do. It was challenging and fun
+thinking of the program's architecture and which functions would do which jobs. PATH searching and char** argument array helper functions were easy and very helpful. 
+Overall, the most challenging part was planning and ironing out bugs in the piping and redirection code. \\
 
 \section{Questions}
-\textbf{Point of assignment}: To go over how the command shell works. To teach how
+\textbf{Point of assignment}: To go over how shells work and how little they actually do on the system. To go over how to use execve() and its wrapper functions. To 
+teach about how processes work on the system, and how forking can be used to make new processes. \\
 
-\textbf{Correctness testing}:
+\textbf{Correctness testing}: To test correctness I ran as many linux commands as I could think of. I ran them on both my machine and the os-class server. I tested
+different piping and redirection combinations. \\
 
-\textbf{Learn}:
+\textbf{Learn}: I learned a lot about how shells work and what they do for the user. I also learned how processes are made. I learned how to manage my time and the
+importance of knowing how long different parts of a project will take. \\
 

### 2015-07-21 A3/latex/writeup.tex
Add writeup.tex mark 2.

@@
 separates them into space-delimited symbols, groups the symbol list into commands separated by redirection operators (|, >, >>, <), and uses the fork and execve
-system calls to perform each command. The program sets STDIN and STDOUT of each child process to accomplish Linux redirection and piping.\\
+system calls to perform each command. The program sets STDIN and STDOUT of each child process to accomplish Linux redirection and piping.\\ \\
 The shell portion of the program sets signal handlers to ignore SIGINT and SIGQUIT system calls. It also listens for the \"exit\" keyword, which prompts the 
@@
 user input. When the enter key is pressed, it reads everything from STDIN, lops off the final newline and sends the input to a symbol parser which separates the
-command into an array of individual symbols. \\
+command into an array of individual symbols. \\ \\
 These symbols are then separated into subcommands. The first symbol in every subcommand is the name of the Linux program to run. The \${PATH} variable is searched
@@
 STDOUT and STDIN values temporarily to the necessary file to accomplish the pipe or redirection. It then returns the streams to their original file descriptors.
-Both of these operations are done via dup2(). \\
+Both of these operations are done via dup2(). \\ \\
 Once execve has run all the subcommands in the command, control returns to the shell loop, which restarts. \\
-This program does NOT implement the stream redirection operator x>&y.
+This program does NOT implement the stream redirection operator x>\&y.
 
 \section{Worklog}
-Sunday July 19th 12:00pm - 4:00pm: Explore exec wrapper functions, build temporary program which executes argv input via execv and fork. \\
-Monday July 20th 7:00am - 8:10am: Implement beginnings of subcommand parsing, supporting only redirection operators. \\
-Monday July 20th 9:30am - 12:00pm: Iron out bugs in redirection operator code, begin on piping. \\
-Monday July 20th 1:00pm - 6:30pm: Finish piping stuff. \\
-Monday July 20th 6:30pm - 9:00pm: Get piping and all other redirection to work fairly seamlessly. \\
-Monday July 20th 9:00pm - 10:00pm: Get shell portion of code to work. Correctly listen for user input and send lines to parser. \\
-Monday July 20th 10:00pm - 11:15pm: Finishing touches. pwd on each new line. Swap out execv for execve very easily. \\
-Monday July 20th 11:15pm: Work on worklog. \\
+\textbf{Sunday July 19th 12:00pm - 4:00pm}: Explore exec wrapper functions, build temporary program which executes argv input via execv and fork. \\
+\textbf{Monday July 20th 7:00am - 8:10am}: Implement beginnings of subcommand parsing, supporting only redirection operators. \\
+\textbf{Monday July 20th 9:30am - 12:00pm}: Iron out bugs in redirection operator code, begin on piping. \\
+\textbf{Monday July 20th 1:00pm - 6:30pm}: Finish piping stuff. \\
+\textbf{Monday July 20th 6:30pm - 9:00pm}: Get piping and all other redirection to work fairly seamlessly. \\
+\textbf{Monday July 20th 9:00pm - 10:00pm}: Get shell portion of code to work. Correctly listen for user input and send lines to parser. \\
+\textbf{Monday July 20th 10:00pm - 11:15pm}: Finishing touches. pwd on each new line. Swap out execv for execve very easily. \\
+\textbf{Monday July 20th 11:15pm}: Work on worklog. \\
 

### 2015-07-21 A3/latex/writeup.tex
mk3

@@
 \begin{document}
+Elliott Capek \\
 
@@
 My shell implementation is named mysh. It is implemented as a shell loop which listens for lines delimited by the enter key. It parses lines from the user, 
-separates them into space-delimited symbols, groups the symbol list into commands separated by redirection operators (|, >, >>, <), and uses the fork and execve
+separates them into space-delimited symbols, groups the symbol list into commands separated by redirection operators (\|, \>, \>\>, \<), and uses the fork and execve
 system calls to perform each command. The program sets STDIN and STDOUT of each child process to accomplish Linux redirection and piping.\\ \\
-The shell portion of the program sets signal handlers to ignore SIGINT and SIGQUIT system calls. It also listens for the \"exit\" keyword, which prompts the 
+The shell portion of the program sets signal handlers to ignore SIGINT and SIGQUIT system calls. It also listens for the exit keyword, which prompts the 
 program to exit. Every step of the main loop (iterating by while(1)) the program reads the current working directory, prints it, flushes STDOUT, then listens for
@@
 Both of these operations are done via dup2(). \\ \\
-Once execve has run all the subcommands in the command, control returns to the shell loop, which restarts. \\
-This program does NOT implement the stream redirection operator x>\&y.
+Once execve has run all the subcommands in the command, control returns to the shell loop, which restarts. \\ \\
+This program does NOT implement the stream redirection operator x\>\&y.
 

### 2015-07-21 A3/latex/writeup.tex
mk3

@@
 My shell implementation is named mysh. It is implemented as a shell loop which listens for lines delimited by the enter key. It parses lines from the user, 
-separates them into space-delimited symbols, groups the symbol list into commands separated by redirection operators (\|, \>, \>\>, \<), and uses the fork and execve
+separates them into space-delimited symbols, groups the symbol list into commands separated by redirection operators and uses the fork and execve
 system calls to perform each command. The program sets STDIN and STDOUT of each child process to accomplish Linux redirection and piping.\\ \\
@@
 Once execve has run all the subcommands in the command, control returns to the shell loop, which restarts. \\ \\
-This program does NOT implement the stream redirection operator x\>\&y.
+This program does NOT implement the stream redirection operator $2>\&1$.
 

### 2015-08-03 A4/prime_procs.c
Adding working prime_procs.c

@@ -1,3 +1,4 @@
+#include <limits.h>
 #include <fcntl.h>
 #include <math.h>
 #include <stdio.h>
@@ -9,7 +10,7 @@
 
 #include "dynarr.h"
 
-#define MAX_PRIME 10000
+#define MAX_PRIME UINT_MAX
 #define PROCESSES 16
 
 int* primes;
@@ -56,9 +57,9 @@ int get_next_sieve_prime(int start) {
 }
 
 static void* find_happiness(int section) {
-	int MSIZE = MAX_PRIME * sizeof(int);
+	long MSIZE = MAX_PRIME * sizeof(int);
 	
-	int i;
+	long i;
 	for (i = section*MAX_PRIME/PROCESSES + 1; i<=(section*MAX_PRIME/PROCESSES + MAX_PRIME/PROCESSES); i++) {
 		if (primes[i] && is_happy(i+1)) { happiness[i] = 1; }
 	}
@@ -73,17 +74,8 @@ static void* find_happiness(int section) {
 }
 
 void mark_multiples(int m) {
-	int MSIZE = MAX_PRIME * sizeof(int);
-	int j = 2;
-	
-	/*sem_t* num_term_children;
-	
-	num_term_children = sem_open("num_term_children", 0);
-	if (num_term_children == SEM_FAILED)
-	{
-		perror("sem_open: ");
-		exit(EXIT_FAILURE);
-	}*/
+	long MSIZE = MAX_PRIME * sizeof(int);
+	long j = 2;
 	
 	while (m*j <= MAX_PRIME)
 	{
@@ -102,18 +94,9 @@ void mark_multiples(int m) {
 }
 
 int main() {
-	int m = 1;									// Current sieve prime, initialized to 1
-	int MSIZE = MAX_PRIME * sizeof(int);
+	long m = 1;									// Current sieve prime, initialized to 1
+	long MSIZE = MAX_PRIME * sizeof(int);
 	int child_pids[PROCESSES];
-	
-	sem_t* num_term_children;
-	
-	num_term_children = sem_open("num_term_children", O_CREAT, S_IRUSR | S_IWUSR, 0);
-	if (num_term_children == SEM_FAILED)
-	{
-		perror("sem_open: ");
-		exit(EXIT_FAILURE);
-	}
 	
[…]

### 2015-08-03 A4/prime_procs.c
Add working copies of both programs.

@@ -10,7 +10,7 @@
 
 #include "dynarr.h"
 
-#define MAX_PRIME UINT_MAX
+#define MAX_PRIME 1000000
 #define PROCESSES 16
 
 int* primes;
@@ -289,7 +289,7 @@ int main() {
 	{
 		if (happiness[e])
 		{
-			printf("p: %ld\t%d\n", p, (unsigned int) e+1);
+			//printf("p: %ld\t%d\n", p, (unsigned int) e+1);
 			out_arr[p] = (unsigned int) e+1;
 			p++;
 		}

### 2015-08-03 A4/prime_threads.c

@@ -195,35 +195,60 @@ int main() {
 		}
 	}
 	
-	int fd;
-	if ((fd = open("happy_primes", O_WRONLY | O_CREAT | O_TRUNC, 0755)) == -1)
+	int num_happy = 0;
+	int y;
+	for (y=0; y<MAX_PRIME; y++)
+	{
+		if (happiness[y])
+		{
+			num_happy++;
+		}
+	}
+	
+	long HSIZE = num_happy*sizeof(unsigned int);
+	unsigned int* out_arr = malloc(HSIZE);
+	
+	int out_fd;
+	if ((out_fd = open("happy_primes_threads", O_RDWR | O_CREAT | O_TRUNC, 0755)) == -1)
 	{
 		perror("error open: ");
 		exit(EXIT_FAILURE);
 	}
 	
+	if (ftruncate(out_fd, HSIZE) == -1)
+	{
+		perror("ftruncate: ");
+		exit(EXIT_FAILURE);
+	}
+	
+	if (lseek(out_fd, 0, SEEK_SET) == -1)
+	{
+		perror("lseek");
+		exit(EXIT_FAILURE);
+	}
+	
+	long p = 0;
 	int e;
-	int p = 1;
 	for (e=0; e<MAX_PRIME; e++)
 	{
 		if (happiness[e])
 		{
-			char buf[100];
-			sprintf(buf, "%d", (unsigned int) e+1);			// Make this print what it's actually supposed to print
-			write(fd, buf, strlen(buf));
+			//printf("p: %ld\t%d\n", p, (unsigned int) e+1);
+			out_arr[p] = (unsigned int) e+1;
 			p++;
 		}
 	}
 	
-	if (close(fd) == -1)
+	if (write(out_fd, (void*) out_arr, HSIZE) == -1)
+	{
+		perror("write");
+		exit(EXIT_FAILURE);
+	}
+	
+	if (close(out_fd) == -1)
 	{
 		perror("close: ");
 		exit(EXIT_FAILURE);
 	}
 	
-	pthread_mutex_unlock(&mutex);
-	pthread_cond_destroy(&thread_termination);
 }
-
-//SORT FROM LEAST TO GREATEST!!
-// Does ALL the program need ot be threaded? or just most? Does my writing need to happen in threads?

### 2015-08-03 A4/prime_procs.c
Going to go test on os-class.

@@ -10,7 +10,7 @@
 
 #include "dynarr.h"
 
-#define MAX_PRIME 1000000
+#define MAX_PRIME 881269
 #define PROCESSES 16
 
 int* primes;
@@ -289,7 +289,7 @@ int main() {
 	{
 		if (happiness[e])
 		{
-			//printf("p: %ld\t%d\n", p, (unsigned int) e+1);
+			printf("%ld %d\n", p+1, (unsigned int) e+1);
 			out_arr[p] = (unsigned int) e+1;
 			p++;
 		}

### 2015-08-03 A4/prime_threads.c

@@ -8,7 +8,7 @@
 
 #include "dynarr.h"
 
-#define MAX_PRIME 1000000
+#define MAX_PRIME 881269
 #define THREADS 16
 
 int primes[MAX_PRIME];
@@ -233,7 +233,7 @@ int main() {
 	{
 		if (happiness[e])
 		{
-			//printf("p: %ld\t%d\n", p, (unsigned int) e+1);
+			printf("%ld %d\n", p+1, (unsigned int) e+1);
 			out_arr[p] = (unsigned int) e+1;
 			p++;
 		}

### 2015-08-03 A4/primes.c

@@ -1,28 +1,0 @@
-#include <math.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-
-#include "utilities.h"
-
-
-
-
-int main() {
-	is_happy(998);
-	return 0;
-}
-
-
-
-
-
-
-
-
-
-
-
-
-
-

### 2015-08-04 A4/prime_procs.c
Made a few changes. Now need to figure out why procs isn't working for big files.

@@ -10,18 +10,18 @@
 
 #include "dynarr.h"
 
-#define MAX_PRIME 881269
+#define MAX_PRIME UINT_MAX
 #define PROCESSES 16
 
 int* primes;
 int* happiness;
 
-int is_happy(int num) {
+int is_happy(long num) {
 	DynIntArr* arr = malloc(sizeof(DynIntArr));
 	arr_init(arr, 10);
 	
 	while(num != 1) {						// Loop until found happy or sad
-		int new_num = 0;
+		long new_num = 0;
 		while (num / 10 != 0) {				// While >1 digit
 			new_num += pow(num % 10, 2);	// num % 10 = least sig. digit
 			num = num / 10;					// Chop off last digit
@@ -29,7 +29,7 @@ int is_happy(int num) {
 		new_num += pow(num, 2);
 		num = new_num;
 		
-		int i;								// Check if already hit this #
+		long i;								// Check if already hit this #
 		for (i=0; i<arr->size; i++)
 		{
 			if (num == arr->data[i])
@@ -47,7 +47,7 @@ int is_happy(int num) {
 }
 
 int get_next_sieve_prime(int start) {
-	int m = start + 1;
+	long m = start + 1;
 	while (m <= floor(sqrt(MAX_PRIME)))
 	{
 	  	if (primes[m-1] == 1) { return m; }		// m is prime, return as next sieve prime
@@ -252,8 +252,8 @@ int main() {
 		waitpid(child_pids[e], NULL, 0);
 	}
 	
-	int num_happy = 0;
-	int y;
+	long num_happy = 0;
+	long y;
 	for (y=0; y<MAX_PRIME; y++)
 	{
 		if (happiness[y])
@@ -289,7 +289,7 @@ int main() {
 	{
 		if (happiness[e])
 		{
-			printf("%ld %d\n", p+1, (unsigned int) e+1);
+			//printf("%ld %d\n", p+1, (unsigned int) e+1);
 			out_arr[p] = (unsigned int) e+1;
 			p++;
 		}

### 2015-08-04 A4/prime_threads.c

@@ -1,4 +1,5 @@
 #include <fcntl.h>
+#include <limits.h>
 #include <math.h>
 #include <pthread.h>
 #include <stdio.h>
@@ -8,7 +9,7 @@
 
 #include "dynarr.h"
 
-#define MAX_PRIME 881269
+#define MAX_PRIME UINT_MAX
 #define THREADS 16
 
 int primes[MAX_PRIME];
@@ -25,17 +26,17 @@ enum t_state {
 
 typedef struct {
 	pthread_t tid;
-	int num;
-	int sieve_prime;
+	long num;
+	long sieve_prime;
 	enum t_state state;
 } thread_struct;
 
-int is_happy(int num) {
+int is_happy(long num) {
 	DynIntArr* arr = malloc(sizeof(DynIntArr));
 	arr_init(arr, 10);
 	
 	while(num != 1) {						// Loop until found happy or sad
-		int new_num = 0;
+		long new_num = 0;
 		while (num / 10 != 0) {				// While >1 digit
 			new_num += pow(num % 10, 2);	// num % 10 = least sig. digit
 			num = num / 10;					// Chop off last digit
@@ -43,7 +44,7 @@ int is_happy(int num) {
 		new_num += pow(num, 2);
 		num = new_num;
 		
-		int i;								// Check if already hit this #
+		long i;								// Check if already hit this #
 		for (i=0; i<arr->size; i++)
 		{
 			if (num == arr->data[i])
@@ -60,8 +61,8 @@ int is_happy(int num) {
 	return 1;
 }
 
-int get_next_sieve_prime(int start) {
-	int m = start + 1;
+int get_next_sieve_prime(long start) {
+	long m = start + 1;
 	while (m <= floor(sqrt(MAX_PRIME)))
 	{
 	  	if (primes[m-1] == 1) { return m; }		// m is prime, return as next sieve prime
@@ -71,9 +72,9 @@ int get_next_sieve_prime(int start) {
 }
 
 static void* mark_multiples(void* arg) {
-	int j = 2;
+	long j = 2;
 	thread_struct* thread_info = (thread_struct*) arg;
-	int m = thread_info->sieve_prime;
+	long m = thread_info->sieve_prime;
 	
 	while (m*j <= MAX_PRIME)
 	{
@@ -92,7 +93,7 @@ static void* mark_multiples(void* arg) {
 static void* find_happiness(void* arg) {
[…]

### 2015-08-04 A4/prime_procs.c
Adding new stuff to run on scratch, not other disc

@@ -89,23 +89,22 @@ void mark_multiples(int m) {
 		exit(EXIT_FAILURE);
 	}
 	
-	//sem_post(num_term_children);
 	exit(EXIT_SUCCESS);
 }
 
 int main() {
-	long m = 1;									// Current sieve prime, initialized to 1
+	long m = 1;											// Current sieve prime, initialized to 1
 	long MSIZE = MAX_PRIME * sizeof(int);
 	int child_pids[PROCESSES];
 	
-	int fd = open("prime_procs_temp_shm", O_RDWR | O_CREAT | O_TRUNC, 0755);
+	int fd = open("/scratch/cs344-su15/capeke/prime_procs_temp_shm", O_RDWR | O_CREAT | O_TRUNC, 0755);
 	if (fd == -1)
 	{
 		perror("open: ");
 		exit(EXIT_FAILURE);
 	}
 	
-	int fd_h = open("prime_procs_temp_shm_happiness", O_RDWR | O_CREAT | O_TRUNC, 0755);
+	int fd_h = open("/scratch/cs344-su15/capeke/prime_procs_temp_shm_happiness", O_RDWR | O_CREAT | O_TRUNC, 0755);
 	if (fd == -1)
 	{
 		perror("open: ");
@@ -266,7 +265,7 @@ int main() {
 	unsigned int* out_arr = malloc(HSIZE);
 	
 	int out_fd;
-	if ((out_fd = open("happy_primes_procs", O_RDWR | O_CREAT | O_TRUNC, 0755)) == -1)
+	if ((out_fd = open("/scratch/cs344-su15/capeke/happy_primes_procs", O_RDWR | O_CREAT | O_TRUNC, 0755)) == -1)
 	{
 		perror("error open: ");
 		exit(EXIT_FAILURE);

### 2015-08-04 A4/prime_threads.c

@@ -210,7 +210,7 @@ int main() {
 	unsigned int* out_arr = malloc(HSIZE);
 	
 	int out_fd;
-	if ((out_fd = open("happy_primes_threads", O_RDWR | O_CREAT | O_TRUNC, 0755)) == -1)
+	if ((out_fd = open("/scratch/cs344-su15/capeke/happy_primes_threads", O_RDWR | O_CREAT | O_TRUNC, 0755)) == -1)
 	{
 		perror("error open: ");
 		exit(EXIT_FAILURE);

### 2015-08-04 A4/prime_procs.c
resolving merge rr

@@ -10,8 +10,8 @@
 
 #include "dynarr.h"
 
-#define MAX_PRIME UINT_MAX
-#define PROCESSES 16
+#define MAX_PRIME 1000000
+#define PROCESSES 1
 
 int* primes;
 int* happiness;
@@ -97,14 +97,14 @@ int main() {
 	long MSIZE = MAX_PRIME * sizeof(int);
 	int child_pids[PROCESSES];
 	
-	int fd = open("/scratch/cs344-su15/capeke/prime_procs_temp_shm", O_RDWR | O_CREAT | O_TRUNC, 0755);
+	int fd = open("prime_procs_temp_shm", O_RDWR | O_CREAT | O_TRUNC, 0755);
 	if (fd == -1)
 	{
 		perror("open: ");
 		exit(EXIT_FAILURE);
 	}
 	
-	int fd_h = open("/scratch/cs344-su15/capeke/prime_procs_temp_shm_happiness", O_RDWR | O_CREAT | O_TRUNC, 0755);
+	int fd_h = open("prime_procs_temp_shm_happiness", O_RDWR | O_CREAT | O_TRUNC, 0755);
 	if (fd == -1)
 	{
 		perror("open: ");
@@ -265,7 +265,7 @@ int main() {
 	unsigned int* out_arr = malloc(HSIZE);
 	
 	int out_fd;
-	if ((out_fd = open("/scratch/cs344-su15/capeke/happy_primes_procs", O_RDWR | O_CREAT | O_TRUNC, 0755)) == -1)
+	if ((out_fd = open("happy_primes_procs", O_RDWR | O_CREAT | O_TRUNC, 0755)) == -1)
 	{
 		perror("error open: ");
 		exit(EXIT_FAILURE);

### 2015-08-04 A4/prime_threads.c

@@ -9,8 +9,8 @@
 
 #include "dynarr.h"
 
-#define MAX_PRIME UINT_MAX
-#define THREADS 16
+#define MAX_PRIME 100000000
+#define THREADS 1
 
 int primes[MAX_PRIME];
 int happiness[MAX_PRIME];
@@ -210,7 +210,7 @@ int main() {
 	unsigned int* out_arr = malloc(HSIZE);
 	
 	int out_fd;
-	if ((out_fd = open("/scratch/cs344-su15/capeke/happy_primes_threads", O_RDWR | O_CREAT | O_TRUNC, 0755)) == -1)
+	if ((out_fd = open("happy_primes_threads", O_RDWR | O_CREAT | O_TRUNC, 0755)) == -1)
 	{
 		perror("error open: ");
 		exit(EXIT_FAILURE);

### 2015-08-04 A4/writeup.tex
Adding changes to writeup.tex.

\documentclass[11pt]{article}

\begin{document}
Elliott Capek \\

\section{Design}

\section{Worklog}
\textbf{Friday July 31st 12:00pm-3:00pm} - Review TLPI book on topics: threads, mutexes, conditional variables, forking, shared memory, semaphores. Write design.\\
\textbf{Friday July 31st 3:30pm-5:00pm} - Wrote is happy function. Wrote dynamic array implementation for use in ishappy. Wrote single-thread Sieve of Eratosthenes implementation. \\
\textbf{Saturday August 1st 11:30pm-1:00pm} - Start on threaded implementation. Get application to successfully create threads and use mutexes on thread counter. Encountered weird bug where multiple threads would die when I expected only one to. Figured out it's because signaling a conditional variable doesn't immediately run the process waiting on it. Multiple threads would die and signal, then the waiting process would run. Fixed this bug by allowing for multiple thread deaths per wait.\\
\textbf{Saturday August 1st 1:30pm-5:00pm} - Debugged another weird bug which had a really easy solution. I found that threads were behaving weirdly and somehow having their states set from alive to terminated even though they hadn't even started running yet. After much searching, found it was because I didn't make sure all threads were dead before going on to the next operation. One thread would sometimes stay alive long enough to make it into the next set of threads and then terminate, causing weird behavior. \\
\textbf{Sunday August 2nd 11:00am-12:00pm} - Put finishing touches on threaded program. \\
\textbf{Sunday August 2nd 12:30pm-3:00pm} - Began implementing multi-process program. Did more research on how I wanted to have the processes interact with each other. Successfully got processes to sieve through a range by modifying threaded program.\\
\textbf{Sunday August 2nd 3:00pm-7:00pm} - Debugged minor issues, like shared memory not being truncated before use. Began implementing the ishappy functionality. \\
\textbf{Sunday August 2nd 7:00pm-9:00pm} - Finished the ishappy functionality. Began researching how to write output to a file. For some reason spent lots of time thinking about how to take an integer and convert it into a binary-representation string, then sprintf to a file. \\
\textbf{Monday August 3rd 11:00am-2:00pm} - Realized that write() was a much better solution than printing to a file. Implemented functionality to open happy primes file, truncate it, and write an unsigned int array to it. Spent a lot of time trying to figure out why my read() couldn't understand what I had just written to the file. Turned out I hadn't lseek()'d to the beginning of the file after I wrote, so of course I couldn't read what I had written. \\
\textbf{Sunday August 2nd 2:30pm-5:00pm} - Put finishing touches on programs, removed unnecessary pieces. \\
\textbf{Sunday August 2nd 5:00pm-7:00pm} - Moved code to server. Hit several problems. First, my programs couldn't be compiled initially because the heap arrays I was making were way too large (range from 0 to UINT MAX). Researched the icc flags to allow this behavior. Second, my programs kept getting killed the second they were executed. Internet says this is because the kernel thought they were too big at the time. Never fixed this bug. Found the key is to wait until not so many other procrastinating students are running their code. \\
\textbf{Sunday August 2nd 7:00pm-10:00pm} - Begin timing code and composing writeup. \\

\section{Challenges}
This was a challenging but very good assignment. I learned a lot about how multithreaded and multiprocess programs work. This was an assignment where I had to spend lots of time reading TLPI to figure out how I wanted to make my program. I was challenged with different implementations - to use semaphores or block on waitpid()s, or to use signals. Etc. \\

The hardest challenge was debugging a stochastic problem in my threaded program. It would show up maybe 1/3 of the time, but sometimes almost never, such that I thought it was gone. The problem was that I wasn't killing all my threads from a previous function (sieving through the primes), so when I tried to coordinate threads to figure out the happy primes of my sieve, some of the threads would die before I could use them. This screwed up the whole execution of the program, since I needed my thread-count variable to accurately reflect the number of threads alive and dead. It took a long time to figure out the source of the error, but I learned a lot. My takeaway from this is that programming slowly and diligently and making sure I think of all edge cases as I write a section of code is much better than quickly writing up a function and debugging it later. \\

\section{Questions}
\textbf{Point of assignment}: The point of this assignment is to have students become familiar with both threads and processes. This assignment is an introduction into parallel programming and sheds light on which approach (threads or processes) is appropriate to what situation. It also teaches how to have different processes communicate via signals, conditional variables, semaphores and mutexes. It also has students learn about how shared memory can be used to easily and quickly link program memory to disk memory.\\

\textbf{Correctness testing}: To test for correctness, I used the OEIS file for the first 10,000 happy primes. I had both my programs output data using the same formatting as the web database. I then did a diff between my outputted files and a downloaded copy of the database. The diff had no output, showing there was no difference between my file and the database. To test my file was being properly binary-dumped into a file, I read() that file and did a line-by-line comparison of the written and read arrays. \\

\textbf{What was learned}: My takeaway from this assignment is that threads are much easier to use than processes for parallel computing. When I was writing the programs I felt that threads were a much more elegant solution, since they already have shared memory implicitly (although I was very surprised by how easy to use shared memory was). Processes were slightly easier to use in that they just signaled when they were about to terminate and didn't need to use a mutex, whereas threads needed to use a mutex to signal with a conditional variable. This is a very minor difference, though. I also learned that shared memory is incredibly powerful, since it allows processes to essentially share data without having to coordinate at all. \\

\end{document}

### 2015-08-04 A4/prime_procs.c
Add signal handling.

@@ -88,7 +88,6 @@ void mark_multiples(int m) {
 		perror("mysnc: ");
 		exit(EXIT_FAILURE);
 	}
-	
 	exit(EXIT_SUCCESS);
 }
 
@@ -169,6 +168,9 @@ int main() {
 		exit(EXIT_FAILURE);
 	}
 		
+	signal(SIGINT, SIG_IGN);
+	signal(SIGQUIT, SIG_IGN);	
+
 	int c;
 	for (c=0; c<PROCESSES; c++)
 	{
@@ -250,6 +252,9 @@ int main() {
 	{
 		waitpid(child_pids[e], NULL, 0);
 	}
+	
+	signal(SIGINT, SIG_DFL);
+	signal(SIGQUIT, SIG_DFL);
 	
 	long num_happy = 0;
 	long y;

### 2015-08-04 A4/writeup.tex
final commit

@@
 \section{Design}
+This project is broken up into two separate pieces: a threaded portion and a multi-process portion.
+Both processes share certain functions, which are implemented almost identically: a sieve function which takes an integer input and marks as composite all multiples of that integer and a function which calculates whether an input integer is happy.\\
+
+The sieve function is simple.
+It just modifies an integer array (memory mapped or normal).
+The happy function is also simple.
+It applies the happy number algorithm repeatedly and after each iteration adds the next number to a dynamic array.
+Once the next number is 1, it returns a 1.
+If the next number is ever found in the dynamic array, that means the cycle is looping and the number will never be happy, so it returns a 0.\\
+
+The threaded program will take as input the number of threads to use and the maximum number for its range.
+This will allow for a script to dynamically time its behavior.
+The program will use mutexes and conditional variables.
+TLPI has a very good example in its second Threads chapter on how to use conditional variables to launch several threads, then stop them all.
+The example uses a conditional variable sort of like a semaphore.
+Every time a thread is about to die, it locks the mutex, decrements a thread counter, changes its state to terminated, unlocks the mutex and dies.
+Meanwhile the main thread wait()s on the worker threads.
+This wait functions works so that the main thread has the death mutex at all times except when it is waiting.
+Thusly threads can only die when the main thread is waiting for them to die.
+After the wait stops blocking, the main thread checks its children for termination.
+I will need to keep these threads working after they terminate, so I will have the program at this point refresh each thread and call it anew.
+I will repeat this loop until the Sieve of Eratosthenes has gone through all the numbers, up to the square root of the max prime.
+After this is done I will use the exact same method to figure out which primes are happy, except I will do no thread refreshing and instead just evenly distribute the work among all threads.
+After this I will somehow write all the happy primes to a file.\\
+
+The multiprocess program will basically work the same way.
+It will do all the same behavior and use the same functions, slightly modified.
+Instead of making a bunch of new threads, it will fork().
+The child process will do the child function (sieving or finding happiness).
+These functions will end with an exit call.
+To sieve, the program will start with two, have a child sieve that number, find the next prime, sieve on that number, etc etc until the square root of the max prime has been sieved.
+The sieve will then be complete.
+To make sure no more than PROCESSES processes are running, I will first make PROCESSES processes.
+I will have a data structure to count the aliveness / deadness of each child.
+I will then use the waitpid() function almost exactly like the conditional wait function of the threaded application to wait for processes to die, then renew them.
+In this way I will keep sieving until the primes are all found.
+I will then evenly distribute the work of calculating happiness among the child processes, write to file and be done.
+\\
 

### 2015-08-04 A4/prime_procs.c

@@ -10,8 +10,8 @@
 
 #include "dynarr.h"
 
-#define MAX_PRIME 1000000
-#define PROCESSES 1
+long MAX_PRIME;
+int PROCESSES;
 
 int* primes;
 int* happiness;
@@ -91,7 +91,11 @@ void mark_multiples(int m) {
 	exit(EXIT_SUCCESS);
 }
 
-int main() {
+int main(int argc, char** argv) {
+	
+	MAX_PRIME = strtol(argv[2], NULL, 10);
+	PROCESSES = atoi(argv[1]);
+	
 	long m = 1;											// Current sieve prime, initialized to 1
 	long MSIZE = MAX_PRIME * sizeof(int);
 	int child_pids[PROCESSES];

### 2015-08-04 A4/prime_threads.c

@@ -9,11 +9,11 @@
 
 #include "dynarr.h"
 
-#define MAX_PRIME 100000000
-#define THREADS 1
+long MAX_PRIME;
+int  THREADS;
 
-int primes[MAX_PRIME];
-int happiness[MAX_PRIME];
+int* primes;
+int* happiness;
 
 static pthread_mutex_t mutex = PTHREAD_MUTEX_INITIALIZER;				// mutex protects thread_termination conditional variable
 static pthread_cond_t thread_termination = PTHREAD_COND_INITIALIZER;
@@ -105,7 +105,14 @@ static void* find_happiness(void* arg) {
 	return NULL;
 }
 
-int main() {
+int main(int argc, char** argv) {
+	
+	MAX_PRIME = strtol(argv[2], NULL, 10);
+	THREADS = atoi(argv[1]);
+	
+	primes = malloc(MAX_PRIME * sizeof(int));
+	happiness = malloc(MAX_PRIME * sizeof(int));
+	
 	long m = 1;									// Current sieve prime, initialized to 1
 	
 	pthread_mutex_lock(&mutex);					// Lock thread_termination, only unlock when blocking on pthread_cond_wait

### 2015-08-10 A4/writeup.tex
Add .gitignore and writeup.

@@
 \documentclass[11pt]{article}
+
+\usepackage{graphicx}
 
@@
 Elliott Capek \\
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.75\textwidth]{image.png}
+\caption{Data from timing threaded versus process programs.
+This is strange behavior.
+For this range of numbers, 100 million, adding more processes makes the situation worse once you have more than eight processes or one thread.
+This probably shows a problem in my implementation, although for reasons described below, this level of work might not actually benefit from more parallelization.
+The main thread calls new threads every time it gets a timeslice.
+However if the threads are completing in between main() timeslices, then more threads doesn't mean more speed..}
+\end{figure}
 

### 2015-08-14 A5/compute.c
Do lots of stuff, namely make compute better and communicate with manage.py. To do: make compute do math.

@@ -8,29 +8,41 @@
 #include <time.h>
 #include <unistd.h>
 
-const char* MANAGE_SOCKNAME = "/tmp/CS344_manage_sock";
 
-pthread_mutex_t mut;
-pthread_cond_t cond;
+// make global variables static so they can't interfere with other files
+static const char* MANAGE_SOCKNAME = "/tmp/CS344_manage_sock";
 
-long compute_flops() {
+static void* g_data;
+static int g_data_len;
+
+static pthread_mutex_t death_mut;
+static pthread_cond_t death_cond;
+
+static pthread_mutex_t data_mut;
+static pthread_cond_t data_cond;
+
+long long compute_flops() {
   printf("starting compute flops\n");
   int* buf = malloc(40000);
-  int i = 0;
-  int j, sum;
+  long long i = 0;
+  int j, c, sum;  // i is the candidate perfect int. j is the possible
+		  // divisor. sum is the sum of all divisors. c makes
+		  // sure we do 1000 perfect calculations before
+		  // checking time.
   int curtime = time(NULL);
   while (time(NULL) < curtime + 10) {
-    sum = 0;
-    for (j=2; j<i; j++) {
-      if (i % j == 0)
-	{ sum += j; }
+    for (c=0; c<1000; c++) {
+      sum = 0;
+      for (j=2; j<i; j++) {
+	if (i % j == 0)
+	  { sum += j; }
+      }
+      if (sum == i)
+	{ buf[i%40000] = 1; }
+      i++;
     }
-    if (sum == i)
-      { buf[i%40000] = 1; }
-    i++;
   }
   free(buf);
-  printf("ending compute flops\n");
   return (i*i/2 - 2*i) / 10;       // Return # of operations done per second
   
 }
@@ -60,7 +72,8 @@ void* manage_communicate(void* arg) {
 
   struct sockaddr_un manage_sock_struct;
   int s;
-  //char read_buf[100];
+  int READ_BUF_MAX = 1000;
+  char read_buf[READ_BUF_MAX];
   int sfd = socket(AF_UNIX, SOCK_STREAM, 0);
   
   manage_sock_struct.sun_family = AF_UNIX;
@@ -69,30 +82,37 @@ void* manage_communicate(void* arg) {
 
   s = connect(sfd, (const struct sockaddr*) &manage_sock_struct, sizeof(struct sockaddr_un));
[…]

### 2015-08-14 A5/manage.py

@@ -1,6 +1,7 @@
 #! /usr/bin/env python2
 
 import atexit
+import math
 import re
 import socket
 import sys
@@ -36,7 +37,14 @@ def handle_report_connection(conn, json_dict):
         terminate_computation()
     
 def handle_compute_connection(conn, json_dict):
-    print json_dict
+
+    r = int(json_dict['flops'])  # rate
+    start = 0   # perfect number lower bound
+    end = math.floor(2 + math.sqrt(4 - 2*(-r*15 + (1/2)*start^2 - 2*start)))
+    response_json = "{\"job_range\":\"" + str(start) + "-" + str(end) + "\"}"
+    print response_json
+    conn.send(response_json)
+    
     print "handling compute connection: " 
 
 def handle_exit(sock):

### 2015-08-14 A5/compute.c
Combine everything into a single thread in compute.c

@@ -12,16 +12,7 @@
 // make global variables static so they can't interfere with other files
 static const char* MANAGE_SOCKNAME = "/tmp/CS344_manage_sock";
 
-static void* g_data;
-static int g_data_len;
-
-static pthread_mutex_t death_mut;
-static pthread_cond_t death_cond;
-
-static pthread_mutex_t data_mut;
-static pthread_cond_t data_cond;
-
-long long compute_flops() {
+long long compute_rate() {
   printf("starting compute flops\n");
   int* buf = malloc(40000);
   long long i = 0;
@@ -47,7 +38,7 @@ long long compute_flops() {
   
 }
 
-void find_perfect_nums(int start, int end, int* nums) {
+void find_perfect_nums(int start, int end, char* is_perfect) {
   int i,j, sum;
   for (i=start; i<=end; i++)
   {
@@ -58,20 +49,25 @@ void find_perfect_nums(int start, int end, int* nums) {
 	{ sum += j; }
     }
     if (sum == i)
-      { nums[i-start] = 1; }
+      { is_perfect[i-start] = 1; }
   }
 }
 
-void* manage_communicate(void* arg) {
-  // Open active socket, connect to manage.py's bound socket, then read from the pipe until a message is received. Signal to the main thread that something happened.
-  // Things to communicate:
-  // First wait for main to cond_signal its done with performance testing. Open a socket with manage, send this information
-  // Get work, use a conditional variable to tell main to start working?
-  // send data when done, then go back to previous line
-  // Listen for kill signal, then raise the terminate signal
+void* listen_for_termination(void* args) {
+  while(1) {
+    ;
+  }
+  return NULL;
+}
+
+int main() {
+  // create new thread to block in manage_communicate, then block until
+  int s;
+  pthread_t kill_signal_listener;
+
+  pthread_create(&kill_signal_listener, NULL, listen_for_termination, NULL);  // Begin communication thread
 
   struct sockaddr_un manage_sock_struct;
-  int s;
   int READ_BUF_MAX = 1000;
   char read_buf[READ_BUF_MAX];
   int sfd = socket(AF_UNIX, SOCK_STREAM, 0);
[…]

### 2015-08-14 A5/manage.py

@@ -40,12 +40,12 @@ def handle_compute_connection(conn, json_dict):
 
     r = int(json_dict['flops'])  # rate
     start = 0   # perfect number lower bound
-    end = math.floor(2 + math.sqrt(4 - 2*(-r*15 + (1/2)*start^2 - 2*start)))
+    end = int(math.floor(2 + math.sqrt(4 - 2*(-r*15 + (1/2)*start^2 - 2*start))))
     response_json = "{\"job_range\":\"" + str(start) + "-" + str(end) + "\"}"
     print response_json
     conn.send(response_json)
-    
-    print "handling compute connection: " 
+    data = conn.recv( (end - start) + 100)
+    print "received data from compute."
 
 def handle_exit(sock):
     print "I'm exiting and handling it."

### 2015-08-14 A5/compute.c
Get compute.c to work. Now need to get kill signals to work, termination to be nice, test multiple instances of compute, and make report actually get nice data.

@@ -13,7 +13,6 @@
 static const char* MANAGE_SOCKNAME = "/tmp/CS344_manage_sock";
 
 long long compute_rate() {
-  printf("starting compute flops\n");
   int* buf = malloc(40000);
   long long i = 0;
   int j, c, sum;  // i is the candidate perfect int. j is the possible
@@ -24,7 +23,7 @@ long long compute_rate() {
   while (time(NULL) < curtime + 10) {
     for (c=0; c<1000; c++) {
       sum = 0;
-      for (j=2; j<i; j++) {
+      for (j=1; j<i; j++) {
 	if (i % j == 0)
 	  { sum += j; }
       }
@@ -38,18 +37,18 @@ long long compute_rate() {
   
 }
 
-void find_perfect_nums(int start, int end, char* is_perfect) {
-  int i,j, sum;
+void find_perfect_nums(long long start, long long end, char* is_perfect) {
+  long long i,j, sum;
   for (i=start; i<=end; i++)
   {
     sum = 0;
-    for (j=2; j<i; j++)
+    for (j=1; j<i; j++)
     {
       if (i % j == 0)
 	{ sum += j; }
     }
     if (sum == i)
-      { is_perfect[i-start] = 1; }
+      { is_perfect[i-start] = '1'; }
   }
 }
 
@@ -62,6 +61,7 @@ void* listen_for_termination(void* args) {
 
 int main() {
   // create new thread to block in manage_communicate, then block until
+
   int s;
   pthread_t kill_signal_listener;
 
@@ -94,33 +94,36 @@ int main() {
   if (s == -1)
     { perror("reading from socket");  exit(EXIT_FAILURE);  }
 
-  printf("manage.py response: %s\n", read_buf);
-
   char* start_range = strchr(read_buf, ':') + 2;
   char* mid_range = strchr(start_range, '-') + 1;
   char* end_range = strchr(mid_range, '"');
 
   char start_buf[100];
   char end_buf[100];
-  snprintf(start_buf, (size_t) (mid_range - start_range - 1), "%s", start_range);
-  snprintf(end_buf, end_range - mid_range, "%s", mid_range);
+
+  strncpy(start_buf, start_range, mid_range - start_range - 1);
+  start_buf[mid_range - start_range - 1] = '\0';
+  strncpy(end_buf, mid_range, end_range - mid_range);
+  end_buf[end_range - mid_range] = '\0';
+ 
   
[…]

### 2015-08-14 A5/manage.py

@@ -39,13 +39,17 @@ def handle_report_connection(conn, json_dict):
 def handle_compute_connection(conn, json_dict):
 
     r = int(json_dict['flops'])  # rate
-    start = 0   # perfect number lower bound
-    end = int(math.floor(2 + math.sqrt(4 - 2*(-r*15 + (1/2)*start^2 - 2*start))))
+    start = 1   # perfect number lower bound
+    end = int(math.floor(1 + math.sqrt(1 - 2*(-r*15 + (1/2)*start^2 - start))))
     response_json = "{\"job_range\":\"" + str(start) + "-" + str(end) + "\"}"
-    print response_json
+    
     conn.send(response_json)
-    data = conn.recv( (end - start) + 100)
-    print "received data from compute."
+    json_data = conn.recv( (end - start) + 13)
+    parsed_data = json_decode(json_data)
+    
+    for i in range(0, len(parsed_data["data"])):
+        if (parsed_data["data"][i] == '1'):
+            print "perfect: " + str(i+start)
 
 def handle_exit(sock):
     print "I'm exiting and handling it."

### 2015-08-14 A5/compute.c
Everything almost works. Just need to make sure everything responds to signals properly.

@@ -1,4 +1,5 @@
 #include <pthread.h>
+#include <signal.h>
 #include <stdio.h>
 #include <stdlib.h>
 #include <string.h>
@@ -33,7 +34,7 @@ long long compute_rate() {
     }
   }
   free(buf);
-  return (i*i/2 - 2*i) / 10;       // Return # of operations done per second
+  return (i*i/2) / 10;       // Return # of operations done per second
   
 }
 
@@ -53,10 +54,31 @@ void find_perfect_nums(long long start, long long end, char* is_perfect) {
 }
 
 void* listen_for_termination(void* args) {
-  while(1) {
-    ;
-  }
-  return NULL;
+  int sfd = socket(AF_UNIX, SOCK_STREAM, 0);
+  int READ_BUF_MAX = 1000;
+  char read_buf[READ_BUF_MAX];
+  struct sockaddr_un manage_sock_struct;
+
+  manage_sock_struct.sun_family = AF_UNIX;
+  strcpy(manage_sock_struct.sun_path, MANAGE_SOCKNAME);
+  
+  int s = connect(sfd, (const struct sockaddr*) &manage_sock_struct, sizeof(struct sockaddr_un));
+  if (s == -1)
+    { perror("creating term manage socket"); exit(EXIT_FAILURE); }
+
+  char* str = "{\"sender\":\"term\"}";
+  
+  s = write(sfd, str, strlen(str));
+  if (s == -1)
+    { perror("writing to socket");  exit(EXIT_FAILURE);  }
+
+  s = read(sfd, read_buf, READ_BUF_MAX);
+  if (s == -1)
+    { perror("reading from socket");  exit(EXIT_FAILURE);  }
+
+  printf("compute.c exiting due to -k flag.\n");
+  
+  exit(EXIT_SUCCESS);
 }
 
 int main() {
@@ -112,7 +134,7 @@ int main() {
   long long perfect_length = end - start;
   char* is_perfect = malloc( (end - start) * sizeof(char) );
   memset(is_perfect, '0', (size_t) (end - start) * sizeof(char));
-  
+
   find_perfect_nums(start, end, is_perfect);
   
   char data_packet[perfect_length + 13];

### 2015-08-14 A5/manage.py

@@ -1,4 +1,4 @@
-#! /usr/bin/env python2
+#! /usr/bin/env python3
 
 import atexit
 import math
@@ -8,23 +8,27 @@ import sys
 import subprocess
 import threading
 import time
+import os
 
 SERVER_SOCKNAME = "/tmp/CS344_manage_sock"
 MAX_READ_SIZE = 8192
 
-def terminate_computation():
-    print "sending termination signals to all compute instances i have knowledge of and then exiting."
+e = threading.Event()
+
+processes = []
+current_num = 1
+perf_nums = []
 
 def json_decode(msg):
     mydict = {}
     rx = re.compile(r'"([a-zA-Z]+)":"([a-zA-Z0-9]+)"')
-    match = rx.findall(msg)
+    match = rx.findall(str(msg))
 
     for l in match:
         mydict[l[0]] = l[1]
 
     rx2 = re.compile(r'"([a-zA-Z]+)":({.+})')
-    match = rx2.findall(msg)
+    match = rx2.findall(str(msg))
 
     for l in match:
         mydict[l[0]] = json_decode(l[1])
@@ -32,62 +36,105 @@ def json_decode(msg):
     return mydict
 
 def handle_report_connection(conn, json_dict):
-    conn.send("[packet containing info about processes, their performance, numbers we've tried and the perfect numbers.]")
+    global perf_nums
+    
+    outstr = "{\"perfects\":\""
+    for p in perf_nums:
+        outstr = outstr + str(p) + ","
+
+    outstr = outstr + "\",\"search-range\":\"1-" + str(current_num-1) + "\",\"processes\":\""
+    for proc in processes:
+        outstr = outstr + str(proc["pid"]) + ","
+
+    outstr = outstr + "\"}"
+        
+    conn.send(str.encode(outstr))
+
     if (json_dict["kill"] == "yes"):
-        terminate_computation()
+        handle_exit()
     
 def handle_compute_connection(conn, json_dict):
 
+    global current_num
+    global processes
+    global perf_nums
+    
+    proc_dict = {}
+    proc_dict["pid"] = json_dict["pid"]
+    proc_dict["flops"] = json_dict["flops"]
+    
     r = int(json_dict['flops'])  # rate
-    start = 1   # perfect number lower bound
-    end = int(math.floor(1 + math.sqrt(1 - 2*(-r*15 + (1/2)*start^2 - start))))
[…]

### 2015-08-14 A5/report.py

@@ -1,10 +1,22 @@
-#! /usr/bin/env python2
+#! /usr/bin/env python3
 import socket
 import getopt
 import sys
+import re
+import binascii
 
 SERVER_SOCKNAME = "/tmp/CS344_manage_sock"
 MAX_READ_SIZE = 8192
+
+def json_decode(msg):
+    mydict = {}
+    rx = re.compile(r'"([a-zA-Z\-]+)":"([a-zA-Z0-9,\-]*)"')
+    match = rx.findall(msg)
+
+    for l in match:
+        mydict[l[0]] = l[1]
+        
+    return mydict
 
 if __name__ == "__main__":
 
@@ -22,8 +34,18 @@ if __name__ == "__main__":
     else:
         msg = msg + '"kill":"no"}'
         
-    sock.send(msg)
+    sock.send(str.encode(msg))
     report = sock.recv(MAX_READ_SIZE)
+    report = json_decode(report.decode("utf-8"))
+    
     sock.close()
 
-    print report
+    if (str(report["processes"]) == ""):
+        print("processes running: None")
+    else:
+        print("processes running: " + str(report["processes"])[:-1])
+    print("perfect numbers found: " + str(report["perfects"])[:-1])
+    if (str(report["search-range"]) == "1-0"):
+        print ("Search range: Nothing searched yet")
+    else:
+        print("numbers searched: " + str(report["search-range"]))

### 2015-08-14 A5/writeup.tex
Add A5
last minute commit.

--- 2015-08-04 A4/writeup.tex
@@
 \documentclass[11pt]{article}
+
+\sepackage{graphicx}
 
@@
 \section{Worklog}
-\textbf{Friday July 31st 12:00pm-3:00pm} - Review TLPI book on topics: threads, mutexes, conditional variables, forking, shared memory, semaphores. Write design.\\
-\textbf{Friday July 31st 3:30pm-5:00pm} - Wrote is happy function. Wrote dynamic array implementation for use in ishappy. Wrote single-thread Sieve of Eratosthenes implementation. \\
-\textbf{Saturday August 1st 11:30pm-1:00pm} - Start on threaded implementation.
-Get application to successfully create threads and use mutexes on thread counter.
-Encountered weird bug where multiple threads would die when I expected only one to.
-Figured out it's because signaling a conditional variable doesn't immediately run the process waiting on it.
-Multiple threads would die and signal, then the waiting process would run.
-Fixed this bug by allowing for multiple thread deaths per wait.\\
-\textbf{Saturday August 1st 1:30pm-5:00pm} - Debugged another weird bug which had a really easy solution.
-I found that threads were behaving weirdly and somehow having their states set from alive to terminated even though they hadn't even started running yet.
-After much searching, found it was because I didn't make sure all threads were dead before going on to the next operation.
-One thread would sometimes stay alive long enough to make it into the next set of threads and then terminate, causing weird behavior.
-\\
-\textbf{Sunday August 2nd 11:00am-12:00pm} - Put finishing touches on threaded program. \\
-\textbf{Sunday August 2nd 12:30pm-3:00pm} - Began implementing multi-process program.
-Did more research on how I wanted to have the processes interact with each other.
-Successfully got processes to sieve through a range by modifying threaded program.\\
-\textbf{Sunday August 2nd 3:00pm-7:00pm} - Debugged minor issues, like shared memory not being truncated before use. Began implementing the ishappy functionality. \\
-\textbf{Sunday August 2nd 7:00pm-9:00pm} - Finished the ishappy functionality.
-Began researching how to write output to a file.
-For some reason spent lots of time thinking about how to take an integer and convert it into a binary-representation string, then sprintf to a file.
-\\
-\textbf{Monday August 3rd 11:00am-2:00pm} - Realized that write() was a much better solution than printing to a file.
-Implemented functionality to open happy primes file, truncate it, and write an unsigned int array to it.
-Spent a lot of time trying to figure out why my read() couldn't understand what I had just written to the file.
-Turned out I hadn't lseek()'d to the beginning of the file after I wrote, so of course I couldn't read what I had written.
-\\
-\textbf{Sunday August 2nd 2:30pm-5:00pm} - Put finishing touches on programs, removed unnecessary pieces. \\
-\textbf{Sunday August 2nd 5:00pm-7:00pm} - Moved code to server.
-Hit several problems.
-First, my programs couldn't be compiled initially because the heap arrays I was making were way too large (range from 0 to UINT MAX).
-Researched the icc flags to allow this behavior.
-Second, my programs kept getting killed the second they were executed.
-Internet says this is because the kernel thought they were too big at the time.
-Never fixed this bug.
-Found the key is to wait until not so many other procrastinating students are running their code.
-\\
-\textbf{Sunday August 2nd 7:00pm-10:00pm} - Begin timing code and composing writeup. \\
 
 \section{Challenges}
-This was a challenging but very good assignment.
-I learned a lot about how multithreaded and multiprocess programs work.
-This was an assignment where I had to spend lots of time reading TLPI to figure out how I wanted to make my program.
-I was challenged with different implementations - to use semaphores or block on waitpid()s, or to use signals.
-Etc.
-\\
-
-The hardest challenge was debugging a stochastic problem in my threaded program.
-It would show up maybe 1/3 of the time, but sometimes almost never, such that I thought it was gone.
-The problem was that I wasn't killing all my threads from a previous function (sieving through the primes), so when I tried to coordinate threads to figure out the happy primes of my sieve, some of the threads would die before I could use them.
-This screwed up the whole execution of the program, since I needed my thread-count variable to accurately reflect the number of threads alive and dead.
-It took a long time to figure out the source of the error, but I learned a lot.
-My takeaway from this is that programming slowly and diligently and making sure I think of all edge cases as I write a section of code is much better than quickly writing up a function and debugging it later.
-\\
 
 \section{Questions}
-\textbf{Point of assignment}: The point of this assignment is to have students become familiar with both threads and processes.
-This assignment is an introduction into parallel programming and sheds light on which approach (threads or processes) is appropriate to what situation.
-It also teaches how to have different processes communicate via signals, conditional variables, semaphores and mutexes.
-It also has students learn about how shared memory can be used to easily and quickly link program memory to disk memory.\\
+\textbf{Point of assignment}:
 
-\textbf{Correctness testing}: To test for correctness, I used the OEIS file for the first 10,000 happy primes.
-I had both my programs output data using the same formatting as the web database.
-I then did a diff between my outputted files and a downloaded copy of the database.
-The diff had no output, showing there was no difference between my file and the database.
-To test my file was being properly binary-dumped into a file, I read() that file and did a line-by-line comparison of the written and read arrays.
-\\
+\textbf{Correctness testing}:
 
-\textbf{What was learned}: My takeaway from this assignment is that threads are much easier to use than processes for parallel computing.
-When I was writing the programs I felt that threads were a much more elegant solution, since they already have shared memory implicitly (although I was very surprised by how easy to use shared memory was).
-Processes were slightly easier to use in that they just signaled when they were about to terminate and didn't need to use a mutex, whereas threads needed to use a mutex to signal with a conditional variable.
-This is a very minor difference, though.
-I also learned that shared memory is incredibly powerful, since it allows processes to essentially share data without having to coordinate at all.
-\\
+\textbf{What was learned}:
 

### 2015-08-14 A5/compute.c

@@ -78,6 +78,7 @@ void* listen_for_termination(void* args) {
 
   printf("compute.c exiting due to -k flag.\n");
   
+  raise(SIGINT);
   exit(EXIT_SUCCESS);
 }
 

### 2015-08-14 A5/manage.py

@@ -87,7 +87,10 @@ def handle_compute_connection(conn, json_dict):
 def handle_term_connection(conn, json_dict):
     global e
     e.wait()
-    conn.send(b'{"kill":""}')
+    try:
+        conn.send(b'{"kill":""}')
+    except:
+        pass
            
 def handle_exit():
     global e
@@ -95,7 +98,7 @@ def handle_exit():
     e.clear()
     time.sleep(1)
     subprocess.call(["rm", SERVER_SOCKNAME])
-    print("manage.py exiting due to -k flag.")
+    print("manage.py exiting.")
     os._exit(0)
 
 def handle_new_connection(conn):

### 2015-08-14 A5/writeup.tex
Add writeup.tex

@@
 
-\sepackage{graphicx}
+\usepackage{graphicx}
 
 \begin{document}
-Elliott Capek \\
+\title{CS344 Assignment 5: Client-Server Architecture}
+\author{Elliott Capek}
+\maketitle
 
 \section{Design}
+As suggested in the assignment description, my code will be broken up into three segments.\\ \\
+manage.py will be the server.
+Its main loop will only listen for new connections.
+The second it gets a new connection in its main socket, it will accept() the connection and wait for the connecting process to identify itself via a JSON package containing its identity.
+Once this is determined, the main thread will create a new thread to handle this connection.
+A connection to compute.c will start a new computation.
+First, this connecting peer process will be added to the list of currently executing compute.c instances, along with its pid.
+compute.c will, in its first communication, also send its performance.
+This performance will be taken into account in determining how much work to give it.
+manage.py will determine how much work to do, then send a JSON packet to its compute.c peer with instructions.
+compute.c will find the perfect numbers in the specified range, send the results back, and quit.
+Its info will then be removed from the manage.py list of running processes and the perfect numbers it found will be added to the manage.py list.\\ \\
+compute.c will begin its life by computing how performant it is.
+It will do this by running the perfect number algorithm on an empty data set for a large number of seconds, then averaging how much work it can do per second.
+It will only then open communication with manage.py.
+It will send its PID and its performance.
+manage.py will then calculate a good range for it to compute, and will respond with a JSOn packet containing this information.
+compute.c will then run this computation, which will hopefully take roughly 15 seconds.
+Once this is over, it will send a response packet to manage.py and terminate.
+compute.c will also spawn a listener thread that will have a separate socket to manage.py.
+It will listen for a kill signal and interrupt itself when it received it.\\ \\
+report.py will be only in communication with manage.py.
+report.py will first open up communication and send an update request.
+manage.py will then gather the required information: currently executing processes, found perfect numbers and currently searched range.
+manage.py will then send this information back to report.py, which will then quit.
+report.py will also have a kill section in its JSON connection packet which will have a value of yes or no.
+If yes, manage.py will first report its info to report, but will then send a kill signal to all its compute.c instances.\\ \\
+I will have a regular expression JSON parser that will identify sections of JSON and parse them. This code will be used in manage.py and report.py.\\ \\
+When I actually implemented my program, I diverged in a couple places.
+I was thinking of having three threads in compute.c: one to do the calculation, one to handle manage.py communication and one to handle listening for termination signals.
+However, it turned out that the first two threads basically never have a chance to do simultaneous work, so no performance boosts are gained by having them run separately.
+So I combined them into a single thread, which made life much easier.
+Also, instead of having manage.py first wait for an introduction write once a connection has been accepted, I ended up just creating a new thread to read the introduction and decide what to do.
+The reason for this is that compute.c first connects, then does its performance calculation, then sends an introduction signal.
+This ten second lag ended up blocking other processes from communicating.\\
 
 \section{Worklog}
+\textbf{Monday August 10th, 5:00pm} - Write framework of three programs. Main threads, communication threads and helper functions.\\
+\textbf{Tuesday August 11th, 5:00pm} - Tie together framework by adding threading functions and conditional variables.\\
+\textbf{Tuesday August 11th, 6:00pm} - Add socket code in compute.c to send and receive signals from manage.py.\\
+\textbf{Tuesday August 11th, 8:00pm} - Add socket code to replace.py as well. All programs can now communicate with dummy messages.\\
+\textbf{Wednesday August 12th, 11:00am} - Begin working on constructing JSON messages in compute.\\
+\textbf{Wednesday August 12th, 12:30pm} - Add more JSON message construction to other programs.\\
+\textbf{Wednesday August 12th, 2:00pm} - Begin working on JSON parser function in manage.py. Parser creates dictionary and populates it with string keys and string/dictionary values.\\
+\textbf{Wednesday August 12th, 3:00pm} - Debug JSON parser. Fix and implement recursive JSON parsing.\\
+\textbf{Wednesday August 12th, 5:00pm} - Implement find\_perfect\_numbers function in compute.c\\
+\textbf{Wednesday August 12th, 7:00pm} - Implement performance tester, which is very similar to find\_perfect\_numbers().
+Basically outputs the number of modulus operations (which I assume to be the most important operations in the perfect finding algorithm) done in a second by the function.\\
+\textbf{Wednesday August 12th, 8:30pm} - Add performance reporting to compute.c JSON communication.
+manage.py can now find the endpoint given a certain starting point of a range that will take 15 seconds to compute.\\
+\textbf{Wednesday August 12th, 9:30pm} - Debug the JSON parser.\\
+\textbf{Thursday August 13th, 11:00am} - Start testing find\_perfect\_numbers() function. Spend some time debugging it.\\
+\textbf{Thursday August 13th, 1:00pm} - Run into trouble with saving results of the perfect finding to a char array.
+I was setting values in a char array to integer 0s and 1s, then having trouble printing and transmitting the string due to length problems.
+It occured to me that '0' and 0 take up the same amount of room in a char array, so I just started using characters and it got much easier.\\
+\textbf{Thursday August 13th, 2:00pm} - Implement compute.c JSON interpretation.
+compute.c can now successfully interpret messages from manage.py.
+Debug problem where, for very high ranges, I only get 20-30 numbers in the range.
+Realize this is actually correct, it takes 15 seconds to calculate the perfectness of huge numbers.\\
+\textbf{Thursday August 13th, 4:00pm} - Fix bug where compute blocks replace.py from communicating with manage.py - latency between compute connecting and sending signal is 15 seconds.
+Add new thread to handle new connections so blocking on waiting for read to work in manage.py doesn't happen.\\
+\textbf{Thursday August 13th, 5:00pm} - Add termination system to manage.py - keep threads open with sockets pointing to all compute instances.
+Signal them when manage.py gets a -k packet from response.py.\\
+\textbf{Thursday August 13th, 7:00pm} - Debugging a problem with termination system.\\
+\textbf{Thursday August 12th, 10:00pm} - Writeup.\\
 
 \section{Challenges}
+This project's biggest challenge was probably layout.
+It required more advanced planning than the other projects.
+I think writing foundation code and prototypes before actually implementing helped me out a lot, since I knew where everything would fit already.
+Planning out the chronology of how communication would happen was another must.
+Planning how compute.c would first calculate its own performance, then send this and receive further instructions before hand allowed me to write out the skeleton of the communication pretty quickly.
+I knew exactly how many socket writes and reads would happen, so implementation was easy.\\ \\
+There were a couple difficult bugs, such as getting the processes to have the right number of threads with open sockets pointing at other programs.
+It was surprisingly easy to do actual socket communication.
+Blocking on reads and writes makes communication incredibly easy.\\ \\
+Another big challenge I ran into was termination.
+It was hard to figure out what was the best way to do it, but I ended up having an open socket connection throughout the life of the compute.c process.
+The second the -k flag was received from report.py, the socket thread in manage.py is prompted by an event (which I take to be something like a conditional variable) to send the kill signal to its compute.c peer process.
+This took some thinking, but I like my implementation.\\ \\
 
 \section{Questions}
-\textbf{Point of assignment}:
-
-\textbf{Correctness testing}:
-
-\textbf{What was learned}:
+\textbf{Point of assignment}: To get students familiar with coordinating programs with socket communication.
+To teach about how the server-client architecture works, and how easy it is to implement.
+To get students used to programming in python.
+To get students to learn about the JSON format.
+To get students used to using regular expressions to parse comlicated text.
+To further teach about signal handling and how signals can be used to terminate programs.\\ \\
+\textbf{Correctness testing}: To test for correctness I was going to do another diff on the output of report.py's perfect number list, but it turned out that the perfect numbers are sparse enough that they can just be eyeballed.
+To test my regular expressions I used online regex testers.
+To verify signals worked properly I used the kill linux command.
+I couldn't think of any good ways to test that the client-server architecture worked properly other than just trying different permutations of launching compute.c instances and report.py instances.\\
+\textbf{What was learned}:I learned that python's implementation of system programming features like threads, sockets and semaphores is similar to C's, although often simpler.
+I learned that I don't really like this.
+It was very easy to implement the python side of opening sockets and launching threads, but I found myself rushing a lot.
+I feel like python has so many features  that I am never using the exact right tool for the job.
+I felt like when I was implementing threads that I wasn't using conditional variables properly, since their API was a lot different.
+I don't really understand the lower levels of python's Threading.Event tool, even though I use it.
+Python seems to lend itself towards quickly using tools, since documentation is so in-depth and the tools themselves are easy to just implement by example.
+I kind of prefer C's approach, where everything is explicit and slow and thorough.\\ \\
+In terms of actual programming skills I gained, I learned that threads should only be used for truly asynchronous tasks.
+I was trying to have two threads coordinate using conditional variables, but this turned ugly very quickly.
+I learned that python's regular expression library is different than C's (I ended up having to use the finall() function, instead of the normal match() function, to find all matches to an expression).
+And I learned that the server-client architecture is very powerful and simple to use.\\
 

### 2015-08-14 A5/manage.py
finished.

@@ -9,6 +9,7 @@ import subprocess
 import threading
 import time
 import os
+import signal
 
 SERVER_SOCKNAME = "/tmp/CS344_manage_sock"
 MAX_READ_SIZE = 8192
@@ -78,10 +79,13 @@ def handle_compute_connection(conn, json_dict):
     json_data = conn.recv( (end - start) + 13)
     parsed_data = json_decode(json_data)
 
-    for i in range(0, len(parsed_data["data"])):
-        if (parsed_data["data"][i] == '1'):
-           perf_nums.append(i+start)
+    try:
+        for i in range(0, len(parsed_data["data"])):
+            if (parsed_data["data"][i] == '1'):
+                perf_nums.append(i+start)
 
+    except:
+        pass
     processes.remove(proc_dict)
 
 def handle_term_connection(conn, json_dict):
@@ -100,6 +104,16 @@ def handle_exit():
     subprocess.call(["rm", SERVER_SOCKNAME])
     print("manage.py exiting.")
     os._exit(0)
+
+def sig_handle_exit(one, two):
+    global e
+    e.set()
+    e.clear()
+    time.sleep(1)
+    subprocess.call(["rm", SERVER_SOCKNAME])
+    print("manage.py exiting.")
+    os._exit(0)
+
 
 def handle_new_connection(conn):
     msg = conn.recv(MAX_READ_SIZE)
@@ -129,6 +143,8 @@ if __name__ == "__main__":
     # thread to listen for new connections, then update internal data structure that maintains compute processes and their performance
     # once a new communication has started, create a thread that sends that process commands, waits for it to finish, then sends more commands
 
+    signal.signal(signal.SIGINT, sig_handle_exit)
+    signal.signal(signal.SIGQUIT, sig_handle_exit)
     subprocess.call(["rm", "-f", SERVER_SOCKNAME])
     sock = socket.socket(socket.AF_UNIX, socket.SOCK_STREAM)
     sock.bind(SERVER_SOCKNAME)

### 2015-08-14 final/writeup.tex
Add beginnings of final project.

\documentclass[11pt]{article}

\begin{document}
\title{CS344 Final: POSIX APIs and their Windows counterparts}
\author{Elliott Capek}
\maketitle

POSIX is a body of rules which specify an interface an operating system can provide to allow programmers to write portable programs. The actual implementation of the interface differs from OS to OS, but any POSIX-adopting OS's interface must have the same behavior. POSIX describes how programs interact with the file system and devices, how processes can access memory, OS information and communicate with each other, how processes can create new processes or threads, and many other things. \\ \\
This document will summarize the following POSIX APIs: memory mapping with \textbf{mmap}, thread creation with \textbf{pthreads}, interprocess communication with \textbf{sockets} and process creating with \textbf{fork}. It will also explain analogous Windows versions of these APIs and how they are different.\\

\section{Sockets}
Sockets are a tool in POSIX that can be used to communicate between processes. A socket is similar to a pipe in that, once two processes have succesfully opened their respective sockets and connected them, reads and writes can be done to the socket and the communication ``just works''. However, this is just one type of socket, known as a Stream Socket. The other major type is called a Datagram Socket, which is different in that individual buffered messages are packaged and sent, and the socket behaves less like a file. A more important difference between sockets and pipes is that sockets are not bound to the same operating system. When a socket is created it is given a protocol to use. The two major ones are the Unix domain and the Inet/Inet6 domain. The former is for communication within the same file system, while the latter is used for communication between networked machines over the internet. Sockets are the major way traffic is sent over the internet. HTTP, SSH, and other important programs/tools use sockets to achieve their inter-machine communication.

\subsection{Sockets in POSIX}

\end{document}

### 2015-08-14 final/writeup.tex
Add subsection on Posix socket usage.

@@
 \subsection{Sockets in POSIX}
+POSIX sockets all have a file descriptor they correspond to. In order to create a new socket for communication, the \textbf{socket(domain, type, protocol} system call must be used.\\ \\
+\textbf{domain} refers to the protocol to be used (local or internet communication).\\ \\
+\textbf{type} refers to the type of socket to be used: stream or datagram.
+Stream sockets are byte-streams.
+This means that read() and write() calls are used to send messages.
+Write() blocks until there is enough room in the socket buffer to hold the message.
+Read blocks until data has been written by the other socket holder.
+Closing the socket sends an EOF.
+Reading an EOF also closes the socket.
+Byte-streams are not conventional files.
+The lseek() command cannot be used on them, since data is destroyed once it is read.
+In a byte-stream, there is no concept of ``messages'' - data is written serially.
+This contrasts with the datagram socket, where messages are passed individually via send and receive calls.
+Datagram sockets do not guarantee messages are actually sent, or arrive in order.
+\textbf{protocol} is for future use, and currently has no influence on behavior.
+It should be 0.\\ \\
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{socket-chronology.png}
+\caption{Diagram of chronology of a typical streaming socket connection between a client and server.
+Image courtesy http://www.cs.uregina.ca/Links/class-info/330/Sockets/sockets.html and ``Unix Network Programming'' by W.
+Richard Stevens.}
+\label{fig:socket-chronology}
+\end{figure}
+
+Once a socket has been created, there is a default timeline of events for communication.
+Typically one process acts as the server, which opens a named socket and waits for a client process to connect.
+This chronology is shown in Figure (\ref{fig:socket-chronology}).
+First, \textbf{bind()} is used to give the socket a name.
+In local communication, this is typically a filename.
+For internet communication, this is a port.
+Other processes use this port, along with your computer's host name, to connect.
+Once the socket is bound, \textbf{listen()} is called to tell the kernel this socket is ready for a connection.
+After this, two things must happen, and they can happen in any order.
+The server must \textbf{accept()} a connection request, and a client must initiate a \textbf{connect()} request.
+The first call blocks until the second call occurs (sort of like a telephone call: one caller blocks until the other answers, or a timeout occurs).
+The \textbf{accept()} call returns the file descriptor of a new socket which is connected to the client application.
+Reading and writing (or sending and receiving) can now occur on the two sockets.\\ \\
+An interesting thing to note is that in POSIX, a socket is actually two different byte streams: one for reading and one for writing, just like a bidirectional pipe.
+One process implicitly writes to its write fd and reads from its read fd, even though only one socket file descriptor is used.
+This behavior is handled by the kernel.
+\\
+
+\subsection{Sockets in Windows}
 

### 2015-08-14 final/writeup.tex
Finish sockets section. Finish intro to mmap section.

@@
 
-\subsection{Sockets in Windows}
+\subsection{Sockets in Windows: WinSock}
+WinSock, the Windows implementation of sockets, is very similar to the POSIX implementation.
+There are \textbf{socket()}, \textbf{bind()}, \textbf{listen()}, \textbf{accept()}, \textbf{connect()}, \textbf{recv()} and \textbf{send()} calls which work in a very similar way to the POSIX calls.
+There are some calls with different names, such as \textbf{closesocket()}.
+The fact that the \textbf{close()} system call cannot be used on WinSock sockets is a result of WinSock sockets not actually being files, but different sorts of data structures.
+Windows does not have the ``everything is a file'' paradigm that Unix adopts.
+This is an advantage in that it puts sockets in a totally different namespace than files, so collisions are less likely.
+However, it limits users to only using the WinSock API to use sockets, instead of using other APIs to interact with the actual file, like in POSIX.
+Using non-socket APIs to interact with sockets probably is not very useful, but there may be some cases when it is nice.
+Another necessary function when using sockets in Windows is the \textbf{WSAStartup()} call, which must be used before any other socket calls are run.
+This function loads the Windows Sockets DLL (shared library).
+This is very different than the way POSIX does things, where shared libraries are automatically loaded.
+This gives more power to the programmer, since it allows them to specify versions and things explicitly.
+However, it also may cause a performance drop, since DLLs probably take time to load.
+As expected, the \textbf{WSACleanup()} function must be called after sockets are done being used to unload the DLL.\\
+
+\subsection{POSIX vs Windows}
+As described above, there are many shared system calls between POSIX and Windows.
+One of the most notable differences is that, because Windows sockets are not files, \textbf{write()} and \textbf{read()} do not work on Windows sockets.
+Everything is done through calls to \textbf{recv()} and \textbf{send()}.
+This allows for less versatility (treating sockets like files makes many things easy in POSIX, like piping a command to a socket) but potentially more confusion, since many of the read() and write() flags used on files do not work on sockets.
+One thing that is immediately obvious when looking at the MSDS documentation for Windows Sockets is that Windows provides a huge number of socket-specific functions: functions to resolve host names via DNS, simultaneously open and read socket connections, send files over sockets, etc.
+These are some very powerful API calls that make programming in Windows nice.\\
+
+\section{Memory Mapping}
+Memory mapping is the act of creating linkages between regions in a program's memory and external data structures, either files or anonymous spaces in memory.
+This is incredibly convenient, since it allows programs to write to files as if they were memory (random access via pointers, memset, etc..).
+It also allows two processes to conveniently share large portions of data, since multiple processes can map their private memory to shared files or anonymous structures.
+There are two types of mappings: private and shared.
+The difference is intuitive: only one process can access a private mapping, whereas many processes can access a shared mapping.
+A private file mapping is a single process taking ownership of a file and using it.
+If another process maps this file, the kernel assures that changes made by other processes are invisible to to each process.
+This is used to initialize memory from a file.
+Private anonymous mappings allocate memory that only the calling process can use.
+Private anonymous mappings are very similar to malloc() calls.
+Shared file mappings allow multiple processes to easily and quickly use and modify the same set of data.
+Shared anonymous mappings use memory that is within the parent process's memory space, and so only children can access it, not other processes.\\
+
+\subsection{POSIX Memory Mapping}
 
 \end{document}
+
+References:
+http://johnnie.jerrata.com/winsocktutorial/
+https://msdn.microsoft.com/en-us/library/windows/desktop/ms740632(v=vs.85).aspx

### 2015-10-01 os-class/A2/writeup.tex
Change structure of directory. Make it a repository for all my school stuff: classes, resume, etc...

--- 2015-07-19 A2/writeup.tex
@@
 \section{Design}
-My program is a state machine which uses getopt to parse through commands. When it reads the "q" option, it puts itself in APPEND mode. This causes it to iterate through
-the further arguments to the program and parse them as input files. When it reads the "x" option, it puts itself in EXTRACT mode. This causes it to read the arfile argument
-and create files from that archive. Etc etc etc. \\
+My program is a state machine which uses getopt to parse through commands.
+When it reads the "q" option, it puts itself in APPEND mode.
+This causes it to iterate through the further arguments to the program and parse them as input files.
+When it reads the "x" option, it puts itself in EXTRACT mode.
+This causes it to read the arfile argument and create files from that archive.
+Etc etc etc.
+\\
 
@@
 
-July 17th 3:00pm - 1:00am: Fixed bugs in outputting archive file properly (\n-padding odd outputs). Spent a long time debugging delete option problem. \\
+July 17th 3:00pm - 1:00am: Fixed bugs in outputting archive file properly (newline-padding odd outputs). Spent a long time debugging delete option problem. \\
 
@@
 
-I also spent a considerable amount of time on the delete code. I was having problems with my pointer references and knowing where I was in the file. I spent an overly long
-amount of time trying to fix a bug, then eventually rewrote the entire section of code in about 20 minutes. I should have done that sooner. I think that for me, it is much 
-easier to rewrite code that involves strings and pointers than to debug existing code. \\
+I also spent a considerable amount of time on the delete code.
+I was having problems with my pointer references and knowing where I was in the file.
+I spent an overly long amount of time trying to fix a bug, then eventually rewrote the entire section of code in about 20 minutes.
+I should have done that sooner.
+I think that for me, it is much easier to rewrite code that involves strings and pointers than to debug existing code.
+\\
 
 \section{Questions}
-Point of assignment: To get us used to the level of difficulty of programming assignments, to give an introduction to system programming and file IO. To show how complicated
-even simple Linux tools are. \\
-Correctness testing: By running the ar command and checking for exact duplication of behavior. Also by checking to see if output from myar could be run with ar and 
-vice versa. I ran most of my tests with simple 3-character .txt files. The program works with more complicated text files; I can't imagine it working differently for larger
-files than 3-character ones. \\
-Learn: I learned next time I need to start earlier. Also I learned lots about IO and string manipulation.\\
+\textbf{Point of assignment}: To get us used to the level of difficulty of programming assignments, to give an introduction to system programming and file IO.
+To show how complicated even simple Linux tools are.
+\\ \\
+\textbf{Correctness testing}: By running the ar command and checking for exact duplication of behavior.
+Also by checking to see if output from myar could be run with ar and vice versa.
+I ran most of my tests with simple 3-character .txt files.
+The program works with more complicated text files; I can't imagine it working differently for larger files than 3-character ones.
+\\ \\
+\textbf{Learn}: I learned next time I need to start earlier. Also I learned lots about IO and string manipulation.\\ \\
 

### 2015-10-01 os-class/final/writeup.tex
Change structure of directory. Make it a repository for all my school stuff: classes, resume, etc...

\documentclass[11pt]{article}

\usepackage{graphicx}

\begin{document}
\title{CS344 Final: POSIX APIs and their Windows counterparts}
\author{Elliott Capek}
\maketitle

POSIX is a body of rules which specify an interface an operating system can provide to allow programmers to write portable programs. The actual implementation of the interface differs from OS to OS, but any POSIX-adopting OS's interface must have the same behavior. POSIX describes how programs interact with the file system and devices, how processes can access memory, OS information and communicate with each other, how processes can create new processes or threads, and many other things. \\ \\
This document will summarize the following POSIX APIs: memory mapping with \textbf{mmap}, thread creation with \textbf{pthreads}, interprocess communication with \textbf{sockets} and process creating with \textbf{fork}. It will also explain analogous Windows versions of these APIs and how they are different.\\

\section{Sockets}
Sockets are a tool in POSIX that can be used to communicate between processes. A socket is similar to a pipe in that, once two processes have succesfully opened their respective sockets and connected them, reads and writes can be done to the socket and the communication ``just works''. However, this is just one type of socket, known as a Stream Socket. The other major type is called a Datagram Socket, which is different in that individual buffered messages are packaged and sent, and the socket behaves less like a file. A more important difference between sockets and pipes is that sockets are not bound to the same operating system. When a socket is created it is given a protocol to use. The two major ones are the Unix domain and the Inet/Inet6 domain. The former is for communication within the same file system, while the latter is used for communication between networked machines over the internet. Sockets are the major way traffic is sent over the internet. HTTP, SSH, and other important programs/tools use sockets to achieve their inter-machine communication.

\subsection{Sockets in POSIX}
POSIX sockets all have a file descriptor they correspond to. In order to create a new socket for communication, the \textbf{socket(domain, type, protocol)} system call must be used. \textbf{domain} refers to the protocol to be used (local or internet communication). \textbf{type} refers to the type of socket to be used: stream or datagram. Stream sockets are byte-streams. This means that read() and write() calls are used to send messages. Write() blocks until there is enough room in the socket buffer to hold the message. Read blocks until data has been written by the other socket holder. Closing the socket sends an EOF. Reading an EOF also closes the socket. Byte-streams are not conventional files. The lseek() command cannot be used on them, since data is destroyed once it is read. In a byte-stream, there is no concept of ``messages'' - data is written serially. This contrasts with the datagram socket, where messages are passed individually via send and receive calls. Datagram sockets do not guarantee messages are actually sent, or arrive in order. \textbf{protocol} is for future use, and currently has no influence on behavior. It should be 0.\\ \\

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{socket-chronology.png}
\caption{Diagram of chronology of a typical streaming socket connection between a client and server. Image courtesy http://www.cs.uregina.ca/Links/class-info/330/Sockets/sockets.html and ``Unix Network Programming'' by W. Richard Stevens.}
\label{fig:socket-chronology}
\end{figure}

Once a socket has been created, there is a default timeline of events for communication. Typically one process acts as the server, which opens a named socket and waits for a client process to connect. This chronology is shown in Figure (\ref{fig:socket-chronology}). First, \textbf{bind()} is used to give the socket a name. In local communication, this is typically a filename. For internet communication, this is a port. Other processes use this port, along with your computer's host name, to connect. Once the socket is bound, \textbf{listen()} is called to tell the kernel this socket is ready for a connection. After this, two things must happen, and they can happen in any order. The server must \textbf{accept()} a connection request, and a client must initiate a \textbf{connect()} request. The first call blocks until the second call occurs (sort of like a telephone call: one caller blocks until the other answers, or a timeout occurs). The \textbf{accept()} call returns the file descriptor of a new socket which is connected to the client application. Reading and writing (or sending and receiving) can now occur on the two sockets.\\ \\
An interesting thing to note is that in POSIX, a socket is actually two different byte streams: one for reading and one for writing, just like a bidirectional pipe. One process implicitly writes to its write fd and reads from its read fd, even though only one socket file descriptor is used. This behavior is handled by the kernel. \\

\subsection{Sockets in Windows: WinSock}
WinSock, the Windows implementation of sockets, is very similar to the POSIX implementation. There are \textbf{socket()}, \textbf{bind()}, \textbf{listen()}, \textbf{accept()}, \textbf{connect()}, \textbf{recv()} and \textbf{send()} calls which work in a very similar way to the POSIX calls. There are some calls with different names, such as \textbf{closesocket()}. The fact that the \textbf{close()} system call cannot be used on WinSock sockets is a result of WinSock sockets not actually being files, but different sorts of data structures. Windows does not have the ``everything is a file'' paradigm that Unix adopts. This is an advantage in that it puts sockets in a totally different namespace than files, so collisions are less likely. However, it limits users to only using the WinSock API to use sockets, instead of using other APIs to interact with the actual file, like in POSIX. Using non-socket APIs to interact with sockets probably is not very useful, but there may be some cases when it is nice. Another necessary function when using sockets in Windows is the \textbf{WSAStartup()} call, which must be used before any other socket calls are run. This function loads the Windows Sockets DLL (shared library). This is very different than the way POSIX does things, where shared libraries are automatically loaded. This gives more power to the programmer, since it allows them to specify versions and things explicitly. However, it also may cause a performance drop, since DLLs probably take time to load. As expected, the \textbf{WSACleanup()} function must be called after sockets are done being used to unload the DLL.\\

\subsection{POSIX vs Windows: Sockets}
As described above, there are many shared system calls between POSIX and Windows. One of the most notable differences is that, because Windows sockets are not files, \textbf{write()} and \textbf{read()} do not work on Windows sockets. Everything is done through calls to \textbf{recv()} and \textbf{send()}. This allows for less versatility (treating sockets like files makes many things easy in POSIX, like piping a command to a socket) but potentially more confusion, since many of the read() and write() flags used on files do not work on sockets. One thing that is immediately obvious when looking at the MSDS documentation for Windows Sockets is that Windows provides a huge number of socket-specific functions: functions to resolve host names via DNS, simultaneously open and read socket connections, send files over sockets, etc. These are some very powerful API calls that make programming in Windows nice.\\

\section{Memory Mapping}
Memory mapping is the act of creating linkages between regions in a program's memory and external data structures, either files or anonymous spaces in memory. This is incredibly convenient, since it allows programs to write to files as if they were memory (random access via pointers, memset, etc..). It also allows two processes to conveniently share large portions of data, since multiple processes can map their private memory to shared files or anonymous structures. There are two types of mappings: private and shared. The difference is intuitive: only one process can access a private mapping, whereas many processes can access a shared mapping. A private file mapping is a single process taking ownership of a file and using it. If another process maps this file, the kernel assures that changes made by other processes are invisible to to each process. This is actually done by the kernel when allocating a program's text area: a mapping is made in the program's virtual memory to the file holding its binary instructions and the program counter points ot the first instruction. This is used to initialize memory from a file. Private anonymous mappings allocate memory that only the calling process can use. Private anonymous mappings are very similar to malloc() calls. Shared file mappings allow multiple processes to easily and quickly use and modify the same set of data. Shared anonymous mappings use memory that is within the parent process's memory space, and so only children can access it, not other processes.\\ \\

\subsection{POSIX Memory Mapping}
The \textbf{mmap(addr, length, prot, flags, fd, offset)} call is how a memory mapping is created in POSIX. addr is an optional argument (can be NULL) which specifies where the mapping should be created. If null, the kernel decides where the memory will exist. length is the size of the mapping in bytes. prot is a mask specifying whether the memory has read, write and execute permissions. flags specifies whether the mapping is private or shared. flags also specifies whether the mapping is anonymous, whether to clear the anonymous mapping with zeros first, etc. fd and offset are for file mappings (ignored for anonymous mappings) and specify the file descriptor and offset (lseek) of the desired file mapping. The \textbf{munmap()} function does the opposite of mmap.\\ \\

\subsection{Windows Memory Mapping: File Mapping}
As the name implies, the closest analogue of memory mapping in Windows is File Mapping, which only works on files. Windows has a very different jargon than POSIX when it comes to discussing mapping. A File View is an object created by File Mapping that points to a mapped file object in physical memory. A File View is analogous to the data pointer returned by mmap in POSIX. This FV can be written and read to just like a data pointer. Every time the file mapping object is swapped back to the hard drive, the changes are commited to disk. Every File View that points to a file can point to a subsection of the file. FVs are coherent in that changes made in one process are visible to all other processes which map to that same file mapping object. \\ \\
The API for actually creating a file mapping and opening it is fairly different in Windows. You must first call \textbf{CreateFile} to bring the file from disk into memory. This is roughly equivalent to the POSIX \textbf{open()} command. Once you have the file loaded into memory, you must then call \textbf{CreateFileMapping}, which goes ahead and creates a mapped version of the file. Both CreateFile and CreateFile Mapping can both have their own file permissions and sizes, which is unique to Windows. Once a CreateFileMapping object has been made, MapViewOfFile creates the actual file view that the program can use. An FV is basically a data pointer, and can be used almost identically.

\subsection{POSIX vs Windows: Memory Mapping}
The first major difference between memory mapping and File Mapping is the latter only works on files, not anonymous memory regions. Extra work is likely required to make an anonymous mapping that is not backed up by files on disk. In POSIX, a file is mapped into memory with open(), and this file object has its own permissions and size. mmap() can then be used to map a part of the process's virtual memory into the file, but this mapping has no permissions. It is just a memory object. In Windows, the CreateFile and CreateFile Mapping objects both have different sizes, which is strange. This allows for more flexibility, in that files can be implicitly stretched if not large enough, and processes can be restricted to their own particular part of the file, not necessarily the whole thing. Another difference is that mmap() returns a file pointer, whereas CreateFileMapping objects can create multiple pointers. In C, creating multiple pointers to a mmap() region must be done explicitly. The separation between the file object and the mapped file object in Windows is more complicated, but gives the user more control in that they can restrict a program's access to the file and prevent data collisions. The fact that Windows uses complicated data structures to represent file mappings (instead of just memory regions) gives Windows the ability to set many more features on a memory region, such as its permissions and size. This creates overhead but more power. \\

\section{Threading}
Multithreading is a common element of every modern operating system. Multithreading allows a process to duplicate its program counter, thusly giving it more flexibility, simplifying implementation of asynchronous processes like socket communication, and allowing it to do more work per unit time by having more processor time and possibly running on multple processors at once.\\ \\
For all their power, threads introduce a good deal of danger and complication to a program. Firstly, all threads share the same global data elements. Any time a program is manipulating the same data in two different ways at once without coordination, bad things are bound to happen. This is why threading implementations come with tools to aid in this coordination. There are different threading libraries. This discussion will focus on the \textbf{pthreads} library. pthreads implements a variety of coordination control mechanisms, the first and foremost of which are mutexes and conditional variables.\\ \\
Mutexes are very simple. They are global data elements that store information about the last thread to interact with them. Mutexes provide two operations: locking and unlocking. Locking a mutex ``checks it out'' to your thread, meaning you own it until you ``return'' the mutex by unlocking it. Locking and unlocking must be done in order by a single thread, unlocking of a mutex by a thread that doesn't own it is not permitted. The key feature of mutexes is attempting to lock a mutex causes a thread to block, or stop executing, until it can lock the mutex (ie the thread owning the mutex unlocks it). Mutexes are used to implicitly protect against threads accessing the same data at the same time. According to the mutex honor system, locking a mutex allows a thread to access some data element. Once it is done, the thread unlocks the mutex, loses access to that data, and permits other threads to access it. This means that no threads will simultaneously use a piece of data, since only one thread can own a mutex at any point in time.\\ \\
A conditional variable takes the mutex one step further by adding a signaling feature. A conditional variable is always associated with a mutex. The conditional variable adds a ``signaling'' feature which threads can use to talk to each other. The ``signal'' is like a switch which gets flipped for threads that are waiting for it to be flipped. Just blocking on a lock() call isn't always the answer to thread coordination. Suppose a thread wants to consume a resource that other threads create. Locking a mutex, checking if the resource has been populated, then exiting if it hasn't is slow. Instead, a consumer thread can wait for a conditional variable's signal. That way, it will only lock the mutex when producer threads signal that they have produced a resource to be consumed. \\

\subsection{POSIX pthreads}
The pthreads library has a variety of useful threading functions: \textbf{pthread\_create()}, \textbf{pthread\_start}, \textbf{pthread\_mutex\_lock()}, \textbf{pthread\_mutex\_unlock()}, \textbf{pthread\_cond\_signal()}, and many many more. Threads are identified by their \textbf{thread\_t} id, which is unique to each thread of a process. When a thread is created with \textbf{pthread\_create()}, it is given a thread ID, a starting function (unlike POSIX forking to create new processes, threads have unique starting locations), and arguments. It must then be started via pthread\_start(). pthread mutexes are lock()ed and unlock()ed just as would be expected. Mutexes must be initialized with \textbf{pthread\_mutex\_init()} and destroyed with \textbf{pthread\_mutex\_destroy()}. pthread conditional variables are initialized and destroyed similarly to mutexes. pthreads offers the functions \textbf{pthread\_cond\_wait()} and \textbf{pthread\_cond\_signal()} to use conditional variables to block and signal. \textbf{pthread\_cond\_signal()} does just that: it signals all threads waiting on \textbf{pthread\_cond\_wait()}. \textbf{pthread\_cond\_wait()} has three stages: it first unlock()s the mutex associated with the conditional variable, then it blocks waiting for the signal(), then it relocks the mutex. The surrounding unlock() and lock() calls are there because typically the waiting thread owns the data, then ``lends'' it out to producer threads to do something with, then reclaims it and operates on it.\\

\subsection{Windows Threads}
In the Windows threading library, \textbf{CreateThread} creates threads. This function takes more arguments than the pthreads version. CreateThread allows the user to specify the thread security, the default stack size, and additional creation flags, such as whether to start paused or unpaused. The security of the thread determines whether, when a child is spawned by the process, its children inherit whatever is returned by the thread. Windows also supplies the \textbf{WaitForObject} and \textbf{WaitForMultipleObjects} functions, which are very powerful. They allow the user to block execution until either one or multiple threads are terminated. pthreads has no single function that does this. Windows also implements things like Events, Critical Sections, Interlocking, and a variety of other tools to manage threads. All in all, the Windows threading library is a lot more featureful than the pthreads library. \\

\subsection{POSIX vs Windows: Threading}
The Windows threading API has several features that pthreads does not. First, threads have security levels. This security descriptor allows individual threads to have different rights and permissions, giving them a lot more nuance than the identically-permissioned pthreads threads. Windows threads can also have their initial stack size specified, which gives the user a lot more power over the performance of their programs. Also, Windows threads can either automatically start on creation, or start out paused and be resume()d later. This makes them slightly more powerful than the always-pause-born pthreads threads. The Windows \textbf{WaitForObject} and \textbf{WaitForMultipleObjects} functions are also very powerful. They allow the user to block execution of a main thread until all of its children are done doing their tasks. pthreads does not have anything like this. With pthreads the implementation of this behavior is a lot more complicated, using a conditional variable to signal the main thread that children threads are dead and increment/decrement a counter. Windows has a lot more functions than pthreads does, like functions to manipulate events and critical sections. \\ \\
The basic takeaway message is that pthreads and Windows can do the same things, but Windows makes things much easier. pthreads is perhaps more lightweight, but also requires much more coding to achieve the same effects. This is both a pro and a con. \\

\section{Process Creation}
In POSIX, all processes are children of other processes. Every process descends from another process, up to the progenitor init process. POSIX uses the \textbf{fork()} system call to essentially duplicate a program. Once duplicated, programs are free to modify their program text and execute different code. This is how all processes call other processes. The only difference between a parent process and its children processes immediately after forking is the parent retains its pid, whereas the child process gets a new pid and has its ppid equal to its parent. This is the way child processes differentiate themselves from their parents. Forking in POSIX is efficient because, until a child process changes its program text, it shares the same program text memory area with its parents, since this section is read only. In addition, after forking, the child process actually shares the same memory as its parent. The only difference is once the child attempts to write to this memory, Linux initiates the Copy-on-Write mechanism to duplicate the memory and let the child operate on it as its own separate process. This delayed duplication is efficient because many times a child program doesn't need to modify the heap, only stack variables. Only duplicating if necessary gives the Copy-on-Write mechanism a performance advantage. \\

\subsection{POSIX Process Creation}
\textbf{fork()} is a simple function - it takes no arguments. Once called, a new process is put into the kernel's process queue and begins executing at the same line as the parent process. Once created, both processes need to tell themselves apart, so they examine the output of fork() to see who is who. This child process then typically jumps to a particular function and begins working on something, or it exec()s into a new function to accomplish some task. \\ \\

In addition to the \textbf{execve()} function, POSIX provides many wrapper functions that provide additional functionality than just executing execve(). Some have a built in system() calls which interpret a command string and breaks it up into different function calls and handles piping and redirection. Other flavors of execve wrapper allow simple command names to be given instead of an actual binary executable to be copied. Other options exist as wel. \\ \\

Linux (not POSIX) implements a special \textbf{clone()} function which behaves similarly to fork(), except it allows the user to specify the program's default stack size, starting location and function parameters. It is more powerful than fork(), but usually not necessary since this level of control is not needed. \\

\subsection{Windows Process Creation}
The Windows version of \textbf{fork()}, \textbf{CreateProcess()}, takes many more arguments. CreateProcess allows the user to specify the command line arguments, thread/process handle inheritability, creation flags, environment variables, starting directory pointer and process info structure. This information allows CreateProcess to really customize the child that is being created. Most importantly, CreateProcess allows the user to specify what the new process is going to be. Created processes are not identical copies of the parent, but have their own program text off the bat.

\subsection{POSIX vs Windows: Process Creation}
CreateProcess is a much heavier function - it takes infinitely more arguments than fork()! As is the usual Windows implementation, CreateProcess gives the user all the control up front in a large function call. The advantage of this is that it allows lots of customization up front, especially if all the user is doing is creating a new process to run a certain program, like ``dir.'' However, the advantage of the fork() then execve() approach of POSIX is that it allows for much more creativity and varied behavior. If the child process is to run the same program as the parent, just a different function inside it, then fork() without execve() is an efficient way to handle this, due to the kernel's Copy-on-Write mechanism. CreateProcess doesn't have an easy way to do this, it always spawns a new process. While the same tasks are accomplishable in either system, the POSIX way usually involves more coding. However, forcing the user to write all this code can be good, since it forces everything to be done properly. The difference between the two calls isn't that large. Windows provides a lot more options for security and customizability, but the POSIX implementation is easier to use.

\section{Sources}

Rose, Johnnie. "Johnnie's Winsock Tutorial by Johnnie Rose, Jr." \textit{Johnnie's Winsock Tutorial by Johnnie Rose, Jr.} Web. 15 Aug. 2015.\\
http://johnnie.jerrata.com/winsocktutorial/. \\ \\

MSDN Socket Documentation. Web. 14 Aug. 2015.\\
https://msdn.microsoft.com/en-us/library/windows/desktop/ms740632(v=vs.85).aspx \\ \\

MSDN Threading Documentation. Web. 14 Aug. 2015. \\
https://msdn.microsoft.com/en-us/library/aa366556(VS.85).aspx \\ \\

MSDN Process Creation Documentation. Web. 14 Aug. 2015. \\
https://msdn.microsoft.com/en-us/library/windows/desktop/ms684863(v=vs.85).aspx \\ \\

Kerrisk, Michael. \textit{The Linux Programming Interface a Linux and UNIX System Programming Handbook}. San Francisco: No Starch, 2010. Print.

\end{document}

### 2015-10-01 A2/myar.c

@@ -1,751 +1,0 @@
-#include <stdio.h>
-#include <stdlib.h>
-#include <ctype.h>
-#include <getopt.h>
-#include <fcntl.h>
-#include <string.h>
-#include <sys/stat.h>
-#include <sys/time.h>
-#include <sys/types.h>
-#include <time.h>
-#include <utime.h>
-#include <dirent.h>
-
-#define MAX_ENTRY_NUMBER_SIZE 100
-#define OUT_DESCRIPTION_MAX_SIZE 200
-
-enum MODES {
-	APPEND = 1,
-	EXTRACT,
-	SUMMARIZE,
-	DELETE,
-	ADD
-};
-
-int main(int argc, char** argv) {
-	
-	char opt;
-	int mode = 0;
-	int verbose = 0;
-	char* ar_file;
-	
-	ar_file = argv[2];
-	
-	while ((opt = getopt(argc, argv, "q:x:t:vd:A")) != -1)
-	{
-		switch(opt)
-		{
-			case 'q':
-				if (mode != 0) {
-					printf("Error. Options q, x, t, d and A are exclusive.\n");
-					exit(EXIT_FAILURE);
-				}
-				if (argc < 4) {
-					printf("Error. Option -q requires two or more arguments: myar -q arfile file [file2 ...]\n");
-					exit(EXIT_FAILURE);
-				}
-				mode = APPEND;
-				break;
-				
-			case 'x':
-				if (mode != 0) {
-					printf("Error. Options q, x, t, d and A are exclusive.\n");
-					exit(EXIT_FAILURE);
-				}
-				if (argc != 3) {
-					printf("Error. -x option takes only one argument: myar -x arfile\n");
-					exit(EXIT_FAILURE);
-				}
-				mode = EXTRACT;
-				break;
-				
-			case 't':
-				if (mode != 0) {
-					printf("Error. Options q, x, t, d and A are exclusive.\n");
-					exit(EXIT_FAILURE);
-				}
-				if (argc != 3) {
-					printf("Error. -t option takes only one argument: myar -t arfile\n");
-					exit(EXIT_FAILURE);
-				}
-				mode = SUMMARIZE;
-				break;
-				
-			case 'v':
-				verbose = 1;
-				break;
-				
-			case 'd':
-				if (mode != 0) {
-					printf("Error. Options q, x, t, d and A are exclusive.\n");
-					exit(EXIT_FAILURE);
-				}
[…]

### 2015-10-01 A3/mysh.c

@@ -1,436 +1,0 @@
-#include <fcntl.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <unistd.h>
-#include <signal.h>
-#include <sys/stat.h>
-#include <sys/types.h>
-#include <sys/wait.h>
-
-enum MODES {
-	REDIR_OUT_APP,
-	REDIR_OUT_OVW,
-	REDIR_IN,
-	REDIR_IN_HERE,
-	PIPE
-};
-
-int mysystem(char* command_string);
-char* get_command_path(char* command);
-void run_command_str(char** arr_words, int num_words, int out_fd, int in_fd);
-void command_string_slice(char* command_string, char*** argument_string_ptr, int* num_args);
-
-void signore(int sig) {
-	;
-}
-
-int main(int argc, char** argv) {
-	// First, implement the system() library call
-	int BUF_SIZE = 100000;
-	char buf[BUF_SIZE];
-	int read_bytes;
-	
-	signal(SIGINT, signore);
-	signal(SIGQUIT, signore);
-	
-	char pwdbuf[2048];
-	
-	while(1)
-	{
-		getcwd(pwdbuf, 2048);
-		printf("%s: ", pwdbuf);
-		fflush(stdout);
-		if ((read_bytes = read(STDOUT_FILENO, buf, BUF_SIZE)) == -1)
-		{
-			perror("Couldn't read from stdout: ");
-			exit(EXIT_FAILURE);
-		}
-		buf[read_bytes-1] = '\0';
-		if (strcmp(buf, "exit") == 0){
-			printf("Exiting shell.\n");
-			exit(EXIT_SUCCESS);
-		} else if (read_bytes > 1) {
-			mysystem(buf);
-		}
-	}
-	
-	return 0;
-}
-
-int mysystem(char* command_string) {
-	int cmd_start = 0;
-	int cmd_current = 0;
-	
-	int saved_stderr = dup(STDERR_FILENO);
-	
-	int piping = 0;
-	int pipe_fds[2];
-	
-	char** argument_string;
-	int num_args;
-
-	command_string_slice(command_string, &argument_string, &num_args);
-	
-	while (cmd_current < num_args)
-	{
-		if (cmd_current == num_args-1)
-		{
-			if (piping)
-			{
-				run_command_str(&argument_string[cmd_start], cmd_current - cmd_start + 1, STDOUT_FILENO, pipe_fds[0]);
-				piping = 0;
-				if (close(pipe_fds[0]) == -1)
-				{
[…]

### 2015-10-01 A4/dynarr.c

@@ -1,35 +1,0 @@
-#include <stdlib.h>
-
-#include "dynarr.h"
-
-void arr_init(DynIntArr* arr, int s) {
-	arr->size = 0;
-	arr->maxsize = s;
-	arr->data = malloc(sizeof(int) * arr->maxsize);
-}
-
-void resize(DynIntArr* arr, int newsize) {
-	int* new_data = malloc(sizeof(int) * newsize);
-	int i;
-	for (i = 0; i<arr->size; i++) 
-	{
-		new_data[i] = arr->data[i];
-	}
-	free(arr->data);
-	arr->data = new_data;
-}
-
-void append(DynIntArr* arr, int n) {
-	arr->size++;
-	if (arr->size < arr->maxsize) {
-		arr->data[arr->size-1] = n;
-	}
-	else {
-		resize(arr, arr->maxsize*2);
-		arr->data[arr->size-1] = n;
-	}
-}
-
-void arr_free(DynIntArr* arr) {
-	free(arr->data);
-}

### 2015-10-01 A4/dynarr.h

@@ -1,10 +1,0 @@
-typedef struct {
-		int* data;
-		int size;
-		int maxsize;
-} DynIntArr;
-
-void arr_init(DynIntArr* arr, int s);
-void resize(DynIntArr* arr, int newsize);
-void append(DynIntArr* arr, int n);
-void arr_free(DynIntArr* arr);

### 2015-10-01 A4/prime_procs.c

@@ -1,321 +1,0 @@
-#include <limits.h>
-#include <fcntl.h>
-#include <math.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <sys/mman.h>
-#include <semaphore.h>
-#include <unistd.h>
-
-#include "dynarr.h"
-
-long MAX_PRIME;
-int PROCESSES;
-
-int* primes;
-int* happiness;
-
-int is_happy(long num) {
-	DynIntArr* arr = malloc(sizeof(DynIntArr));
-	arr_init(arr, 10);
-	
-	while(num != 1) {						// Loop until found happy or sad
-		long new_num = 0;
-		while (num / 10 != 0) {				// While >1 digit
-			new_num += pow(num % 10, 2);	// num % 10 = least sig. digit
-			num = num / 10;					// Chop off last digit
-		}
-		new_num += pow(num, 2);
-		num = new_num;
-		
-		long i;								// Check if already hit this #
-		for (i=0; i<arr->size; i++)
-		{
-			if (num == arr->data[i])
-			{
-				return 0; // is sad
-			}
-		}
-		append(arr, num);
-	}
-	return 1; // is happy
-	arr_free(arr);
-	free(arr);
-	
-	return 1;
-}
-
-int get_next_sieve_prime(int start) {
-	long m = start + 1;
-	while (m <= floor(sqrt(MAX_PRIME)))
-	{
-	  	if (primes[m-1] == 1) { return m; }		// m is prime, return as next sieve prime
-		m++;
-	}
-	return -1;
-}
-
-static void* find_happiness(int section) {
-	long MSIZE = MAX_PRIME * sizeof(int);
-	
-	long i;
-	for (i = section*MAX_PRIME/PROCESSES + 1; i<=(section*MAX_PRIME/PROCESSES + MAX_PRIME/PROCESSES); i++) {
-		if (primes[i] && is_happy(i+1)) { happiness[i] = 1; }
-	}
-	
-	if (msync(happiness, MSIZE, MS_SYNC) == -1)
-	{
-		perror("mysnc: ");
-		exit(EXIT_FAILURE);
-	}
-	
-	exit(EXIT_SUCCESS);
-}
-
[…]

### 2015-10-01 A4/prime_threads.c

@@ -1,262 +1,0 @@
-#include <fcntl.h>
-#include <limits.h>
-#include <math.h>
-#include <pthread.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <unistd.h>
-
-#include "dynarr.h"
-
-long MAX_PRIME;
-int  THREADS;
-
-int* primes;
-int* happiness;
-
-static pthread_mutex_t mutex = PTHREAD_MUTEX_INITIALIZER;				// mutex protects thread_termination conditional variable
-static pthread_cond_t thread_termination = PTHREAD_COND_INITIALIZER;
-
-enum t_state {
-	THREAD_LIVE,
-	THREAD_DEAD,
-	THREAD_JOINED
-};
-
-typedef struct {
-	pthread_t tid;
-	long num;
-	long sieve_prime;
-	enum t_state state;
-} thread_struct;
-
-int is_happy(long num) {
-	DynIntArr* arr = malloc(sizeof(DynIntArr));
-	arr_init(arr, 10);
-	
-	while(num != 1) {						// Loop until found happy or sad
-		long new_num = 0;
-		while (num / 10 != 0) {				// While >1 digit
-			new_num += pow(num % 10, 2);	// num % 10 = least sig. digit
-			num = num / 10;					// Chop off last digit
-		}
-		new_num += pow(num, 2);
-		num = new_num;
-		
-		long i;								// Check if already hit this #
-		for (i=0; i<arr->size; i++)
-		{
-			if (num == arr->data[i])
-			{
-				return 0; // is sad
-			}
-		}
-		append(arr, num);
-	}
-	return 1; // is happy
-	arr_free(arr);
-	free(arr);
-	
-	return 1;
-}
-
-int get_next_sieve_prime(long start) {
-	long m = start + 1;
-	while (m <= floor(sqrt(MAX_PRIME)))
-	{
-	  	if (primes[m-1] == 1) { return m; }		// m is prime, return as next sieve prime
-		m++;
-	}
-	return -1;
-}
-
-static void* mark_multiples(void* arg) {
-	long j = 2;
-	thread_struct* thread_info = (thread_struct*) arg;
-	long m = thread_info->sieve_prime;
-	
-	while (m*j <= MAX_PRIME)
-	{
-		primes[m*j-1] = 0;
-		j++;
-	}
-	
[…]

### 2015-10-01 A5/compute.c

@@ -1,152 +1,0 @@
-#include <pthread.h>
-#include <signal.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <sys/types.h>
-#include <sys/un.h>
-#include <sys/socket.h>
-#include <time.h>
-#include <unistd.h>
-
-
-// make global variables static so they can't interfere with other files
-static const char* MANAGE_SOCKNAME = "/tmp/CS344_manage_sock";
-
-long long compute_rate() {
-  int* buf = malloc(40000);
-  long long i = 0;
-  int j, c, sum;  // i is the candidate perfect int. j is the possible
-		  // divisor. sum is the sum of all divisors. c makes
-		  // sure we do 1000 perfect calculations before
-		  // checking time.
-  int curtime = time(NULL);
-  while (time(NULL) < curtime + 10) {
-    for (c=0; c<1000; c++) {
-      sum = 0;
-      for (j=1; j<i; j++) {
-	if (i % j == 0)
-	  { sum += j; }
-      }
-      if (sum == i)
-	{ buf[i%40000] = 1; }
-      i++;
-    }
-  }
-  free(buf);
-  return (i*i/2) / 10;       // Return # of operations done per second
-  
-}
-
-void find_perfect_nums(long long start, long long end, char* is_perfect) {
-  long long i,j, sum;
-  for (i=start; i<=end; i++)
-  {
-    sum = 0;
-    for (j=1; j<i; j++)
-    {
-      if (i % j == 0)
-	{ sum += j; }
-    }
-    if (sum == i)
-      { is_perfect[i-start] = '1'; }
-  }
-}
-
-void* listen_for_termination(void* args) {
-  int sfd = socket(AF_UNIX, SOCK_STREAM, 0);
-  int READ_BUF_MAX = 1000;
-  char read_buf[READ_BUF_MAX];
-  struct sockaddr_un manage_sock_struct;
-
-  manage_sock_struct.sun_family = AF_UNIX;
-  strcpy(manage_sock_struct.sun_path, MANAGE_SOCKNAME);
-  
-  int s = connect(sfd, (const struct sockaddr*) &manage_sock_struct, sizeof(struct sockaddr_un));
-  if (s == -1)
-    { perror("creating term manage socket"); exit(EXIT_FAILURE); }
-
-  char* str = "{\"sender\":\"term\"}";
-  
[…]

### 2015-10-01 A5/manage.py

@@ -1,159 +1,0 @@
-#! /usr/bin/env python3
-
-import atexit
-import math
-import re
-import socket
-import sys
-import subprocess
-import threading
-import time
-import os
-import signal
-
-SERVER_SOCKNAME = "/tmp/CS344_manage_sock"
-MAX_READ_SIZE = 8192
-
-e = threading.Event()
-
-processes = []
-current_num = 1
-perf_nums = []
-
-def json_decode(msg):
-    mydict = {}
-    rx = re.compile(r'"([a-zA-Z]+)":"([a-zA-Z0-9]+)"')
-    match = rx.findall(str(msg))
-
-    for l in match:
-        mydict[l[0]] = l[1]
-
-    rx2 = re.compile(r'"([a-zA-Z]+)":({.+})')
-    match = rx2.findall(str(msg))
-
-    for l in match:
-        mydict[l[0]] = json_decode(l[1])
-        
-    return mydict
-
-def handle_report_connection(conn, json_dict):
-    global perf_nums
-    
-    outstr = "{\"perfects\":\""
-    for p in perf_nums:
-        outstr = outstr + str(p) + ","
-
-    outstr = outstr + "\",\"search-range\":\"1-" + str(current_num-1) + "\",\"processes\":\""
-    for proc in processes:
-        outstr = outstr + str(proc["pid"]) + ","
-
-    outstr = outstr + "\"}"
-        
-    conn.send(str.encode(outstr))
-
-    if (json_dict["kill"] == "yes"):
-        handle_exit()
-    
-def handle_compute_connection(conn, json_dict):
-
-    global current_num
-    global processes
-    global perf_nums
-    
-    proc_dict = {}
-    proc_dict["pid"] = json_dict["pid"]
-    proc_dict["flops"] = json_dict["flops"]
-    
-    r = int(json_dict['flops'])  # rate
-    start = current_num   # perfect number lower bound
-    end = int(math.sqrt(2*r*15+start*start))
-    current_num = end + 1
-    
-    proc_dict["start"] = start
-    proc_dict["end"] = end
-    processes.append(proc_dict)
-    
-    response_json = str.encode("{\"job_range\":\"" + str(start) + "-" + str(end) + "\"}")
-    
-    conn.send(response_json)
-    json_data = conn.recv( (end - start) + 13)
-    parsed_data = json_decode(json_data)
-
-    try:
-        for i in range(0, len(parsed_data["data"])):
[…]

### 2015-10-01 A5/report.py

@@ -1,51 +1,0 @@
-#! /usr/bin/env python3
-import socket
-import getopt
-import sys
-import re
-import binascii
-
-SERVER_SOCKNAME = "/tmp/CS344_manage_sock"
-MAX_READ_SIZE = 8192
-
-def json_decode(msg):
-    mydict = {}
-    rx = re.compile(r'"([a-zA-Z\-]+)":"([a-zA-Z0-9,\-]*)"')
-    match = rx.findall(msg)
-
-    for l in match:
-        mydict[l[0]] = l[1]
-        
-    return mydict
-
-if __name__ == "__main__":
-
-    killmode = False
-    
-    if (sys.argv[1:] == ['-k']):
-        killmode = True
-    
-    sock = socket.socket(socket.AF_UNIX, socket.SOCK_STREAM)
-    sock.connect(SERVER_SOCKNAME)
-
-    msg = '{"sender":"report",'
-    if (killmode):
-        msg = msg + '"kill":"yes"}'
-    else:
-        msg = msg + '"kill":"no"}'
-        
-    sock.send(str.encode(msg))
-    report = sock.recv(MAX_READ_SIZE)
-    report = json_decode(report.decode("utf-8"))
-    
-    sock.close()
-
-    if (str(report["processes"]) == ""):
-        print("processes running: None")
-    else:
-        print("processes running: " + str(report["processes"])[:-1])
-    print("perfect numbers found: " + str(report["perfects"])[:-1])
-    if (str(report["search-range"]) == "1-0"):
-        print ("Search range: Nothing searched yet")
-    else:
-        print("numbers searched: " + str(report["search-range"]))

### 2015-10-01 os-class/A2/myar.c

@@ -1,0 +1,751 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <ctype.h>
+#include <getopt.h>
+#include <fcntl.h>
+#include <string.h>
+#include <sys/stat.h>
+#include <sys/time.h>
+#include <sys/types.h>
+#include <time.h>
+#include <utime.h>
+#include <dirent.h>
+
+#define MAX_ENTRY_NUMBER_SIZE 100
+#define OUT_DESCRIPTION_MAX_SIZE 200
+
+enum MODES {
+	APPEND = 1,
+	EXTRACT,
+	SUMMARIZE,
+	DELETE,
+	ADD
+};
+
+int main(int argc, char** argv) {
+	
+	char opt;
+	int mode = 0;
+	int verbose = 0;
+	char* ar_file;
+	
+	ar_file = argv[2];
+	
+	while ((opt = getopt(argc, argv, "q:x:t:vd:A")) != -1)
+	{
+		switch(opt)
+		{
+			case 'q':
+				if (mode != 0) {
+					printf("Error. Options q, x, t, d and A are exclusive.\n");
+					exit(EXIT_FAILURE);
+				}
+				if (argc < 4) {
+					printf("Error. Option -q requires two or more arguments: myar -q arfile file [file2 ...]\n");
+					exit(EXIT_FAILURE);
+				}
+				mode = APPEND;
+				break;
+				
+			case 'x':
+				if (mode != 0) {
+					printf("Error. Options q, x, t, d and A are exclusive.\n");
+					exit(EXIT_FAILURE);
+				}
+				if (argc != 3) {
+					printf("Error. -x option takes only one argument: myar -x arfile\n");
+					exit(EXIT_FAILURE);
+				}
+				mode = EXTRACT;
+				break;
+				
+			case 't':
+				if (mode != 0) {
+					printf("Error. Options q, x, t, d and A are exclusive.\n");
+					exit(EXIT_FAILURE);
+				}
+				if (argc != 3) {
+					printf("Error. -t option takes only one argument: myar -t arfile\n");
+					exit(EXIT_FAILURE);
+				}
+				mode = SUMMARIZE;
+				break;
+				
+			case 'v':
+				verbose = 1;
+				break;
+				
+			case 'd':
+				if (mode != 0) {
+					printf("Error. Options q, x, t, d and A are exclusive.\n");
+					exit(EXIT_FAILURE);
+				}
[…]

### 2015-10-01 os-class/A3/test/mysh.c

@@ -1,0 +1,436 @@
+#include <fcntl.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <unistd.h>
+#include <signal.h>
+#include <sys/stat.h>
+#include <sys/types.h>
+#include <sys/wait.h>
+
+enum MODES {
+	REDIR_OUT_APP,
+	REDIR_OUT_OVW,
+	REDIR_IN,
+	REDIR_IN_HERE,
+	PIPE
+};
+
+int mysystem(char* command_string);
+char* get_command_path(char* command);
+void run_command_str(char** arr_words, int num_words, int out_fd, int in_fd);
+void command_string_slice(char* command_string, char*** argument_string_ptr, int* num_args);
+
+void signore(int sig) {
+	;
+}
+
+int main(int argc, char** argv) {
+	// First, implement the system() library call
+	int BUF_SIZE = 100000;
+	char buf[BUF_SIZE];
+	int read_bytes;
+	
+	signal(SIGINT, signore);
+	signal(SIGQUIT, signore);
+	
+	char pwdbuf[2048];
+	
+	while(1)
+	{
+		getcwd(pwdbuf, 2048);
+		printf("%s: ", pwdbuf);
+		fflush(stdout);
+		if ((read_bytes = read(STDOUT_FILENO, buf, BUF_SIZE)) == -1)
+		{
+			perror("Couldn't read from stdout: ");
+			exit(EXIT_FAILURE);
+		}
+		buf[read_bytes-1] = '\0';
+		if (strcmp(buf, "exit") == 0){
+			printf("Exiting shell.\n");
+			exit(EXIT_SUCCESS);
+		} else if (read_bytes > 1) {
+			mysystem(buf);
+		}
+	}
+	
+	return 0;
+}
+
+int mysystem(char* command_string) {
+	int cmd_start = 0;
+	int cmd_current = 0;
+	
+	int saved_stderr = dup(STDERR_FILENO);
+	
+	int piping = 0;
+	int pipe_fds[2];
+	
+	char** argument_string;
+	int num_args;
+
+	command_string_slice(command_string, &argument_string, &num_args);
+	
+	while (cmd_current < num_args)
+	{
+		if (cmd_current == num_args-1)
+		{
+			if (piping)
+			{
+				run_command_str(&argument_string[cmd_start], cmd_current - cmd_start + 1, STDOUT_FILENO, pipe_fds[0]);
+				piping = 0;
+				if (close(pipe_fds[0]) == -1)
+				{
[…]

### 2015-10-01 os-class/A4/dynarr.c

@@ -1,0 +1,35 @@
+#include <stdlib.h>
+
+#include "dynarr.h"
+
+void arr_init(DynIntArr* arr, int s) {
+	arr->size = 0;
+	arr->maxsize = s;
+	arr->data = malloc(sizeof(int) * arr->maxsize);
+}
+
+void resize(DynIntArr* arr, int newsize) {
+	int* new_data = malloc(sizeof(int) * newsize);
+	int i;
+	for (i = 0; i<arr->size; i++) 
+	{
+		new_data[i] = arr->data[i];
+	}
+	free(arr->data);
+	arr->data = new_data;
+}
+
+void append(DynIntArr* arr, int n) {
+	arr->size++;
+	if (arr->size < arr->maxsize) {
+		arr->data[arr->size-1] = n;
+	}
+	else {
+		resize(arr, arr->maxsize*2);
+		arr->data[arr->size-1] = n;
+	}
+}
+
+void arr_free(DynIntArr* arr) {
+	free(arr->data);
+}

### 2015-10-01 os-class/A4/dynarr.h

@@ -1,0 +1,10 @@
+typedef struct {
+		int* data;
+		int size;
+		int maxsize;
+} DynIntArr;
+
+void arr_init(DynIntArr* arr, int s);
+void resize(DynIntArr* arr, int newsize);
+void append(DynIntArr* arr, int n);
+void arr_free(DynIntArr* arr);

### 2015-10-01 os-class/A4/prime_procs.c

@@ -1,0 +1,321 @@
+#include <limits.h>
+#include <fcntl.h>
+#include <math.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <sys/mman.h>
+#include <semaphore.h>
+#include <unistd.h>
+
+#include "dynarr.h"
+
+long MAX_PRIME;
+int PROCESSES;
+
+int* primes;
+int* happiness;
+
+int is_happy(long num) {
+	DynIntArr* arr = malloc(sizeof(DynIntArr));
+	arr_init(arr, 10);
+	
+	while(num != 1) {						// Loop until found happy or sad
+		long new_num = 0;
+		while (num / 10 != 0) {				// While >1 digit
+			new_num += pow(num % 10, 2);	// num % 10 = least sig. digit
+			num = num / 10;					// Chop off last digit
+		}
+		new_num += pow(num, 2);
+		num = new_num;
+		
+		long i;								// Check if already hit this #
+		for (i=0; i<arr->size; i++)
+		{
+			if (num == arr->data[i])
+			{
+				return 0; // is sad
+			}
+		}
+		append(arr, num);
+	}
+	return 1; // is happy
+	arr_free(arr);
+	free(arr);
+	
+	return 1;
+}
+
+int get_next_sieve_prime(int start) {
+	long m = start + 1;
+	while (m <= floor(sqrt(MAX_PRIME)))
+	{
+	  	if (primes[m-1] == 1) { return m; }		// m is prime, return as next sieve prime
+		m++;
+	}
+	return -1;
+}
+
+static void* find_happiness(int section) {
+	long MSIZE = MAX_PRIME * sizeof(int);
+	
+	long i;
+	for (i = section*MAX_PRIME/PROCESSES + 1; i<=(section*MAX_PRIME/PROCESSES + MAX_PRIME/PROCESSES); i++) {
+		if (primes[i] && is_happy(i+1)) { happiness[i] = 1; }
+	}
+	
+	if (msync(happiness, MSIZE, MS_SYNC) == -1)
+	{
+		perror("mysnc: ");
+		exit(EXIT_FAILURE);
+	}
+	
+	exit(EXIT_SUCCESS);
+}
+
[…]

### 2015-10-01 os-class/A4/prime_threads.c

@@ -1,0 +1,262 @@
+#include <fcntl.h>
+#include <limits.h>
+#include <math.h>
+#include <pthread.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <unistd.h>
+
+#include "dynarr.h"
+
+long MAX_PRIME;
+int  THREADS;
+
+int* primes;
+int* happiness;
+
+static pthread_mutex_t mutex = PTHREAD_MUTEX_INITIALIZER;				// mutex protects thread_termination conditional variable
+static pthread_cond_t thread_termination = PTHREAD_COND_INITIALIZER;
+
+enum t_state {
+	THREAD_LIVE,
+	THREAD_DEAD,
+	THREAD_JOINED
+};
+
+typedef struct {
+	pthread_t tid;
+	long num;
+	long sieve_prime;
+	enum t_state state;
+} thread_struct;
+
+int is_happy(long num) {
+	DynIntArr* arr = malloc(sizeof(DynIntArr));
+	arr_init(arr, 10);
+	
+	while(num != 1) {						// Loop until found happy or sad
+		long new_num = 0;
+		while (num / 10 != 0) {				// While >1 digit
+			new_num += pow(num % 10, 2);	// num % 10 = least sig. digit
+			num = num / 10;					// Chop off last digit
+		}
+		new_num += pow(num, 2);
+		num = new_num;
+		
+		long i;								// Check if already hit this #
+		for (i=0; i<arr->size; i++)
+		{
+			if (num == arr->data[i])
+			{
+				return 0; // is sad
+			}
+		}
+		append(arr, num);
+	}
+	return 1; // is happy
+	arr_free(arr);
+	free(arr);
+	
+	return 1;
+}
+
+int get_next_sieve_prime(long start) {
+	long m = start + 1;
+	while (m <= floor(sqrt(MAX_PRIME)))
+	{
+	  	if (primes[m-1] == 1) { return m; }		// m is prime, return as next sieve prime
+		m++;
+	}
+	return -1;
+}
+
+static void* mark_multiples(void* arg) {
+	long j = 2;
+	thread_struct* thread_info = (thread_struct*) arg;
+	long m = thread_info->sieve_prime;
+	
+	while (m*j <= MAX_PRIME)
+	{
+		primes[m*j-1] = 0;
+		j++;
+	}
+	
[…]

### 2015-10-01 os-class/A4/pyrunner.py

@@ -1,0 +1,57 @@
+#!/usr/bin/env python
+
+import subprocess
+
+subprocess.call(["echo", "procs 32 procs 100000 items"])
+subprocess.call(["time", "./mprocs", "32", "100000"])
+subprocess.call(["echo", "procs 16 procs 100000 items"])
+subprocess.call(["time", "./mprocs", "16", "100000"])
+subprocess.call(["echo", "procs 8 procs 100000 items"])
+subprocess.call(["time", "./mprocs", "8", "100000"])
+subprocess.call(["echo", "procs 4 procs 100000 items"])
+subprocess.call(["time", "./mprocs", "4", "100000"])
+subprocess.call(["echo", "procs 2 procs 100000 items"])
+subprocess.call(["time", "./mprocs", "2", "100000"])
+subprocess.call(["echo", "procs 1 procs 100000 items"])
+subprocess.call(["time", "./mprocs", "1", "100000"])
+
+
+subprocess.call(["echo", "procs 32 procs 1000000 items"])
+subprocess.call(["time", "./mprocs", "32", "1000000"])
+subprocess.call(["echo", "procs 16 procs 1000000 items"])
+subprocess.call(["time", "./mprocs", "16", "1000000"])
+subprocess.call(["echo", "procs 8 procs 1000000 items"])
+subprocess.call(["time", "./mprocs", "8", "1000000"])
+subprocess.call(["echo", "procs 4 procs 1000000 items"])
+subprocess.call(["time", "./mprocs", "4", "1000000"])
+subprocess.call(["echo", "procs 2 procs 1000000 items"])
+subprocess.call(["time", "./mprocs", "2", "1000000"])
+subprocess.call(["echo", "procs 1 procs 1000000 items"])
+subprocess.call(["time", "./mprocs", "1", "1000000"])
+
+subprocess.call(["echo", "procs 32 procs 10000000 items"])
+subprocess.call(["time", "./mprocs", "32", "10000000"])
+subprocess.call(["echo", "procs 16 procs 10000000 items"])
+subprocess.call(["time", "./mprocs", "16", "10000000"])
+subprocess.call(["echo", "procs 8 procs 10000000 items"])
+subprocess.call(["time", "./mprocs", "8", "10000000"])
+subprocess.call(["echo", "procs 4 procs 10000000 items"])
+subprocess.call(["time", "./mprocs", "4", "10000000"])
+subprocess.call(["echo", "procs 2 procs 10000000 items"])
[…]

### 2015-10-01 os-class/A5/compute.c

@@ -1,0 +1,152 @@
+#include <pthread.h>
+#include <signal.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <sys/types.h>
+#include <sys/un.h>
+#include <sys/socket.h>
+#include <time.h>
+#include <unistd.h>
+
+
+// make global variables static so they can't interfere with other files
+static const char* MANAGE_SOCKNAME = "/tmp/CS344_manage_sock";
+
+long long compute_rate() {
+  int* buf = malloc(40000);
+  long long i = 0;
+  int j, c, sum;  // i is the candidate perfect int. j is the possible
+		  // divisor. sum is the sum of all divisors. c makes
+		  // sure we do 1000 perfect calculations before
+		  // checking time.
+  int curtime = time(NULL);
+  while (time(NULL) < curtime + 10) {
+    for (c=0; c<1000; c++) {
+      sum = 0;
+      for (j=1; j<i; j++) {
+	if (i % j == 0)
+	  { sum += j; }
+      }
+      if (sum == i)
+	{ buf[i%40000] = 1; }
+      i++;
+    }
+  }
+  free(buf);
+  return (i*i/2) / 10;       // Return # of operations done per second
+  
+}
+
+void find_perfect_nums(long long start, long long end, char* is_perfect) {
+  long long i,j, sum;
+  for (i=start; i<=end; i++)
+  {
+    sum = 0;
+    for (j=1; j<i; j++)
+    {
+      if (i % j == 0)
+	{ sum += j; }
+    }
+    if (sum == i)
+      { is_perfect[i-start] = '1'; }
+  }
+}
+
+void* listen_for_termination(void* args) {
+  int sfd = socket(AF_UNIX, SOCK_STREAM, 0);
+  int READ_BUF_MAX = 1000;
+  char read_buf[READ_BUF_MAX];
+  struct sockaddr_un manage_sock_struct;
+
+  manage_sock_struct.sun_family = AF_UNIX;
+  strcpy(manage_sock_struct.sun_path, MANAGE_SOCKNAME);
+  
+  int s = connect(sfd, (const struct sockaddr*) &manage_sock_struct, sizeof(struct sockaddr_un));
+  if (s == -1)
+    { perror("creating term manage socket"); exit(EXIT_FAILURE); }
+
+  char* str = "{\"sender\":\"term\"}";
+  
[…]

### 2015-10-01 os-class/A5/manage.py

@@ -1,0 +1,159 @@
+#! /usr/bin/env python3
+
+import atexit
+import math
+import re
+import socket
+import sys
+import subprocess
+import threading
+import time
+import os
+import signal
+
+SERVER_SOCKNAME = "/tmp/CS344_manage_sock"
+MAX_READ_SIZE = 8192
+
+e = threading.Event()
+
+processes = []
+current_num = 1
+perf_nums = []
+
+def json_decode(msg):
+    mydict = {}
+    rx = re.compile(r'"([a-zA-Z]+)":"([a-zA-Z0-9]+)"')
+    match = rx.findall(str(msg))
+
+    for l in match:
+        mydict[l[0]] = l[1]
+
+    rx2 = re.compile(r'"([a-zA-Z]+)":({.+})')
+    match = rx2.findall(str(msg))
+
+    for l in match:
+        mydict[l[0]] = json_decode(l[1])
+        
+    return mydict
+
+def handle_report_connection(conn, json_dict):
+    global perf_nums
+    
+    outstr = "{\"perfects\":\""
+    for p in perf_nums:
+        outstr = outstr + str(p) + ","
+
+    outstr = outstr + "\",\"search-range\":\"1-" + str(current_num-1) + "\",\"processes\":\""
+    for proc in processes:
+        outstr = outstr + str(proc["pid"]) + ","
+
+    outstr = outstr + "\"}"
+        
+    conn.send(str.encode(outstr))
+
+    if (json_dict["kill"] == "yes"):
+        handle_exit()
+    
+def handle_compute_connection(conn, json_dict):
+
+    global current_num
+    global processes
+    global perf_nums
+    
+    proc_dict = {}
+    proc_dict["pid"] = json_dict["pid"]
+    proc_dict["flops"] = json_dict["flops"]
+    
+    r = int(json_dict['flops'])  # rate
+    start = current_num   # perfect number lower bound
+    end = int(math.sqrt(2*r*15+start*start))
+    current_num = end + 1
+    
+    proc_dict["start"] = start
+    proc_dict["end"] = end
+    processes.append(proc_dict)
+    
+    response_json = str.encode("{\"job_range\":\"" + str(start) + "-" + str(end) + "\"}")
+    
+    conn.send(response_json)
+    json_data = conn.recv( (end - start) + 13)
+    parsed_data = json_decode(json_data)
+
+    try:
+        for i in range(0, len(parsed_data["data"])):
[…]

### 2015-10-01 os-class/A5/report.py

@@ -1,0 +1,51 @@
+#! /usr/bin/env python3
+import socket
+import getopt
+import sys
+import re
+import binascii
+
+SERVER_SOCKNAME = "/tmp/CS344_manage_sock"
+MAX_READ_SIZE = 8192
+
+def json_decode(msg):
+    mydict = {}
+    rx = re.compile(r'"([a-zA-Z\-]+)":"([a-zA-Z0-9,\-]*)"')
+    match = rx.findall(msg)
+
+    for l in match:
+        mydict[l[0]] = l[1]
+        
+    return mydict
+
+if __name__ == "__main__":
+
+    killmode = False
+    
+    if (sys.argv[1:] == ['-k']):
+        killmode = True
+    
+    sock = socket.socket(socket.AF_UNIX, socket.SOCK_STREAM)
+    sock.connect(SERVER_SOCKNAME)
+
+    msg = '{"sender":"report",'
+    if (killmode):
+        msg = msg + '"kill":"yes"}'
+    else:
+        msg = msg + '"kill":"no"}'
+        
+    sock.send(str.encode(msg))
+    report = sock.recv(MAX_READ_SIZE)
+    report = json_decode(report.decode("utf-8"))
+    
+    sock.close()
+
+    if (str(report["processes"]) == ""):
+        print("processes running: None")
+    else:
+        print("processes running: " + str(report["processes"])[:-1])
+    print("perfect numbers found: " + str(report["perfects"])[:-1])
+    if (str(report["search-range"]) == "1-0"):
+        print ("Search range: Nothing searched yet")
+    else:
+        print("numbers searched: " + str(report["search-range"]))

### 2015-10-02 extra/latex/template.tex
Add new stuff to school fs.

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{marvosym} % Required for symbols in the colored box
\usepackage{ifsym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Classical Mechanics Assignment One}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\end{document}

### 2015-10-05 hablog/hablog.py
Add hablog stuff.

@@ -1,0 +1,52 @@
+#! /usr/bin/env python
+import sys
+import time
+
+def read_loop(read_inc):
+	print "Ready, set, read!"
+	time.sleep(60*read_inc)
+	print "\a\a\a\a\a\a"
+	pages = float(raw_input("How many pages did you read? "))
+	comments = raw_input("Any special comments? ")
+	return (pages, comments)
+
+def read():
+	read_inc = int(raw_input("How long should the reading increment be in minutes? "))
+	fd = open("log.txt", 'a')
+	read_mins = 0
+	rerun = 'y'
+	
+	raw_input("Hit enter to start!")
+	
+	while (rerun == 'y'):
+		(pages, comments) = read_loop(read_inc)
+		ts = time.time()
+		fd.write("{}\t{}\t{}\t{}\n".format(*[ts, read_inc, pages, comments]))
+		read_mins += read_inc
+		rerun = raw_input("Again (y/n)? ")
+	
+	print "Read a total of {} minutes!".format(*[read_mins])
+
+
+def rate():
+	print "rating"
+
+def main(argv):
+	if (len(argv) > 1):
+		print "Too many arguments. Correct usage: ./hablog.py [read/rate]"
+		exit(1)
+		
+	if argv[0] == "read":
+		read()
+		exit(0)
+		
+	elif argv[0] == "rate":
+		rate()
+		exit(0)
+		
+	else:
+		print "Unrecognized argument. Correct usage: ./hablog.py [read/rate]"
+		exit(1)
+
+if __name__ == "__main__":
+	main(sys.argv[1:])

### 2015-10-05 classical/latex/a1.tex

--- 2015-10-02 extra/latex/template.tex
@@
 
+\section{Problem One}
+\textbf{What do you expect from this course?} \\ \\
+       {From this course I want to learn the physics techniques for modeling more complicated systems than what was taught in the introductory physics sequence.
+I want to get a better understanding for how physical modeling works and get a better intution for when to use particular techniques to model physical systems.
+I'd also like to learn how to use advanced techniques, like the Lagrangian, to characterizing physical systems.}
+
+\vspace{1 cm}
+
+\section{Problem Two}
+\textbf{What do you expect from the instructor?} \\ \\
+       {I expect the professor to answer questions during lecture and be available for office hours for help and clarification on homework assignments.}
+
+\vspace{1 cm}
+
+\section{Problem Three}
+\textbf{What do you expect from your classmates?} \\ \\
+       {I expect my classmates to not be disruptive during lecture and work to create an environment conducive to learning.
+I expect them not to ask for homework solutions, but to ask for help on homework when needed.}
+
+\vspace{1 cm}
+
+\section{Problem Four}
+\textbf{How prepared do you feel that you are for this class?} \\ \\
+       {I feel adequately prepared.
+I have taken the Paradigms series in the past year, in addition to the two linear algebra courses and calculus series recommended for this course.
+I am not worried about the material, and know that if anything challenging comes up, I can always do research, ask my classmates or the professor for help.}
+
 \end{document}

### 2015-10-09 classical/latex/a2.tex
Add new stuff.

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{amsmath}
\usepackage{marvosym} % Required for symbols in the colored box
\usepackage{ifsym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Classical Mechanics Assignment One}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem One}
\textbf{Consider  a  ball  that  moves  vertically  under  the  influences  of  both  gravity  and  air resistance.  For  the  purposes  of  this  problem,  take  vertically  upward  as  the  positive 
direction. (For instance, the gravitational force on the ball would be expressed as -mg in this case.) For each equation of motion below, determine whether that equation applies 
to (i) a situation in which the ball moves upward, (ii) a situation in which the ball moves 
downward, (iii) either of these, or (iv) neither of these. Explain your reasoning for each 
case. c is a positive constant. Justify each answer, no need  to add a  final paragraph of 
comment for the whole problem.} \\ \\

\begin{equation}
  m\frac{dv}{dt} = -mg + cv\\
\end{equation}
This motion equation corresponds to case (iii), either upwards or downwards motion. $\frac{dv}{dt}$ is positive when $cv > mg$, and negative when $cv < mg$. This means that in the case that v is positive and $cv$ has a larger magnitude than the downward acceleration $mg$, the acceleration will be positive and make velocity more positive, leading to upwards motion. When $cv$ is less than the magnitude of $mg$, gravity is either overcoming an upwards force or reinforcing a downward force and accelerating the ball downward.\\

\begin{equation}
  m\frac{dv}{dt} = -mg - cv\\
\end{equation}
This situation is the opposite of the above one: neither upwards nor downwards motion will ensue. This is because acceleration now moves opposite velocity. In the case where $v > \frac{-mg}{c}$, the ball will be accelerated downwards until $v < \frac{-mg}{c}$. Once this happens, the ball will be accelerated upwards until $v > \frac{-mg}{c}$ once again, and this will continue in a cycle.\\

\begin{equation}
  m\frac{dv}{dt} = -mg + cv^2\\
\end{equation}
This situation corresponds to (iii), either upwards or downwards motion. If $\sqrt{\frac{mg}{c}} < v_0$, then the acceleration will be upwards and an already upward-moving ball will move that way faster. If $\sqrt{\frac{mg}{c}} > v_0$, then the ball will be caught in a loop of its downward velocity being pushed upward or downward, but it will never be positive enough to have both v' and v positive at once.\\

\begin{equation}
  m\frac{dv}{dt} = -mg - cv^2\\
\end{equation}

\vspace{1 cm}

\section{Problem Two}
\textbf{Express  the  terminal  velocity  of  objects  falling  vertically  in  air  and  subject  to  the following (fictitious) drag  forces (no need to add a  final paragraph of comment  for the 
whole problem):} \\ \\

a.)
\begin{align}
  \vec{F}_{drag} &= -\alpha\log(v)\vec{v}\\
  \frac{dv}{dt} &= -mg + \vec{F}_{drag}\\
  \frac{dv}{dt} &= -mg - \alpha\log(v)v\\  
  \frac{dv}{dt} &= -mg - \alpha\log(v_{term})v_{term} = 0\\
  -mg &= \alpha\log(v_{term})v_{term}\\
\end{align}

b.)
\begin{align}
  \vec{F}_{drag} &= -\alpha v^3\vec{v}\\
  \frac{dv}{dt} &= -mg + \vec{F}_{drag}\\
  \frac{dv}{dt} &= -mg -\alpha v^3\vec{v}\\
  \frac{dv}{dt} &= -mg -\alpha v^4\\
  \frac{dv}{dt} &= -mg -\alpha v_{term}^4 = 0\\
  -\alpha v_{term}^4 &= -mg\\
  v_{term}^4 &= \frac{-mg}{\alpha}\\
  v_{term} &= \sqrt[4]{\frac{-mg}{\alpha}}\\
 \end{align}

\vspace{1 cm}

\section{Problem Three}
\textbf{Consider a softball of diameter 10cm and mass 200g and one of Millikan’s oil droplets 
with  diameter  1µm  and  mass  1pg.  They  are  both subject  to  air  drag  of  equation
$\vec{F}_{drag} = −\beta D\vec{v} −\gamma D^2v\vec{v}$, with $\beta = 1.6 \times 10^{−4}$ and $\gamma = 0.25$ (SI units).} \\ \\

a. For  the  ball  and  the  droplet,  find  the  velocity  range  for  which  the  linear  (viscous) 
term of the drag force dominates and the one for which the quadratic (momentum) 
term dominates.\\

\begin{align}
  F_{vis} &> F_{mom}\\
  -\beta D\vec{v_{vis}} &> -\gamma D^2v_{vis}\vec{v}_{vis}\\
  \beta D\vec{v_{vis}} &< \gamma D^2v_{vis}\vec{v}_{vis}\\
  \beta D\vec{v_{vis}} &< \gamma D^2v_{vis}\vec{v}_{vis}\\
  \beta v_{vis} &< \gamma Dv^2_{vis}\\
  \frac{\beta}{\gamma D} &< v_{vis}\\
  \frac{\beta}{\gamma D} &< v_{vis}\\
\end{align}
Therefore, $\frac{\beta}{\gamma D}  > v_{mom}$.

Ball:
\begin{align}
  \frac{\beta}{\gamma D} < v_{vis}\\
  \frac{1.6*10^{-4}}{0.25*0.1} < v_{vis}\\
  0.0064 \frac{m}{s} < v_{vis}
\end{align}

Oil droplet:
\begin{align}
  \frac{\beta}{\gamma D} < v_{vis}\\
  \frac{1.6*10^{-4}}{0.25*10^{-6}} < v_{vis}\\
  640 \frac{m}{s} < v_{vis}
\end{align}

b. Calculate the terminal velocity of the ball and of the droplets when they are falling 
vertically.\\

\begin{align}
  \frac{dv}{dt} &= -mg −\beta D\vec{v} −\gamma D^2v\vec{v}\\
  0 &= -mg −\beta Dv_{term} −\gamma D^2v^2_{term}\\
  0 &= mg + \beta Dv_{term} + \gamma D^2v^2_{term}\\
  0 &= \gamma D^2v^2_{term} + \beta Dv_{term} + mg\\
  v_{term} &= \frac{-\beta D \pm \sqrt{\beta^2 D^2 - 4\gamma D^2mg}}{2\gamma D^2}\\
\end{align}

Ball:
\begin{align}
  v_{term} &= \frac{-\beta D \pm \sqrt{\beta^2 D^2 - 4\gamma D^2mg}}{2\gamma D^2}\\
  v_{term} &= \frac{-1.6*10^{-4}*0.1 \pm \sqrt{1.6^2*10^{-8}*0.1^2 - 4*0.25*0.1^2*0.2*9.8}}{2*0.25*0.1^2}\\  
\end{align}

c. Reflect on your result. What term of the drag  force is dominant  for the  free-falling 
softball? What term is dominant for the oil droplet?\\

\vspace{1 cm}

\section{Problem Four}
\textbf{How prepared do you feel that you are for this class?} \\ \\

\end{document}

### 2015-10-09 classical/latex/a2.tex
Adding new stuffs.

@@
 
+\usepackage{graphicx}
 \usepackage{amsmath}
@@
 \textbf{Consider  a  ball  that  moves  vertically  under  the  influences  of  both  gravity  and  air resistance.
-For  the  purposes  of  this  problem,  take  vertically  upward  as  the  positive
-direction. (For instance, the gravitational force on the ball would be expressed as -mg in this case.) For each equation of motion below, determine whether that equation applies 
+For  the  purposes  of  this  problem,  take  vertically  upward  as  the  positive direction.
+(For instance, the gravitational force on the ball would be expressed as -mg in this case.) For each equation of motion below, determine whether that equation applies
 to (i) a situation in which the ball moves upward, (ii) a situation in which the ball moves 
@@
 \end{equation}
-This situation is the opposite of the above one: neither upwards nor downwards motion will ensue.
-This is because acceleration now moves opposite velocity.
-In the case where $v > \frac{-mg}{c}$, the ball will be accelerated downwards until $v < \frac{-mg}{c}$.
-Once this happens, the ball will be accelerated upwards until $v > \frac{-mg}{c}$ once again, and this will continue in a cycle.\\
+This situation corresponds to (ii), downwards motion.
+When velocity is positive, it feels downward acceleration.
+When velocity is negative, assuming air drag is small compared to gravity, the object will travel downwards until it reaches terminal velocity.
 
@@
 \end{equation}
-This situation corresponds to (iii), either upwards or downwards motion.
-If $\sqrt{\frac{mg}{c}} < v_0$, then the acceleration will be upwards and an already upward-moving ball will move that way faster.
-If $\sqrt{\frac{mg}{c}} > v_0$, then the ball will be caught in a loop of its downward velocity being pushed upward or downward, but it will never be positive enough to have both v' and v positive at once.\\
+This situation corresponds to either upward or downward motion.
+If $\sqrt(\frac{mg}{c}) < v$, then the object will constantly feel upward motion.
+If $\sqrt(\frac{mg}{c}) > v$, then the object will at first feel deceleration until v is so negative that $\sqrt(\frac{mg}{c}) < v$, at which point it will begin to accelerate until this is no longer true.
+It will oscillate around the negative v where $\sqrt(\frac{mg}{c}) > v$ indefinitely.
 
@@
 \end{equation}
+This situation corresponds to downward motion. $-cv^2$ will always be negative, and so the acceleration will always be downwards.\\
 
@@
 \section{Problem Two}
-\textbf{Express  the  terminal  velocity  of  objects  falling  vertically  in  air  and  subject  to  the following (fictitious) drag  forces (no need to add a  final paragraph of comment  for the 
-whole problem):} \\ \\
+\textbf{Express  the  terminal  velocity  of  objects  falling  vertically  in  air  and  subject to the following (fictitious) drag  forces (no need to add a  final paragraph of comment  for the whole problem):} \\ \\
 
-a.)
+a.)\\
+Here we treat $\hat{v}$ as a unit vector which always points in the direction of velocity, in this case downwards.\\
+
 \begin{align}
-  \vec{F}_{drag} &= -\alpha\log(v)\vec{v}\\
+  \vec{F}_{drag} &= -\alpha\log(v)\hat{v}\\
   \frac{dv}{dt} &= -mg + \vec{F}_{drag}\\
-  \frac{dv}{dt} &= -mg - \alpha\log(v)v\\  
-  \frac{dv}{dt} &= -mg - \alpha\log(v_{term})v_{term} = 0\\
-  -mg &= \alpha\log(v_{term})v_{term}\\
+  \frac{dv}{dt} &= -mg + \alpha\log(v)\\  
+  \frac{dv}{dt} &= -mg + \alpha\log(v_{term}) = 0\\
+  \frac{mg}{\alpha} &= \log(v_{term})\\
+  \frac{mg}{\alpha} &= \log(v_{term})\\
+  v_{term} &= e^{\frac{mg}{\alpha}}\\
 \end{align}
@@
   \vec{F}_{drag} &= -\alpha v^3\vec{v}\\
+  \vec{F}_{drag} &= -\alpha v^4\hat{v}\\  
   \frac{dv}{dt} &= -mg + \vec{F}_{drag}\\
-  \frac{dv}{dt} &= -mg -\alpha v^3\vec{v}\\
-  \frac{dv}{dt} &= -mg -\alpha v^4\\
-  \frac{dv}{dt} &= -mg -\alpha v_{term}^4 = 0\\
-  -\alpha v_{term}^4 &= -mg\\
-  v_{term}^4 &= \frac{-mg}{\alpha}\\
-  v_{term} &= \sqrt[4]{\frac{-mg}{\alpha}}\\
+  \frac{dv}{dt} &= -mg -\alpha v^4\hat{v}\\
+  \frac{dv}{dt} &= -mg +\alpha v_{term}^4 = 0\\
+  \alpha v_{term}^4 &= mg\\
+  v_{term}^4 &= \frac{mg}{\alpha}\\
+  v_{term} &= \sqrt[4]{\frac{mg}{\alpha}}\\
  \end{align}
@@
 
-b. Calculate the terminal velocity of the ball and of the droplets when they are falling 
-vertically.\\
+b. Calculate the terminal velocity of the ball and of the droplets when they are falling vertically.\\
+
+We let $\vec{v}$ point downward, and expect positive velocities.\\
 
@@
   \frac{dv}{dt} &= -mg −\beta D\vec{v} −\gamma D^2v\vec{v}\\
-  0 &= -mg −\beta Dv_{term} −\gamma D^2v^2_{term}\\
-  0 &= mg + \beta Dv_{term} + \gamma D^2v^2_{term}\\
-  0 &= \gamma D^2v^2_{term} + \beta Dv_{term} + mg\\
-  v_{term} &= \frac{-\beta D \pm \sqrt{\beta^2 D^2 - 4\gamma D^2mg}}{2\gamma D^2}\\
+  0 &= -mg + \beta Dv_{term} +\gamma D^2v^2_{term}\\
+  0 &= \gamma D^2v^2_{term} + \beta Dv_{term} - mg\\
+  v_{term} &= \frac{-\beta D \pm \sqrt{\beta^2 D^2 + 4\gamma D^2mg}}{2\gamma D^2}\\
 \end{align}
@@
   v_{term} &= \frac{-\beta D \pm \sqrt{\beta^2 D^2 - 4\gamma D^2mg}}{2\gamma D^2}\\
-  v_{term} &= \frac{-1.6*10^{-4}*0.1 \pm \sqrt{1.6^2*10^{-8}*0.1^2 - 4*0.25*0.1^2*0.2*9.8}}{2*0.25*0.1^2}\\  
+  v_{term} &= \frac{-1.6*10^{-4}*0.1 + \sqrt{1.6^2*10^{-8}*0.1^2 + 4*0.25*0.1^2*0.2*9.8}}{2*0.25*0.1^2}\\
+  &= 27.99 \frac{m}{s}\\
 \end{align}
 
-c. Reflect on your result. What term of the drag  force is dominant  for the  free-falling 
-softball? What term is dominant for the oil droplet?\\
+Oil droplet:
+\begin{align}
+  v_{term} &= \frac{-\beta D \pm \sqrt{\beta^2 D^2 - 4\gamma D^2mg}}{2\gamma D^2}\\
+  v_{term} &= \frac{-1.6*10^{-4}*10^{-6} + \sqrt{1.6^2*10^{-8}*10^{-12} + 4*0.25*10^{-12}*10^{-15}*9.8}}{2*0.25*10^{-12}}\\
+  = 6.1*10^{-5} \frac{m}{s}\\
+\end{align}
+
+c. Reflect on your result. What term of the drag force is dominant  for the  free-falling softball? What term is dominant for the oil droplet?\\
+
+The ball's terminal velocity is about 28 meters per second, which means quadratic drag is the dominating force.
+The oil droplet's terminal velocity is well below 640 meters per second, so it is under almost entirely linear drag.
+In fact, oil droplets never feel momentum/quadratic drag.\\
+
+This problem was cool because it showed that some objects never really feel quadratic drag.
+Oil drops, which have very low radius, feel very little quadratic drag.
+Basically all objects on the size scale of an oil-drop or smaller feel mostly viscous drag from the air they move.
+This problem also reinforced how to calculate terminal velocity.
+I also had to spend some time thinking about how to deal with the coordinate system and velocity vectors and what direction they point in.
+I decided to make them point down, and this made the sign of the drag forces upwards.
+\\
 
@@
 \section{Problem Four}
-\textbf{How prepared do you feel that you are for this class?} \\ \\
+\textbf{When a cyclist comes to a stop, without using the brakes, she is subject to two forces: the quadratic force of air resistance  $F_{air} = −cv^22$ and a constant friction force  $F_f$.} \\ \\
+
+a. Write and solve the equation of motion for the velocity as a function of time while the cyclist is coasting to a halt.\\
+
+\begin{align}
+  m\frac{dv}{dt} = F_f - cv^2\\
+  m\frac{dv}{F_f - cv^2} = dt\\
+  \int_{v0}^v m\frac{dv}{F_f - cv^2} = \int dt\\
+  \int_{v0}^v \frac{mdv}{F_f - cv^2} = t\\
+  \frac{m*\tanh^{-1}(\frac{\sqrt{c}v}{\sqrt{F}})}{\sqrt{cF}} - \frac{m*\tanh^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})}{\sqrt{cF}} = t\\
+  \tanh^{-1}(\frac{\sqrt{c}v}{\sqrt{F}}) - \tanh^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}}) = \frac{t\sqrt{cF}}{m}\\
+  \tanh^{-1}(\frac{\sqrt{c}v}{\sqrt{F}}) = \frac{t\sqrt{cF}}{m} + \tanh^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})\\  
+  \tanh(\frac{t\sqrt{cF}}{m} + \tanh^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})) = \frac{\sqrt{c}v}{\sqrt{F}}\\
+  \sqrt{\frac{F}{c}}\tanh(\frac{t\sqrt{cF}}{m} + \tanh^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})) = v(t)\\    
+\end{align}
+
+b. Plot  a  graph  of  the  result  for  a  cyclist  plus  bike  mass  of  80kg,  an  initial  velocity 20m/s, a quadratic drag coefficient  c = 0.2 (in SI units) and a  friction  force of 3 N.\\
+
+\begin{figure}[h!]
+  \caption{Plot of cyclist velocity over time.}
+  \centering
+    \includegraphics[width=0.7\textwidth]{images/cyclist_velocity_m1.png}
+\end{figure}
+
+Answer the following graphically:\\
+- How long will it take the cyclist to slow down to 15 m/s?\\
+
+By examining the graph, we can see that it takes roughly ten seconds for the cyclist to slow from 20 m/s to 15 m/s.\\
+
+- How long to slow from 10m/s to 5m/s?\\
+
+Also by inspection of teh graph, we can see that it takes roughly 25 seconds to slow from 10m/s to 5m/s.\\
+
+- How long does it take for a complete stop (from the initial velocity to 0m/s)?\\
+It takes roughly 140 seconds to reach a complete stop, starting at 20m/s.\\
+
+This was a harder problem, but also intersting.
+We had to use the basic technique of physics, finding a differential equation to model a system and solving for velocity (or usually position), then use a computer to plot this solution, since it is difficult to interpret otherwise.
+I learned that from very simple initial forces (a constant force and a velocity squared force, both opposing motion), the equation for velocity can become complex (this one involved a hyperbolic tangent).
+
+\section{Problem Five}
+
+\textbf{A particle is fired vertically upward in a constant gravitational field with an initial speed $v_0$.
+Show  that  if  there  is  a  retarding  force  proportional  to  the  square  of  the instantaneous speed, the speed of the particle when it returns to the initial position is:}
+
+$$ v = \frac{v_ov_{term}}{\sqrt{v_0^2+v^2_{term}}}  $$
+
+\textbf{where $v_{term}$ is the terminal velocity.}
+
+We need to solve the following differential equation:\\
+
+When going up:\\
+\begin{align}
+  m\frac{dv}{dt} = -mg - cv^2\\
+  m\frac{dv}{dy}\frac{dy}{dt} = -mg - cv^2\\
+  m\frac{dv}{dy}v = -mg - cv^2\\
+  m\frac{dv}{-mg - cv^2}v = dy\\
+  \int_{v_0}^0 m\frac{dv}{-mg - cv^2}v = \int dy\\
+  \int_{v_0}^0 m\frac{mv}{-mg - cv^2}dv = y_{max}\\
+  \int_{v_0^2}^0 \frac{mv}{-mg - cu}\frac{du}{2v} = y_{max}\\
+  \frac{1}{2} \int_{v_0^2}^0 \frac{m}{-mg - cu} du = y_{max}\\
+  \frac{m \log(cv_0^2 + gm)}{2c} - \frac{m \log(gm)}{2c} = y_{max}\\
+  \frac{m}{2c} \log(\frac{cv_0^2 + gm}{gm}) = y_{max}\\  
+\end{align}
+
+When falling back down:\\
+\begin{align}
+  m\frac{dv}{dt} = -mg + cv^2\\
+  m\frac{dv}{dy}\frac{dy}{dt} = -mg + cv^2\\
+  m\frac{dv}{dy}v = -mg + cv^2\\
+  m\frac{dv}{-mg + cv^2}v = dy\\
+  \int_{v_0}^0 m\frac{dv}{-mg + cv^2}v = \int dy\\
+  \int_{v_0}^0 m\frac{mv}{-mg + cv^2}dv = -y_{max}\\
+  \int_{v_0^2}^0 \frac{mv}{-mg + cu}\frac{du}{2v} = -y_{max}\\
+  \frac{1}{2} \int_{v_0^2}^0 \frac{m}{-mg + cu} du = -y_{max}\\
+  \frac{m \log(-gm)}{2c} - \frac{m \log(cv_0^2 - gm)}{2c} = -y_{max}\\
+  \frac{m}{2c} \log(\frac{gm}{gm - cv_0^2}) = -y_{max}\\
+\end{align}
+
+We now have two expressions for $y_{max}$. We set them equal and see what we get:
+
+\begin{align}
+  \frac{m}{2c} \log(\frac{gm}{cv_0^2 + gm}) = -\frac{m}{2c} \log(\frac{gm}{gm - cv_0^2})\\
+  \log(\frac{gm}{cv_0^2 + gm}) = -\log(\frac{gm}{gm - cv_0^2})\\
+  \log(\frac{gm}{cv_0^2 + gm}) = \log(\frac{gm - cv_0^2}{gm})\\
+  \frac{gm}{cv_0^2 + gm} = \frac{gm - cv_0^2}{gm}\\
+  \frac{g^2m^2}{cv_0^2 + gm} = gm - cv_0^2\\    
+\end{align}
+
+We now find $v_{term}$:
+
+\begin{align}
+  m\frac{dv}{dt} = -mg + cv^2 = 0\\
+  -mg + cv_{term}^2 = 0\\
+  cv_{term}^2 = mg\\
+  v_{term} = \sqrt{\frac{mg}{c}}\\
+\end{align}
+
+This was a really interesting problem in that it had us do a fairly lengthy derivation of something.
+We had to apply lots of different knowledge - tricky integration, terminal velocity calculation and logarithm manipulation.
+It was really intersting to see how all the little parts fit together.
+This was probably the hardest part of the problem set in that it was really hard to get the correct answer, but it was a fun problem and a good test of math skills.\\
 

### 2015-10-09 classical/latex/a2.tex
Push errthang.

@@
 
-\title{Classical Mechanics Assignment One}
+\title{Classical Mechanics Assignment Two}
 \author{Elliott Capek}
@@
 This situation corresponds to either upward or downward motion.
-If $\sqrt(\frac{mg}{c}) < v$, then the object will constantly feel upward motion.
-If $\sqrt(\frac{mg}{c}) > v$, then the object will at first feel deceleration until v is so negative that $\sqrt(\frac{mg}{c}) < v$, at which point it will begin to accelerate until this is no longer true.
-It will oscillate around the negative v where $\sqrt(\frac{mg}{c}) > v$ indefinitely.
+If $\sqrt{\frac{mg}{c}} < v$, then the object will constantly feel upward motion.
+If $\sqrt{\frac{mg}{c}} > v$, then the object will at first feel deceleration until v is so negative that $\sqrt{\frac{mg}{c}} < v$, at which point it will begin to accelerate until this is no longer true.
+It will oscillate around the negative v where $\sqrt{\frac{mg}{c}} > v$ indefinitely.
 
@@
 
+This solution makes sense. It has positive sign, which makes sense, since positive sign corresponds to downward motion. Alpha is inversely proportional to terminal velocity, and mass is proportional.
+
 b.)
@@
 
+This solution makes sense.
+Terminal velocity is proportional to mass and inversely proportional to alpha, as expected.
+For this particular system, where force is highly velocity-dependent, the terminal velocity grows a lot slower than in the above case, where force is less strongly velocity-dependent.
+
 \vspace{1 cm}
@@
 
-a. For  the  ball  and  the  droplet,  find  the  velocity  range  for  which  the  linear  (viscous) 
-term of the drag force dominates and the one for which the quadratic (momentum) 
-term dominates.\\
+a. For  the  ball  and  the  droplet,  find  the  velocity  range  for  which  the  linear  (viscous) term of the drag force dominates and the one for which the quadratic (momentum) term dominates.\\
 
@@
 \begin{align}
-  m\frac{dv}{dt} = F_f - cv^2\\
-  m\frac{dv}{F_f - cv^2} = dt\\
-  \int_{v0}^v m\frac{dv}{F_f - cv^2} = \int dt\\
-  \int_{v0}^v \frac{mdv}{F_f - cv^2} = t\\
-  \frac{m*\tanh^{-1}(\frac{\sqrt{c}v}{\sqrt{F}})}{\sqrt{cF}} - \frac{m*\tanh^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})}{\sqrt{cF}} = t\\
-  \tanh^{-1}(\frac{\sqrt{c}v}{\sqrt{F}}) - \tanh^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}}) = \frac{t\sqrt{cF}}{m}\\
-  \tanh^{-1}(\frac{\sqrt{c}v}{\sqrt{F}}) = \frac{t\sqrt{cF}}{m} + \tanh^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})\\  
-  \tanh(\frac{t\sqrt{cF}}{m} + \tanh^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})) = \frac{\sqrt{c}v}{\sqrt{F}}\\
-  \sqrt{\frac{F}{c}}\tanh(\frac{t\sqrt{cF}}{m} + \tanh^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})) = v(t)\\    
+  m\frac{dv}{dt} = -F_f - cv^2\\
+  m\frac{dv}{-F_f - cv^2} = dt\\
+  \int_{v0}^v m\frac{dv}{-F_f - cv^2} = \int dt\\
+  \int_{v0}^v \frac{mdv}{F_f + cv^2} = -t\\
+  \frac{m*\tan^{-1}(\frac{\sqrt{c}v}{\sqrt{F}})}{\sqrt{cF}} - \frac{m*\tan^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})}{\sqrt{cF}} = -t\\
+  \tan^{-1}(\frac{\sqrt{c}v}{\sqrt{F}}) - \tan^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}}) = \frac{-t\sqrt{cF}}{m}\\
+  \tan^{-1}(\frac{\sqrt{c}v}{\sqrt{F}}) = \frac{-t\sqrt{cF}}{m} + \tan^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})\\  
+  \tan(\frac{-t\sqrt{cF}}{m} + \tan^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})) = \frac{\sqrt{c}v}{\sqrt{F}}\\
+  \sqrt{\frac{F}{c}}\tan(\frac{-t\sqrt{cF}}{m} + \tan^{-1}(\frac{\sqrt{c}v_0}{\sqrt{F}})) = v(t)\\    
 \end{align}
@@
 We had to use the basic technique of physics, finding a differential equation to model a system and solving for velocity (or usually position), then use a computer to plot this solution, since it is difficult to interpret otherwise.
-I learned that from very simple initial forces (a constant force and a velocity squared force, both opposing motion), the equation for velocity can become complex (this one involved a hyperbolic tangent).
+I learned that from very simple initial forces (a constant force and a velocity squared force, both opposing motion), the equation for velocity can become complex (this one involved a tangent function).
 
@@
   m\frac{dv}{-mg + cv^2}v = dy\\
-  \int_{v_0}^0 m\frac{dv}{-mg + cv^2}v = \int dy\\
-  \int_{v_0}^0 m\frac{mv}{-mg + cv^2}dv = -y_{max}\\
-  \int_{v_0^2}^0 \frac{mv}{-mg + cu}\frac{du}{2v} = -y_{max}\\
-  \frac{1}{2} \int_{v_0^2}^0 \frac{m}{-mg + cu} du = -y_{max}\\
-  \frac{m \log(-gm)}{2c} - \frac{m \log(cv_0^2 - gm)}{2c} = -y_{max}\\
-  \frac{m}{2c} \log(\frac{gm}{gm - cv_0^2}) = -y_{max}\\
-\end{align}
-
-We now have two expressions for $y_{max}$. We set them equal and see what we get:
-
-\begin{align}
-  \frac{m}{2c} \log(\frac{gm}{cv_0^2 + gm}) = -\frac{m}{2c} \log(\frac{gm}{gm - cv_0^2})\\
-  \log(\frac{gm}{cv_0^2 + gm}) = -\log(\frac{gm}{gm - cv_0^2})\\
-  \log(\frac{gm}{cv_0^2 + gm}) = \log(\frac{gm - cv_0^2}{gm})\\
-  \frac{gm}{cv_0^2 + gm} = \frac{gm - cv_0^2}{gm}\\
-  \frac{g^2m^2}{cv_0^2 + gm} = gm - cv_0^2\\    
+  \int_{0}^{vf} m\frac{dv}{-mg + cv^2}v = \int dy\\
+  \int_{0}^{vf} m\frac{mv}{-mg + cv^2}dv = -y_{max}\\
+  \int_{0}^{vf} \frac{mv}{-mg + cu}\frac{du}{2v} = -y_{max}\\
+  \frac{1}{2} \int_{0}^{vf} \frac{m}{-mg + cu} du = -y_{max}\\
+  \frac{m \log(cv_f^2 - gm)}{2c} - \frac{m \log(-gm)}{2c} = -y_{max}\\
+  \frac{m}{2c} (\log(cv_f^2 - gm) - \log(-gm)) = -y_{max}\\
+  \frac{m}{2c} \log(\frac{cv_f^2 - gm}{-gm}) = -y_{max}\\
+  \frac{m}{2c} \log(\frac{gm - cv_f^2}{gm}) = y_{max}\\
 \end{align}
@@
 
+We now have two expressions for $y_{max}$. We set them equal and see what we get:
+
+\begin{align}
+  \frac{m}{2c} \log(\frac{gm}{cv_0^2 + gm}) = \frac{m}{2c} \log(\frac{gm - cv_f^2}{gm})\\
+  \frac{gm}{cv_0^2 + gm} = \frac{gm - cv_f^2}{gm}\\
+  \frac{(gm)^2}{cv_0^2 + gm} = gm - cv_f^2\\
+  cv_f^2 = gm - \frac{(gm)^2}{cv_0^2 + gm}\\
+  v_f^2 = \frac{gm}{c} - \frac{(gm)^2}{c^2v_0^2 + cgm}\\
+  v_f^2 = v_{term}^2 - \frac{v_{term}^4}{v_0^2 + v_{term}^2}\\
+  v_f^2 = \frac{v_{term}^2(v_0^2 + v_{term}^2) - v_{term}^4}{v_0^2 + v_{term}^2}\\
+  v_f^2 = \frac{v_{term}^2v_0^2}{v_0^2 + v_{term}^2}\\      
+\end{align}
+
 This was a really interesting problem in that it had us do a fairly lengthy derivation of something.

### 2015-10-15 classical/latex/a3.tex
Add new sterf

--- 2015-10-09 classical/latex/a2.tex
@@
 
-\title{Classical Mechanics Assignment One}
+\title{Classical Mechanics Assignment X}
 \author{Elliott Capek}
@@
 \section{Problem One}
-\textbf{Consider  a  ball  that  moves  vertically  under  the  influences  of  both  gravity  and  air resistance.
-For  the  purposes  of  this  problem,  take  vertically  upward  as  the  positive
-direction. (For instance, the gravitational force on the ball would be expressed as -mg in this case.) For each equation of motion below, determine whether that equation applies 
-to (i) a situation in which the ball moves upward, (ii) a situation in which the ball moves 
-downward, (iii) either of these, or (iv) neither of these. Explain your reasoning for each 
-case. c is a positive constant. Justify each answer, no need  to add a  final paragraph of 
-comment for the whole problem.} \\ \\
+\textbf{Consider an object that is thrown vertically up with initial speed $v_0$ in a linear medium.} \\ \\
 
-\begin{equation}
-  m\frac{dv}{dt} = -mg + cv\\
-\end{equation}
-This motion equation corresponds to case (iii), either upwards or downwards motion.
-$\frac{dv}{dt}$ is positive when $cv > mg$, and negative when $cv < mg$.
-This means that in the case that v is positive and $cv$ has a larger magnitude than the downward acceleration $mg$, the acceleration will be positive and make velocity more positive, leading to upwards motion.
-When $cv$ is less than the magnitude of $mg$, gravity is either overcoming an upwards force or reinforcing a downward force and accelerating the ball downward.\\
+\textbf{a. Measuring y upward from the point of release, write expressions for the object velocity and position as a function of time.}
 
-\begin{equation}
-  m\frac{dv}{dt} = -mg - cv\\
-\end{equation}
-This situation is the opposite of the above one: neither upwards nor downwards motion will ensue.
-This is because acceleration now moves opposite velocity.
-In the case where $v > \frac{-mg}{c}$, the ball will be accelerated downwards until $v < \frac{-mg}{c}$.
-Once this happens, the ball will be accelerated upwards until $v > \frac{-mg}{c}$ once again, and this will continue in a cycle.\\
+\begin{align*}
+  a(t) = -g - bv\\ ro ro diff eq!
+  v(t) = v_0 -gt\\
+  y(t) = v_0t - \frac{1}{2}gt^2\\
+\end{align*}
 
-\begin{equation}
-  m\frac{dv}{dt} = -mg + cv^2\\
-\end{equation}
-This situation corresponds to (iii), either upwards or downwards motion.
-If $\sqrt{\frac{mg}{c}} < v_0$, then the acceleration will be upwards and an already upward-moving ball will move that way faster.
-If $\sqrt{\frac{mg}{c}} > v_0$, then the ball will be caught in a loop of its downward velocity being pushed upward or downward, but it will never be positive enough to have both v' and v positive at once.\\
+\textbf{b. Find the time at which the object reaches its highest point and its position at that time.}
 
-\begin{equation}
-  m\frac{dv}{dt} = -mg - cv^2\\
-\end{equation}
+\textbf{c. Show that the limiting case for very small drag coefficient coincides with the solution for the frictionless motion.}
 
@@
 \section{Problem Two}
-\textbf{Express  the  terminal  velocity  of  objects  falling  vertically  in  air  and  subject  to  the following (fictitious) drag  forces (no need to add a  final paragraph of comment  for the 
-whole problem):} \\ \\
+\textbf{A point mass of mass m slides down an inclined plane under the influence of gravity.
+If the motion is resisted by a force F=kmv2, show that the time required to move a distance d after starting from rest is:}
 
-a.)
-\begin{align}
-  \vec{F}_{drag} &= -\alpha\log(v)\vec{v}\\
-  \frac{dv}{dt} &= -mg + \vec{F}_{drag}\\
-  \frac{dv}{dt} &= -mg - \alpha\log(v)v\\  
-  \frac{dv}{dt} &= -mg - \alpha\log(v_{term})v_{term} = 0\\
-  -mg &= \alpha\log(v_{term})v_{term}\\
-\end{align}
+\begin{equation*}
+  t = \frac{\cosh^{-1}(e^{kd}}{\sqrt{kg\sin(\Theta)}}\\
+\end{equation*}
 
-b.)
-\begin{align}
-  \vec{F}_{drag} &= -\alpha v^3\vec{v}\\
-  \frac{dv}{dt} &= -mg + \vec{F}_{drag}\\
-  \frac{dv}{dt} &= -mg -\alpha v^3\vec{v}\\
-  \frac{dv}{dt} &= -mg -\alpha v^4\\
-  \frac{dv}{dt} &= -mg -\alpha v_{term}^4 = 0\\
-  -\alpha v_{term}^4 &= -mg\\
-  v_{term}^4 &= \frac{-mg}{\alpha}\\
-  v_{term} &= \sqrt[4]{\frac{-mg}{\alpha}}\\
- \end{align}
+\textbf{Where $\Theta$ is the angle of inclination of the plane.} \\ \\
 
@@
 \section{Problem Three}
-\textbf{Consider a softball of diameter 10cm and mass 200g and one of Millikan’s oil droplets 
-with  diameter  1µm  and  mass  1pg.  They  are  both subject  to  air  drag  of  equation
-$\vec{F}_{drag} = −\beta D\vec{v} −\gamma D^2v\vec{v}$, with $\beta = 1.6 \times 10^{−4}$ and $\gamma = 0.25$ (SI units).} \\ \\
-
-a. For  the  ball  and  the  droplet,  find  the  velocity  range  for  which  the  linear  (viscous) 
-term of the drag force dominates and the one for which the quadratic (momentum) 
-term dominates.\\
-
-\begin{align}
-  F_{vis} &> F_{mom}\\
-  -\beta D\vec{v_{vis}} &> -\gamma D^2v_{vis}\vec{v}_{vis}\\
-  \beta D\vec{v_{vis}} &< \gamma D^2v_{vis}\vec{v}_{vis}\\
-  \beta D\vec{v_{vis}} &< \gamma D^2v_{vis}\vec{v}_{vis}\\
-  \beta v_{vis} &< \gamma Dv^2_{vis}\\
-  \frac{\beta}{\gamma D} &< v_{vis}\\
-  \frac{\beta}{\gamma D} &< v_{vis}\\
-\end{align}
-Therefore, $\frac{\beta}{\gamma D}  > v_{mom}$.
-
-Ball:
-\begin{align}
-  \frac{\beta}{\gamma D} < v_{vis}\\
-  \frac{1.6*10^{-4}}{0.25*0.1} < v_{vis}\\
-  0.0064 \frac{m}{s} < v_{vis}
-\end{align}
-
-Oil droplet:
-\begin{align}
-  \frac{\beta}{\gamma D} < v_{vis}\\
-  \frac{1.6*10^{-4}}{0.25*10^{-6}} < v_{vis}\\
-  640 \frac{m}{s} < v_{vis}
-\end{align}
-
-b. Calculate the terminal velocity of the ball and of the droplets when they are falling 
-vertically.\\
-
-\begin{align}
-  \frac{dv}{dt} &= -mg −\beta D\vec{v} −\gamma D^2v\vec{v}\\
-  0 &= -mg −\beta Dv_{term} −\gamma D^2v^2_{term}\\
-  0 &= mg + \beta Dv_{term} + \gamma D^2v^2_{term}\\
-  0 &= \gamma D^2v^2_{term} + \beta Dv_{term} + mg\\
-  v_{term} &= \frac{-\beta D \pm \sqrt{\beta^2 D^2 - 4\gamma D^2mg}}{2\gamma D^2}\\
-\end{align}
-
-Ball:
-\begin{align}
-  v_{term} &= \frac{-\beta D \pm \sqrt{\beta^2 D^2 - 4\gamma D^2mg}}{2\gamma D^2}\\
-  v_{term} &= \frac{-1.6*10^{-4}*0.1 \pm \sqrt{1.6^2*10^{-8}*0.1^2 - 4*0.25*0.1^2*0.2*9.8}}{2*0.25*0.1^2}\\  
-\end{align}
-
-c. Reflect on your result. What term of the drag  force is dominant  for the  free-falling 
-softball? What term is dominant for the oil droplet?\\
-
+\textbf{A badly designed rocket has an initial mass of 7x104 kg and on firing it burns its fuel at a rate of 250 kg/s.
+The exhaust speed is 2500 m/s.
+If the rocket has a vertical ascent from rest on the Earth, how long after the rocket engines fire will the rocket lift off?} \\ \\
 \vspace{1 cm}
@@
 \section{Problem Four}
-\textbf{How prepared do you feel that you are for this class?} \\ \\
+\textbf{A projectile of mass M explodes while in flight at speed v into three fragments.
+The first fragment, with mass m1=M/2 travels in the original direction of the projectile; the second, with mass m2=M/6 travels in the opposite direction, and the third comes to rest.
+The energy released in the explosion is equal to five times the kinetic energy of the projectile just before exploding.
+What are the velocities of the first two fragments?} \\ \\
+\vspace{1 cm}
+
+\section{Problem Five}
+\textbf{Find the center of mass of a uniform hemispherical shell of inner and outer radii a and b and mass M. Comment on the limiting cases. (5 bonus points)} \\ \\
+\vspace{1 cm}
 

### 2015-10-15 classical/latex/template.tex
Add new sterf

--- 2015-10-15 classical/latex/a3.tex
@@
 \section{Problem One}
-\textbf{Consider an object that is thrown vertically up with initial speed $v_0$ in a linear medium.} \\ \\
-
-\textbf{a. Measuring y upward from the point of release, write expressions for the object velocity and position as a function of time.}
-
-\begin{align*}
-  a(t) = -g - bv\\ ro ro diff eq!
-  v(t) = v_0 -gt\\
-  y(t) = v_0t - \frac{1}{2}gt^2\\
-\end{align*}
-
-\textbf{b. Find the time at which the object reaches its highest point and its position at that time.}
-
-\textbf{c. Show that the limiting case for very small drag coefficient coincides with the solution for the frictionless motion.}
-
+\textbf{} \\ \\
 \vspace{1 cm}
@@
 \section{Problem Two}
-\textbf{A point mass of mass m slides down an inclined plane under the influence of gravity.
-If the motion is resisted by a force F=kmv2, show that the time required to move a distance d after starting from rest is:}
-
-\begin{equation*}
-  t = \frac{\cosh^{-1}(e^{kd}}{\sqrt{kg\sin(\Theta)}}\\
-\end{equation*}
-
-\textbf{Where $\Theta$ is the angle of inclination of the plane.} \\ \\
-
+\textbf{} \\ \\
 \vspace{1 cm}
@@
 \section{Problem Three}
-\textbf{A badly designed rocket has an initial mass of 7x104 kg and on firing it burns its fuel at a rate of 250 kg/s.
-The exhaust speed is 2500 m/s.
-If the rocket has a vertical ascent from rest on the Earth, how long after the rocket engines fire will the rocket lift off?} \\ \\
+\textbf{} \\ \\
 \vspace{1 cm}
@@
 \section{Problem Four}
-\textbf{A projectile of mass M explodes while in flight at speed v into three fragments.
-The first fragment, with mass m1=M/2 travels in the original direction of the projectile; the second, with mass m2=M/6 travels in the opposite direction, and the third comes to rest.
-The energy released in the explosion is equal to five times the kinetic energy of the projectile just before exploding.
-What are the velocities of the first two fragments?} \\ \\
+\textbf{} \\ \\
 \vspace{1 cm}
@@
 \section{Problem Five}
-\textbf{Find the center of mass of a uniform hemispherical shell of inner and outer radii a and b and mass M. Comment on the limiting cases. (5 bonus points)} \\ \\
+\textbf{} \\ \\
 \vspace{1 cm}

### 2015-10-15 comp_model/latex/a1.tex
Add new sterf

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage{amsmath}
\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Computational Modeling HW 1}
\author{Elliott Capek}

\begin{document}

\maketitle{}

\section{1.4: Trapezoidal method coefficients}
\begin{equation}
  \sum_{j=0}^r \alpha_jV_{n+j} = h \sum_{j=0}^r \beta_j f(t_{n+j}, V_{n+j})\\
\end{equation}

The trapezoidal method is this:\\
\begin{equation}
  V_{n+1} = V_n + h\Big( \frac{1}{2}f(V_n, t_n) + \frac{1}{2}f(V_{n+1}, t_{n+1}) \Big)
\end{equation}

By examining these two equations, we can see that:

\begin{align}
  \alpha_0 = -1 &\hspace{2cm} \beta_0 = \frac{1}{2}\\
  \alpha_1 = 0 &\hspace{2cm} \beta_1 = 1\\
  \alpha_2 = 0 &\hspace{2cm} \beta_2 = 1\\
  ...  &\hspace{2cm} ...\\
  \alpha_{r-1} = 0 &\hspace{2cm} \beta_{r-1} = 1\\
  \alpha_r = 1 &\hspace{2cm} \beta_1 = \frac{1}{2}\\    
\end{align}

\begin{equation}
  V_{n+1} = V_n + h\Big( \frac{1}{2}f(V_n, t_n) + \frac{1}{2}f(V_{n+1}, t_{n+1}) \Big)
\end{equation}

If we let our function f take the following form: $f(x, t) = \alpha x$ where $x(0) = x_0$, we can see that:

\begin{align}
  V_{n+1} &= V_n + \frac{h}{2}\Big( f(V_n, t_n) + f(V_{n+1}, t_{n+1}) \Big)\\
  V_{n+1} &= V_n + \frac{h}{2}\Big( \alpha V_n + \alpha V_{n+1} \Big)\\
  V_{n+1} &= V_n + \frac{\alpha hV_n}{2} + \frac{\alpha hV_{n+1}}{2}\\
  V_{n+1} \Big(1 - \frac{\alpha h}{2}\Big) &= V_n \Big(1 + \frac{\alpha h}{2}\Big) \\
  V_{n+1} &= V_n \frac{1 + \frac{\alpha h}{2}}{1 - \frac{\alpha h}{2}} \\
  V_{n+1} &= V_n \frac{2 + \alpha h}{2 - \alpha h} \\
  V_{n+1} &= V_0 \Big(\frac{2 + \alpha h}{2 - \alpha h}\Big)^{n+1} \\    
\end{align}

From this we can see that as long as the following condition is satisfied, we will have a converging solution:

\begin{align}
  |2 + \alpha h| \leq |2 - \alpha h|\\
\end{align}
\section{1.9: Order of Accuracy for Trapezoidal and Euler Methods}
We want to analytically represent the error terms for the Euler and Trapezoidal methods and show how they depend on our timestep, $h$.\\

First we solve for the behavior of our function using a Taylor Series. Let $x(t)$ be our function and $v(t) = \frac{dx(t)}{dt}$. Let $N = \frac{t}{h}$, where t is the current time and h the timestep. Finally, let $x(t_0) = 0$.\\

\begin{align}
  x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} v(t)dt\\
  v(t) &= v(t_i) + (t-t_i)v'(t_i) + (t-t_i)^2v''(t_i) + ...\\
  \int_{t_i}^{t_{i+1}} v(t)dt &\approx \int_{t_i}^{t_{i+1}} v(t_i) + (t-t_i)v'(t_i) + \frac{1}{2}(t-t_i)^2v''(t_i)  dt\\
\end{align}

We now do a u-substitution, where $S = \frac{t - t_i}{t_{i+1} - t_i}$ and $h = t_{i+1} - t_i$:

\begin{align}
  \int_{t_i}^{t_{i+1}} v(t)dt &\approx h\int_i^1 v(t_i) + hSv'(t_i) + \frac{1}{2}h^2S^2v''(t_i) dS\\
  \int_{t_i}^{t_{i+1}} v(t)dt &\approx h \big( v(t_i)S + \frac{hS^2}{2}v'(t_i) + \frac{h^2S^3}{6}v''(t_i) \big) |_0^1 \\
  \int_{t_i}^{t_{i+1}} v(t)dt &\approx h \big( v(t_i) + \frac{h}{2}v'(t_i) + \frac{h^2}{6}v''(t_i) \big) \\
  \int_{t_i}^{t_{i+1}} v(t)dt &\approx hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{6}v''(t_i) \\
\end{align}

This gives us a fairly exact value for $\int_{t_i}^{t_{i+1}} v(t)dt$. We now need to compare this with the approximations that our two methods make. For Euler's method, $\int_{t_i}^{t_{i+1}} v(t)dt = hf(t_i)$. For the Trapezoidal method, $\int_{t_i}^{t_{i+1}} v(t)dt = \frac{h}{2}\Big( f(t_{i+1}) + f(t_{i}) \Big)$. We will first plug in the Euler's method approximation to the above expression to find its error term.

\begin{align}
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i)\\
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{6}v''(t_i)\\
  hv(t_i) &= hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{6}v''(t_i)\\
  -\frac{h}{3}v''(t_i) &= v'(t_i)\\  
\end{align}

We now substitute this back into our original result:

\begin{align}
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{6}v''(t_i) \\
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + \frac{h^2}{2}(-\frac{h}{3}v''(t_i)) + \frac{h^3}{6}v''(t_i) \\
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + -\frac{h^3}{4}(v''(t_i)) + \frac{h^3}{6}v''(t_i) \\
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + -\frac{h^3}{12}(v''(t_i))\\
\end{align}

As we can see, the error for the forward Euler's method is $-\frac{h^3}{12}(v''(t_i))$ per timestep. We now plug this result into our summed equation to find the total error for all timesteps:

\begin{align}
  x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} v(t)dt\\
  x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} hv(t_i) + -\frac{h^3}{12}(v''(t_i))\\
  x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} hv(t_i) + -\frac{h^3}{12}(v''(t_i))\\  
\end{align}

We only care about the error term, so we will focus on that. We substitute $N$ with $\frac{t - t_0}{h}$ and replace $v''(t_i)$ with a term that represents the average value of $v''(t_i)$ over the whole sum: $v''_{ave}$.

\begin{align}
  E &= \sum_{i=0}^{i=N-1} -\frac{h^3}{12}(v''(t_i))\\
  E &= -N\frac{h^3}{12}v''_{ave}\\
  E &= -\frac{t - t_0}{h} \frac{h^3}{12}v''_{ave}\\
  E &= -(t - t_0)\frac{h^2}{12}v''_{ave}\\    
\end{align}

We do the same thing for the Trapezoidal method. We first find the error term:

\begin{align}
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i)\\
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{4}v''(t_i)\\
  \frac{h}{2}\Big(v(t_{i+1}) + v(t_i)\Big) &= hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{4}v''(t_i)\\
  \frac{h}{2}\Big(-v(t_{i+1}) + v(t_i)\Big) - \frac{h^3}{4}v''(t_i) &= \frac{h^2}{2}v'(t_i)\\
  \frac{1}{h}\Big(-v(t_{i+1}) + v(t_i)\Big) - \frac{h}{2}v''(t_i) &= v'(t_i)\\
\end{align}

We then substitute this back into our original result:

\begin{align}
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{6}v''(t_i) \\
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + \frac{h^2}{2}\Big(\frac{1}{h}\Big(v(t_{i+1}) - v(t_i)\Big) - \frac{h}{4}v''(t_i) \Big) + \frac{h^3}{6}v''(t_i) \\
  \int_{t_i}^{t_{i+1}} v(t)dt &= \frac{h}{2}\big(v(t_{i+1}) + v(t_i)\big) - \frac{h^3}{4}v''(t_i) + \frac{h^3}{6}v''(t_i) \\
  \int_{t_i}^{t_{i+1}} v(t)dt &= \frac{h}{2}\big(v(t_{i+1}) + v(t_i)\big) - \frac{h^3}{12}v''(t_i)\\    
\end{align}

From this we can see that the error term for the trapezoidal rule is $-\frac{h^3}{12}v''(t_i)$ per timestep. We then sum over all timesteps as above:

\begin{align}
    x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} v(t)dt\\
  x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} \frac{h}{2}\big(v(t_{i+1}) + v(t_i)\big) - \frac{h^3}{12}v''(t_i)\\
  x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} \frac{h}{2}\big(v(t_{i+1}) + v(t_i)\big) + \sum_{i=0}^{i=N-1} -\frac{h^3}{12}v''(t_i)\\
\end{align}

We finally solve for the error:

\begin{align}
  E &= \sum_{i=0}^{i=N-1} -\frac{h^3}{12}(v''(t_i))\\
  E &= -N\frac{h^3}{12}v''_{ave}\\
  E &= -\frac{t - t_0}{h} \frac{h^3}{12}v''_{ave}\\
  E &= -(t - t_0)\frac{h^2}{12}v''_{ave}\\
\end{align}

From this we can see that the Trapezoidal method's error is proportional to $h^2$.

\end{document}

### 2015-10-15 classical/latex/a3.tex
stuff

@@
 
-\title{Classical Mechanics Assignment X}
+\title{Classical Mechanics Assignment 3}
 \author{Elliott Capek}
@@
 \section{Problem One}
-\textbf{Consider an object that is thrown vertically up with initial speed $v_0$ in a linear medium.} \\ \\
+\textbf{Consider an object that is thrown vertically up with initial speed $v_0$ in a linear medium.} \\
 
@@
 \begin{align*}
-  a(t) = -g - bv\\ ro ro diff eq!
-  v(t) = v_0 -gt\\
-  y(t) = v_0t - \frac{1}{2}gt^2\\
+  m\ddot{y}(t) &= -mg - b\dot{y}(t)\\
+  m\ddot{y}(t) &= -b\big(\frac{mg}{b} + \dot{y}(t))\\
+  m\ddot{y}(t) &= -b\big(\dot{y}_{term} + \dot{y}(t))
+  \hspace{1in}\ddot{y} = 0 \rightarrow \dot{y}_{term} = \frac{mg}{b}\\
+  m\ddot{y}(t) &= -b\phi(t)
+  \hspace{1.6in} \phi(t) = \dot{y}_{term} + \dot{y}(t)\\
+  m\ddot{\phi(t)} &= -b\phi(t)
+  \hspace{1.6in} \dot{\phi(t)} = \ddot{y}(t)\\
+  \phi(t) &= Ae^{-\frac{bt}{m}}\\
+  \dot{y}(t) &= Ae^{-\frac{b}{m}t} - \dot{y}_{term}
+  \hspace{1.6in} \dot{y}(0) = \dot{y}_0 = A - \dot{y}_{term}\\
+  \dot{y}(t) &= (\dot{y}_0 + \dot{y}_{term})e^{-\frac{b}{m}t} - \dot{y}_{term}\\
+  \dot{y}(t) &= \dot{y}_0e^{-\frac{b}{m}t} + \dot{y}_{term}(e^{-\frac{b}{m}t} - 1)\\
+  y(t) &= -\frac{m}{b}e^{-\frac{b}{m}t}\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}t\\
 \end{align*}
@@
 
-\textbf{c. Show that the limiting case for very small drag coefficient coincides with the solution for the frictionless motion.}
+\begin{align*}
+  \dot{y}(t) &= 0 = \dot{y}_0e^{-\frac{b}{m}t} + \dot{y}_{term}(e^{-\frac{b}{m}t} - 1)\\
+  \dot{y}_{term}(1 - e^{-\frac{b}{m}t}) &= \dot{y}_0e^{-\frac{b}{m}t}\\
+  \dot{y}_{term} - \dot{y}_{term}e^{-\frac{b}{m}t} &= \dot{y}_0e^{-\frac{b}{m}t}\\
+  \dot{y}_{term}e^{\frac{b}{m}t} &= \dot{y}_0 + \dot{y}_{term}\\
+  e^{\frac{b}{m}t} &= \frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}\\
+  t_{max} &= \frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\\
+  y(t_{max}) &= -\frac{m}{b}e^{-\frac{b}{m}\frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}}\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}\frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\\
+    y(t_{max}) &= -\frac{m}{b\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}\frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\\
+\end{align*}
+
+\textbf{c. Show that the limiting case for very small drag coefficient coincides with the solution for the frictionless motion.}\\
+For frictionless motion:
+\begin{align*}
+  \ddot{y}(t) &= -g\\
+  \dot{y}(t) &= \dot{y}_0 - gt\\
+  y(t) &= \dot{y}_0t - \frac{1}{2}gt^2\\
+  \dot{y}(t_{max}) &= 0 \rightarrow t_{max} = \frac{\dot{y}}{g}\\
+  y_{max} &= \frac{\dot{y}_0^2}{2g}\\
+\end{align*}
+
+For motion in a linear medium when $\lim_{b\rightarrow0}$:
+
+\begin{align*}
+  \lim_{b\rightarrow0}y(t_{max}) &= \lim_{b->0} \hspace{0.2cm} -\frac{m}{b\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}\frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\\
+  \lim_{b\rightarrow0}y(t_{max}) &= \lim_{b->0} \hspace{0.2cm} -\frac{m}{b\frac{\dot{y}_0 + \frac{mg}{b}}{\frac{mg}{b}}}\Big(\dot{y}_0 + \frac{mg}{b} \Big) - \frac{mg}{b}\frac{m}{b}\log{\frac{\dot{y}_0 + \frac{mg}{b}}{\frac{mg}{b}}}\\
+  \lim_{b\rightarrow0}y(t_{max}) &= \lim_{b->0} \hspace{0.2cm} -\frac{m}{b\frac{\dot{y}_0 + \frac{mg}{b}}{\frac{mg}{b}}}\Big(\dot{y}_0 + \frac{mg}{b} \Big) - \frac{mg}{b}\frac{m}{b}\log{\frac{\dot{y}_0 + \frac{mg}{b}}{\frac{mg}{b}}}\\
+\end{align*}
 
@@
 \textbf{A point mass of mass m slides down an inclined plane under the influence of gravity.
-If the motion is resisted by a force F=kmv2, show that the time required to move a distance d after starting from rest is:}
+If the motion is resisted by a force $F=kmv^2$, show that the time required to move a distance d after starting from rest is:}
 
 \begin{equation*}
-  t = \frac{\cosh^{-1}(e^{kd}}{\sqrt{kg\sin(\Theta)}}\\
+  t = \frac{\cosh^{-1}(e^{kd})}{\sqrt{kg\sin(\Theta)}}\\
 \end{equation*}

### 2015-10-15 classical/latex/a3.tex
add

@@
 
+We define a one-dimensional coordinate system where the positive axis points in the direction of motion of the point.\\
+
+\begin{align*}
+  \vec{F} &= \sin(\Theta)mg - kmv^2\\
+  m\dot{v} &= \sin(\Theta)mg - kmv^2\\
+  \frac{dv}{dt} &= \sin(\Theta)g - kv^2\\
+  \frac{dv}{\sin(\Theta)g - kv^2} &= dt\\
+  \int_0^{v} \frac{dv}{\sin(\Theta)g - kv^2} &= \int_0^t dt\\
+  \int_0^{v} \frac{\csc(\Theta)}{g(1 - \frac{\csc(\Theta)k}{g}v^2)}dv &= \int_0^t dt\\
+  \frac{\csc(\Theta)}{g}\int_0^{v} \frac{1}{1 - \frac{\csc(\Theta)k}{g}v^2}dv &= \int_0^t dt\\
+  \frac{\csc(\Theta)}{g}\int_0^{v} \frac{1}{1 - \frac{\csc(\Theta)k}{g}v^2}dv &= \int_0^t dt
+  \hspace{1in} u = \mathrm{i}v\sqrt{\frac{\csc(\Theta)k}{g}}\\
+  -\frac{\csc(\Theta)}{g}\mathrm{i}\sqrt{\frac{g}{\csc(\Theta)k}}\int_0^{\mathrm{i}v\sqrt{\frac{\csc(\Theta)k}{g}}} \frac{1}{1 + u^2}du &= \int_0^t dt
+  \hspace{1in} dv = -\mathrm{i}\sqrt{\frac{g}{\csc(\Theta)k}}du\\
+  -\mathrm{i}\frac{\sqrt{\csc(\Theta)}}{\sqrt{kg}} \int_0^{\mathrm{i}v\sqrt{\frac{\csc(\Theta)k}{g}}} \frac{1}{1 + u^2}du &= \int_0^t dt\\
+  -\mathrm{i}\frac{\sqrt{\csc(\Theta)}}{\sqrt{kg}} \tan^{-1}\Bigg(\mathrm{i}v\sqrt{\frac{\csc(\Theta)k}{g}}\Bigg) &= \int_0^t dt\\
+  \frac{\sqrt{\csc(\Theta)}}{\sqrt{kg}} \tanh^{-1}\Bigg(v\sqrt{\frac{\csc(\Theta)k}{g}}\Bigg) &= t\\
+  \frac{\tanh^{-1}\Bigg(v\sqrt{\frac{\csc(\Theta)k}{g}}\Bigg)}{\sqrt{kg\sin(\Theta)}} &= t\\ 
+\end{align*}
+
+Now we try this...
+
+\begin{align*}
+  \frac{\tanh^{-1}\Bigg(v\sqrt{\frac{\csc(\Theta)k}{g}}\Bigg)}{\sqrt{kg\sin(\Theta)}} &= t\\
+  \tanh^{-1}\Bigg(v\sqrt{\frac{\csc(\Theta)k}{g}}\Bigg) &= t\sqrt{kg\sin(\Theta)}\\
+  v\sqrt{\frac{\csc(\Theta)k}{g}} &= \tanh\Big(t\sqrt{kg\sin(\Theta)}\Big)\\
+  v &= \frac{\tanh\Big(t\sqrt{kg\sin(\Theta)}\Big)}{\sqrt{\frac{\csc(\Theta)k}{g}}}\\    
+\end{align*}
+
+Now we try integrating...
+
+\begin{align*}
+  d &= \int_0^t v(t) dt\\
+  d &= \int_0^t \frac{\tanh\Big(t\sqrt{kg\sin(\Theta)}\Big)}{\sqrt{\frac{\csc(\Theta)k}{g}}} dt\\
+  d &= \frac{\log(\cosh\Big(t\sqrt{kg\sin(\Theta)}\Big))}{\sqrt{\frac{\csc(\Theta)k}{g}}}\\
+  d\sqrt{\frac{\csc(\Theta)k}{g}} &= \log(\cosh\Big(t\sqrt{kg\sin(\Theta)}\Big))\\
+  e^{d\sqrt{\frac{\csc(\Theta)k}{g}}} &= \cosh\Big(t\sqrt{kg\sin(\Theta)}\Big)\\
+  \cosh^{-1}\Bigg(e^{d\sqrt{\frac{\csc(\Theta)k}{g}}}) &= t\sqrt{kg\sin(\Theta)}\\
+  \frac{\cosh^{-1}\Big(e^{d\sqrt{\frac{\csc(\Theta)k}{g}}}\Big)}{\sqrt{kg\sin(\Theta)}} &= t\\        
+\end{align*}
+
 \vspace{1 cm}

### 2015-10-16 classical/latex/a3.tex
new stuff

@@
 \usepackage{titlesec} % Allows creating custom \section's
+
+\allowdisplaybreaks
 
@@
   \frac{\tanh^{-1}\Bigg(v\sqrt{\frac{\csc(\Theta)k}{g}}\Bigg)}{\sqrt{kg\sin(\Theta)}} &= t\\ 
-\end{align*}
-
-Now we try this...
-
-\begin{align*}
   \frac{\tanh^{-1}\Bigg(v\sqrt{\frac{\csc(\Theta)k}{g}}\Bigg)}{\sqrt{kg\sin(\Theta)}} &= t\\
@@
 
-Now we try integrating...
+Now we try integrating out the v:
 
@@
   d &= \int_0^t \frac{\tanh\Big(t\sqrt{kg\sin(\Theta)}\Big)}{\sqrt{\frac{\csc(\Theta)k}{g}}} dt\\
-  d &= \frac{\log(\cosh\Big(t\sqrt{kg\sin(\Theta)}\Big))}{\sqrt{\frac{\csc(\Theta)k}{g}}}\\
-  d\sqrt{\frac{\csc(\Theta)k}{g}} &= \log(\cosh\Big(t\sqrt{kg\sin(\Theta)}\Big))\\
-  e^{d\sqrt{\frac{\csc(\Theta)k}{g}}} &= \cosh\Big(t\sqrt{kg\sin(\Theta)}\Big)\\
-  \cosh^{-1}\Bigg(e^{d\sqrt{\frac{\csc(\Theta)k}{g}}}) &= t\sqrt{kg\sin(\Theta)}\\
-  \frac{\cosh^{-1}\Big(e^{d\sqrt{\frac{\csc(\Theta)k}{g}}}\Big)}{\sqrt{kg\sin(\Theta)}} &= t\\        
+  d &= \frac{\log(\cosh\Big(t\sqrt{kg\sin(\Theta)}\Big))}{\sqrt{kg\sin(\Theta)}\sqrt{\frac{\csc(\Theta)k}{g}}}\\
+  d &= \frac{\log(\cosh\Big(t\sqrt{kg\sin(\Theta)}\Big))}{\sqrt{\frac{kg}{\csc(\Theta)}}\sqrt{\frac{\csc(\Theta)k}{g}}}\\
+ d &= \frac{\log(\cosh\Big(t\sqrt{kg\sin(\Theta)}\Big))}{k}\\    
+  kd &= \log(\cosh\Big(t\sqrt{kg\sin(\Theta)}\Big))\\
+  e^{kd} &= \cosh\Big(t\sqrt{kg\sin(\Theta)}\Big)\\
+  \cosh^{-1}(e^{kd}) &= t\sqrt{kg\sin(\Theta)}\\
+  \frac{\cosh^{-1}(e^{kd})}{\sqrt{kg\sin(\Theta)}} &= t\\        
 \end{align*}

### 2015-10-16 classical/latex/a3.tex
daddsf

@@
 If the rocket has a vertical ascent from rest on the Earth, how long after the rocket engines fire will the rocket lift off?} \\ \\
-\vspace{1 cm}
+%\vspace{1 cm}
+Here $m$ is the mass of the rocket, $m_0$ is the initial mass of the rocket, $dm$ is the mass of the particles being ejected per timestep $dt$, $v$ is the rocket's velocity, $v_{ex}$ is the exhaust's velocity and $dv$ is the rocket's change in velocity per timestep $dt$.
+
+\begin{align*}
+  m\ddot{y} &= -mg - \dot{m}v_{ex}\\
+  m\ddot{y} &= -mg + |\dot{m}|v_{ex}\\
+  \ddot{y} &= -g + \frac{|\dot{m}|}{m}v_{ex}\\
+  \frac{dv}{dt} &= -g + \frac{|\dot{m}|}{m}v_{ex}\\
+  dv &= \Big(-g + \frac{|\dot{m}|}{m}v_{ex}\Big)dt\\
+  dv &= \Big(-g + \frac{|\dot{m}|}{m}v_{ex}\Big)\Big(-\frac{dm}{|\dot{m}|}\Big)\\
+  dv &= \Big(\frac{g}{|\dot{m}|} - \frac{v_{ex}}{m}\Big)dm\\
+  \int_0^v dv &= \int_{m_0}^m \Big(\frac{g}{|\dot{m}|} - \frac{v_{ex}}{m}\Big)dm\\
+  v &= \frac{g}{|\dot{m}|}\big(m-m_0) - v_{ex}\log(\frac{m}{m_0})\\
+  v &= \frac{-g}{|\dot{m}|}\big(m_0-m) + v_{ex}\log(\frac{m_0}{m})\\  
+\end{align*}
+
+This is an incorrect equation in that it predicts negative velocity when the gravitational force is stronger than the exhaust force.
+This is because we didn't include the normal force term in our initial enumeration of the forces acting on the rocket.
+The second thrust exceeds gravity the normal force disappears and our equation becomes correct.
+This coincides with the time the rocket lifts off:\\
+
+\begin{align*}
+  0 &= \frac{-g}{|\dot{m}|}\big(m_0-m) + v_{ex}\log(\frac{m_0}{m})\\
+  \frac{-g}{|\dot{m}|}\big(m_0-m) &= v_{ex}\log(\frac{m_0}{m})\\
+\end{align*}
+
+We plug in our values:\\
+
+\begin{align*}
+  \Big(\frac{-9.8\frac{m}{s^2}}{250\frac{kg}{s}}\Big)\big(7*10^4kg - 250\frac{kg}{s}t) &= \big(2500\frac{m}{s}\big)\log\Big(\frac{7*10^{4}kg}{250\frac{kg}{s}t}\Big)\\
+  \Big(1.568^{e-5}\frac{1}{kg}\Big)\big(7e4 kg - 250*t kg) &= \log\Big(\frac{7e4}{250*t}\Big)\\
+  1.0976 - 0.00392t &= \log\Big(\frac{280}{t}\Big)\\
+  -4.5372 - 0.00392t &= -\log(t)\\
+  t &= 231.752
+\end{align*}
 
@@
 
+Using conservation of momentum and conservation of energy, we can form two equations:\\
+
+\begin{align*}
+  MV &= \frac{1}{2}MV_1 + \frac{1}{6}MV_2\\
+  3MV^2 &= \frac{1}{4}MV_1^2 + \frac{1}{12}MV_2^2\\
+\end{align*}
+
+We have two equations and two unknowns. We can solve for $V_1$ and $V_2$:\\
+
+\begin{align*}
+  V &= \frac{1}{2}V_1 + \frac{1}{6}V_2\\
+  V^2 &= \frac{1}{4}V_1^2 + \frac{1}{36}V_2^2 + \frac{1}{6}V_1V_2\\
+  3V^2 &= \frac{1}{4}V_1^2 + \frac{1}{12}V_2^2\\
+  V^2 &= \frac{1}{12}V_1^2 + \frac{1}{36}V_2^2\\
+  \frac{1}{4}V_1^2 + \frac{1}{36}V_2^2 + \frac{1}{6}V_1V_2 &= \frac{1}{12}V_1^2 + \frac{1}{36}V_2^2\\
+  \frac{1}{6}V_1^2 + \frac{1}{6}V_1V_2 &= 0\\
+  V_1 &= -V_2\\
+\end{align*}
+
+We plug this back into our first equation...\\
+
+\begin{align*}
+  V &= \frac{1}{2}V_1 + -\frac{1}{6}V_1\\
+  V &= \frac{1}{3}V_1\\
+  V_1 &= 3V\\
+  V_2 &= -3V\\
+\end{align*}
+
 \section{Problem Five}

### 2015-10-22 classical/latex/a3.tex
ad

@@
   \dot{y}(t) &= \dot{y}_0e^{-\frac{b}{m}t} + \dot{y}_{term}(e^{-\frac{b}{m}t} - 1)\\
-  y(t) &= -\frac{m}{b}e^{-\frac{b}{m}t}\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}t\\
+  y(t) &= -\frac{m}{b^2}e^{-\frac{b}{m}t}\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}t\\
 \end{align*}
@@
   t_{max} &= \frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\\
-  y(t_{max}) &= -\frac{m}{b}e^{-\frac{b}{m}\frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}}\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}\frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\\
-    y(t_{max}) &= -\frac{m}{b\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}\frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\\
+  t_{max} &= \frac{m}{b}\log\Big(\frac{b\dot{y}_0 + mg}{mg}\Big)\\
+  t_{max} &= \frac{-m}{b}\log\Big(\frac{mg}{b\dot{y}_0 + mg}\Big)\\
+  y(t) &= -\frac{m}{b^2}e^{-\frac{b}{m}\frac{-m}{b}\log\Big(\frac{mg}{b\dot{y}_0 + mg}\Big)}\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}\frac{-m}{b}\log\Big(\frac{mg}{b\dot{y}_0 + mg}\Big)\\
+  y(t) &= -\frac{m}{b^2}\Big(\frac{mg}{b\dot{y}_0 + mg}\Big)\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}\frac{-m}{b}\log\Big(\frac{mg}{b\dot{y}_0 + mg}\Big)\\
+  y(t) &= -\frac{m}{b^2}\Big(\frac{mg}{b\dot{y}_0 + mg}\Big)\Big(b\dot{y}_0 + mg\Big) - \dot{y}_{term}\frac{-m}{b}\log\Big(\frac{mg}{b\dot{y}_0 + mg}\Big)\\
+  y(t) &= -\frac{gm^2}{b^2} - \frac{-m^2g}{b^2}\log\Big(\frac{mg}{b\dot{y}_0 + mg}\Big)\\
+  y(t) &= \frac{gm^2}{b^2}\Bigg(\log\Big(\frac{mg}{b\dot{y}_0 + mg}\Big) - 1\Bigg)\\
 \end{align*}
@@
   y(t) &= \dot{y}_0t - \frac{1}{2}gt^2\\
-  \dot{y}(t_{max}) &= 0 \rightarrow t_{max} = \frac{\dot{y}}{g}\\
+  \dot{y}(t_{max}) &= 0 \rightarrow t_{max} = \frac{\dot{y}_0}{g}\\
   y_{max} &= \frac{\dot{y}_0^2}{2g}\\
@@
 \begin{align*}
-  \lim_{b\rightarrow0}y(t_{max}) &= \lim_{b->0} \hspace{0.2cm} -\frac{m}{b\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}\frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\\
-  \lim_{b\rightarrow0}y(t_{max}) &= \lim_{b->0} \hspace{0.2cm} -\frac{m}{b\frac{\dot{y}_0 + \frac{mg}{b}}{\frac{mg}{b}}}\Big(\dot{y}_0 + \frac{mg}{b} \Big) - \frac{mg}{b}\frac{m}{b}\log{\frac{\dot{y}_0 + \frac{mg}{b}}{\frac{mg}{b}}}\\
-  \lim_{b\rightarrow0}y(t_{max}) &= \lim_{b->0} \hspace{0.2cm} -\frac{m}{b\frac{\dot{y}_0 + \frac{mg}{b}}{\frac{mg}{b}}}\Big(\dot{y}_0 + \frac{mg}{b} \Big) - \frac{mg}{b}\frac{m}{b}\log{\frac{\dot{y}_0 + \frac{mg}{b}}{\frac{mg}{b}}}\\
+  \lim_{b\rightarrow0}\Bigg[\frac{gm^2}{b^2}\Bigg(\log\Big(\frac{mg}{b\dot{y}_0 + mg}\Big) - 1\Bigg)\Bigg] \rightarrow WolframAlpha &= \dot{y}_0 - gt\\
 \end{align*}
+
+This was a cool problem.
+It was good to go through the entire derivation of motion through a linear medium using a different coordinate system than the one we used in class.
+It made sure I really understood how that kind of motion worked and how to use separation of variables to solve for position.
+I had a bit of trouble with the algebra, and found that my solution didn't agree with the frictionless solution when b goes to zero for position, for some strange reason.
+I checked with others and they got the same thing I did, so I'm not sure what the problem is.\\
 
@@
 
+This was a really cool problem.
+I had a lot of fun trying to do the integral by myself and then using Mathematica to help me out.
+I think I'm a little better at doing trigonometric integration after doing this problem.
+I also know a little more about hyperbolic trigonometric functions now.\\
+
 \vspace{1 cm}
@@
 
+I liked this problem.
+After doing it I understand the rocket equation we're using pretty well.
+I think it's interesting that the equation only works when there is no normal force on the rocket, ie when the rocket is just taking off.
+This makes sense, since it would be impossible to include the normal force in the force declaration without using a picewise function or a step function.\\
+
 \section{Problem Four}
@@
 
+This problem taught me more about how to do simple explosion calculations.
+If we make the assumption that we can just say the KE of a system increases when the explosion takes place, it is a very easy problem to figure out the velocities and trajectories of the particles using Physics 211-style momentum and energy conservation equations.
+I feel like it wouldn't be possible to apply in the real world since lots more factors would go into the calculation, since we wouldn't know what pieces would break.
+But if we did know what pieces would break, by doing structural analysis or something, maybe this wouldn't be such a bad approximation.
+We would know the chemical energy in the bomb, so the particle velocities would be easy to calculate.\\
+
 \section{Problem Five}

### 2015-10-23 classical/latex/a4.tex
Add stuff.

--- 2015-10-16 classical/latex/a3.tex
@@
 
-\title{Classical Mechanics Assignment 3}
+\title{Classical Mechanics Assignment 4}
 \author{Elliott Capek}
@@
 \section{Problem One}
-\textbf{Consider an object that is thrown vertically up with initial speed $v_0$ in a linear medium.} \\
+\textbf{Consider a rocket accelerating in free space by burning its fuel at a constant rate. Find the velocity at which the momentum of the rocket (neglecting the already spent fuel) is maximized.} \\
 
-\textbf{a. Measuring y upward from the point of release, write expressions for the object velocity and position as a function of time.}
+We want to find the velocity at which $m\dot{y}$ is maximized. We find when the derivative of this with respect to time is zero, plug in for the values we find and go from there:
 
 \begin{align*}
-  m\ddot{y}(t) &= -mg - b\dot{y}(t)\\
-  m\ddot{y}(t) &= -b\big(\frac{mg}{b} + \dot{y}(t))\\
-  m\ddot{y}(t) &= -b\big(\dot{y}_{term} + \dot{y}(t))
-  \hspace{1in}\ddot{y} = 0 \rightarrow \dot{y}_{term} = \frac{mg}{b}\\
-  m\ddot{y}(t) &= -b\phi(t)
-  \hspace{1.6in} \phi(t) = \dot{y}_{term} + \dot{y}(t)\\
-  m\ddot{\phi(t)} &= -b\phi(t)
-  \hspace{1.6in} \dot{\phi(t)} = \ddot{y}(t)\\
-  \phi(t) &= Ae^{-\frac{bt}{m}}\\
-  \dot{y}(t) &= Ae^{-\frac{b}{m}t} - \dot{y}_{term}
-  \hspace{1.6in} \dot{y}(0) = \dot{y}_0 = A - \dot{y}_{term}\\
-  \dot{y}(t) &= (\dot{y}_0 + \dot{y}_{term})e^{-\frac{b}{m}t} - \dot{y}_{term}\\
-  \dot{y}(t) &= \dot{y}_0e^{-\frac{b}{m}t} + \dot{y}_{term}(e^{-\frac{b}{m}t} - 1)\\
-  y(t) &= -\frac{m}{b}e^{-\frac{b}{m}t}\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}t\\
+  \frac{\partial}{\partial t} m\dot{y} &= \dot{m}\dot{y} + m\ddot{y} = 0\\
+  m\ddot{y} &= -mg - \dot{m}v_{ex}\\
+  \dot{m}\dot{y} &= mg + \dot{m}v_{ex}\\
+  \dot{y} &= \frac{mg + \dot{m}v_{ex}}{\dot{m}}\\
 \end{align*}
 
-\textbf{b. Find the time at which the object reaches its highest point and its position at that time.}
+\section{Problem Two}
+\textbf{Consider a rocket subject to a linear resistive force $\vec{F} = -b\vec{v}$, but no other external forces.
+Find the velocity as a function of the remaining mass for a rocket that starts from rest and ejects mass at a constant rate $k = -\dot{m}$ and velocity $v_{ex}$.} \\
 
 \begin{align*}
-  \dot{y}(t) &= 0 = \dot{y}_0e^{-\frac{b}{m}t} + \dot{y}_{term}(e^{-\frac{b}{m}t} - 1)\\
-  \dot{y}_{term}(1 - e^{-\frac{b}{m}t}) &= \dot{y}_0e^{-\frac{b}{m}t}\\
-  \dot{y}_{term} - \dot{y}_{term}e^{-\frac{b}{m}t} &= \dot{y}_0e^{-\frac{b}{m}t}\\
-  \dot{y}_{term}e^{\frac{b}{m}t} &= \dot{y}_0 + \dot{y}_{term}\\
-  e^{\frac{b}{m}t} &= \frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}\\
-  t_{max} &= \frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\\
-  y(t_{max}) &= -\frac{m}{b}e^{-\frac{b}{m}\frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}}\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}\frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\\
-    y(t_{max}) &= -\frac{m}{b\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}\frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\\
+  m\ddot{y} &= -\frac{b}{m}\dot{y} - \dot{m}v_{ex}\\
 \end{align*}
 
-\textbf{c. Show that the limiting case for very small drag coefficient coincides with the solution for the frictionless motion.}\\
-For frictionless motion:
+\section{Problem Three}
+\textbf{Consider the surface generated by revolving a line connecting two points ($x_1$,$y_1$) and ($x_2$,$y_2$) about an axis coplanar with the two points.
+Find the equation of the line connecting the points such that the surface area generated by the revolution is minimum.}
+
 \begin{align*}
-  \ddot{y}(t) &= -g\\
-  \dot{y}(t) &= \dot{y}_0 - gt\\
-  y(t) &= \dot{y}_0t - \frac{1}{2}gt^2\\
-  \dot{y}(t_{max}) &= 0 \rightarrow t_{max} = \frac{\dot{y}}{g}\\
-  y_{max} &= \frac{\dot{y}_0^2}{2g}\\
+  SA &= \int_{y_1}^{y_2} dA = \int_{y_1}^{y_2} 2\pi\sqrt{x^2+y^2}\sqrt{x'^2+1}dy\\
+  f(x, x', y) &= 2\pi\sqrt{x^2+y^2}\sqrt{x'^2+1}\\
 \end{align*}
 
-For motion in a linear medium when $\lim_{b\rightarrow0}$:
+\begin{align*}
+  \frac{\partial f}{\partial x} &- \frac{\partial}{\partial y} \frac{\partial f}{\partial x'} = 0\\
+  \frac{\partial f}{\partial x} &= \frac{2\pi x\sqrt{x'^2+1}}{\sqrt{x^2+y^2}}\\
+  \frac{\partial}{\partial y} \frac{\partial f}{\partial x'} &= 0\\
+\end{align*}
+
+So...
+\begin{align*}
+  \frac{2\pi x\sqrt{x'^2+1}}{\sqrt{x^2+y^2}} = 0\\
+\end{align*}
+
+\section{Problem Four}
+\textbf{Consider a medium in which the refractive index $n$ is inversely proportional to $r_2$.
+That is, $n=a/r2$, where $r$ is the distance from the origin.
+Use Fermat’s principle to find the path of light traveling between two points that lie in the xy plane.
+Hint: use 2-dimensional polar coordinates.
+Show that the path found is part of a circle that intersects the origin of the reference frame.}
+
+We use polar coordinates to describe our two points. Our differential is $ds = \sqrt{r^2+r'^2}d\theta$.
 
 \begin{align*}
-  \lim_{b\rightarrow0}y(t_{max}) &= \lim_{b->0} \hspace{0.2cm} -\frac{m}{b\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}\frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\\
-  \lim_{b\rightarrow0}y(t_{max}) &= \lim_{b->0} \hspace{0.2cm} -\frac{m}{b\frac{\dot{y}_0 + \frac{mg}{b}}{\frac{mg}{b}}}\Big(\dot{y}_0 + \frac{mg}{b} \Big) - \frac{mg}{b}\frac{m}{b}\log{\frac{\dot{y}_0 + \frac{mg}{b}}{\frac{mg}{b}}}\\
-  \lim_{b\rightarrow0}y(t_{max}) &= \lim_{b->0} \hspace{0.2cm} -\frac{m}{b\frac{\dot{y}_0 + \frac{mg}{b}}{\frac{mg}{b}}}\Big(\dot{y}_0 + \frac{mg}{b} \Big) - \frac{mg}{b}\frac{m}{b}\log{\frac{\dot{y}_0 + \frac{mg}{b}}{\frac{mg}{b}}}\\
+  T &= \int_{\theta_1}^{\theta_2} \frac{1}{v}ds = \int_{\theta_1}^{\theta_2} \frac{a}{cr^2}\sqrt{r^2+r'^2}d\theta\\
+  f(r,r',\theta) &= \frac{a}{cr^2}\sqrt{r^2+r'^2}d\theta\\
 \end{align*}
 
-\vspace{1 cm}
+\begin{align*}
+  \frac{\partial f}{\partial r} - \frac{\partial}{\partial \theta} \frac{\partial f}{\partial r'} &= 0\\
+  \frac{\partial}{\partial \theta} \frac{\partial f}{\partial r'} &= 0\\
+\end{align*}
 
-\section{Problem Two}
-\textbf{A point mass of mass m slides down an inclined plane under the influence of gravity.
-If the motion is resisted by a force $F=kmv^2$, show that the time required to move a distance d after starting from rest is:}
-
-\begin{equation*}
-  t = \frac{\cosh^{-1}(e^{kd})}{\sqrt{kg\sin(\Theta)}}\\
-\end{equation*}
-
-\textbf{Where $\Theta$ is the angle of inclination of the plane.} \\ \\
-
-We define a one-dimensional coordinate system where the positive axis points in the direction of motion of the point.\\
+And so:
 
 \begin{align*}
-  \vec{F} &= \sin(\Theta)mg - kmv^2\\
-  m\dot{v} &= \sin(\Theta)mg - kmv^2\\
-  \frac{dv}{dt} &= \sin(\Theta)g - kv^2\\
-  \frac{dv}{\sin(\Theta)g - kv^2} &= dt\\
-  \int_0^{v} \frac{dv}{\sin(\Theta)g - kv^2} &= \int_0^t dt\\
-  \int_0^{v} \frac{\csc(\Theta)}{g(1 - \frac{\csc(\Theta)k}{g}v^2)}dv &= \int_0^t dt\\
-  \frac{\csc(\Theta)}{g}\int_0^{v} \frac{1}{1 - \frac{\csc(\Theta)k}{g}v^2}dv &= \int_0^t dt\\
-  \frac{\csc(\Theta)}{g}\int_0^{v} \frac{1}{1 - \frac{\csc(\Theta)k}{g}v^2}dv &= \int_0^t dt
-  \hspace{1in} u = \mathrm{i}v\sqrt{\frac{\csc(\Theta)k}{g}}\\
-  -\frac{\csc(\Theta)}{g}\mathrm{i}\sqrt{\frac{g}{\csc(\Theta)k}}\int_0^{\mathrm{i}v\sqrt{\frac{\csc(\Theta)k}{g}}} \frac{1}{1 + u^2}du &= \int_0^t dt
-  \hspace{1in} dv = -\mathrm{i}\sqrt{\frac{g}{\csc(\Theta)k}}du\\
-  -\mathrm{i}\frac{\sqrt{\csc(\Theta)}}{\sqrt{kg}} \int_0^{\mathrm{i}v\sqrt{\frac{\csc(\Theta)k}{g}}} \frac{1}{1 + u^2}du &= \int_0^t dt\\
-  -\mathrm{i}\frac{\sqrt{\csc(\Theta)}}{\sqrt{kg}} \tan^{-1}\Bigg(\mathrm{i}v\sqrt{\frac{\csc(\Theta)k}{g}}\Bigg) &= \int_0^t dt\\
-  \frac{\sqrt{\csc(\Theta)}}{\sqrt{kg}} \tanh^{-1}\Bigg(v\sqrt{\frac{\csc(\Theta)k}{g}}\Bigg) &= t\\
-  \frac{\tanh^{-1}\Bigg(v\sqrt{\frac{\csc(\Theta)k}{g}}\Bigg)}{\sqrt{kg\sin(\Theta)}} &= t\\ 
-  \frac{\tanh^{-1}\Bigg(v\sqrt{\frac{\csc(\Theta)k}{g}}\Bigg)}{\sqrt{kg\sin(\Theta)}} &= t\\
-  \tanh^{-1}\Bigg(v\sqrt{\frac{\csc(\Theta)k}{g}}\Bigg) &= t\sqrt{kg\sin(\Theta)}\\
-  v\sqrt{\frac{\csc(\Theta)k}{g}} &= \tanh\Big(t\sqrt{kg\sin(\Theta)}\Big)\\
-  v &= \frac{\tanh\Big(t\sqrt{kg\sin(\Theta)}\Big)}{\sqrt{\frac{\csc(\Theta)k}{g}}}\\    
+    \frac{\partial f}{\partial r} = 0\\
 \end{align*}
 
-Now we try integrating out the v:
-
-\begin{align*}
-  d &= \int_0^t v(t) dt\\
-  d &= \int_0^t \frac{\tanh\Big(t\sqrt{kg\sin(\Theta)}\Big)}{\sqrt{\frac{\csc(\Theta)k}{g}}} dt\\
-  d &= \frac{\log(\cosh\Big(t\sqrt{kg\sin(\Theta)}\Big))}{\sqrt{kg\sin(\Theta)}\sqrt{\frac{\csc(\Theta)k}{g}}}\\
-  d &= \frac{\log(\cosh\Big(t\sqrt{kg\sin(\Theta)}\Big))}{\sqrt{\frac{kg}{\csc(\Theta)}}\sqrt{\frac{\csc(\Theta)k}{g}}}\\
- d &= \frac{\log(\cosh\Big(t\sqrt{kg\sin(\Theta)}\Big))}{k}\\    
-  kd &= \log(\cosh\Big(t\sqrt{kg\sin(\Theta)}\Big))\\
-  e^{kd} &= \cosh\Big(t\sqrt{kg\sin(\Theta)}\Big)\\
-  \cosh^{-1}(e^{kd}) &= t\sqrt{kg\sin(\Theta)}\\
-  \frac{\cosh^{-1}(e^{kd})}{\sqrt{kg\sin(\Theta)}} &= t\\        
-\end{align*}
-
-\vspace{1 cm}
-
-\section{Problem Three}
-\textbf{A badly designed rocket has an initial mass of 7x104 kg and on firing it burns its fuel at a rate of 250 kg/s.
-The exhaust speed is 2500 m/s.
-If the rocket has a vertical ascent from rest on the Earth, how long after the rocket engines fire will the rocket lift off?} \\ \\
-\vspace{1 cm}
-
-\section{Problem Four}
-\textbf{A projectile of mass M explodes while in flight at speed v into three fragments.
-The first fragment, with mass m1=M/2 travels in the original direction of the projectile; the second, with mass m2=M/6 travels in the opposite direction, and the third comes to rest.
-The energy released in the explosion is equal to five times the kinetic energy of the projectile just before exploding.
-What are the velocities of the first two fragments?} \\ \\
-\vspace{1 cm}
-
-\section{Problem Five}
-\textbf{Find the center of mass of a uniform hemispherical shell of inner and outer radii a and b and mass M. Comment on the limiting cases. (5 bonus points)} \\ \\
-\vspace{1 cm}
+Meaning
 

### 2015-10-29 classical/latex/a5.tex
Add Classical HW 5

--- 2015-10-23 classical/latex/a4.tex
@@
 
-\title{Classical Mechanics Assignment 4}
+\title{Classical Mechanics Assignment 5}
 \author{Elliott Capek}
@@
 \section{Problem One}
-\textbf{Consider a rocket accelerating in free space by burning its fuel at a constant rate. Find the velocity at which the momentum of the rocket (neglecting the already spent fuel) is maximized.} \\
-
-We want to find the velocity at which $m\dot{y}$ is maximized. We find when the derivative of this with respect to time is zero, plug in for the values we find and go from there:
+\textbf{Find the shape assumed by an infinitely flexible rope of linear density $\rho$ and length L with both sides attached to the ceiling of a room.
+(Hint, the shape will minimize the gravitational potential energy of the rope).}
 
 \begin{align*}
-  \frac{\partial}{\partial t} m\dot{y} &= \dot{m}\dot{y} + m\ddot{y} = 0\\
-  m\ddot{y} &= -mg - \dot{m}v_{ex}\\
-  \dot{m}\dot{y} &= mg + \dot{m}v_{ex}\\
-  \dot{y} &= \frac{mg + \dot{m}v_{ex}}{\dot{m}}\\
+  PE_{aug} &= \int_{-a}^{a} \left[-gy\rho \sqrt{1 + x'^2} - \lambda\sqrt{1+x'^2}\right]dx\\
+  f(y,y',x) &= -gy\rho \sqrt{1 + x'^2} - \lambda\sqrt{1+x'^2}\\
+\end{align*}
+
+We use the $\frac{df}{dx'} = k$ form of the Euler-Lagrange equation:
+
+\begin{align*}
+  \frac{-gy\rho x'}{\sqrt{1 + x'^2}} - \frac{\lambda x'}{\sqrt{1+x'^2}} &= k\\
+  \frac{-gy\rho x' - \lambda x'}{\sqrt{1 + x'^2}} &= k\\
 \end{align*}
@@
 \section{Problem Two}
-\textbf{Consider a rocket subject to a linear resistive force $\vec{F} = -b\vec{v}$, but no other external forces.
-Find the velocity as a function of the remaining mass for a rocket that starts from rest and ejects mass at a constant rate $k = -\dot{m}$ and velocity $v_{ex}$.} \\
+\textbf{Demonstrate that the shortest distance between two points in 3D is a straight line.}
 
 \begin{align*}
-  m\ddot{y} &= -\frac{b}{m}\dot{y} - \dot{m}v_{ex}\\
+  L &= \int_{z_1}^{z_2} \sqrt{1 + x'^2 + y'^2} dz\\
+  f(x, x', y, y', z) &= \sqrt{1 + x'^2 + y'^2}\\
 \end{align*}
 
-\section{Problem Three}
-\textbf{Consider the surface generated by revolving a line connecting two points ($x_1$,$y_1$) and ($x_2$,$y_2$) about an axis coplanar with the two points.
-Find the equation of the line connecting the points such that the surface area generated by the revolution is minimum.}
+Since there is no x, y or z dependence in $f$, we can use the condensed Euler-Lagrange equation $\frac{df}{dn'} = k$:\\
 
 \begin{align*}
-  SA &= \int_{y_1}^{y_2} dA = \int_{y_1}^{y_2} 2\pi\sqrt{x^2+y^2}\sqrt{x'^2+1}dy\\
-  f(x, x', y) &= 2\pi\sqrt{x^2+y^2}\sqrt{x'^2+1}\\
+  \frac{df}{dx'} &= K_1\\
+  \frac{x'}{\sqrt{1 + x'^2 + y'^2}} &= K_1\\
+  x'^2 &= K_1^2\left(1 + x'^2 + y'^2\right)\\
+  x'^2 &= \frac{K_1^2\left(1 + y'^2\right)}{1-K_1}\\
+  y'^2 &= \frac{K_2^2\left(1 + x'^2\right)}{1-K_2}
+  \hspace*{1in} \mbox{Solutions for x' and y' are symmetric}\\
+  x'^2 &= \frac{K_1^2\left(1 + \frac{K_2^2\left(1 + x'^2\right)}{1-K_2}\right)}{1-K_1}\\
+  x'^2\left(1-K_1\right) &= K_1^2\left(1 + \frac{K_2^2\left(1 + x'^2\right)}{1-K_2}\right)\\
+  x'^2\left(1-K_1\right) &= K_1^2 + \frac{K_1^2K_2^2\left(1 + x'^2\right)}{1-K_2}\\
+  x'^2\left(1-K_1\right) &= K_1^2 + \frac{K_1^2K_2^2}{1-K_2} + \frac{x'^2K_1^2K_2^2}{1-K_2}\\
+  x'^2\left(1-K_1 - \frac{K_1^2K_2^2}{1-K_2}\right) &= K_1^2 + \frac{K_1^2K_2^2}{1-K_2}\\
+  x' &= \sqrt{\frac{K_1^2 + \frac{K_1^2K_2^2}{1-K_2}}{1-K_1 - \frac{K_1^2K_2^2}{1-K_2}}}\\
 \end{align*}
 
-\begin{align*}
-  \frac{\partial f}{\partial x} &- \frac{\partial}{\partial y} \frac{\partial f}{\partial x'} = 0\\
-  \frac{\partial f}{\partial x} &= \frac{2\pi x\sqrt{x'^2+1}}{\sqrt{x^2+y^2}}\\
-  \frac{\partial}{\partial y} \frac{\partial f}{\partial x'} &= 0\\
-\end{align*}
+Because $x'$ is constant and $y'^2 = \frac{K_2^2\left(1 + x'^2\right)}{1-K_2}$, y' is also constant. Thus we have the equation of a line!\\
 
-So...
-\begin{align*}
-  \frac{2\pi x\sqrt{x'^2+1}}{\sqrt{x^2+y^2}} = 0\\
-\end{align*}
-
-\section{Problem Four}
-\textbf{Consider a medium in which the refractive index $n$ is inversely proportional to $r_2$.
-That is, $n=a/r2$, where $r$ is the distance from the origin.
-Use Fermat’s principle to find the path of light traveling between two points that lie in the xy plane.
-Hint: use 2-dimensional polar coordinates.
-Show that the path found is part of a circle that intersects the origin of the reference frame.}
-
-We use polar coordinates to describe our two points. Our differential is $ds = \sqrt{r^2+r'^2}d\theta$.
-
-\begin{align*}
-  T &= \int_{\theta_1}^{\theta_2} \frac{1}{v}ds = \int_{\theta_1}^{\theta_2} \frac{a}{cr^2}\sqrt{r^2+r'^2}d\theta\\
-  f(r,r',\theta) &= \frac{a}{cr^2}\sqrt{r^2+r'^2}d\theta\\
-\end{align*}
-
-\begin{align*}
-  \frac{\partial f}{\partial r} - \frac{\partial}{\partial \theta} \frac{\partial f}{\partial r'} &= 0\\
-  \frac{\partial}{\partial \theta} \frac{\partial f}{\partial r'} &= 0\\
-\end{align*}
-
-And so:
-
-\begin{align*}
-    \frac{\partial f}{\partial r} = 0\\
-\end{align*}
-
-Meaning
-
+This was a fun easy problem.
+It was interesting that it was so easy to add arbitrary numbers of variables to the Euler-Lagrange equation.
+To add a variable, you just add another Euler-Lagrange equation to the system.
+It is strange that the equations are not coupled at all.
+It kind of makes sense, since you are just minimizing the path in two dimensions for each equation.
+You could take the 3D plot and look at its XZ side and minimize that.
+Then you could minimize the YZ plot.
+The XZ and YZ plots describe the same curve, but are not related.
 \end{document}

### 2015-11-01 classical/latex/a3.tex
HI

--- 2015-10-16 classical/latex/a3.tex
@@
 \usepackage{amsmath}
-\usepackage{marvosym} % Required for symbols in the colored box
-\usepackage{ifsym} % Required for symbols in the colored box
 
@@
 
-\allowdisplaybreaks
-
 % Format of the section titles
@@
 
-\title{Classical Mechanics Assignment 3}
+\title{Classical Mechanics Assignment X}
 \author{Elliott Capek}
@@
 \section{Problem One}
-\textbf{Consider an object that is thrown vertically up with initial speed $v_0$ in a linear medium.} \\
+\textbf{Consider an object that is thrown vertically up with initial speed $v_0$ in a linear medium.} \\ \\
 
@@
 \begin{align*}
-  m\ddot{y}(t) &= -mg - b\dot{y}(t)\\
-  m\ddot{y}(t) &= -b\big(\frac{mg}{b} + \dot{y}(t))\\
-  m\ddot{y}(t) &= -b\big(\dot{y}_{term} + \dot{y}(t))
-  \hspace{1in}\ddot{y} = 0 \rightarrow \dot{y}_{term} = \frac{mg}{b}\\
-  m\ddot{y}(t) &= -b\phi(t)
-  \hspace{1.6in} \phi(t) = \dot{y}_{term} + \dot{y}(t)\\
-  m\ddot{\phi(t)} &= -b\phi(t)
-  \hspace{1.6in} \dot{\phi(t)} = \ddot{y}(t)\\
-  \phi(t) &= Ae^{-\frac{bt}{m}}\\
-  \dot{y}(t) &= Ae^{-\frac{b}{m}t} - \dot{y}_{term}
-  \hspace{1.6in} \dot{y}(0) = \dot{y}_0 = A - \dot{y}_{term}\\
-  \dot{y}(t) &= (\dot{y}_0 + \dot{y}_{term})e^{-\frac{b}{m}t} - \dot{y}_{term}\\
-  \dot{y}(t) &= \dot{y}_0e^{-\frac{b}{m}t} + \dot{y}_{term}(e^{-\frac{b}{m}t} - 1)\\
-  y(t) &= -\frac{m}{b}e^{-\frac{b}{m}t}\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}t\\
+  v'(t) = -g - bv\\
+
 \end{align*}
@@
 
-\begin{align*}
-  \dot{y}(t) &= 0 = \dot{y}_0e^{-\frac{b}{m}t} + \dot{y}_{term}(e^{-\frac{b}{m}t} - 1)\\
-  \dot{y}_{term}(1 - e^{-\frac{b}{m}t}) &= \dot{y}_0e^{-\frac{b}{m}t}\\
-  \dot{y}_{term} - \dot{y}_{term}e^{-\frac{b}{m}t} &= \dot{y}_0e^{-\frac{b}{m}t}\\
-  \dot{y}_{term}e^{\frac{b}{m}t} &= \dot{y}_0 + \dot{y}_{term}\\
-  e^{\frac{b}{m}t} &= \frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}\\
-  t_{max} &= \frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\\
-  y(t_{max}) &= -\frac{m}{b}e^{-\frac{b}{m}\frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}}\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}\frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\\
-    y(t_{max}) &= -\frac{m}{b\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}\frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\\
-\end{align*}
-
-\textbf{c. Show that the limiting case for very small drag coefficient coincides with the solution for the frictionless motion.}\\
-For frictionless motion:
-\begin{align*}
-  \ddot{y}(t) &= -g\\
-  \dot{y}(t) &= \dot{y}_0 - gt\\
-  y(t) &= \dot{y}_0t - \frac{1}{2}gt^2\\
-  \dot{y}(t_{max}) &= 0 \rightarrow t_{max} = \frac{\dot{y}}{g}\\
-  y_{max} &= \frac{\dot{y}_0^2}{2g}\\
-\end{align*}
-
-For motion in a linear medium when $\lim_{b\rightarrow0}$:
-
-\begin{align*}
-  \lim_{b\rightarrow0}y(t_{max}) &= \lim_{b->0} \hspace{0.2cm} -\frac{m}{b\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\Big(\dot{y}_0 + \dot{y}_{term} \Big) - \dot{y}_{term}\frac{m}{b}\log{\frac{\dot{y}_0 + \dot{y}_{term}}{\dot{y}_{term}}}\\
-  \lim_{b\rightarrow0}y(t_{max}) &= \lim_{b->0} \hspace{0.2cm} -\frac{m}{b\frac{\dot{y}_0 + \frac{mg}{b}}{\frac{mg}{b}}}\Big(\dot{y}_0 + \frac{mg}{b} \Big) - \frac{mg}{b}\frac{m}{b}\log{\frac{\dot{y}_0 + \frac{mg}{b}}{\frac{mg}{b}}}\\
-  \lim_{b\rightarrow0}y(t_{max}) &= \lim_{b->0} \hspace{0.2cm} -\frac{m}{b\frac{\dot{y}_0 + \frac{mg}{b}}{\frac{mg}{b}}}\Big(\dot{y}_0 + \frac{mg}{b} \Big) - \frac{mg}{b}\frac{m}{b}\log{\frac{\dot{y}_0 + \frac{mg}{b}}{\frac{mg}{b}}}\\
-\end{align*}
+\textbf{c. Show that the limiting case for very small drag coefficient coincides with the solution for the frictionless motion.}
 
@@
 \textbf{A point mass of mass m slides down an inclined plane under the influence of gravity.
-If the motion is resisted by a force $F=kmv^2$, show that the time required to move a distance d after starting from rest is:}
+If the motion is resisted by a force F=kmv2, show that the time required to move a distance d after starting from rest is:}
 
 \begin{equation*}
-  t = \frac{\cosh^{-1}(e^{kd})}{\sqrt{kg\sin(\Theta)}}\\
+  t = \frac{\cosh^{-1}(e^{kd}}{\sqrt{kg\sin(\Theta)}}\\
 \end{equation*}
@@
 \textbf{Where $\Theta$ is the angle of inclination of the plane.} \\ \\
-
-We define a one-dimensional coordinate system where the positive axis points in the direction of motion of the point.\\
-
-\begin{align*}
-  \vec{F} &= \sin(\Theta)mg - kmv^2\\
-  m\dot{v} &= \sin(\Theta)mg - kmv^2\\
-  \frac{dv}{dt} &= \sin(\Theta)g - kv^2\\
-  \frac{dv}{\sin(\Theta)g - kv^2} &= dt\\
-  \int_0^{v} \frac{dv}{\sin(\Theta)g - kv^2} &= \int_0^t dt\\
-  \int_0^{v} \frac{\csc(\Theta)}{g(1 - \frac{\csc(\Theta)k}{g}v^2)}dv &= \int_0^t dt\\
-  \frac{\csc(\Theta)}{g}\int_0^{v} \frac{1}{1 - \frac{\csc(\Theta)k}{g}v^2}dv &= \int_0^t dt\\
-  \frac{\csc(\Theta)}{g}\int_0^{v} \frac{1}{1 - \frac{\csc(\Theta)k}{g}v^2}dv &= \int_0^t dt
-  \hspace{1in} u = \mathrm{i}v\sqrt{\frac{\csc(\Theta)k}{g}}\\
-  -\frac{\csc(\Theta)}{g}\mathrm{i}\sqrt{\frac{g}{\csc(\Theta)k}}\int_0^{\mathrm{i}v\sqrt{\frac{\csc(\Theta)k}{g}}} \frac{1}{1 + u^2}du &= \int_0^t dt
-  \hspace{1in} dv = -\mathrm{i}\sqrt{\frac{g}{\csc(\Theta)k}}du\\
-  -\mathrm{i}\frac{\sqrt{\csc(\Theta)}}{\sqrt{kg}} \int_0^{\mathrm{i}v\sqrt{\frac{\csc(\Theta)k}{g}}} \frac{1}{1 + u^2}du &= \int_0^t dt\\
-  -\mathrm{i}\frac{\sqrt{\csc(\Theta)}}{\sqrt{kg}} \tan^{-1}\Bigg(\mathrm{i}v\sqrt{\frac{\csc(\Theta)k}{g}}\Bigg) &= \int_0^t dt\\
-  \frac{\sqrt{\csc(\Theta)}}{\sqrt{kg}} \tanh^{-1}\Bigg(v\sqrt{\frac{\csc(\Theta)k}{g}}\Bigg) &= t\\
-  \frac{\tanh^{-1}\Bigg(v\sqrt{\frac{\csc(\Theta)k}{g}}\Bigg)}{\sqrt{kg\sin(\Theta)}} &= t\\ 
-  \frac{\tanh^{-1}\Bigg(v\sqrt{\frac{\csc(\Theta)k}{g}}\Bigg)}{\sqrt{kg\sin(\Theta)}} &= t\\
-  \tanh^{-1}\Bigg(v\sqrt{\frac{\csc(\Theta)k}{g}}\Bigg) &= t\sqrt{kg\sin(\Theta)}\\
-  v\sqrt{\frac{\csc(\Theta)k}{g}} &= \tanh\Big(t\sqrt{kg\sin(\Theta)}\Big)\\
-  v &= \frac{\tanh\Big(t\sqrt{kg\sin(\Theta)}\Big)}{\sqrt{\frac{\csc(\Theta)k}{g}}}\\    
-\end{align*}
-
-Now we try integrating out the v:
-
-\begin{align*}
-  d &= \int_0^t v(t) dt\\
-  d &= \int_0^t \frac{\tanh\Big(t\sqrt{kg\sin(\Theta)}\Big)}{\sqrt{\frac{\csc(\Theta)k}{g}}} dt\\
-  d &= \frac{\log(\cosh\Big(t\sqrt{kg\sin(\Theta)}\Big))}{\sqrt{kg\sin(\Theta)}\sqrt{\frac{\csc(\Theta)k}{g}}}\\
-  d &= \frac{\log(\cosh\Big(t\sqrt{kg\sin(\Theta)}\Big))}{\sqrt{\frac{kg}{\csc(\Theta)}}\sqrt{\frac{\csc(\Theta)k}{g}}}\\
- d &= \frac{\log(\cosh\Big(t\sqrt{kg\sin(\Theta)}\Big))}{k}\\    
-  kd &= \log(\cosh\Big(t\sqrt{kg\sin(\Theta)}\Big))\\
-  e^{kd} &= \cosh\Big(t\sqrt{kg\sin(\Theta)}\Big)\\
-  \cosh^{-1}(e^{kd}) &= t\sqrt{kg\sin(\Theta)}\\
-  \frac{\cosh^{-1}(e^{kd})}{\sqrt{kg\sin(\Theta)}} &= t\\        
-\end{align*}
 

### 2015-11-02 classical/latex/a4.tex

@@
 
-We want to find the velocity at which $m\dot{y}$ is maximized. We find when the derivative of this with respect to time is zero, plug in for the values we find and go from there:
+\begin{align*}
+  m\ddot{y} &= -\dot{m}V_{ex}\\
+  m\frac{dv}{dt} &= -\frac{dm}{dt}V_{ex}\\
+  dv &= -\frac{dm}{m}V_{ex}\\
+  \int_0^v dv &= \int_{m_0}^m-\frac{dm}{m}V_{ex}\\
+  v &= -\log(\frac{m}{m_0})V_{ex}\\
+  m &= m_0e^{\frac{-v}{V_{ex}}}\\
+\end{align*}
+
+We want to find the velocity which minimizes momentum...
 
 \begin{align*}
-  \frac{\partial}{\partial t} m\dot{y} &= \dot{m}\dot{y} + m\ddot{y} = 0\\
-  m\ddot{y} &= -mg - \dot{m}v_{ex}\\
-  \dot{m}\dot{y} &= mg + \dot{m}v_{ex}\\
-  \dot{y} &= \frac{mg + \dot{m}v_{ex}}{\dot{m}}\\
+  p &= mv\\
+  \frac{dp}{dv} &= m + v\frac{dm}{dv} = 0\\
+  m + \frac{m_0v}{V_{ex}}e^{\frac{-v}{V_{ex}}} &= 0\\
+  m &= \frac{m_0v}{V_{ex}}e^{\log(\frac{m}{m_0})}\\
+  v &= V_{ex}\\
 \end{align*}
+
+This problem has an interesting result.
+It isn't intuitive to me why it is the case - it probably has something to do with momentum conservation.
+The derivation itself was fairly simple, just another application of separation of variables.
+I like getting practice using this and other methods to solve differential equations.\\
 
@@
 \begin{align*}
-  m\ddot{y} &= -\frac{b}{m}\dot{y} - \dot{m}v_{ex}\\
+  (m_0+\dot{m}t)\ddot{y} &= -b\dot{y} - \dot{m}v_{ex}\\
+  \frac{d\dot{y}}{dt} &= \frac{-b\dot{y} - \dot{m}v_{ex}}{(m_0+\dot{m}t)}\\
+  \int_0^v \frac{d\dot{y}}{-b\dot{y} - \dot{m}v_{ex}} &= \int_0^t \frac{1}{(m_0+\dot{m}t)}dt\\
+  \frac{-1}{b}\log(\frac{bv + \dot{m}v_{ex}}{\dot{m}v_{ex}}) &= \frac{1}{\dot{m}}\log(\frac{m_0+\dot{m}t}{m_0})\\
+  -\dot{m}\log(\frac{bv + \dot{m}v_{ex}}{\dot{m}v_{ex}}) &= b\log(\frac{m_0+\dot{m}t}{m_0})\\
+  \frac{bv + \dot{m}v_{ex}}{\dot{m}v_{ex}} &= (\frac{m_0+\dot{m}t}{m_0})^{\frac{b}{\dot{m}}}\\
+  v &= \frac{1}{b}\Big(\dot{m}v_{ex}(\frac{m_0+\dot{m}t}{m_0})^{\frac{-b}{\dot{m}}} - \dot{m}v_{ex}\Big)\\
+  v &= \frac{kV_{ex}}{b}\Big(1 - (\frac{m}{m_0})^{\frac{b}{k}} \Big)\\
 \end{align*}
+
+This was another interesting problem.
+I think I'm getting really experienced at using separation of variables to solve equations.
+The equation for velocity makes sense - as the $m/m_0$ fraction decreases, the velocity increases.
+Since the fraction m/m0 is always less than one, the velocity is always possible.
+It is interesting that the terminal velocity once all the mass has been expended (if all the rocket is actually fuel) is $kV_{ex}/b$ - this is good to know.\\
 
@@
 \begin{align*}
-  SA &= \int_{y_1}^{y_2} dA = \int_{y_1}^{y_2} 2\pi\sqrt{x^2+y^2}\sqrt{x'^2+1}dy\\
-  f(x, x', y) &= 2\pi\sqrt{x^2+y^2}\sqrt{x'^2+1}\\
+  SA &= \int_{y_1}^{y_2} dA = \int_{y_1}^{y_2} 2\pi y\sqrt{y'^2+1}dx\\
+  f(y, y', x) &= 2\pi y\sqrt{y'^2+1}\\
 \end{align*}
@@
 \begin{align*}
-  \frac{\partial f}{\partial x} &- \frac{\partial}{\partial y} \frac{\partial f}{\partial x'} = 0\\
-  \frac{\partial f}{\partial x} &= \frac{2\pi x\sqrt{x'^2+1}}{\sqrt{x^2+y^2}}\\
-  \frac{\partial}{\partial y} \frac{\partial f}{\partial x'} &= 0\\
+  f - y'\frac{\partial f}{\partial y'} &= C\\
+  y\sqrt{y'^2+1} - \frac{yy'^2}{\sqrt{y'^2+1}} &= C\\
+  y\sqrt{y'^2+1} &= C + \frac{yy'^2}{\sqrt{y'^2+1}}\\
+  y(y'^2+1) &= C\sqrt{y'^2+1} + yy'^2\\
+  y &= C\sqrt{y'^2+1}\\
+  \frac{y^2}{C^2} - 1 &= y'^2\\
+  y^2 - C^2 &= y'^2C^2\\
+  \frac{\sqrt{y^2 - C^2}}{C} &= y'\\
+  \frac{\sqrt{y^2 - C^2}}{C} &= y'\\
+  \frac{dx}{dy} = \frac{1}{y'} &= \frac{C}{\sqrt{y^2 - C^2}}\\
+  x = \int \frac{dx}{dy}dy &= \int \frac{1}{y'} = \frac{C}{\sqrt{y^2 - C^2}}dy\\
+  x = C\cosh^{-1}(\frac{y}{C})+b\\
+  y = C\cosh(\frac{x-b}{C})\\
 \end{align*}
 
-So...
-\begin{align*}
-  \frac{2\pi x\sqrt{x'^2+1}}{\sqrt{x^2+y^2}} = 0\\
-\end{align*}
+So we can see that a hyperbolic cosine describes the smallest surface which connects two points.
+We can find the values of b and C by solving our result for the initial and final points, which must be done with a computer.
+This corresponds to a surface which starts at one point, bows inwards, then arcs outward again.
+This problem was a cool introduction to the power of using the Euler-Lagrange equation to find surfaces.
+It was actually fairly easy to get the solution, given the difficulty of solving the problem by any other means.\\
 
@@
   T &= \int_{\theta_1}^{\theta_2} \frac{1}{v}ds = \int_{\theta_1}^{\theta_2} \frac{a}{cr^2}\sqrt{r^2+r'^2}d\theta\\
-  f(r,r',\theta) &= \frac{a}{cr^2}\sqrt{r^2+r'^2}d\theta\\
+  f(r,r',\theta) &= \frac{a}{cr^2}\sqrt{r^2+r'^2}\\
 \end{align*}
@@
 \begin{align*}
-  \frac{\partial f}{\partial r} - \frac{\partial}{\partial \theta} \frac{\partial f}{\partial r'} &= 0\\
-  \frac{\partial}{\partial \theta} \frac{\partial f}{\partial r'} &= 0\\
+  f - r'\frac{\partial f}{\partial r'} &= C\\
+  \frac{a}{cr^2}\sqrt{r^2+r'^2} - \frac{ar'^2}{cr^2\sqrt{r^2+r'^2}} &= C\\
+  \frac{1}{r^2}\sqrt{r^2+r'^2} - \frac{r'^2}{r^2\sqrt{r^2+r'^2}} &= \frac{Cc}{a}\\
+  \sqrt{r^2+r'^2} &= \frac{Ccr^2}{a} + \frac{r'^2}{\sqrt{r^2+r'^2}}\\
+  r^2+r'^2 &= \frac{Ccr^2}{a}\sqrt{r^2+r'^2} + r'^2\\
+  \frac{a^2}{C^2c^2} &= r^2+r'^2\\
+  \sqrt{\frac{a^2}{C^2c^2} - r^2} &= \frac{dr}{d\theta}\\
+  \int d\theta &= \int \frac{dr}{\sqrt{\frac{a^2}{C^2c^2} - r^2}}\\
+  \theta &= \tan^{-1}\Big( \frac{c^2r\sqrt(\frac{a^2C^2-c^2r^2}{c^2})}{-a^2C^2+c^2r^2}   \Big)\\
+  \tan(\theta) &= \frac{c^2r\sqrt{\frac{a^2C^2-c^2r^2}{c^2}}}{-a^2C^2+c^2r^2}\\
+  r(\theta) &= \frac{C\tan(\theta)}{\sqrt{1+\tan(\theta)^2}}
 \end{align*}
 
-And so:
-
-\begin{align*}
-    \frac{\partial f}{\partial r} = 0\\
-\end{align*}
-
-Meaning
-
+This was a very interesting problem.
+We arrived at a concise solution for a very complicated physical problem using relatively straightforward techniques.
+It was interesting to note that the actual value of the physical constants is not important, since the path is minimized this particular way for any value of a and c.
+The C in our final expression is just a constant multiple that modifies the radius of the circle.
 \end{document}

### 2015-11-02 comp_model/latex/a2.tex

--- 2015-10-15 comp_model/latex/a1.tex
@@
 
-\section{1.4: Trapezoidal method coefficients}
-\begin{equation}
-  \sum_{j=0}^r \alpha_jV_{n+j} = h \sum_{j=0}^r \beta_j f(t_{n+j}, V_{n+j})\\
-\end{equation}
+\section{3.2: Steady-state Population Model}
 
-The trapezoidal method is this:\\
-\begin{equation}
-  V_{n+1} = V_n + h\Big( \frac{1}{2}f(V_n, t_n) + \frac{1}{2}f(V_{n+1}, t_{n+1}) \Big)
-\end{equation}
+We will let $u_j < 1$ for all $j$. We also let the following be true:
 
-By examining these two equations, we can see that:
+\begin{equation*}
+  \sum_j \beta_j \Big(\prod_{k<j}(1-u_k) \Big) = 1\\
+\end{equation*}
 
-\begin{align}
-  \alpha_0 = -1 &\hspace{2cm} \beta_0 = \frac{1}{2}\\
-  \alpha_1 = 0 &\hspace{2cm} \beta_1 = 1\\
-  \alpha_2 = 0 &\hspace{2cm} \beta_2 = 1\\
-  ...  &\hspace{2cm} ...\\
-  \alpha_{r-1} = 0 &\hspace{2cm} \beta_{r-1} = 1\\
-  \alpha_r = 1 &\hspace{2cm} \beta_1 = \frac{1}{2}\\    
-\end{align}
+Consider the following matrix:
 
 \begin{equation}
-  V_{n+1} = V_n + h\Big( \frac{1}{2}f(V_n, t_n) + \frac{1}{2}f(V_{n+1}, t_{n+1}) \Big)
+  P = 
+   \begin{pmatrix}
+     \beta_1 & \beta_2 & ... & \beta_{N-1} & \beta_N\\
+     1-u_1 & & & &\\
+     & 1-u_2 & & &\\
+     & & ... & &\\
+     & & 1-u_{N-1} & &\\
+     & & & 1-u_{N} & 0\\
+   \end{pmatrix}
 \end{equation}
-
-If we let our function f take the following form: $f(x, t) = \alpha x$ where $x(0) = x_0$, we can see that:
-
-\begin{align}
-  V_{n+1} &= V_n + \frac{h}{2}\Big( f(V_n, t_n) + f(V_{n+1}, t_{n+1}) \Big)\\
-  V_{n+1} &= V_n + \frac{h}{2}\Big( \alpha V_n + \alpha V_{n+1} \Big)\\
-  V_{n+1} &= V_n + \frac{\alpha hV_n}{2} + \frac{\alpha hV_{n+1}}{2}\\
-  V_{n+1} \Big(1 - \frac{\alpha h}{2}\Big) &= V_n \Big(1 + \frac{\alpha h}{2}\Big) \\
-  V_{n+1} &= V_n \frac{1 + \frac{\alpha h}{2}}{1 - \frac{\alpha h}{2}} \\
-  V_{n+1} &= V_n \frac{2 + \alpha h}{2 - \alpha h} \\
-  V_{n+1} &= V_0 \Big(\frac{2 + \alpha h}{2 - \alpha h}\Big)^{n+1} \\    
-\end{align}
-
-From this we can see that as long as the following condition is satisfied, we will have a converging solution:
-
-\begin{align}
-  |2 + \alpha h| \leq |2 - \alpha h|\\
-\end{align}
-\section{1.9: Order of Accuracy for Trapezoidal and Euler Methods}
-We want to analytically represent the error terms for the Euler and Trapezoidal methods and show how they depend on our timestep, $h$.\\
-
-First we solve for the behavior of our function using a Taylor Series.
-Let $x(t)$ be our function and $v(t) = \frac{dx(t)}{dt}$.
-Let $N = \frac{t}{h}$, where t is the current time and h the timestep.
-Finally, let $x(t_0) = 0$.\\
-
-\begin{align}
-  x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} v(t)dt\\
-  v(t) &= v(t_i) + (t-t_i)v'(t_i) + (t-t_i)^2v''(t_i) + ...\\
-  \int_{t_i}^{t_{i+1}} v(t)dt &\approx \int_{t_i}^{t_{i+1}} v(t_i) + (t-t_i)v'(t_i) + \frac{1}{2}(t-t_i)^2v''(t_i)  dt\\
-\end{align}
-
-We now do a u-substitution, where $S = \frac{t - t_i}{t_{i+1} - t_i}$ and $h = t_{i+1} - t_i$:
-
-\begin{align}
-  \int_{t_i}^{t_{i+1}} v(t)dt &\approx h\int_i^1 v(t_i) + hSv'(t_i) + \frac{1}{2}h^2S^2v''(t_i) dS\\
-  \int_{t_i}^{t_{i+1}} v(t)dt &\approx h \big( v(t_i)S + \frac{hS^2}{2}v'(t_i) + \frac{h^2S^3}{6}v''(t_i) \big) |_0^1 \\
-  \int_{t_i}^{t_{i+1}} v(t)dt &\approx h \big( v(t_i) + \frac{h}{2}v'(t_i) + \frac{h^2}{6}v''(t_i) \big) \\
-  \int_{t_i}^{t_{i+1}} v(t)dt &\approx hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{6}v''(t_i) \\
-\end{align}
-
-This gives us a fairly exact value for $\int_{t_i}^{t_{i+1}} v(t)dt$.
-We now need to compare this with the approximations that our two methods make.
-For Euler's method, $\int_{t_i}^{t_{i+1}} v(t)dt = hf(t_i)$.
-For the Trapezoidal method, $\int_{t_i}^{t_{i+1}} v(t)dt = \frac{h}{2}\Big( f(t_{i+1}) + f(t_{i}) \Big)$.
-We will first plug in the Euler's method approximation to the above expression to find its error term.
-
-\begin{align}
-  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i)\\
-  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{6}v''(t_i)\\
-  hv(t_i) &= hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{6}v''(t_i)\\
-  -\frac{h}{3}v''(t_i) &= v'(t_i)\\  
-\end{align}
-
-We now substitute this back into our original result:
-
-\begin{align}
-  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{6}v''(t_i) \\
-  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + \frac{h^2}{2}(-\frac{h}{3}v''(t_i)) + \frac{h^3}{6}v''(t_i) \\
-  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + -\frac{h^3}{4}(v''(t_i)) + \frac{h^3}{6}v''(t_i) \\
-  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + -\frac{h^3}{12}(v''(t_i))\\
-\end{align}
-
-As we can see, the error for the forward Euler's method is $-\frac{h^3}{12}(v''(t_i))$ per timestep. We now plug this result into our summed equation to find the total error for all timesteps:
-
-\begin{align}
-  x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} v(t)dt\\
-  x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} hv(t_i) + -\frac{h^3}{12}(v''(t_i))\\
-  x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} hv(t_i) + -\frac{h^3}{12}(v''(t_i))\\  
-\end{align}
-
-We only care about the error term, so we will focus on that.
-We substitute $N$ with $\frac{t - t_0}{h}$ and replace $v''(t_i)$ with a term that represents the average value of $v''(t_i)$ over the whole sum: $v''_{ave}$.
-
-\begin{align}
-  E &= \sum_{i=0}^{i=N-1} -\frac{h^3}{12}(v''(t_i))\\
-  E &= -N\frac{h^3}{12}v''_{ave}\\
-  E &= -\frac{t - t_0}{h} \frac{h^3}{12}v''_{ave}\\
-  E &= -(t - t_0)\frac{h^2}{12}v''_{ave}\\    
-\end{align}
-
-We do the same thing for the Trapezoidal method. We first find the error term:
-
-\begin{align}
-  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i)\\
-  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{4}v''(t_i)\\
-  \frac{h}{2}\Big(v(t_{i+1}) + v(t_i)\Big) &= hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{4}v''(t_i)\\
-  \frac{h}{2}\Big(-v(t_{i+1}) + v(t_i)\Big) - \frac{h^3}{4}v''(t_i) &= \frac{h^2}{2}v'(t_i)\\
-  \frac{1}{h}\Big(-v(t_{i+1}) + v(t_i)\Big) - \frac{h}{2}v''(t_i) &= v'(t_i)\\
-\end{align}
-
-We then substitute this back into our original result:
-
-\begin{align}
-  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{6}v''(t_i) \\
-  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + \frac{h^2}{2}\Big(\frac{1}{h}\Big(v(t_{i+1}) - v(t_i)\Big) - \frac{h}{4}v''(t_i) \Big) + \frac{h^3}{6}v''(t_i) \\
-  \int_{t_i}^{t_{i+1}} v(t)dt &= \frac{h}{2}\big(v(t_{i+1}) + v(t_i)\big) - \frac{h^3}{4}v''(t_i) + \frac{h^3}{6}v''(t_i) \\
-  \int_{t_i}^{t_{i+1}} v(t)dt &= \frac{h}{2}\big(v(t_{i+1}) + v(t_i)\big) - \frac{h^3}{12}v''(t_i)\\    
-\end{align}
-
-From this we can see that the error term for the trapezoidal rule is $-\frac{h^3}{12}v''(t_i)$ per timestep. We then sum over all timesteps as above:
-
-\begin{align}
-    x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} v(t)dt\\
-  x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} \frac{h}{2}\big(v(t_{i+1}) + v(t_i)\big) - \frac{h^3}{12}v''(t_i)\\
-  x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} \frac{h}{2}\big(v(t_{i+1}) + v(t_i)\big) + \sum_{i=0}^{i=N-1} -\frac{h^3}{12}v''(t_i)\\
-\end{align}
-
-We finally solve for the error:
-
-\begin{align}
-  E &= \sum_{i=0}^{i=N-1} -\frac{h^3}{12}(v''(t_i))\\
-  E &= -N\frac{h^3}{12}v''_{ave}\\
-  E &= -\frac{t - t_0}{h} \frac{h^3}{12}v''_{ave}\\
-  E &= -(t - t_0)\frac{h^2}{12}v''_{ave}\\
-\end{align}
-
-From this we can see that the Trapezoidal method's error is proportional to $h^2$.
-
 \end{document}

### 2015-11-03 classical/latex/a5.tex
Add new stuff Monday 11/2

@@
 \begin{align*}
-  PE_{aug} &= \int_{-a}^{a} \left[-gy\rho \sqrt{1 + x'^2} - \lambda\sqrt{1+x'^2}\right]dx\\
-  f(y,y',x) &= -gy\rho \sqrt{1 + x'^2} - \lambda\sqrt{1+x'^2}\\
+  PE_{aug} &= \int_{-a}^{a} \left[-gy\rho \sqrt{1 + y'^2} - \lambda\sqrt{1+y'^2}\right]dx\\
+  f(y,y',x) &= -\left(gy\rho + \lambda\right)\sqrt{1 + y'^2}\\
 \end{align*}
 
-We use the $\frac{df}{dx'} = k$ form of the Euler-Lagrange equation:
+We use the $f - y'\frac{df}{dy'} = k$ form of the Euler-Lagrange equation:
 
 \begin{align*}
-  \frac{-gy\rho x'}{\sqrt{1 + x'^2}} - \frac{\lambda x'}{\sqrt{1+x'^2}} &= k\\
-  \frac{-gy\rho x' - \lambda x'}{\sqrt{1 + x'^2}} &= k\\
+  -\left(gy\rho + \lambda\right)\sqrt{1 + y'^2} + \frac{\left(gy\rho + \lambda\right)y'^2}{\sqrt{1 + y'^2}} &= k\\
+  -\left(gy\rho + \lambda\right)\left(1 + y'^2\right) + \left(gy\rho + \lambda\right)y'^2 &= k\sqrt{1 + y'^2}\\
+  -\left(gy\rho + \lambda\right) &= k\sqrt{1 + y'^2}\\
+  \frac{\left(gy\rho + \lambda\right)^2 - k^2}{k^2} &= y'^2\\
+  \sqrt{\frac{\left(gy\rho + \lambda\right)^2 - k^2}{k^2}} &= \frac{dy}{dx}\\
+  \int dx &= \int \frac{k}{\sqrt{\left(gy\rho + \lambda\right)^2 - k^2}}dy\\
+  x &= \frac{1}{g\rho}\int \frac{1}{\sqrt{u^2/k^2 - 1}}du\\
+  x &= \frac{1}{g\rho}\cosh^{-1}\left(\frac{u}{k}\right) + C\\
+  x &= \frac{1}{g\rho}\cosh^{-1}\left(\frac{gy\rho + \lambda}{k}\right) + C\\
+  \frac{k\cosh(g\rho x - C) - \lambda}{g\rho} &= y(x)\\
 \end{align*}
+
+This was an interesting problem.
+In our solution k, C and $\lambda$ are all arbitrary constants which can be found by taking into account the initial conditions L, $x_1$ and $x_2$.
+It was tough to get the integral to work out.
+I knew it should have been an inverse hyperbolic cosine, so I had to intermediate solutions pretty heavily to get them in a form that would integrate to what I needed.\\
 

### 2015-11-04 classical/latex/a5.tex
add

--- 2015-10-29 classical/latex/a5.tex
@@
 \begin{align*}
-  PE_{aug} &= \int_{-a}^{a} \left[-gy\rho \sqrt{1 + x'^2} - \lambda\sqrt{1+x'^2}\right]dx\\
-  f(y,y',x) &= -gy\rho \sqrt{1 + x'^2} - \lambda\sqrt{1+x'^2}\\
+  PE_{aug} &= \int_{-a}^{a} \left[-gy\rho \sqrt{1 + x'^2} - \lambda\sqrt{1+x'^2}\right]dy\\
+  f(x,x',y) &= -gy\rho \sqrt{1 + x'^2} - \lambda\sqrt{1+x'^2}\\
 \end{align*}
@@
   \frac{-gy\rho x' - \lambda x'}{\sqrt{1 + x'^2}} &= k\\
+  \left(g^2y^2\rho^2 + \lambda^2 + 2\lambda gy\rho\right)x'^2 &= k^2\left(1 + x'^2\right)\\
+  \left(g^2y^2\rho^2 + \lambda^2 + 2\lambda gy\rho -k^2\right)x'^2 &= k^2\\
+  x' &= \sqrt{\frac{k^2}{\left(g^2y^2\rho^2 + \lambda^2 + 2\lambda gy\rho -k^2\right)}}\\
+  \int dx &= \int \sqrt{\frac{k^2}{\left(g^2y^2\rho^2 + \lambda^2 + 2\lambda gy\rho -k^2\right)}}dy\\
 \end{align*}

### 2015-11-06 classical/latex/a6.tex
Add classical assignment 6.

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{marvosym} % Required for symbols in the colored box
\usepackage{ifsym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

\allowdisplaybreaks

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Classical Mechanics Assignment 6}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem One: Ball on an Inclined Plane}
\textbf{Write the Lagrangian for a cylinder (mass m, radius R, and moment of inertia I) that rolls without slipping straight down an inclined plane at an angle $\alpha$ from the horizontal. Use as your generalized coordinates the cylinder’s distance x measured down the plane from its starting point. Solve the Lagrange equation for the cylinder’s acceleration $\ddot{x}$. Remember that $T = \frac{1}{2} mv^2 + \frac{1}{2} I\omega^2$ where $v$ is the velocity of the center of mass and $\omega$ is the cylinder’s angular velocity.}

\begin{align*}
  x' &= x\cos(\alpha)\\
  y' &= -x\sin(\alpha)\\
  \omega &= \frac{v}{2\pi R}\\
  \vspace{1cm}\\
  U &= mgy' = -mgx\sin(\alpha)\\
  T &= \frac{1}{2}mv^2 + \frac{1}{2}I\omega^2\\
  &= \frac{1}{2}m\left(\dot{x}'^2 + \dot{y}'^2\right) + \frac{1}{2}I\left(\frac{\dot{x}'^2 + \dot{y}'^2}{2\pi R}\right)\\
  &= \frac{1}{2}m\left(\dot{x}^2\sin^2(\alpha) + \dot{x}^2\cos^2(\alpha)\right) + \frac{1}{2}I\left(\frac{\dot{x}^2\sin^2(\alpha) + \dot{x}^2\cos^2(\alpha)}{4\pi^2R^2}\right)\\
  &= m\left(\dot{x}^2\right) + I\left(\frac{\dot{x}^2}{4\pi^2R^2}\right)\\
  &= \dot{x}^2\left(\frac{I + 4\pi^2R^2m}{4\pi^2R^2}\right)\\
  L &= \dot{x}^2\left(\frac{I + 4\pi^2R^2m}{4\pi^2R^2}\right) + mgx\sin(\alpha)\\
  \vspace{1cm}\\
  &\frac{dL}{dx} - \frac{d}{dt}\frac{dL}{d\dot{x}} = 0\\
  &mg\sin(\alpha) - \ddot{x}\left(\frac{I + 4\pi^2R^2m}{\pi^2R^2}\right) = 0\\
  \ddot{x} &= \frac{\pi^2R^2mg\sin(\alpha)}{I + 4\pi^2R^2m}\\
  x &= \frac{\pi^2R^2mg\sin(\alpha)}{2I + 8\pi^2R^2m}t^2
\end{align*}

\section{Problem Two: Bead on a Spiral}
\textbf{A smooth wire is bent into the shape of a helix, with cylindrical coordinates $\rho=R$ and $z=\lambda\phi$, where $R$ and $\lambda$ are constants and the z axis is vertically up (and gravity is vertically down). Using z as your generalized coordinate, write down the Lagrangian for a bead of mass m threaded on the wire. Find the Lagrange equation and the bead’s vertical acceleration $\ddot{z}$. In the limit that $R → 0$, what is $\ddot{z}$? Does this make sense?}

\begin{align*}
  v &= \sqrt{\dot{\rho}^2 + \rho^2\dot{\phi}^2 + \dot{z}^2}\\
  &= \sqrt{\rho^2\frac{\dot{z}^2}{\lambda^2} + \dot{z}^2}\\
  \vspace{1cm}\\
  T &= \frac{1}{2}mv^2\\
  &= \frac{1}{2}m\left(\rho^2\frac{\dot{z}^2}{\lambda^2} + \dot{z}^2\right)\\
  U &= mgz\\
  L &= \frac{1}{2}m\left(\rho^2\frac{\dot{z}^2}{\lambda^2} + \dot{z}^2\right) - mgz\\
  \vspace{1cm}\\
  &\frac{dL}{dz} - \frac{d}{dt}\frac{dL}{d\dot{z}} = 0\\
  -mg - \ddot{z}m\left(\frac{\rho^2}{\lambda^2}+1\right) &= 0\\
  \ddot{z} &= \frac{-g}{\frac{\rho^2}{\lambda^2}+1}\\
\end{align*}

\section{Problem Three: }
\textbf{Write the Lagrangian for a cylinder (mass m, radius r, and moment of inertia $I=\frac{mr^2}{2}$) that rolls without slipping in the bottom of a circular shaped hole of equation $z = −\sqrt{R^2 − x^2}$ , where z is the vertical axis (see figure below). Find the angular frequency of small oscillations about the bottom of the hole.}

\end{document}

### 2015-11-07 comp_model/crystal_phase.c
Add crystal stuff.

@@ -1,0 +1,81 @@
+#include <fcntl.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <sys/stat.h>
+#include <unistd.h>
+
+const int size = 10;
+
+void log_data(double data[size][size], int fd) {
+  printf("size: ", );
+  char* str;
+  printf("%s, fd: %d\n", str, fd);
+  if (write(fd, str, sizeof(char) * strlen(str)) == -1)
+    {
+      perror("write to arfile: ");
+      exit(EXIT_FAILURE);
+    }
+}
+
+int main() {
+  double dx = 0.03; //m
+  double dy = 0.03; //m
+  double dt = 0.0003; //s
+  double tau = 0.0003;
+  double epsilonbar = 0.01;
+  double mu = 1.0;
+  double k = 4;
+  double delta = 0.02;
+  double anisotropy = 4.0;
+  double alpha = 0.9;
+  double gamma = 10.0;
+  double Teq = 1.0;
+  int r = 10;
+
+  int fd;
+  if ((fd = open("data.txt", O_WRONLY | O_CREAT, S_IRUSR | S_IWUSR | S_IRGRP | S_IWGRP | S_IWOTH | S_IROTH)) == -1)
+    {
+      perror("open failed: ");
+      exit(EXIT_FAILURE);
+    }
+
+  printf("fd start: %d\n", fd);
+
+  double T[size-1][size-1];
+  double ep[size-1][size-1];
+  double dep_dt[size-1][size-1];
+  double phi[size-1][size-1];
+  double dphi_dx[size-1][size-1];
+  double dphi_dy[size-1][size-1];
+  double lap_T[size-1][size-1];
+  double lap_phi[size-1][size-1];
+  double phi_new[size-1][size-1];
+  double T_new[size-1][size-1];
+  double phi_grad_angle[size-1][size-1];
+  double t = 0;
+  double t_final = 100.0;
+
+  for (int x = 0; x<size; x++) {
+    for (int y = 0; y<size; y++) {
+      T[x][y] = 0;
+      ep[x][y] = 0;
+      dep_dt[x][y] = 0;
+      phi[x][y] = 0;
+      dphi_dx[x][y] = 0;
+      dphi_dy[x][y] = 0;
+      lap_T[x][y] = 0;
+      lap_phi[x][y] = 0;
+      phi_new[x][y] = 0;
[…]

### 2015-11-08 classical/latex/a6.tex
Add more stuff.

@@
 
+It took me a while to figure out why I needed to set up kinetic energy in cartesian coordinates.
+I thought I could just take my axes to be in the x direction (down the slope), but that isn't how we've done other problems.
+Although I did use the cartesian coordinates method, I found that using other cartesian coordinates (ones more natural to the problem, where x is in the direction of the wedge) would have yielded the same answer.
+I will try to apply these smarter cartesian coordinates to further problems.\\
+
 \section{Problem Two: Bead on a Spiral}
@@
   -mg - \ddot{z}m\left(\frac{\rho^2}{\lambda^2}+1\right) &= 0\\
-  \ddot{z} &= \frac{-g}{\frac{\rho^2}{\lambda^2}+1}\\
+  \ddot{z} &= \frac{-g}{\frac{R^2}{\lambda^2}+1}\\
 \end{align*}
+
+Fun problem.
+It was simple and easy to set up.
+It was interesting to think about how the problem would be different if the helix was moving.
+Also, when R goes to zero, the rhs of the equation just turns to -g, which is exactly what you would expect: a frictionless bead falling down a straight loop.
+It was cool that this works out.\\
 
@@
 
+\begin{align*}
+  x &= (R-r)\sin(\theta)\\
+  z &= -(R-r)\cos(\theta)\\
+  v^2 &= (R-r)^2\dot{\theta}^2\\
+  \omega^2 &= \frac{v^2}{r^2} = \frac{(R-r)^2\dot{\theta}^2}{r^2}\\
+  \vspace{1cm}\\
+  T &= \frac{1}{2}mv^2 + \frac{1}{2}I\omega^2\\
+  &= \frac{1}{2}m(R-r)^2\dot{\theta}^2 + \frac{1}{2}I\frac{(R-r)^2\dot{\theta}^2}{r^2}\\
+  &= \frac{1}{2}\left(m + \frac{I}{r^2}\right)(R-r)^2\dot{\theta}^2\\
+  U &= -mg(R-r)\cos(\theta)\\
+  L &= \frac{1}{2}\left(m + \frac{I}{r^2}\right)(R-r)^2\dot{\theta}^2 + mg(R-r)\cos(\theta)\\
+  \vspace{1cm}\\
+  \frac{dL}{d\theta} - \frac{d}{dt}\frac{dL}{d\dot{\theta}} &= 0\\
+  -mg(R-r)\sin(\theta) - \left(m + \frac{I}{r^2}\right)(R-r)^2\ddot{\theta} &= 0\\
+  \left(m + \frac{I}{r^2}\right)(R-r)\ddot{\theta} &= -mg\sin(\theta)\\
+  \ddot{\theta} &= \frac{-mg\sin(\theta)}{\left(m + \frac{I}{r^2}\right)(R-r)}\\
+  \ddot{\theta} &= \frac{-2g\sin(\theta)}{3(R-r)}\\
+\end{align*}
+
+We then take the small angle approximation:
+
+\begin{align*}
+  \ddot{\theta} &= \frac{-2g\theta}{3(R-r)}\\
+\end{align*}
+
+Suppose our solution for $\theta$ takes the form $\theta = e^{wt}$. Then we can see that $\omega = \sqrt{\frac{2g}{3(R-r)}}$.\\
+
+Cool problem!
+It felt nice to get a final result that looks like that of a pendulum, even though we used a completely different method.
+The small angle approximation helped a lot for this problem, I was struggling at applying it for a while.
+Remembering how to find the angular frequency was good for me, the method of guessing the solution takes the form of a pendulum solution and comparing the solutions together to create new information is cool.\\
+
+\section{Problem Four: }
+\textbf{The figure below shows a simple pendulum (mass m, length l) whose point of support P is attached to the edge of a wheel (center O, radius R) that is forced to rotate at a fixed angular velocity $\omega$.
+At t=0, the point P is level with O on the right.
+Write the Lagrangian and find the equation of motion for the angle $\phi$.
+[Hint: Be careful writing down the kinetic energy T.
+A safe way to get the velocity right is to write down the position of the bob at time t, and then differentiate.] Check that your answer makes sense in the special case that $\omega=0$.}
+\begin{align*}
+  x &= R\cos(\omega t) + l\sin(\phi)\\
+  y &= R\sin(\omega t) - l\cos(\phi)\\
+  \dot{x} &= -\omega R\sin(\omega t) + l\cos(\phi)\dot{\phi}\\
+  \dot{y} &= \omega R\cos(\omega t) + l\sin(\phi)\dot{\phi}\\
+  \vspace{1cm}\\
+  T &= \frac{1}{2}mv^2\\
+  &= \frac{1}{2}m\left(\omega^2R^2\sin^2(\omega t) + l^2\cos^2(\phi)\dot{\phi}^2 - 2\omega Rl\sin(\omega t)\cos(\phi)\dot{\phi} + \omega^2 R^2\cos^2(\omega t) + l^2\sin^2(\phi)\dot{\phi}^2 + 2\omega Rl\cos(\omega t)\sin(\phi)\dot{\phi}\right)\\
+  &= \frac{1}{2}m\left(\omega^2R^2 + l^2\dot{\phi}^2 + 2\dot{\phi}\omega Rl\left(\cos(\omega t)\sin(\phi) - \sin(\omega t)\cos(\phi)\right)\right)\\
+  &= \frac{1}{2}m\left(\omega^2R^2 + l^2\dot{\phi}^2 + 2\dot{\phi}\omega Rl\sin(\phi - \omega t)\right)\\
+  U &= mgy = mg\left(R\sin(\omega t) - l\cos(\phi)\right)\\
+  L &= \frac{1}{2}m\left(\omega^2R^2 + l^2\dot{\phi}^2 + 2\dot{\phi}\omega Rl\sin(\phi - \omega t)\right) - mg\left(R\sin(\omega t) - l\cos(\phi)\right)\\
+  &= \frac{m\omega^2R^2}{2} + \frac{ml^2\dot{\phi}^2}{2} + m\dot{\phi}\omega Rl\sin(\phi - \omega t) - mg\left(R\sin(\omega t) - l\cos(\phi)\right)\\
+  \vspace{1cm}\\
+  &\frac{dL}{d\phi} - \frac{d}{dt}\frac{dL}{d\dot{\phi}} = 0\\
+  &\frac{dL}{d\phi} = m\dot{\phi}\omega Rl\cos(\phi - \omega t) - mgl\sin(\phi)\\
+  &\frac{dL}{d\dot{\phi}} = ml^2\dot{\phi} + m\omega Rl\sin(\phi - \omega t)\\
+  &\frac{d}{dt}\frac{dL}{d\dot{\phi}} = ml^2\ddot{\phi} - m\omega Rl\cos(\phi - \omega t)\left(\dot{\phi} - \omega\right)\\
+  &m\dot{\phi}\omega Rl\cos(\phi - \omega t) - mgl\sin(\phi) - ml^2\ddot{\phi} + m\omega Rl\cos(\phi - \omega t)\left(\dot{\phi} - \omega\right) = 0\\
+  &\omega^2R\cos(\phi - \omega t) + g\sin(\phi) = l\ddot{\phi}\\
+\end{align*}
+
+This solution makes sense because, for the case where $\omega = 0$, the equation would be just $g\sin(\phi) = l\ddot{\phi}$, which is the equation for a pendulum.
+This was a cool problem, I feel like I'm getting experienced at using Lagrangians.
+Applying the equation of tangential velocity was also something I had to remember how to do.
+I think it's interesting that coming at the problem from different angles (cartesian coordinates in one direction or another) will yield the same answer.\\
+
 \end{document}

### 2015-11-08 comp_model/crystal_phase.c

@@ -7,15 +7,19 @@
 
 const int size = 10;
 
-void log_data(double data[size][size], int fd) {
-  printf("size: ", );
-  char* str;
-  printf("%s, fd: %d\n", str, fd);
-  if (write(fd, str, sizeof(char) * strlen(str)) == -1)
-    {
-      perror("write to arfile: ");
-      exit(EXIT_FAILURE);
+void log_data(double* data, int fd) {
+  char* str = malloc(100 * sizeof(char));
+  for (int y = 0; y < size; y++) {
+    for (int x = 0; x < size; x++) {
+      sprintf(str, "%2.5f", data[x+y*size]);
+      strcat(str, " ");
+      if (write(fd, str, sizeof(char) * strlen(str)) == -1)
+	{
+	  perror("write to arfile: ");
+	  exit(EXIT_FAILURE);
+	}
     }
+  }
 }
 
 int main() {
@@ -31,7 +35,7 @@ int main() {
   double alpha = 0.9;
   double gamma = 10.0;
   double Teq = 1.0;
-  int r = 10;
+  int r = 3;
 
   int fd;
   if ((fd = open("data.txt", O_WRONLY | O_CREAT, S_IRUSR | S_IWUSR | S_IRGRP | S_IWGRP | S_IWOTH | S_IROTH)) == -1)
@@ -42,36 +46,37 @@ int main() {
 
   printf("fd start: %d\n", fd);
 
-  double T[size-1][size-1];
-  double ep[size-1][size-1];
-  double dep_dt[size-1][size-1];
-  double phi[size-1][size-1];
-  double dphi_dx[size-1][size-1];
-  double dphi_dy[size-1][size-1];
-  double lap_T[size-1][size-1];
-  double lap_phi[size-1][size-1];
-  double phi_new[size-1][size-1];
-  double T_new[size-1][size-1];
-  double phi_grad_angle[size-1][size-1];
+  int arrsize = size*size;
+  double T[arrsize];
+  double ep[arrsize];
+  double dep_dt[arrsize];
+  double phi[arrsize];
+  double dphi_dx[arrsize];
+  double dphi_dy[arrsize];
+  double lap_T[arrsize];
+  double lap_phi[arrsize];
+  double phi_new[arrsize];
+  double T_new[arrsize];
+  double phi_grad_angle[arrsize];
[…]

### 2015-11-08 comp_model/plot.py

@@ -1,0 +1,40 @@
+#! /usr/bin/env python
+
+import math
+import numpy as np
+import time
+import matplotlib.pyplot as plt
+
+data = np.genfromtxt("data.txt", delimiter=" ")
+plt.ion()
+
+ax = plt.gca()
+ax.set_aspect("equal", adjustable="box")
+#ax.set_xlim(-40,40)
+#ax.set_ylim(-40,40)
+
+plt.xlabel('$x$ (m)')
+plt.ylabel('$y$ (m)')
+
+i = 0
+
+x = range(0, int(math.sqrt(len(data))))
+y = range(0, int(math.sqrt(len(data))))
+X, Y = np.meshgrid(x, y)
+Z = np.reshape(data, (math.sqrt(len(data)), math.sqrt(len(data))))
+
+print X, Z
+
+plt.contour(X, Y, Z)
+
+# while i < len(data) or loop:
+  
+#   if len(sys.argv) >= 2:
+#     if sys.argv[1] == "step":
+#       raw_input("Hit enter to step.")
+#       i += 1
+#     else
+#       i += 1
+#       plt.pause(0.001)
+
+#   plt.draw()

### 2015-11-09 comp_model/plot.py
add plot.py

@@ -4,6 +4,7 @@ import math
 import numpy as np
 import time
 import matplotlib.pyplot as plt
+import matplotlib.mlab as mlab
 
 data = np.genfromtxt("data.txt", delimiter=" ")
 plt.ion()
@@ -21,7 +22,11 @@ i = 0
 x = range(0, int(math.sqrt(len(data))))
 y = range(0, int(math.sqrt(len(data))))
 X, Y = np.meshgrid(x, y)
-Z = np.reshape(data, (math.sqrt(len(data)), math.sqrt(len(data))))
+#Z = np.reshape(data, (math.sqrt(len(data)), math.sqrt(len(data))))
+
+Z1 = mlab.bivariate_normal(X, Y, 1.0, 1.0, 0.0, 0.0)
+Z2 = mlab.bivariate_normal(X, Y, 1.5, 0.5, 1, 1)
+Z = 10.0 * (Z2 - Z1)
 
 print X, Z
 

### 2015-11-09 comp_model/latex/a4.tex

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage{amsmath}
\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Computational Modeling HW 4}
\author{Elliott Capek}

\begin{document}

\maketitle{}

\section{4.2: Dimensional Analysis of Ship Drag Model}
We want to create dimensionless $\pi$-groups for a model using the following variables and units: [D] = kg m/$\mbox{s}^2$, [L] = m, [v] = m/s, [$\mu$] = kg/(m s), [g] = m/$\mbox{s}^2$, and [$\rho$] = kg/$\mbox{m}^3$\\

We lay out our dimensional matrix like so:

\begin{equation*}
  \begin{tabular}{c|c c c c c c}
    ~ & L & v & $\mu$ & g & $\rho$ & D\\
    \hline
    kg & 0 &  0 &  1 &  0 &  1 & 1\\
    m  & 1 &  1 & -1 &  1 & -3 & 1 \\
    s  & 0 & -1 & -1 & -2 &  0 & 2 \\
  \end{tabular}
\end{equation*}

We then perform row reduction to put our matrix in a simplified form:

\begin{equation*}
  \begin{pmatrix}
    1 &  0 &  0 & -1 & -1 &  5\\
    0 &  1 &  0 &  2 & -1 & -3\\
    0 &  0 &  1 &  0 &  1 &  1\\
  \end{pmatrix}
\end{equation*}

This system has six columns and three dimensions, so by the Buckingham Pi theorem, it should have three basis vectors for its null space (or three pi groups). Three good basis vectors are these:

\begin{align*}
  S_1 =
  \begin{pmatrix}
    1\\
    -2\\
    0\\
    1\\
    0\\
    0\\
  \end{pmatrix}
  \hspace{1cm}
  S_2 =
  \begin{pmatrix}
    1\\
    1\\
    -1\\
    0\\
    1\\
    0\\
  \end{pmatrix}
  \hspace{1cm}
  S_3 =
  \begin{pmatrix}
    -5\\
    3\\
    -1\\
    0\\
    0\\
    1\\
  \end{pmatrix}
\end{align*}

From this we can derive $\pi$-groups of $\pi_1 = \frac{Lg}{v^2}$, $\pi_2 = \frac{Lv\rho}{\mu}$, and $\pi_3 = \frac{v^3D}{L^5\mu}$. These three vectors can be linearly combined in any way we want. If we just add the three together, we get the $\pi$ group $\pi_{1+2+3} = \frac{v^2g\rho D}{L^3\mu^2}$, which is unitless as desired. From this we can derive an equation for drag: $D = C\frac{L^3\mu}{v^2g\rho}$. Here C is a unitless proportionality constant.\\

This particular expression for drag, while unitless, doesn't make a lot of sense. Drag has the strongest proportionality to velocity, but we see here that it is inversely proportional. Water density $\rho$ should also be proportional to drag, not inversely proportional. It also isn't clear why the drag would be so strongly related to the ship's length L. But we don't expect intuitive answers through this method, there were an infinite number of linear combinations of our $\pi$-groups we could have chosen to represent D.\\

The $\pi$-groups themselves are also arbitrary. We could have chosen a different basis for our null space. The groups are all dimensionless, as desired, but this method of generating dimensionless parameters is not based in physics but math, and so there is no guarentee that the $\pi$-groups will mean anything.
\end{document}

### 2015-11-09 comp_model/crystal_phase.c
Add nonworking crystal code. Figure out the phiold/phi business?

@@ -1,11 +1,13 @@
 #include <fcntl.h>
+#include <math.h>
 #include <stdio.h>
 #include <stdlib.h>
 #include <string.h>
 #include <sys/stat.h>
 #include <unistd.h>
 
-const int size = 10;
+const double M_PI =  3.14159265358979323846;
+const int size = 100;
 
 void log_data(double* data, int fd) {
   char* str = malloc(100 * sizeof(char));
@@ -20,6 +22,12 @@ void log_data(double* data, int fd) {
 	}
     }
   }
+  sprintf(str, "\n");
+  if (write(fd, str, sizeof(char) * strlen(str)) == -1)
+    {
+      perror("write to arfile: ");
+      exit(EXIT_FAILURE);
+    }
 }
 
 int main() {
@@ -35,7 +43,7 @@ int main() {
   double alpha = 0.9;
   double gamma = 10.0;
   double Teq = 1.0;
-  int r = 3;
+  int r = 50;
 
   int fd;
   if ((fd = open("data.txt", O_WRONLY | O_CREAT, S_IRUSR | S_IWUSR | S_IRGRP | S_IWGRP | S_IWOTH | S_IROTH)) == -1)
@@ -43,8 +51,6 @@ int main() {
       perror("open failed: ");
       exit(EXIT_FAILURE);
     }
-
-  printf("fd start: %d\n", fd);
 
   int arrsize = size*size;
   double T[arrsize];
@@ -59,7 +65,7 @@ int main() {
   double T_new[arrsize];
   double phi_grad_angle[arrsize];
   double t = 0;
-  double t_final = 100.0;
+  double t_final = 0.005;
 
   for (int y = 0; y<size; y++) {
     for (int x = 0; x<size; x++) {
@@ -80,7 +86,72 @@ int main() {
       }
     }
   }
-  log_data(phi, fd);
+
+  int ym, yp, xm, xp;
+  double dep2_dx, dep2_dy, term1, term2, term3, m;
+  for (t=0; t<t_final; t+=dt) {
+    printf("t=%f\n", t);
+    for (int y = 0; y<size; y++) {
[…]

### 2015-11-09 comp_model/plot.py

@@ -6,7 +6,8 @@ import time
 import matplotlib.pyplot as plt
 import matplotlib.mlab as mlab
 
-data = np.genfromtxt("data.txt", delimiter=" ")
+data = np.genfromtxt("data.txt")
+print len(data)
 plt.ion()
 
 ax = plt.gca()
@@ -18,28 +19,11 @@ plt.xlabel('$x$ (m)')
 plt.ylabel('$y$ (m)')
 
 i = 0
+while i < len(data):
+    print str(i) + " of " + str(len(data))
+    Z = np.reshape(data[i], (math.sqrt(len(data[i])), math.sqrt(len(data[i]))))
+    plt.imshow(Z)
+    plt.pause(1)
+    i+=1
 
-x = range(0, int(math.sqrt(len(data))))
-y = range(0, int(math.sqrt(len(data))))
-X, Y = np.meshgrid(x, y)
-#Z = np.reshape(data, (math.sqrt(len(data)), math.sqrt(len(data))))
-
-Z1 = mlab.bivariate_normal(X, Y, 1.0, 1.0, 0.0, 0.0)
-Z2 = mlab.bivariate_normal(X, Y, 1.5, 0.5, 1, 1)
-Z = 10.0 * (Z2 - Z1)
-
-print X, Z
-
-plt.contour(X, Y, Z)
-
-# while i < len(data) or loop:
-  
-#   if len(sys.argv) >= 2:
-#     if sys.argv[1] == "step":
-#       raw_input("Hit enter to step.")
-#       i += 1
-#     else
-#       i += 1
-#       plt.pause(0.001)
-
-#   plt.draw()
+plt.show()

### 2015-11-10 comp_model/crystal_phase.c
add new stuff.

@@ -7,7 +7,7 @@
 #include <unistd.h>
 
 const double m_pi =  3.14159265358979323846;
-const int size = 15;
+const int size = 300;
 
 void log_data(double* data, int fd) {
   char* str = malloc(100 * sizeof(char));
@@ -34,6 +34,7 @@ int main() {
   double dx = 0.03; //m
   double dy = 0.03; //m
   double dt = 0.0003; //s
+  double t_final = 100*dt;
   double tau = 0.0003;
   double epsilonbar = 0.01;
   double mu = 1.0;
@@ -43,10 +44,10 @@ int main() {
   double alpha = 0.9;
   double gamma = 10.0;
   double Teq = 1.0;
-  int r = 10;
+  int r = sqrt(20);
 
   int fd;
-  if ((fd = open("data.txt", O_WRONLY | O_CREAT, S_IRUSR | S_IWUSR | S_IRGRP | S_IWGRP | S_IWOTH | S_IROTH)) == -1)
+  if ((fd = open("data.txt", O_WRONLY | O_CREAT | O_TRUNC, S_IRUSR | S_IWUSR | S_IRGRP | S_IWGRP | S_IWOTH | S_IROTH)) == -1)
     {
       perror("open failed: ");
       exit(EXIT_FAILURE);
@@ -65,7 +66,6 @@ int main() {
   double T_new[arrsize];
   double phi_grad_angle[arrsize];
   double t = 0;
-  double t_final = 0.0007;
 
   for (int y = 0; y<size; y++) {
     for (int x = 0; x<size; x++) {
@@ -87,10 +87,12 @@ int main() {
     }
   }
 
+  log_data(phi, fd);
+
   int ym, yp, xm, xp;
   double dep2_dx, dep2_dy, term1, term2, term3, m;
   for (t=0; t<t_final; t+=dt) {
-    printf("t=%f\n", t);
+    if ((int)(t/dt) % (int)(t_final/(dt*10)) == 0) printf("step %g of %g\n", t/dt, t_final/dt);
     for (int y = 0; y<size; y++) {
       for (int x = 0; x<size; x++) {
 	ym = y-1;

### 2015-11-10 comp_model/crystal_phase.c
Fix problem with array sizes...somehow?

@@ -7,7 +7,7 @@
 #include <unistd.h>
 
 const double m_pi =  3.14159265358979323846;
-const int size = 300;
+const int size = 1000;
 
 void log_data(double* data, int fd) {
   char* str = malloc(100 * sizeof(char));
@@ -34,7 +34,7 @@ int main() {
   double dx = 0.03; //m
   double dy = 0.03; //m
   double dt = 0.0003; //s
-  double t_final = 100*dt;
+  double t_final = 10*dt;
   double tau = 0.0003;
   double epsilonbar = 0.01;
   double mu = 1.0;
@@ -54,17 +54,17 @@ int main() {
     }
 
   int arrsize = size*size;
-  double T[arrsize];
-  double ep[arrsize];
-  double dep_dt[arrsize];
-  double phi[arrsize];
-  double dphi_dx[arrsize];
-  double dphi_dy[arrsize];
-  double lap_T[arrsize];
-  double lap_phi[arrsize];
-  double phi_new[arrsize];
-  double T_new[arrsize];
-  double phi_grad_angle[arrsize];
+  double* T = malloc(sizeof(double) * arrsize);
+  double* ep = malloc(sizeof(double) * arrsize);
+  double* dep_dt = malloc(sizeof(double) * arrsize);
+  double* phi = malloc(sizeof(double) * arrsize);
+  double* dphi_dx = malloc(sizeof(double) * arrsize);
+  double* dphi_dy = malloc(sizeof(double) * arrsize);
+  double* lap_T = malloc(sizeof(double) * arrsize);
+  double* lap_phi = malloc(sizeof(double) * arrsize);
+  double* phi_new = malloc(sizeof(double) * arrsize);
+  double* T_new = malloc(sizeof(double) * arrsize);
+  double* phi_grad_angle = malloc(sizeof(double) * arrsize);
   double t = 0;
 
   for (int y = 0; y<size; y++) {

### 2015-11-10 comp_model/crystal_phase.c
Simulator works, but doesn't create pretty snowflakes. Not sure why!

@@ -7,13 +7,14 @@
 #include <unistd.h>
 
 const double m_pi =  3.14159265358979323846;
-const int size = 1000;
+const int size = 500;
+const int animate = 0;
 
 void log_data(double* data, int fd) {
-  char* str = malloc(100 * sizeof(char));
+  char* str = malloc(10 * sizeof(char));
   for (int y = 0; y < size; y++) {
     for (int x = 0; x < size; x++) {
-      sprintf(str, "%2.5f", data[x+y*size]);
+      sprintf(str, "%2.2f", data[x+y*size]);
       strcat(str, " ");
       if (write(fd, str, sizeof(char) * strlen(str)) == -1)
 	{
@@ -34,17 +35,18 @@ int main() {
   double dx = 0.03; //m
   double dy = 0.03; //m
   double dt = 0.0003; //s
-  double t_final = 10*dt;
+  double t_final = 100*dt;
   double tau = 0.0003;
   double epsilonbar = 0.01;
   double mu = 1.0;
   double k = 4;
-  double delta = 0.02;
+  double delta = 0.04;
   double anisotropy = 4.0;
   double alpha = 0.9;
   double gamma = 10.0;
   double Teq = 1.0;
-  int r = sqrt(20);
+  double T0 = 0;
+  int r = sqrt(40);
 
   int fd;
   if ((fd = open("data.txt", O_WRONLY | O_CREAT | O_TRUNC, S_IRUSR | S_IWUSR | S_IRGRP | S_IWGRP | S_IWOTH | S_IROTH)) == -1)
@@ -87,7 +89,7 @@ int main() {
     }
   }
 
-  log_data(phi, fd);
+  if (animate) log_data(phi, fd);
 
   int ym, yp, xm, xp;
   double dep2_dx, dep2_dy, term1, term2, term3, m;
@@ -102,58 +104,69 @@ int main() {
 	
 	if (ym==-1) ym = size-1;
 	if (xm==-1) xm = size-1;
-	if (yp==size) ym = 0;
-	if (xp==size) xm = 0;
+	if (yp==size) yp = 0;
+	if (xp==size) xp = 0;
 
[…]

### 2015-11-10 comp_model/plot.py

@@ -4,6 +4,7 @@ from __future__ import division
 
 import math
 import numpy as np
+import sys
 import time
 import matplotlib.pyplot as plt
 import matplotlib.mlab as mlab
@@ -12,7 +13,7 @@ plt.ion()
 
 data = np.genfromtxt("data.txt")
 time = 1 # s
-frames = 10
+frames = 50
 
 ax = plt.gca()
 ax.set_aspect("equal", adjustable="box")
@@ -21,13 +22,19 @@ ax.set_aspect("equal", adjustable="box")
 
 plt.xlabel('$x$ (m)')
 plt.ylabel('$y$ (m)')
-
 i = 0
-while i < len(data):
-    print str(i) + " of " + str(len(data))
-    Z = np.reshape(data[i], (math.sqrt(len(data[i])), math.sqrt(len(data[i]))))
-    plt.imshow(Z)
-    plt.pause(time/frames)
-    i += int(len(data)/frames)
 
-plt.show()
+datafile = open("data.txt", "r").read()
+
+if  len(data.shape) == 1:
+    Z = np.reshape(data, (math.sqrt(len(data)), math.sqrt(len(data))))
+    plt.imshow(Z)
+    input("Press enter to continue.")
+    sys.exit(0)
+else:
+    while i < len(data):
+        print str(i) + " of " + str(len(data))
+        Z = np.reshape(data[i], (math.sqrt(len(data[i])), math.sqrt(len(data[i]))))
+        plt.imshow(Z)
+        plt.pause(time/frames)
+        i += int(len(data)/frames)

### 2015-11-10 comp_model/latex/paper1.tex
Add paper.

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage{amsmath}
\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Effect of Parameters in Phase-Field Model for simulating Solid-Liquid Transitions}
\author{Elliott Capek}

\begin{document}
\maketitle{}

\section{Introduction}
The Phase-Field model is a method for simulating phase transitions in systems. Phase transitions in real life are discontinuous, making them hard to represent analytically. The Phase-Field method overcomes this by representing the phase of a region with a space- and time-dependent function $\phi(r,t)$. $\phi$ ranges from $0$ to $1$, with either end of the range being one phase or the other, and numbers between $0$ and $1$ representing transitional regions. These transitional regions, while not physical, can approximate reality well if the distance between a 0-zone and a 1-zone is as small as possible. \\

In this paper we seek to reproduce the results of Sanal [1], where a solid-liquid phase transition with anisotropic dependence is time-evolved. The anisotropy (angular dependence) of the equations used in this paper allow for radially-symmetric crystals to form. In this formulation, $\phi=1$ corresponds to solid phase and $\phi=0$ corresponds to liquid phase. The equation for time-evolving $\phi$ is given in Sanal as:

\begin{equation}  \label{eq:dPhidt}
  \tau \frac{\partial \phi}{\partial t} = -\frac{\partial}{\partial x} \left(\epsilon \frac{d\epsilon}{d\theta}\frac{\partial\phi}{\partial y}\right) + \frac{\partial}{\partial y} \left(\epsilon \frac{d\epsilon}{d\theta}\frac{\partial \phi}{\partial x}\right) + \nabla \cdot \left(\epsilon^2\nabla\phi\right) + \phi\left(1-\phi\right)\left(\phi-0.5+m\right)
\end{equation}

Here $\tau$ is the relaxation time, or the time it takes $\phi$ to come into equilibrium with its neighbors. $\epsilon$ is a small $\theta$-dependent parameter which controls how quickly the solid-liquid interface moves. $\theta$ is the angle from the origin measured from the horizontal. Making $\epsilon$ depend on $\theta$ introduces anisotropy into the model, allowing different angles to spread at different rates. $m$ is the thermodynamic driving force, a temperature-dependent number which determines how favorable solidification is. Larger values of temperature (and hence $m$) correspond to quicker solidification. \\

The equation for $\epsilon(\theta)$ is given as follows:

\begin{equation}
  \epsilon = 1 + \delta\cos\left(j\left(\theta_0-\theta\right)\right)
\end{equation}

where $\delta$ is the strength of anisotropy, $j$ is the harmonic number and $\theta_0$ is the angular offset. The higher $j$ is, the more angular extrema will exist in the system.\\

Temperature is also important to this model, as it decides the value of m, the rate of interface front movement. Temperature's time-evolution is given by:

\begin{equation}  \label{eq:dTdt}
  \frac{\partial T}{\partial t} = \nabla^2T + k\frac{\partial \phi}{\partial t}
\end{equation}

where k is the latent heat of the phase change. Liquid freezing into a solid is exergonic, so heat is evolved when $\frac{\partial\phi}{\partial t}$ is positive. \\

\section{Method}
To simulate the behavior of the differential equation given by Sanal, we use the forwards-Euler finite difference method to time-evolve Equations (\ref{eq:dPhidt}) and (\ref{eq:dTdt}). We implement this in a C program (see Code section) and a Python script using matplotlib to plot our results.

\section{Results}

\section{Next Steps}

\end{document}

### 2015-11-10 comp_model/crystal_phase.c

@@ -8,13 +8,15 @@
 
 const double m_pi =  3.14159265358979323846;
 const int size = 500;
-const int animate = 0;
+const int animate = 1;
+const int skiprate = 1;
+const int debug = 0;
 
 void log_data(double* data, int fd) {
   char* str = malloc(10 * sizeof(char));
   for (int y = 0; y < size; y++) {
     for (int x = 0; x < size; x++) {
-      sprintf(str, "%2.2f", data[x+y*size]);
+      sprintf(str, "%2.4f", data[x+y*size]);
       strcat(str, " ");
       if (write(fd, str, sizeof(char) * strlen(str)) == -1)
 	{
@@ -35,12 +37,12 @@ int main() {
   double dx = 0.03; //m
   double dy = 0.03; //m
   double dt = 0.0003; //s
-  double t_final = 100*dt;
+  double t_final = 500*dt;
   double tau = 0.0003;
   double epsilonbar = 0.01;
   double mu = 1.0;
   double k = 4;
-  double delta = 0.04;
+  double delta = 0.01;
   double anisotropy = 4.0;
   double alpha = 0.9;
   double gamma = 10.0;
@@ -145,7 +147,7 @@ int main() {
 
 	term3 = dep2_dx*dphi_dx[x+y*size] + dep2_dy*dphi_dy[x+y*size];
 
-	double phiold = phi[x+y*size]; // wrong stuff
+	double phiold = phi[x+y*size];
 	
 	m = alpha/m_pi * atan(gamma*(Teq-T[x+y*size]));
 	
@@ -154,18 +156,20 @@ int main() {
                phiold*(1.0-phiold)*(phiold - 0.5 + m)) * dt / tau;
 
 	T_new[x+y*size] = T[x+y*size] + lap_T[x+y*size]*dt + k*(phi_new[x+y*size] - phiold);
-	//phi[x+y*size] = phi_new[x+y*size]; // wrong stuff
-	//T[x+y*size] = T_new[x+y*size];    // wrong stuff
+	/* if (x==3 && y==3) printf("%2.2g\n", lap_T[x+y*size]*dt); */
       }
     }
     
     for (int y = 0; y<size; y++) {
       for (int x = 0; x<size; x++) {
+	if (debug) printf("%2.4g ", T[x+y*size]);
     	phi[x+y*size] = phi_new[x+y*size];
     	T[x+y*size] = T_new[x+y*size];
       }
+      if (debug) printf("\n");
     }
[…]

### 2015-11-10 comp_model/plot.py

@@ -13,7 +13,7 @@ plt.ion()
 
 data = np.genfromtxt("data.txt")
 time = 1 # s
-frames = 50
+frames = 20
 
 ax = plt.gca()
 ax.set_aspect("equal", adjustable="box")
@@ -37,4 +37,4 @@ else:
         Z = np.reshape(data[i], (math.sqrt(len(data[i])), math.sqrt(len(data[i]))))
         plt.imshow(Z)
         plt.pause(time/frames)
-        i += int(len(data)/frames)
+        i += math.ceil(len(data)/frames)

### 2015-11-12 comp_model/latex/paper1.tex
Add additions to paper.

@@
 
-where $\delta$ is the strength of anisotropy, $j$ is the harmonic number and $\theta_0$ is the angular offset. The higher $j$ is, the more angular extrema will exist in the system.\\
+where $\delta$ is the strength of anisotropy, $j$ is the harmonic number and $\theta_0$ is the angular offset. The higher $j$ is, the more angular ``modes'' will exist in the system.\\
 
@@
 
-\section{Method}
-To simulate the behavior of the differential equation given by Sanal, we use the forwards-Euler finite difference method to time-evolve Equations (\ref{eq:dPhidt}) and (\ref{eq:dTdt}).
-We implement this in a C program (see Code section) and a Python script using matplotlib to plot our results.
+\section{Derivation of $\phi$-evolution}
+The derivation of the equation for $\frac{d\phi}{dt}$ begins with a free energy functional:
 
-\section{Results}
+\begin{equation}
+  F = \int f(\phi,m) + \frac{\epsilon(\phi)^2}{2}|\nabla\phi|^2 dV
+\end{equation}
 
-\section{Next Steps}
+F, the free energy, is a functional of $\phi(r,t)$ and $\epsilon(\phi)$.
+The above equation has two components: function $f(\phi,m)$, which is the potential energy of the system, and the $\epsilon$ term, which represents the surface tension of the crystal.\\
+
+Potential f only needs to be defined between $\phi=0$ and $\phi=1$, due to the definition of $\phi$.
+f should have local minima at $\phi=0$ and $\phi=1$ to encourage settling at either phase and discourage transitional values of $\phi$.
+The potential is also m-dependent.
+m is the thermodynamic driving force of the phase-change reaction.
+It decides if it is favorable or unfavorable to be a solid or liquid, and is a function of T.
+It will be defined later.
+Potential depends on m in that the higher m is, the more favorable the liquid phase should be.
+Sanal presents the following as the phase-transition potential:
+
+\begin{equation}
+  f(\phi,m) = \frac{1}{4}\phi^4 - (\frac{1}{2} - \frac{1}{3}m)\phi^3 + (\frac{1}{4}-\frac{1}{2}m)\phi^2
+\end{equation}
+
+The epsilon term in the F functional exists to make the phase-transition region as small as possible, and also to introduce anisotropy.
+The potential f discourages $\phi$ from taking transition values, but it isn't detailed enough.
+It discourages transition values system-wide, but we want to make the transition region as small as possible.
+This requires a $\nabla\phi$, which is present in the $\epsilon$-term.
+The larger $\nabla\phi$, the smaller the solid-liquid interface will be.
+Because we want to encourage small solid-liquid as well as liquid-solid transitions, the gradient term becomes $|\nabla\phi|$.
+The actual $\epsilon(\phi)$ function controls where $\phi$ should change.
+This $\epsilon$ is actually a function of the negative gradient of phi, $\epsilon(-\nabla\phi)$ since it only depends on the angle of the interface.
+This is to introduce anisotropy.
+$\epsilon$ is defined as follows:
+
+\begin{equation}
+  \epsilon = 1 + \delta\cos(j(\phi_0-\phi))
+\end{equation}
+
+where $\delta$ is a small value which controls strength of anisotropy, $j$ controls how many angular nodes show up, and $\phi_0$ rotates the crystal.
+The basic function of $\epsilon$ is to control where the phase-transition should appear, or in this case, how expensive certain angles of the phase transition should be.
+Certain angles are cheap, and so the system will have the barrier take those on, and some will be expensive, so the system will try and minimize the area of the interface at that angle.
+In particular, values of $j(\phi_0-\phi)$ close to $-1$ will be favored and values close to $1$ will be disfavored.
+
+The free energy functional is a value that is minimized when a system evolves in time. We can use this fact to time-evolve $\phi$.
 

### 2015-11-12 comp_model/crystal_phase.c
Add stuff for paper.

@@ -7,9 +7,9 @@
 #include <unistd.h>
 
 const double m_pi =  3.14159265358979323846;
-const int size = 500;
+const int size = 300;
 const int animate = 1;
-const int skiprate = 1;
+const int skiprate = 10;
 const int debug = 0;
 
 void log_data(double* data, int fd) {
@@ -34,21 +34,21 @@ void log_data(double* data, int fd) {
 }
 
 int main() {
-  double dx = 0.03; //m
-  double dy = 0.03; //m
+  double dx = 0.034; //m
+  double dy = 0.034; //m
   double dt = 0.0003; //s
-  double t_final = 500*dt;
+  double t_final = 1000*dt; // Just forced temperature to 1, unphysical!
   double tau = 0.0003;
   double epsilonbar = 0.01;
   double mu = 1.0;
-  double k = 4;
-  double delta = 0.01;
+  double k = 1;
+  double delta = 0.02;
   double anisotropy = 4.0;
   double alpha = 0.9;
   double gamma = 10.0;
   double Teq = 1.0;
   double T0 = 0;
-  int r = sqrt(40);
+  int r = 70;
 
   int fd;
   if ((fd = open("data.txt", O_WRONLY | O_CREAT | O_TRUNC, S_IRUSR | S_IWUSR | S_IRGRP | S_IWGRP | S_IWOTH | S_IROTH)) == -1)
@@ -139,7 +139,7 @@ int main() {
 	dep2_dx = (ep[xp+y*size]*ep[xp+y*size] - ep[xm+y*size]*ep[xm+y*size]) / dx;
 	dep2_dy = (ep[x+yp*size]*ep[x+yp*size] - ep[x+ym*size]*ep[x+ym*size]) / dy;
 
-	term1 = -(ep[xp+y*size]*dep_dt[xp+y*size]*dphi_dy[xp+y*size]
+	term1 = (ep[xp+y*size]*dep_dt[xp+y*size]*dphi_dy[xp+y*size]
 		 - ep[xm+y*size]*dep_dt[xm+y*size]*dphi_dy[xm+y*size]) / dx;
 	
 	term2 = (ep[x+yp*size]*dep_dt[x+yp*size]*dphi_dx[x+yp*size]
@@ -162,7 +162,7 @@ int main() {
     
     for (int y = 0; y<size; y++) {
       for (int x = 0; x<size; x++) {
-	if (debug) printf("%2.4g ", T[x+y*size]);
[…]

### 2015-11-12 comp_model/plot.py

@@ -12,7 +12,10 @@ import matplotlib.mlab as mlab
 plt.ion()
 
 data = np.genfromtxt("data.txt")
-time = 1 # s
+
+hi = raw_input("Press any key to play animation.")
+
+time = 10 # s
 frames = 20
 
 ax = plt.gca()
@@ -28,7 +31,11 @@ datafile = open("data.txt", "r").read()
 
 if  len(data.shape) == 1:
     Z = np.reshape(data, (math.sqrt(len(data)), math.sqrt(len(data))))
+    fig = plt.figure()
     plt.imshow(Z)
+    fig.suptitle('Anisotropy j=2', fontsize=20)
+    plt.xlabel('x')
+    plt.ylabel('y')
     input("Press enter to continue.")
     sys.exit(0)
 else:

### 2015-11-13 classical/latex/a7.tex
add new a7

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{graphicx}
\usepackage{amsmath}

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

\allowdisplaybreaks

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Classical Mechanics Assignment 7}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem One: Shrinking Pendulum Hamiltonian}

\begin{align*}
  x &= \ell\sin\theta\\
  y &= -\ell\cos\theta\\
  \dot{x} &= \dot{\ell}\sin\theta + \ell\cos\theta\dot{\theta}\\
  \dot{y} &= -\dot{\ell}\cos\theta + \ell\sin\theta\dot{\theta}\\
  \vspace{1cm}\\
  T &= \frac{1}{2}m\left(\dot{\ell}^2 + \ell^2\dot{\theta}^2\right)\\
  U &= -mg\ell\cos\theta\\
  \mathcal{L} &= \frac{1}{2}m\left(\dot{\ell}^2 + \ell^2\dot{\theta}^2\right) + mg\ell\cos\theta\\
  \vspace{1cm}\\
  p_{\theta} &= m\ell^2\dot{\theta}\\
  p_{\ell} &= m\dot{\ell}\\
  \mathcal{H} &= p_{\theta}\dot{\theta} + p_{\ell}\dot{\ell} - \mathcal{L} = m\ell^2\dot{\theta}^2 + m\dot{\ell}^2 - \frac{1}{2}m\left(\dot{\ell}^2 + \ell^2\dot{\theta}^2\right) - mg\ell\cos\theta\\
  &= \frac{1}{2}m\left(\alpha^2 + \ell^2\dot{\theta}^2\right) - mg\ell\cos\theta\\
  \vspace{1cm}\\
  T + U &= \frac{1}{2}m\left(\alpha^2 + \ell^2\dot{\theta}^2\right) - mg\ell\cos\theta\\
\end{align*}

As we can see, the Hamiltonian and the total energy for this system are identical. Because the Lagrangian for this system has no time-dependence, we know that $\mathcal{H}$ must be conserved, and thus total energy is conserved.\\

This problem was a basic tour of how to set up problems using the Hamiltonian. We didn't solve it to completion, but the answer could have been found by deriving $\mathcal{H}$ with respect to momentum and the velocities. It took me a second to figure out this problem, since I wasn't considering that there are actually two generalized coordinates in this problem - $\theta$ and $l$. Once I figured this out, the problem became much easier.\\

\section{Problem Two}

\textbf{A spherical pendulum consists of a bob of mass m attached to a weightless extension rod of length l. The end of the rod opposite the bob pivots freely (in all directions) about some fixed point. Set up the Hamiltonian function in spherical coordinates. (If $p_{\phi}=0$, the result is the same as that for the plane pendulum.) Combine the term that depends on $p_{\phi}$ with the ordinary potential energy to find the effective potential $V(\theta,p_{\phi})$. Sketch V as a function of $\theta$ for several values of $p_{\phi}$, including $p_{\phi}$ = 0. Discuss the features of the motion, pointing out the differences between $p_{\phi}=0$ and $p_{\phi} \neq 0$. Discuss the limiting case of the conical pendulum ($\theta = $constant) with reference to the V − $\theta$ diagram.}

\begin{align*}
  x &= \ell\sin\phi\cos\theta\\
  y &= \ell\sin\phi\sin\theta\\
  \dot{x} &= \ell\cos\phi\cos\theta\dot{\phi} - \ell\sin\phi\sin\theta\dot{\theta}\\
  \dot{y} &= \ell\cos\phi\sin\theta\dot{\phi} + \ell\sin\phi\cos\theta\dot{\theta}\\
  T &= \frac{1}{2}m\left(\ell^2\cos^2\phi\dot{\phi}^2 + \ell^2\sin^2\phi\dot{\theta}^2\right)\\
\end{align*}

\end{document}

### 2015-11-19 classical/latex/a7.tex
Addign new stuff.

@@
 \usepackage{amsmath}
+\usepackage{marvosym} % Required for symbols in the colored box
+\usepackage{ifsym} % Required for symbols in the colored box
 
@@
 
-\section{Problem One: Shrinking Pendulum Hamiltonian}
+\section{Problem One}
+\textbf{Consider a simple plane pendulum consisting of a mass m attached to a string of length l. After the pendulum is set into motion, the length of the string is shortened at a constant rate:}
+\begin{equation}
+  \frac{dl}{dt} = −\alpha
+\end{equation}
+
+\textbf{The suspension point remains fixed. Compute the Lagrangian and Hamiltonian functions. Compare the Hamiltonian and the total energy, and discuss the conservation of energy for the system.}
 
@@
   y &= -\ell\cos\theta\\
-  \dot{x} &= \dot{\ell}\sin\theta + \ell\cos\theta\dot{\theta}\\
-  \dot{y} &= -\dot{\ell}\cos\theta + \ell\sin\theta\dot{\theta}\\
+  \dot{x} &= -\alpha\sin\theta + \ell\cos\theta\dot{\theta}\\
+  \dot{y} &= \alpha\cos\theta + \ell\sin\theta\dot{\theta}\\
   \vspace{1cm}\\
-  T &= \frac{1}{2}m\left(\dot{\ell}^2 + \ell^2\dot{\theta}^2\right)\\
-  U &= -mg\ell\cos\theta\\
-  \mathcal{L} &= \frac{1}{2}m\left(\dot{\ell}^2 + \ell^2\dot{\theta}^2\right) + mg\ell\cos\theta\\
-  \vspace{1cm}\\
-  p_{\theta} &= m\ell^2\dot{\theta}\\
-  p_{\ell} &= m\dot{\ell}\\
-  \mathcal{H} &= p_{\theta}\dot{\theta} + p_{\ell}\dot{\ell} - \mathcal{L} = m\ell^2\dot{\theta}^2 + m\dot{\ell}^2 - \frac{1}{2}m\left(\dot{\ell}^2 + \ell^2\dot{\theta}^2\right) - mg\ell\cos\theta\\
-  &= \frac{1}{2}m\left(\alpha^2 + \ell^2\dot{\theta}^2\right) - mg\ell\cos\theta\\
-  \vspace{1cm}\\
-  T + U &= \frac{1}{2}m\left(\alpha^2 + \ell^2\dot{\theta}^2\right) - mg\ell\cos\theta\\
-\end{align*}
-
-As we can see, the Hamiltonian and the total energy for this system are identical.
-Because the Lagrangian for this system has no time-dependence, we know that $\mathcal{H}$ must be conserved, and thus total energy is conserved.\\
-
-This problem was a basic tour of how to set up problems using the Hamiltonian.
-We didn't solve it to completion, but the answer could have been found by deriving $\mathcal{H}$ with respect to momentum and the velocities.
-It took me a second to figure out this problem, since I wasn't considering that there are actually two generalized coordinates in this problem - $\theta$ and $l$.
-Once I figured this out, the problem became much easier.\\
-
-\section{Problem Two}
-
-\textbf{A spherical pendulum consists of a bob of mass m attached to a weightless extension rod of length l.
-The end of the rod opposite the bob pivots freely (in all directions) about some fixed point.
-Set up the Hamiltonian function in spherical coordinates.
-(If $p_{\phi}=0$, the result is the same as that for the plane pendulum.) Combine the term that depends on $p_{\phi}$ with the ordinary potential energy to find the effective potential $V(\theta,p_{\phi})$.
-Sketch V as a function of $\theta$ for several values of $p_{\phi}$, including $p_{\phi}$ = 0.
-Discuss the features of the motion, pointing out the differences between $p_{\phi}=0$ and $p_{\phi} \neq 0$.
-Discuss the limiting case of the conical pendulum ($\theta = $constant) with reference to the V − $\theta$ diagram.}
-
-\begin{align*}
-  x &= \ell\sin\phi\cos\theta\\
-  y &= \ell\sin\phi\sin\theta\\
-  \dot{x} &= \ell\cos\phi\cos\theta\dot{\phi} - \ell\sin\phi\sin\theta\dot{\theta}\\
-  \dot{y} &= \ell\cos\phi\sin\theta\dot{\phi} + \ell\sin\phi\cos\theta\dot{\theta}\\
-  T &= \frac{1}{2}m\left(\ell^2\cos^2\phi\dot{\phi}^2 + \ell^2\sin^2\phi\dot{\theta}^2\right)\\
+  T &= \frac{1}{2}mv^2 = \frac{1}{2}m\left(\alpha^2+\ell^2\dot{\theta}^2\right)\\
+  U &= mgy = -mg\ell\cos\theta\\
+  \mathcal{L} &= \frac{1}{2}m\left(\alpha^2+\ell^2\dot{\theta}^2\right) + mg\ell\cos\theta\\
 \end{align*}

### 2015-11-19 comp_model/crystal_phase_old_real.c

@@ -34,14 +34,14 @@ void log_data(double* data, int fd) {
 }
 
 int main() {
-  double dx = 0.03; //m
-  double dy = 0.03; //m
+  double dx = 0.034; //m
+  double dy = 0.034; //m
   double dt = 0.0003; //s
-  double t_final = 500*dt;
+  double t_final = 1000*dt; // Just forced temperature to 1, unphysical!
   double tau = 0.0003;
   double epsilonbar = 0.01;
   double mu = 1.0;
-  double k = 4;
+  double k = 1;
   double delta = 0.02;
   double anisotropy = 4.0;
   double alpha = 0.9;

### 2015-11-19 classical/latex/a7.tex

--- 2015-11-13 classical/latex/a7.tex
@@
 
+\section{Problem One}
+\textbf{Consider a simple plane pendulum consisting of a mass m attached to a string of length l. After the pendulum is set into motion, the length of the string is shortened at a constant rate:}
+\begin{equation}
+  \frac{dl}{dt} = −\alpha
+\end{equation}
+
+\textbf{The suspension point remains fixed. Compute the Lagrangian and Hamiltonian functions. Compare the Hamiltonian and the total energy, and discuss the conservation of energy for the system.}
+
 \section{Problem One: Shrinking Pendulum Hamiltonian}
@@
   y &= -\ell\cos\theta\\
+
+  \dot{x} &= -\alpha\sin\theta + \ell\cos\theta\dot{\theta}\\
+  \dot{y} &= \alpha\cos\theta + \ell\sin\theta\dot{\theta}\\
+  \vspace{1cm}\\
+  T &= \frac{1}{2}mv^2 = \frac{1}{2}m\left(\alpha^2+\ell^2\dot{\theta}^2\right)\\
+  U &= mgy = -mg\ell\cos\theta\\
+  \mathcal{L} &= \frac{1}{2}m\left(\alpha^2+\ell^2\dot{\theta}^2\right) + mg\ell\cos\theta\\
+\end{align*}
+
   \dot{x} &= \dot{\ell}\sin\theta + \ell\cos\theta\dot{\theta}\\

### 2015-11-22 classical/latex/a7.tex
stuff

--- 2015-11-13 classical/latex/a7.tex
@@
 
-\section{Problem Two}
+\section{Problem Two: 3D Pendulum Hamiltonian}
 
@@
 \begin{align*}
+  x &= \ell\cos\phi\sin\theta\\
+  y &= \ell\sin\phi\sin\theta\\
+  z &= \ell\cos\theta\\
+  \dot{x} &=  -\ell\sin\phi\sin\theta\dot{\phi} + \ell\cos\phi\cos\theta\dot{\theta}\\
+  \dot{y} &= \ell\cos\phi\sin\theta\dot{\phi} + \ell\sin\phi\cos\theta\dot{\theta}\\
+  \dot{z} &= -\ell\sin\theta\dot{\theta}\\
+  \vspace{1cm}\\
+  T &= \frac{1}{2}m\left(\dot{x}^2+\dot{y}^2+\dot{z}^2\right)\\
+  &= \frac{1}{2}m\ell^2\left(\sin^2\theta\dot{\phi}^2 + \dot{\theta}^2\right)\\
+  U &= mgz = mg\ell\cos\theta\\
+  \mathcal{L} &= \frac{1}{2}m\ell^2\left(\sin^2\theta\dot{\phi}^2 + \dot{\theta}^2\right) - mg\ell\cos\theta\\
+  \vspace{1cm}\\
+  P_{\theta} &= \frac{d\mathcal{L}}{d\dot{\theta}} = m\ell^2\dot{\theta}\\
+  P_{\phi} &= \frac{d\mathcal{L}}{d\dot{\phi}} = m\ell^2\sin^2(\theta)\dot{\phi}\\
+  \vspace{1cm}\\
+  \vspace{1cm}\\
+  \mathcal{H} &= P_{\theta}\dot{\theta} + P_{\phi}\dot{\phi} - \mathcal{L}\\
+  &= \frac{1}{2}m\ell^2\left(\sin^2\theta\dot{\phi}^2 + \dot{\theta}^2\right) + mg\ell\cos\theta\\
+  &= \frac{1}{2}m\ell^2\left(\sin^2\theta\dot{\phi}^2 + \frac{P_{\theta}^2}{m^2\ell^4}\right) + mg\ell\cos\theta\\
+
+    V_{eff}(\theta, P_{\phi}) &=
+\end{align*}
+
+\begin{figure}[h]
+\caption{Plot of effective potential based on $p_{\phi}$ and gravity potential. Angular momentum linearly adds to the sinusoidal gravity potential.}
+\centering
+\includegraphics[width=0.5\textwidth]{../v-plot_legend.png}
+\end{figure}
+
+\section{Problem Three}
+\textbf{A bead of mass m is threaded on a frictionless wire that is bent into a helix with cylindrical polar coordinates (ρ, ϕ, z) satisfying z = cϕ and ρ = R with c and R constants.
+The z axis points vertically up and gravity vertically down.
+Using ϕ as your generalized coordinate, write down the kinetic and potential energies and the Hamiltonian H as a function of ϕ and its conjugate momentum p.
+Write down Hamilton’s equations and solve for φ!!and hence !z!.
+Explain your result in terms of Newtonian mechanics and discuss the special case that R = 0.}
+
+\end{document}
+
+\begin{align*}
   x &= \ell\sin\phi\cos\theta\\
@@
   \dot{y} &= \ell\cos\phi\sin\theta\dot{\phi} + \ell\sin\phi\cos\theta\dot{\theta}\\
+  \vspace{1cm}\\
   T &= \frac{1}{2}m\left(\ell^2\cos^2\phi\dot{\phi}^2 + \ell^2\sin^2\phi\dot{\theta}^2\right)\\
+  U &= mgz = mg\ell\cos\theta\\
+  \mathcal{L} &= \frac{1}{2}m\left(\ell^2\cos^2\phi\dot{\phi}^2 + \ell^2\sin^2\phi\dot{\theta}^2\right) - mg\ell\cos\theta\\
+  \vspace{1cm}\\
+  \mathcal{H} &= p_{\theta}\dot{\theta} + p_{\phi}\dot{\phi} - \mathcal{L}\\
+  &= m\ell\sin\phi\dot{\theta}^2 + m\ell\cos\phi\dot{\phi}^2 - \frac{1}{2}m\left(\ell^2\cos^2\phi\dot{\phi}^2 + \ell^2\sin^2\phi\dot{\theta}^2\right) + mg\ell\cos\theta\\
+  &= \frac{1}{2}m\left(\ell^2\cos^2\phi\dot{\phi}^2 + \ell^2\sin^2\phi\dot{\theta}^2\right) + mg\ell\cos\theta\\
+  &= \frac{p_{\theta}^2}{2m} + \frac{p_{\phi}^2}{2m} + mg\ell\cos\theta\\
+  V(\theta,p_{\phi}) &= \frac{p_{\phi}^2}{2m} + mg\ell\cos\theta = \frac{1}{2}m\ell^2\cos^2\phi\dot{\phi}^2 + mg\ell\cos\theta\\
 \end{align*}
-
-\end{document}

### 2015-11-22 comp_model/latex/paper1.tex
stuff

@@
 \renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header
+\renewcommand{\baselinestretch}{1.2}
 
@@
 
-\title{Effect of Parameters in Phase-Field Model for simulating Solid-Liquid Transitions}
+\title{Using the Phase-Field model for Creating Anisotropic Crystals}
 \author{Elliott Capek}
@@
 
-The equation for $\epsilon(\theta)$ is given as follows:
-
-\begin{equation}
-  \epsilon = 1 + \delta\cos\left(j\left(\theta_0-\theta\right)\right)
-\end{equation}
-
-where $\delta$ is the strength of anisotropy, $j$ is the harmonic number and $\theta_0$ is the angular offset. The higher $j$ is, the more angular ``modes'' will exist in the system.\\
-
-Temperature is also important to this model, as it decides the value of m, the rate of interface front movement. Temperature's time-evolution is given by:
-
-\begin{equation}  \label{eq:dTdt}
-  \frac{\partial T}{\partial t} = \nabla^2T + k\frac{\partial \phi}{\partial t}
-\end{equation}
-
-where k is the latent heat of the phase change. Liquid freezing into a solid is exergonic, so heat is evolved when $\frac{\partial\phi}{\partial t}$ is positive. \\
-
 \section{Derivation of $\phi$-evolution}
@@
 It decides if it is favorable or unfavorable to be a solid or liquid, and is a function of T.
-It will be defined later.
 Potential depends on m in that the higher m is, the more favorable the liquid phase should be.
@@
 \begin{equation}
-  f(\phi,m) = \frac{1}{4}\phi^4 - (\frac{1}{2} - \frac{1}{3}m)\phi^3 + (\frac{1}{4}-\frac{1}{2}m)\phi^2
+  f(\phi,m) = \frac{1}{4}\phi^4 - \left(\frac{1}{2} - \frac{1}{3}m\right)\phi^3 + \left(\frac{1}{4}-\frac{1}{2}m\right)\phi^2
 \end{equation}
+
+This polynomial meets all the above requirements. See Figure (\ref{fig:f}) for a plot of potential.\\
+
+Thermodynamic driving force m is given by the following equation:
+
+\begin{equation}
+  m = \frac{\alpha \tan^{-1}\left(\gamma\left(T_{eq}-T\right)\right)}{\pi}
+\end{equation}
+
+where $\alpha \leq 1$ and $\gamma$ are positive constants. By this definition, $|m| \leq \frac{1}{2}$.
+
+\begin{figure}[h!]
+  \caption{\textbf{A plot of $f(\phi,m)$ for different values of m.
+Notice that all lines have minima at $\phi=0$ and $\phi=1$.
+The general trend is higher values of m have higher potentials for the solid phase.
+At m=0, there is an equal potential for solid and liquid phase, hence this is the equilibrium m (corresponding to an equilibrium temperature).}}
+  \centering
+  \includegraphics[width=0.5\textwidth]{../f_plot.png}
+  \label{fig:f}
+\end{figure}
 
@@
 Certain angles are cheap, and so the system will have the barrier take those on, and some will be expensive, so the system will try and minimize the area of the interface at that angle.
-In particular, values of $j(\phi_0-\phi)$ close to $-1$ will be favored and values close to $1$ will be disfavored.
+In particular, values of $j(\phi_0-\phi)$ close to $-1$ will be favored and values close to $1$ will be disfavored.\\
 
-The free energy functional is a value that is minimized when a system evolves in time. We can use this fact to time-evolve $\phi$.
+The free energy functional is a value that is minimized when a system evolves in time.
+We can use this fact to time-evolve $\phi$.
+Intuitively we know that, if we want to minimize F by changing $\phi$, we should couple $\frac{\partial\phi}{\partial t}$ to the functional derivative $\frac{\delta F}{\delta \phi}$, or how much free energy is changing with respect to a change in $\phi$.
+Sanal does this as follows:
+
+\begin{equation}
+  \tau \frac{\partial \phi}{\partial t} = \frac{\delta F}{\delta \phi}
+\end{equation}
+
+where $\tau$ is a small positive parameter which controls the proportionality of the two derivatives.
+This functional derivative is difficult to evaluate.
+We can transform this equation into a simpler numerical one like so:
+
+\begin{equation}
+  \tau \frac{d\phi_{ij}}{dt} = \frac{dF_{ij}}{d\phi_{ij}}
+  \label{eq:phi-evol}
+\end{equation}
+
+Becuse we are on a two-dimensional grid, we can compartmentalize $\phi$, so $\phi_{ij}$ is the $(i,j)$ element on our grid.
+Free energy F is an integral over the entire system, but we can compartmentalize it to the following:
+
+\begin{equation}
+  F_{ij}(\phi_{ij}) = f(\phi_{ij},m) + \frac{\epsilon(\phi_{ij})^2}{2}|\nabla\phi_{ij}|^2
+\end{equation}
+
+where we have removed the volume integral on the total F.
+We can see that $F = \sum_i \sum_j F_{ij}$, so minimizing F is equivalent to minimizing $F_{ij}$ for all values of $i,j$.
+We have transformed the functional F into a simple function $F_{ij}$.\\
+
+NOTE: One problem with Equation (\ref{eq:phi-evol}) to me is that $\tau$ should be negative.
+It seems that if $\frac{dF_{ij}}{d\phi_{ij}}$ is positive, then $\frac{d\phi_{ij}}{dt}$ should be negative, so as to make sure $F_{ij}$ is always decreasing as time increases.
+However, this equation seems to predict the opposite.\\
+
+We then take the $\phi$ derivative of the RHS of Equation (\ref{eq:phi-evol}) and get the following:
+
+\begin{equation}
+  \tau \frac{d\phi_{ij}}{dt} = \frac{\partial}{\partial x}\left(\epsilon_{ij}\frac{\partial \epsilon_{ij}}{\partial \phi_{ij}}\frac{\partial \phi_{ij}}{\partial y}\right) + \frac{\partial}{\partial y}\left(\epsilon_{ij}\frac{\partial \epsilon_{ij}}{\partial \phi}\frac{\partial \phi_{ij}}{\partial x}\right) + \epsilon_{ij}^2\nabla^2\phi_{ij} + \phi_{ij}\left(1-\phi_{ij}\right)\left(\phi_{ij}-0.5+m\right)
+\end{equation}
+
+The fourth term is a rearrangement of $-\frac{\partial f}{\partial\phi}$.
+The first three terms are part of the product rule derivation of $\frac{\epsilon^2}{2}|\nabla\phi|^2$.
+This equation is in a form that we can evaluate numerically.
+
+\section{Derivation of T-evolution}
+
+Temperature evolution is done with the following equation:
+
+\begin{equation}
+  \frac{\partial T}{\partial t} = \nabla^2 T + K\frac{\partial\phi}{\partial t}
+\end{equation}
+
+The first term is the normal heat equation, where heat is passed from high-temperature elements to low-temperature elements.
+The second term is the heat evolved from the transition from liquid to solid.
+K is the latent heat, which describes how much heat is released at this transition.
+
+\section{Simulation}
+Now that we have our time-evolution equations, it is a simple matter to solve for $\phi$ over time.
+We do this by using the forwards-Euler method on equations (11) and (12), which requires us to compute spacial derivatives of $\frac{\partial}{\partial x}\left(\epsilon_{ij}\frac{\partial \epsilon_{ij}}{\partial \phi_{ij}}\frac{\partial \phi_{ij}}{\partial y}\right)$ and $\frac{\partial}{\partial y}\left(\epsilon_{ij}\frac{\partial \epsilon_{ij}}{\partial \phi}\frac{\partial \phi_{ij}}{\partial x}\right)$, as well as the Laplacian of both $\epsilon$ and $\phi$.
+We compute the spacial derivatives of the functions by taking the difference between the values at neighboring points (using periodic boundary conditions).
+We use a nine-point Laplacian using the considered point and the eight points adjacent to it, weighting the adjacent points $+\frac{2}{3}$ and the considered point $-4$.
+
+\section{Reference}
+        \hspace{2cm}[1] Sanal, Rahul.
+Numerical Simulation of Dendritic crystal growth using phase field method and investigating the effects of different physical parameter on the growth of the dendrite.
+Arxiv.<arxiv.org/ftp/arxiv/papers/1412/1412.3197.pdf>\\
 

### 2015-11-22 physical_chemistry/latex/pchem_tips.tex
stuff

--- 2015-11-06 classical/latex/a6.tex
@@
 
+\usepackage{amsmath}
 \usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins
-
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{marvosym} % Required for symbols in the colored box
-\usepackage{ifsym} % Required for symbols in the colored box
 
@@
 
-\title{Classical Mechanics Assignment 6}
-\author{Elliott Capek}
-\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles
-
+\title{Physical Chemistry Tips}
 \begin{document}
@@
 
-\section{Problem One: Ball on an Inclined Plane}
-\textbf{Write the Lagrangian for a cylinder (mass m, radius R, and moment of inertia I) that rolls without slipping straight down an inclined plane at an angle $\alpha$ from the horizontal.
-Use as your generalized coordinates the cylinder’s distance x measured down the plane from its starting point.
-Solve the Lagrange equation for the cylinder’s acceleration $\ddot{x}$.
-Remember that $T = \frac{1}{2} mv^2 + \frac{1}{2} I\omega^2$ where $v$ is the velocity of the center of mass and $\omega$ is the cylinder’s angular velocity.}
+\section{Random relationships}
+
+\subsection{Weird unknown axioms}
 
 \begin{align*}
-  x' &= x\cos(\alpha)\\
-  y' &= -x\sin(\alpha)\\
-  \omega &= \frac{v}{2\pi R}\\
-  \vspace{1cm}\\
-  U &= mgy' = -mgx\sin(\alpha)\\
-  T &= \frac{1}{2}mv^2 + \frac{1}{2}I\omega^2\\
-  &= \frac{1}{2}m\left(\dot{x}'^2 + \dot{y}'^2\right) + \frac{1}{2}I\left(\frac{\dot{x}'^2 + \dot{y}'^2}{2\pi R}\right)\\
-  &= \frac{1}{2}m\left(\dot{x}^2\sin^2(\alpha) + \dot{x}^2\cos^2(\alpha)\right) + \frac{1}{2}I\left(\frac{\dot{x}^2\sin^2(\alpha) + \dot{x}^2\cos^2(\alpha)}{4\pi^2R^2}\right)\\
-  &= m\left(\dot{x}^2\right) + I\left(\frac{\dot{x}^2}{4\pi^2R^2}\right)\\
-  &= \dot{x}^2\left(\frac{I + 4\pi^2R^2m}{4\pi^2R^2}\right)\\
-  L &= \dot{x}^2\left(\frac{I + 4\pi^2R^2m}{4\pi^2R^2}\right) + mgx\sin(\alpha)\\
-  \vspace{1cm}\\
-  &\frac{dL}{dx} - \frac{d}{dt}\frac{dL}{d\dot{x}} = 0\\
-  &mg\sin(\alpha) - \ddot{x}\left(\frac{I + 4\pi^2R^2m}{\pi^2R^2}\right) = 0\\
-  \ddot{x} &= \frac{\pi^2R^2mg\sin(\alpha)}{I + 4\pi^2R^2m}\\
-  x &= \frac{\pi^2R^2mg\sin(\alpha)}{2I + 8\pi^2R^2m}t^2
+  \frac{dH}{dP}_t &= -T\frac{dV}{dT}_P + V\\
 \end{align*}
 
-\section{Problem Two: Bead on a Spiral}
-\textbf{A smooth wire is bent into the shape of a helix, with cylindrical coordinates $\rho=R$ and $z=\lambda\phi$, where $R$ and $\lambda$ are constants and the z axis is vertically up (and gravity is vertically down).
-Using z as your generalized coordinate, write down the Lagrangian for a bead of mass m threaded on the wire.
-Find the Lagrange equation and the bead’s vertical acceleration $\ddot{z}$.
-In the limit that $R → 0$, what is $\ddot{z}$?
-Does this make sense?}
-
 \begin{align*}
-  v &= \sqrt{\dot{\rho}^2 + \rho^2\dot{\phi}^2 + \dot{z}^2}\\
-  &= \sqrt{\rho^2\frac{\dot{z}^2}{\lambda^2} + \dot{z}^2}\\
-  \vspace{1cm}\\
-  T &= \frac{1}{2}mv^2\\
-  &= \frac{1}{2}m\left(\rho^2\frac{\dot{z}^2}{\lambda^2} + \dot{z}^2\right)\\
-  U &= mgz\\
-  L &= \frac{1}{2}m\left(\rho^2\frac{\dot{z}^2}{\lambda^2} + \dot{z}^2\right) - mgz\\
-  \vspace{1cm}\\
-  &\frac{dL}{dz} - \frac{d}{dt}\frac{dL}{d\dot{z}} = 0\\
-  -mg - \ddot{z}m\left(\frac{\rho^2}{\lambda^2}+1\right) &= 0\\
-  \ddot{z} &= \frac{-g}{\frac{\rho^2}{\lambda^2}+1}\\
+  \frac{dQ}{T} &= dS\\
+  TdS &= (dH)_p = C_pdT\\
+  \frac{dS}{dT} &=\frac{C_p}{T}\\
 \end{align*}
 
-\section{Problem Three: }
-\textbf{Write the Lagrangian for a cylinder (mass m, radius r, and moment of inertia $I=\frac{mr^2}{2}$) that rolls without slipping in the bottom of a circular shaped hole of equation $z = −\sqrt{R^2 − x^2}$ , where z is the vertical axis (see figure below).
-Find the angular frequency of small oscillations about the bottom of the hole.}
-
 \end{document}

### 2015-11-23 physical_chemistry/latex/a4.tex
New a4.tex.

--- 2015-11-06 classical/latex/a6.tex
@@
 
-\title{Classical Mechanics Assignment 6}
+\title{Physical Chemistry Assignment Four}
 \author{Elliott Capek}
-\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles
-
 \begin{document}
@@
 
-\section{Problem One: Ball on an Inclined Plane}
-\textbf{Write the Lagrangian for a cylinder (mass m, radius R, and moment of inertia I) that rolls without slipping straight down an inclined plane at an angle $\alpha$ from the horizontal.
-Use as your generalized coordinates the cylinder’s distance x measured down the plane from its starting point.
-Solve the Lagrange equation for the cylinder’s acceleration $\ddot{x}$.
-Remember that $T = \frac{1}{2} mv^2 + \frac{1}{2} I\omega^2$ where $v$ is the velocity of the center of mass and $\omega$ is the cylinder’s angular velocity.}
+\section{4.10}
 
+a.)
 \begin{align*}
-  x' &= x\cos(\alpha)\\
-  y' &= -x\sin(\alpha)\\
-  \omega &= \frac{v}{2\pi R}\\
-  \vspace{1cm}\\
-  U &= mgy' = -mgx\sin(\alpha)\\
-  T &= \frac{1}{2}mv^2 + \frac{1}{2}I\omega^2\\
-  &= \frac{1}{2}m\left(\dot{x}'^2 + \dot{y}'^2\right) + \frac{1}{2}I\left(\frac{\dot{x}'^2 + \dot{y}'^2}{2\pi R}\right)\\
-  &= \frac{1}{2}m\left(\dot{x}^2\sin^2(\alpha) + \dot{x}^2\cos^2(\alpha)\right) + \frac{1}{2}I\left(\frac{\dot{x}^2\sin^2(\alpha) + \dot{x}^2\cos^2(\alpha)}{4\pi^2R^2}\right)\\
-  &= m\left(\dot{x}^2\right) + I\left(\frac{\dot{x}^2}{4\pi^2R^2}\right)\\
-  &= \dot{x}^2\left(\frac{I + 4\pi^2R^2m}{4\pi^2R^2}\right)\\
-  L &= \dot{x}^2\left(\frac{I + 4\pi^2R^2m}{4\pi^2R^2}\right) + mgx\sin(\alpha)\\
-  \vspace{1cm}\\
-  &\frac{dL}{dx} - \frac{d}{dt}\frac{dL}{d\dot{x}} = 0\\
-  &mg\sin(\alpha) - \ddot{x}\left(\frac{I + 4\pi^2R^2m}{\pi^2R^2}\right) = 0\\
-  \ddot{x} &= \frac{\pi^2R^2mg\sin(\alpha)}{I + 4\pi^2R^2m}\\
-  x &= \frac{\pi^2R^2mg\sin(\alpha)}{2I + 8\pi^2R^2m}t^2
+  \Delta H &= -T^2\frac{\partial (\Delta G/T)}{\partial T}\\
+  -\frac{\Delta H}{T^2} &= \frac{\partial (\Delta G/T)}{\partial T}\\
+  \int_{T_1}^{T_2} -\frac{\Delta H}{T^2}dT &= \frac{\Delta G}{T}\Bigg|_{\frac{\Delta G_1}{T_1}}^{\frac{\Delta G_2}{T_2}}\\
+  \Delta H\left(\frac{1}{T_2} - \frac{1}{T_1}\right) &= \frac{\Delta G_2}{T_2} - \frac{\Delta G_1}{T_1}\\
+  \Delta G_2 &= \Delta H\left(1 - \frac{T_2}{T_1}\right) + \Delta G_1 \frac{T_2}{T_1}\\
 \end{align*}
 
-\section{Problem Two: Bead on a Spiral}
-\textbf{A smooth wire is bent into the shape of a helix, with cylindrical coordinates $\rho=R$ and $z=\lambda\phi$, where $R$ and $\lambda$ are constants and the z axis is vertically up (and gravity is vertically down).
-Using z as your generalized coordinate, write down the Lagrangian for a bead of mass m threaded on the wire.
-Find the Lagrange equation and the bead’s vertical acceleration $\ddot{z}$.
-In the limit that $R → 0$, what is $\ddot{z}$?
-Does this make sense?}
-
+b.)
 \begin{align*}
-  v &= \sqrt{\dot{\rho}^2 + \rho^2\dot{\phi}^2 + \dot{z}^2}\\
-  &= \sqrt{\rho^2\frac{\dot{z}^2}{\lambda^2} + \dot{z}^2}\\
-  \vspace{1cm}\\
-  T &= \frac{1}{2}mv^2\\
-  &= \frac{1}{2}m\left(\rho^2\frac{\dot{z}^2}{\lambda^2} + \dot{z}^2\right)\\
-  U &= mgz\\
-  L &= \frac{1}{2}m\left(\rho^2\frac{\dot{z}^2}{\lambda^2} + \dot{z}^2\right) - mgz\\
-  \vspace{1cm}\\
-  &\frac{dL}{dz} - \frac{d}{dt}\frac{dL}{d\dot{z}} = 0\\
-  -mg - \ddot{z}m\left(\frac{\rho^2}{\lambda^2}+1\right) &= 0\\
-  \ddot{z} &= \frac{-g}{\frac{\rho^2}{\lambda^2}+1}\\
+  \Delta H &= -T^2\frac{\partial (\Delta G/T)}{\partial T}\\
+  -\frac{\Delta H}{T^2} &= \frac{\partial (\Delta G/T)}{\partial T}\\
+  \int_{T_1}^{T_2} &\left(-\frac{\Delta H_1 + (T-T_1)\Delta C_P}{T^2}\right)dT = \frac{\Delta G}{T}\Bigg|_{\frac{\Delta G_1}{T_1}}^{\frac{\Delta G_2}{T_2}}\\
+  &\Delta H_1\left(\frac{1}{T_2}-\frac{1}{T_1}\right) + \ln\left(\frac{T_1}{T_2}\right)\Delta C_P - T_1\Delta C_P\left(\frac{1}{T_2}-\frac{1}{T_1}\right) = \frac{\Delta G_2}{T_2} - \frac{\Delta G_1}{T_1}\\
+  \Delta G_2 &= \left(\Delta H_1 - T_1\Delta C_P\right)\left(1 - \frac{T_2}{T_1}\right) + T_2\ln\left(\frac{T_1}{T_2}\right)\Delta C_P + \frac{\Delta G_1}{T_1}\\
 \end{align*}
 
-\section{Problem Three: }
-\textbf{Write the Lagrangian for a cylinder (mass m, radius r, and moment of inertia $I=\frac{mr^2}{2}$) that rolls without slipping in the bottom of a circular shaped hole of equation $z = −\sqrt{R^2 − x^2}$ , where z is the vertical axis (see figure below).
-Find the angular frequency of small oscillations about the bottom of the hole.}
+\section{4.11}
+\begin{align*}
+  \bar{V} &= 1 \left(\frac{1 m}{10^2 g}\right)^3 \frac{18.01 g}{1 mol} = 1.801 * 10^{-5} \frac{m^3}{mol}\\
+  \Delta\bar{G} &= \left(1.801 * 10^{-5} \frac{m^3}{mol}\right)\left(10^3 bar * 10^5 \frac{Pa}{bar}\right) = 1.8 \frac{kJ}{mol}\\
+\end{align*}
+
+\section{4.12}
+\begin{align*}
+  \Delta \bar{G} &= -\bar{S}dT + \bar{V}dP\\
+  \Delta \bar{G} &= RT\ln\left(\frac{P_2}{P_1}\right)\\
+  &= (8.314 \frac{J}{K mol})(298K)\ln(0.1) = -5.7 \frac{kJ}{mol}\\
+\end{align*}
+
+G is a state function, so all systems with the same initial and final states have the same $\Delta G$.
+
+\section{4.15}
+\begin{align*}
+  \Delta H_{vap} &= 33.2 \frac{kJ}{mol}\\
+  w &= -P_{ext}\Delta V = -(10^5 Pa)\frac{(1 mol)(8.314 \frac{J}{K mol} * 383 K)}{10^5 Pa}\\
+  &= -3.184 \frac{kJ}{mol}\\
+  q &= dH_P = 33.2 \frac{kJ}{mol}\\
+  \Delta \bar{H} &= 33.2 \frac{kJ}{mol}\\
+  \Delta \bar{U} &= -3.184 \frac{kJ}{mol} + 33.2 \frac{kJ}{mol} = 30.02 \frac{kJ}{mol}\\
+  \Delta \bar{G} &= 0\\
+  \Delta \bar{S} &= \frac{q}{T} = 86.6 \frac{J}{mol K}\\
+\end{align*}
 

### 2015-11-23 classical/latex/a8.tex
Add stuffs.

--- 2015-11-06 classical/latex/a6.tex
@@
 
-\title{Classical Mechanics Assignment 6}
+\title{Classical Mechanics Assignment 8}
 \author{Elliott Capek}
@@
 
-\section{Problem One: Ball on an Inclined Plane}
-\textbf{Write the Lagrangian for a cylinder (mass m, radius R, and moment of inertia I) that rolls without slipping straight down an inclined plane at an angle $\alpha$ from the horizontal.
-Use as your generalized coordinates the cylinder’s distance x measured down the plane from its starting point.
-Solve the Lagrange equation for the cylinder’s acceleration $\ddot{x}$.
-Remember that $T = \frac{1}{2} mv^2 + \frac{1}{2} I\omega^2$ where $v$ is the velocity of the center of mass and $\omega$ is the cylinder’s angular velocity.}
+\section{Problem One: }
+\textbf{A pendulum consists of a mass m suspended by a massless spring with un-extended length $b$ and spring constant $k$. Find Hamilton’s equations of motion.}
 
-\begin{align*}
-  x' &= x\cos(\alpha)\\
-  y' &= -x\sin(\alpha)\\
-  \omega &= \frac{v}{2\pi R}\\
-  \vspace{1cm}\\
-  U &= mgy' = -mgx\sin(\alpha)\\
-  T &= \frac{1}{2}mv^2 + \frac{1}{2}I\omega^2\\
-  &= \frac{1}{2}m\left(\dot{x}'^2 + \dot{y}'^2\right) + \frac{1}{2}I\left(\frac{\dot{x}'^2 + \dot{y}'^2}{2\pi R}\right)\\
-  &= \frac{1}{2}m\left(\dot{x}^2\sin^2(\alpha) + \dot{x}^2\cos^2(\alpha)\right) + \frac{1}{2}I\left(\frac{\dot{x}^2\sin^2(\alpha) + \dot{x}^2\cos^2(\alpha)}{4\pi^2R^2}\right)\\
-  &= m\left(\dot{x}^2\right) + I\left(\frac{\dot{x}^2}{4\pi^2R^2}\right)\\
-  &= \dot{x}^2\left(\frac{I + 4\pi^2R^2m}{4\pi^2R^2}\right)\\
-  L &= \dot{x}^2\left(\frac{I + 4\pi^2R^2m}{4\pi^2R^2}\right) + mgx\sin(\alpha)\\
-  \vspace{1cm}\\
-  &\frac{dL}{dx} - \frac{d}{dt}\frac{dL}{d\dot{x}} = 0\\
-  &mg\sin(\alpha) - \ddot{x}\left(\frac{I + 4\pi^2R^2m}{\pi^2R^2}\right) = 0\\
-  \ddot{x} &= \frac{\pi^2R^2mg\sin(\alpha)}{I + 4\pi^2R^2m}\\
-  x &= \frac{\pi^2R^2mg\sin(\alpha)}{2I + 8\pi^2R^2m}t^2
-\end{align*}
+\section{Problem Two: }
+\textbf{The potential for an non-harmonic oscillator is:}
 
-\section{Problem Two: Bead on a Spiral}
-\textbf{A smooth wire is bent into the shape of a helix, with cylindrical coordinates $\rho=R$ and $z=\lambda\phi$, where $R$ and $\lambda$ are constants and the z axis is vertically up (and gravity is vertically down).
-Using z as your generalized coordinate, write down the Lagrangian for a bead of mass m threaded on the wire.
-Find the Lagrange equation and the bead’s vertical acceleration $\ddot{z}$.
-In the limit that $R → 0$, what is $\ddot{z}$?
-Does this make sense?}
+\begin{equation*}
+  U(x) = \frac{kx^2}{2} + \frac{bx^4}{4}\\
+\end{equation*}
 
-\begin{align*}
-  v &= \sqrt{\dot{\rho}^2 + \rho^2\dot{\phi}^2 + \dot{z}^2}\\
-  &= \sqrt{\rho^2\frac{\dot{z}^2}{\lambda^2} + \dot{z}^2}\\
-  \vspace{1cm}\\
-  T &= \frac{1}{2}mv^2\\
-  &= \frac{1}{2}m\left(\rho^2\frac{\dot{z}^2}{\lambda^2} + \dot{z}^2\right)\\
-  U &= mgz\\
-  L &= \frac{1}{2}m\left(\rho^2\frac{\dot{z}^2}{\lambda^2} + \dot{z}^2\right) - mgz\\
-  \vspace{1cm}\\
-  &\frac{dL}{dz} - \frac{d}{dt}\frac{dL}{d\dot{z}} = 0\\
-  -mg - \ddot{z}m\left(\frac{\rho^2}{\lambda^2}+1\right) &= 0\\
-  \ddot{z} &= \frac{-g}{\frac{\rho^2}{\lambda^2}+1}\\
-\end{align*}
+\textbf{where k and b are positive constants. Find Hamilton’s equations of motion. No final discussion needed.}
 
 \section{Problem Three: }
-\textbf{Write the Lagrangian for a cylinder (mass m, radius r, and moment of inertia $I=\frac{mr^2}{2}$) that rolls without slipping in the bottom of a circular shaped hole of equation $z = −\sqrt{R^2 − x^2}$ , where z is the vertical axis (see figure below).
-Find the angular frequency of small oscillations about the bottom of the hole.}
+\textbf{Reconsider the problem of the two masses and three springs (seen in class) in the event that the three springs all have different force constants (but the masses are equal).
+Find the two characteristic frequencies and compare the magnitudes with the natural frequencies of the two oscillators in the absence of coupling.}
+
+\section{Problem Four: }
+\textbf{A thin hoop of radius R and mass M oscillates in its own plane hanging from a single fixed point.
+Attached to the hoop is a small mass M constrained to move (in a frictionless manner) along the hoop.
+The inertia of the loop for rotation around a point on the loop is $I=2MR^2$.
+Consider only small oscillations, and show that the eigenfrequencies are:}
+
+\begin{equation*}
+  \omega_1 = \sqrt{\frac{2g}{R}}
+\end{equation*}
+
+\begin{equation*}
+  \omega_2 = \sqrt{\frac{g}{2R}}
+\end{equation*}
 

### 2015-11-28 comp_model/crystal_phase.c
cmmitting changies.

@@ -37,7 +37,7 @@ int main() {
   double dx = 0.03; //m
   double dy = 0.03; //m
   double dt = 0.0003; //s
-  double t_final = 500*dt;
+  double t_final = 1500*dt;
   double tau = 0.0003;
   double epsilonbar = 0.01;
   double mu = 1.0;
@@ -48,7 +48,7 @@ int main() {
   double gamma = 10.0;
   double Teq = 1.0;
   double T0 = 0;
-  int r = 70;
+  int r = 50;
 
   int fd;
   if ((fd = open("data.txt", O_WRONLY | O_CREAT | O_TRUNC, S_IRUSR | S_IWUSR | S_IRGRP | S_IWGRP | S_IWOTH | S_IROTH)) == -1)
@@ -139,7 +139,7 @@ int main() {
 	dep2_dx = (ep[xp+y*size]*ep[xp+y*size] - ep[xm+y*size]*ep[xm+y*size]) / dx;
 	dep2_dy = (ep[x+yp*size]*ep[x+yp*size] - ep[x+ym*size]*ep[x+ym*size]) / dy;
 
-	term1 = (ep[xp+y*size]*dep_dt[xp+y*size]*dphi_dy[xp+y*size]
+	term1 = -(ep[xp+y*size]*dep_dt[xp+y*size]*dphi_dy[xp+y*size]
 		 - ep[xm+y*size]*dep_dt[xm+y*size]*dphi_dy[xm+y*size]) / dx;
 	
 	term2 = (ep[x+yp*size]*dep_dt[x+yp*size]*dphi_dx[x+yp*size]

### 2015-11-28 comp_model/crystal_phase.c
Stuff

@@ -7,7 +7,7 @@
 #include <unistd.h>
 
 const double m_pi =  3.14159265358979323846;
-const int size = 300;
+const int size = 500;
 const int animate = 1;
 const int skiprate = 10;
 const int debug = 0;
@@ -34,21 +34,21 @@ void log_data(double* data, int fd) {
 }
 
 int main() {
-  double dx = 0.034; //m
-  double dy = 0.034; //m
+  double dx = 0.03; //m
+  double dy = 0.03; //m
   double dt = 0.0003; //s
-  double t_final = 1000*dt; // Just forced temperature to 1, unphysical!
+  double t_final = 2000*dt; // Just forced temperature to 1, unphysical!
   double tau = 0.0003;
   double epsilonbar = 0.01;
   double mu = 1.0;
   double k = 1;
-  double delta = 0.02;
+  double delta = 0.01;
   double anisotropy = 4.0;
   double alpha = 0.9;
   double gamma = 10.0;
   double Teq = 1.0;
   double T0 = 0;
-  int r = 70;
+  int r = 50;
 
   int fd;
   if ((fd = open("data.txt", O_WRONLY | O_CREAT | O_TRUNC, S_IRUSR | S_IWUSR | S_IRGRP | S_IWGRP | S_IWOTH | S_IROTH)) == -1)
@@ -139,7 +139,7 @@ int main() {
 	dep2_dx = (ep[xp+y*size]*ep[xp+y*size] - ep[xm+y*size]*ep[xm+y*size]) / dx;
 	dep2_dy = (ep[x+yp*size]*ep[x+yp*size] - ep[x+ym*size]*ep[x+ym*size]) / dy;
 
-	term1 = (ep[xp+y*size]*dep_dt[xp+y*size]*dphi_dy[xp+y*size]
+	term1 = -(ep[xp+y*size]*dep_dt[xp+y*size]*dphi_dy[xp+y*size]
 		 - ep[xm+y*size]*dep_dt[xm+y*size]*dphi_dy[xm+y*size]) / dx;
 	
 	term2 = (ep[x+yp*size]*dep_dt[x+yp*size]*dphi_dx[x+yp*size]

### 2015-11-28 comp_model/latex/der.tex
stuff

--- 2015-11-23 physical_chemistry/latex/a4.tex
@@
 
-\title{Physical Chemistry Assignment Four}
+\title{Free energy derivative}
 \author{Elliott Capek}
+
 \begin{document}
-
 \maketitle{}
 
-\section{4.10}
+\section{Introduction}
+Doing the derivation of free energy:
 
-a.)
 \begin{align*}
-  \Delta H &= -T^2\frac{\partial (\Delta G/T)}{\partial T}\\
-  -\frac{\Delta H}{T^2} &= \frac{\partial (\Delta G/T)}{\partial T}\\
-  \int_{T_1}^{T_2} -\frac{\Delta H}{T^2}dT &= \frac{\Delta G}{T}\Bigg|_{\frac{\Delta G_1}{T_1}}^{\frac{\Delta G_2}{T_2}}\\
-  \Delta H\left(\frac{1}{T_2} - \frac{1}{T_1}\right) &= \frac{\Delta G_2}{T_2} - \frac{\Delta G_1}{T_1}\\
-  \Delta G_2 &= \Delta H\left(1 - \frac{T_2}{T_1}\right) + \Delta G_1 \frac{T_2}{T_1}\\
+  G &= V(\phi(r)) + \frac{\delta^2\left(\nabla\phi(r)\right)^2}{2}
 \end{align*}
 
-b.)
 \begin{align*}
-  \Delta H &= -T^2\frac{\partial (\Delta G/T)}{\partial T}\\
-  -\frac{\Delta H}{T^2} &= \frac{\partial (\Delta G/T)}{\partial T}\\
-  \int_{T_1}^{T_2} &\left(-\frac{\Delta H_1 + (T-T_1)\Delta C_P}{T^2}\right)dT = \frac{\Delta G}{T}\Bigg|_{\frac{\Delta G_1}{T_1}}^{\frac{\Delta G_2}{T_2}}\\
-  &\Delta H_1\left(\frac{1}{T_2}-\frac{1}{T_1}\right) + \ln\left(\frac{T_1}{T_2}\right)\Delta C_P - T_1\Delta C_P\left(\frac{1}{T_2}-\frac{1}{T_1}\right) = \frac{\Delta G_2}{T_2} - \frac{\Delta G_1}{T_1}\\
-  \Delta G_2 &= \left(\Delta H_1 - T_1\Delta C_P\right)\left(1 - \frac{T_2}{T_1}\right) + T_2\ln\left(\frac{T_1}{T_2}\right)\Delta C_P + \frac{\Delta G_1}{T_1}\\
-\end{align*}
-
-\section{4.11}
-\begin{align*}
-  \bar{V} &= 1 \left(\frac{1 m}{10^2 g}\right)^3 \frac{18.01 g}{1 mol} = 1.801 * 10^{-5} \frac{m^3}{mol}\\
-  \Delta\bar{G} &= \left(1.801 * 10^{-5} \frac{m^3}{mol}\right)\left(10^3 bar * 10^5 \frac{Pa}{bar}\right) = 1.8 \frac{kJ}{mol}\\
-\end{align*}
-
-\section{4.12}
-\begin{align*}
-  \Delta \bar{G} &= -\bar{S}dT + \bar{V}dP\\
-  \Delta \bar{G} &= RT\ln\left(\frac{P_2}{P_1}\right)\\
-  &= (8.314 \frac{J}{K mol})(298K)\ln(0.1) = -5.7 \frac{kJ}{mol}\\
-\end{align*}
-
-G is a state function, so all systems with the same initial and final states have the same $\Delta G$.
-
-\section{4.15}
-\begin{align*}
-  \Delta H_{vap} &= 33.2 \frac{kJ}{mol}\\
-  w &= -P_{ext}\Delta V = -(10^5 Pa)\frac{(1 mol)(8.314 \frac{J}{K mol} * 383 K)}{10^5 Pa}\\
-  &= -3.184 \frac{kJ}{mol}\\
-  q &= dH_P = 33.2 \frac{kJ}{mol}\\
-  \Delta \bar{H} &= 33.2 \frac{kJ}{mol}\\
-  \Delta \bar{U} &= -3.184 \frac{kJ}{mol} + 33.2 \frac{kJ}{mol} = 30.02 \frac{kJ}{mol}\\
-  \Delta \bar{G} &= 0\\
-  \Delta \bar{S} &= \frac{q}{T} = 86.6 \frac{J}{mol K}\\
+  \frac{dG}{d\phi} &= \frac{d}{d\phi}V(\phi(r)) + \frac{d}{d\phi}\frac{\delta^2\left(\nabla\phi(r)\right)^2}{2}\\
+  &= -\phi(1-\phi)(\phi-\frac{1}{2}+n) + \frac{d}{d\phi}\frac{\delta^2\left(\nabla\phi(r)\right)^2}{2}\\
 \end{align*}

### 2015-11-28 classical/latex/a8.tex

@@
 
-\section{Problem One: }
+\section{Problem One: Springy Pendulum}
 \textbf{A pendulum consists of a mass m suspended by a massless spring with un-extended length $b$ and spring constant $k$. Find Hamilton’s equations of motion.}
 
-\section{Problem Two: }
+\begin{align*}
+  x &= (b+\ell)\sin\theta\\
+  y &= -(b+\ell)\cos\theta\\
+  \dot{x} &= \dot{\ell}\sin\theta + (b+\ell)\cos\theta\dot{\theta}\\
+  \dot{y} &= -\dot{\ell}\cos\theta + (b+\ell)\sin\theta\dot{\theta}\\
+  \vspace{1cm}\\
+  T &= \frac{1}{2}m\left(\dot{x}^2 + \dot{y}^2\right)\\
+  &= \frac{1}{2}m\left(\dot{\ell}^2 + (b+\ell)^2\dot{\theta}^2\right)\\
+  U &= mgy = -mg(b + \ell)\cos\theta + \frac{1}{2}k\ell^2\\
+  \mathcal{L} &= \frac{1}{2}m\left(\dot{\ell}^2 + (b+\ell)^2\dot{\theta}^2\right) + mg(b + \ell)\cos\theta - \frac{1}{2}k\ell^2\\
+  \vspace{1cm}\\
+  p_{\ell} &= \frac{d\mathcal{L}}{d\dot{\ell}} = m\dot{\ell}\\
+  p_{\theta} &= \frac{d\mathcal{L}}{d\dot{\theta}} = m(b + \ell)\dot{\theta}\\
+  \mathcal{H} &= m\dot{\ell}^2 + m(b + \ell)\dot{\theta}^2 - \frac{1}{2}m\left(\dot{\ell}^2 + (b+\ell)^2\dot{\theta}^2\right) - mg(b + \ell)\cos\theta + \frac{1}{2}k\ell^2\\
+  &= \frac{p_{\ell}^2}{m} + \frac{p_{\theta}^2}{m(b + \ell)} - \frac{1}{2}m\left(\frac{p_{\ell}^2}{m^2} + \frac{p_{\theta}^2}{m^2}\right) - mg(b + \ell)\cos\theta + \frac{1}{2}k\ell^2\\
+  \vspace{1cm}\\
+  \dot{p_{\theta}} &= -\frac{d\mathcal{H}}{d\theta}\\
+  &= m\ddot{\theta} = -mg(b + \ell)\sin\theta\\
+  \dot{p_{\ell}} &= -\frac{d\mathcal{H}}{d\ell}\\
+  &= m\ddot{\ell} = -m\dot{\theta}^2 + m\left(b+\ell\right)\dot{\theta}^2 + mg\cos\theta + 2k\ell\\
+\end{align*}
+
+This was an interesting usage of Hamiltonian mechanics to solve a very weird physical system.
+This would have been a difficult problem to solve using Newtonian mechanics.
+It probably would have been easier to solve with Lagrangian mechanics due to there being fewer steps, but Hamiltonian worked well.
+The end result is a set of coupled motion equations with no obvious solution, which is in line with what we were told about in class.
+
+\section{Problem Two: Non-harmonic Oscillator}
 \textbf{The potential for an non-harmonic oscillator is:}
@@
 
-\section{Problem Three: }
+\begin{align*}
+  T &= \frac{1}{2}m\dot{x}^2\\
+  \mathcal{L} &= \frac{1}{2}m\dot{x}^2 - \frac{kx^2}{2} - \frac{bx^4}{4}\\
+  \vspace{1cm}
+  p &= \frac{d\mathcal{L}}{d\dot{x}} = m\dot{x}\\
+  \mathcal{H} &= m\dot{x} - \frac{1}{2}m\dot{x}^2 + \frac{kx^2}{2} + \frac{bx^4}{4}\\
+  \dot{p} &= -\frac{d\mathcal{H}}{dx} = -kx - bx^3\\
+\end{align*}
+
+We combine this equation with our knowledge of the definition of momentum $p = m\dot{x}$ to find the Hamilton Equation of Motion:
+
+\begin{align*}
+  m\ddot{x} &= -kx - bx^3\\
+\end{align*}
+
+\section{Problem Three: Two masses Three springs}
 \textbf{Reconsider the problem of the two masses and three springs (seen in class) in the event that the three springs all have different force constants (but the masses are equal).
 Find the two characteristic frequencies and compare the magnitudes with the natural frequencies of the two oscillators in the absence of coupling.}
+
+\begin{align*}
+  T &= \frac{1}{2}m\dot{x}^2 + \frac{1}{2}m\dot{y}^2\\
+  U &= \frac{1}{2}k_1x^2 + \frac{1}{2}k_3y^2 + \frac{1}{2}k_2\left(y - x\right)^2\\
+  \mathcal{L} &= \frac{1}{2}\left(k_1x^2 + k_3y^2 + k_2\left(y-x\right)^2 + m\dot{x}^2 + m\dot{y}^2\right)\\
+\end{align*}
+
+By solving the Euler-Lagrange equation for this Lagrangian we find the following motion equations:
+
+\begin{align*}
+  m\ddot{x} &= k_1x - k_2\left(y-x\right)\\
+  m\ddot{y} &= k_3y + k_2\left(y-x\right)\\
+\end{align*}
+
+We transfer these motion equations to matrix form and solve their determinant:
+
+\begin{align*}
+  det
+  \begin{pmatrix}
+    k_1+k_2-\omega^2m & -k_2\\
+    -k_2 & k_2 + k_3-\omega^2m\\
+  \end{pmatrix}
+  = 0\\
+  \left(k_1+k_2-\omega^2m\right)\left(k_2 + k_3 - \omega^2m\right) - k_2^2 = 0\\
+  &\omega^4m^2 + \omega^2m\left(-k_1-2k_2-k_3\right) + \left(k_1k_2 + k_1k_3 + k_2k_3\right) = 0\\
+  \omega^2 &= \frac{m\left(k_1+2k_2+k_3\right) \pm \sqrt{m^2\left(k_1+2k_2+k_3\right)^2 - 4m^2\left(k_1k_2+k_1k_3+k_2k_3\right)}}{2m^2}\\
+  \omega_1 &= \sqrt{\frac{m\left(k_1+2k_2+k_3\right) + \sqrt{m^2\left(k_1+2k_2+k_3\right)^2 - 4m^2\left(k_1k_2+k_1k_3+k_2k_3\right)}}{2m^2}}\\
+  &= \sqrt{\frac{\left(k_1+2k_2+k_3\right) + \sqrt{k_1^2+4k_2^2+k_3^2-2k_1k_3}}{2m}}\\
+  \omega_2 &= \sqrt{\frac{m\left(k_1+2k_2+k_3\right) - \sqrt{m^2\left(k_1+2k_2+k_3\right)^2 - 4m^2\left(k_1k_2+k_1k_3+k_2k_3\right)}}{2m^2}}\\
+  &= \sqrt{\frac{k_1+2k_2+k_3 - \sqrt{k_1^2+4k_2^2+k_3^2-2k_1k_3}}{2m}}\\
+\end{align*}
+
+We can compare the uncoupled frequencies (where $k_2=0$) to these:
+
+\begin{align*}
+  \lim_{k_2\rightarrow0}\omega_1 &= \sqrt{\frac{\left(k_1+k_3\right) + \sqrt{k_1^2+k_3^2-2k_1k_3}}{2m}} = \sqrt{\frac{k_1}{m}}\\
+  \lim_{k_2\rightarrow0}\omega_2 &= \sqrt{\frac{\left(k_1+k_3\right) - \sqrt{k_1^2+k_3^2-2k_1k_3}}{2m}} = \sqrt{\frac{k_3}{m}}\\
+\end{align*}
+
+This is just what we expect, a system of two independent uncoupled harmonic oscillators.\\
+
+This was a neat problem.
+There wasn't really anything different about it compared with other similar problems we've done, except perhaps the determinant equation being too difficult to factor, so having to use the quadratic formula on a fourth-order polynomial.
+It was nice that the uncoupled limit worked out so well.\\
 
@@
 
+We begin by defining our generalized coordinates. $\Theta$ is the hoop's CM-connection point angle with the horizontal and $\phi$ is the mass-hoop CM angle with respect to the horizontal.
+
+\begin{align*}
+  x &= R\sin\theta + R\sin\phi\\
+  y &= -R\cos\theta + -R\cos\phi\\
+  \dot{x}^2 + \dot{y}^2 &= R^2\left(\dot{\theta}^2 + \dot{\phi}^2 + 2\dot{\theta}\dot{\phi}\cos\left(\theta-\phi\right)\right)\\
+  \vspace{1cm}\\
+  T &= \frac{1}{2}M\left(\dot{x}^2+\dot{y}^2\right) + \frac{1}{2}I\omega^2\\
+  &= \frac{1}{2}MR^2\left(\dot{\theta}^2 + \dot{\phi}^2 + 2\dot{\phi}\dot{\theta}\cos(\theta-\phi) + 2\dot{\theta}^2\right)\\
+  \mathcal{L} &= \frac{1}{2}MR^2\left(\dot{\theta}^2 + \dot{\phi}^2 + 2\dot{\phi}\dot{\theta}\cos(\theta-\phi) + 2\dot{\theta}^2\right) + mgR\left(\cos\theta + \cos\phi\right) + mgR\cos\theta\\
+  &= \frac{1}{2}MR^2\left(3\dot{\theta}^2 + \dot{\phi}^2 + 2\dot{\phi}\dot{\theta}\cos(\theta-\phi)\right) + mgR\left(2\cos\theta + \cos\phi\right)\\
+  &\hspace{1cm} cos(\theta) = 1 - \frac{\theta^2}{2},\hspace{1cm} cos(\phi) = 1 - \frac{\phi^2}{2}\\
+  &= \frac{1}{2}MR^2\left(3\dot{\theta}^2 + \dot{\phi}^2 + 2\dot{\phi}\dot{\theta}\right) + mgR\left(3 - \theta^2 - \frac{\phi^2}{2}\right)\\
+  \vspace{1cm}\\
+  \frac{d}{dt}\frac{d\mathcal{L}}{d\dot{\theta}} &= \frac{d\mathcal{L}}{d\theta}\\
+  \frac{1}{2}MR^2\left(6\ddot{\theta} + 2\ddot{\phi}\right) &= -2MgR\theta\\
+  R\left(6\ddot{\theta} + 2\ddot{\phi}\right) &= -2g\theta\\
+  3\ddot{\theta} + \ddot{\phi} &= \frac{-2g}{R}\theta\\
+  \frac{d}{dt}\frac{d\mathcal{L}}{d\dot{\phi}} &= \frac{d\mathcal{L}}{d\phi}\\
+  \frac{1}{2}MR^2\left(2\ddot{\phi} + 2\ddot{\theta}\right) &= -MgR\phi\\
+  \ddot{\phi} + \ddot{\theta} &= \frac{-g}{R}\phi\\
+\end{align*}
+
+We have obtained our equations of motion using Lagrangian mechanics.
+We now guess the solution using the complex exponential function.
+This will give us a system of two equations and three unknowns.
+However, this particular system allows us to cancel out the A and B coefficients and solve for $\omega^2$.\\
+
+\begin{align*}
+  \theta &= Ae^{i\omega t}\\
+  \phi &= Be^{i\omega t}\\
+  3\omega^2A + \omega^2B &= \frac{-2g}{R}A\\
+  \omega^2A + \omega^2B &= \frac{-g}{R}B\\
+  A = B\left(\frac{g}{R\omega^2}-1\right)\\
+  3\omega^2\left(\frac{g}{R\omega^2}-1\right) + \omega^2 &= \frac{2g}{R}\left(\frac{g}{R\omega^2}-1\right)\\
+  \frac{3g}{R} - 3\omega^2 + \omega^2 = \frac{2g^2}{R^2}\frac{1}{\omega^2} - \frac{2g}{R}\\
+  -2\omega^4 + \frac{5g}{R}\omega^2 - \frac{2g^2}{R^2} &= 0\\
+  \omega^2&= \frac{-\frac{5g}{R} \pm \sqrt{\frac{25g^2}{R^2} - 8\frac{2g^2}{R^2}}}{-4} = \frac{g}{2R}, \frac{2g}{R}\\
+  \omega_1 &= \sqrt{\frac{g}{2R}}\\
+  \omega_2 &= \sqrt{\frac{2g}{R}}\\
+\end{align*}
+
+This was a very cool problem.
+We applied a bunch of different physics to solve it: Taylor approximations to the small-angle cosine, Lagrangian mechanics, and a complex set of generalized coordinates.
+This was a long and difficult problem, but I enjoyed thinking about how to set it up.
+The trick of using the small-angle approximation as a Taylor series, but only applying it to the large terms (not the terms involving the angular velocities) was very tricky.
+I think that thinking about how values in your equations influence the behavior of the equation, instead of just operating on them, is important to doing good physics.\\
 \end{document}

### 2015-11-28 classical/latex/a9.tex

--- 2015-11-23 classical/latex/a8.tex
@@
 
-\title{Classical Mechanics Assignment 8}
+\title{Classical Mechanics Assignment 9}
 \author{Elliott Capek}
@@
 \section{Problem One: }
-\textbf{A pendulum consists of a mass m suspended by a massless spring with un-extended length $b$ and spring constant $k$. Find Hamilton’s equations of motion.}
-
-\section{Problem Two: }
-\textbf{The potential for an non-harmonic oscillator is:}
-
-\begin{equation*}
-  U(x) = \frac{kx^2}{2} + \frac{bx^4}{4}\\
-\end{equation*}
-
-\textbf{where k and b are positive constants. Find Hamilton’s equations of motion. No final discussion needed.}
-
-\section{Problem Three: }
-\textbf{Reconsider the problem of the two masses and three springs (seen in class) in the event that the three springs all have different force constants (but the masses are equal).
-Find the two characteristic frequencies and compare the magnitudes with the natural frequencies of the two oscillators in the absence of coupling.}
-
-\section{Problem Four: }
-\textbf{A thin hoop of radius R and mass M oscillates in its own plane hanging from a single fixed point.
-Attached to the hoop is a small mass M constrained to move (in a frictionless manner) along the hoop.
-The inertia of the loop for rotation around a point on the loop is $I=2MR^2$.
-Consider only small oscillations, and show that the eigenfrequencies are:}
-
-\begin{equation*}
-  \omega_1 = \sqrt{\frac{2g}{R}}
-\end{equation*}
-
-\begin{equation*}
-  \omega_2 = \sqrt{\frac{g}{2R}}
-\end{equation*}
 

### 2015-11-28 physical_chemistry/latex/a4.tex
add
Add stuffs.

@@
   &\Delta H_1\left(\frac{1}{T_2}-\frac{1}{T_1}\right) + \ln\left(\frac{T_1}{T_2}\right)\Delta C_P - T_1\Delta C_P\left(\frac{1}{T_2}-\frac{1}{T_1}\right) = \frac{\Delta G_2}{T_2} - \frac{\Delta G_1}{T_1}\\
-  \Delta G_2 &= \left(\Delta H_1 - T_1\Delta C_P\right)\left(1 - \frac{T_2}{T_1}\right) + T_2\ln\left(\frac{T_1}{T_2}\right)\Delta C_P + \frac{\Delta G_1}{T_1}\\
+  \Delta G_2 &= \left(\Delta H_1 - T_1\Delta C_P\right)\left(1 - \frac{T_2}{T_1}\right) + T_2\ln\left(\frac{T_1}{T_2}\right)\Delta C_P + \frac{T_2\Delta G_1}{T_1}\\
 \end{align*}
@@
 
+\section{4.16}
+\begin{align*}
+  G &= aT + b + \frac{c}{T}\\
+  S &= -\left(\frac{\partial G}{\partial T}\right)_P = \frac{c}{T^2} - a\\
+  H &= G + TS = b + \frac{2c}{T}\\
+\end{align*}
+
+\section{4.17}
+First we take the water from atmospheric pressure to its vapor pressure, so we can do the gas transition reversibly.
+This transition is $\Delta G_1$.
+We then convert it to a gas, this is $\Delta G_2$.
+We then bring the gas to the vapor pressure of the solid, this is $\Delta G_3$.
+We then sublimate the gas, this is $\Delta G_4$.
+Finally for $\Delta G_5$ we take the solid back up to the starting pressure.
+\begin{align*}
+  V_l &= 1.8*10^{-5} m^3\\
+  V_s &= 1.96*10^{-5} m^3\\
+  \Delta G_1 &= \int VdP = 1.8*10^{-5} m^3 * \left(489Pa - 10^5 Pa\right) = -1.79J\\
+  \Delta G_2 &= 0 \hspace{1cm}\mbox{State transitions @ vapor pressure isenthalpic}\\
+  \Delta G_3 &= \int VdP = \int \frac{RT}{P}dP = (8.314)(298)\ln\left(\frac{475}{489}\right) = -65.2J\\
+  \Delta G_4 &= 0 \hspace{1cm}\mbox{State transitions @ vapor pressure isenthalpic}\\
+  \Delta G_5 &= \int VdP = 1.96*10^{-5} m^3 * \left(10^5Pa - 475Pa\right) = 1.95J\\
+  \Delta G &= -1.79J + -65.2J + 1.95J = -65.04\frac{J}{mol}\\
+\end{align*}
+
+\section{4.24}
+a.)
+\begin{align*}
+  f &= Pe^{\int_0^p \frac{Z-1}{p}dp}\\
+  &= Pe^{\int_0^p \frac{B}{RT}dp}\\
+  &= Pe^{\frac{Bp}{RT}}\\
+  &= Pe^{Z-1}\\
+\end{align*}
+
+b.)
+\begin{align*}
+  f_{H_2} &= PZ = P + \frac{1}{RT}\left(b - \frac{a}{RT}\right)P^2\\
+  &= 50 \mbox{bar} + \frac{1}{\left(8.314*10^{-5} \frac{m^3 \mbox{bar}}{\mbox{K mol}}\right)298K}\left(0.026  - \frac{0.247}{\left(8.314*10^{-5} \frac{m^3 \mbox{bar}}{\mbox{K mol}}\right)298K}\right)\left(2500 \mbox{bar}^2\right)\\
+  &= 52 bar\\
+\end{align*}
+
+\section{4.27}
+To separate, take each gas and separate it. This is a zero-energy process. Then take each separated gas back to 1 mol. Use the ideal gas approximation:
+
+\begin{align*}
+  w &= -2\int pdV = -2nRT \int \frac{1}{V}dV = -2nRT\ln\left(\frac{V_2}{V_1}\right)\\
+  &= 2(8.314)(298)\ln(2) = 3.4kJ\\
+\end{align*}
+
+\section{4.50}
+\begin{align*}
+  Z &= 1 + \left(b - \frac{a}{RT}\right)\frac{P}{RT}\\
+  V &= \frac{RT}{P} + b - \frac{a}{RT}\\
+  \Delta G &= \int_{P_1}^{P_2} dG = \int_{P_1}^{P_2} \frac{\partial G}{\partial P} dP\\
+  &= \int_{P_1}^{P_2} VdP - \int_{P_1}^{P_2} \left(\frac{RT}{P} + b - \frac{a}{RT}\right) dP\\
+  &= RT\ln\left(\frac{P_2}{P_1}\right) + \left(b-\frac{a}{RT}\right)\left(P_2-P_1\right)\\
+\end{align*}
+
+\section{4.54}
+\begin{align*}
+  \Delta S &= -R\sum n_i\ln y_i\\
+  &= -\left(8.314\right)\left(\ln\left(\frac{1}{3}\right) + 2\ln\left(\frac{2}{3}\right)\right) = 15.87 \frac{J}{K}\\
+  \Delta G &= -TS = -298*18.87 = -4.7kJ\\
+\end{align*}
+
 \end{document}

### 2015-12-02 classical/latex/a9.tex
Add working crystal movie generation.

@@
 
-\section{Problem One: }
+\section{Problem One: Inertia tensor of 4-particle system}
+
+\textbf{A four particles system consists of masses $m_i$ at coordinates:}
+\begin{align*}
+  m_1 &= 5m\\
+  \vec{r_1} &= (a, 2a, 0)\\
+  m_2 &= 5m\\
+  \vec{r_2} &= (0, -2a, a)\\
+  m_3 &= 5m\\
+  \vec{r_3} &= (a, a, -a)\\
+  m_4 &= 5m\\
+  \vec{r_4} &= (3a, a, -2a)\\
+\end{align*}
+\textbf{Find the inertia tensor, the principal axes, and the moments of inertia.
+Explicitly check that the principal axes are orthogonal to each other.
+(Hint: you can use an online solver for the third degree polynomial, such as: http://www.calculatorsoup.com/calculators/algebra/cubicequation.php)}
 

### 2015-12-02 comp_model/crystal_phase.c

@@ -7,9 +7,9 @@
 #include <unistd.h>
 
 const double m_pi =  3.14159265358979323846;
-const int size = 215;
+const int size = 500;
 const int animate = 1;
-const int skiprate = 20;
+const int skiprate = 10;
 const int debug = 0;
 
 void log_data(double* data, int fd) {
@@ -37,19 +37,19 @@ int main() {
   double dx = 0.03; //m
   double dy = 0.03; //m
   double dt = 0.0001; //s
-  double t_final = 0.2; //s
+  double t_final = 0.5; //s
   
   double tau = 0.0003;
-  double delta_bar = 0.01; // Thickness of layer (?)
+  double delta_bar = 0.011; // Average thickness of layer (?)
   double F = 1.8; // Latent heat of fusion
   double mu = 0.02; // Strength of anisotropy
-  double anisotropy = 4.0;
+  double anisotropy = 6.0;
   double beta = 0.9; // n-shifting coefficient
   double eta = 10.0; // n-shifting coefficient
   double TM = 1.0;
   double T0 = 0.0;
   double theta_0 = 1.57;
-  int r = 5;
+  int r = 2;
 
   int fd;
   if ((fd = open("data.txt", O_WRONLY | O_CREAT | O_TRUNC, S_IRUSR | S_IWUSR | S_IRGRP | S_IWGRP | S_IWOTH | S_IROTH)) == -1)

### 2015-12-02 comp_model/plot.py

@@ -4,6 +4,7 @@ from __future__ import division
 
 import math
 import numpy as np
+import os
 import sys
 import time
 import matplotlib.pyplot as plt
@@ -13,8 +14,7 @@ plt.ion()
 
 data = np.genfromtxt("data.txt")
 
-if len(sys.argv[1:]) == 0:
-    hi = raw_input("Press any key to play animation.")
+hi = raw_input("Press any key to play animation.")
 
 time = 10 # s
 frames = len(data) / 10
@@ -35,9 +35,19 @@ if  len(data.shape) == 1:
     sys.exit(0)
     
 else:
+    f = 1
     while i < len(data):
         print str(i) + " of " + str(len(data))
         Z = np.reshape(data[i], (math.sqrt(len(data[i])), math.sqrt(len(data[i]))))
         plt.imshow(Z)
-        plt.pause(0.001)
+        
+        if len(sys.argv[1:]) == 1:        
+            fname = 'PNGs/%s-%03d.png' % (sys.argv[1], f)
+            plt.savefig(fname)
+
+        plt.pause(0.0001)
         i += math.ceil(len(data)/frames)
+        f += 1
+    
+if len(sys.argv[1:]) == 1:        
+    os.system("convert -delay 50 PNGs/%s-*.png GIFs/%s.gif" % (sys.argv[1], sys.argv[1]))

### 2015-12-03 comp_model/crystal_phase.c
Adding some GIFs.

@@ -9,7 +9,7 @@
 const double m_pi =  3.14159265358979323846;
 const int size = 500;
 const int animate = 1;
-const int skiprate = 10;
+const int skiprate = 150;
 const int debug = 0;
 
 void log_data(double* data, int fd) {
@@ -37,10 +37,10 @@ int main() {
   double dx = 0.03; //m
   double dy = 0.03; //m
   double dt = 0.0001; //s
-  double t_final = 0.5; //s
+  double t_final = 0.7; //s
   
   double tau = 0.0003;
-  double delta_bar = 0.011; // Average thickness of layer (?)
+  double delta_bar = 0.01; // Average thickness of layer (?)
   double F = 1.8; // Latent heat of fusion
   double mu = 0.02; // Strength of anisotropy
   double anisotropy = 6.0;
@@ -49,7 +49,7 @@ int main() {
   double TM = 1.0;
   double T0 = 0.0;
   double theta_0 = 1.57;
-  int r = 2;
+  int r = 3;
 
   int fd;
   if ((fd = open("data.txt", O_WRONLY | O_CREAT | O_TRUNC, S_IRUSR | S_IWUSR | S_IRGRP | S_IWGRP | S_IWOTH | S_IROTH)) == -1)

### 2015-12-03 comp_model/plot.py

@@ -35,19 +35,17 @@ if  len(data.shape) == 1:
     sys.exit(0)
     
 else:
-    f = 1
     while i < len(data):
         print str(i) + " of " + str(len(data))
         Z = np.reshape(data[i], (math.sqrt(len(data[i])), math.sqrt(len(data[i]))))
         plt.imshow(Z)
         
         if len(sys.argv[1:]) == 1:        
-            fname = 'PNGs/%s-%03d.png' % (sys.argv[1], f)
+            fname = 'PNGs/%s-%03d.png' % (sys.argv[1], i)
             plt.savefig(fname)
-
+            
         plt.pause(0.0001)
-        i += math.ceil(len(data)/frames)
-        f += 1
+        i += 1
     
 if len(sys.argv[1:]) == 1:        
     os.system("convert -delay 50 PNGs/%s-*.png GIFs/%s.gif" % (sys.argv[1], sys.argv[1]))

### 2015-12-03 comp_model/crystal_phase.c
Adding more figures.

@@ -9,7 +9,7 @@
 const double m_pi =  3.14159265358979323846;
 const int size = 500;
 const int animate = 1;
-const int skiprate = 150;
+const int skiprate = 450;
 const int debug = 0;
 
 void log_data(double* data, int fd) {
@@ -43,7 +43,7 @@ int main() {
   double delta_bar = 0.01; // Average thickness of layer (?)
   double F = 1.8; // Latent heat of fusion
   double mu = 0.02; // Strength of anisotropy
-  double anisotropy = 6.0;
+  double anisotropy = 8.0;
   double beta = 0.9; // n-shifting coefficient
   double eta = 10.0; // n-shifting coefficient
   double TM = 1.0;

### 2015-12-06 physical_chemistry/latex/pchem_tips.tex
Adding everything.

@@
 \usepackage{amsmath}
+\usepackage{gensymb}
 \usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins
@@
 
-\section{Random relationships}
-
-\subsection{Weird unknown axioms}
+\section{Equations}
+\begin{align*}
+  \left(\frac{dU}{dV}\right)_T &= T\left(\frac{dP}{dT}\right)_V - P\\
+  \left(\frac{dH}{dP}\right)_T &= -T\left(\frac{dV}{dT}\right)_P + V\\
+  &\mbox{Arrived at via first law, expanding}\\
+  &\mbox{a differential and a Maxwell relation}\\
+\end{align*}
 
 \begin{align*}
-  \frac{dH}{dP}_t &= -T\frac{dV}{dT}_P + V\\
+  \frac{d\left(\frac{G}{T}\right)}{dT} &= -\frac{H}{T^2}\\
+\end{align*}
+
+\begin{align*}
+  H &= -T^2\left(\frac{\partial\left(G/T\right)}{\partial T}\right)_P
+\end{align*}
+
+\begin{align*}
+  \kappa &= -\frac{1}{V}\left(\frac{\partial V}{\partial P}\right)_T \hspace{1cm} \mbox{Cubic expansion}\\
+\alpha &= \frac{1}{V}\left(\frac{\partial V}{\partial T}\right)_P \hspace{1cm} \mbox{Cubic expansion}\\
+\end{align*}
+
+\begin{align*}
+  \left(\frac{\partial U}{\partial V}\right)_T &= \left(\frac{\partial H}{\partial P}\right)_T = 0\\
+  \left(\frac{\partial U}{\partial T}\right)_V &= C_V \hspace{1cm} \mbox{For an ideal gas}\\
 \end{align*}
@@
 
+\begin{align*}
+  \frac{\partial G}{\partial P}_T &= V\\
+  \Delta \overline{G}_{ideal} &= \Delta G_{ideal}\degree + RT\ln\left(\frac{P}{P\degree}\right)\\
+  \Delta \overline{G} &= \Delta G\degree + RT\ln\left(\frac{f}{P\degree}\right)\\
+  \frac{f}{P} &= e^{\frac{1}{RT}\int_0^P(\bar{V}-\bar{V}^{id})dP} = e^{\int_0^P\frac{Z-1}{P}dP}\\
+  \mu_i &= \mu_i\degree + RT\ln(a_i)\\
+  \mu_i &= \mu_i\degree + RT\ln\left(\frac{f_i}{P\degree}\right)\\
+\end{align*}
+
+\begin{align*}
+  K &= \Pi a_{i,eq}^{\nu_i}\\
+  Q &= \Pi a_{i}^{\nu_i}\\
+  K &= e^{\frac{-\Delta_r G\degree}{RT}}\\
+  K_{gas} &= \Pi\left(\frac{f_{i,eq}}{P\degree}\right)^{\nu_i}\\
+  K_{gas,ideal} &= \Pi\left(\frac{P_{i,eq,ideal}}{P\degree}\right)^{\nu_i}\\
+\end{align*}
+Equilibrium constant. The product of all the concentrations of the reaction. At standard state, the Gibbs free energy is $\sum \mu_i\degree\nu_i$.
+
+\begin{align*}
+  P_i &= \frac{n_iRT}{V} = C_iRT\\
+  K_p &= \Pi \left(\frac{C_iRT}{P\degree}\right)^{\nu_i} = \left(\frac{C_i\degree RT}{P\degree}\right)^{\nu_i} \Pi \left(\frac{C_i}{C_i\degree}\right)^{\nu_i}\\
+\end{align*}
+
+\begin{align*}
+  \mbox{Ideal gas expansion}&\hspace{2cm}\Delta S = R\ln(\frac{V_2}{V_1})\\
+  \mbox{Ideal gas mixing}&\hspace{2cm}\Delta S = -R\sum y_i\ln(y_i)\\
+  \mbox{Nonideal phase transition}&\hspace{2cm} \Delta S = \frac{\Delta H}{T}\\
+\end{align*}
+
+\begin{align*}
+  \mbox{Boltzmann Hypothesis}&\hspace{2cm} S = k\ln(\Omega)\\
+\end{align*}
+
+\begin{align*}
+  \mbox{Second law of thermodynamics}&\hspace{2cm} dS \geq \frac{dQ_{rev}}{T}\\
+  \mbox{Third law of thermodynamics}&\hspace{2cm}\mbox{Crystals have zero entropy}\\
+\end{align*}
+
+\begin{align*}
+  \mbox{Carnot cycle}&\hspace{2cm}\eta = \frac{w}{q_1} = \frac{\Delta T}{T_1}\\
+\end{align*}
+
+\begin{align*}
+  \mu_i = \mu_i\degree + RT\ln\left(\frac{P_i}{P\degree}\right)\\
+  \overline{G}_i = \overline{G}_i\degree + RT\ln\left(\frac{P_i}{P\degree}\right)\\
+\end{align*}
+
+\section{Things to know}
+\begin{align*}
+  &\mbox{Prove entropy is exact for both ideal and nonideal gasses}\\
+  &\mbox{Derive first relations in document, the ones with dU/dV}\\
+\end{align*}
+
 \end{document}

### 2015-12-06 comp_model/crystal_phase.c

@@ -7,9 +7,9 @@
 #include <unistd.h>
 
 const double m_pi =  3.14159265358979323846;
-const int size = 500;
+const int size = 1000;
 const int animate = 1;
-const int skiprate = 450;
+const int skiprate = 250;
 const int debug = 0;
 
 void log_data(double* data, int fd) {
@@ -43,7 +43,7 @@ int main() {
   double delta_bar = 0.01; // Average thickness of layer (?)
   double F = 1.8; // Latent heat of fusion
   double mu = 0.02; // Strength of anisotropy
-  double anisotropy = 8.0;
+  double anisotropy = 6.0;
   double beta = 0.9; // n-shifting coefficient
   double eta = 10.0; // n-shifting coefficient
   double TM = 1.0;

### 2015-12-06 comp_model/crystal_phase.c
none

@@ -7,7 +7,7 @@
 #include <unistd.h>
 
 const double m_pi =  3.14159265358979323846;
-const int size = 500;
+const int size = 215;
 const int animate = 1;
 const int skiprate = 10;
 const int debug = 0;
@@ -37,10 +37,10 @@ int main() {
   double dx = 0.03; //m
   double dy = 0.03; //m
   double dt = 0.0001; //s
-  double t_final = 0.5; //s
+  double t_final = 0.2; //s
   
   double tau = 0.0003;
-  double delta_bar = 0.011; // Average thickness of layer (?)
+  double delta_bar = 0.01; // Average thickness of layer (?)
   double F = 1.8; // Latent heat of fusion
   double mu = 0.02; // Strength of anisotropy
   double anisotropy = 6.0;
@@ -49,7 +49,7 @@ int main() {
   double TM = 1.0;
   double T0 = 0.0;
   double theta_0 = 1.57;
-  int r = 2;
+  int r = 4;
 
   int fd;
   if ((fd = open("data.txt", O_WRONLY | O_CREAT | O_TRUNC, S_IRUSR | S_IWUSR | S_IRGRP | S_IWGRP | S_IWOTH | S_IROTH)) == -1)

### 2015-12-08 physical_chemistry/latex/pchem_tips.tex
blah

@@
   \frac{f}{P} &= e^{\frac{1}{RT}\int_0^P(\bar{V}-\bar{V}^{id})dP} = e^{\int_0^P\frac{Z-1}{P}dP}\\
-  \mu_i &= \mu_i\degree + RT\ln(a_i)\\
+  \mu_i &= \mu_i\degree + RT\ln(a_i) \hspace{2cm}\mbox{Activity ~ purity, caps at 1}\\
+  &\mbox{Pure liquids have an activity of 1}\\
   \mu_i &= \mu_i\degree + RT\ln\left(\frac{f_i}{P\degree}\right)\\
@@
   K_p &= \Pi \left(\frac{C_iRT}{P\degree}\right)^{\nu_i} = \left(\frac{C_i\degree RT}{P\degree}\right)^{\nu_i} \Pi \left(\frac{C_i}{C_i\degree}\right)^{\nu_i}\\
+  K_c &= \Pi\left(\frac{c_i}{c\degree}\right)^{\nu_i}\\
+  k &= \frac{c}{K}\\
 \end{align*}
@@
 \begin{align*}
-  \mbox{Ideal gas expansion}&\hspace{2cm}\Delta S = R\ln(\frac{V_2}{V_1})\\
+  V &= \frac{mRT}{MP}\\
+  &\mbox{For dissociation A -> 2B}\\
+  K &= \frac{4\xi^2}{1-\xi^2}\\
+  \xi &= \frac{M_1-M_2}{M_2}\\
+\end{align*}
+
+\begin{align*}
+  k &= e^{\frac{-\Delta G}{RT}}\\
+\end{align*}
+
+\begin{align*}
+  \Delta_f G_{isomer} &= -RT\ln\left(\sum e^{\frac{-\Delta_f G_i}{RT}}\right)\\
+  \Delta_f H_{isomer}\degree &= \sum r_i \Delta_f H_i\degree\\
+  r_i &= e^{\frac{\Delta_f G_{isomer} - \Delta_f G_i}{RT}}\\
+\end{align*}
+
+\begin{align*}
+  \mbox{Ideal gas expansion - isothermal?}&\hspace{2cm}\Delta S = R\ln(\frac{V_2}{V_1}) = R\ln(\frac{P_1}{P_2})\\
   \mbox{Ideal gas mixing}&\hspace{2cm}\Delta S = -R\sum y_i\ln(y_i)\\
   \mbox{Nonideal phase transition}&\hspace{2cm} \Delta S = \frac{\Delta H}{T}\\
+  \mbox{Constant V heating}&\hspace{2cm} \Delta S = \int_{T1}^{T2} \frac{C_V}{T}dT\\
+  \mbox{Constant P heating}&\hspace{2cm} \Delta S = \int_{T1}^{T2} \frac{C_P}{T}dT\\
+  \mbox{Phase change constant T,P}&\hspace{2cm} \Delta S = \frac{\Delta H}{T}\\
 \end{align*}
@@
 \begin{align*}
-  \mu_i = \mu_i\degree + RT\ln\left(\frac{P_i}{P\degree}\right)\\
-  \overline{G}_i = \overline{G}_i\degree + RT\ln\left(\frac{P_i}{P\degree}\right)\\
+  \mu_i &= \mu_i\degree + RT\ln\left(\frac{P_i}{P\degree}\right)\\
+  \overline{G}_i &= \overline{G}_i\degree + RT\ln\left(\frac{P_i}{P\degree}\right)\\
+  \Delta_r G &= \sum \nu_i\mu_i = \sum \nu_i \Delta_f G_i\\
+\end{align*}
+
+\begin{align*}
+  \ln\left(\frac{K_2}{K_1}\right) &= \frac{\Delta_r H\degree(T_2-T_1)}{RT_1T_2}\\
+\end{align*}
+
+\begin{align*}
+  \mbox{Clapeyron Equation}\\
+  \frac{dP}{dT} &= \frac{\Delta S}{\Delta V} = \frac{\Delta H}{T\Delta V}\\
+  a_i &= \frac{P_i}{P_i*}\hspace{2cm}\mbox{The activity is the vapor pressure over the pure vapor pressure}\\
+  P_i &= x_iP_i*\hspace{2cm}\mbox{xi is mol frawction of i in liquid}\\
+  a_i &= x_i\hspace{2cm}\mbox{Ideal solutions only}\\
+  \mu_i(l) &= \mu_i(l)\degree + RT\ln(x_i)\hspace{2cm}\mbox{Ideal solutions only}\\
+  P_i &= y_iP = K_ix_i\hspace{2cm}\mbox{Henry's law}\\
+\end{align*}
+
+\begin{align*}
+  n_i &= nz_i = n_{\ell}x_i + n_{v}y_i\hspace{2cm}\mbox{x: solution fraction, y: vapor fraction}\\
+  \frac{n_\ell}{n_v} &= \frac{z_i - y_i}{x_i - z_i}\hspace{2cm}\mbox{Lever rule, tells you liquid:gas, since z in between x and y}\\
+  \mu_i &= \mu_i\degree + RT\ln(v_ix_i)\\
+  v_i &= \frac{P_i}{x_iP_i*} = \frac{y_iP}{x_iP_i*}\hspace{2cm}\mbox{Rault's law pure}\\
+  v_i' &= \frac{y_iP}{x_iK_i}\hspace{2cm}\mbox{Henry's law dilute}\\
+\end{align*}
+
+\begin{align*}
+  \Delta T &= \frac{RT^2_{f/b}x_B}{\Delta H}\hspace{2cm}\mbox{Boiling/Freezing pt change}\\
+  \Delta T &= \frac{RT^2_{f/b}m_Am_B}{\Delta H}\\
 \end{align*}

### 2015-12-08 physical_chemistry/latex/pchem_tips.tex
blah

@@
   a_i &= \frac{P_i}{P_i*}\hspace{2cm}\mbox{The activity is the vapor pressure over the pure vapor pressure}\\
+  \frac{dP}{dT} &= \frac{P\Delta_{vap}H}{RT^2}\hspace{2cm}\mbox{Clausius-Clapeyron}\\
+  \ln\left(\frac{P_2}{P_1}\right) &= \frac{\Delta_{vap}H(T_2-T_1)}{RT_1T_2}\\
   P_i &= x_iP_i*\hspace{2cm}\mbox{xi is mol frawction of i in liquid}\\
@@
   P_i &= y_iP = K_ix_i\hspace{2cm}\mbox{Henry's law}\\
+\end{align*}
+
+\begin{align*}
+  \Delta T_f &= \frac{RT^2_{fus,A}M_am_B}{\Delta_{fus}H_A\degree} = K_fm_B\\
+\end{align*}
+
+\begin{align*}
+  \Delta_{sub} H &= \Delta_{vap} H + \Delta_{fus} H\\
 \end{align*}

### 2015-12-12 comp_model/latex/paper2.tex
add everything, writing paper 2

--- 2015-11-10 comp_model/latex/paper1.tex
@@
 
-\title{Effect of Parameters in Phase-Field Model for simulating Solid-Liquid Transitions}
+\title{Exploring Theory and Anisotropy Effects in Dendritic Phase-Field Model for Crystal Formation}
 \author{Elliott Capek}
@@
 
-\section{Introduction}
-The Phase-Field model is a method for simulating phase transitions in systems.
-Phase transitions in real life are discontinuous, making them hard to represent analytically.
-The Phase-Field method overcomes this by representing the phase of a region with a space- and time-dependent function $\phi(r,t)$.
-$\phi$ ranges from $0$ to $1$, with either end of the range being one phase or the other, and numbers between $0$ and $1$ representing transitional regions.
-These transitional regions, while not physical, can approximate reality well if the distance between a 0-zone and a 1-zone is as small as possible.
-\\
+\section{Variations on Model}
+\subsection*{Zhao Model}
+The Zhao model, described in [2], is most notable because of its use of noise to simulate thermal fluctuations in the crystal formation. The
 
-In this paper we seek to reproduce the results of Sanal [1], where a solid-liquid phase transition with anisotropic dependence is time-evolved.
-The anisotropy (angular dependence) of the equations used in this paper allow for radially-symmetric crystals to form.
-In this formulation, $\phi=1$ corresponds to solid phase and $\phi=0$ corresponds to liquid phase.
-The equation for time-evolving $\phi$ is given in Sanal as:
+\section{Anisotropy Effects}
+\subsection*{Original Kanazawa Model}
+The original model used in Kobayashi [1] introduces anisotropy by making interface layer thickness $\delta$ a function of $\theta$, the angle between $\nabla\phi$ and the x-axis.
+This dependence is given by:\\
 
-\begin{equation}  \label{eq:dPhidt}
-  \tau \frac{\partial \phi}{\partial t} = -\frac{\partial}{\partial x} \left(\epsilon \frac{d\epsilon}{d\theta}\frac{\partial\phi}{\partial y}\right) + \frac{\partial}{\partial y} \left(\epsilon \frac{d\epsilon}{d\theta}\frac{\partial \phi}{\partial x}\right) + \nabla \cdot \left(\epsilon^2\nabla\phi\right) + \phi\left(1-\phi\right)\left(\phi-0.5+m\right)
-\end{equation}
+\begin{equation*}
+  \delta(\theta) = \delta_0(1 + \mu\cos(a_0(\theta-\theta_0)))
+\end{equation*}
 
-Here $\tau$ is the relaxation time, or the time it takes $\phi$ to come into equilibrium with its neighbors.
-$\epsilon$ is a small $\theta$-dependent parameter which controls how quickly the solid-liquid interface moves.
-$\theta$ is the angle from the origin measured from the horizontal.
-Making $\epsilon$ depend on $\theta$ introduces anisotropy into the model, allowing different angles to spread at different rates.
-$m$ is the thermodynamic driving force, a temperature-dependent number which determines how favorable solidification is.
-Larger values of temperature (and hence $m$) correspond to quicker solidification.
-\\
+where $\delta_0$ is the average layer width, $\mu$ is the effect of anisotropy and $\theta_0$ is the rotation of the crystal.
+In Figure ($\ref{fig:r-delta-4}$) we show how $\delta$ varies with $\theta$ and the effect this has on the resultant crystal.
+Figures ($\ref{fig:r-delta-3}$) and ($\ref{fig:r-delta-6}$) show the same effect for anisotropies of $6$ and $3$, respectively.\\
 
-The equation for $\epsilon(\theta)$ is given as follows:
+These plots carry a lot of explanatory power.
+The lighter regions represent the lowest values of $\delta$.
+The system tends to minimize free energy, and $\delta$ makes a positive contribution to the value of $G$.
+We can think of $\delta(\theta)$ as a cost function that makes certain angles of $\nabla\phi$ more expensive than others.
+If we look at Figure ($\ref{fig-r-delta-4}$), the plot for $a=4$, we can see that
 
-\begin{equation}
-  \epsilon = 1 + \delta\cos\left(j\left(\theta_0-\theta\right)\right)
-\end{equation}
+\begin{figure}[h!]
+  \centering
+  \textbf{a. }\includegraphics[width=0.2\textwidth]{../radial-anisotropy-4.png}
+  \hspace{1cm}\textbf{b. }\includegraphics[width=0.2\textwidth]{../anis-4.png}
+  \caption{\textbf{a.} Values of $\delta(\theta)$, where $\theta$ is the angular polar coordinate of points in the XY plane.
+Darker values correspond to larger values of $\delta$.
+\textbf{b.} The resulting crystal, plotted for the following parameters: $\a=4$,$dx=0.0003$, $F=1.8$ and $t=0.2$.}
+  \label{fig:r-delta-4}
+\end{figure}
 
-where $\delta$ is the strength of anisotropy, $j$ is the harmonic number and $\theta_0$ is the angular offset. The higher $j$ is, the more angular extrema will exist in the system.\\
+\begin{figure}[h!]
+  \centering
+  \textbf{a. }\includegraphics[width=0.2\textwidth]{../radial-anisotropy-3.png}
+  \hspace{1cm}\textbf{b. }\includegraphics[width=0.2\textwidth]{../anis-3.png}
+  \caption{$\delta$ and $\phi$, this time for $a=3$.}
+  \label{fig:r-delta-3}
+\end{figure}
 
-Temperature is also important to this model, as it decides the value of m, the rate of interface front movement. Temperature's time-evolution is given by:
+\begin{figure}[h!]
+  \centering
+  \textbf{a. }\includegraphics[width=0.2\textwidth]{../radial-anisotropy-6.png}
+  \hspace{1cm}\textbf{b. }\includegraphics[width=0.2\textwidth]{../anis-6.png}
+  \caption{$\delta$ and $\phi$, this time for $a=6$.}
+  \label{fig:r-delta-6}
+\end{figure}
 
-\begin{equation}  \label{eq:dTdt}
-  \frac{\partial T}{\partial t} = \nabla^2T + k\frac{\partial \phi}{\partial t}
-\end{equation}
-
-where k is the latent heat of the phase change. Liquid freezing into a solid is exergonic, so heat is evolved when $\frac{\partial\phi}{\partial t}$ is positive. \\
-
-\section{Method}
-To simulate the behavior of the differential equation given by Sanal, we use the forwards-Euler finite difference method to time-evolve Equations (\ref{eq:dPhidt}) and (\ref{eq:dTdt}).
-We implement this in a C program (see Code section) and a Python script using matplotlib to plot our results.
-
-\section{Results}
-
-\section{Next Steps}
+\section{Reference}
+\textbf{[1]} Kobayashi citation here\\
 

### 2015-12-12 comp_model/crystal_phase.c

@@ -7,13 +7,9 @@
 #include <unistd.h>
 
 const double m_pi =  3.14159265358979323846;
-<<<<<<< HEAD
-const int size = 215;
-=======
-const int size = 1000;
->>>>>>> bcadcb6ba16aac3549056064faf90d3ae51a63d9
-const int animate = 1;
-const int skiprate = 250;
+const int size = 300;
+const int animate = 0;
+const int skiprate = 1000;
 const int debug = 0;
 
 void log_data(double* data, int fd) {
@@ -41,31 +37,20 @@ int main() {
   double dx = 0.03; //m
   double dy = 0.03; //m
   double dt = 0.0001; //s
-<<<<<<< HEAD
-  double t_final = 0.2; //s
+  double t_final = 0.4; //s
   
   double tau = 0.0003;
   double delta_bar = 0.01; // Average thickness of layer (?)
-  double F = 1.8; // Latent heat of fusion
-=======
-  double t_final = 0.7; //s
-  
-  double tau = 0.0003;
-  double delta_bar = 0.01; // Average thickness of layer (?)
-  double F = 0.5; // Latent heat of fusion
->>>>>>> bcadcb6ba16aac3549056064faf90d3ae51a63d9
+  double F = 2.0; // Latent heat of fusion
+
   double mu = 0.02; // Strength of anisotropy
-  double anisotropy = 6.0;
+  double anisotropy = 4.0;
   double beta = 0.9; // n-shifting coefficient
   double eta = 10.0; // n-shifting coefficient
   double TM = 1.0;
   double T0 = 0.0;
   double theta_0 = 1.57;
-<<<<<<< HEAD
-  int r = 4;
-=======
-  int r = 3;
->>>>>>> bcadcb6ba16aac3549056064faf90d3ae51a63d9
+  int r = 10;
 
   int fd;
   if ((fd = open("data.txt", O_WRONLY | O_CREAT | O_TRUNC, S_IRUSR | S_IWUSR | S_IRGRP | S_IWGRP | S_IWOTH | S_IROTH)) == -1)

### 2015-12-12 comp_model/latex/paper2.tex
Add lots of stuff.

@@
 
-\section{Variations on Model}
-\subsection*{Zhao Model}
-The Zhao model, described in [2], is most notable because of its use of noise to simulate thermal fluctuations in the crystal formation. The
+\section{Introduction}
+This is the continuation of a two-part paper exploring a Ginzburg-Landau-inspired model for describing the dendritic behavior of crystal formation using an order parameter originally formulated by Koboyashi [1].
+This section of the paper will focus on the parameters in the model, giving conceptual explanations for the effects of varying these parameters.
+It also describes some further work done on the Kobayashi model.
 
-\section{Anisotropy Effects}
-\subsection*{Original Kanazawa Model}
+\section{Effect of Parameters}
+In this section we describe the influence of various parameters on the Kobayashi dendritic crystal model.
+
+\subsection{Latent Heat}
+Latent heat K enters the model in the heat evolution equation:
+
+\begin{equation}
+  \frac{\partial u}{\partial t} = \nabla^2u + K\frac{\partial\phi}{\partial t}\\
+\end{equation}
+
+The larger K is, the greater the heat evolved from the solidifying crystal.
+Our results for crystal growth at different latent heats, shown in Figure ($\ref{fig:latent-heat}$), agrees with Shah [5].
+The general effects we observe are an increase in anisotropic effects with increase in latent heat, and a decrease in growth speed and crystal size with latent heat.
+These results are intuitive.
+The $K\dot{\phi}$ term acts to magnify the effects of crystal formation.
+The initial conditions are set such that the field begins colder than the liquid-solid equilibrium, so crystal formation is preferred.
+Solidification is the only means of introducing higher temperature to the system, since the normal heat equation only serves to spread heat out.
+\\
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.15\textwidth]{../F-0.5-10.png}
+  \includegraphics[width=0.15\textwidth]{../F-1.0-10.png}
+  \includegraphics[width=0.15\textwidth]{../F-1.5-10.png}
+  \includegraphics[width=0.15\textwidth]{../F-1.8-10.png}
+  \includegraphics[width=0.15\textwidth]{../F-2.5-10.png}
+  \caption{Effect of latent heat on crystal growth for the same parameters. Latent heat values of 0.5, 1.0, 1.5, 1.8 and 2.0 from left to right.}
+  \label{fig:latent-heat}
+\end{figure}
+
+If latent heat is low, then little head is added to the system via crystallization and all angles of $\nabla\phi$ can crystallize.
+While the $\delta(\theta)$ term does add an energy cost to certain angles, we can see from Figure ($\ref{fig:latent-heat}.a$) that minimizing potential f, which heavily favors the solid phase at low temperatures, outweighs the cost of anisotropic angles.\\
+
+At higher values of K the opposite is true.
+When crystallization introduces lots of heat to the system, melting begins to take effect and we no longer see rampant crystal growth.
+We can assume that because size decreases with K, at higher temperatures our f is shifted to favor the liquid phase, so the magnitude of $\frac{\delta G}{\delta\phi}$ is lower, leading to slower growth.
+The reason anisotropy is more pronounced at high K is that at low K, the magnitude of potential f is much larger than the anisotropic interfacial width term, so anisotropy is less important.
+At temperatures closer to equilibrium f becomes much closer to zero, so anisotropy plays a larger role.\\
+
+Figure ($\ref{fig:latent-heat-u}$) shows the temperature field for K=2.0.
+From this we can see that dimensionless temperature is fairly uniform around the boundary, at a value of about -0.9.
+At this value f still favors solidification, but it is less energetically favorable, which can explain why crystal formation is slower.
+\\
+
+Figure ($\ref{fig:latent-heat-cold}$) shows the same information for K=0.8.
+Here we can see that at this lower value of K the bulk of the crystal is about $0.025$ degreeless temperature units colder than the K=2.0 crystal.
+This is a small amount, but it may be enough to explain why such rampant growth occurs at such a low value of K.\\
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.45\textwidth]{../u-field-6.png}
+  \includegraphics[width=0.45\textwidth]{../phi-field-6.png}
+  \caption{Plots of both temperature and phi for a K=2.0 crystal growth.}
+  \label{fig:latent-heat}
+\end{figure}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.45\textwidth]{../u-cold-field-3.png}
+  \includegraphics[width=0.45\textwidth]{../phi-cold-field-3.png}
+  \caption{Plots of both temperature and phi for a K=0.8 crystal growth.}
+  \label{fig:latent-heat-cold}
+\end{figure}
+
+\subsection{Anisotropy}
 The original model used in Kobayashi [1] introduces anisotropy by making interface layer thickness $\delta$ a function of $\theta$, the angle between $\nabla\phi$ and the x-axis.
@@
 
-\begin{equation*}
+\begin{equation}
   \delta(\theta) = \delta_0(1 + \mu\cos(a_0(\theta-\theta_0)))
-\end{equation*}
+  \label{eq:kobayoshi-delta}
+\end{equation}
 
 where $\delta_0$ is the average layer width, $\mu$ is the effect of anisotropy and $\theta_0$ is the rotation of the crystal.
+Anisotropy is the most obvious parameter to change in this model.
+As expected, changing the modes in our $\delta(\theta)$ interface width changes the behavior based on circular initial conditions.
+Figure ($\ref{fig:anisotropy}$) shows crystal growth behavior for different anisotropies.\\
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.15\textwidth]{../A-3-17.png}
+  \includegraphics[width=0.15\textwidth]{../A-4-17.png}
+  \includegraphics[width=0.15\textwidth]{../A-6-17.png}
+  \includegraphics[width=0.15\textwidth]{../A-8-17.png}
+  \caption{Effect of anisotropy on crystal formation. Anisotropy values of 3, 4, 6 and 8 from left to right.}
+  \label{fig:anisotropy}
+\end{figure}
+
 In Figure ($\ref{fig:r-delta-4}$) we show how $\delta$ varies with $\theta$ and the effect this has on the resultant crystal.
@@
 We can think of $\delta(\theta)$ as a cost function that makes certain angles of $\nabla\phi$ more expensive than others.
-If we look at Figure ($\ref{fig-r-delta-4}$), the plot for $a=4$, we can see that
+If we look at Figure ($\ref{fig:r-delta-4}$.b), the plot for $a=4$, we can see that nearly all interface angles are at multiples of $\pi/2$.
+From Figure ($\ref{fig:r-delta-4}$.a) we can see that these are the least expensive angles.
+This same effect is seen by looking at Figure ($\ref{fig:r-delta-6}$), where $a=6$.\\
+
+Interestingly, the effect is much less pronounced for Figure ($\ref{fig:r-delta-3}$) at $a=3$.
+By looking at Figure ($\ref{fig:r-delta-3}$.b) we see weaker anisotropy, or more angles in the crystal interface.
+This can be explained by looking at Equation ($\ref{eq:kobayoshi-delta}$).
+When anisotropy $a$ is low, angles near the minima are less expensive than when anisotropy is high.
+This means that, while the system will still tend to maximize the lowest-cost angles, those near it will be more available, and thus the final crystal will have a wider variance in angles.\\
 
@@
 Darker values correspond to larger values of $\delta$.
-\textbf{b.} The resulting crystal, plotted for the following parameters: $\a=4$,$dx=0.0003$, $F=1.8$ and $t=0.2$.}
+\textbf{b.} The resulting crystal, plotted for the following parameters: $a=4$,$dx=0.0003$, $F=1.8$ and $t=0.2$.}
   \label{fig:r-delta-4}
@@
 
+\subsection{Layer Thickness}
+Layer thickness has an interesting effect on the crystal growth.
+Increasing average layer thickness $\delta_0$ in Equation ($\ref{eq:kobayoshi-delta}$) effectively increases the magnitude of the anisotropy term in our free energy definition.
+We see this pattern reflected in Figure ($\ref{fig:dleta}$).
+The highest $\delta_0$ crystal on the right seems to have the most ``anisotropy'' - it has the most offshoots, all travelling in the same direction (a minimum-energy angle in Equation ($\ref{eq:delta-6}$)).
+However, ``increased anisotropy'' doesn't fully describe the weirdness that arises in the $\delta_0 = 0.11$ crystal.
+It isn't intuitive why holes would show up in the middle of crystals.
+The large non-phase regions at the edges of the main dendrites, the green regions between $\phi=0$ and $\phi=1$ can be explained by accepting that at high values of anisotropy the penalties of being at a non-phase $\phi$ value imposed by potential f are less important, since the anisotropy term is being minimized.\\
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.15\textwidth]{../d-0.008-5.png}
+  \includegraphics[width=0.15\textwidth]{../d-0.09-5.png}
+  \includegraphics[width=0.15\textwidth]{../d-0.1-5.png}
+  \includegraphics[width=0.15\textwidth]{../d-0.11-5.png}
+  \caption{Effect of boundary width on crystal formation. $\delta$ values of 0.008, 0.009, 0.1 and 0.11 from left to right.}
+  \label{fig:delta}
+\end{figure}
+
+\section{Variations on Model}
+\subsection*{Zhao Model}
+The Zhao model, described in [2], is most notable because of its use of noise to simulate thermal fluctuations in the crystal formation. The evolution equation for $\phi$ is given by:
+
+\begin{equation}
+  \tau \frac{d\phi}{dt} = L_0\left(\frac{1}{T_M} - \frac{1}{F}\right)p'(\phi) - \frac{W_f}{\lambda^2}\frac{\partial^2\phi}{\partial\phi\partial T} + \frac{\epsilon_f^2}{T}\nabla^2\phi + f_0\\
+\end{equation}
+
+This model has some significant deviations from the original Kobayashi model.
+First, there is a latent heat term modifying the evolution of $\phi$.
+This is interesting because latent heat, and therefore the heat evolved from the crystal interface, has a direct impact on the change in $\phi$.
+This model does not use the same potential polynomial used in Kobayashi, so perhaps this is how the boundary movement is taken into account.\\
+
+Another interesting aspect of this model is that, in an unfortunate symbol collision with the Sanal [3] crystal paper, $\epsilon_f$ is not the boundary width, but a constant.
+$W_f$ is the boundary width.
+This means that the all-important $\nabla^2\phi$ term is separated from the boundary width, which is very different from the traditional term found in Kobayashi:
+
+\begin{equation}
+  \frac{\delta^2(\nabla\phi)^2}{2}
+\end{equation}
+
+This suggests that perhaps the semi-unintuitive linking of the interface width term with $\nabla^2\phi$ is unnecessary.\\
+
+The most important part of the Zhao model is that it introduces randomness into the deterministic model.
+The $f_0$ term on the right is a Gaussian-sampled zero-centered term which randomly speeds up or slows down crystal formation.
+The effects of this added randomness are shown in [2].
+The crystals evolved with noise show more small dendrites shooting off the main ones, which makes sense.
+Small nucleations of new dendrites can be caused by the $f_0$ term which would otherwise never show up.\\
+
+\subsection{LDA Model}
+This is less a new model and more a limiting case of the Kobayashi model.
+When the discussed material is particularly inert, we can make the Local Density Approximation and discount the effects of surface tension.
+Thus, the only energy we need to worry about is our potential, which can be given by a simple polynomial, such as the equilibrium case $f_{eq} = (1-\phi^2)^2/4$, as given in Biben [4], which gives more on this topic.
+In the case of this model, already existing boundaries would solidify, but extra terms are needed to actually cause crystal formation.\\
+
+In the section on latent heat we discussed how the magnitude of the phase potential vs surface tension (anisotropy) terms in our free energy definition changed the behavior.
+Figure ($\ref{fig:latent-heat}.a$) is an approximation to the LDA model.
+As we have argued, the potential term must outweigh the anisotropy term since we see little anisotropy in this crystal and rampant growth.\\
+
+\section{Further Questions}
+The major puzzle still present in this model is the reason for angular assymetry.
+It seems the model should be completely radially symmetric for all values of anisotropy.
+Figures $\ref{fig:r-delta-4}.a$, $\ref{fig:r-delta-3}.a$ and $\ref{fig:r-delta-6}.a$ are all angularly symmetric.
+While the angular plot for $a=3$ is less ``well-behaved'' than the even anisotropy plots, it is still symmetric.
+The more chaotic behavior of this plot may at least partially explain why the anisotropy-3 crystal is so much more strange-looking than the others, but it is not a complete explanation.\\
+
+It is unlikely that the asymmetry comes from a defect in the simulation itself, since angular asymmetry can also be seen in Biben [4] and Shah [5]. It remains to be see what the source of this is.\\
+
 \section{Reference}
-\textbf{[1]} Kobayashi citation here\\
-
+\textbf{[1]} Kobayashi, R. (1993) Modeling and Numerical Simulations of Dendritic Crystal Growth. \textit{Physica D}, 63, 410-423. http://dx.doi.org/10.1016/0167-2789(93)90120-P\\
+\textbf{[2]} Zhao, Y and Hou, H (2013). Simulation of Dendritic Crystal Growth of Pure Ni using the Phase-Field Model. \textit{Rev. Adv. Mater. Sci.}, 33, 246-250\\
+\textbf{[3]} Sanal, R (2014).
+Numerical Simulation of Dendritic crystal growth using phase field method and investigating the effects of different physical parameter on the growth of the dendrite.
+UBid 50094548.
+Arxiv: http://arxiv.org/ftp/arxiv/papers/1412/1412.3197.pdf\\
+\textbf{[4]} Biben, T. (2005) Phase-Field Models for Free-Boundary Problems. \textit{European Journal of Physics}, 26, 47-55. http://dx.doi.org/10.1088/0143-0807/26/5/S06
+\textbf{[5]}Shah, A \textit{et.
+al} (2014).
+Numerical Simulation of Two-Dimensional Dendritic Growth Using Phase-Field Model.
+\textit{World Journal of Mechanics}, 128-136.
+http://dx.doi.org/10.4236/wjm.2014.45015
 \end{document}

### 2015-12-12 comp_model/crystal_phase.c

@@ -7,8 +7,8 @@
 #include <unistd.h>
 
 const double m_pi =  3.14159265358979323846;
-const int size = 300;
-const int animate = 0;
+const int size = 350;
+const int animate = 1;
 const int skiprate = 1000;
 const int debug = 0;
 
@@ -37,20 +37,22 @@ int main() {
   double dx = 0.03; //m
   double dy = 0.03; //m
   double dt = 0.0001; //s
-  double t_final = 0.4; //s
+  double t_final = 0.2; //s
   
   double tau = 0.0003;
   double delta_bar = 0.01; // Average thickness of layer (?)
-  double F = 2.0; // Latent heat of fusion
+  double F = 0.8; // Latent heat of fusion
 
   double mu = 0.02; // Strength of anisotropy
-  double anisotropy = 4.0;
+  double anisotropy = 6.0;
   double beta = 0.9; // n-shifting coefficient
   double eta = 10.0; // n-shifting coefficient
   double TM = 1.0;
   double T0 = 0.0;
   double theta_0 = 1.57;
-  int r = 10;
+  int r = 3;
+
+  double eccentricity = 1.0;
 
   int fd;
   if ((fd = open("data.txt", O_WRONLY | O_CREAT | O_TRUNC, S_IRUSR | S_IWUSR | S_IRGRP | S_IWGRP | S_IWOTH | S_IROTH)) == -1)
@@ -87,7 +89,7 @@ int main() {
       u_new[x+y*size] = 0;
       phi_grad_angle[x+y*size] = 0;
 
-      if ((x-size/2.0)*(x-size/2.0) + (y-size/2.0)*(y-size/2.0) < r*r) {
+      if (eccentricity*(x-size/2.0)*(x-size/2.0) + (y-size/2.0)*(y-size/2.0) < r*r) {
 	phi[x+y*size] = 1.0;
       }
     }

### 2015-12-12 comp_model/plot.py

@@ -35,6 +35,12 @@ if  len(data.shape) == 1:
     sys.exit(0)
     
 else:
+    Z = np.reshape(data[0], (math.sqrt(len(data[0])), math.sqrt(len(data[0]))))
+    cax = plt.imshow(Z)
+
+    cbar = plt.colorbar(cax)#, ticks=[-1, 0, 1])
+    #cbar.ax.set_yticklabels(['< -1', '0', '> 1'])  # vertically oriented colorbar
+        
     while i < len(data):
         print str(i) + " of " + str(len(data))
         Z = np.reshape(data[i], (math.sqrt(len(data[i])), math.sqrt(len(data[i]))))

### 2015-12-12 comp_model/latex/paper2.tex
done with term?

@@
 
-\title{Exploring Theory and Anisotropy Effects in Dendritic Phase-Field Model for Crystal Formation}
+\title{Exploring Parameter Effects in the Dendritic Phase-Field Model for Crystal Formation: Part II}
 \author{Elliott Capek}
@@
 The larger K is, the greater the heat evolved from the solidifying crystal.
-Our results for crystal growth at different latent heats, shown in Figure ($\ref{fig:latent-heat}$), agrees with Shah [5].
+Our results for crystal growth at different latent heats, shown in Figure ($\ref{fig:latent-heat}$), agree with Shah [5].
 The general effects we observe are an increase in anisotropic effects with increase in latent heat, and a decrease in growth speed and crystal size with latent heat.
@@
 If latent heat is low, then little head is added to the system via crystallization and all angles of $\nabla\phi$ can crystallize.
-While the $\delta(\theta)$ term does add an energy cost to certain angles, we can see from Figure ($\ref{fig:latent-heat}.a$) that minimizing potential f, which heavily favors the solid phase at low temperatures, outweighs the cost of anisotropic angles.\\
+While the $\delta(\theta)$ term does add an energy cost to certain angles, we can see from Figure ($\ref{fig:latent-heat}.a$) that minimizing potential f, which heavily favors the solid phase at low temperatures, outweighs the cost of certain angles.\\
 
@@
 
-Figure ($\ref{fig:latent-heat-u}$) shows the temperature field for K=2.0.
+Figure ($\ref{fig:latent-heat}$) shows the temperature field for K=2.0.
 From this we can see that dimensionless temperature is fairly uniform around the boundary, at a value of about -0.9.
@@
 
-where $\delta_0$ is the average layer width, $\mu$ is the effect of anisotropy and $\theta_0$ is the rotation of the crystal.
+where $\delta_0$ is the average layer width, $\mu$ is the strength of anisotropy and $\theta_0$ is the rotation of the crystal.
 Anisotropy is the most obvious parameter to change in this model.
@@
 This can be explained by looking at Equation ($\ref{eq:kobayoshi-delta}$).
-When anisotropy $a$ is low, angles near the minima are less expensive than when anisotropy is high.
-This means that, while the system will still tend to maximize the lowest-cost angles, those near it will be more available, and thus the final crystal will have a wider variance in angles.\\
+When anisotropy $a$ is low, angles near the minima are less expensive than they would be when anisotropy is high.
+This means that, while the system will still tend to maximize the lowest-cost angles, those near it will be more available, and thus the final crystal will have a wider variance in angles.
+The cosine function for lower modes is ``broader'', so the angles immediately near the minimum angles, which are dramatically more accessible to the system than those far from minima, are cheaper and therefore more represented.\\
 
@@
   \centering
-  \textbf{a. }\includegraphics[width=0.2\textwidth]{../radial-anisotropy-4.png}
-  \hspace{1cm}\textbf{b. }\includegraphics[width=0.2\textwidth]{../anis-4.png}
-  \caption{\textbf{a.} Values of $\delta(\theta)$, where $\theta$ is the angular polar coordinate of points in the XY plane.
+  \textbf{a. }\includegraphics[width=0.25\textwidth]{../radial-anisotropy-4.png}
+  \hspace{1cm}\textbf{b. }\includegraphics[width=0.25\textwidth]{../anis-4.png}
+  \caption{\textbf{a.} Values of $\delta(\theta)$, where $\theta$ is the angular polar coordinate, translated from the rectangular coordinates centered in the middle, of points in the XY plane.
 Darker values correspond to larger values of $\delta$.
-\textbf{b.} The resulting crystal, plotted for the following parameters: $a=4$,$dx=0.0003$, $F=1.8$ and $t=0.2$.}
+\textbf{b.} The resulting crystal, plotted for the following parameters: $a=4$, $dx=0.0003$, $F=1.8$ and $t=0.2$.}
   \label{fig:r-delta-4}
@@
   \centering
-  \textbf{a. }\includegraphics[width=0.2\textwidth]{../radial-anisotropy-3.png}
-  \hspace{1cm}\textbf{b. }\includegraphics[width=0.2\textwidth]{../anis-3.png}
+  \textbf{a. }\includegraphics[width=0.25\textwidth]{../radial-anisotropy-3.png}
+  \hspace{1cm}\textbf{b. }\includegraphics[width=0.25\textwidth]{../anis-3.png}
   \caption{$\delta$ and $\phi$, this time for $a=3$.}
@@
   \centering
-  \textbf{a. }\includegraphics[width=0.2\textwidth]{../radial-anisotropy-6.png}
-  \hspace{1cm}\textbf{b. }\includegraphics[width=0.2\textwidth]{../anis-6.png}
+  \textbf{a. }\includegraphics[width=0.25\textwidth]{../radial-anisotropy-6.png}
+  \hspace{1cm}\textbf{b. }\includegraphics[width=0.25\textwidth]{../anis-6.png}
   \caption{$\delta$ and $\phi$, this time for $a=6$.}
@@
 Increasing average layer thickness $\delta_0$ in Equation ($\ref{eq:kobayoshi-delta}$) effectively increases the magnitude of the anisotropy term in our free energy definition.
-We see this pattern reflected in Figure ($\ref{fig:dleta}$).
-The highest $\delta_0$ crystal on the right seems to have the most ``anisotropy'' - it has the most offshoots, all travelling in the same direction (a minimum-energy angle in Equation ($\ref{eq:delta-6}$)).
+We see this pattern reflected in Figure ($\ref{fig:delta}$).
+The highest $\delta_0$ crystal on the right seems to have the most ``anisotropy'' - it has the most offshoots, all travelling in the same direction (a minimum-energy angle in Figure ($\ref{fig:r-delta-6}$).a).
+The large non-phase regions at the edges of the main dendrites, the green regions between $\phi=0$ and $\phi=1$, can be explained by accepting that at high values of anisotropy the penalties of being at a non-phase $\phi$ value imposed by potential f are less important, since the anisotropy term has a larger magnitude.
 However, ``increased anisotropy'' doesn't fully describe the weirdness that arises in the $\delta_0 = 0.11$ crystal.
 It isn't intuitive why holes would show up in the middle of crystals.
-The large non-phase regions at the edges of the main dendrites, the green regions between $\phi=0$ and $\phi=1$ can be explained by accepting that at high values of anisotropy the penalties of being at a non-phase $\phi$ value imposed by potential f are less important, since the anisotropy term is being minimized.\\
+This is a yet unknown mystery of the $\delta_0$ parameter.\\
 
@@
   \centering
-  \includegraphics[width=0.15\textwidth]{../d-0.008-5.png}
-  \includegraphics[width=0.15\textwidth]{../d-0.09-5.png}
-  \includegraphics[width=0.15\textwidth]{../d-0.1-5.png}
-  \includegraphics[width=0.15\textwidth]{../d-0.11-5.png}
+  \includegraphics[width=0.2\textwidth]{../d-0.008-5.png}
+  \includegraphics[width=0.2\textwidth]{../d-0.09-5.png}
+  \includegraphics[width=0.2\textwidth]{../d-0.1-5.png}
+  \includegraphics[width=0.2\textwidth]{../d-0.11-5.png}
   \caption{Effect of boundary width on crystal formation. $\delta$ values of 0.008, 0.009, 0.1 and 0.11 from left to right.}
@@
 \section{Variations on Model}
-\subsection*{Zhao Model}
+\subsection{Zhao Model}
 The Zhao model, described in [2], is most notable because of its use of noise to simulate thermal fluctuations in the crystal formation. The evolution equation for $\phi$ is given by:
@@
 
-The most important part of the Zhao model is that it introduces randomness into the deterministic model.
+The most important part of the Zhao model is that it introduces randomness into the deterministic Kobayashi model.
 The $f_0$ term on the right is a Gaussian-sampled zero-centered term which randomly speeds up or slows down crystal formation.
@@
 When the discussed material is particularly inert, we can make the Local Density Approximation and discount the effects of surface tension.
-Thus, the only energy we need to worry about is our potential, which can be given by a simple polynomial, such as the equilibrium case $f_{eq} = (1-\phi^2)^2/4$, as given in Biben [4], which gives more on this topic.
+Thus, the only energy we need to worry about is our potential, which can be given by a simple polynomial, such as the equilibrium case $f_{eq} = (1-\phi^2)^2/4$, as given in Biben [4].
 In the case of this model, already existing boundaries would solidify, but extra terms are needed to actually cause crystal formation.\\
@@
 Arxiv: http://arxiv.org/ftp/arxiv/papers/1412/1412.3197.pdf\\
-\textbf{[4]} Biben, T. (2005) Phase-Field Models for Free-Boundary Problems. \textit{European Journal of Physics}, 26, 47-55. http://dx.doi.org/10.1088/0143-0807/26/5/S06
-\textbf{[5]}Shah, A \textit{et.
+\textbf{[4]} Biben, T. (2005) Phase-Field Models for Free-Boundary Problems. \textit{European Journal of Physics}, 26, 47-55. http://dx.doi.org/10.1088/0143-0807/26/5/S06\\
+\textbf{[5]} Shah, A \textit{et.
 al} (2014).

### 2016-01-12 quantum/latex/a1.tex
Add quantum

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{marvosym} % Required for symbols in the colored box
\usepackage{ifsym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\usepackage{physics}
\usepackage{amssymb}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Quantum Mechanics Assignment One}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem 2.17}
This problem asks us to take the following spin-1 state and figure out the probabilities of
measuring z- and x-component spins, and also to find their expectation value.\\

\begin{align*}
  \ket{\Psi} \doteq &\frac{1}{\sqrt{30}}
  \begin{pmatrix}
    1\\
    2\\
    5i\\
  \end{pmatrix}
\end{align*}

We use the normal $P_x = \abs{\braket{\Psi_x}{\Psi}}^2$ to find probabilities. It is simple to
calculate the braket since both states are in the $S_z$ representation.

\begin{align*}
  P_{1} = \abs{\braket{\Psi_{1}}{\Psi}}^2 &=
  \abs{\frac{1}{\sqrt{30}}\begin{pmatrix}1&0&0\end{pmatrix}\begin{pmatrix}1\\2\\5i\\\end{pmatrix}}^2
  = \frac{1}{30}\\
  P_{0} = \abs{\braket{\Psi_{0}}{\Psi}}^2 &=
  \abs{\frac{1}{\sqrt{30}}\begin{pmatrix}0&1&0\end{pmatrix}\begin{pmatrix}1\\2\\5i\\\end{pmatrix}}^2
  = \frac{4}{30}\\
  P_{-1} = \abs{\braket{\Psi_{-1}}{\Psi}}^2 &=
  \abs{\frac{1}{\sqrt{30}}\begin{pmatrix}0&0&1\end{pmatrix}\begin{pmatrix}1\\2\\5i\\\end{pmatrix}}^2
  = \frac{25}{30}\\
\end{align*}
\begin{align*}
  \expval{S_z}{\Psi} &= \frac{1}{30}\begin{pmatrix}1&2&-5i\end{pmatrix}\begin{pmatrix}
      \hslash&0&0\\0&0&0\\0&0&-\hslash\\\end{pmatrix}\begin{pmatrix}1\\2\\5i\end{pmatrix}
        = \frac{1}{30}\begin{pmatrix}1&2&-5i\end{pmatrix}
          \begin{pmatrix}\hslash\\0\\5\hslash i\end{pmatrix}
            = \frac{1}{30}\left(1 + -25\right) = \frac{-4}{5}\\
\end{align*}

We then want to compute the expectation value of the spin in the x-direction. We use the
$S_z$ representation of $S_x$:

\begin{align*}
  \expval{S_x}{\Psi} &= \frac{1}{30}\begin{pmatrix}1&2&-5i\end{pmatrix}\frac{\hslash}{\sqrt{2}}
    \begin{pmatrix}0&1&0\\1&0&1\\0&1&0\\\end{pmatrix}\begin{pmatrix}1\\2\\5i\end{pmatrix}
        = \frac{1}{30\sqrt{2}}\begin{pmatrix}1&2&-5i\end{pmatrix}
          \begin{pmatrix}2\\1+5i\\2\end{pmatrix}\\
            = \frac{4}{30\sqrt{2}}\\
\end{align*}

\section{Problem 5.11}
This problem deals with an infinite square well wave which suddenly has its right potential wall
move from $L$ to $3L$. We are asked to find the probability of finding the wave in the ground or
first excited state. This problem is basically asking us to find the $c_1$ and $c_2$ coefficients
of the old wave function in the new well, and from them find probabilities.

\begin{align*}
  \Psi(x) &= \sqrt{\frac{2}{L}}\sin\left(\frac{\pi x}{L}\right)\\
  \ket{E_1} &= \sqrt{\frac{2}{3L}}\sin\left(\frac{\pi x}{3L}\right)\\
  \ket{E_2} &= \sqrt{\frac{2}{3L}}\sin\left(\frac{2\pi x}{3L}\right)\\
\end{align*}

We wish to find the $c_1$ and $c_2$ coefficients for this system.

\begin{align*}
  c_1 &= \braket{E_1}{\Psi} = \int_0^{3L} \sqrt{\frac{2}{3L}}\sin\left(\frac{\pi x}{3L}\right)*
  \sqrt{\frac{2}{L}}\sin\left(\frac{\pi x}{L}\right)dx\\
  &= \sqrt{\frac{4}{3L^3}}\int_0^{3L}
  \sin\left(\frac{\pi x}{3L}\right)\sin\left(\frac{\pi x}{L}\right)\\
  &= \sqrt{\frac{2}{3L^3}}\left(\int_0^{3L} \cos\left(\frac{2\pi x}{3L}\right)dx
  - \int_0^{3L} \cos\left(\frac{4\pi x}{3L}\right)dx\right)\\
  &= 0\\
\end{align*}

\begin{align*}
  c_2 &= \braket{E_2}{\Psi} = \int_0^{3L} \sqrt{\frac{2}{3L}}\sin\left(\frac{2\pi x}{3L}\right)*
  \sqrt{\frac{2}{L}}\sin\left(\frac{\pi x}{L}\right)dx\\
  &= \sqrt{\frac{4}{3L^3}}\int_0^{3L}
  \sin\left(\frac{2\pi x}{3L}\right)\sin\left(\frac{\pi x}{L}\right)\\
  &= \sqrt{\frac{2}{3L^3}}\left(\int_0^{3L} \cos\left(\frac{\pi x}{3L}\right)dx
  - \int_0^{3L} \cos\left(\frac{5\pi x}{3L}\right)dx\right)\\
  &= 0\\
\end{align*}

From this we can see that the probability of measuring the wave in the new tripled well to be in
$\ket{E_1}$ or $\ket{E_2}$ is zero, since $P_n = \abs{c_n}^2$. This result is to be expected, since
the wave should be completely in the second excited state, $\ket{E_3}$. This is because extending
the $\Psi(x)$ wave function from $0$ to $3L$ results in a two-node wave function identical to the
$\ket{E_3}$ state.\\

\section{8.7}

\end{document}

### 2016-01-13 comp_model/crystal_phase.c
Add everything.

@@ -7,9 +7,9 @@
 #include <unistd.h>
 
 const double m_pi =  3.14159265358979323846;
-const int size = 350;
+const int size = 250;
 const int animate = 1;
-const int skiprate = 1000;
+const int skiprate = 10;
 const int debug = 0;
 
 void log_data(double* data, int fd) {
@@ -41,10 +41,10 @@ int main() {
   
   double tau = 0.0003;
   double delta_bar = 0.01; // Average thickness of layer (?)
-  double F = 0.8; // Latent heat of fusion
+  double F = 1.8; // Latent heat of fusion
 
   double mu = 0.02; // Strength of anisotropy
-  double anisotropy = 6.0;
+  double anisotropy = 4.0;
   double beta = 0.9; // n-shifting coefficient
   double eta = 10.0; // n-shifting coefficient
   double TM = 1.0;

### 2016-01-20 quantum/latex/a2.tex
stuff.

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{marvosym} % Required for symbols in the colored box
\usepackage{ifsym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\usepackage{physics}
\usepackage{amssymb}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Quantum Mechanics Assignment Two}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{9.4: Forbidden region probability for QHO}
For this problem we want to calculate the probability of measuring a QHO ground-state object outside
of the classically allowed region. We do this by calculating the turning points of a classical HO,
ie finding the x-values where energy equals potential, and then integrating $\psi^*\psi$ over the
forbidden region to find the chance of forbidden observation.\\

We first find the turning point for the ground state $n=0$:
\begin{align*}
  V_{HO} &= E_0\\
  \frac{1}{2}m\omega^2x^2 &= \hbar\omega\left(0+\frac{1}{2}\right)\\
  x_0 &= \sqrt{\frac{\hbar}{m\omega}}\\
\end{align*}

We then find the probability of finding the particle in this region:

\begin{align*}
  P_{forb} &= \int_{\infty}^{-x_0} \psi_0^*\psi_0 dx + \int_{x_0}^{\infty} \psi_0^*\psi_0 dx\\
  &= \int_{\infty}^{-x_0} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar}dx
  + \int_{x_0}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar} dx\\
  &= \sqrt{\frac{m\omega}{\pi\hbar}} \left(\frac{-\hbar}{m\omega}\right)\left(
    \frac{1}{x}e^{-m\omega x^2 / \hbar} \Bigg|_{-\infty}^{-\sqrt{\frac{\hbar}{m\omega}}} +
    \frac{1}{x}e^{-m\omega x^2 / \hbar} \Bigg|_{\sqrt{\frac{\hbar}{m\omega}}}^{\infty}\right)\\
  &= \frac{2}{\sqrt{\pi}} = 0.41\\
\end{align*}

\section{9.9: QHO expectation values via integration, operator method}
In this problem we use both wave function integration and the operator method to find the
expectation values of various observables: x, p, $x^2$, $p^2$.

\textbf{a.) Ground state wave function calculations}
\begin{align*}
  <x> &= \int_{-\infty}^{\infty} \psi^*x\psi dx\\
  &= \int_{-\infty}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar}x dx = 0\\
\end{align*}
We don't need to perform the $<x>$ integral because it involves integrating an odd function over
an even domain, which always results in zero. This is to be expected, because the wave ground-state
wave function is symmetric across zero.\\

\begin{align*}
  <p> &= \int_{-\infty}^{\infty} \psi^* -i\hbar \frac{d}{dx}\psi dx\\
  &= \int_{-\infty}^{\infty}
  -i\sqrt{\frac{m\omega}{\pi\hbar}}(-m\omega x)e^{\frac{-m\omega x^2}{2\hbar}}dx = 0\\
\end{align*}
This integral is zero since we are again integrating an odd function over an even domain. This
makes sense because, since the spacial distribution of the wave function is equal on both sides
of the x-axis, the velocities (and hence momenta) should be equal as well.\\

\begin{align*}
  <x^2> &= \int_{-\infty}^{\infty} \psi^*x^2\psi dx\\
  &= \int_{-\infty}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar}x^2 dx\\
  &= 2\sqrt{\frac{m\omega}{\pi\hbar}}\int_{0}^{\infty} e^{-m\omega x^2 / \hbar}x^2 dx\\
  &= 2\sqrt{\frac{m\omega}{\pi\hbar}} *
  \frac{\sqrt{\pi}}{4(\frac{m\omega}{\hbar})^{3/2}} \hspace{2cm}\mbox{Wolfram}\\
  &= \frac{\hbar}{2m\omega}\\
\end{align*}
Here we exploit the fact that integrating an even function over an even domain is equivalent to
doubling the integral of the function over half the domain. The answer makes sense, since increasing
$m\omega$ corresponds to making the quadratic potential steeper, hence making the wave function
tighter.\\

\begin{align*}
  <p^2> &= \int_{-\infty}^{\infty} \psi^* \left(-\hbar^2 \frac{d^2}{dx^2}\right)\psi dx\\
  &= \sqrt{\frac{m\omega}{\pi\hbar}}\int_{-\infty}^{\infty} -\hbar^2e^{\frac{-m\omega x^2}{\hbar}}
  \left(\frac{m^2\omega^2x^2}{\hbar^2} - \frac{m\omega x}{\hbar}\right)dx\\
  &= -\hbar^2\sqrt{\frac{m\omega}{\pi\hbar}}\left(
  \int_{-\infty}^{\infty} e^{\frac{-m\omega x^2}{\hbar}}\frac{m^2\omega^2x^2}{\hbar^2}dx -
  \int_{-\infty}^{\infty} e^{\frac{-m\omega x^2}{\hbar}}\frac{m\omega}{\hbar}\right)dx\\
  &= -\hbar^2\sqrt{\frac{m\omega}{\pi\hbar}}
  \left(\frac{1}{4}\sqrt{\frac{m\pi\omega}{\hbar}} -
  \frac{1}{2}\sqrt{\frac{m\pi\omega}{\hbar}}\right) \hspace{2cm}\mbox{Wolfram}\\
  &= \frac{\hbar m\omega}{4}\\
\end{align*}
This answer is intuitive, since increasing mass as well as oscillatory speed should both increase
the expected momentum.

\textbf{b.) Any-state operator method calculations}
Here we define the position and momentum operators in terms of the raising and lowering operators:
\begin{align*}
  \hat{x} &= \sqrt{\frac{\hbar}{2m\omega}}\left(a + a^\dagger\right)\\
  \hat{p} &= i\sqrt{\frac{m\omega\hbar}{2}}\left(a^\dagger - a\right)\\
\end{align*}
We then solve for the various expectation values using these definitions:

\begin{align*}
  <x> &= \bra{n}x\ket{n}\\
  &= \sqrt{\frac{\hbar}{2m\omega}}\bra{n}\left(a + a^\dagger\right)\ket{n}\\
  &= \sqrt{\frac{\hbar}{2m\omega}}
  \left(\sqrt{n}\bra{n}\ket{n-1} + \sqrt{n+1}\bra{n}\ket{n+1}\right) = 0\\
\end{align*}
This answer indicates that all pure QHO wave functions are centered around zero. This makes
sense since the QHO wave functions are all made up of either even or odd functions. When squared,
as is done when finding the probability, the equation becomes entirely even, meaning it is
symmetric across the y-axis. From this we can see that all QHO wave functions have an expectation
value of zero for x-measurement.

\begin{align*}
  <p> &= \bra{n}\hat{p}\ket{n}\\
  &= i\sqrt{\frac{m\hbar\omega}{2}}\bra{n}\left(a^\dagger-a\right)\ket{n}\\
  &= i\sqrt{\frac{m\hbar\omega}{2}}
  \left(\sqrt{n+1}\bra{n}\ket{n+1} - \sqrt{n}\bra{n}\ket{n-1}\right) = 0\\
\end{align*}
This is an interesting result. It is intuitive that any wave function should have an expectation of
zero for momentum, else the wave function would move through space.

How does this reconcile w/ fact that p involves a spacial derivative?

\end{document}

### 2016-01-20 classical/latex/a9.tex

@@
   \vec{r_1} &= (a, 2a, 0)\\
-  m_2 &= 5m\\
+  m_2 &= 2m\\
   \vec{r_2} &= (0, -2a, a)\\
-  m_3 &= 5m\\
+  m_3 &= 3m\\
   \vec{r_3} &= (a, a, -a)\\
-  m_4 &= 5m\\
+  m_4 &= m\\
   \vec{r_4} &= (3a, a, -2a)\\
@@
 
+\begin{align*}
+  I &=
+  \begin{pmatrix}
+    \int_V \rho(y^2+z^2)dV & -\int_V xydV & -\int_V xzdV\\
+     -\int_V xydV & \int_V \rho(x^2+z^2)dV & -\int_V yzdV\\
+     -\int_V xzdV & -\int_V yzdV & \int_V \rho(x^2+y^2)dV\\
+  \end{pmatrix}\\
+  I &=
+  \begin{pmatrix}
+    M\sum (y_i^2+z_i^2) & -\sum x_iy_i & -\sum x_iz_i\\
+     -\sum x_iy_i & M\sum (x_i^2+z_i^2) & -\sum y_iz_i\\
+     -\sum x_iz_i & -\sum y_iz_i & M\sum (x_i^2+y_i^2)\\
+  \end{pmatrix}\\
+  I &=
+  Ma^2
+  \begin{pmatrix}
+    41 & -16 & 9\\
+    -16 & 26 & 9\\
+    9 & 9 & 49\\
+  \end{pmatrix}
+\end{align*}
+
+We now want to find the solutions to the eigenvalue equation:
+
+\begin{align*}
+  &\det
+  \begin{pmatrix}
+    Ma^241-\lambda & -Ma^216 & Ma^29\\
+    -Ma^216 & Ma^226-\lambda & Ma^29\\
+    Ma^29 & Ma^29 & Ma^249-\lambda\\
+  \end{pmatrix} = 0\\
+  \lambda_1 &= 54.9631\hspace{0.1cm}Ma^2\\
+  \lambda_2 &= 49.3639\hspace{0.1cm}Ma^2\\
+  \lambda_3 &= 11.673\hspace{0.1cm}Ma^2\\
+\end{align*}
+
+These are our moments of inertia. We can then find our angular eigenvectors, which represent our principal axes:
+
+\begin{align*}
+  \omega_1 &=
+  \begin{pmatrix}
+    0.62\\
+    -0.1\\
+    0.78\\
+  \end{pmatrix}\hspace{1cm}
+  \omega_2 =
+  \begin{pmatrix}
+    -0.59\\
+    0.6\\
+    0.54\\
+  \end{pmatrix}\hspace{1cm}
+  \omega_3 =
+  \begin{pmatrix}
+    -0.53\\
+    -0.79\\
+    0.32\\
+  \end{pmatrix}
+\end{align*}
+
+We now verify this is an orthogonal basis:
+
+\begin{align*}
+  \omega_1 \cdot \omega_2 = \omega_1 \cdot \omega_3 = \omega_2 \cdot \omega_3 \approx 0\\
+\end{align*}
+
+This was a cool problem.
+I think it's really interesting that all objects have three principal axes.
+This is not at all intuitive.
+I can see that all objects have at least one principal axis.
+And by this, there must be two vectors orthogonal to it, so it makes sense given that assumption that there are three principal axes.
+This is still unintuitive, though.
+This problem involved using linear algebra solvers online since the actual eigenvalues were so ugly.
+But it was a good exercise.
+It was interesting to take the continuous form of the I equation and discretize it.\\
+
 \end{document}

### 2016-01-20 quantum/latex/a1.tex
Add everything.
add a1.work
add all

@@
 \section{Problem 5.11}
+This problem asks us to find the probabilities to find $n=2$ Hydrogen electrons in the classically
+forbidden region.\\
+
 This problem deals with an infinite square well wave which suddenly has its right potential wall
@@
 \section{8.7}
+We begin with the potential and find the r-values at which $E=V$. We then calculate
+the probability using $P = \int_r^\infty \psi^* \psi dr$.
+
+\begin{align*}
+  V(r) &= \frac{-e^2}{4\pi\epsilon_0r}\\
+\end{align*}
+
+We first solve for the forbidden distance:
+
+\begin{align*}
+  V(r) &= E_2\\
+  -\frac{Z\alpha\hslash c}{r} &= -\frac{1}{2n^2}\alpha^2m_ec^2\\
+  \frac{-1.44 \mbox{eV nm}}{r} &= -3.4 \mbox{eV}\\
+  r &= 2.36 \mbox{nm}
+\end{align*}
+
+Next we find the probability of our particles being further than this:
+
+\begin{align*}
+  P_{200} &= \int_{2.36}^\infty \int_0^{2\pi} \int_0^\pi
+  \frac{r^2}{32\pi a_0^3}\left(2-\frac{r}{a_0}\right)^2e^{\frac{-r}{a_0}}
+  \sin(\theta)drd\phi d\theta = 2.09*10^{-14}\\
+  P_{21-1} &= \int_{2.36}^\infty \int_0^{2\pi} \int_0^\pi
+  \frac{r^4}{64\pi a_0^5}e^{\frac{-r}{a_0}}\sin^3(\theta)
+  drd\phi d\theta = 7.62*10^{-15}\\
+  P_{210} &= \int_{2.36}^\infty \int_0^{2\pi} \int_0^\pi
+  \frac{r^4}{32\pi a_0^5}e^{\frac{-r}{a_0}}\cos^2(\theta)
+  \sin(\theta)drd\phi d\theta = 7.62*10^{-15}\\
+  P_{211} &= \int_{2.36}^\infty \int_0^{2\pi} \int_0^\pi
+  \frac{r^4}{64\pi a_0^5}e^{\frac{-r}{a_0}}\sin^3(\theta)
+  drd\phi d\theta = 7.62*10^{-15}\\
+\end{align*}
+
+We note that it is much more likely for the s-shell electron to be found in the forbidden (far
+from atom) zone than the p-shell electrons. This is indicative of the shape of the s vs p shells
+and how spread-out they are from the nucleus. We also note that, though we started off with 3
+different wave functions for the $\ell = 1$ states, we get the same probabilities in the end.
+This is to be expected, since the different $m_\ell$-numbered states are just rotations of each
+other around the atom's axis, their shapes are the same.\\
+
+I was initially confused as to why the units of the wave function were in $m^{-3/2}$, but then
+recalled that complex conjugating the wave function and integrating over three dimensions would
+cancel out these units.\\
 

### 2016-01-20 quantum/latex/a2.tex
Add changes from last nite

@@
 \usepackage{marvosym} % Required for symbols in the colored box
-\usepackage{ifsym} % Required for symbols in the colored box
 
@@
 \usepackage{fancyhdr}
-\usepackage{physics}
+\usepackage{amsmath}
+% \usepackage{physics}
 \usepackage{amssymb}
@@
 
-\section{9.9: QHO expectation values via integration, operator method}
-In this problem we use both wave function integration and the operator method to find the
-expectation values of various observables: x, p, $x^2$, $p^2$.
+%% \section{9.9: QHO expectation values via integration, operator method}
+%% In this problem we use both wave function integration and the operator method to find the
+%% expectation values of various observables: x, p, $x^2$, $p^2$.
 
-\textbf{a.) Ground state wave function calculations}
+%% \textbf{a.) Ground state wave function calculations}
+%% \begin{align*}
+%%   <x> &= \int_{-\infty}^{\infty} \psi^*x\psi dx\\
+%%   &= \int_{-\infty}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar}x dx = 0\\
+%% \end{align*}
+%% We don't need to perform the $<x>$ integral because it involves integrating an odd function over
+%% an even domain, which always results in zero. This is to be expected, because the wave ground-state
+%% wave function is symmetric across zero.\\
+
+%% \begin{align*}
+%%   <p> &= \int_{-\infty}^{\infty} \psi^* -i\hbar \frac{d}{dx}\psi dx\\
+%%   &= \int_{-\infty}^{\infty}
+%%   -i\sqrt{\frac{m\omega}{\pi\hbar}}(-m\omega x)e^{\frac{-m\omega x^2}{2\hbar}}dx = 0\\
+%% \end{align*}
+%% This integral is zero since we are again integrating an odd function over an even domain. This
+%% makes sense because, since the spacial distribution of the wave function is equal on both sides
+%% of the x-axis, the velocities (and hence momenta) should be equal as well.\\
+
+%% \begin{align*}
+%%   <x^2> &= \int_{-\infty}^{\infty} \psi^*x^2\psi dx\\
+%%   &= \int_{-\infty}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar}x^2 dx\\
+%%   &= 2\sqrt{\frac{m\omega}{\pi\hbar}}\int_{0}^{\infty} e^{-m\omega x^2 / \hbar}x^2 dx\\
+%%   &= 2\sqrt{\frac{m\omega}{\pi\hbar}} *
+%%   \frac{\sqrt{\pi}}{4(\frac{m\omega}{\hbar})^{3/2}} \hspace{2cm}\mbox{Wolfram}\\
+%%   &= \frac{\hbar}{2m\omega}\\
+%% \end{align*}
+%% Here we exploit the fact that integrating an even function over an even domain is equivalent to
+%% doubling the integral of the function over half the domain. The answer makes sense, since increasing
+%% $m\omega$ corresponds to making the quadratic potential steeper, hence making the wave function
+%% tighter.\\
+
+%% \begin{align*}
+%%   <p^2> &= \int_{-\infty}^{\infty} \psi^* \left(-\hbar^2 \frac{d^2}{dx^2}\right)\psi dx\\
+%%   &= \sqrt{\frac{m\omega}{\pi\hbar}}\int_{-\infty}^{\infty} -\hbar^2e^{\frac{-m\omega x^2}{\hbar}}
+%%   \left(\frac{m^2\omega^2x^2}{\hbar^2} - \frac{m\omega}{\hbar}\right)dx\\
+%%   &= -\hbar^2\sqrt{\frac{m\omega}{\pi\hbar}}\left(
+%%   \int_{-\infty}^{\infty} e^{\frac{-m\omega x^2}{\hbar}}\frac{m^2\omega^2x^2}{\hbar^2}dx -
+%%   \int_{-\infty}^{\infty} e^{\frac{-m\omega x^2}{\hbar}}\frac{m\omega}{\hbar}\right)dx\\
+%%   &= -\hbar^2\sqrt{\frac{m\omega}{\pi\hbar}}
+%%   \left(\frac{1}{2}\sqrt{\frac{m\pi\omega}{\hbar}} -
+%%   \sqrt{\frac{m\pi\omega}{\hbar}}\right) \hspace{2cm}\mbox{Wolfram}\\
+%%   &= \frac{\hbar m\omega}{2}\\
+%% \end{align*}
+%% This answer is intuitive, since increasing mass as well as oscillatory speed should both increase
+%% the expected momentum.
+
+%% \textbf{b.) Any-state operator method calculations}
+%% Here we define the position and momentum operators in terms of the raising and lowering operators:
+%% \begin{align*}
+%%   \hat{x} &= \sqrt{\frac{\hbar}{2m\omega}}\left(a + a^\dagger\right)\\
+%%   \hat{p} &= i\sqrt{\frac{m\omega\hbar}{2}}\left(a^\dagger - a\right)\\
+%% \end{align*}
+%% We then solve for the various expectation values using these definitions:
+
+%% \begin{align*}
+%%   <x> &= \bra{n}x\ket{n}\\
+%%   &= \sqrt{\frac{\hbar}{2m\omega}}\bra{n}\left(a + a^\dagger\right)\ket{n}\\
+%%   &= \sqrt{\frac{\hbar}{2m\omega}}
+%%   \left(\sqrt{n}\bra{n}\ket{n-1} + \sqrt{n+1}\bra{n}\ket{n+1}\right) = 0\\
+%% \end{align*}
+%% This answer indicates that all pure QHO wave functions are centered around zero. This makes
+%% sense since the QHO wave functions are all made up of either even or odd functions. When squared,
+%% as is done when finding the probability, the equation becomes entirely even, meaning it is
+%% symmetric across the y-axis. From this we can see that all QHO wave functions have an expectation
+%% value of zero for x-measurement.
+
+%% \begin{align*}
+%%   <p> &= \bra{n}\hat{p}\ket{n}\\
+%%   &= i\sqrt{\frac{m\hbar\omega}{2}}\bra{n}\left(a^\dagger-a\right)\ket{n}\\
+%%   &= i\sqrt{\frac{m\hbar\omega}{2}}
+%%   \left(\sqrt{n+1}\bra{n}\ket{n+1} - \sqrt{n}\bra{n}\ket{n-1}\right) = 0\\
+%% \end{align*}
+%% This is an interesting result. It is intuitive that any wave function should have an expectation of
+%% zero for momentum, else the wave function would move through space due to a nonzero average
+%% velocity.
+
+%% \begin{align*}
+%%   <x^2> &= \bra{n}x^2\ket{n}\\
+%%   &= \frac{\hbar}{2m\omega}\bra{n}\left(a + a^\dagger\right)\left(a + a^\dagger\right)\ket{n}\\
+%%   &= \frac{\hbar}{2m\omega}\bra{n}
+%%   \left(aa + aa^\dagger + a^\dagger a + a^\dagger a^\dagger\right)\ket{n}\\
+%%   &= \frac{\hbar}{2m\omega}
+%%   \left(aa + aa^\dagger + a^\dagger a + a^\dagger a^\dagger\right)\ket{n}\\
+%%   &= \frac{\hbar}{2m\omega} \left(\sqrt{n-1}\sqrt{n-2}\bra{n}\ket{n-2} + (n+1)\bra{n}\ket{n} +
+%%   (n)\bra{n}\ket{n} + \sqrt{n+1}\sqrt{n+2}\bra{n}\ket{n+2}\right)\\
+%%   &= \frac{\hbar}{2m\omega} \left(2n+1\right)\\
+%% \end{align*}
+%% This is an interesting result which indicates that for each extra rung on the ladder, the
+%% expectation of the squared position increases linearly, meaning the expectation of the magnitude
+%% of x increases as $\sqrt{n}$. This makes sense, since increasing energy levels have less and less
+%% room to expand in the quadratic potential well. The $n=0$ state corresponds to the wave function
+%% solution for the ground state found above.\\
+
+%% \begin{align*}
+%%   <p^2> &= \bra{n}p^2\ket{n}\\
+%%   &= -\frac{m\hbar\omega}{2}\bra{n}\left(a^\dagger-a\right)\left(a^\dagger-a\right)\ket{n}\\
+%%   &= -\frac{m\hbar\omega}{2}\bra{n}
+%%   \left(a^\dagger a^\dagger - a^\dagger a - aa^\dagger + aa\right)\ket{n}\\
+%%   &= -\frac{m\hbar\omega}{2}\bra{n}\left(\sqrt{n-1}\sqrt{n-2}\bra{n}\ket{n-2}
+%%   - (n)\bra{n}\ket{n} - (n+1)\bra{n}\ket{n} + \sqrt{n+1}\sqrt{n+2}\bra{n}\ket{n+2}\right)\\
+%%   &= \frac{\hbar m\omega}{2}\left(2n + 1\right)
+%% \end{align*}
+%% This result is similar to the above one for $x^2$ with different units. It is still a $\sqrt{n}$
+%% relationship for momentum, which makes sense.\\
+
+%% \textbf{c.) Uncertainty principle}\\
+%% Since the above two cases agree with each other, we check that the more general case of part
+%% \textbf{b.} agrees with the uncertainty principle.
+
+%% \begin{align*}
+%%   \Delta x &= \sqrt{<x^2>-<x>^2}\\
+%%   &= \sqrt{\frac{\hbar}{2m\omega} \left(2n+1\right)}\\
+%%   \Delta p &= \sqrt{<p^2>-<p>^2}\\
+%%   &= \sqrt{\frac{\hbar m\omega}{2}\left(2n + 1\right)}\\
+%%   \Delta x \Delta p &= \sqrt{\frac{\hbar}{2m\omega} \left(2n+1\right)}
+%%   \sqrt{\frac{\hbar m\omega}{2}\left(2n + 1\right)}\\
+%%   &= \frac{\hbar}{2}(2n+1)\\
+%% \end{align*}
+%% This is an interesting result which indicates that uncertainty increases as the pure wave function
+%% gets more energetic. Ground state is the state with maximum certainty. The higher states get
+%% progressively more uncertain, which makes sense since both the velocity and the well get larger
+%% for higher states.\\
+
+%% \section{9.14: Unknown QHO wavefunction}
+%% We are given the wave function $\psi(x,0) = A\left[1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
+%%   2\frac{m\omega}{\hbar}x^2\right]$. Because its highest-order polynomial term is $x^2$, we know
+%% that it must be made up of the $n = 0, 1, 2$ energy states. We now figure out the coefficients for
+%% each:
+
+%% \begin{align*}
+%%   c_0 &= \int_{-\infty}^{\infty} \phi_0^*\psi dx\\
+%%   &= A\int_{-\infty}^{\infty} \left(\frac{m\omega}{\pi\hbar}\right)^{1/2}e^{-m\omega x^2/\hbar}
+%%   \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
+%%   2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/(2\hbar)}dx\\
+%%   &= 2\\
+%%   c_1 &= A\int_{-\infty}^{\infty} \frac{2\sqrt{\frac{m\omega}{\hbar}x}}{\sqrt{2}}
+%%     \left(\frac{m\omega}{\pi\hbar}\right)^{1/2} \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
+%%     2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
+%%  &= -\frac{3}{\sqrt{2}}\\
+%%  c_2 &= A\int_{-\infty}^{\infty} \frac{\frac{4m\omega}{\hbar}x^2 - 2}{\sqrt{8}}
+%%     \left(\frac{m\omega}{\pi\hbar}\right)^{1/2} \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
+%%     2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
+%%  &= \sqrt{2}\\
+%% \end{align*}
+
+%% From this we can see that $\psi(x,0) = \sqrt{\frac{21}{2}}
+%% \left(2\ket{0} + -\frac{3}{\sqrt{2}}\ket{1} + \sqrt{2}\ket{2}\right)$. It is then easy to compute
+%% the expectation value: \\
+
+%% \begin{align*}
+%%   <E> &= \sigma \hbar\omega\left(n+\frac{1}{2}\right)|c_n|^2\\
+%%   &= \frac{2 + \frac{27}{4} + 5}{\frac{21}{2}} = 1.3 eV\\
+%% \end{align*}
+
+\section{6: Position, momentum commutator}
+Here we find the value of the position, momentum commutator:
 \begin{align*}
-  <x> &= \int_{-\infty}^{\infty} \psi^*x\psi dx\\
-  &= \int_{-\infty}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar}x dx = 0\\
+  [\hat{x}, \hat{p}] &= \hat{x}\hat{p} - \hat{p}\hat{x}\\
+  &= \left(\hat{x}\hat{p}\psi(x) - \hat{p}\hat{x}\psi(x)\right)\frac{1}{\psi(x)}\\
+  &= \left(\hat{x}\left(-i\hbar\frac{d}{dx}\right)\psi(x) - \hat{p}x\psi(x)\right)\frac{1}{\psi(x)}\\
+  &= \left(-i\hbar x \frac{d\psi}{dx} + i\hbar\left(\psi(x)
+  + x\frac{d\psi}{dx}\right)\right)\frac{1}{\psi(x)}\\
+  &= \left(i\hbar\psi(x)\right)\frac{1}{\psi(x)}\\
+  &= i\hbar\\
 \end{align*}
-We don't need to perform the $<x>$ integral because it involves integrating an odd function over
-an even domain, which always results in zero. This is to be expected, because the wave ground-state
-wave function is symmetric across zero.\\
+Here we use the definitions of the position and momentum operators to find the value of their
+commutator. We need to operate on a generic wave function in order to do algebra on the
+operators, since the differential in the momentum operator is meaningless without a function
+to act on. This result is critical, since it is at the heart of the Heisenberg Uncertainty
+Principle: position and momentum don't commute, so they do not share a basis and they are not
+co-measurable. This is the mathematical representation of the popular saying ``you cannot know
+the position and momentum of a particle simultaneously.''
+
+\section{7: Commutator algebra}
 
 \begin{align*}
-  <p> &= \int_{-\infty}^{\infty} \psi^* -i\hbar \frac{d}{dx}\psi dx\\
-  &= \int_{-\infty}^{\infty}
-  -i\sqrt{\frac{m\omega}{\pi\hbar}}(-m\omega x)e^{\frac{-m\omega x^2}{2\hbar}}dx = 0\\
+  \left[\hat{A},\hat{A}^n\right] &= AA^n - A^nA = A^{n+1} - A^{n+1} = 0\\
+  \left[\hat{A},\hat{B}\right] &= \hat{A}\hat{B} - \hat{B}\hat{A}
+  = -\left(\hat{B}\hat{A} - \hat{A}\hat{B}\right) = -\left[\hat{B},\hat{A}\right]\\
+  \left[\hat{A},c\hat{B}\right] &= \hat{A}c\hat{B} - c\hat{B}\hat{A}
+  = c\left(\hat{A}\hat{B} - \hat{B}\hat{A}\right) = c\left[\hat{A},\hat{B}\right]\\
+  \left[\hat{A},\left(\hat{B}+\hat{C}\right)\right] &= \hat{A}\left(\hat{B} + \hat{C}\right)
+  - \left(\hat{B} + \hat{C}\right)\hat{A} = \hat{A}\hat{B} - \hat{B}\hat{A}
+  + \hat{A}\hat{C} + \hat{C}\hat{A} = \left[\hat{A},\hat{B}\right] + \left[\hat{A},\hat{C}\right]\\
+  \left[\hat{A},\hat{B}\hat{C}\right] &= \hat{A}\hat{B}\hat{C} - \hat{B}\hat{C}\hat{A}
+  = \hat{A}\hat{B}\hat{C} - \hat{B}\hat{A}\hat{C} + \hat{B}\hat{A}\hat{C} - \hat{B}\hat{C}\hat{A}
+  = \left[\hat{A},\hat{B}\right]\hat{C} + \hat{B}\left[\hat{A},\hat{C}\right]\\
+  \left[\hat{A},\left[\hat{B},\hat{C}\right]\right]
+  + \left[\hat{B},\left[\hat{C},\hat{A}\right]\right]
+  &+ \left[\hat{C},\left[\hat{A},\hat{B}\right]\right] =
+  \left[\hat{A},\hat{B}\hat{C} - \hat{C}\hat{B}\right]
+  + \left[\hat{B},\hat{C}\hat{A} - \hat{A}\hat{C}\right]
+  + \left[\hat{C},\hat{A}\hat{B} - \hat{B}\hat{A}\right] = \\
+  &= \hat{A}\hat{B}\hat{C} - \hat{A}\hat{C}\hat{B}
+  + \hat{B}\hat{C}\hat{A} - \hat{B}\hat{A}\hat{C}
+  + \hat{C}\hat{A}\hat{B} - \hat{C}\hat{B}\hat{A}\\
 \end{align*}
-This integral is zero since we are again integrating an odd function over an even domain. This
-makes sense because, since the spacial distribution of the wave function is equal on both sides
-of the x-axis, the velocities (and hence momenta) should be equal as well.\\
-
-\begin{align*}
-  <x^2> &= \int_{-\infty}^{\infty} \psi^*x^2\psi dx\\
-  &= \int_{-\infty}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar}x^2 dx\\
-  &= 2\sqrt{\frac{m\omega}{\pi\hbar}}\int_{0}^{\infty} e^{-m\omega x^2 / \hbar}x^2 dx\\
-  &= 2\sqrt{\frac{m\omega}{\pi\hbar}} *
-  \frac{\sqrt{\pi}}{4(\frac{m\omega}{\hbar})^{3/2}} \hspace{2cm}\mbox{Wolfram}\\
-  &= \frac{\hbar}{2m\omega}\\
-\end{align*}
-Here we exploit the fact that integrating an even function over an even domain is equivalent to
-doubling the integral of the function over half the domain. The answer makes sense, since increasing
-$m\omega$ corresponds to making the quadratic potential steeper, hence making the wave function
-tighter.\\
-
-\begin{align*}
-  <p^2> &= \int_{-\infty}^{\infty} \psi^* \left(-\hbar^2 \frac{d^2}{dx^2}\right)\psi dx\\
-  &= \sqrt{\frac{m\omega}{\pi\hbar}}\int_{-\infty}^{\infty} -\hbar^2e^{\frac{-m\omega x^2}{\hbar}}
-  \left(\frac{m^2\omega^2x^2}{\hbar^2} - \frac{m\omega x}{\hbar}\right)dx\\
-  &= -\hbar^2\sqrt{\frac{m\omega}{\pi\hbar}}\left(
-  \int_{-\infty}^{\infty} e^{\frac{-m\omega x^2}{\hbar}}\frac{m^2\omega^2x^2}{\hbar^2}dx -
-  \int_{-\infty}^{\infty} e^{\frac{-m\omega x^2}{\hbar}}\frac{m\omega}{\hbar}\right)dx\\
-  &= -\hbar^2\sqrt{\frac{m\omega}{\pi\hbar}}
-  \left(\frac{1}{4}\sqrt{\frac{m\pi\omega}{\hbar}} -
-  \frac{1}{2}\sqrt{\frac{m\pi\omega}{\hbar}}\right) \hspace{2cm}\mbox{Wolfram}\\
-  &= \frac{\hbar m\omega}{4}\\
-\end{align*}
-This answer is intuitive, since increasing mass as well as oscillatory speed should both increase
-the expected momentum.
-
-\textbf{b.) Any-state operator method calculations}
-Here we define the position and momentum operators in terms of the raising and lowering operators:
-\begin{align*}
-  \hat{x} &= \sqrt{\frac{\hbar}{2m\omega}}\left(a + a^\dagger\right)\\
-  \hat{p} &= i\sqrt{\frac{m\omega\hbar}{2}}\left(a^\dagger - a\right)\\
-\end{align*}
-We then solve for the various expectation values using these definitions:
-
-\begin{align*}
-  <x> &= \bra{n}x\ket{n}\\
-  &= \sqrt{\frac{\hbar}{2m\omega}}\bra{n}\left(a + a^\dagger\right)\ket{n}\\
-  &= \sqrt{\frac{\hbar}{2m\omega}}
-  \left(\sqrt{n}\bra{n}\ket{n-1} + \sqrt{n+1}\bra{n}\ket{n+1}\right) = 0\\
-\end{align*}
-This answer indicates that all pure QHO wave functions are centered around zero. This makes
-sense since the QHO wave functions are all made up of either even or odd functions. When squared,
-as is done when finding the probability, the equation becomes entirely even, meaning it is
-symmetric across the y-axis. From this we can see that all QHO wave functions have an expectation
-value of zero for x-measurement.
-
-\begin{align*}
-  <p> &= \bra{n}\hat{p}\ket{n}\\
-  &= i\sqrt{\frac{m\hbar\omega}{2}}\bra{n}\left(a^\dagger-a\right)\ket{n}\\
-  &= i\sqrt{\frac{m\hbar\omega}{2}}
-  \left(\sqrt{n+1}\bra{n}\ket{n+1} - \sqrt{n}\bra{n}\ket{n-1}\right) = 0\\
-\end{align*}
-This is an interesting result. It is intuitive that any wave function should have an expectation of
-zero for momentum, else the wave function would move through space.
-
-How does this reconcile w/ fact that p involves a spacial derivative?
 

### 2016-01-20 quantum/latex/a2.tex
b la

--- 2016-01-20 quantum/latex/a2.tex
@@
   &= \sqrt{\frac{m\omega}{\pi\hbar}}\int_{-\infty}^{\infty} -\hbar^2e^{\frac{-m\omega x^2}{\hbar}}
-  \left(\frac{m^2\omega^2x^2}{\hbar^2} - \frac{m\omega x}{\hbar}\right)dx\\
+  \left(\frac{m^2\omega^2x^2}{\hbar^2} - \frac{m\omega}{\hbar}\right)dx\\
   &= -\hbar^2\sqrt{\frac{m\omega}{\pi\hbar}}\left(
@@
   &= -\hbar^2\sqrt{\frac{m\omega}{\pi\hbar}}
-  \left(\frac{1}{4}\sqrt{\frac{m\pi\omega}{\hbar}} -
-  \frac{1}{2}\sqrt{\frac{m\pi\omega}{\hbar}}\right) \hspace{2cm}\mbox{Wolfram}\\
-  &= \frac{\hbar m\omega}{4}\\
+  \left(\frac{1}{2}\sqrt{\frac{m\pi\omega}{\hbar}} -
+  \sqrt{\frac{m\pi\omega}{\hbar}}\right) \hspace{2cm}\mbox{Wolfram}\\
+  &= \frac{\hbar m\omega}{2}\\
 \end{align*}
@@
 This is an interesting result. It is intuitive that any wave function should have an expectation of
-zero for momentum, else the wave function would move through space.
+zero for momentum, else the wave function would move through space due to a nonzero average
+velocity.
 
-How does this reconcile w/ fact that p involves a spacial derivative?
+\begin{align*}
+  <x^2> &= \bra{n}x^2\ket{n}\\
+  &= \frac{\hbar}{2m\omega}\bra{n}\left(a + a^\dagger\right)\left(a + a^\dagger\right)\ket{n}\\
+  &= \frac{\hbar}{2m\omega}\bra{n}
+  \left(aa + aa^\dagger + a^\dagger a + a^\dagger a^\dagger\right)\ket{n}\\
+  &= \frac{\hbar}{2m\omega}
+  \left(aa + aa^\dagger + a^\dagger a + a^\dagger a^\dagger\right)\ket{n}\\
+  &= \frac{\hbar}{2m\omega} \left(\sqrt{n-1}\sqrt{n-2}\bra{n}\ket{n-2} + (n+1)\bra{n}\ket{n} +
+  (n)\bra{n}\ket{n} + \sqrt{n+1}\sqrt{n+2}\bra{n}\ket{n+2}\right)\\
+  &= \frac{\hbar}{2m\omega} \left(2n+1\right)\\
+\end{align*}
+This is an interesting result which indicates that for each extra rung on the ladder, the
+expectation of the squared position increases linearly, meaning the expectation of the magnitude
+of x increases as $\sqrt{n}$. This makes sense, since increasing energy levels have less and less
+room to expand in the quadratic potential well. The $n=0$ state corresponds to the wave function
+solution for the ground state found above.\\
 
+\begin{align*}
+  <p^2> &= \bra{n}p^2\ket{n}\\
+  &= -\frac{m\hbar\omega}{2}\bra{n}\left(a^\dagger-a\right)\left(a^\dagger-a\right)\ket{n}\\
+  &= -\frac{m\hbar\omega}{2}\bra{n}
+  \left(a^\dagger a^\dagger - a^\dagger a - aa^\dagger + aa\right)\ket{n}\\
+  &= -\frac{m\hbar\omega}{2}\bra{n}\left(\sqrt{n-1}\sqrt{n-2}\bra{n}\ket{n-2}
+  - (n)\bra{n}\ket{n} - (n+1)\bra{n}\ket{n} + \sqrt{n+1}\sqrt{n+2}\bra{n}\ket{n+2}\right)\\
+  &= \frac{\hbar m\omega}{2}\left(2n + 1\right)
+\end{align*}
+This result is similar to the above one for $x^2$ with different units. It is still a $\sqrt{n}$
+relationship for momentum, which makes sense.\\
+
+\textbf{c.) Uncertainty principle}\\
+Since the above two cases agree with each other, we check that the more general case of part
+\textbf{b.} agrees with the uncertainty principle.
+
+\begin{align*}
+  \Delta x &= \sqrt{<x^2>-<x>^2}\\
+  &= \sqrt{\frac{\hbar}{2m\omega} \left(2n+1\right)}\\
+  \Delta p &= \sqrt{<p^2>-<p>^2}\\
+  &= \sqrt{\frac{\hbar m\omega}{2}\left(2n + 1\right)}\\
+  \Delta x \Delta p &= \sqrt{\frac{\hbar}{2m\omega} \left(2n+1\right)}
+  \sqrt{\frac{\hbar m\omega}{2}\left(2n + 1\right)}\\
+  &= \frac{\hbar}{2}(2n+1)\\
+\end{align*}
+This is an interesting result which indicates that uncertainty increases as the pure wave function
+gets more energetic. Ground state is the state with maximum certainty. The higher states get
+progressively more uncertain, which makes sense since both the velocity and the well get larger
+for higher states.\\
+
+\section{9.13: }
+
+\section{9.14: Unknown QHO wavefunction}
+We are given the wave function $\psi(x,0) = A\left[1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
+  2\frac{m\omega}{\hbar}x^2\right]$. Because its highest-order polynomial term is $x^2$, we know
+that it must be made up of the $n = 0, 1, 2$ energy states. We now figure out the coefficients for
+each:
+
+\begin{align*}
+  c_0 &= \int_{-\infty}^{\infty} \phi_0^*\psi dx\\
+  &= A\int_{-\infty}^{\infty} \left(\frac{m\omega}{\pi\hbar}\right)^{1/2}e^{-m\omega x^2/\hbar}
+  \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
+  2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/(2\hbar)}dx\\
+  &= 2\\
+  c_1 &= A\int_{-\infty}^{\infty} \frac{2\sqrt{\frac{m\omega}{\hbar}x}}{\sqrt{2}}
+    \left(\frac{m\omega}{\pi\hbar}\right)^{1/2} \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
+    2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
+ &= -\frac{3}{\sqrt{2}}\\
+ c_2 &= A\int_{-\infty}^{\infty} \frac{\frac{4m\omega}{\hbar}x^2 - 2}{\sqrt{8}}
+    \left(\frac{m\omega}{\pi\hbar}\right)^{1/2} \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
+    2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
+ &= \sqrt{2}\\
+\end{align*}
+
+From this we can see that $\psi(x,0) = \sqrt{\frac{21}{2}}
+\left(2\ket{0} + -\frac{3}{\sqrt{2}}\ket{1} + \sqrt{2}\ket{2}\right)$. It is then easy to compute
+the expectation value: \\
+
+\begin{align*}
+  <E> &= \sigma \hbar\omega\left(n+\frac{1}{2}\right)|c_n|^2\\
+  &= \frac{2 + \frac{27}{4} + 5}{\frac{21}{2}} = 1.3 eV\\
+\end{align*}
 \end{document}

### 2016-01-20 quantum/latex/a2.tex
stuff.

--- 2016-01-20 quantum/latex/a2.tex
@@
 %% \end{align*}
+\section{5: 2D QHO}
+To take the QHO to two dimensions, we use the following Hamiltonian:
+
+\begin{align*}
+  H &= \frac{\hat{p}_x^2}{2m} + \frac{\hat{p}_y^2}{2m}
+  + \frac{1}{2}m\omega^2\left(x^2+y^2\right)\\
+\end{align*}
+
+We notice that none of the terms in this Hamiltonian are coupled, so we break it
+up into two Hamiltonians:
+
+\begin{align*}
+  H_x &= \frac{\hat{p}_x^2}{2m} \frac{1}{2}m\omega^2x^2)\\
+  H_y &= \frac{\hat{p}_y^2}{2m} \frac{1}{2}m\omega^2y^2)\\
+  H &= H_x + H_y\\
+\end{align*}
+
+our energy eigenvalue equation then takes the form:
+
+\begin{align*}
+  \left(\hat{H}_x + \hat{H}_y\right)X(x)Y(y) &= (E_x + E_y)X(x)Y(y)\\
+\end{align*}
+
+This equation is basically the sum of two uncoupled QHOs. Because of this, we
+can calculate the energy fairly easily:
+
+\begin{align*}
+  E_{xy} &= E_x + E_y = \left(n_x+n_y + 1\right)\hbar\omega\\
+\end{align*}
+
+From this we can see that the system is very degenerate. For the $n^{th}$-highest
+energy level there are $\frac{n}{2}$ possible states that have that energy: for
+the tenth energy level, the possible n-numbers are $1,9$, $2,8$, $3,7$, $4,6$ and
+$5,5$.\\
+
+We can construct the ground level energy state by taking the product of the two
+uncoupled QHOs:
+
+\begin{align*}
+  \psi_0(x,y) &= \left(\frac{m\omega}{\pi\hbar}\right)
+  e^{-m\omega\left(x^2+y^2\right)/2\hbar}\\
+\end{align*}
 

### 2016-01-20 quantum/latex/a2.tex
bla

@@
   P_{forb} &= \int_{\infty}^{-x_0} \psi_0^*\psi_0 dx + \int_{x_0}^{\infty} \psi_0^*\psi_0 dx\\
-  &= \int_{\infty}^{-x_0} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar}dx
-  + \int_{x_0}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar} dx\\
-  &= \sqrt{\frac{m\omega}{\pi\hbar}} \left(\frac{-\hbar}{m\omega}\right)\left(
-    \frac{1}{x}e^{-m\omega x^2 / \hbar} \Bigg|_{-\infty}^{-\sqrt{\frac{\hbar}{m\omega}}} +
-    \frac{1}{x}e^{-m\omega x^2 / \hbar} \Bigg|_{\sqrt{\frac{\hbar}{m\omega}}}^{\infty}\right)\\
-  &= \frac{2}{\sqrt{\pi}} = 0.41\\
+  &= 2\int_{x_0}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar} dx\\
+  &= 0.16\\
 \end{align*}
+
+Thus we see that there is a small chance of finding the particle in the classically
+forbidden region, which is expected for the ground state of the QHO. To do the
+calculation we exploit the fact that even functions can be integrated on only one
+side of the y-axis and doubled to get the total integral answer.\\
 
@@
 %%   2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/(2\hbar)}dx\\
-%%   &= 2\\
+%%   &= 2A\\
 %%   c_1 &= A\int_{-\infty}^{\infty} \frac{2\sqrt{\frac{m\omega}{\hbar}x}}{\sqrt{2}}
@@
 %%     2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
-%%  &= -\frac{3}{\sqrt{2}}\\
+%%  &= -\frac{3}{\sqrt{2}}A\\
 %%  c_2 &= A\int_{-\infty}^{\infty} \frac{\frac{4m\omega}{\hbar}x^2 - 2}{\sqrt{8}}
@@
 %%     2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
-%%  &= \sqrt{2}\\
+%%  &= \sqrt{2}A\\
 %% \end{align*}
 
-%% From this we can see that $\psi(x,0) = \sqrt{\frac{21}{2}}
-%% \left(2\ket{0} + -\frac{3}{\sqrt{2}}\ket{1} + \sqrt{2}\ket{2}\right)$. It is then easy to compute
+%% From this we can see that $\psi(x,0) = \sqrt{
+%% \left(\frac{2\sqrt{2}}{\sqrt{21}\ket{0}
+%%+ \frac{-3}{\sqrt{21}}\ket{1} + \frac{sqrt{2}}{\sqrt{21}}\ket{2}\right)$. It is then easy to compute
 %% the expectation value: \\
@@
 %% \begin{align*}
-%%   <E> &= \sigma \hbar\omega\left(n+\frac{1}{2}\right)|c_n|^2\\
-%%   &= \frac{2 + \frac{27}{4} + 5}{\frac{21}{2}} = 1.3 eV\\
+%%   <E> &= \sum \hbar\omega\left(n+\frac{1}{2}\right)|c_n|^2\\
+%%   &= \frac{2 + \frac{27}{4} + 5}{\frac{21}{2}} = 1.3\hbar\omega\\
 %% \end{align*}
+\section{9.13: Half-QHO well}
+(Inspired by http://www.physicspages.com/2012/08/18/half-harmonic-oscillator/)
+
+This system is identical to the 1D QHO, except all positive x values have infinite
+potential. So the system is a mixture of the infinite well and SHO potentials. We
+thus expect the wave function to have characteristics of both systems. For positive
+x values we expect the wave function to be zero, as it is in the infinite well.
+Because the negative x-values have a potential identical to QHO, we expect the wave
+function to be identical as well. So our total wave function is a Hermite/Gaussian
+for negative x values and zero for positive x values.\\
+
+However, we're not done. We notice that the x=0 boundary condition is not obeyed
+for all Hermite/Gaussian wave functions. We require the wave function to be zero
+at x=0, which is only true for odd QHO wave functions. Thus our energy spectrum
+is identical to the QHO spectru, but only for odd values of n:
+
+\begin{align*}
+  E_n &= \left(n + \frac{1}{2}\right)\hbar\omega\hspace{2cm}\mbox{n=1,3,5,...}\\
+\end{align*}
+
+For this problem we just applied the basic logic that the value of a wave function
+is only based on its current location, not the overall shape of the potential. The
+wave function for a system that is identical to the QHO will have a wave function
+identical to QHO in the similar regions. This is a powerful result.\\
+
 \section{5: 2D QHO}
+(Derivation inspired by
+$\mbox{nyu.edu/classes/tuckerman/adv.chem/lectures/lecture\_18/node6.html)}$\\
 To take the QHO to two dimensions, we use the following Hamiltonian:
@@
 
-We notice that none of the terms in this Hamiltonian are coupled, so we break it
-up into two Hamiltonians:
-
-\begin{align*}
-  H_x &= \frac{\hat{p}_x^2}{2m} \frac{1}{2}m\omega^2x^2)\\
-  H_y &= \frac{\hat{p}_y^2}{2m} \frac{1}{2}m\omega^2y^2)\\
-  H &= H_x + H_y\\
-\end{align*}
-
+We notice that none of the cartesian terms in this Hamiltonian are coupled, so
 our energy eigenvalue equation then takes the form:
@@
 \begin{align*}
-  \left(\hat{H}_x + \hat{H}_y\right)X(x)Y(y) &= (E_x + E_y)X(x)Y(y)\\
+  H_{xy}X(x)Y(y) &= (E_x + E_y)X(x)Y(y)\\
 \end{align*}
@@
 \begin{align*}
-  \psi_0(x,y) &= \left(\frac{m\omega}{\pi\hbar}\right)
+  \psi_0(x,y) &= \left(\frac{m\omega}{\pi\hbar}\right)^{1/4}
   e^{-m\omega\left(x^2+y^2\right)/2\hbar}\\
 \end{align*}
+
+The ease with which we go from the 1D to 2D harmonic oscillators comes from the fact
+that the x and y coordinates of the oscillator are not coupled. The x and y cross
+sections of the wave can take on any allowed 1D state. It is not intuitive to me
+why this method exactly works. Why are the energies of the x and y components
+just summed together to get total energy, as opposed to being multiplied or
+something? The degeneracy of the system is interesting. Because a sum is used,
+for large energies there are a huge number of ways to configure the system.\\
 

### 2016-01-24 quantum/latex/a3.tex
Add quantum hw 3 beginnings.

--- 2016-01-20 quantum/latex/a2.tex
@@
 \usepackage{fancyhdr}
+\usepackage{amsmath}
 \usepackage{physics}
@@
 
-\title{Quantum Mechanics Assignment Two}
+\title{Quantum Mechanics Assignment Three}
 \author{Elliott Capek}
@@
 
-\section{9.4: Forbidden region probability for QHO}
-For this problem we want to calculate the probability of measuring a QHO ground-state object outside
-of the classically allowed region. We do this by calculating the turning points of a classical HO,
-ie finding the x-values where energy equals potential, and then integrating $\psi^*\psi$ over the
-forbidden region to find the chance of forbidden observation.\\
+\section{9.16: Time evolving QHO state}
+We are given the probabilities of measuring certain energies for a QHO state $\ket{\psi}$. We
+are asked to find the energy basis representation for this state and compute the expectations for
+momentum and energy.\\
 
-We first find the turning point for the ground state $n=0$:
+\textbf{Finding energy representation}
 \begin{align*}
-  V_{HO} &= E_0\\
-  \frac{1}{2}m\omega^2x^2 &= \hbar\omega\left(0+\frac{1}{2}\right)\\
-  x_0 &= \sqrt{\frac{\hbar}{m\omega}}\\
+  |\bra{1}\ket{\psi}|^2 = 0.36 \rightarrow |c_1|^2 &= 0.36 \rightarrow c_1 = e^{ai}0.6\\
+  |\bra{2}\ket{\psi}|^2 = 0.64 \rightarrow |c_1|^2 &= 0.64 \rightarrow c_1 = e^{bi}0.8\\
 \end{align*}
 
-We then find the probability of finding the particle in this region:
+\textbf{Time evolution}
+\begin{align*}
+  \ket{\psi} &= 0.6e^{ai}e^{-iE_1t/\hbar}\ket{1} + 0.8e^{bi}e^{-iE_2t/\hbar}\ket{2}\\
+  &= 0.6e^{ai}e^{-3i\omega t/2}\ket{1} + 0.8e^{bi}e^{-5i\omega t/2}\ket{2}\\
+\end{align*}
+
+We are told that the expectation value for position is minimized at $t=0$. We can use this to find
+the relative phases of the wave function coefficients:\\
+
+\textbf{Calculating relative phase}
+\begin{align*}
+  <x> &= \bra{\psi}\left(\sqrt{\frac{\hbar}{2m\omega}}
+  \left(a + a^\dagger\right)\right)\ket{\psi}\\
+  &= \sqrt{\frac{\hbar}{2m\omega}}\left(\bra{\psi}a\ket{\psi}
+  + \bra{\psi}a^\dagger\ket{\psi}\right)\\
+  &= \sqrt{\frac{\hbar}{2m\omega}}\left(\sqrt{2}c_1^*(t)c_2(t)\bra{1}\ket{1}
+  + \sqrt{2}c_1(t)c_2^*(t)\bra{2}\ket{2}\right)\\
+  &= \sqrt{\frac{\hbar}{m\omega}}\left(c_1^*(t)c_2(t) + c_1(t)c_2^*(t)\right)\\
+  &= \sqrt{\frac{\hbar}{m\omega}}\left(0.6*0.8*e^{i(b-a)} + 0.6*0.8e^{i(a-b)}\right)
+  \hspace{2cm}\mbox{Time-dependent terms are 1 when t=0}\\
+  &= 2*0.6*0.8*\sqrt{\frac{\hbar}{m\omega}}\left(\cos(b-a) + i\sin(b-a)\right)\\
+\end{align*}
+
+To minimize this expectation value, $a-b = b-a = \pi$. Thus we represent our system as follows:
 
 \begin{align*}
-  P_{forb} &= \int_{\infty}^{-x_0} \psi_0^*\psi_0 dx + \int_{x_0}^{\infty} \psi_0^*\psi_0 dx\\
-  &= \int_{\infty}^{-x_0} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar}dx
-  + \int_{x_0}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar} dx\\
-  &= \sqrt{\frac{m\omega}{\pi\hbar}} \left(\frac{-\hbar}{m\omega}\right)\left(
-    \frac{1}{x}e^{-m\omega x^2 / \hbar} \Bigg|_{-\infty}^{-\sqrt{\frac{\hbar}{m\omega}}} +
-    \frac{1}{x}e^{-m\omega x^2 / \hbar} \Bigg|_{\sqrt{\frac{\hbar}{m\omega}}}^{\infty}\right)\\
-  &= \frac{2}{\sqrt{\pi}} = 0.41\\
+  \psi(t) &= 0.6e^{i\pi}e^{-3i\omega t/2}\ket{1} + 0.8e^{-5i\omega t/2}\ket{2}\\
+  &= -0.6e^{-3i\omega t/2}\ket{1} + 0.8e^{-5i\omega t/2}\ket{2}\\
 \end{align*}
 
-\section{9.9: QHO expectation values via integration, operator method}
-In this problem we use both wave function integration and the operator method to find the
-expectation values of various observables: x, p, $x^2$, $p^2$.
+We then calculate expectation values for the system:\\
 
-\textbf{a.) Ground state wave function calculations}
+\textbf{Momentum}
 \begin{align*}
-  <x> &= \int_{-\infty}^{\infty} \psi^*x\psi dx\\
-  &= \int_{-\infty}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar}x dx = 0\\
-\end{align*}
-We don't need to perform the $<x>$ integral because it involves integrating an odd function over
-an even domain, which always results in zero. This is to be expected, because the wave ground-state
-wave function is symmetric across zero.\\
-
-\begin{align*}
-  <p> &= \int_{-\infty}^{\infty} \psi^* -i\hbar \frac{d}{dx}\psi dx\\
-  &= \int_{-\infty}^{\infty}
-  -i\sqrt{\frac{m\omega}{\pi\hbar}}(-m\omega x)e^{\frac{-m\omega x^2}{2\hbar}}dx = 0\\
-\end{align*}
-This integral is zero since we are again integrating an odd function over an even domain. This
-makes sense because, since the spacial distribution of the wave function is equal on both sides
-of the x-axis, the velocities (and hence momenta) should be equal as well.\\
-
-\begin{align*}
-  <x^2> &= \int_{-\infty}^{\infty} \psi^*x^2\psi dx\\
-  &= \int_{-\infty}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar}x^2 dx\\
-  &= 2\sqrt{\frac{m\omega}{\pi\hbar}}\int_{0}^{\infty} e^{-m\omega x^2 / \hbar}x^2 dx\\
-  &= 2\sqrt{\frac{m\omega}{\pi\hbar}} *
-  \frac{\sqrt{\pi}}{4(\frac{m\omega}{\hbar})^{3/2}} \hspace{2cm}\mbox{Wolfram}\\
-  &= \frac{\hbar}{2m\omega}\\
-\end{align*}
-Here we exploit the fact that integrating an even function over an even domain is equivalent to
-doubling the integral of the function over half the domain. The answer makes sense, since increasing
-$m\omega$ corresponds to making the quadratic potential steeper, hence making the wave function
-tighter.\\
-
-\begin{align*}
-  <p^2> &= \int_{-\infty}^{\infty} \psi^* \left(-\hbar^2 \frac{d^2}{dx^2}\right)\psi dx\\
-  &= \sqrt{\frac{m\omega}{\pi\hbar}}\int_{-\infty}^{\infty} -\hbar^2e^{\frac{-m\omega x^2}{\hbar}}
-  \left(\frac{m^2\omega^2x^2}{\hbar^2} - \frac{m\omega}{\hbar}\right)dx\\
-  &= -\hbar^2\sqrt{\frac{m\omega}{\pi\hbar}}\left(
-  \int_{-\infty}^{\infty} e^{\frac{-m\omega x^2}{\hbar}}\frac{m^2\omega^2x^2}{\hbar^2}dx -
-  \int_{-\infty}^{\infty} e^{\frac{-m\omega x^2}{\hbar}}\frac{m\omega}{\hbar}\right)dx\\
-  &= -\hbar^2\sqrt{\frac{m\omega}{\pi\hbar}}
-  \left(\frac{1}{2}\sqrt{\frac{m\pi\omega}{\hbar}} -
-  \sqrt{\frac{m\pi\omega}{\hbar}}\right) \hspace{2cm}\mbox{Wolfram}\\
-  &= \frac{\hbar m\omega}{2}\\
-\end{align*}
-This answer is intuitive, since increasing mass as well as oscillatory speed should both increase
-the expected momentum.
-
-\textbf{b.) Any-state operator method calculations}
-Here we define the position and momentum operators in terms of the raising and lowering operators:
-\begin{align*}
-  \hat{x} &= \sqrt{\frac{\hbar}{2m\omega}}\left(a + a^\dagger\right)\\
-  \hat{p} &= i\sqrt{\frac{m\omega\hbar}{2}}\left(a^\dagger - a\right)\\
-\end{align*}
-We then solve for the various expectation values using these definitions:
-
-\begin{align*}
-  <x> &= \bra{n}x\ket{n}\\
-  &= \sqrt{\frac{\hbar}{2m\omega}}\bra{n}\left(a + a^\dagger\right)\ket{n}\\
-  &= \sqrt{\frac{\hbar}{2m\omega}}
-  \left(\sqrt{n}\bra{n}\ket{n-1} + \sqrt{n+1}\bra{n}\ket{n+1}\right) = 0\\
-\end{align*}
-This answer indicates that all pure QHO wave functions are centered around zero. This makes
-sense since the QHO wave functions are all made up of either even or odd functions. When squared,
-as is done when finding the probability, the equation becomes entirely even, meaning it is
-symmetric across the y-axis. From this we can see that all QHO wave functions have an expectation
-value of zero for x-measurement.
-
-\begin{align*}
-  <p> &= \bra{n}\hat{p}\ket{n}\\
-  &= i\sqrt{\frac{m\hbar\omega}{2}}\bra{n}\left(a^\dagger-a\right)\ket{n}\\
-  &= i\sqrt{\frac{m\hbar\omega}{2}}
-  \left(\sqrt{n+1}\bra{n}\ket{n+1} - \sqrt{n}\bra{n}\ket{n-1}\right) = 0\\
-\end{align*}
-This is an interesting result. It is intuitive that any wave function should have an expectation of
-zero for momentum, else the wave function would move through space due to a nonzero average
-velocity.
-
-\begin{align*}
-  <x^2> &= \bra{n}x^2\ket{n}\\
-  &= \frac{\hbar}{2m\omega}\bra{n}\left(a + a^\dagger\right)\left(a + a^\dagger\right)\ket{n}\\
-  &= \frac{\hbar}{2m\omega}\bra{n}
-  \left(aa + aa^\dagger + a^\dagger a + a^\dagger a^\dagger\right)\ket{n}\\
-  &= \frac{\hbar}{2m\omega}
-  \left(aa + aa^\dagger + a^\dagger a + a^\dagger a^\dagger\right)\ket{n}\\
-  &= \frac{\hbar}{2m\omega} \left(\sqrt{n-1}\sqrt{n-2}\bra{n}\ket{n-2} + (n+1)\bra{n}\ket{n} +
-  (n)\bra{n}\ket{n} + \sqrt{n+1}\sqrt{n+2}\bra{n}\ket{n+2}\right)\\
-  &= \frac{\hbar}{2m\omega} \left(2n+1\right)\\
-\end{align*}
-This is an interesting result which indicates that for each extra rung on the ladder, the
-expectation of the squared position increases linearly, meaning the expectation of the magnitude
-of x increases as $\sqrt{n}$. This makes sense, since increasing energy levels have less and less
-room to expand in the quadratic potential well. The $n=0$ state corresponds to the wave function
-solution for the ground state found above.\\
-
-\begin{align*}
-  <p^2> &= \bra{n}p^2\ket{n}\\
-  &= -\frac{m\hbar\omega}{2}\bra{n}\left(a^\dagger-a\right)\left(a^\dagger-a\right)\ket{n}\\
-  &= -\frac{m\hbar\omega}{2}\bra{n}
-  \left(a^\dagger a^\dagger - a^\dagger a - aa^\dagger + aa\right)\ket{n}\\
-  &= -\frac{m\hbar\omega}{2}\bra{n}\left(\sqrt{n-1}\sqrt{n-2}\bra{n}\ket{n-2}
-  - (n)\bra{n}\ket{n} - (n+1)\bra{n}\ket{n} + \sqrt{n+1}\sqrt{n+2}\bra{n}\ket{n+2}\right)\\
-  &= \frac{\hbar m\omega}{2}\left(2n + 1\right)
-\end{align*}
-This result is similar to the above one for $x^2$ with different units. It is still a $\sqrt{n}$
-relationship for momentum, which makes sense.\\
-
-\textbf{c.) Uncertainty principle}\\
-Since the above two cases agree with each other, we check that the more general case of part
-\textbf{b.} agrees with the uncertainty principle.
-
-\begin{align*}
-  \Delta x &= \sqrt{<x^2>-<x>^2}\\
-  &= \sqrt{\frac{\hbar}{2m\omega} \left(2n+1\right)}\\
-  \Delta p &= \sqrt{<p^2>-<p>^2}\\
-  &= \sqrt{\frac{\hbar m\omega}{2}\left(2n + 1\right)}\\
-  \Delta x \Delta p &= \sqrt{\frac{\hbar}{2m\omega} \left(2n+1\right)}
-  \sqrt{\frac{\hbar m\omega}{2}\left(2n + 1\right)}\\
-  &= \frac{\hbar}{2}(2n+1)\\
-\end{align*}
-This is an interesting result which indicates that uncertainty increases as the pure wave function
-gets more energetic. Ground state is the state with maximum certainty. The higher states get
-progressively more uncertain, which makes sense since both the velocity and the well get larger
-for higher states.\\
-
-\section{9.13: }
-
-\section{9.14: Unknown QHO wavefunction}
-We are given the wave function $\psi(x,0) = A\left[1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
-  2\frac{m\omega}{\hbar}x^2\right]$. Because its highest-order polynomial term is $x^2$, we know
-that it must be made up of the $n = 0, 1, 2$ energy states. We now figure out the coefficients for
-each:
-
-\begin{align*}
-  c_0 &= \int_{-\infty}^{\infty} \phi_0^*\psi dx\\
-  &= A\int_{-\infty}^{\infty} \left(\frac{m\omega}{\pi\hbar}\right)^{1/2}e^{-m\omega x^2/\hbar}
-  \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
-  2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/(2\hbar)}dx\\
-  &= 2\\
-  c_1 &= A\int_{-\infty}^{\infty} \frac{2\sqrt{\frac{m\omega}{\hbar}x}}{\sqrt{2}}
-    \left(\frac{m\omega}{\pi\hbar}\right)^{1/2} \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
-    2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
- &= -\frac{3}{\sqrt{2}}\\
- c_2 &= A\int_{-\infty}^{\infty} \frac{\frac{4m\omega}{\hbar}x^2 - 2}{\sqrt{8}}
-    \left(\frac{m\omega}{\pi\hbar}\right)^{1/2} \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
-    2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
- &= \sqrt{2}\\
+  <p> &= -i\sqrt{\frac{\hbar}{2m\omega}}\bra{\psi}\left(a - a^\dagger)\right)\ket{\psi}\\
+  &= -i\sqrt{\frac{\hbar}{2m\omega}}\left(\bra{\psi}a\ket{\psi}
+  - \bra{\psi}a^\dagger\ket{\psi}\right)\\
+  &= -i\sqrt{\frac{\hbar}{2m\omega}}\left(\sqrt{2}c_1^*(t)c_2(t)\bra{1}\ket{1}
+  - \sqrt{2}c_1(t)c_2^*(t)\bra{2}\ket{2}\right)\\
+  &= -i*0.6*0.8*\sqrt{\frac{\hbar}{m\omega}}\left(e^{i\omega t} - e^{-i\omega t}\right)\\
+  &= -0.48\sqrt{\frac{\hbar}{m\omega}}\sin(\omega t)\\
 \end{align*}
 
-From this we can see that $\psi(x,0) = \sqrt{\frac{21}{2}}
-\left(2\ket{0} + -\frac{3}{\sqrt{2}}\ket{1} + \sqrt{2}\ket{2}\right)$. It is then easy to compute
-the expectation value: \\
+\textbf{Energy}
+\begin{align*}
+  \braket{p} &= 
+\end{align*}
 
-\begin{align*}
-  <E> &= \sigma \hbar\omega\left(n+\frac{1}{2}\right)|c_n|^2\\
-  &= \frac{2 + \frac{27}{4} + 5}{\frac{21}{2}} = 1.3 eV\\
-\end{align*}
 \end{document}

### 2016-01-25 bioinformatics/hw1.py
Add bioinformatics dir.

@@ -1,0 +1,7 @@
+#! /usr/bin/python2.7
+
+from Bio import SeqIO
+
+seq = SeqIO.parse(open("sequence"), "fasta")
+for ref in seq:
+    print ref.seq

### 2016-01-25 quantum/latex/a3.tex

@@
   <p> &= -i\sqrt{\frac{\hbar}{2m\omega}}\bra{\psi}\left(a - a^\dagger)\right)\ket{\psi}\\
-  &= -i\sqrt{\frac{\hbar}{2m\omega}}\left(\bra{\psi}a\ket{\psi}
+  &= -i\sqrt{\frac{\hbar m\omega}{2}}\left(\bra{\psi}a\ket{\psi}
   - \bra{\psi}a^\dagger\ket{\psi}\right)\\
-  &= -i\sqrt{\frac{\hbar}{2m\omega}}\left(\sqrt{2}c_1^*(t)c_2(t)\bra{1}\ket{1}
+  &= -i\sqrt{\frac{\hbar m\omega}{2}}\left(\sqrt{2}c_1^*(t)c_2(t)\bra{1}\ket{1}
   - \sqrt{2}c_1(t)c_2^*(t)\bra{2}\ket{2}\right)\\
-  &= -i*0.6*0.8*\sqrt{\frac{\hbar}{m\omega}}\left(e^{i\omega t} - e^{-i\omega t}\right)\\
-  &= -0.48\sqrt{\frac{\hbar}{m\omega}}\sin(\omega t)\\
+  &= -i*0.6*0.8*\sqrt{\hbar m\omega}\left(e^{i\omega t} - e^{-i\omega t}\right)\\
+  &= -0.48\sqrt{\hbar m\omega}\sin(\omega t)\\
 \end{align*}
+
+An interesting result! First, the units look good. Second, the magnitude of the expected momentum
+integrated over time is zero, meaning that the particle won't travel anywhere, which is good. The
+trigonometric function suggests that the particle's momentum oscillates sinusoidally at the
+classical frequency, just as we expect for a classical oscillator.\\
 
@@
 \begin{align*}
-  \braket{p} &= 
+  <H> &= \bra{1}(\frac{3}{2}\hbar\omega)\ket{1} + \bra{2}(\frac{5}{2}\hbar\omega)\ket{2}\\
+  &= 0.36 * \frac{3}{2}\hbar\omega + 0.64 * \frac{5}{2}\hbar\omega\\
+  &= 2.08\hbar\omega\\
 \end{align*}
 
+This is an expected result closest to the most-represented eigenstate in the system. It makes sense
+that there is no time-dependence, since that would mean energy would be entering and leaving the
+system, and energy is conserved for QM expectation values.
+
+\section{9.17: More Expectation Values}
+Similar to the above problem we are given a wavefunction and asked to calculate expectation values
+from it. We begin by figuring out its normalization constant:
+
+\textbf{Normalization constant}
+\begin{align*}
+  \psi(x,0) &= A\left[\phi_0(x) + 2\phi_1(x) + 2\phi_2(x)\right]\\
+  \braket{\psi} &= 1
+  A^2\left(1 + 4 + 4\right) &= 1 \rightarrow A = \frac{1}{3}\\
+\end{align*}
+
+Therefore our full time-evolved wavefunction is:
+
+\begin{align*}
+  \psi(x,t) &= \frac{1}{3}e^{\frac{-i\omega t}{2}}\phi_0(x)
+  + \frac{2}{3}e^{\frac{-3i\omega t}{2}}\phi_1(x)
+  + \frac{2}{3}e^{\frac{-5i\omega t}{2}}\phi_2(x)\\
+\end{align*}
+
+\textbf{Energy measurements}
+By inspection we see that the first three energy levels of the QHO spectrum are represented in the
+wave function, so these are the only ones we calculate the energy for.
+
+\begin{align*}
+  \mathbb{P}(\frac{1}{2}\omega) &= |\bra{0}\ket{\psi}|^2 = |c_0|^2 = \frac{1}{9}\\
+  \mathbb{P}(\frac{3}{2}\omega) &= |\bra{1}\ket{\psi}|^2 = |c_1|^2 = \frac{4}{9}\\
+  \mathbb{P}(\frac{5}{2}\omega) &= |\bra{2}\ket{\psi}|^2 = |c_2|^2 = \frac{4}{9}\\
+\end{align*}
+
+We can see the probabilities sum up to 1, as desired.\\
+
+\textbf{Momentum expectation value}
+\begin{align*}
+  <p> &= -i\sqrt{\frac{\hbar}{2m\omega}}\bra{\psi}\left(a - a^\dagger)\right)\ket{\psi}\\
+  &= -i\sqrt{\frac{\hbar m\omega}{2}}\left(\bra{\psi}a\ket{\psi}
+  - \bra{\psi}a^\dagger\ket{\psi}\right)\\
+  &= -i\sqrt{\frac{\hbar m\omega}{2}}
+  \Bigg(
+  \sqrt{1}\frac{2}{9}e^{\frac{i\omega t}{2}}e^{\frac{-3i\omega t}{2}}\bra{0}\ket{0}
+  + \sqrt{2}\frac{4}{9}e^{\frac{3i\omega t}{2}}e^{\frac{-5i\omega t}{2}}\bra{1}\ket{1}\\
+  &-\sqrt{1}\frac{2}{9}e^{\frac{3i\omega t}{2}}e^{\frac{-i\omega t}{2}}\bra{1}\ket{1}
+  - \sqrt{2}\frac{4}{9}e^{\frac{5i\omega t}{2}}e^{\frac{-3i\omega t}{2}}\bra{2}\ket{2}
+  \Bigg)\\
+  &= -i\sqrt{\frac{\hbar m\omega}{2}}
+  \left(\frac{2}{9}e^{-i\omega t} + \sqrt{2}\frac{4}{9}e^{-i\omega t}
+  - \frac{2}{9}e^{i\omega t} - \sqrt{2}\frac{4}{9}e^{i\omega t}\right)\\
+  &= \frac{-2i}{9}\sqrt{\frac{\hbar m\omega}{2}}
+  \Bigg(
+  \cos(\omega t) - i\sin(\omega t)
+  + \sqrt{8}\cos(\omega t) - i\sqrt{8}\sin(\omega t)\\
+  &- \cos(\omega t) - i\sin(\omega t)
+  - \sqrt{8}\cos(\omega t) - i\sqrt{8}\sin(\omega t)
+  \Bigg)\\
+  &= \frac{4i}{9}\sqrt{\frac{\hbar m\omega}{2}}\left(1 + \sqrt{8}\right)\sin(\omega t)\\
+\end{align*}
+
+This is another cool result. It is messier than the previous momentum expectation value due to
+a larger number of states, but still has proper units and displays the sinusoidal oscillation
+typical of systems with adjacent energies. I wonder why systems without adjacent energies have a
+zero expectation value for momentum. What is special about adjacent states that makes them
+``couple''?\\
+
+\textbf{Energy expectation}
+\begin{align*}
+  <H> &= \frac{1}{9}\frac{1}{2}\hbar\omega + \frac{4}{9}\hbar{3}{2}\hbar\omega
+  + \frac{4}{9}\frac{5}{2}\hbar\omega\\
+  &= \frac{33}{18}\hbar\omega = 1.83\hbar\omega\\
+\end{align*}
+
+A cool and expected result.\\
+
+\textbf{Energy standard deviation}
+We must first find the value of $<H^2>$:
+\begin{align*}
+  <H^2> &= \frac{1}{9}\frac{1}{4}\hbar^2\omega^2 + \frac{4}{9}\hbar{9}{4}\hbar^2\omega^2
+  + \frac{4}{9}\frac{25}{4}\hbar^2\omega^2\\
+  &= \frac{137}{36}\hbar^2\omega^2 = 3.8\hbar^2\omega^2\\
+\end{align*}
+
+We then calculate the standard deviation:
+\begin{align*}
+  \Delta E &= \sqrt{<H^2> - <H>^2} = \sqrt{(3.8 - 3.34)\hbar^2\omega^2}\\
+  &= \sqrt{0.46\hbar^2\omega^2} \approx \frac{2}{3}\hbar\omega\\
+\end{align*}
+
+Standard deviations in QHO systems are interesting because all the energies are evenly spaced. This
+value represents how spread-out the states are. It is expected that because the states are close
+together for this system, the standard deviation is small.
 \end{document}

### 2016-01-27 quantum/latex/a3.tex
Add most of hw 3

@@
   &= 0.36 * \frac{3}{2}\hbar\omega + 0.64 * \frac{5}{2}\hbar\omega\\
-  &= 2.08\hbar\omega\\
+  &= 2.14\hbar\omega\\
 \end{align*}
@@
   \psi(x,0) &= A\left[\phi_0(x) + 2\phi_1(x) + 2\phi_2(x)\right]\\
-  \braket{\psi} &= 1
+  \braket{\psi} &= 1\\
   A^2\left(1 + 4 + 4\right) &= 1 \rightarrow A = \frac{1}{3}\\
@@
 together for this system, the standard deviation is small.
+
+\section{9.22: Fictional System}
+Here we are given the energy spectrum $E_n = n^3\hbar\omega$ and asked to give some information
+about the system: its Hamiltonian and eigenstates in energy representation (matrix form)
+and the matrix representation of the operator $A\ket{n} = 3n^2\ket{n+2}$.\\
+
+Because any operator in its own basis is diagonal and because any eigenvector in its own basis is
+a unit vector, we represent the Hamiltonian and eigenvectors as so:
+
+\begin{align*}
+  H \doteq
+  \begin{pmatrix}
+    \hbar\omega & 0 & 0 & ...\\
+    0 & 8\hbar\omega & 0 & ...\\
+    0 & 0 & 27\hbar\omega & ...\\
+    ... & ... & ... & ...\\
+  \end{pmatrix}
+\end{align*}
+
+\begin{align*}
+  \ket{1} \doteq
+  \begin{pmatrix}
+    1\\
+    0\\
+    0\\
+    ...\\
+  \end{pmatrix}
+  \hspace{1cm}
+  \ket{2} \doteq
+  \begin{pmatrix}
+    0\\
+    1\\
+    0\\
+    ...\\
+  \end{pmatrix}
+  \hspace{1cm}
+  \ket{3} \doteq
+  \begin{pmatrix}
+    0\\
+    0\\
+    1\\
+    ...\\
+  \end{pmatrix}
+\end{align*}
+
+To find the matrix representation of A we exploit the fact that in the energy basis eigenstates
+are represented by vectors with a single unit entry. Thus the first column of our matrix will only
+interact with the n=1 vector, the second column with the n=2 vector, etc. We also know that the
+result of our matrix-vector multiplication should produce a vector two energies higher than the
+input state. Thus it is easy to see that we will construct a matrix with elements two spaces below
+diagonal with values of $3n^2$, where n is the current column.
+
+\begin{align*}
+  A \doteq
+  \begin{pmatrix}
+    0 & 0 & 0 & 0 & 0 & ...  \\
+    0 & 0 & 0 & 0 & 0 & ...  \\
+    3 & 0 & 0 & 0 & 0 & ...  \\
+    0 & 12 & 0 & 0 & 0 & ... \\
+    0 & 0 & 27 & 0 & 0 & ... \\
+    0 & 0 & 0 & 48 & 0 & ... \\
+    ... & ... & ... & ... & ... & ... \\
+  \end{pmatrix}
+\end{align*}
+
+\section{Paper}
+The paper \textit{Quantum ground state and single-phonon control of a mechanical resonator} is a
+demonstration that a macroscopic, harmonic-oscillating ``quantum drum'' can be cooled enough to
+display behavior normally only predicted for microscopic systems. This is interesting because it
+is an easy validation of quantum mechanics. The project entails cooling a resonator until its
+heat energy is much smaller than that predicted for its oscillation energy, ensuring that
+energy measurements are only of the oscillation. The energy is then measured, and it is shown that
+only discrete values are taken on, not a continuum, as predicted by QM.\\
+
+It is interesting to me that the microscopic details of the system need not be taken into account.
+It seems like the individual wave functions of each atom in the oscillator can just be summed
+into one giant wave function for the oscillator. Maybe that simplification can only be done at
+very cold temperatures?\\
+
+\section{10.1: Diagonalize a perturbed Hamiltonian}
+We begin with the perturbed Hamiltonian in the $S_z$ basis and convert it into its own perturbed
+basis. We then show that the eigenvalues in this basis are identical to those found by redefining
+the direction of Z, aka $E_n = \pm \frac{\hbar}{2} \sqrt{(\omega_0 + \omega_1)^2 + \omega_2^2}$.
+
+We begin with the perturbed matrix and diagonalize it:\\
+
+\begin{align*}
+  H &\doteq
+  \pm\frac{\hbar}{2}
+  \begin{pmatrix}
+    \omega_0 + \omega_1 & \omega_2\\
+    \omega_2 & -\omega_0 - \omega_1\\
+  \end{pmatrix}
+  =
+  \pm\frac{\hbar}{2}
+  \begin{pmatrix}
+    -\omega_2 & \omega_2\left(\frac{-\omega_2}{\omega_0 + \omega_1}\right)\\
+    \omega_2 & -\omega_0 - \omega_1\\
+  \end{pmatrix}
+  =
+  \pm\frac{\hbar}{2}
+  \begin{pmatrix}
+    -\omega_2 & \frac{-\omega_2^2}{\omega_0 + \omega_1}\\
+    0 & -\omega_0 - \omega_1 - \frac{-\omega_2^2}{\omega_0 + \omega_1}\\
+  \end{pmatrix}\\
+  &=
+  \pm\frac{\hbar}{2}
+  \begin{pmatrix}
+    -\omega_2 & \frac{-\omega_2^2}{\omega_0 + \omega_1}\\
+    0 & \frac{\omega_2^2}{\omega_0 + \omega_1}\\
+  \end{pmatrix}
+  =
+  \pm\frac{\hbar}{2}
+  \begin{pmatrix}
+    -\omega_2 & 0\\
+    0 & \frac{\omega_2^2}{\omega_0 + \omega_1}\\
+  \end{pmatrix}
+\end{align*}
+
 \end{document}

### 2016-01-27 quantum/latex/a3.tex
finish hw 3

@@
   \Bigg)\\
-  &= \frac{4i}{9}\sqrt{\frac{\hbar m\omega}{2}}\left(1 + \sqrt{8}\right)\sin(\omega t)\\
+  &= \frac{-4}{9}\sqrt{\frac{\hbar m\omega}{2}}\left(1 + \sqrt{8}\right)\sin(\omega t)\\
 \end{align*}
@@
 
+This was a cool problem that shows that it is very easy to construct a Hamiltonian for a given
+quantum system in matrix representation (not necessarily position representation). It is
+important to note that operator A is not an actual physical operator, since it has no eigenvalue
+equation.\\
+
+\clearpage
+
 \section{Paper}
@@
 
-We begin with the perturbed matrix and diagonalize it:\\
+We begin with the perturbed matrix and use the fact that the eigenvalues of a matrix will be the
+diagonal elements in the diagonalized form:\\
 
@@
     \omega_2 & -\omega_0 - \omega_1\\
-  \end{pmatrix}
-  =
+  \end{pmatrix}\\
+\end{align*}
+
+\begin{align*}
+  \det\left(H - \lambda I\right) &= 0\\
+  \left(\omega_0 + \omega_1 - \lambda\right)
+  \left(-\omega_0 - \omega_1 - \lambda\right) - \omega_2^2 &= 0\\
+  \lambda^2 - \left(\omega_0 + \omega_1\right)^2 - \omega_2^2 &= 0\\
+  \lambda = \pm\sqrt{\left(\omega_0 + \omega_1\right)^2 + \omega_2^2}\\
+\end{align*}
+
+We then use these eigenvalues to construct a diagonal matrix. We know that the basis for the
+perturbed Hamiltonian will be very similar to the unperturbed Hamiltonian, so we arrange our
+perturbed Hamiltonian with the top left element positive and bottom right negative:
+
+\begin{align*}
+  H_p &\doteq
   \pm\frac{\hbar}{2}
   \begin{pmatrix}
-    -\omega_2 & \omega_2\left(\frac{-\omega_2}{\omega_0 + \omega_1}\right)\\
-    \omega_2 & -\omega_0 - \omega_1\\
-  \end{pmatrix}
-  =
-  \pm\frac{\hbar}{2}
-  \begin{pmatrix}
-    -\omega_2 & \frac{-\omega_2^2}{\omega_0 + \omega_1}\\
-    0 & -\omega_0 - \omega_1 - \frac{-\omega_2^2}{\omega_0 + \omega_1}\\
+    \sqrt{\left(\omega_0 + \omega_1\right)^2 + \omega_2^2} & 0\\
+    0 & -\sqrt{\left(\omega_0 + \omega_1\right)^2 + \omega_2^2}\\
   \end{pmatrix}\\
-  &=
-  \pm\frac{\hbar}{2}
-  \begin{pmatrix}
-    -\omega_2 & \frac{-\omega_2^2}{\omega_0 + \omega_1}\\
-    0 & \frac{\omega_2^2}{\omega_0 + \omega_1}\\
-  \end{pmatrix}
-  =
-  \pm\frac{\hbar}{2}
-  \begin{pmatrix}
-    -\omega_2 & 0\\
-    0 & \frac{\omega_2^2}{\omega_0 + \omega_1}\\
-  \end{pmatrix}
 \end{align*}
 
+We have already shown that the eigenvalues agree with those in McIntyre (10.9). What we have done in
+this problem is taken a matrix in the unperturbed energy basis and transformed it into its own
+perturbed basis. We know this is a new basis because the perturbed Hamiltonian in the original
+basis was not diagonal.\\
+
+It is a good confirmation that the method used in McIntyre to find the eigenvalues of the perturbed
+system, redefining the direction of Z to point in the direction of the new B-field, yielded the same
+result as creating a new basis for the perturbed Hamiltonian.\\
+
 \end{document}

### 2016-01-31 quantum/latex/a4.tex

--- 2016-01-25 quantum/latex/a3.tex
@@
 \usepackage{amsmath}
-\usepackage{physics}
+\usepackage{braket}
 \usepackage{amssymb}
@@
 
-\section{9.16: Time evolving QHO state}
-We are given the probabilities of measuring certain energies for a QHO state $\ket{\psi}$. We
-are asked to find the energy basis representation for this state and compute the expectations for
-momentum and energy.\\
+\section{10.8: Magnetic Perturbation Spin 1}
+For this problem we want to find the energy perturbations due to a magnetic field perturbation
+in the x direction, while remaining in the z basis. To do this we first find the
+perturbed Hamiltonian, which is just the x spin-1 matrix times a factor, and
+then use the formula
+%$E_n^{(2)} = \sum_{m\noteq n}\frac{|\bra{n^{(0)}}H'\ket{n^{(0)}|^2}{(E_n^{(0)}-E_m^{(0)}}}}$
+to find the second-order energy corrections.\\
 
-\textbf{Finding energy representation}
+We first find the perturbed Hamiltonian:
+
 \begin{align*}
-  |\bra{1}\ket{\psi}|^2 = 0.36 \rightarrow |c_1|^2 &= 0.36 \rightarrow c_1 = e^{ai}0.6\\
-  |\bra{2}\ket{\psi}|^2 = 0.64 \rightarrow |c_1|^2 &= 0.64 \rightarrow c_1 = e^{bi}0.8\\
+  H' &= \vec{\mu} \cdot \vec{B_2}\\
+  &= \frac{g e}{2m} \hat{\mu} \cdot B_2 \hat{x}\\
+  &= \frac{\hbar\omega_2}{2} S_x
+  \hspace{2cm}\omega_2 = \frac{geB}{2m}\\
+  &=
+  \begin{pmatrix}
+    0 & \frac{\hbar\omega_2}{2} & 0\\
+    \frac{\hbar\omega_2}{2} & 0 & \frac{\hbar\omega_2}{2}\\
+    0 & \frac{\hbar\omega_2}{2} & 0\\
+  \end{pmatrix}
 \end{align*}
 
-\textbf{Time evolution}
+Then we just apply the second-order perturbation energy equation to find our
+energy changes, noticing that there will be no first-order corrections because
+$H'$ has zeroes along its diagonal.
+
 \begin{align*}
-  \ket{\psi} &= 0.6e^{ai}e^{-iE_1t/\hbar}\ket{1} + 0.8e^{bi}e^{-iE_2t/\hbar}\ket{2}\\
-  &= 0.6e^{ai}e^{-3i\omega t/2}\ket{1} + 0.8e^{bi}e^{-5i\omega t/2}\ket{2}\\
+  E_n^{(2)} = \sum_{m\noteq n}\frac{|\bra{n^{(0)}}H'\ket{n^{(0)}|^2}{(E_n^{(0)}-E_m^{(0)}}}}\\
 \end{align*}
 
-We are told that the expectation value for position is minimized at $t=0$. We can use this to find
-the relative phases of the wave function coefficients:\\
-
-\textbf{Calculating relative phase}
 \begin{align*}
-  <x> &= \bra{\psi}\left(\sqrt{\frac{\hbar}{2m\omega}}
-  \left(a + a^\dagger\right)\right)\ket{\psi}\\
-  &= \sqrt{\frac{\hbar}{2m\omega}}\left(\bra{\psi}a\ket{\psi}
-  + \bra{\psi}a^\dagger\ket{\psi}\right)\\
-  &= \sqrt{\frac{\hbar}{2m\omega}}\left(\sqrt{2}c_1^*(t)c_2(t)\bra{1}\ket{1}
-  + \sqrt{2}c_1(t)c_2^*(t)\bra{2}\ket{2}\right)\\
-  &= \sqrt{\frac{\hbar}{m\omega}}\left(c_1^*(t)c_2(t) + c_1(t)c_2^*(t)\right)\\
-  &= \sqrt{\frac{\hbar}{m\omega}}\left(0.6*0.8*e^{i(b-a)} + 0.6*0.8e^{i(a-b)}\right)
-  \hspace{2cm}\mbox{Time-dependent terms are 1 when t=0}\\
-  &= 2*0.6*0.8*\sqrt{\frac{\hbar}{m\omega}}\left(\cos(b-a) + i\sin(b-a)\right)\\
+  E_{\frac{-1}{2}} &=
+  \frac{|H'_{\frac{-1}{2},0}|^2}{E_{0} - E_{\frac{-1}{2}}} +
+  \frac{|H'_{\frac{-1}{2},\frac{1}{2}}|^2}{E_{\frac{1}{2}} - E_{\frac{-1}{2}}}\\
+  &= \frac{\hbar^2\omega_2^2}{2\hbar\omega_0}
+  = \frac{\hbar\omega_2^2}{2\omega_0}\\
 \end{align*}
 
-To minimize this expectation value, $a-b = b-a = \pi$. Thus we represent our system as follows:
+\begin{align*}
+  E_{0} &=
+  \frac{|H'_{\frac{-1}{2},0}|^2}{E_{0} - E_{\frac{-1}{2}}} +
+  \frac{|H'_{\frac{-1}{2},\frac{1}{2}}|^2}{E_{\frac{1}{2}} - E_{\frac{-1}{2}}}\\
+  &= \frac{\hbar^2\omega_2^2}{2\hbar\omega_0}
+  = \frac{\hbar\omega_2^2}{2\omega_0}\\
+\end{align*}
 
 \begin{align*}
-  \psi(t) &= 0.6e^{i\pi}e^{-3i\omega t/2}\ket{1} + 0.8e^{-5i\omega t/2}\ket{2}\\
-  &= -0.6e^{-3i\omega t/2}\ket{1} + 0.8e^{-5i\omega t/2}\ket{2}\\
+  E_{\frac{1}{2}} &=
+  \frac{|H'_{\frac{-1}{2},0}|^2}{E_{0} - E_{\frac{-1}{2}}} +
+  \frac{|H'_{\frac{-1}{2},\frac{1}{2}}|^2}{E_{\frac{1}{2}} - E_{\frac{-1}{2}}}\\
+  &= \frac{\hbar^2\omega_2^2}{2\hbar\omega_0}
+  = \frac{\hbar\omega_2^2}{2\omega_0}\\
 \end{align*}
 
-We then calculate expectation values for the system:\\
-
-\textbf{Momentum}
-\begin{align*}
-  <p> &= -i\sqrt{\frac{\hbar}{2m\omega}}\bra{\psi}\left(a - a^\dagger)\right)\ket{\psi}\\
-  &= -i\sqrt{\frac{\hbar m\omega}{2}}\left(\bra{\psi}a\ket{\psi}
-  - \bra{\psi}a^\dagger\ket{\psi}\right)\\
-  &= -i\sqrt{\frac{\hbar m\omega}{2}}\left(\sqrt{2}c_1^*(t)c_2(t)\bra{1}\ket{1}
-  - \sqrt{2}c_1(t)c_2^*(t)\bra{2}\ket{2}\right)\\
-  &= -i*0.6*0.8*\sqrt{\hbar m\omega}\left(e^{i\omega t} - e^{-i\omega t}\right)\\
-  &= -0.48\sqrt{\hbar m\omega}\sin(\omega t)\\
-\end{align*}
-
-An interesting result! First, the units look good. Second, the magnitude of the expected momentum
-integrated over time is zero, meaning that the particle won't travel anywhere, which is good. The
-trigonometric function suggests that the particle's momentum oscillates sinusoidally at the
-classical frequency, just as we expect for a classical oscillator.\\
-
-\textbf{Energy}
-\begin{align*}
-  <H> &= \bra{1}(\frac{3}{2}\hbar\omega)\ket{1} + \bra{2}(\frac{5}{2}\hbar\omega)\ket{2}\\
-  &= 0.36 * \frac{3}{2}\hbar\omega + 0.64 * \frac{5}{2}\hbar\omega\\
-  &= 2.08\hbar\omega\\
-\end{align*}
-
-This is an expected result closest to the most-represented eigenstate in the system. It makes sense
-that there is no time-dependence, since that would mean energy would be entering and leaving the
-system, and energy is conserved for QM expectation values.
-
-\section{9.17: More Expectation Values}
-Similar to the above problem we are given a wavefunction and asked to calculate expectation values
-from it. We begin by figuring out its normalization constant:
-
-\textbf{Normalization constant}
-\begin{align*}
-  \psi(x,0) &= A\left[\phi_0(x) + 2\phi_1(x) + 2\phi_2(x)\right]\\
-  \braket{\psi} &= 1
-  A^2\left(1 + 4 + 4\right) &= 1 \rightarrow A = \frac{1}{3}\\
-\end{align*}
-
-Therefore our full time-evolved wavefunction is:
-
-\begin{align*}
-  \psi(x,t) &= \frac{1}{3}e^{\frac{-i\omega t}{2}}\phi_0(x)
-  + \frac{2}{3}e^{\frac{-3i\omega t}{2}}\phi_1(x)
-  + \frac{2}{3}e^{\frac{-5i\omega t}{2}}\phi_2(x)\\
-\end{align*}
-
-\textbf{Energy measurements}
-By inspection we see that the first three energy levels of the QHO spectrum are represented in the
-wave function, so these are the only ones we calculate the energy for.
-
-\begin{align*}
-  \mathbb{P}(\frac{1}{2}\omega) &= |\bra{0}\ket{\psi}|^2 = |c_0|^2 = \frac{1}{9}\\
-  \mathbb{P}(\frac{3}{2}\omega) &= |\bra{1}\ket{\psi}|^2 = |c_1|^2 = \frac{4}{9}\\
-  \mathbb{P}(\frac{5}{2}\omega) &= |\bra{2}\ket{\psi}|^2 = |c_2|^2 = \frac{4}{9}\\
-\end{align*}
-
-We can see the probabilities sum up to 1, as desired.\\
-
-\textbf{Momentum expectation value}
-\begin{align*}
-  <p> &= -i\sqrt{\frac{\hbar}{2m\omega}}\bra{\psi}\left(a - a^\dagger)\right)\ket{\psi}\\
-  &= -i\sqrt{\frac{\hbar m\omega}{2}}\left(\bra{\psi}a\ket{\psi}
-  - \bra{\psi}a^\dagger\ket{\psi}\right)\\
-  &= -i\sqrt{\frac{\hbar m\omega}{2}}
-  \Bigg(
-  \sqrt{1}\frac{2}{9}e^{\frac{i\omega t}{2}}e^{\frac{-3i\omega t}{2}}\bra{0}\ket{0}
-  + \sqrt{2}\frac{4}{9}e^{\frac{3i\omega t}{2}}e^{\frac{-5i\omega t}{2}}\bra{1}\ket{1}\\
-  &-\sqrt{1}\frac{2}{9}e^{\frac{3i\omega t}{2}}e^{\frac{-i\omega t}{2}}\bra{1}\ket{1}
-  - \sqrt{2}\frac{4}{9}e^{\frac{5i\omega t}{2}}e^{\frac{-3i\omega t}{2}}\bra{2}\ket{2}
-  \Bigg)\\
-  &= -i\sqrt{\frac{\hbar m\omega}{2}}
-  \left(\frac{2}{9}e^{-i\omega t} + \sqrt{2}\frac{4}{9}e^{-i\omega t}
-  - \frac{2}{9}e^{i\omega t} - \sqrt{2}\frac{4}{9}e^{i\omega t}\right)\\
-  &= \frac{-2i}{9}\sqrt{\frac{\hbar m\omega}{2}}
-  \Bigg(
-  \cos(\omega t) - i\sin(\omega t)
-  + \sqrt{8}\cos(\omega t) - i\sqrt{8}\sin(\omega t)\\
-  &- \cos(\omega t) - i\sin(\omega t)
-  - \sqrt{8}\cos(\omega t) - i\sqrt{8}\sin(\omega t)
-  \Bigg)\\
-  &= \frac{4i}{9}\sqrt{\frac{\hbar m\omega}{2}}\left(1 + \sqrt{8}\right)\sin(\omega t)\\
-\end{align*}
-
-This is another cool result. It is messier than the previous momentum expectation value due to
-a larger number of states, but still has proper units and displays the sinusoidal oscillation
-typical of systems with adjacent energies. I wonder why systems without adjacent energies have a
-zero expectation value for momentum. What is special about adjacent states that makes them
-``couple''?\\
-
-\textbf{Energy expectation}
-\begin{align*}
-  <H> &= \frac{1}{9}\frac{1}{2}\hbar\omega + \frac{4}{9}\hbar{3}{2}\hbar\omega
-  + \frac{4}{9}\frac{5}{2}\hbar\omega\\
-  &= \frac{33}{18}\hbar\omega = 1.83\hbar\omega\\
-\end{align*}
-
-A cool and expected result.\\
-
-\textbf{Energy standard deviation}
-We must first find the value of $<H^2>$:
-\begin{align*}
-  <H^2> &= \frac{1}{9}\frac{1}{4}\hbar^2\omega^2 + \frac{4}{9}\hbar{9}{4}\hbar^2\omega^2
-  + \frac{4}{9}\frac{25}{4}\hbar^2\omega^2\\
-  &= \frac{137}{36}\hbar^2\omega^2 = 3.8\hbar^2\omega^2\\
-\end{align*}
-
-We then calculate the standard deviation:
-\begin{align*}
-  \Delta E &= \sqrt{<H^2> - <H>^2} = \sqrt{(3.8 - 3.34)\hbar^2\omega^2}\\
-  &= \sqrt{0.46\hbar^2\omega^2} \approx \frac{2}{3}\hbar\omega\\
-\end{align*}
-
-Standard deviations in QHO systems are interesting because all the energies are evenly spaced. This
-value represents how spread-out the states are. It is expected that because the states are close
-together for this system, the standard deviation is small.
 \end{document}

### 2016-02-01 bioinformatics/hw1.py
Ssstuffff

@@ -1,6 +1,6 @@
 #! /usr/bin/python2.7
 
-from Bio import SeqIO
+from biopython import SeqIO
 
 seq = SeqIO.parse(open("sequence"), "fasta")
 for ref in seq:

### 2016-02-02 quantum/latex/a4.tex
all
Ssstuffff

@@
 then use the formula
-%$E_n^{(2)} = \sum_{m\noteq n}\frac{|\bra{n^{(0)}}H'\ket{n^{(0)}|^2}{(E_n^{(0)}-E_m^{(0)}}}}$
+$E_n^{(2)} = \sum_{m\neq n}\frac{|\bra{n^{(0)}}H'\ket{n^{(0)}}|^2}{(E_n^{(0)}-E_m^{(0)})}$
 to find the second-order energy corrections.\\
@@
 \begin{align*}
-  E_n^{(2)} = \sum_{m\noteq n}\frac{|\bra{n^{(0)}}H'\ket{n^{(0)}|^2}{(E_n^{(0)}-E_m^{(0)}}}}\\
+  E_n^{(2)} = \sum_{m\neq n}\frac{|\bra{n^{(0)}}H'\ket{n^{(0)}|^2}}{(E_n^{(0)}-E_m^{(0)})}
 \end{align*}
@@
 \begin{align*}
-  E_{\frac{-1}{2}} &=
-  \frac{|H'_{\frac{-1}{2},0}|^2}{E_{0} - E_{\frac{-1}{2}}} +
-  \frac{|H'_{\frac{-1}{2},\frac{1}{2}}|^2}{E_{\frac{1}{2}} - E_{\frac{-1}{2}}}\\
-  &= \frac{\hbar^2\omega_2^2}{2\hbar\omega_0}
-  = \frac{\hbar\omega_2^2}{2\omega_0}\\
+  E_1^{(2)} &=
+  \frac{|H'_{1,2}|^2}{E_2 - E_1} +
+  \frac{|H'_{1,3}|^2}{E_{3} - E_{1}}\\
+  &= \frac{\hbar^2\omega_2^2}{4\hbar\omega_0}\\
+  &= \frac{\hbar\omega_2^2}{4\omega_0}\\
 \end{align*}
@@
 \begin{align*}
-  E_{0} &=
-  \frac{|H'_{\frac{-1}{2},0}|^2}{E_{0} - E_{\frac{-1}{2}}} +
-  \frac{|H'_{\frac{-1}{2},\frac{1}{2}}|^2}{E_{\frac{1}{2}} - E_{\frac{-1}{2}}}\\
-  &= \frac{\hbar^2\omega_2^2}{2\hbar\omega_0}
-  = \frac{\hbar\omega_2^2}{2\omega_0}\\
+  E_2^{(2)} &=
+  \frac{|H'_{2,1}|^2}{E_1 - E_2} +
+  \frac{|H'_{2,3}|^2}{E_3 - E_2}\\
+  &= -\frac{\hbar^2\omega_2^2}{4\hbar\omega_0} + \frac{\hbar^2\omega_2^2}{4\hbar\omega_0}\\
+  &= 0\\
 \end{align*}
@@
 \begin{align*}
-  E_{\frac{1}{2}} &=
-  \frac{|H'_{\frac{-1}{2},0}|^2}{E_{0} - E_{\frac{-1}{2}}} +
-  \frac{|H'_{\frac{-1}{2},\frac{1}{2}}|^2}{E_{\frac{1}{2}} - E_{\frac{-1}{2}}}\\
-  &= \frac{\hbar^2\omega_2^2}{2\hbar\omega_0}
-  = \frac{\hbar\omega_2^2}{2\omega_0}\\
+  E_3^{(2)} &=
+  \frac{|H'_{3,1}|^2}{E_1 - E_3} +
+  \frac{|H'_{3,2}|^2}{E_2 - E_3}\\
+  &= -\frac{\hbar\omega_2^2}{4\omega_0}\\
 \end{align*}
 
+These are nicely symmetric energy corrections typical of physics problems. From this we can see
+that the ratio between the initial and perturbation Larmor frequencies is the only factor in
+deciding the magnitude of the perturbation. It is kind of weird that having a big initial
+Larmor frequency decreases the magnitude of the perturbation energy.\\
+
+We now plot our results:\\
+\vspace{7cm}
+
+As expected, we see quadratic dependence of the energy change on on perturbation strength for the
+second-order correction.
+\section{9.17: Ramp correction to Infinite Square Well}
+Here we find the first-order energy corrections for an infinite square well wavefunction where a
+ramp potential $H' = V'(x) = \beta x$ is applied. Because the correction Hamiltonian is in the
+x basis, we'll need to use wave function notation to find the ground-state energy correction:
+
+\begin{align*}
+  E_1^{(1)} &= \bra{1}H'\ket{1}\\
+  &= \frac{2\beta}{L}\int_0^L \sin^2\left(\frac{\pi x}{L}\right)x dx\\
+  &= \frac{\beta L}{2}\\
+\end{align*}
+
+This shows how easy it is to apply perturbation theory to find the energy corrections for simple
+systems. The units make sense and as we expect, both $\beta$ and L increase the energy shift.\\
+
+\section{10.18: Dirac delta Infinite Square Well}
+In this problem we deal with an infinite square well which has been perturbed by a delta function:
+$H' = V(x) = LV_0\delta(x-L/2)$\\
+
+\textbf{a.} First, we calculate the nth energy perturbation $E_n^1$ using wave function notation:
+
+\begin{align*}
+  E_n^{(1)} &= \bra{n}H'\ket{n}\\
+  &= \int_0^L 2\sin^2\left(\frac{n\pi x}{L}\right)V_0\delta(x-L/2)dx\\
+  &= 2V_0\sin^2\left(\frac{n\pi}{2}\right)\\
+  \begin{cases}
+    2V_0 & \mbox{n is odd}\\
+    0 & \mbox{n is even}\\
+  \end{cases}\\
+\end{align*}
+
+\textbf{b.}
+This result is based on the parity of the function. Even-parity functions have zero density at the
+center of the well, whereas odd-parity functions have nonzero density there. Therefore conceptually
+we can say that the delta potential only interacts with the odd functions; the even functions tunnel
+right through.\\
+
+\textbf{c.}
+The corrected wave function is given by:
+
+\begin{align*}
+  \ket{1} = \sum_{m\neq1} \frac{2mL^2\bra{m}H'\ket{m}}{\left(m^2-1\right)\hbar^2\pi^2}\\
+\end{align*}
+
+This value will be maximized when m is the smallest odd number larger than 1, i.e, 3. Thus the
+ground state is corrected most by the $\ket{3^{(0)}}$ state. This makes sense, since in general
+similar-energy waves correct each other the most. This conceptually makes sense, since the wave
+would take on characteristics of those similar to it in energy.\\
+
 \end{document}

### 2016-02-02 quantum/latex/a4.tex
almost done!

@@
 
+\section{10.24: Degenerate Perturbation}
+Here we are given just the perturbed Hamiltonian of a system and are asked to find the first
+nonvanishing energy correction to each of the four states. The perturbed Hamiltonian is:
+
+\begin{equation*}
+  H \doteq
+  \begin{pmatrix}
+    3 & \epsilon & 0 & 0\\
+    \epsilon & 3 & 2\epsilon & 0\\
+    0 & 2\epsilon & 5 & \epsilon\\
+    0 & 0 & \epsilon & 7\\
+  \end{pmatrix}
+\end{equation*}
+
+\textbf{a.}
+We are told that there are no diagonal perturbation elements, and so because the Hamiltonian is in
+the original basis we write our states as follows:
+
+\begin{equation*}
+  3V_0,
+  \begin{pmatrix}
+    1 \\ 0 \\ 0 \\ 0\\
+  \end{pmatrix}
+  \hspace{2cm}
+  3V_0,
+  \begin{pmatrix}
+    0 \\ 1 \\ 0 \\ 0\\
+  \end{pmatrix}
+  \hspace{2cm}
+  5V_0,
+  \begin{pmatrix}
+    0 \\ 0 \\ 1 \\ 0\\
+  \end{pmatrix}
+  \hspace{2cm}
+  7V_0,
+  \begin{pmatrix}
+    0 \\ 0 \\ 0 \\ 1\\
+  \end{pmatrix}
+  \hspace{2cm}
+\end{equation*}
+
+\textbf{b.}
+First-order corrections always manifest as terms on the Hamiltonian's diagonal. Our diagonal has
+no such terms, and so all corrections are 2nd-order or higher. We first solve for the nondegenerate
+$\ket{3}$ and $\ket{4}$ state energy corrections:
+
+\begin{align*}
+  E_3^{(2)} &=
+  \frac{|H'_{3,1}|^2}{E_1 - E_3} +
+  \frac{|H'_{3,2}|^2}{E_2 - E_3} +
+  \frac{|H'_{3,4}|^2}{E_4 - E_3}\\
+  &= \frac{4\epsilon^2}{-2V_0} + \frac{\epsilon^2}{-4V_0}\\
+  &= \frac{-9\epsilon^2}{4V_0}
+\end{align*}
+
+\begin{align*}
+  E_4^{(2)} &=
+  \frac{|H'_{4,1}|^2}{E_1 - E_4} +
+  \frac{|H'_{4,2}|^2}{E_2 - E_4} +
+  \frac{|H'_{4,3}|^2}{E_3 - E_4}\\
+  &= \frac{-\epsilon^2}{2V_0}
+\end{align*}
+
 \end{document}

### 2016-02-02 quantum/latex/a4.tex
stuff
Add completed? a4.tex

@@
   &= 2V_0\sin^2\left(\frac{n\pi}{2}\right)\\
-  \begin{cases}
+  &= \begin{cases}
     2V_0 & \mbox{n is odd}\\
@@
 
+Now we solve for the perturbed states. We begin with the order-one correction equation:\\
+
+\begin{align*}
+  \left(H_0 - E_n^0\right)\ket{2^{(1)}} &= \left(E_n^{(1)} - H'\right)\ket{n^{(0)}}\\
+\end{align*}
+
+We then populate this equation with our values, solving for $E_1^1$:\\
+
+\begin{align*}
+  \begin{pmatrix}
+    0 & 0 & 0 & 0\\
+    0 & 0 & 0 & 0\\
+    0 & 0 & 2 & 0\\
+    0 & 0 & 0 & 4\\
+  \end{pmatrix}
+  \begin{pmatrix}
+    c_{11}\\
+    c_{22}\\
+    c_{33}\\
+    c_{44}\\
+  \end{pmatrix}
+  &=
+  \begin{pmatrix}
+    E_1^1 - H_{11}' & -H_{12}' & -H_{13}' & -H_{14}'\\
+    -H_{21}' & E_1^1 - H_{22}' & -H_{23}' & -H_{24}'\\
+    -H_{31}' & -H_{32}' & E_1^1 - H_{33}' & -H_{34}'\\
+    -H_{41}' & -H_{42}' & -H_{43}' & E_1^1 - H_{44}'\\
+  \end{pmatrix}
+  \begin{pmatrix}
+    \alpha\\
+    \beta\\
+    0\\
+    0\\
+  \end{pmatrix}
+\end{align*}
+
+This leads to the following system of equations:\\
+
+\begin{align*}
+  \left(E_1^{(1)} - H_{11}'\right)\alpha + H_{12}'\beta &= 0\\
+  H_{21}'\alpha + \left(E_1^{(1)} - H_{22}'\right)\beta &= 0\\
+\end{align*}
+
+which can be turned into an energy eigenvalue equation:\\
+
+\begin{align*}
+  \begin{pmatrix}
+    H_{11}' & H_{12}'\\
+    H_{21}' & H_{22}'\\
+  \end{pmatrix}
+  \begin{pmatrix}
+    \alpha\\
+    \beta\\
+  \end{pmatrix}
+  &=
+  E_1^{(1)}
+  \begin{pmatrix}
+    \alpha\\
+    \beta\\
+  \end{pmatrix}\\
+  \begin{pmatrix}
+    0 & \epsilon\\
+    \epsilon & 0\\
+  \end{pmatrix}
+  \begin{pmatrix}
+    \alpha\\
+    \beta\\
+  \end{pmatrix}
+  &=
+  E_1^{(1)}
+  \begin{pmatrix}
+    \alpha\\
+    \beta\\
+  \end{pmatrix}\\
+\end{align*}
+
+We then solve for the eigenvalues of this system and get the following result:
+
+\begin{align*}
+  E_1^{(1)} &= \pm \epsilon\\
+\end{align*}
+
 \end{document}

### 2016-02-03 bioinformatics/latex/a2.tex
biomatics stuff
all

--- 2015-11-28 classical/latex/a9.tex
@@
 \usepackage{amsmath}
-\usepackage{marvosym} % Required for symbols in the colored box
-\usepackage{ifsym} % Required for symbols in the colored box
+\usepackage{listings}
 
@@
 
-\title{Classical Mechanics Assignment 9}
+\title{Bioinformatics Assignment 2}
 \author{Elliott Capek}
@@
 
-\section{Problem One: }
+\section{Problem One: Scoring Alignments}
+Here we are asked to compute the score for different given alignments.\\
+
+\textbf{A.} 3\\
+\textbf{B.} 1\\
+
+\section{Problem Two: BLAST}
+
+\section{Problem Four: Database size}
+We use the definition of E-value to compute database size, using an E-value of $e^{-27}:\\
+
+\begin{align*}
+  m &= \frac{E2^{S'}}{n} = \frac{e^{-27}2^{121}}{150} = 3.3*10^{22}\\
+\end{align*}
+
+\section{Problem Seven: GC Content}
+GC content is conserved when taking reverse complement, since Gs and Cs map to
+each other.\\
+
+For original probability $p_{N0}$ and reverse complement probability $P_N'$:\\
+
+\begin{align*}
+  p_{G0} &= p_C'\\
+  p_{C0} &= p_G'\\
+  p_G &= \frac{p_{G0} + p_G'}{2} = \frac{p_{G0} + p_{C0}}{2} = \frac{0.63}{2} = 0.315\\
+  p_C &= 0.315\\
+  p_A &= 0.185\\
+  p_T &= 0.185\\
+\end{align*}
+
+\section{Problem Eight: MEME}
+Here we enter some MEME commands to find motifs:\\
+\textbf{A.} meme proteins.fasta -protein -mod zoops -minw 10\\
+\textbf{B.} meme sequences.fa -rna -mod anr -minw 5 -maxw 10\\
+\textbf{C.} meme cores.fasta -dna -mod oops -nmotifs 5 \# -revcomp not specified,
+only search forward strand\\
 

### 2016-02-04 bioinformatics/latex/a2.tex
all

@@
 Here we are asked to compute the score for different given alignments.\\
-
 \textbf{A.} 3\\
@@
 \section{Problem Two: BLAST}
+When BLASTp'ing the \textit{Drosophila melanogaster} hedgehog gene, six results come up when
+compared to the human protein database. These all seme like ``valid'' hits just by looking
+at their names (they are all ``hedgehogs'' of some type). When looking for valid hits, identity
+and query cover are very important. Having high query cover, as do the ``desert'' and ``indian''
+hedgehog genes, means the genes occupy roughly the same space, ie one is not a subcomponent of
+the other. Having a high identity means the genes are similar. Another important factor is having
+a low E-value. All the elements in this list have low E-values, meaning it is unlikely any of
+them are just random hits. Total score changes based on the lengths of the sequences, so is
+more difficult to interpret.\\
+
+\section{Problem Three: Computing frequency matrix elements}
+In this problem we use the following information: (2S, 4N, 2T) to compute various frequency,
+normalized frequency and probability elements:\\
+
+\begin{align*}
+  f &= \mbox{Number of occurrences}\\
+  q &= \mbox{Number of occurences normalized}\\
+  p &= \mbox{Number of pairs involving this element, double cound identical pairs}\\
+\end{align*}
+
+\begin{align*}
+  f_{SS} &= 1\\
+  f_{TT} &= 4 * 3 = 12\\
+  f_{NN} &= 1\\
+  f_{ST} &= 2 * 4 = 8\\
+  f_{NS} &= 2 * 2 = 4\\
+  f_{NT} &= 2 * 4 = 8\\
+  q_{SS} &= 1 / 34\\
+  q_{TT} &= 12 / 34 = 6 / 17\\
+  q_{NN} &= 1 / 34\\
+  q_{ST} &= 8 / 34 = 4 / 17\\
+  q_{NS} &= 4 / 34 = 2 / 17\\
+  q_{NT} &= 4 / 17\\
+  p_S &= 1 + 6 = 7\\
+  p_T &= 12 + 1 = 13\\
+  p_N &= 1 + 6 = 7\\
+\end{align*}
 
 \section{Problem Four: Database size}
-We use the definition of E-value to compute database size, using an E-value of $e^{-27}:\\
+We use the definition of E-value to compute database size, using an E-value of $10^{-27}$:\\
 
 \begin{align*}
-  m &= \frac{E2^{S'}}{n} = \frac{e^{-27}2^{121}}{150} = 3.3*10^{22}\\
+  m &= \frac{E*2^{S'}}{n} = \frac{10^{-27}*2^{121}}{150} \approx 1.77*10^7\\
 \end{align*}
+
+This is about 18 million residues. This seems small for a database, since this would be around
+60 million nucleotides, which is two orders of magnitude smaller than the number of nucleotides
+in the human genome. However, perhaps it is only a section of the genome, or perhaps it is for
+a smaller organism.\\
+
+\section{Problem Five: Dinosaur protein}
+Here we begin with the given sequence and use BioPython to transcribe() and translate() it into:\\
+
+MVHWTALITGHWGGKAECGAEADDARVVILYPLAVFASFGNLSDTAILGNPMVRAHKLSFGDAVNLDNIKN
+TIIFSWQLLHCDLHENFRLLGDIDAVEWASHEREIVSKDFDPECCQAAWQKLVRVAHAKYH\\
+
+We then run a BLAST nonredundant database protein search on it. We get dozens of hits with 100\%
+query cover and roughly 65\% identity. Such a high query cover indicates that the entire sequence
+is used in the alignment, meaning the given and matched sequences are likely to be the same gene.
+The low identity means that there is a lot of difference between the given and matched sequences.
+Most of these sequences have a very low E-value (roughly $10^{-35}$), meaning that it is highly
+unlikely that our results are due to chance.\\
+
+This all comes together to mean that this could be a dinosaur Hemoglobin, since the matches are
+of high accuracy and there is lots of mutation.\\
+
+\section{Problem Six: Orca homeobox conservation}
+After running a BLASTp nonredundant search on the \textit{Orcinus orca} protein sequence, I found
+that there was a 99\% identity between the orca and \textit{Sus scrofa} sequences, and only a
+96\% similarity between orca and \textit{Bos taurus} sequences. There was also a 100\% sequence
+cover with the boar, but only 92\% with the cow. This means that the orca sequence is almost
+identical to the boar sequence, meaning orcas and boars are a lot more closely related. The
+E-values were very low $10^{-175}$, meaning it is incredibly unlikely that these sequences were
+randomly related.\\
 

### 2016-02-07 bioinformatics/latex/midterm.tex
Add bioinformatics midterm.

--- 2016-02-03 bioinformatics/latex/a2.tex
@@
 
-\title{Bioinformatics Assignment 2}
+\title{Bioinformatics Midterm}
 \author{Elliott Capek}
@@
 \maketitle{}
+\section{One: DP Alignment Algorithms}
+Smith-Waterman and Needleman-Wunsch are two very similar dynamic programming
+algorithms for finding the optimal alignment of a set of text sequences. The
+basic premise of the algorithms is filling in a table that corresponds to every
+possible alignment that can be made, then choosing the optimal alignment based
+on scoring guidelines. The reason this isn't a ridiculously slow process is
+because the table represents not every possible alignment, but the branch points
+between different alignments.\\
 
-\section{Problem One: Scoring Alignments}
-Here we are asked to compute the score for different given alignments.\\
+Smith-Waterman (SW) is used to find global alignments, so the entire sequences matter.
+Needleman-Wunsch (NW) is used to find local alignments, where only a subset of the
+sequence is to be considered. In order to only consider a subset of the sequence,
+two modifications to SW are made to achieve NW. These changes effectively limit
+the portion of the DP table being considered. The first change is that no section in
+the table can have a score lower than the score of the top-left (start of global
+alignment) element. This conceptually means that the alignment can begin anywhere
+in the sequence, since having the same score as the global alignment start means
+that any table element can be treated as the alignment start.\\
 
-\textbf{A.} 3\\
-\textbf{B.} 1\\
+The second change is that the NW alignment can end anywhere, not at the bottom-right
+corner of the table. This effectively means that the local alignmemnt can end before
+the entire sequence is considered.\\
 
-\section{Problem Two: BLAST}
-
-\section{Problem Four: Database size}
-We use the definition of E-value to compute database size, using an E-value of $e^{-27}:\\
-
-\begin{align*}
-  m &= \frac{E2^{S'}}{n} = \frac{e^{-27}2^{121}}{150} = 3.3*10^{22}\\
-\end{align*}
-
-\section{Problem Seven: GC Content}
-GC content is conserved when taking reverse complement, since Gs and Cs map to
-each other.\\
-
-For original probability $p_{N0}$ and reverse complement probability $P_N'$:\\
-
-\begin{align*}
-  p_{G0} &= p_C'\\
-  p_{C0} &= p_G'\\
-  p_G &= \frac{p_{G0} + p_G'}{2} = \frac{p_{G0} + p_{C0}}{2} = \frac{0.63}{2} = 0.315\\
-  p_C &= 0.315\\
-  p_A &= 0.185\\
-  p_T &= 0.185\\
-\end{align*}
-
-\section{Problem Eight: MEME}
-Here we enter some MEME commands to find motifs:\\
-\textbf{A.} meme proteins.fasta -protein -mod zoops -minw 10\\
-\textbf{B.} meme sequences.fa -rna -mod anr -minw 5 -maxw 10\\
-\textbf{C.} meme cores.fasta -dna -mod oops -nmotifs 5 \# -revcomp not specified,
-only search forward strand\\
+\section{Two: }
 

### 2016-02-07 bioinformatics/latex/midterm.tex
Add work on 1-3 for ba midterm.

@@
 \section{One: DP Alignment Algorithms}
-Smith-Waterman and Needleman-Wunsch are two very similar dynamic programming
+Needleman-Wunsch and Smith-Waterman are two very similar dynamic programming
 algorithms for finding the optimal alignment of a set of text sequences. The
@@
 
-Smith-Waterman (SW) is used to find global alignments, so the entire sequences matter.
-Needleman-Wunsch (NW) is used to find local alignments, where only a subset of the
-sequence is to be considered. In order to only consider a subset of the sequence,
-two modifications to SW are made to achieve NW. These changes effectively limit
+Needleman-Wunsch (NW) is used to find global alignments, so the entire sequences
+matter. Smith-Waterman (SW) is used to find local alignments, where only a subset of
+the sequence is to be considered. In order to only consider a subset of the sequence,
+two modifications to NW are made to achieve SW. These changes effectively limit
 the portion of the DP table being considered. The first change is that no section in
@@
 
-The second change is that the NW alignment can end anywhere, not at the bottom-right
-corner of the table. This effectively means that the local alignmemnt can end before
-the entire sequence is considered.\\
+The second change is that the SW alignment ``ends'' at the highest-scoring value,
+not at the bottom-right corner of the table. This effectively means that the local
+alignment can end before the entire sequence is considered.\\
 
-\section{Two: }
+\section{Two: BLAST Scores}
+\textbf{E-Value:} The expected number of hits that score equal to or higher than a
+particular hit for a given query sequence. Conceptually similar to how unreliable a
+score is, but also takes into account the size of the database and sequence length.
+
+\begin{align*}
+  E &= Knme^{-\lambda S}\\
+\end{align*}
+
+Inverse exponential relationship to score means larger scores have lower E-values,
+as expected since higher scores correspond to better alignments. Query sequence length
+n also important, since longer sequences are more likely to have a random local
+alignment in them. Database size m is important for the same reason: more sequences
+in the database lead to a higher chance of getting a random hit.\\
+
+The E-value is a weird number in that it isn't for a particular score, but for any
+score greater than a particular number. It is useful for assessing how useful a hit
+is. If the hit has a high E-value, then there is a good chance that it is a random
+hit that shouldn't be trusted.\\
+
+\textbf{p-Value:} The probability of getting a score equal to or higher than the score
+of a given hit for a particular query due to chance. Given by:
+
+\begin{align*}
+  P(S\geq x) &= 1 - e^{-Knme^{-\lambda x}}\\
+\end{align*}
+
+As score threshold x gets larger, the probability gets smaller, as expected. The two
+terms in this presumably come from integrating over the Poisson distribution, where
+1 is the probability of having less than the maximum score and the term on the right
+is the probability of having a score less than x.\\
+The p-value is used to calculate E, which is more useful when evaluating the validity
+of queries.\\
+
+\section{Three: BLOSUM Matrix}
+Here we calculate BLOSUM matrix elements for an alignment with
+\textbf{5 R, 2H, 2 L}.\\
+
+\begin{align*}
+  f_{RR} &= \begin{pmatrix} n_R\\ 2\\ \end{pmatrix} = 10\\
+  f_{HH} &= \begin{pmatrix} n_H\\ 2\\ \end{pmatrix} = 1\\
+  f_{LL} &= \begin{pmatrix} n_L\\ 2\\ \end{pmatrix} = 1\\
+  f_{HL} &= n_H \times n_L = 4\\
+  f_{HR} &= n_H \times n_R = 10\\
+  f_{LR} &= n_L \times n_R = 10\\
+  q_{RR} &= \frac{f_{RR}}{f_{RR} + f_{HH} + f_{LL} + f_{HL} + f_{HR} + f_{LR}}
+  = \frac{10}{36} = 0.277\\
+  q_{HH} &= \frac{f_{HH}}{f_{RR} + f_{HH} + f_{LL} + f_{HL} + f_{HR} + f_{LR}}
+  = \frac{1}{36} = 0.027\\
+  q_{LL} &= \frac{f_{LL}}{f_{RR} + f_{HH} + f_{LL} + f_{HL} + f_{HR} + f_{LR}}
+  = \frac{1}{36} = 0.027\\
+  q_{HL} &= \frac{f_{HL}}{f_{RR} + f_{HH} + f_{LL} + f_{HL} + f_{HR} + f_{LR}}
+  = \frac{4}{36} = 0.111\\
+  q_{HR} &= \frac{f_{HR}}{f_{RR} + f_{HH} + f_{LL} + f_{HL} + f_{HR} + f_{LR}}
+  = \frac{10}{36} = 0.277\\
+  q_{LR} &= \frac{f_{LR}}{f_{RR} + f_{HH} + f_{LL} + f_{HL} + f_{HR} + f_{LR}}
+  = \frac{10}{36} = 0.277\\
+  p_R &= q_{RR} + \frac{1}{2}\left(q_{RH} + q_{RL}\right) = 0.277 +
+  \frac{0.277 + 0.277}{2} = 0.558\\
+  p_H &= q_{HH} + \frac{1}{2}\left(q_{HR} + q_{HL}\right) = 0.027 +
+  \frac{0.111 + 0.277}{2} = 0.221\\
+  p_L &= q_{LL} + \frac{1}{2}\left(q_{LH} + q_{LR}\right) = 0.027 +
+  \frac{0.111 + 0.277}{2} = 0.221\\
+\end{align*}
+
+Here all q-terms and p-terms sum to one, as desired.
 

### 2016-02-08 bioinformatics/latex/midterm.tex
add stufffff
Add changes to midterm.

@@
 
+\section{Four: BLAST flubber}
+
+\textbf{A.)} blastall -p blastp -i flubber.fasta -o flubber\_blast.txt\\
+\textbf{B.)} The top his is for protein FluB of \textit{Actinomadura vulgaris}.
+The highest scoring alignment has a score of 754, the total score is 1213 and the
+E-value is 0.\\
+\textbf{C.)} E-value calculation:
+
+\begin{align*}
+  E &= Knme^{-\lambda S}\\
+  &= K(2.5*10^{10})(417)e^{-\lambda * 1213}\\
+\end{align*}
+
+\section{Five: MEME commands}
+
+Check this with the homework on your desk and online if possible
+
+\textbf{A.)} meme peptides.fasta -protein -mod anr -minw 7 -maxw 7\\
+\textbf{B.)} meme enhancers.fa -dna -mod oops -maxw 10\\
+
+\section{Six: blastall commands}
+\textbf{A.)} formatdb -p F -t midichlorians -n midichlorians -i yoda.microbiome.fasta\\
+\textbf{B.)} blastall -p blastp -i transcripts.fa -d myProteins  -g 2 -o transcripts.txt # gap pen rite??\\
+\textbf{C.)} blastall -p blastp -i proteins.fa -d myGenome -o proteins.txt\\
+
+\section{Eight: Revcomp GC}
+
+from Bio import SeqIO
+
+seq = Bio.Seq(``sequence'')
+print ``Is GC content of sequence and sequence.revcomp equal?''
+print seq.cg() == seq.revcomp.cg()
+
+Then try it on sequence...
+
+\section{Nine: }
+
 \end{document}

### 2016-02-11 bioinformatics/latex/midterm.tex
I don't knwo what I'm doing.

@@
 \begin{align*}
-  E &= Knme^{-\lambda S}\\
-  &= K(2.5*10^{10})(417)e^{-\lambda * 1213}\\
+  E &= nm*2^{-S'}\\
+  &= (2.5*10^{10})(417)2^{-1213}\\
+  &= 7.4*10^{-353} \approx 0\\
 \end{align*}
@@
 
-\textbf{A.)} meme peptides.fasta -protein -mod anr -minw 7 -maxw 7\\
-\textbf{B.)} meme enhancers.fa -dna -mod oops -maxw 10\\
+\textbf{A.)} meme peptides.fasta -protein -mod anr -nmotifs 3 -minw 7 -maxw 7\\
+\textbf{B.)} meme enhancers.fa -dna -mod oops -nmotifs 1 -maxw 10\\
 
@@
 \textbf{A.)} formatdb -p F -t midichlorians -n midichlorians -i yoda.microbiome.fasta\\
-\textbf{B.)} blastall -p blastp -i transcripts.fa -d myProteins  -g 2 -o transcripts.txt # gap pen rite??\\
+\textbf{B.)} blastall -p blastp -i transcripts.fa -d myProteins -G 2 -o transcripts.txt\\
 \textbf{C.)} blastall -p blastp -i proteins.fa -d myGenome -o proteins.txt\\
@@
 \section{Eight: Revcomp GC}
+from Bio.SeqUtils import GC\\
+from Bio import SeqIO\\
+seq = Bio.Seq(``GCGTAGCTGATCACCAGCTAGCGACGTATCAAAAACGCTGACGGGGTA'')\\
+print ``Are GC contents of sequence and sequence.reverse\_complement() equal?''\\
+print GC(seq) == GC(seq.reverse\_complement())\\
+>>> True!\\
 
-from Bio import SeqIO
+\section{Nine: DNA concatenation}
+from Bio.Seq import Seq\\
+seq = Bio.Seq(``ACGTTACGAGGAGAGTC'')\\
+seq\_RNA = seq.transcribe()\\
+print seq\_RNA + seq\_RNA.reverse\_complement() + seq\_RNA\\
 
-seq = Bio.Seq(``sequence'')
-print ``Is GC content of sequence and sequence.revcomp equal?''
-print seq.cg() == seq.revcomp.cg()
+\section{Ten: PHRED scores}
 
-Then try it on sequence...
-
-\section{Nine: }
+\begin{align*}
+  p_{err}(Z) &= 10^{-\frac{ord(Z)+33}{10}} = 10^{-12.3} = 5*10^{-13}\\
+  p_{err}(z) &= 10^{-\frac{ord(z)+33}{10}} = 10^{-15.5} = 3.16*10^{-16}\\
+  p_{err}(:) &= 10^{-\frac{ord(:)+33}{10}} = 10^{-9.1} = 1.7*10^{-10}\\
+  p_{err}(+) &= 10^{-\frac{ord(+)+33}{10}} = 10^{-7.6} = 7.94*10^{-10}\\
+\end{align*}
 

### 2016-02-11 quantum/latex/a5.tex

--- 2016-02-02 quantum/latex/a4.tex
@@
 
-\title{Quantum Mechanics Assignment Three}
+\title{Quantum Mechanics Assignment Five}
 \author{Elliott Capek}
@@
 
-\section{10.8: Magnetic Perturbation Spin 1}
-For this problem we want to find the energy perturbations due to a magnetic field perturbation
-in the x direction, while remaining in the z basis. To do this we first find the
-perturbed Hamiltonian, which is just the x spin-1 matrix times a factor, and
-then use the formula
-$E_n^{(2)} = \sum_{m\neq n}\frac{|\bra{n^{(0)}}H'\ket{n^{(0)}}|^2}{(E_n^{(0)}-E_m^{(0)})}$
-to find the second-order energy corrections.\\
+\section{11.7: $\vec{S} \cdot \vec{I}$ simplified}
 
-We first find the perturbed Hamiltonian:
+Here we show how using angular momentum ladder operators can simplify the dot product
+$\vec{S} \cdot \vec{I}$:\\
 
 \begin{align*}
-  H' &= \vec{\mu} \cdot \vec{B_2}\\
-  &= \frac{g e}{2m} \hat{\mu} \cdot B_2 \hat{x}\\
-  &= \frac{\hbar\omega_2}{2} S_x
-  \hspace{2cm}\omega_2 = \frac{geB}{2m}\\
-  &=
-  \begin{pmatrix}
-    0 & \frac{\hbar\omega_2}{2} & 0\\
-    \frac{\hbar\omega_2}{2} & 0 & \frac{\hbar\omega_2}{2}\\
-    0 & \frac{\hbar\omega_2}{2} & 0\\
-  \end{pmatrix}
+  \vec{S} \cdot \vec{I} &= \hat{S}_x\hat{I}_x + \hat{S}_y\hat{I}_y + \hat{S}_z\hat{I}_z\\
+  &= \frac{1}{2}\left(\hat{S}_x\hat{I}_x -i\hat{S}_x\hat{I}_y + i\hat{S}_y\hat{I}_x
+  + \hat{S}_y\hat{I}_y + \hat{S}_x\hat{I}_x + i\hat{S}_x\hat{I}_y
+  + -i\hat{S}_y\hat{I}_x\hat{S}_y\hat{I}_y\right)
+  + \hat{S}_z\hat{I}_z\\
+  &= \frac{1}{2}\left(\hat{S}_+\hat{I}_- + \hat{S}_-\hat{I}_+\right) + \hat{S}_z\hat{I}_z\\
 \end{align*}
 
-Then we just apply the second-order perturbation energy equation to find our
-energy changes, noticing that there will be no first-order corrections because
-$H'$ has zeroes along its diagonal.
-
-\begin{align*}
-  E_n^{(2)} = \sum_{m\neq n}\frac{|\bra{n^{(0)}}H'\ket{n^{(0)}|^2}}{(E_n^{(0)}-E_m^{(0)})}
-\end{align*}
-
-\begin{align*}
-  E_1^{(2)} &=
-  \frac{|H'_{1,2}|^2}{E_2 - E_1} +
-  \frac{|H'_{1,3}|^2}{E_{3} - E_{1}}\\
-  &= \frac{\hbar^2\omega_2^2}{4\hbar\omega_0}\\
-  &= \frac{\hbar\omega_2^2}{4\omega_0}\\
-\end{align*}
-
-\begin{align*}
-  E_2^{(2)} &=
-  \frac{|H'_{2,1}|^2}{E_1 - E_2} +
-  \frac{|H'_{2,3}|^2}{E_3 - E_2}\\
-  &= -\frac{\hbar^2\omega_2^2}{4\hbar\omega_0} + \frac{\hbar^2\omega_2^2}{4\hbar\omega_0}\\
-  &= 0\\
-\end{align*}
-
-\begin{align*}
-  E_3^{(2)} &=
-  \frac{|H'_{3,1}|^2}{E_1 - E_3} +
-  \frac{|H'_{3,2}|^2}{E_2 - E_3}\\
-  &= -\frac{\hbar\omega_2^2}{4\omega_0}\\
-\end{align*}
-
-These are nicely symmetric energy corrections typical of physics problems. From this we can see
-that the ratio between the initial and perturbation Larmor frequencies is the only factor in
-deciding the magnitude of the perturbation. It is kind of weird that having a big initial
-Larmor frequency decreases the magnitude of the perturbation energy.\\
-
-We now plot our results:\\
-\vspace{7cm}
-
-As expected, we see quadratic dependence of the energy change on on perturbation strength for the
-second-order correction.
-\section{9.17: Ramp correction to Infinite Square Well}
-Here we find the first-order energy corrections for an infinite square well wavefunction where a
-ramp potential $H' = V'(x) = \beta x$ is applied. Because the correction Hamiltonian is in the
-x basis, we'll need to use wave function notation to find the ground-state energy correction:
-
-\begin{align*}
-  E_1^{(1)} &= \bra{1}H'\ket{1}\\
-  &= \frac{2\beta}{L}\int_0^L \sin^2\left(\frac{\pi x}{L}\right)x dx\\
-  &= \frac{\beta L}{2}\\
-\end{align*}
-
-This shows how easy it is to apply perturbation theory to find the energy corrections for simple
-systems. The units make sense and as we expect, both $\beta$ and L increase the energy shift.\\
-
-\section{10.18: Dirac delta Infinite Square Well}
-In this problem we deal with an infinite square well which has been perturbed by a delta function:
-$H' = V(x) = LV_0\delta(x-L/2)$\\
-
-\textbf{a.} First, we calculate the nth energy perturbation $E_n^1$ using wave function notation:
-
-\begin{align*}
-  E_n^{(1)} &= \bra{n}H'\ket{n}\\
-  &= \int_0^L 2\sin^2\left(\frac{n\pi x}{L}\right)V_0\delta(x-L/2)dx\\
-  &= 2V_0\sin^2\left(\frac{n\pi}{2}\right)\\
-  \begin{cases}
-    2V_0 & \mbox{n is odd}\\
-    0 & \mbox{n is even}\\
-  \end{cases}\\
-\end{align*}
-
-\textbf{b.}
-This result is based on the parity of the function. Even-parity functions have zero density at the
-center of the well, whereas odd-parity functions have nonzero density there. Therefore conceptually
-we can say that the delta potential only interacts with the odd functions; the even functions tunnel
-right through.\\
-
-\textbf{c.}
-The corrected wave function is given by:
-
-\begin{align*}
-  \ket{1} = \sum_{m\neq1} \frac{2mL^2\bra{m}H'\ket{m}}{\left(m^2-1\right)\hbar^2\pi^2}\\
-\end{align*}
-
-This value will be maximized when m is the smallest odd number larger than 1, i.e, 3. Thus the
-ground state is corrected most by the $\ket{3^{(0)}}$ state. This makes sense, since in general
-similar-energy waves correct each other the most. This conceptually makes sense, since the wave
-would take on characteristics of those similar to it in energy.\\
+This is a useful math trick with no physical meaning! Multiplying raising and lowering operators
+together has no physical significance, let alone operators for different particles. Nevertheless
+this will probably come in handy later.\\
 

### 2016-02-12 quantum/latex/a5.tex
Add 15.tex
errthang

@@
 
+\section{1: Effect of E-field on H spectrum}
+Here we examine how a perturbing electric field influences the n=2 to n=1 transition for
+the Hydrogen atom.\\
+
+\textbf{a.)} First we find the unperturbed transition energy:\\
+
+\begin{align*}
+  E_2 - E_1 &= -13.6 \mbox{eV} \left(\frac{1}{2^2} - \frac{1}{1^2}\right) = 10.2 \mbox{eV}\\
+\end{align*}
+
+\textbf{b.)} Now we apply a 0.5 V/m electric field and see how the spectrum changes. The Stark
+Effect tells us that the energy levels will be perturbed by either zero or
+$\pm3e\epsilon a_0$, which for this electric field would be $\pm 2.64*10^{-2}eV$. This comes out
+to be a change of roughly $0.25$ percent, which is a fairly small value.\\
+
+The perturbation due to the electric field can be interpreted as the energy change of the
+electric dipole moment interacting with the applied field. Some of the n=2 states are radially
+symmetric, and so have no electric dipole to alter the energy. However certain combinations of
+orbitals (ie $\bra{200}+\bra{210}$) do have a bias to part of the electron, and so do have an
+energy change.\\
+
 \section{11.7: $\vec{S} \cdot \vec{I}$ simplified}
@@
 
+\section{11.9: Commutators}
+Here we show that the spin magnitude operators \textbf{$S^2$} and \textbf{$I^2$} commute with
+the perturbation Hamiltonian, while the component operators \textbf{$S_z$} and \textbf{$I_z$}
+do not.
+
+\begin{align*}
+  \left[S^2, H'_{hf}\right] &= \left(S^2H'_{hf} - H'_{hf}S^2\right)\ket{m_p,m_s}\\
+  &= S^2H'_{hf}\ket{m_p,m_s} - H'_{hf}S^2\ket{m_p,m_s}\\
+  &= S^2E_{m_p,m_s}\ket{m_p,m_s} - H'_{hf}S^2\frac{3\hbar^2}{4}\ket{m_p,m_s}\\
+  &= \frac{3\hbar^2}{4}E_{m_p,m_s}\ket{m_p,m_s} - E_{m_p,m_s}S^2\frac{3\hbar^2}{4}\ket{m_p,m_s}\\
+  &= 0\\
+\end{align*}
+
 \end{document}

### 2016-02-12 quantum/latex/a5.tex
bla

@@
 Effect tells us that the energy levels will be perturbed by either zero or
-$\pm3e\epsilon a_0$, which for this electric field would be $\pm 2.64*10^{-2}eV$. This comes out
-to be a change of roughly $0.25$ percent, which is a fairly small value.\\
+$\pm3e\epsilon a_0$, which for this electric field would be $\pm 7.94*10^{-2}eV$. This comes out
+to be a change of roughly $0.77$ percent, which is a fairly small value.\\
 
@@
 energy change.\\
+
+\section{11.2: Proving raising / lowering operators}
+Here we show that the raising and lowering operators for angular momentum behave similarly
+to the QHO raising and lowering operators.\\
+
+First, we must show that the operators raise and lower the z-component momentum measurement
+by one unit:\\
+
+\begin{align*}
+  J_z(J_+\ket{j,m_j}) &= (\hbar J_+ + J_+J_z)\ket{j,m_j}\\
+  &= \hbar J_+\ket{j,m_j} + m_j \hbar J_+\ket{j,m_j}\\
+  J_z(J_+\ket{j,m_j}) &= (m_j+1)\hbar(J_z\ket{j,m_j})
+  \hspace{2cm}\mbox{J+ raises energy by $\hbar$}\\
+\end{align*}
+
+\begin{align*}
+  J_z(J_-\ket{j,m_j}) &= (-\hbar J_- + J_-J_z)\ket{j,m_j}\\
+  &= \hbar J_-\ket{j,m_j} + m_j \hbar J_-\ket{j,m_j}\\
+  J_z(J_-\ket{j,m_j}) &= (m_j-1)\hbar(J_z\ket{j,m_j})
+  \hspace{2cm}\mbox{J- lowers energy by $\hbar$}\\
+\end{align*}
+
+This shows that the operation of $J_\pm$ on a state before the operation of $J_z$ effectively
+raises or lowers the measurement of that state by $\hbar$. We now need to find the relationship
+between $J_\pm\ket{j,m_j}$ and $\ket{j,m_j\pm1}$.\\
+
+\begin{align*}
+  |J_+\ket{j,m_j}|^2 &= c^2\\
+  &= \bra{j,m_j}J_-J_+\ket{j,m_j}\\
+  &= \bra{j,m_j}(J_x-iJ_y)(J_x+iJ_y)\ket{j,m_j}\\
+  &= \bra{j,m_j}J_xJ_x + J_yJ_y + i[J_x,J_y]\ket{j,m_j}\\
+  &= \bra{j,m_j}J^2 - J_zJ_z + i[J_x,J_y]\ket{j,m_j}\\
+  &= \bra{j,m_j}J^2 - J_zJ_z - \hbar J_z\ket{j,m_j}\\
+  &= j(j+1)\hbar - m_j^2\hbar^2 - m_j\hbar^2\\
+  &= \hbar^2\left(j(j+1)\hbar - m_j(m_j + 1)\right)\\
+  &\rightarrow J_+\ket{j,m_j} = \hbar\sqrt{j(j+1) - m_j(m_j+1)}\ket{j,m_j+1}\\
+\end{align*}
+
+\begin{align*}
+  |J_-\ket{j,m_j}|^2 &= c^2\\
+  &= \bra{j,m_j}J_+J_-\ket{j,m_j}\\
+  &= \bra{j,m_j}(J_x+iJ_y)(J_x-iJ_y)\ket{j,m_j}\\
+  &= \bra{j,m_j}J_xJ_x + J_yJ_y - i[J_x,J_y]\ket{j,m_j}\\
+  &= \bra{j,m_j}J^2 - J_zJ_z - i[J_x,J_y]\ket{j,m_j}\\
+  &= \bra{j,m_j}J^2 - J_zJ_z + \hbar J_z\ket{j,m_j}\\
+  &= j(j+1)\hbar^2 - m_j^2\hbar^2 + m_j\hbar&2\\
+  &= \hbar^2\left(j(j+1) - m_j(m_j-1)\right)\\
+  &\rightarrow J_-\ket{j,m_j} = \hbar\sqrt{j(j+1) - m_j(m_j-1)}\ket{j,m_j-1}\\
+\end{align*}
+
+We've just shown that the angular momentum ladder operators behave the same way the quantum
+harmonic oscillator operators do. This is powerful, since it will allow us to simplify math
+later on. It is important to note that the raising and lowering operators only act on the
+Z-component of spin, and that they have two termination conditions: at the minimum and
+maximum Z-component.\\
+
+\textit{Inspired by Jared Cayton}\\
 
@@
 
-\section{11.9: Commutators}
-Here we show that the spin magnitude operators \textbf{$S^2$} and \textbf{$I^2$} commute with
-the perturbation Hamiltonian, while the component operators \textbf{$S_z$} and \textbf{$I_z$}
-do not.
+\section{11.8: Hyperfine Hamiltonian}
+Here we use ladder operators to find the hyperfine Hamiltonian for the Hydrogen atom:\\
 
 \begin{align*}
-  \left[S^2, H'_{hf}\right] &= \left(S^2H'_{hf} - H'_{hf}S^2\right)\ket{m_p,m_s}\\
-  &= S^2H'_{hf}\ket{m_p,m_s} - H'_{hf}S^2\ket{m_p,m_s}\\
-  &= S^2E_{m_p,m_s}\ket{m_p,m_s} - H'_{hf}S^2\frac{3\hbar^2}{4}\ket{m_p,m_s}\\
-  &= \frac{3\hbar^2}{4}E_{m_p,m_s}\ket{m_p,m_s} - E_{m_p,m_s}S^2\frac{3\hbar^2}{4}\ket{m_p,m_s}\\
-  &= 0\\
+  H_{hf}' &=
+  \begin{pmatrix}
+    \bra{++}H_{hf}'\ket{++} &
+    \bra{++}H_{hf}'\ket{+-} &
+    \bra{++}H_{hf}'\ket{-+} &
+    \bra{++}H_{hf}'\ket{--} &\\
+    \bra{+-}H_{hf}'\ket{++} &
+    \bra{+-}H_{hf}'\ket{+-} &
+    \bra{+-}H_{hf}'\ket{-+} &
+    \bra{+-}H_{hf}'\ket{--} &\\
+    \bra{-+}H_{hf}'\ket{++} &
+    \bra{-+}H_{hf}'\ket{+-} &
+    \bra{-+}H_{hf}'\ket{-+} &
+    \bra{-+}H_{hf}'\ket{--} &\\
+    \bra{--}H_{hf}'\ket{++} &
+    \bra{--}H_{hf}'\ket{+-} &
+    \bra{--}H_{hf}'\ket{-+} &
+    \bra{--}H_{hf}'\ket{--} &\\
+  \end{pmatrix}\\
+  &=
+  \frac{A}{\hbar^2}
+  \begin{pmatrix}
+    \bra{++}\vec{S}\cdot\vec{I}\ket{++} &
+    \bra{++}\vec{S}\cdot\vec{I}\ket{+-} &
+    \bra{++}\vec{S}\cdot\vec{I}\ket{-+} &
+    \bra{++}\vec{S}\cdot\vec{I}\ket{--} &\\
+    \bra{+-}\vec{S}\cdot\vec{I}\ket{++} &
+    \bra{+-}\vec{S}\cdot\vec{I}\ket{+-} &
+    \bra{+-}\vec{S}\cdot\vec{I}\ket{-+} &
+    \bra{+-}\vec{S}\cdot\vec{I}\ket{--} &\\
+    \bra{-+}\vec{S}\cdot\vec{I}\ket{++} &
+    \bra{-+}\vec{S}\cdot\vec{I}\ket{+-} &
+    \bra{-+}\vec{S}\cdot\vec{I}\ket{-+} &
+    \bra{-+}\vec{S}\cdot\vec{I}\ket{--} &\\
+    \bra{--}\vec{S}\cdot\vec{I}\ket{++} &
+    \bra{--}\vec{S}\cdot\vec{I}\ket{+-} &
+    \bra{--}\vec{S}\cdot\vec{I}\ket{-+} &
+    \bra{--}\vec{S}\cdot\vec{I}\ket{--} &\\
+  \end{pmatrix}\\
+   = & \frac{A}{\hbar^2}
+  \begin{pmatrix}
+    \frac{\hbar^2}{4}&
+    \bra{++}\frac{1}{2}S_-I_+\ket{+-}&
+    \bra{++}\frac{1}{2}S_+I_-\ket{-+}&
+    0&\\
+    0&
+    \bra{+-}\frac{1}{2}S_-I_+\ket{+-} - \frac{\hbar^2}{4}&
+    \bra{+-}\frac{1}{2}S_+I_-\ket{-+}&
+    0&\\
+    0&
+    \bra{-+}\frac{1}{2}S_-I_+\ket{+-}&
+    \bra{-+}\frac{1}{2}S_+I_-\ket{-+} - \frac{\hbar^2}{4}&
+    0&\\
+    0&
+    \bra{--}\frac{1}{2}S_-I_+\ket{+-}&
+    \bra{--}\frac{1}{2}S_+I_-\ket{-+}&
+    \frac{\hbar^2}{4} &\\
+  \end{pmatrix}\\
+  = & \frac{A}{\hbar^2}
+  \begin{pmatrix}
+    \frac{\hbar^2}{4}&
+    0&
+    0&
+    0&\\
+    0&
+    -\frac{\hbar^2}{4}&
+    \frac{2\hbar^2}{4}&
+    0&\\
+    0&
+    \frac{2\hbar^2}{4}&
+    -\frac{\hbar^2}{4}&
+    0&\\
+    0&
+    0&
+    0&
+    \frac{\hbar^2}{4} &\\
+  \end{pmatrix}\\
+  = & \frac{A}{4}
+  \begin{pmatrix}
+    1&
+    0&
+    0&
+    0&\\
+    0&
+    -1&
+    2&
+    0&\\
+    0&
+    2&
+    -1&
+    0&\\
+    0&
+    0&
+    0&
+    1&\\
+  \end{pmatrix}\\
 \end{align*}
 
+We have succesfully shown how expanding the angular momentum dot product and using ladder
+operators allows each element in our perturbation Hamiltonian to be solved for. Our end
+product is a nondiagonal matrix with first and second-order corrections. Unfortunately our
+original energies are degenerate, as this is the n=2 state. Thus we must diagonalize this
+matrix to get a new diagonal perturbation matrix representing the changes to our original
+system.\\
+
 \end{document}

### 2016-02-24 quantum/latex/a7.tex
all agin actually

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{marvosym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\usepackage{amsmath}
\usepackage{braket}
\usepackage{amssymb}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Quantum Mechanics Assignment Six}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{11.12: blah}

\textbf{d.)}
Here we expand McIntyre (12.26) and show that it is equivalent to (12.30):\\

\begin{align*}
  \braket{H'_{rel}} &=
  -\frac{1}{2mc^2}\left((E_n^0)^2 +
  2\frac{e^2}{4\pi\epsilon_0}E_n^0\braket{\frac{1}{r}} +
  \frac{e^4}{16\pi^2\epsilon_0^2}\braket{\frac{1}{r^2}}\right)\\
  &= -\frac{1}{2mc^2}\left(\frac{\alpha^4m^2c^4}{4n^4}
  + 2\frac{e^2}{4\pi\epsilon_0}\left(\frac{-\alpha^2mc^2}{2n^2}\right)
  \left(\frac{1}{n^2a_0}\right) +
  \frac{e^4}{16\pi^2\epsilon_0^2}\left(\frac{1}{(\ell+1/2)n^3a_0^2}\right)\right)\\
  &= \frac{-\alpha^4mc^2}{8n^4}
  + \frac{e^2}{4\pi\epsilon_0}\left(\frac{\alpha^2}{2n^2}\right)
  \left(\frac{1}{n^2a_0}\right) +
  \frac{-e^4}{2*16\pi^2\epsilon_0^2}\left(\frac{1}{(\ell+1/2)n^3a_0^2mc^2}\right)\\
  &= \frac{-\alpha^4mc^2}{8n^4}
  + a_0\alpha^2mc^2\left(\frac{\alpha^2}{2n^2}\right)
  \left(\frac{1}{n^2a_0}\right) +
  \frac{-1}{2}a_0^2\alpha^4m^2c^4\left(\frac{1}{(\ell+1/2)n^3a_0^2mc^2}\right)\\
  &= \alpha^4mc^2\left(\frac{3}{8n^4}\right) +
  \frac{1}{2}\alpha^4mc^2\left(\frac{-1}{(\ell+1/2)n^3}\right)\\
  &= -\frac{1}{2}\alpha^4mc^2\left(\frac{1}{n^3\left(\ell+\frac{1}{2}\right)} -
    \frac{3}{4n^4}\right)\\
\end{align*}

\section{2: Spectroscopy Numbers}

\section{3: Energies}
Here we calculate many different energies and show how they are related.\\

\textbf{a.) Unperturbed energies}
\begin{align*}
  E_1^0 - E_2^0 &= -13.7eV\left(\frac{1}{4} - \frac{1}{1}\right)\\
  &= 10.275eV\\
  &= \frac{3}{8}\alpha^2mc^2\\
  &= 2.48 * 10^6 GHz\\
\end{align*}

\textbf{b.) SO corrections to energy}
\begin{align*}
  E_{S}' &= \frac{1}{4}\alpha^4mc^2\frac{j(j+1)
    - \ell(\ell+1) - \frac{3}{4}}{n^3\ell(\ell+\frac{1}{2})(\ell+1)}\\
  E_{n=2,\ell=0} &= 0\\
  E_{n=2,\ell=1} &= \frac{-29}{384}\alpha^4mc^2\\
  &= -1.1 *10^{-4} eV\\
  &= 0.026 GHz\\
  E_{n=1,\ell=0} &= 0\\
\end{align*}

\textbf{c.) Relativistic corrections to energy}
\begin{align*}
  E_{rel}' &= -\frac{1}{2}\alpha^4mc^2
  \left(\frac{1}{n^3(\ell+\frac{1}{2})} - \frac{3}{4n^4}\right)\\
  E_{n=2,\ell=1} &= -\frac{7}{384}\alpha^4mc^2\\
  &= -2.6*10^{-5} eV\\
  &= 6.39 GHz\\
  E_{n=2,\ell=0} &= -\frac{13}{128}\alpha^4mc^2\\
  &= 1.4*10^{-4} eV\\
  &= -35.6 GHz\\
  E_{n=1,\ell=0} &= -\frac{5}{8}\alpha^4mc^2\\
  &= 9.1*10^{-4} eV\\
  &= -219.2 GHz\\
\end{align*}

\textbf{d.) Total fine-structure corrections}

\begin{align*}
  E_{n=2,\ell=1} &= -\frac{7}{384}\alpha^4mc^2 = -2.6*10^{-5} eV = -6.39 GHz\\
  E_{n=2,\ell=0} &= -\frac{17}{96}\alpha^4mc^2 = -2.5*10^{-4} = -35.6 GHz\\
  E_{n=1,\ell=0} &= -\frac{5}{8}\alpha^4mc^2 = -9.1*10^{-4} eV = -219.2 GHz\\
\end{align*}

\textbf{e.) }
\begin{align*}
  E_{n=2,\ell=1}: 6.39 GHz \rightarrow \lambda = 4.69*10^7 nm = .0496m\\
  E_{n=2,\ell=0}: 35.6 GHz \rightarrow \lambda = 8.42*10^6 nm = 0.00842m\\
  E_{n=1,\ell=0}: 219.2 GHz \rightarrow \lambda = 1.37*10^6 nm = 0.0137m\\
\end{align*}

\textbf{f.) Rest mass}
The reduced mass must be used! The electrons are travelling at roughly 1\% the speed
of light, so they feel relativistic effects non-negligibly.\\

This was a cool, long problem! It is interesting that the $\ell=0$ states have no
SO corrections, since they have no angular momentum and hence there is no apparent
B-field that they see due to ``orbiting'' the proton. The numbers are weird, it is odd
to me that there are such large differences (nearly a factor of ten) betwen the
corrections between the first three quantum states, but I guess this is because
the orbitals behave pretty differently. The fact that the relativistic differences
have nearly a ten-fold difference in energy might indicate that the electrons have
vastly different speeds, which is an interesting result.\\

\end{document}

### 2016-02-24 quantum/latex/a7.tex
h

@@
 
+This was an interesting problem. It was good to go over how the relativistic fine-
+structure corrections are derived, especially how to deal with a $1/\hat{r}$
+operator. At first I was confused about how to deal with an inverse operator, but
+then I realized that it was just a normal operator, like $\hat{r^2}$. Exploiting
+the fact that the complex conjugate Hamiltonian acts on bras the same way that the
+normal Hamiltonian acts on kets was another important method to solving this problem.\\
+
 \section{2: Spectroscopy Numbers}
+Here we discuss the spectroscopic labels and splitting we see in the figure in the
+problem.\\
+
+The labels are kind of like spectroscopic notation. The upper term (3d, 2p, etc...) is
+the n and $\ell$ number for the state, where $\ell$ is translated into alphabet
+notation. The subscript is the
 

### 2016-02-24 quantum/latex/a7.tex
hh

--- 2016-02-24 quantum/latex/a7.tex
@@
 
-\title{Quantum Mechanics Assignment Six}
+\title{Quantum Mechanics Assignment Five}
 \author{Elliott Capek}
@@
 
-\section{11.12: blah}
+\section{Problem 1: Relativistic Fine Correction}
+
+This problem asks us to work through parts of the derivation of perturbation theory applied to
+the relativistic correction to the Hydrogen energy spectrum.\\
+
+\textbf{a.)}Here we show how to find the expectation value of $p^4$ more easily:\\
+
+\begin{align*}
+  &\bra{n,\ell,m}p^4\ket{n,\ell,m}\\
+  &= \bra{n,\ell,m}p^2p^2\ket{n,\ell,m}\\
+  &= 4m^2\bra{n,\ell,m}\left(H_0 + \frac{e^2}{4\pi\epsilon r}\right)
+  \left(H_0 + \frac{e^2}{4\pi\epsilon r}\right)\ket{n,\ell,m}\\
+  &= 4m^2\left(\bra{n,\ell,m}H_0^2\ket{n,\ell,m}
+  + \bra{n,\ell,m}H_0\frac{e^2}{4\pi\epsilon r}\ket{n,\ell,m}
+  + \bra{n,\ell,m}\frac{e^2}{4\pi\epsilon r}H_0\ket{n,\ell,m}
+  + \bra{n,\ell,m}\frac{e^4}{16\pi^2\epsilon^2r^2}\ket{n,\ell,m}\right)\\
+  &= 4m^2\left((E_n^{(0)})^2
+  + E_n^{(0)}\bra{n,\ell,m}\frac{e^2}{4\pi\epsilon r}\ket{n,\ell,m}
+  + \bra{n,\ell,m}\frac{e^2}{4\pi\epsilon r}\ket{n,\ell,m}E_n^{(0)}
+  + \frac{e^4}{16\pi^2\epsilon^2r^2}\bra{n,\ell,m}\frac{1}{r^2}\ket{n,\ell,m}\right)\\
+  &= 4m^2\left((E_n^{(0)})^2
+  + \frac{e^2E_n^{(0)}}{4\pi\epsilon}\braket{\frac{1}{r}}
+  + \frac{e^2}{4\pi\epsilon}\braket{\frac{1}{r}}E_n^{(0)}
+  + \frac{e^4}{16\pi^2\epsilon^2r^2}\braket{\frac{1}{r^2}}\right)\\
+  &= 4m^2\left((E_n^{(0)})^2
+  + \frac{e^2E_n^{(0)}}{2\pi\epsilon}\braket{\frac{1}{r}}
+  + \frac{e^4}{16\pi^2\epsilon^2r^2}\braket{\frac{1}{r^2}}\right)\\
+\end{align*}
+
+\textbf{b.)}\\
+
+\textbf{c.)} We perform the following integrals using Mathematica:
+
+\begin{align*}
+  \braket{\frac{1}{r}}_{20} &= \left(\frac{Z}{2a}\right)^3\int_0^\infty
+  4r\left(1-\frac{Zr}{2a}\right)^2e^{-Zr/a}dr = \frac{Z}{4a}\\
+  \braket{\frac{1}{r^2}}_{20} &= \left(\frac{Z}{2a}\right)^3\int_0^\infty
+  4\left(1-\frac{Zr}{2a}\right)^2e^{-Zr/a}dr = \frac{Z^2}{4a^2}\\
+  \braket{\frac{1}{r}}_{21} &= \left(\frac{Z}{2a}\right)^3\int_0^\infty
+  \frac{Z^2r}{3a^2}e^{Zr/a}dr = \frac{Z}{4a}\\
+  \braket{\frac{1}{r^2}}_{21} &= \left(\frac{Z^5}{24a^5}\right)^3\int_0^\infty
+  r^2e^{-Zr/a}dr = \frac{Z^2}{12a^2}\\
+\end{align*}
+
+These are the desired results. We then assume this is true for all n, $\ell$ quantum numbers.\\
 
 \textbf{d.)}
-Here we expand McIntyre (12.26) and show that it is equivalent to (12.30):\\
-
-\begin{align*}
-  \braket{H'_{rel}} &=
-  -\frac{1}{2mc^2}\left((E_n^0)^2 +
-  2\frac{e^2}{4\pi\epsilon_0}E_n^0\braket{\frac{1}{r}} +
-  \frac{e^4}{16\pi^2\epsilon_0^2}\braket{\frac{1}{r^2}}\right)\\
-  &= -\frac{1}{2mc^2}\left(\frac{\alpha^4m^2c^4}{4n^4}
-  + 2\frac{e^2}{4\pi\epsilon_0}\left(\frac{-\alpha^2mc^2}{2n^2}\right)
-  \left(\frac{1}{n^2a_0}\right) +
-  \frac{e^4}{16\pi^2\epsilon_0^2}\left(\frac{1}{(\ell+1/2)n^3a_0^2}\right)\right)\\
-  &= \frac{-\alpha^4mc^2}{8n^4}
-  + \frac{e^2}{4\pi\epsilon_0}\left(\frac{\alpha^2}{2n^2}\right)
-  \left(\frac{1}{n^2a_0}\right) +
-  \frac{-e^4}{2*16\pi^2\epsilon_0^2}\left(\frac{1}{(\ell+1/2)n^3a_0^2mc^2}\right)\\
-  &= \frac{-\alpha^4mc^2}{8n^4}
-  + a_0\alpha^2mc^2\left(\frac{\alpha^2}{2n^2}\right)
-  \left(\frac{1}{n^2a_0}\right) +
-  \frac{-1}{2}a_0^2\alpha^4m^2c^4\left(\frac{1}{(\ell+1/2)n^3a_0^2mc^2}\right)\\
-  &= \alpha^4mc^2\left(\frac{3}{8n^4}\right) +
-  \frac{1}{2}\alpha^4mc^2\left(\frac{-1}{(\ell+1/2)n^3}\right)\\
-  &= -\frac{1}{2}\alpha^4mc^2\left(\frac{1}{n^3\left(\ell+\frac{1}{2}\right)} -
-    \frac{3}{4n^4}\right)\\
-\end{align*}
-
-\section{2: Spectroscopy Numbers}
-
-\section{3: Energies}
-Here we calculate many different energies and show how they are related.\\
-
-\textbf{a.) Unperturbed energies}
-\begin{align*}
-  E_1^0 - E_2^0 &= -13.7eV\left(\frac{1}{4} - \frac{1}{1}\right)\\
-  &= 10.275eV\\
-  &= \frac{3}{8}\alpha^2mc^2\\
-  &= 2.48 * 10^6 GHz\\
-\end{align*}
-
-\textbf{b.) SO corrections to energy}
-\begin{align*}
-  E_{S}' &= \frac{1}{4}\alpha^4mc^2\frac{j(j+1)
-    - \ell(\ell+1) - \frac{3}{4}}{n^3\ell(\ell+\frac{1}{2})(\ell+1)}\\
-  E_{n=2,\ell=0} &= 0\\
-  E_{n=2,\ell=1} &= \frac{-29}{384}\alpha^4mc^2\\
-  &= -1.1 *10^{-4} eV\\
-  &= 0.026 GHz\\
-  E_{n=1,\ell=0} &= 0\\
-\end{align*}
-
-\textbf{c.) Relativistic corrections to energy}
-\begin{align*}
-  E_{rel}' &= -\frac{1}{2}\alpha^4mc^2
-  \left(\frac{1}{n^3(\ell+\frac{1}{2})} - \frac{3}{4n^4}\right)\\
-  E_{n=2,\ell=1} &= -\frac{7}{384}\alpha^4mc^2\\
-  &= -2.6*10^{-5} eV\\
-  &= 6.39 GHz\\
-  E_{n=2,\ell=0} &= -\frac{13}{128}\alpha^4mc^2\\
-  &= 1.4*10^{-4} eV\\
-  &= -35.6 GHz\\
-  E_{n=1,\ell=0} &= -\frac{5}{8}\alpha^4mc^2\\
-  &= 9.1*10^{-4} eV\\
-  &= -219.2 GHz\\
-\end{align*}
-
-\textbf{d.) Total fine-structure corrections}
-
-\begin{align*}
-  E_{n=2,\ell=1} &= -\frac{7}{384}\alpha^4mc^2 = -2.6*10^{-5} eV = -6.39 GHz\\
-  E_{n=2,\ell=0} &= -\frac{17}{96}\alpha^4mc^2 = -2.5*10^{-4} = -35.6 GHz\\
-  E_{n=1,\ell=0} &= -\frac{5}{8}\alpha^4mc^2 = -9.1*10^{-4} eV = -219.2 GHz\\
-\end{align*}
-
-\textbf{e.) }
-\begin{align*}
-  E_{n=2,\ell=1}: 6.39 GHz \rightarrow \lambda = 4.69*10^7 nm = .0496m\\
-  E_{n=2,\ell=0}: 35.6 GHz \rightarrow \lambda = 8.42*10^6 nm = 0.00842m\\
-  E_{n=1,\ell=0}: 219.2 GHz \rightarrow \lambda = 1.37*10^6 nm = 0.0137m\\
-\end{align*}
-
-\textbf{f.) Rest mass}
-The reduced mass must be used! The electrons are travelling at roughly 1\% the speed
-of light, so they feel relativistic effects non-negligibly.\\
-
-This was a cool, long problem! It is interesting that the $\ell=0$ states have no
-SO corrections, since they have no angular momentum and hence there is no apparent
-B-field that they see due to ``orbiting'' the proton. The numbers are weird, it is odd
-to me that there are such large differences (nearly a factor of ten) betwen the
-corrections between the first three quantum states, but I guess this is because
-the orbitals behave pretty differently. The fact that the relativistic differences
-have nearly a ten-fold difference in energy might indicate that the electrons have
-vastly different speeds, which is an interesting result.\\
 

### 2016-02-24 quantum/latex/a7.tex

--- 2016-02-24 quantum/latex/a7.tex
@@
 
-\title{Quantum Mechanics Assignment Six}
+\title{Quantum Mechanics Assignment Seven}
 \author{Elliott Capek}
@@
 
-\section{11.12: blah}
+\section{Problem 1: Relativistic Fine Correction}
+
+This problem asks us to work through parts of the derivation of perturbation theory applied to
+the relativistic correction to the Hydrogen energy spectrum.\\
+
+\textbf{a.)}Here we show how to find the expectation value of $p^4$ more easily:\\
+
+\begin{align*}
+  &\bra{n,\ell,m}p^4\ket{n,\ell,m}\\
+  &= \bra{n,\ell,m}p^2p^2\ket{n,\ell,m}\\
+  &= 4m^2\bra{n,\ell,m}\left(H_0 + \frac{e^2}{4\pi\epsilon r}\right)
+  \left(H_0 + \frac{e^2}{4\pi\epsilon r}\right)\ket{n,\ell,m}\\
+  &= 4m^2\left(\bra{n,\ell,m}H_0^2\ket{n,\ell,m}
+  + \bra{n,\ell,m}H_0\frac{e^2}{4\pi\epsilon r}\ket{n,\ell,m}
+  + \bra{n,\ell,m}\frac{e^2}{4\pi\epsilon r}H_0\ket{n,\ell,m}
+  + \bra{n,\ell,m}\frac{e^4}{16\pi^2\epsilon^2r^2}\ket{n,\ell,m}\right)\\
+  &= 4m^2\left((E_n^{(0)})^2
+  + E_n^{(0)}\bra{n,\ell,m}\frac{e^2}{4\pi\epsilon r}\ket{n,\ell,m}
+  + \bra{n,\ell,m}\frac{e^2}{4\pi\epsilon r}\ket{n,\ell,m}E_n^{(0)}
+  + \frac{e^4}{16\pi^2\epsilon^2r^2}\bra{n,\ell,m}\frac{1}{r^2}\ket{n,\ell,m}\right)\\
+  &= 4m^2\left((E_n^{(0)})^2
+  + \frac{e^2E_n^{(0)}}{4\pi\epsilon}\braket{\frac{1}{r}}
+  + \frac{e^2}{4\pi\epsilon}\braket{\frac{1}{r}}E_n^{(0)}
+  + \frac{e^4}{16\pi^2\epsilon^2r^2}\braket{\frac{1}{r^2}}\right)\\
+  &= 4m^2\left((E_n^{(0)})^2
+  + \frac{e^2E_n^{(0)}}{2\pi\epsilon}\braket{\frac{1}{r}}
+  + \frac{e^4}{16\pi^2\epsilon^2r^2}\braket{\frac{1}{r^2}}\right)\\
+\end{align*}
+
+\textbf{b.)}\\
+
+\textbf{c.)} We perform the following integrals using Mathematica:
+
+\begin{align*}
+  \braket{\frac{1}{r}}_{20} &= \left(\frac{Z}{2a}\right)^3\int_0^\infty
+  4r\left(1-\frac{Zr}{2a}\right)^2e^{-Zr/a}dr = \frac{Z}{4a}\\
+  \braket{\frac{1}{r^2}}_{20} &= \left(\frac{Z}{2a}\right)^3\int_0^\infty
+  4\left(1-\frac{Zr}{2a}\right)^2e^{-Zr/a}dr = \frac{Z^2}{4a^2}\\
+  \braket{\frac{1}{r}}_{21} &= \left(\frac{Z}{2a}\right)^3\int_0^\infty
+  \frac{Z^2r}{3a^2}e^{Zr/a}dr = \frac{Z}{4a}\\
+  \braket{\frac{1}{r^2}}_{21} &= \left(\frac{Z^5}{24a^5}\right)^3\int_0^\infty
+  r^2e^{-Zr/a}dr = \frac{Z^2}{12a^2}\\
+\end{align*}
+
+These are the desired results. We then assume this is true for all n, $\ell$ quantum numbers.\\
 
@@
 the n and $\ell$ number for the state, where $\ell$ is translated into alphabet
-notation. The subscript is the
+notation. The subscript is the F-number (L+S) for that orbital. The spin of the electron can be
+$\pm\frac{1}{2}$, so the F-number plus or minus $\frac{1}{2}$ is the $\ell$-number. $3d_{5/2}$ is
+the n=3, $\ell$=2 orbital \textit{where the angular momentum and spin momentum combine}. There is
+another orbital where they are subtracted.\\
+
+The $3p_{3/2}$ and $3d_{3/2}$ orbitals both have corrections of 1.62GHz, as seen in the table. They
+are made up of the following linear combinations of uncoupled states:\\
+
+\begin{align*}
+  \ket{n,F,M_F} &= \ket{n,M_I,M_s}\\
+  \ket{3,\frac{3}{2}, \frac{ 3}{2}} &= \ket{3,1,\frac{1}{2}}\\
+  \ket{3,\frac{3}{2}, \frac{ 1}{2}} &= \frac{1}{\sqrt{3}}\ket{3,1,\frac{1}{2}}
+  + \sqrt{\frac{2}{3}}\ket{2,0,\frac{1}{2}}\\
+  \ket{3,\frac{3}{2}, \frac{-1}{2}} &= \sqrt{\frac{2}{3}}\ket{3,0,-\frac{1}{2}}
+  + \frac{1}{\sqrt{3}}\ket{3,-1,\frac{1}{2}}\\
+  \ket{3,\frac{3}{2}, \frac{-3}{2}} &= \ket{3,-1,-\frac{1}{2}}\\
+\end{align*}
+
+I liked this problem. I didn't get to finish it, but it was interesting thinking about how to use
+Clebsch-Gordan tables to compare bases. It was also helpful thinking about how there are different
+quantum states for when angular momenta combine vs subtract from each other. This doesn't make
+sense but is an important part of quantum mechanics.\\
 

### 2016-02-24 bioinformatics/latex/a3.tex
blarbl

--- 2016-02-03 bioinformatics/latex/a2.tex
@@
 
-\section{Problem One: Scoring Alignments}
-Here we are asked to compute the score for different given alignments.\\
-
-\textbf{A.} 3\\
-\textbf{B.} 1\\
-
-\section{Problem Two: BLAST}
-
-\section{Problem Four: Database size}
-We use the definition of E-value to compute database size, using an E-value of $e^{-27}:\\
+\section{Problem One: Phylogenetic Trees}
+For the following set of nucleotides, we compute the number of substitutions between
+each sequence and plug it into the Jukes-Cantor distance formula to estimate the
+phylogenetic tree for the four species:
 
 \begin{align*}
-  m &= \frac{E2^{S'}}{n} = \frac{e^{-27}2^{121}}{150} = 3.3*10^{22}\\
+  \mbox{Human} \hspace{1cm}ACGTGCTGCTAGCTGACTGATCGATCGTACGTCTAG\\
+  \mbox{Chimp} \hspace{1cm}ACGTCCTGCAAGCTGACAGATCGATCCTACGTCTAG\\
+  \mbox{Mouse} \hspace{1cm}ACGACCAGCTAGGACAGACTTGGATCCTACCTTTAC\\
+  \mbox{Rat}   \hspace{1cm}TCGACCAGCTAGGAAAGACTTGCATCCTACCATTAC\\
 \end{align*}
 
-\section{Problem Seven: GC Content}
-GC content is conserved when taking reverse complement, since Gs and Cs map to
-each other.\\
-
-For original probability $p_{N0}$ and reverse complement probability $P_N'$:\\
-
 \begin{align*}
-  p_{G0} &= p_C'\\
-  p_{C0} &= p_G'\\
-  p_G &= \frac{p_{G0} + p_G'}{2} = \frac{p_{G0} + p_{C0}}{2} = \frac{0.63}{2} = 0.315\\
-  p_C &= 0.315\\
-  p_A &= 0.185\\
-  p_T &= 0.185\\
+  \mbox{Substitutions} \hspace{1cm}
+  \begin{pmatrix}
+    * & 4 & 15 & 18\\
+    & * & 13 & 16\\
+    & & * & 4\\
+    & & & *\\
+  \end{pmatrix}
 \end{align*}
 
-\section{Problem Eight: MEME}
-Here we enter some MEME commands to find motifs:\\
-\textbf{A.} meme proteins.fasta -protein -mod zoops -minw 10\\
-\textbf{B.} meme sequences.fa -rna -mod anr -minw 5 -maxw 10\\
-\textbf{C.} meme cores.fasta -dna -mod oops -nmotifs 5 \# -revcomp not specified,
-only search forward strand\\
-
 \end{document}

### 2016-02-24 bioinformatics/latex/a3.tex
hopefully everything

@@
 
-\title{Bioinformatics Assignment 2}
+\title{Bioinformatics Assignment 3}
 \author{Elliott Capek}
@@
 \begin{align*}
-  \mbox{Substitutions} \hspace{1cm}
+  \mbox{\# of Substitutions (Hamming distance)} \hspace{1cm}
   \begin{pmatrix}
@@
 
+We divide each Hamming distance by the length of the sequence to get the probability per site
+of a mutation. We use this to compute the phylogenetic distances:\\
+
+\begin{align*}
+  \mbox{Pairwise substitution probability} \hspace{1cm}
+  \begin{pmatrix}
+    * & 0.111 & 0.416 & 0.5\\
+    & * & 0.361 & 0.444\\
+    & & * & 0.111\\
+    & & & *\\
+  \end{pmatrix}
+\end{align*}
+
+\begin{align*}
+  \hat{d} = -\frac{3}{4}\log\left(1-\frac{4}{3}p\right) \rightarrow \hspace{1cm}
+  \mbox{Phylogenetic Distance}& \hspace{1cm}
+  \begin{pmatrix}
+    * & 0.12 & 0.60 & 0.82\\
+    & * & 0.49 & 0.67\\
+    & & * & 0.12\\
+    & & & *\\
+  \end{pmatrix}
+\end{align*}\\
+
+From this we can see that:\\
+$\cdot$Humans and chimps are closely related\\
+$\cdot$Mice and rats are closely related\\
+$\cdot$Chimps and mice are slightly related\\
+$\cdot$Humans and rats are very unrelated\\
+
+Unfortunately there is lots of redundant information here, so it is impossible to construct
+an accurate unrooted tree. We know that the chimp/human pair and mouse/rat pair were the last
+to diverge, and probably diverged at similar times, since both pairs have the same distance (0.12).
+However this conflicts with the much closer distance of humans to mice than humans to rats. The best
+we can do is average the Chimp-mouse, chimp-rat, human-mouse and human-rat distances to get 0.645.
+Subtracting off the two pair distances (0.24) we get an inner node length of 0.405:
+
+\begin{verbatim}
+     0.12          0.12
+Chimp ---|  0.405  |---Mouse
+         ----------
+Human ---|         |---Rat
+\end{verbatim}
+
+\section{Problem Two: Histones}
+Here we do multiple-sequence alignments on various h3 protein sequences and analyze the trees made
+by Clustalw and Phyml.
+
+\textbf{a.)}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.45\textwidth]{h3-clustalw.png}
+  \includegraphics[width=0.45\textwidth]{h3-phyml.png}
+  \caption{ClustalW tree on left, Phyml tree on right. 332-Arabidopsis, 156-human, 688-zebrafish,
+  387-mouse, 787-drosophila melanogaster, 327-tobacco.}
+\end{figure}
+
+Phyml Tree:
+\begin{verbatim}
+(((gi|7870716:0.000000,gi|6886177:0.000000):0.000000,(gi|387198|:0.000000,gi|1568559:0.007289)
+:0.007228):0.036831,gi|3273350:0.000000,gi|3326617:0.000000);
+\end{verbatim}
+
+ClustalW Tree:
+\begin{verbatim}
+((gi|332661759|gb|AEE87159.1|:0.00000,gi|3273350|dbj|BAA31218.1|:0.00000):0.03309,
+(gi|1568559|emb|CAB02546.1|:0.00735,gi|387198|gb|AAA37813.1|:0.00000):0.00368,
+(gi|688617769|ref|XP_009296462.:0.00000,gi|78707164|ref|NP_001027387.1:0.00000):0.00368);
+\end{verbatim}
+
+The PhyML tree accurately groups the mammals together, then the other animals, then the plants.
+However, ClustalW does a better job of grouping each specific group together (mammals, non-mammals,
+plants). PhyML doesn't show flies, zebrafish being more related to mammals than each other,
+which seems innaccurate. So in that regard PhyML does a better job. Although PhyML doesn't seem
+to be generating a binary tree, whereas ClustalW does. This could be because PhyML has a minimum
+threshold for grouping things together, whereas ClustalW just takes the most related things and
+groups them without a threshold.\\
+
+\textbf{b.)}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.75\textwidth]{oops-logo.png}\\
+  \includegraphics[width=0.75\textwidth]{anr-logo.png}
+  \caption{OOPS on left, ANR on right.}
+\end{figure}
+
+These two found motifs look almost identical. This suggests that there is only one
+main motif in the sequences, so that one-occurence-per-sequence and any-number-of-occurrences
+both finds the same number: one.\\
+
+\section{Problem Three: More Trees}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.45\textwidth]{clustalw.png}
+  \includegraphics[width=0.45\textwidth]{phyml.png}
+  \caption{ClustalW on left, PhyML on right.}
+\end{figure}
+
+We see similar behavior here, where ClustalW insists on making a binary tree, whereas PhyML does
+not.
+
 \end{document}

### 2016-02-25 bioinformatics/latex/a3.tex
all

@@
   \includegraphics[width=0.45\textwidth]{h3-phyml.png}
-  \caption{ClustalW tree on left, Phyml tree on right. 332-Arabidopsis, 156-human, 688-zebrafish,
+  \caption{ClustalW tree on left for H3, Phyml tree on right. 332-Arabidopsis, 156-human, 688-zebrafish,
   387-mouse, 787-drosophila melanogaster, 327-tobacco.}
@@
   \includegraphics[width=0.45\textwidth]{phyml.png}
-  \caption{ClustalW on left, PhyML on right.}
+  \caption{ClustalW on left for HoxA, PhyML on right.}
 \end{figure}
 
-We see similar behavior here, where ClustalW insists on making a binary tree, whereas PhyML does
-not.
+We see similar behavior here, where ClustalW insists on making a binary tree,
+whereas PhyML does not. The trees look similar and seem to convey the same
+information, which is good. It is interesting that zebrafish have such a
+large distance from the other organisms. I guess it makes sense, since they
+are an early type of fish? Plus the mRNA sequence for the zebrafish was much shorter
+than the other sequences.\\
+
+ClustalW Newick:
+\begin{verbatim}
+((gi|human:0.08731,gi|chimpanzee:-0.00104):0.09698,
+(gi|mouse:0.10978,gi|rat:-0.05153):0.09846,gi|zebrafish:0.58169);
+\end{verbatim}
+
+PhyML Newick:
+\begin{verbatim}
+(((gi|chimpan:0.017786,gi|human:0.039499):0.032820,gi|zebrafi:0.886506)
+:0.075342,gi|rat:0.019008,gi|mouse:0.012854);
+\end{verbatim}
+
+\section{Four: BED, GFF}
+
+\textbf{a.)}
+chr1 72 145 region1 100 + $\rightarrow$ chr1 progname feature 73 145 100 +\\
+chr1 221 342 region2 100 + $\rightarrow$ chr1 progname feature 222 221 100 +\\
+chr1 473 577 region3 100 + $\rightarrow$ chr1 progname feature 474 473 100 +\\
+
+Here progname is the name of the generating program, BED does not specify it.\\
+
+\textbf{b.)}
+chr1 myDB gene 743 1200 . + . ID=gene1 $\rightarrow$ chr1 742 1200 name . +
+chr1 myDB gene 3405 5345 . + . ID=gene2 $\rightarrow$ chr1 3404 5345 name . +
+chr3 myDB mRNA 9143 9778 . + . ID=transcript1 $\rightarrow$ chr1 9142 9778 name . +
+
+\section{Problem Five: Stalling Indices}
+
+Pol II Chip Tissue 2:\\
+We find the median of the gene body to be 1.3. We find the max $\pm$ 100 of the TSS
+to be 1.3 as well.\\
+\begin{align*}
+  \frac{Max}{Median} &= \frac{1.3}{1.3} \approx 1
+\end{align*}
+
+We find the median to be 1.4 and the max to be 5.1.\\
+Pol II Chip Tissue 1:\\
+\begin{align*}
+  \frac{Max}{Median} &= \frac{5.1}{1.4} \approx 3.64
+\end{align*}
 

### 2016-02-29 quantum/latex/a8.tex

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{marvosym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\usepackage{amsmath}
\usepackage{braket}
\usepackage{amssymb}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Quantum Mechanics Assignment Eight}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem 1: Two Spin-1 Particle states}
Here we use a Clebsch-Gordan table and our knowledge of the P interchange operator
to list the spin states a system of two spin-1 particles could take on, and whether
they are symmetric or antisymmetric with respect to particle interchange.\\

Two spin particles combine to have a total angular momentum of 0, 1 or 2. We use a CG
table to list the states, in the format $\ket{J,M_J} = \ket{M_{S1}, M_{S2}}$:\\

\begin{align*}
  \mbox{Symmetric}\hspace{2cm} &\ket{2,2} = \ket{1,1} = \ket{1,1}\\
  \mbox{Symmetric}\hspace{2cm}&\ket{2,1} = \frac{1}{\sqrt{2}}\left(\ket{1,0}
  + \ket{0,1}\right)
  = \frac{1}{\sqrt{2}}\left(\ket{0,1} + \ket{1,0}\right)\\
  \mbox{Symmetric}\hspace{2cm} &\ket{2,0} =
  \frac{1}{\sqrt{2}}\left(\ket{1,-1} + \ket{-1,1}\right) + \sqrt{\frac{2}{3}}\ket{0,0}
  = \frac{1}{\sqrt{2}}\left(\ket{-1,1} + \ket{1,-1}\right)
  + \sqrt{\frac{2}{3}}\ket{0,0}\\
  \mbox{Symmetric}\hspace{2cm} &\ket{2,-1} =
  \frac{1}{\sqrt{2}}\left(\ket{-1,0} + \ket{0,-1}\right)
  = \frac{1}{\sqrt{2}}\left(\ket{0,-1} + \ket{-1,0}\right)\\
  \mbox{Symmetric}\hspace{2cm} &\ket{2,-2} = \ket{-1,-1} = \ket{-1,-1}\\
\end{align*}

\begin{align*}
  \mbox{Antisymmetric}\hspace{2cm} &\ket{1,1}
  = \frac{1}{\sqrt{2}}\left(\ket{1,0} - \ket{0,1}\right)
  = -\frac{1}{\sqrt{2}}\left(\ket{0,1} - \ket{1,0}\right)\\
  \mbox{Antisymmetric}\hspace{2cm} &\ket{1,0}
  = \frac{1}{\sqrt{2}}\left(\ket{1,-1} - \ket{-1,1}\right)
  = -\frac{1}{\sqrt{2}}\left(\ket{-1,1} - \ket{1,-1}\right)\\
  \mbox{Antisymmetric}\hspace{2cm} &\ket{1,-1}
  = \frac{1}{\sqrt{2}}\left(\ket{-1,0} - \ket{0,-1}\right)
  = -\frac{1}{\sqrt{2}}\left(\ket{0,-1} - \ket{-1,0}\right)\\
\end{align*}

\begin{align*}
  \mbox{Symmetric}\hspace{2cm} &\ket{0,0} = \ket{-1,-1} =
  \frac{1}{\sqrt{3}}\left(\ket{1,-1} - \ket{0,0} \ket{-1,1}\right)
  = \frac{1}{\sqrt{3}}\left(\ket{-1,1} - \ket{0,0} \ket{1,-1}\right)\\
\end{align*}

This is a demonstration of how the coupled basis states for 2x spin-1 particles
obey the Symmetrization Postulate. This is an advantage that the uncoupled basis
does not share: $\ket{1,-1}$ does not obey the SP, since the first particle is always
in the $M_S = 1$ state, and so is effectively ``labelled''. This is disallowed
by QM.\\

An interesting result we can glean from this is that two fermions with
spin-1 (eg protons) will always go to the F=1 state when they have a symmetric
wave function (ie they are in the ground state), since all wavefunctions must be
antisymmetric for fermions. This is a weird case that is not intuitive - why is
there such an arbitrary set of rules on what the spin can and cannot be?\\

\end{document}

### 2016-02-29 quantum/latex/a8.tex
a8

@@
 
+\section{Problem 13.7: Noninteracting particles in a HO potential}
+Here we find the interparticle spacing for three types of systems: a distinguishable,
+fermonic and bosonic pair of particles in a QHO potential where one particle is in state $n$
+and the other in state $k$. We construct our states as follows, noting that fermions have AS
+descriptions, bosons have S descriptions and distinguishable particles have neither:
+
+\begin{align*}
+  \mbox{Distinguishable}&\hspace{1cm} \ket{\psi} = \ket{n}_1\ket{k}_2\\
+  \mbox{Fermions}&\hspace{1cm} \ket{\psi}
+  = \frac{1}{\sqrt{2}}\left(\ket{n}_1\ket{k}_2 - \ket{k}_1\ket{n}_2\right)\\
+  \mbox{Bosons}&\hspace{1cm} \ket{\psi}
+  = \frac{1}{\sqrt{2}}\left(\ket{n}_1\ket{k}_2 + \ket{k}_1\ket{n}_2\right)\\
+\end{align*}
+
+\textbf{a.) Distinguishable}
+
+\begin{align*}
+  \braket{(x_1-x_2)^2} &= \bra{n}_1\bra{k}_2(x_1-x_2)^2\ket{n}_1\ket{k}_2\\
+  &= \bra{n}_1\bra{k}_2(x_1^2 + x_2^2 - 2x_1x_2)\ket{n}_1\ket{k}_2\\
+  &= \bra{n}_1\bra{k}_2x_1^2\ket{n}_1\ket{k}_2 +
+  \bra{n}_1\bra{k}_2x_2^2\ket{n}_1\ket{k}_2 - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
+  &= \bra{n}_1x_1^2\ket{n}_1\bra{k}_2\ket{k}_2 +
+  \bra{n}_1\ket{n}_1\bra{k}_2x_2^2\ket{k}_2 - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
+  &= \bra{n}_1x_1^2\ket{n}_1 +
+  \bra{k}_2x_2^2\ket{k}_2 - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
+  now look up the first two terms and turn the third into an integral\\
+\end{align*}
+
 \end{document}

### 2016-02-29 bioinformatics/latex/a3.tex
bla

@@
   \includegraphics[width=0.45\textwidth]{h3-phyml.png}
-  \caption{ClustalW tree on left for H3, Phyml tree on right. 332-Arabidopsis, 156-human, 688-zebrafish,
-  387-mouse, 787-drosophila melanogaster, 327-tobacco.}
+  \caption{ClustalW tree on left for H3, Phyml tree on right. 332-Arabidopsis,
+    156-human, 688-zebrafish,387-mouse, 787-drosophila melanogaster, 327-tobacco.}
 \end{figure}
@@
 
-These two found motifs look almost identical. This suggests that there is only one
+These two found motifs in Figure 2 look almost identical. This suggests that there is only one
 main motif in the sequences, so that one-occurence-per-sequence and any-number-of-occurrences
-both finds the same number: one.\\
+both finds the same number: one. If there were more motifs that met the motifness-threshold,
+they would be found by ANR but not OOPS. The fact that there is only one main motif is plausible:
+there is one active site on the coded protein, the rest of the protein is free to mutate to meet
+the species' needs.\\
 
@@
 
-\textbf{a.)}
+\textbf{a.)}\\
 chr1 72 145 region1 100 + $\rightarrow$ chr1 progname feature 73 145 100 +\\
-chr1 221 342 region2 100 + $\rightarrow$ chr1 progname feature 222 221 100 +\\
-chr1 473 577 region3 100 + $\rightarrow$ chr1 progname feature 474 473 100 +\\
+chr1 221 342 region2 100 + $\rightarrow$ chr1 progname feature 222 342 100 +\\
+chr1 473 577 region3 100 + $\rightarrow$ chr1 progname feature 474 577 100 +\\
 
-Here progname is the name of the generating program, BED does not specify it.\\
+Here progname, feature are the names of the generating program and feature,
+as BED does not specify them.\\
 
-\textbf{b.)}
-chr1 myDB gene 743 1200 . + . ID=gene1 $\rightarrow$ chr1 742 1200 name . +
-chr1 myDB gene 3405 5345 . + . ID=gene2 $\rightarrow$ chr1 3404 5345 name . +
-chr3 myDB mRNA 9143 9778 . + . ID=transcript1 $\rightarrow$ chr1 9142 9778 name . +
+\textbf{b.)}\\
+chr1 myDB gene 743 1200 . + . ID=gene1 $\rightarrow$ chr1 742 1200 name . +\\
+chr1 myDB gene 3405 5345 . + . ID=gene2 $\rightarrow$ chr1 3404 5345 name . +\\
+chr3 myDB mRNA 9143 9778 . + . ID=transcript1 $\rightarrow$ chr1 9142 9778 name . +\\
 
@@
 
-Pol II Chip Tissue 2:\\
-We find the median of the gene body to be 1.3. We find the max $\pm$ 100 of the TSS
-to be 1.3 as well.\\
+\textbf{Pol II Chip Tissue 2:}\\
+We find the median of the gene body {1.1, 1.1, 1.1, 1.5, 1.7} to be 1.1.
+We find the max $\pm$ 100 of the TSS to be 1.3 by inspection.\\
 \begin{align*}
-  \frac{Max}{Median} &= \frac{1.3}{1.3} \approx 1
+  SI = \frac{Max}{Median} &= \frac{1.3}{1.1} \approx 1.18
 \end{align*}
 
-We find the median to be 1.4 and the max to be 5.1.\\
-Pol II Chip Tissue 1:\\
+\textbf{Pol II Chip Tissue 1:}\\
+We find the median of {1.3, 1.3, 1.3, 1.5, 1.5} to be 1.3 and the max within 100pm of
+the TSS to be 5.1 by inspection.\\
 \begin{align*}
-  \frac{Max}{Median} &= \frac{5.1}{1.4} \approx 3.64
+  SI = \frac{Max}{Median} &= \frac{5.1}{1.3} \approx 3.92
 \end{align*}

### 2016-03-01 quantum/latex/a8.tex
work on a8

@@
 
+\section{Problem 2: Neutron Star}
+Neutron stars are celestial objects made up almost exclusively of neutrons, or spin-1/2 fermions.\\
+
 \section{Problem 13.7: Noninteracting particles in a HO potential}
@@
   \bra{n}_1\ket{n}_1\bra{k}_2x_2^2\ket{k}_2 - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
-  &= \bra{n}_1x_1^2\ket{n}_1 +
-  \bra{k}_2x_2^2\ket{k}_2 - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
-  now look up the first two terms and turn the third into an integral\\
+  &= \bra{n}_1x_1^2\ket{n}_1 + \bra{k}_2x_2^2\ket{k}_2
+  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
+  &= \frac{\hbar}{2m\omega}\bra{n}_1
+  \left(a^\dagger a^\dagger + aa + aa^\dagger + a^\dagger a\right)\ket{n}_1
+  + \frac{\hbar}{2m\omega}\bra{k}_2
+  \left(a^\dagger a^\dagger + aa + aa^\dagger + a^\dagger a\right)\ket{k}_2
+  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
+  &= \frac{\hbar}{2m\omega}\bra{n}_1\left(aa^\dagger + a^\dagger a\right)\ket{n}_1
+  + \frac{\hbar}{2m\omega}\bra{k}_2\left(aa^\dagger + a^\dagger a\right)\ket{k}_2
+  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
+  &= \frac{\hbar}{2m\omega}\left(2n + 2k + 2\right)
+  - \frac{2\hbar}{2m\omega}\bra{n}_1\bra{k}_2
+  \left(a_1+a_1^\dagger\right)\left(a_2+a_2^\dagger\right)\ket{n}_1\ket{k}_2\\
+  &= \frac{\hbar}{2m\omega}\left(2n + 2k + 2\right)
+  - \frac{2\hbar}{2m\omega}\bra{n}_1\bra{k}_2
+  \left(a_1a_2 + a_1^\dagger a_2^\dagger + a_1a_2^\dagger + a_2a_1^\dagger\right)\ket{k}_2\\
+  &= \frac{\hbar}{2m\omega}\left(2n + 2k + 2\right)
+  - \frac{2\hbar}{2m\omega}\bra{n}_1\bra{k}_2
+  \left(a_1a_2 + a_1^\dagger a_2^\dagger + a_1a_2^\dagger + a_2a_1^\dagger\right)\bra{n}_1\ket{k}_2\\
+  &= \frac{\hbar}{2m\omega}\left(2n + 2k + 2\right)\\
+  &- \frac{2\hbar}{2m\omega}
+  \left(\bra{n}_1a_1\ket{n}_1\ket{k}_2a_2\ket{k}_2
+  + \bra{n}_1a_1^\dagger\ket{n}_1\bra{k}_2a_2^\dagger\ket{k}_2
+  + \bra{n}_1a_1\ket{n}_1\bra{k}_2a_2^\dagger\ket{k}_2
+  + \bra{k}_2a_2\ket{k}_2\bra{n}_1a_1^\dagger\ket{n}_1\right)\\
+  &= \frac{\hbar}{m\omega}(2n+2k+2)\\
 \end{align*}
 
+\textbf{b.) Boson}
+
+\begin{align*}
+  \braket{(x_1-x_2)^2} &=
+  \frac{1}{2}\left(\bra{n}_1\bra{k}_2 + \bra{k}_1\bra{n}_2\right)
+  \left(x_1^2 + x_2^2 - 2x_1x_2\right)
+  \left(\ket{n}_1\ket{k}_2 + \ket{k}_1\ket{n}_2\right)\\
+  &= \frac{1}{2}\left(
+  \bra{n}_1x_1^2\ket{n}_1 + \bra{k}_2x_2^2\ket{k}_2
+  + \bra{k}_1x_1^2\ket{k}_1 + \bra{n}_2x_2^2\ket{n}_2
+  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2 - 2\bra{n}_2\bra{k}_1x_1x_2\ket{n}_2\ket{k}_1\right)\\
+  &= \frac{1}{2}\left(4n + 4k + 4\right)
+  - \bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2 - \bra{n}_2\bra{k}_1x_1x_2\ket{n}_2\ket{k}_1\\
+  &= \frac{\hbar}{2m\omega}\left(4n + 4k + 4\right)\\
+\end{align*}
+
+\textbf{c.) Fermion}
+
+\begin{align*}
+  \braket{(x_1-x_2)^2} &=
+  \frac{1}{2}\left(\bra{n}_1\bra{k}_2 - \bra{k}_1\bra{n}_2\right)
+  \left(x_1^2 + x_2^2 - 2x_1x_2\right)
+  \left(\ket{n}_1\ket{k}_2 - \ket{k}_1\ket{n}_2\right)\\
+  &= \frac{1}{2}\left(
+  \bra{n}_1x_1^2\ket{n}_1 + \bra{k}_2x_2^2\ket{k}_2
+  - \bra{k}_1x_1^2\ket{k}_1 - \bra{n}_2x_2^2\ket{n}_2
+  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2 + 2\bra{n}_2\bra{k}_1x_1x_2\ket{n}_2\ket{k}_1\right)\\
+  &= \frac{\hbar}{2m\omega}\frac{1}{2}\left(2n+1 + 2k+1 - 2n-1 - 2k-1 - 0\right)\\
+  &= 0\\
+\end{align*}
+
+This is a really interesting problem that demonstrates how systems that are identical save
+for their symmetry rules can behave totally differently. The boson system has twice the expectation
+value of the distinguishable particle system, and the fermion system has zero expectation value
+for particle distance. This is a very unintuitive result, since it predicts that on average the
+two particles are on top of each other.\\
+
+\section{Problem 4: Three-particle system}
+In this problem we practice writing the wave functions for three-particle systems of particles
+with different symmetry rules:\\
+
+\section{Problem 5: Zeeman Effect}
+
 \end{document}

### 2016-03-01 quantum/latex/a8.tex
a8 prob5

@@
 \section{Problem 5: Zeeman Effect}
+When doing Zeeman Effect calculations and deciding whether the Zeeman corrections should go in the
+first-order Hamiltonian correction or be put in the unperturbed Hamiltonian, a ``strong'' B-field
+is one that is too big to be a first-order correction. For the n=2 state of Hydrogen, the
+first-order fine structure corrections have a size of:
+
+\begin{equation*}
+  E_{fs}^{(1)}
+  = -\frac{1}{2}\alpha^4mc^2\frac{1}{2^3}\left(\frac{1}{j+\frac{1}{2}}-\frac{3}{8}\right)\\
+  \approx -\frac{1}{16}\alpha^4mc^2\\
+\end{equation*}
+
+We say that Zeeman corrections of this size and larger should be considered due to ``strong''
+magnetic fields. The energy perturbation due to the B-field is given by:\\
+
+\begin{equation*}
+  \Delta E_{Z} = \frac{\mu_B B}{\hbar}\left(g_\ell L_z + g_e S_z\right)
+  \approx \frac{\mu_B B}{\hbar}\\
+\end{equation*}
+
+We equate these expressions to find the smallest ``strong'' B-field:\\
+
+\begin{align*}
+  B &\approx \frac{-\hbar}{16mu_B}\alpha^4mc^2
+  = \frac{-6.5*10^{-16}}{16*6*10^{-5}}*7^4*10^{-12}*5*10^5\\
+  \approx 8*10^{-26}T\\
+\end{align*}
+
+We now find the B-field which perturbes the H atom enough to equate the energies of the 2p and 3d
+states:\\
+
+\begin{align*}
+  E_{2p} + \Delta E_{Zeeman, 2p} + E_{fs}^{(1)} &= E_{3d} + \Delta E_{Zeeman, 3d} + E_{fs}^{(1)}\\
+  \frac{-13.6eV}{4} + \mu_BB(m_\ell+2m_s)
+  + \frac{1}{2}\alpha^4mc^2\left(\frac{3}{4n^3}
+  -\frac{\ell(\ell+1)-m_\ell m_s}{n^3\ell(\ell+\frac{1}{2})(\ell+1)}\right)
+  &=\\ \frac{-13.6eV}{8} + \mu_BB(m_\ell+2m_s) + \frac{1}{2}\alpha^4mc^2\left(\frac{3}{4n^3}
+  -\frac{\ell(\ell+1)-m_\ell m_s}{n^3\ell(\ell+\frac{1}{2})(\ell+1)}\right)\\
+  \frac{-13.6eV}{4} + 2\mu_BB
+  + \alpha^4mc^2\left(\frac{3}{64}
+  -\frac{2-\frac{1}{2}}{24}\right)
+  &= \frac{-13.6eV}{8} + -3\mu_BB + \frac{1}{2}\alpha^4mc^2\left(\frac{3}{108}
+  -\frac{6+1}{405}\right)\\
+  \frac{13.6eV}{4} + 5\mu_BB
+  &= \alpha^4mc^2
+  \left(\frac{3}{216} - \frac{7}{405} - \frac{3}{64} + \frac{3}{48}\right)\\
+  B &= \frac{\alpha^4mc^2
+    \left(\frac{3}{216} - \frac{7}{405} - \frac{3}{64} + \frac{3}{48}\right) - \frac{13.6eV}{4}}{5\mu_B}\\
+  &= \frac{\frac{317\alpha^4mc^2}{25920} - \frac{13.6eV}{4}}{5\mu_B}\\
+  &\approx \frac{0.68eV}{5.7*10^{-5}eV/T} = 10^4 T\\
+\end{align*}
+
+This is a huge magnetic field!\\
 

### 2016-03-02 quantum/latex/a8.tex
finish hw?

@@
 \section{Problem 2: Neutron Star}
-Neutron stars are celestial objects made up almost exclusively of neutrons, or spin-1/2 fermions.\\
+Neutron stars are celestial objects made up almost exclusively of neutrons, or spin-1/2 fermions.
+They are incredibly massive, and without the Pauli Exclusion Principle would overcome the
+inter-particle repulsions and collapse into a black hole. However, because of the PEP, no two
+particles in a system can take on the same quantum state. I don't know the quantum numbers of
+a neutron in a system like this, but I do know that no two neutrons in the system are allowed to
+have the same spacial wave function, because this would violate the PEP. This requirement for each
+particle to have a different spacial wave function translates to the particles needing to be in
+different spaces, and hence not collapsing into the center and forming a black hole. Thus there
+is a ``quantum force'' which opposes gravity and keeps the particles stable.\\
+
+This is a really weird thing to think about. It is odd that the entire star is treated as a single
+system, meaning each neutron must have different quantum numbers, since this translates to an
+immense number of numbers. It is odd that the PEP manifests itself as a force which opposes gravity.
+\\
+
+Another rival explanation is that, due to the Heisenberg Uncertainty Principle, the position of
+each particle is well known due to tight packing, so the momentum uncertainty must be very large.
+This translates to the particles having a large range of momenta. This momenta could possibly be
+the means by which the particles oppose the gravitational force, since the ``quantum force''
+explanation seems weird to me.\\
 
@@
 In this problem we practice writing the wave functions for three-particle systems of particles
-with different symmetry rules:\\
+with different symmetry rules. Here A, B and C represent spacial states and $S_A, S_B$ and $S_C$ are
+spin states. Particles 1, 2 and 3 are the three particles in the system.\\
+
+\textbf{Distinguishable Particles}
+
+The distinguishable particle case is simple. Because we are allowed to label particles, we say that
+each particle has its own spacial and spin states A, B and C. There is no need for us to make our
+wave function have equal probabilities of measuring 1, 2 and 3 in any state, we are allowed to say
+particle 1 is in state A:\\
+
+\begin{align*}
+  \Psi_D &= \left(\Psi_1^A\Psi_2^B\Psi_3^C\right)\ket{S_A}_1\ket{S_B}_2\ket{S_C}_3\\
+\end{align*}
+
+\textbf{Bosons and Fermions}
+
+The calculation for the boson and fermion states is more difficult. Our wave functions must reflect
+that measuring a particle in state A cannot tell us whether this particle is particle 1, 2 or 3. To
+achieve this we sum all the possible combinations of states to create a completely ambiguous
+superposition state. This is the symmetric state.\\
+
+\begin{align*}
+  \Psi^{S} &=
+  \Psi_1^A\Psi_2^B\Psi_3^C + \Psi_1^A\Psi_2^C\Psi_3^B\\
+  &+ \Psi_1^B\Psi_2^A\Psi_3^C + \Psi_1^B\Psi_2^C\Psi_3^A\\
+  &+ \Psi_1^C\Psi_2^A\Psi_3^B + \Psi_1^C\Psi_2^B\Psi_3^A\\
+  \ket{S}^{S} &=
+  \ket{S_A}_1\ket{S_B}_2\ket{S_C}_3 + \ket{S_A}_1\ket{S_C}_2\ket{S_B}_3\\
+  &+ \ket{S_B}_1\ket{S_A}_2\ket{S_C}_3 + \ket{S_B}_1\ket{S_C}_2\ket{S_A}_3\\
+  &+ \ket{S_C}_1\ket{S_A}_2\ket{S_B}_3 + \ket{S_C}_1\ket{S_B}_2\ket{S_A}_3\\
+\end{align*}
+
+We then use a Slater determinant to construct the antisymmetric spacial and spin states:\\
+
+\begin{align*}
+  \Psi^{A} &= \det
+  \begin{vmatrix}
+    \Psi_1^A & \Psi_2^A & \Psi_3^A\\
+    \Psi_1^B & \Psi_2^B & \Psi_3^B\\
+    \Psi_1^C & \Psi_2^C & \Psi_3^C\\
+  \end{vmatrix}\\
+  &= \Psi_1^A\Psi_2^B\Psi_3^C - \Psi_1^A\Psi_2^C\Psi_3^B\\
+  &- \Psi_1^B\Psi_2^A\Psi_3^C + \Psi_1^C\Psi_2^A\Psi_3^B\\
+  &+ \Psi_1^B\Psi_2^C\Psi_3^A - \Psi_1^C\Psi_2^B\Psi_3^A\\
+  \ket{S}^A &= \det
+  \begin{vmatrix}
+    \ket{S_A}_1^A & \ket{S_A}_2 & \ket{S_A}_3\\
+    \ket{S_B}_1^B & \ket{S_B}_2 & \ket{S_B}_3\\
+    \ket{S_C}_1^C & \ket{S_C}_2 & \ket{S_C}_3\\
+  \end{vmatrix}\\
+  &= \ket{S_A}_1\ket{S_B}_2\ket{S_C}_3 - \ket{S_A}_1\ket{S_C}_2\ket{S_B}_3\\
+  &- \ket{S_B}_1\ket{S_A}_2\ket{S_C}_3 + \ket{S_C}_1\ket{S_A}_2\ket{S_B}_3\\
+  &+ \ket{S_B}_1\ket{S_C}_2\ket{S_A}_3 - \ket{S_C}_1\ket{S_B}_2\ket{S_A}_3\\
+\end{align*}
+
+To create our boson wave functions, we take any combination of individual wave functions that result
+in an overall symmetric wave function:\\
+
+\begin{align*}
+  \Psi_B^{SS} &= \Psi_B^S\ket{S}^S\\
+  &= \left(\Psi_1^A\Psi_2^B\Psi_3^C + \Psi_1^A\Psi_2^C\Psi_3^B
+  + \Psi_1^B\Psi_2^A\Psi_3^C + \Psi_1^B\Psi_2^C\Psi_3^A
+  + \Psi_1^C\Psi_2^A\Psi_3^B + \Psi_1^C\Psi_2^B\Psi_3^A\right)\\
+  &\left(\ket{S_A}_1\ket{S_B}_2\ket{S_C}_3 + \ket{S_A}_1\ket{S_C}_2\ket{S_B}_3
+  + \ket{S_B}_1\ket{S_A}_2\ket{S_C}_3 + \ket{S_B}_1\ket{S_C}_2\ket{S_A}_3
+  + \ket{S_C}_1\ket{S_A}_2\ket{S_B}_3 + \ket{S_C}_1\ket{S_B}_2\ket{S_A}_3\right)
+\end{align*}
+
+\begin{align*}
+  \Psi_B^{AA} &= \Psi_B^A\ket{S}^A\\
+  &= \left(\Psi_1^A\Psi_2^B\Psi_3^C - \Psi_1^A\Psi_2^C\Psi_3^B
+  - \Psi_1^B\Psi_2^A\Psi_3^C + \Psi_1^C\Psi_2^A\Psi_3^B
+  + \Psi_1^B\Psi_2^C\Psi_3^A - \Psi_1^C\Psi_2^B\Psi_3^A\right)\\
+  &\left(\ket{S_A}_1\ket{S_B}_2\ket{S_C}_3 - \ket{S_A}_1\ket{S_C}_2\ket{S_B}_3
+  - \ket{S_B}_1\ket{S_A}_2\ket{S_C}_3 + \ket{S_C}_1\ket{S_A}_2\ket{S_B}_3
+  + \ket{S_B}_1\ket{S_C}_2\ket{S_A}_3 - \ket{S_C}_1\ket{S_B}_2\ket{S_A}_3\right)\\
+\end{align*}
+
+To make our fermion states, we take all the possible antisymmetric overall wavefunctions:\\
+
+\begin{align*}
+  \Psi_F^{AS} &= \Psi^A\ket{S}^S\\
+  &= \left(\Psi_1^A\Psi_2^B\Psi_3^C - \Psi_1^A\Psi_2^C\Psi_3^B
+  - \Psi_1^B\Psi_2^A\Psi_3^C + \Psi_1^C\Psi_2^A\Psi_3^B
+  + \Psi_1^B\Psi_2^C\Psi_3^A - \Psi_1^C\Psi_2^B\Psi_3^A\right)\\
+  &\left(\ket{S_A}_1\ket{S_B}_2\ket{S_C}_3 + \ket{S_A}_1\ket{S_C}_2\ket{S_B}_3
+  + \ket{S_B}_1\ket{S_A}_2\ket{S_C}_3 + \ket{S_B}_1\ket{S_C}_2\ket{S_A}_3
+  + \ket{S_C}_1\ket{S_A}_2\ket{S_B}_3 + \ket{S_C}_1\ket{S_B}_2\ket{S_A}_3\right)\\
+\end{align*}
+
+\begin{align*}
+  \Psi_F^{SA} &= \Psi^S\ket{S}^A\\
+  &= \left(\Psi_1^A\Psi_2^B\Psi_3^C + \Psi_1^A\Psi_2^C\Psi_3^B
+  + \Psi_1^B\Psi_2^A\Psi_3^C + \Psi_1^B\Psi_2^C\Psi_3^A
+  + \Psi_1^C\Psi_2^A\Psi_3^B + \Psi_1^C\Psi_2^B\Psi_3^A\right)\\
+  &\left(\ket{S_A}_1\ket{S_B}_2\ket{S_C}_3 - \ket{S_A}_1\ket{S_C}_2\ket{S_B}_3
+  - \ket{S_B}_1\ket{S_A}_2\ket{S_C}_3 + \ket{S_C}_1\ket{S_A}_2\ket{S_B}_3
+  + \ket{S_B}_1\ket{S_C}_2\ket{S_A}_3 - \ket{S_C}_1\ket{S_B}_2\ket{S_A}_3\right)\\
+\end{align*}
+
+\textbf{Many-electron system}
+To describe a many-electron system, we use the following sets:\\
+
+\begin{align*}
+  A_{spacial} &= [\Psi^A, \Psi^B, ...]\\
+  A_{spin} &= [\ket{S_A}, \ket{S_B}, ...]\\
+\end{align*}
+
+We also use function P(q), which returns all possible combinations of the elements of set q. Our
+many-electron system is then represented by:\\
+
+\begin{align*}
+  \Psi_F^{AS} &= \Psi^A\ket{S}^S\\
+  &= \left(\sum_p^{P(A_{spacial})}\prod_{i=0}^N p(i)_i\right)\det
+  \begin{vmatrix}
+    \ket{S_A}_1 & \ket{S_A}_2 & ... & \ket{S_A}_N\\
+    \ket{S_B}_1 & \ket{S_B}_2 & ... & \ket{S_B}_N\\
+    \vdots & \vdots & \vdots & \vdots\\
+    \ket{S_N}_1 & \ket{S_N}_2 & ... & \ket{S_N}_N\\
+  \end{vmatrix}
+\end{align*}
+
+\begin{align*}
+  \Psi_F^{SA} &= \Psi^S\ket{S}^A\\
+  &= \det
+  \begin{vmatrix}
+    \Psi^A_1 & \Psi^A_2 & ... & \Psi^A_N\\
+    \Psi^B_1 & \Psi^B_2 & ... & \Psi^B_N\\
+    \vdots & \vdots & \vdots & \vdots\\
+    \Psi^N_1 & \Psi^N_2 & ... & \Psi^N_N\\
+  \end{vmatrix}
+  \left(\sum_p^{P(A_{spin})}\prod_{i=0}^N p(i)_i\right)
+\end{align*}
+
+This is a cool problem which made me think about how to generalize wave functions given the
+Symmetrization Principle. Using Slater determinants was really helpful here, I'm not sure I could
+have done the problem without them. My solution for representing many-electron systems is a bit
+impractical, I'm interested in if and how such systems actually are represented.\\
 
@@
 
-This is a huge magnetic field!\\
+This is a huge magnetic field! I'm not entirely sure of it, but perhaps a very strong field is
+necessary to split the orbitals to the level that they agree. It's important to understand when
+the Zeeman effect can be treated as a perturbation and when it has to be actually included in
+the original Hamiltonian. I said that whenever the Zeeman term is of a similar magnitude to the
+fine structure correction, it is too big to include as a perturbation.\\
 

### 2016-03-06 quantum/latex/a9.tex
start final quantum assignment.

--- 2016-03-01 quantum/latex/a8.tex
@@
 
-\title{Quantum Mechanics Assignment Eight}
+\title{Quantum Mechanics Assignment Nine}
 \author{Elliott Capek}
@@
 
-\section{Problem 1: Two Spin-1 Particle states}
-Here we use a Clebsch-Gordan table and our knowledge of the P interchange operator
-to list the spin states a system of two spin-1 particles could take on, and whether
-they are symmetric or antisymmetric with respect to particle interchange.\\
-
-Two spin particles combine to have a total angular momentum of 0, 1 or 2. We use a CG
-table to list the states, in the format $\ket{J,M_J} = \ket{M_{S1}, M_{S2}}$:\\
+\section{Problem One: Gaussian Perturbation of Infinite Square Well}
+Here we deal with an object in an infinite square well subject to a perturbation $V_0x^2e^{-t\tau}$.
+The particle begins at $t=0$ in the ground state $E_1$. We wish to find the probability of a
+transition to $E_2$ as a function of time $t$. To do this, we first calculate the perturbed
+state coefficient $c_2(t)$:
 
 \begin{align*}
-  \mbox{Symmetric}\hspace{2cm} &\ket{2,2} = \ket{1,1} = \ket{1,1}\\
-  \mbox{Symmetric}\hspace{2cm}&\ket{2,1} = \frac{1}{\sqrt{2}}\left(\ket{1,0}
-  + \ket{0,1}\right)
-  = \frac{1}{\sqrt{2}}\left(\ket{0,1} + \ket{1,0}\right)\\
-  \mbox{Symmetric}\hspace{2cm} &\ket{2,0} =
-  \frac{1}{\sqrt{2}}\left(\ket{1,-1} + \ket{-1,1}\right) + \sqrt{\frac{2}{3}}\ket{0,0}
-  = \frac{1}{\sqrt{2}}\left(\ket{-1,1} + \ket{1,-1}\right)
-  + \sqrt{\frac{2}{3}}\ket{0,0}\\
-  \mbox{Symmetric}\hspace{2cm} &\ket{2,-1} =
-  \frac{1}{\sqrt{2}}\left(\ket{-1,0} + \ket{0,-1}\right)
-  = \frac{1}{\sqrt{2}}\left(\ket{0,-1} + \ket{-1,0}\right)\\
-  \mbox{Symmetric}\hspace{2cm} &\ket{2,-2} = \ket{-1,-1} = \ket{-1,-1}\\
+  c_2(t) &= \frac{1}{i\hbar}\int_0^t\bra{2}H'\ket{1}e^{i(E_2-E_1)t'/\hbar}dt'\\
+  &= \frac{2}{i\hbar L}\int_0^t
+  \left(\int_0^L\sin\left(\frac{2\pi x}{L}\right)
+  V_0x^2e^{-t/\tau}\sin\left(\frac{\pi x}{L}\right)dx\right)
+  e^{\frac{i\hbar\pi^2(2^2-1^2)}{2mL^2}t}dt'\\
+  &= \frac{2V_0}{i\hbar L}\int_0^t
+  \left(\int_0^L\sin\left(\frac{2\pi x}{L}\right)
+  x^2\sin\left(\frac{\pi x}{L}\right)dx\right)
+  e^{-t/\tau}e^{\frac{i3\hbar\pi^2}{2mL^2}t}dt'\\
+  &= \frac{2V_0}{i\hbar L}\int_0^t\frac{-8L^3}{\pi^2}e^{-t/\tau}e^{\frac{i3\hbar\pi^2}{2mL^2}t}dt'\\
+  &= \frac{-16V_0L^2}{i\hbar\pi^2}
+  \int_0^te^{\left(\frac{i3\hbar\pi^2}{2mL^2} - \frac{1}{\tau}\right)t}dt'\\
+  &= i\frac{16V_0L^2}{\hbar\pi^2}\left(\frac{i3\hbar\pi^2}{2mL^2} - \frac{1}{\tau}\right)^{-1}
+  e^{\frac{i3\hbar\pi^2}{2mL^2}t}e^{-\frac{t}{\tau}}\\
 \end{align*}
 
+We then find the probability of measurement:
+
 \begin{align*}
-  \mbox{Antisymmetric}\hspace{2cm} &\ket{1,1}
-  = \frac{1}{\sqrt{2}}\left(\ket{1,0} - \ket{0,1}\right)
-  = -\frac{1}{\sqrt{2}}\left(\ket{0,1} - \ket{1,0}\right)\\
-  \mbox{Antisymmetric}\hspace{2cm} &\ket{1,0}
-  = \frac{1}{\sqrt{2}}\left(\ket{1,-1} - \ket{-1,1}\right)
-  = -\frac{1}{\sqrt{2}}\left(\ket{-1,1} - \ket{1,-1}\right)\\
-  \mbox{Antisymmetric}\hspace{2cm} &\ket{1,-1}
-  = \frac{1}{\sqrt{2}}\left(\ket{-1,0} - \ket{0,-1}\right)
-  = -\frac{1}{\sqrt{2}}\left(\ket{0,-1} - \ket{-1,0}\right)\\
+  P_{1\rightarrow2} &= |c_2(t)|^2\\
+  &= \left(\frac{16V_0L^2}{\hbar\pi^2}\right)^2
+  \left(\frac{i3\hbar\pi^2}{2mL^2} - \frac{1}{\tau}\right)^{-2}
+  e^{i\frac{6\hbar\pi^2}{2mL^2}t}e^{-2\frac{t}{\tau}}\\
 \end{align*}
 
+Because
+
 \begin{align*}
-  \mbox{Symmetric}\hspace{2cm} &\ket{0,0} = \ket{-1,-1} =
-  \frac{1}{\sqrt{3}}\left(\ket{1,-1} - \ket{0,0} \ket{-1,1}\right)
-  = \frac{1}{\sqrt{3}}\left(\ket{-1,1} - \ket{0,0} \ket{1,-1}\right)\\
+  \lim_{t\rightarrow\infty} P_{1\rightarrow2}(t) = 0\\
 \end{align*}
 
-This is a demonstration of how the coupled basis states for 2x spin-1 particles
-obey the Symmetrization Postulate. This is an advantage that the uncoupled basis
-does not share: $\ket{1,-1}$ does not obey the SP, since the first particle is always
-in the $M_S = 1$ state, and so is effectively ``labelled''. This is disallowed
-by QM.\\
+we can say that it becomes less and less likely for the system to be measured with energy
+$E_2$ as time goes on.\\
 
-An interesting result we can glean from this is that two fermions with
-spin-1 (eg protons) will always go to the F=1 state when they have a symmetric
-wave function (ie they are in the ground state), since all wavefunctions must be
-antisymmetric for fermions. This is a weird case that is not intuitive - why is
-there such an arbitrary set of rules on what the spin can and cannot be?\\
-
-\section{Problem 2: Neutron Star}
-Neutron stars are celestial objects made up almost exclusively of neutrons, or spin-1/2 fermions.\\
-
-\section{Problem 13.7: Noninteracting particles in a HO potential}
-Here we find the interparticle spacing for three types of systems: a distinguishable,
-fermonic and bosonic pair of particles in a QHO potential where one particle is in state $n$
-and the other in state $k$. We construct our states as follows, noting that fermions have AS
-descriptions, bosons have S descriptions and distinguishable particles have neither:
-
-\begin{align*}
-  \mbox{Distinguishable}&\hspace{1cm} \ket{\psi} = \ket{n}_1\ket{k}_2\\
-  \mbox{Fermions}&\hspace{1cm} \ket{\psi}
-  = \frac{1}{\sqrt{2}}\left(\ket{n}_1\ket{k}_2 - \ket{k}_1\ket{n}_2\right)\\
-  \mbox{Bosons}&\hspace{1cm} \ket{\psi}
-  = \frac{1}{\sqrt{2}}\left(\ket{n}_1\ket{k}_2 + \ket{k}_1\ket{n}_2\right)\\
-\end{align*}
-
-\textbf{a.) Distinguishable}
-
-\begin{align*}
-  \braket{(x_1-x_2)^2} &= \bra{n}_1\bra{k}_2(x_1-x_2)^2\ket{n}_1\ket{k}_2\\
-  &= \bra{n}_1\bra{k}_2(x_1^2 + x_2^2 - 2x_1x_2)\ket{n}_1\ket{k}_2\\
-  &= \bra{n}_1\bra{k}_2x_1^2\ket{n}_1\ket{k}_2 +
-  \bra{n}_1\bra{k}_2x_2^2\ket{n}_1\ket{k}_2 - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
-  &= \bra{n}_1x_1^2\ket{n}_1\bra{k}_2\ket{k}_2 +
-  \bra{n}_1\ket{n}_1\bra{k}_2x_2^2\ket{k}_2 - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
-  &= \bra{n}_1x_1^2\ket{n}_1 + \bra{k}_2x_2^2\ket{k}_2
-  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
-  &= \frac{\hbar}{2m\omega}\bra{n}_1
-  \left(a^\dagger a^\dagger + aa + aa^\dagger + a^\dagger a\right)\ket{n}_1
-  + \frac{\hbar}{2m\omega}\bra{k}_2
-  \left(a^\dagger a^\dagger + aa + aa^\dagger + a^\dagger a\right)\ket{k}_2
-  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
-  &= \frac{\hbar}{2m\omega}\bra{n}_1\left(aa^\dagger + a^\dagger a\right)\ket{n}_1
-  + \frac{\hbar}{2m\omega}\bra{k}_2\left(aa^\dagger + a^\dagger a\right)\ket{k}_2
-  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
-  &= \frac{\hbar}{2m\omega}\left(2n + 2k + 2\right)
-  - \frac{2\hbar}{2m\omega}\bra{n}_1\bra{k}_2
-  \left(a_1+a_1^\dagger\right)\left(a_2+a_2^\dagger\right)\ket{n}_1\ket{k}_2\\
-  &= \frac{\hbar}{2m\omega}\left(2n + 2k + 2\right)
-  - \frac{2\hbar}{2m\omega}\bra{n}_1\bra{k}_2
-  \left(a_1a_2 + a_1^\dagger a_2^\dagger + a_1a_2^\dagger + a_2a_1^\dagger\right)\ket{k}_2\\
-  &= \frac{\hbar}{2m\omega}\left(2n + 2k + 2\right)
-  - \frac{2\hbar}{2m\omega}\bra{n}_1\bra{k}_2
-  \left(a_1a_2 + a_1^\dagger a_2^\dagger + a_1a_2^\dagger + a_2a_1^\dagger\right)\bra{n}_1\ket{k}_2\\
-  &= \frac{\hbar}{2m\omega}\left(2n + 2k + 2\right)\\
-  &- \frac{2\hbar}{2m\omega}
-  \left(\bra{n}_1a_1\ket{n}_1\ket{k}_2a_2\ket{k}_2
-  + \bra{n}_1a_1^\dagger\ket{n}_1\bra{k}_2a_2^\dagger\ket{k}_2
-  + \bra{n}_1a_1\ket{n}_1\bra{k}_2a_2^\dagger\ket{k}_2
-  + \bra{k}_2a_2\ket{k}_2\bra{n}_1a_1^\dagger\ket{n}_1\right)\\
-  &= \frac{\hbar}{m\omega}(2n+2k+2)\\
-\end{align*}
-
-\textbf{b.) Boson}
-
-\begin{align*}
-  \braket{(x_1-x_2)^2} &=
-  \frac{1}{2}\left(\bra{n}_1\bra{k}_2 + \bra{k}_1\bra{n}_2\right)
-  \left(x_1^2 + x_2^2 - 2x_1x_2\right)
-  \left(\ket{n}_1\ket{k}_2 + \ket{k}_1\ket{n}_2\right)\\
-  &= \frac{1}{2}\left(
-  \bra{n}_1x_1^2\ket{n}_1 + \bra{k}_2x_2^2\ket{k}_2
-  + \bra{k}_1x_1^2\ket{k}_1 + \bra{n}_2x_2^2\ket{n}_2
-  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2 - 2\bra{n}_2\bra{k}_1x_1x_2\ket{n}_2\ket{k}_1\right)\\
-  &= \frac{1}{2}\left(4n + 4k + 4\right)
-  - \bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2 - \bra{n}_2\bra{k}_1x_1x_2\ket{n}_2\ket{k}_1\\
-  &= \frac{\hbar}{2m\omega}\left(4n + 4k + 4\right)\\
-\end{align*}
-
-\textbf{c.) Fermion}
-
-\begin{align*}
-  \braket{(x_1-x_2)^2} &=
-  \frac{1}{2}\left(\bra{n}_1\bra{k}_2 - \bra{k}_1\bra{n}_2\right)
-  \left(x_1^2 + x_2^2 - 2x_1x_2\right)
-  \left(\ket{n}_1\ket{k}_2 - \ket{k}_1\ket{n}_2\right)\\
-  &= \frac{1}{2}\left(
-  \bra{n}_1x_1^2\ket{n}_1 + \bra{k}_2x_2^2\ket{k}_2
-  - \bra{k}_1x_1^2\ket{k}_1 - \bra{n}_2x_2^2\ket{n}_2
-  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2 + 2\bra{n}_2\bra{k}_1x_1x_2\ket{n}_2\ket{k}_1\right)\\
-  &= \frac{\hbar}{2m\omega}\frac{1}{2}\left(2n+1 + 2k+1 - 2n-1 - 2k-1 - 0\right)\\
-  &= 0\\
-\end{align*}
-
-This is a really interesting problem that demonstrates how systems that are identical save
-for their symmetry rules can behave totally differently. The boson system has twice the expectation
-value of the distinguishable particle system, and the fermion system has zero expectation value
-for particle distance. This is a very unintuitive result, since it predicts that on average the
-two particles are on top of each other.\\
-
-\section{Problem 4: Three-particle system}
-In this problem we practice writing the wave functions for three-particle systems of particles
-with different symmetry rules:\\
-
-\section{Problem 5: Zeeman Effect}
+This was a cool introduction to applying time-dependent perturbation theory equations to solving
+problems. The answer is intuitive. The perturbation decays exponentially, so at large times it
+should behave exactly like the unperturbed infinite square well and stay in the ground state with
+100\% probability.\\
 

### 2016-03-07 bioinformatics/latex/a4.tex
Add all

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{listings}

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

\allowdisplaybreaks

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Bioinformatics Assignment 4}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem one: RNA duplexes}
We put the 5' and 3' microRNA sequences into a fasta file and use the command
\texttt{RNAduplex} to find their duplex energies. Example command:
\texttt{cat duplex.fa | RNAfold}\\

\textbf{hsa-let-7a: } $\Delta G = -23.50$\\
\textbf{hsa-mir-9: } $\Delta G = -23.00$\\
\textbf{hsa-let-7a: } $\Delta G = -15.50$\\

\section{Problem two: Inter-species RNA duplexes}
We put the 5' human RNA sequences and the 3' Drosophila sequence into a fasta file
and recomputed the duplex energies in the same way as in Problem One:
\textbf{dme-let-7: } $\Delta G = -20.4$\\
\textbf{dme-mir-9a: } $\Delta G = -17.3$\\
\textbf{dme-mir-100: } $\Delta G = -16.5$\\

As we can see, these duplexes are a higher energy than the human-human duplexes, which
is to be expected because DNA sequences drift over time, and divergences from the
original sequence are more likely to increase the duplex energy than decrease it.\\

\section{Problem three: RNA structure commands}
\textbf{a.)} \texttt{cat hsa-mir-100.fa | RNAduplex -T 26}\\
\textbf{b.)} \texttt{ HOW YOU DO DIS??????????????????????}\\

\section{Problem four: }
Here we take a normal RNA sequence and compute its energy by using RNAduplex on it with
its reverse complement. We then take the sequence and concatenate it with its reverse
complement, and successively add ``CA'' non-bonding pairs to the center to form a
loop. We then use RNAfold on each length of the loop to find its energy. We then plot
these energies along with the $1.75*RT*ln(l)$, the theoretical energy, to see how
well the relationship holds.\\

\begin{figure}[h!]
  \caption{}
  \centering
  \includegraphics[width=0.5\textwidth]{../duplex_energy.png}
\end{figure}

From this figure we can see that the trend is mostly logarithmic, except there is a
slight dip in energy at loops of length 6 and 8. This is weird. This could be due to
a geometry thing - perhaps loops of this size are stabilized due to orbitals aligning,
like the plot of cycloalkane stability with length.\\

\section{Problem five: RNA duplex free energy}
Here we take 40-length AU or GC sequences and compute their duplex energy with their
reverse complements. We calculate an energy of -108.28 kcal/mol for 40 C-G alternating
bonds and -41.8 kcal/mol for 40 A-U alternating bonds. We divide by 40 to get -2.7
kcal/mol for a C-G bond (assuming neighbors of C-G bonds, value will be different
for different neighbors) and -1.045 kcal/mol for A-U bonds.\\

\section{Problem six: TopHat command}
Here we use the following command to align reads\_rep1.fastq to the mm10 mouse genome
index, using the mouseGenes.gtf annotation file:\\

\texttt{tophat -o reads\_mm10\_tophat
  -G mouseGenes.gtf mm10 reads\_rep1.fastq}

\section{Problem seven: Cuffdiff}

\end{document}

### 2016-03-09 bioinformatics/latex/a4.tex
Add a4 stuff.

@@
 We put the 5' human RNA sequences and the 3' Drosophila sequence into a fasta file
-and recomputed the duplex energies in the same way as in Problem One:
+and recomputed the duplex energies in the same way as in Problem One:\\
 \textbf{dme-let-7: } $\Delta G = -20.4$\\
@@
 \section{Problem three: RNA structure commands}
-\textbf{a.)} \texttt{cat hsa-mir-100.fa | RNAduplex -T 26}\\
-\textbf{b.)} \texttt{ HOW YOU DO DIS??????????????????????}\\
+\textbf{a.)} \texttt{cat duplex.fasta | RNAduplex -T 26}\\
+\textbf{b.)} \texttt{cat hotair.fasta | RNAfold -T 35}\\
 
@@
 \section{Problem seven: Cuffdiff}
+RPKM stands for \textbf{Reads per Kilobase} of gene per length per \textbf{Million}.
+This is basically taking the number of reads mapped to a gene over the total number
+of reads, which is the ``activity'' of that gene. This number then needs to be corrected
+for the length of the transcript, so we divide by the transcript length.\\
+
+FPKM stands for \textbf{Fragments per Kilobase} of gene per length per \textbf{Million}.
+This is the exact same as RPKM, except fragments of cDNA are used instead of just reads.
+Fragments are better because they take into account things like paired-end effects.\\
+
+\section{Problem eight: Cuffdiff options}
+
+\textbf{a.)} The \texttt{-L/--labels} option can be used to specify the labels used
+in the \texttt{cuffdiff} output. This is a comma-separated lsit if multiple labels
+are to be used.\\
+
+\textbf{b.)} \texttt{cuffdiff} finds expression levels for transcripts, it doesn't create
+a map of transcripts, or transcriptome. \texttt{cufflinks} is the tool to create a
+transcriptome assembly.\\
+
+\section{Problem nine: Intersection vs Union}
+Here we use the tutorial described in Trapnell 2013 to compute the log fold-change.\\
+
+To compute via the union method, we just count the number of hits over the union of
+the exons. We find that there are 25 hits for both Condition A and Condition B, so:
+
+\begin{align*}
+  \log_2\left(\frac{25}{25}\right) &= 0\\
+\end{align*}
+
+To find via the intersection method, we do the same with the intersection of the two
+exons and find there are 19 hits for both Condition A and Condition B, so:
+
+\begin{align*}
+  \log_2\left(\frac{19}{19}\right) &= 0\\
+\end{align*}
+
+Finally to find the true expression, we count the hits for each exon they belong to
+and divide by the exon's length:
+
+\begin{align*}
+  \log_2\left(\frac{\frac{14}{3L}+\frac{11}{2L}}{\frac{15}{3L}+\frac{9}{2L}}\right)
+  &= \log_2\left(\frac{\frac{28+33}{6L}}{\frac{30+27}{6L}}\right)\\
+  &= \log_2\left(\frac{61}{57}\right)\\
+  &= 0.098\\
+\end{align*}
 

### 2016-03-09 quantum/latex/a9.tex
a9

@@
 
+\section{Problem Two: Cubed Perturbation}
+Here we deal with a time-dependent perturbation $H'(x,t) = Ax^3e^{-\gamma t}$
+of a harmonic oscillator. We wish to find the probability of transitioning from the ground
+state after a long time.\\
+
+\textbf{a.)} First we find the probability of transitioning, using energies given by the unperturbed
+harmonic oscillator:
+
+\begin{align*}
+  P_{i\rightarrow f}(t)
+  &= \frac{1}{\hbar^2}\left|\int_0^t\bra{f}H'(t')\ket{0}e^{i(E_f-E_0)t/\hbar}dt'\right|^2\\
+  &= \frac{1}{\hbar^2}\left|\int_0^t
+  \left(\int_{-\infty}^\infty Ax^3e^{-\gamma t}dx\right)e^{if\omega t}dt'\right|^2\\
+  &= \frac{A}{\hbar^2}\left|\int_0^t\bra{f}x^3\ket{0}e^{-\gamma t}e^{if\omega t}dt'\right|^2\\
+  &= \frac{A^2}{\hbar^2}\left|\left(\frac{\hbar}{2m\omega}\right)^{3/2}
+  \int_0^t\bra{f}\left(a^\dagger+a\right)^3\ket{0}e^{-\gamma t}e^{if\omega t}dt'\right|^2\\
+  &= \frac{A^2\hbar}{8m^3\omega^3}
+  \left|\int_0^t\bra{f}
+  \left(a^\dagger a^\dagger a^\dagger + a^\dagger a^\dagger a + a^\dagger aa^\dagger
+  + aa^\dagger a^\dagger + aaa^\dagger + aa^\dagger a + a^\dagger aa + aaa\right)
+  \ket{0}e^{-\gamma t}e^{if\omega t}dt'\right|^2\\
+\end{align*}
+
+From this we can see that the only possible values of $\ket{f}$ are $\ket{1}$ and $\ket{3}$. We
+solve for their transition probabilities:
+
+\begin{align*}
+  P_{0\rightarrow 1}(t) &= \frac{A^2\hbar}{8m^3\omega^3}
+  \left|\int_0^t\bra{1}\left(a^\dagger aa^\dagger + aa^\dagger a^\dagger\right)
+  \ket{0}e^{-\gamma t}e^{if\omega t}dt'\right|^2\\
+  &= \frac{A^2\hbar}{8m^3\omega^3}\left|\int_0^t3e^{-\gamma t}e^{if\omega t}dt'\right|^2\\
+\end{align*}
+
+\begin{align*}
+  P_{0\rightarrow 3}(t) &= \frac{A^2\hbar}{8m^3\omega^3}
+  \left|\int_0^t\bra{3}a^\dagger a^\dagger a^\dagger\ket{0}e^{-\gamma t}e^{if\omega t}dt'\right|^2\\
+  &= \frac{A^2\hbar}{8m^3\omega^3}\left|\int_0^t\sqrt{6}e^{-\gamma t}e^{if\omega t}dt'\right|^2\\
+\end{align*}
+
 \end{document}

### 2016-03-13 bioinformatics/latex/a5.tex
crystal files and final BI hw.

--- 2016-02-24 bioinformatics/latex/a3.tex
@@
 
-\section{Problem One: Phylogenetic Trees}
-For the following set of nucleotides, we compute the number of substitutions between
-each sequence and plug it into the Jukes-Cantor distance formula to estimate the
-phylogenetic tree for the four species:
+\section{Problem One: Sequence Complexity}
+We use the Wooton-Federhen complexity score to quantify the complexity of four nucleotide sequences.
+The formula is this:
 
 \begin{align*}
-  \mbox{Human} \hspace{1cm}ACGTGCTGCTAGCTGACTGATCGATCGTACGTCTAG\\
-  \mbox{Chimp} \hspace{1cm}ACGTCCTGCAAGCTGACAGATCGATCCTACGTCTAG\\
-  \mbox{Mouse} \hspace{1cm}ACGACCAGCTAGGACAGACTTGGATCCTACCTTTAC\\
-  \mbox{Rat}   \hspace{1cm}TCGACCAGCTAGGAAAGACTTGCATCCTACCATTAC\\
+  C_{wf} &= \frac{1}{N}\log_4\left(\frac{N!}{n_A!n_C!n_T!n_G!}\right)\\
 \end{align*}
 
+Complexity Scores:
 \begin{align*}
-  \mbox{\# of Substitutions (Hamming distance)} \hspace{1cm}
-  \begin{pmatrix}
-    * & 4 & 15 & 18\\
-    & * & 13 & 16\\
-    & & * & 4\\
-    & & & *\\
-  \end{pmatrix}
+  &ATATATATATATATATATAT: N=20, n_A=10, n_T=10 \rightarrow C_{wf}
+  = \frac{1}{20}\log_4\left(\frac{20!}{10!10!}\right) = 0.43\\
+  &GCGCGCGCGCGCGCGCGCGC: N=20, n_G=10, n_C=10 \rightarrow C_{wf}
+  = \frac{1}{20}\log_4\left(\frac{20!}{10!10!}\right) = 0.43\\
+  &ACGTACGTACGTACGTACGT: N=20, n_A=5, n_T=5 n_G=5, n_C=5 \rightarrow C_{wf}
+  = \frac{1}{20}\log_4\left(\frac{20!}{5!5!5!5!}\right) = 0.83\\
+  &AAAAAAAAAAAAAAAAAAAA: N=20, n_A=20 \rightarrow C_{wf}
+  = \frac{1}{20}\log_4\left(\frac{20!}{20!}\right) = 0\\
 \end{align*}
 
-We divide each Hamming distance by the length of the sequence to get the probability per site
-of a mutation. We use this to compute the phylogenetic distances:\\
+These results are as we expect. The first two sequences are medium complexity, and the one with
+an even distribution of the four nucleotides is the highest complexity.\\
+
+\section{Problem Two: }
+Here we use the Benjamini-Hochberg procedure for an FDR of 0.01. First, we rank the p-values
+ascendingly:
 
 \begin{align*}
-  \mbox{Pairwise substitution probability} \hspace{1cm}
-  \begin{pmatrix}
-    * & 0.111 & 0.416 & 0.5\\
-    & * & 0.361 & 0.444\\
-    & & * & 0.111\\
-    & & & *\\
-  \end{pmatrix}
+  &p_1 \leq p_7 \leq p_2 \leq p_3 \leq p_4 \leq p_5 \leq p_6 \leq p_{10} \leq p_8 \leq p_9\\
+  &1\hspace{0.65cm}2\hspace{0.65cm}3\hspace{0.65cm}4\hspace{0.65cm}5\hspace{0.65cm}6\hspace{0.65cm}
+  7\hspace{0.65cm}8\hspace{0.65cm}9\hspace{0.65cm}10\\
 \end{align*}
 
+We find the maximum significant rank through:
+
 \begin{align*}
-  \hat{d} = -\frac{3}{4}\log\left(1-\frac{4}{3}p\right) \rightarrow \hspace{1cm}
-  \mbox{Phylogenetic Distance}& \hspace{1cm}
-  \begin{pmatrix}
-    * & 0.12 & 0.60 & 0.82\\
-    & * & 0.49 & 0.67\\
-    & & * & 0.12\\
-    & & & *\\
-  \end{pmatrix}
-\end{align*}\\
+  \frac{p_{r*}N}{r*} \leq FDR\\
+  \frac{p_{r*}}{r*} \leq 0.001\\
+\end{align*}
 
-From this we can see that:\\
-$\cdot$Humans and chimps are closely related\\
-$\cdot$Mice and rats are closely related\\
-$\cdot$Chimps and mice are slightly related\\
-$\cdot$Humans and rats are very unrelated\\
+we find that the $p_1$ and $p_7$ are the only p-values that make this cut, so they are the only
+ones we consider significant.\\
 
-Unfortunately there is lots of redundant information here, so it is impossible to construct
-an accurate unrooted tree. We know that the chimp/human pair and mouse/rat pair were the last
-to diverge, and probably diverged at similar times, since both pairs have the same distance (0.12).
-However this conflicts with the much closer distance of humans to mice than humans to rats. The best
-we can do is average the Chimp-mouse, chimp-rat, human-mouse and human-rat distances to get 0.645.
-Subtracting off the two pair distances (0.24) we get an inner node length of 0.405:
+\section{Problem 3: }
+Do this last
+
+\section{Problem 4: Phylogenetic Tree NN interchange}
+Here we use the NNI procedure to come up with new ways of representing the following tree:
 
 \begin{verbatim}
-     0.12          0.12
-Chimp ---|  0.405  |---Mouse
-         ----------
-Human ---|         |---Rat
+A    B
+I----I
+C    D
 \end{verbatim}
 
-\section{Problem Two: Histones}
-Here we do multiple-sequence alignments on various h3 protein sequences and analyze the trees made
-by Clustalw and Phyml.
+For any four-subtree tree, there are only two ways it can be rearranged. These are the only
+two possible rearrangements:
 
-\textbf{a.)}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=0.45\textwidth]{h3-clustalw.png}
-  \includegraphics[width=0.45\textwidth]{h3-phyml.png}
-  \caption{ClustalW tree on left, Phyml tree on right. 332-Arabidopsis, 156-human, 688-zebrafish,
-  387-mouse, 787-drosophila melanogaster, 327-tobacco.}
-\end{figure}
-
-Phyml Tree:
 \begin{verbatim}
-(((gi|7870716:0.000000,gi|6886177:0.000000):0.000000,(gi|387198|:0.000000,gi|1568559:0.007289)
-:0.007228):0.036831,gi|3273350:0.000000,gi|3326617:0.000000);
+A    C        A    C
+I----I  and   I----I
+B    D        D    B
 \end{verbatim}
 
-ClustalW Tree:
-\begin{verbatim}
-((gi|332661759|gb|AEE87159.1|:0.00000,gi|3273350|dbj|BAA31218.1|:0.00000):0.03309,
-(gi|1568559|emb|CAB02546.1|:0.00735,gi|387198|gb|AAA37813.1|:0.00000):0.00368,
-(gi|688617769|ref|XP_009296462.:0.00000,gi|78707164|ref|NP_001027387.1:0.00000):0.00368);
-\end{verbatim}
+Each tree can be simplified to a pairing between A and one of the other three subtrees, since
+you only need one pair to define a tree of size 4. There are three possible pairs involving A,
+so three possible trees.\\
 
-The PhyML tree accurately groups the mammals together, then the other animals, then the plants.
-However, ClustalW does a better job of grouping each specific group together (mammals, non-mammals,
-plants). PhyML doesn't show flies, zebrafish being more related to mammals than each other,
-which seems innaccurate. So in that regard PhyML does a better job. Although PhyML doesn't seem
-to be generating a binary tree, whereas ClustalW does. This could be because PhyML has a minimum
-threshold for grouping things together, whereas ClustalW just takes the most related things and
-groups them without a threshold.\\
-
-\textbf{b.)}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=0.75\textwidth]{oops-logo.png}\\
-  \includegraphics[width=0.75\textwidth]{anr-logo.png}
-  \caption{OOPS on left, ANR on right.}
-\end{figure}
-
-These two found motifs look almost identical. This suggests that there is only one
-main motif in the sequences, so that one-occurence-per-sequence and any-number-of-occurrences
-both finds the same number: one.\\
-
-\section{Problem Three: More Trees}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=0.45\textwidth]{clustalw.png}
-  \includegraphics[width=0.45\textwidth]{phyml.png}
-  \caption{ClustalW on left, PhyML on right.}
-\end{figure}
-
-We see similar behavior here, where ClustalW insists on making a binary tree, whereas PhyML does
-not.
+\section{Problem Six: ChipSeq}
 

### 2016-03-14 bioinformatics/latex/a5.tex
finish bi a5.pdf

@@
 
+\begin{align*}
+  \frac{0.001 * 10}{1} \leq \frac{0.001 * 10}{2} \leq \frac{0.005 * 10}{3} \leq
+  \frac{0.01 * 10}{4} \leq \frac{0.01 * 10}{5} \leq \frac{0.05 * 10}{6} \leq
+  \frac{0.02 * 10}{7} \le \frac{0.03 * 10}{8} \leq \frac{0.2 * 10}{9} \leq
+  \frac{0.5 * 10}{10}\\
+\end{align*}
+
+\begin{align*}
+  0.01 - 0.005 - \frac{0.0}5{3} \leq
+  \frac{0.01 * 10}{4} \leq \frac{0.01 * 10}{5} \leq \frac{0.05 * 10}{6} \leq
+  \frac{0.02 * 10}{7} \le \frac{0.03 * 10}{8} \leq \frac{0.2 * 10}{9} \leq
+  \frac{0.5 * 10}{10}\\
+\end{align*}
+
 we find that the $p_1$ and $p_7$ are the only p-values that make this cut, so they are the only
@@
 \section{Problem 3: }
-Do this last
+We download the sonic hedgehog gene and run \texttt{hmmscan} to find the location of likely domains.
+We use Pfam database human domains from Lab 9 to do this. When we run \texttt{hmmscan} we get
+the following (trimmed) results:
+
+\begin{verbatim}
+#                                      -------------- this domain ---
+# target name        accession   tlen  i-Evalue  score  bias   description of target
+#------------------- ---------- -----  -------- ------ ----- -----------------------
+HH_signal            PF01085.15   161   2.7e-85  283.8   0.0   Hedgehog amino-terminal
+Hint                 PF01079.17   214   7.8e-72  241.1   0.0   Hint module
+Peptidase_M15_3      PF08291.8    111     0.001   19.0   0.0   Peptidase M15
+Hint_2               PF13403.3    143         3    8.0   0.0   Hint domain
+\end{verbatim}
+
+Here we see the top two domains, the ``Hedgehog amino-terminal'' and ``Hint module'' have
+pretty small E-values compared to the other two. The tlens are similar for all of these
+domains, so we can compare their E-values and say the top two domains are significant and
+the bottom two aren't.\\
 
@@
 
-\section{Problem Six: ChipSeq}
+\section{Problem Five: jnet domains}
+Here we use the \texttt{jnet} program to find the predicted secondary structure of human
+beta-defensin-1 sequence. Here are the results:\\
 
+\begin{verbatim}
+ RES	: MRTSYLLLFTLCLLLSEMASGGNFLTGLGHRSDHYNCVSSGGQCLYSACPIFTKIQGTCYRGKAK
+ ALIGN	: ---HHHHHHHHHHHHHHHH----EEE---------EEE-----EE-E----EEEE----------
+ CONF	: 83146888899999876405884010246668753030587632101255231211200023322
+ FINAL	: ---HHHHHHHHHHHHHHHH----EEE---------EEE-----EEEE----EEEE----------
+\end{verbatim}
+
+Here we see there is likely a large alpha-helix (H) near the N-terminus and 4
+beta-sheets evenly spaced throughout.\\
+
+This compares pretty well to the RCSB structure prediction, which predicts an
+alpha-helix near the N-term and three alpha-helices evenly spaced through the
+protein.\\
+
+\section{Problem Six: ChIPSeq}
+To find the p-value of the ChIPSeq peak we are given, we use the following formula:\\
+
+\begin{align*}
+  \mbox{p-val} &= 1 - F(N_{ChIP}-1,N_{ChIP}+N_{cont}, 0.5)\\
+  \mbox{p-val} &= 1 - F(46, 65, 0.5)\\
+  &= 0.000211 = 2.1*10^{-4}\\
+\end{align*}
+
+This is a fairly small p-value, which means this is likely a legitimate peak.\\
 \end{document}

### 2016-04-05 solid_state/latex/a1.tex
Add SS stuff.

--- 2016-01-12 quantum/latex/a1.tex
@@
 
-\title{Quantum Mechanics Assignment One}
+\title{Solid State Assignment One}
 \author{Elliott Capek}
@@
 
-\section{Problem 2.17}
-This problem asks us to take the following spin-1 state and figure out the probabilities of
-measuring z- and x-component spins, and also to find their expectation value.\\
+\section{Problem 1: D functions as linear combinations}
+For each of the five d-orbitals wavefunctions, create a 3D plot of the angular wave and express
+the angular wave as a linear combination of spherical harmonics, like so:
 
 \begin{align*}
-  \ket{\Psi} \doteq &\frac{1}{\sqrt{30}}
-  \begin{pmatrix}
-    1\\
-    2\\
-    5i\\
-  \end{pmatrix}
+  d_{x^2-y^2} &= a_{00}Y_{00} + a_{1-1}Y_{1-1} + a_{10}Y_{10} + a_{11}Y_{11} + a_{20}Y_{20} + ...\\
 \end{align*}
 
-We use the normal $P_x = \abs{\braket{\Psi_x}{\Psi}}^2$ to find probabilities. It is simple to
-calculate the braket since both states are in the $S_z$ representation.
-
-\begin{align*}
-  P_{1} = \abs{\braket{\Psi_{1}}{\Psi}}^2 &=
-  \abs{\frac{1}{\sqrt{30}}\begin{pmatrix}1&0&0\end{pmatrix}\begin{pmatrix}1\\2\\5i\\\end{pmatrix}}^2
-  = \frac{1}{30}\\
-  P_{0} = \abs{\braket{\Psi_{0}}{\Psi}}^2 &=
-  \abs{\frac{1}{\sqrt{30}}\begin{pmatrix}0&1&0\end{pmatrix}\begin{pmatrix}1\\2\\5i\\\end{pmatrix}}^2
-  = \frac{4}{30}\\
-  P_{-1} = \abs{\braket{\Psi_{-1}}{\Psi}}^2 &=
-  \abs{\frac{1}{\sqrt{30}}\begin{pmatrix}0&0&1\end{pmatrix}\begin{pmatrix}1\\2\\5i\\\end{pmatrix}}^2
-  = \frac{25}{30}\\
-\end{align*}
-\begin{align*}
-  \expval{S_z}{\Psi} &= \frac{1}{30}\begin{pmatrix}1&2&-5i\end{pmatrix}\begin{pmatrix}
-      \hslash&0&0\\0&0&0\\0&0&-\hslash\\\end{pmatrix}\begin{pmatrix}1\\2\\5i\end{pmatrix}
-        = \frac{1}{30}\begin{pmatrix}1&2&-5i\end{pmatrix}
-          \begin{pmatrix}\hslash\\0\\5\hslash i\end{pmatrix}
-            = \frac{1}{30}\left(1 + -25\right) = \frac{-4}{5}\\
-\end{align*}
-
-We then want to compute the expectation value of the spin in the x-direction. We use the
-$S_z$ representation of $S_x$:
-
-\begin{align*}
-  \expval{S_x}{\Psi} &= \frac{1}{30}\begin{pmatrix}1&2&-5i\end{pmatrix}\frac{\hslash}{\sqrt{2}}
-    \begin{pmatrix}0&1&0\\1&0&1\\0&1&0\\\end{pmatrix}\begin{pmatrix}1\\2\\5i\end{pmatrix}
-        = \frac{1}{30\sqrt{2}}\begin{pmatrix}1&2&-5i\end{pmatrix}
-          \begin{pmatrix}2\\1+5i\\2\end{pmatrix}\\
-            = \frac{4}{30\sqrt{2}}\\
-\end{align*}
-
-\section{Problem 5.11}
-This problem deals with an infinite square well wave which suddenly has its right potential wall
-move from $L$ to $3L$. We are asked to find the probability of finding the wave in the ground or
-first excited state. This problem is basically asking us to find the $c_1$ and $c_2$ coefficients
-of the old wave function in the new well, and from them find probabilities.
-
-\begin{align*}
-  \Psi(x) &= \sqrt{\frac{2}{L}}\sin\left(\frac{\pi x}{L}\right)\\
-  \ket{E_1} &= \sqrt{\frac{2}{3L}}\sin\left(\frac{\pi x}{3L}\right)\\
-  \ket{E_2} &= \sqrt{\frac{2}{3L}}\sin\left(\frac{2\pi x}{3L}\right)\\
-\end{align*}
-
-We wish to find the $c_1$ and $c_2$ coefficients for this system.
-
-\begin{align*}
-  c_1 &= \braket{E_1}{\Psi} = \int_0^{3L} \sqrt{\frac{2}{3L}}\sin\left(\frac{\pi x}{3L}\right)*
-  \sqrt{\frac{2}{L}}\sin\left(\frac{\pi x}{L}\right)dx\\
-  &= \sqrt{\frac{4}{3L^3}}\int_0^{3L}
-  \sin\left(\frac{\pi x}{3L}\right)\sin\left(\frac{\pi x}{L}\right)\\
-  &= \sqrt{\frac{2}{3L^3}}\left(\int_0^{3L} \cos\left(\frac{2\pi x}{3L}\right)dx
-  - \int_0^{3L} \cos\left(\frac{4\pi x}{3L}\right)dx\right)\\
-  &= 0\\
-\end{align*}
-
-\begin{align*}
-  c_2 &= \braket{E_2}{\Psi} = \int_0^{3L} \sqrt{\frac{2}{3L}}\sin\left(\frac{2\pi x}{3L}\right)*
-  \sqrt{\frac{2}{L}}\sin\left(\frac{\pi x}{L}\right)dx\\
-  &= \sqrt{\frac{4}{3L^3}}\int_0^{3L}
-  \sin\left(\frac{2\pi x}{3L}\right)\sin\left(\frac{\pi x}{L}\right)\\
-  &= \sqrt{\frac{2}{3L^3}}\left(\int_0^{3L} \cos\left(\frac{\pi x}{3L}\right)dx
-  - \int_0^{3L} \cos\left(\frac{5\pi x}{3L}\right)dx\right)\\
-  &= 0\\
-\end{align*}
-
-From this we can see that the probability of measuring the wave in the new tripled well to be in
-$\ket{E_1}$ or $\ket{E_2}$ is zero, since $P_n = \abs{c_n}^2$. This result is to be expected, since
-the wave should be completely in the second excited state, $\ket{E_3}$. This is because extending
-the $\Psi(x)$ wave function from $0$ to $3L$ results in a two-node wave function identical to the
-$\ket{E_3}$ state.\\
-
-\section{8.7}
+\subsection{d
 

### 2016-04-05 solid_state/latex/a4.tex
Add SS stuff.

--- 2016-02-02 quantum/latex/a4.tex
@@
 \begin{align*}
-  H' &= \vec{\mu} \cdot \vec{B_2}\\
-  &= \frac{g e}{2m} \hat{\mu} \cdot B_2 \hat{x}\\
-  &= \frac{\hbar\omega_2}{2} S_x
+  H' &= -\vec{\mu} \cdot \vec{B_2}\\
+  &= -\frac{g e}{2m} \hat{\mu} \cdot B_2 \hat{x}\\
+  &= -\frac{\hbar\omega_2}{2} S_x
   \hspace{2cm}\omega_2 = \frac{geB}{2m}\\
   &=
-  \begin{pmatrix}
-    0 & \frac{\hbar\omega_2}{2} & 0\\
-    \frac{\hbar\omega_2}{2} & 0 & \frac{\hbar\omega_2}{2}\\
-    0 & \frac{\hbar\omega_2}{2} & 0\\
+  -\begin{pmatrix}
+    0 & \frac{\hbar\omega_2}{\sqrt{2}} & 0\\
+    \frac{\hbar\omega_2}{\sqrt{2}} & 0 & \frac{\hbar\omega_2}{\sqrt{2}}\\
+    0 & \frac{\hbar\omega_2}{\sqrt{2}} & 0\\
   \end{pmatrix}
@@
   &= \frac{\hbar\omega_2^2}{4\omega_0}\\
+  &= \frac{geB_2^2\hbar}{4mB_0}
 \end{align*}
@@
   &= -\frac{\hbar\omega_2^2}{4\omega_0}\\
+  &= \frac{-geB_2^2\hbar}{4mB_0}
 \end{align*}
@@
 
+\textbf{d.}
+Here we must use wave function notation since the perturbed Hamiltonian is in the
+spacial representation:
+
+\begin{align*}
+  E_1^{(1)} &= \bra{1^0}H'\ket{1^0}\\
+  &= \frac{2V_0}{L\epsilon}
+  \int_{\frac{L}{2}-\frac{\epsilon L}{2}}^{\frac{L}{2}+\frac{\epsilon L}{2}}
+  \sin^2\left(\frac{\pi x}{L}\right)dx\\
+  &= V_0 + \frac{2V_0\sin(\pi\epsilon)}{\pi\epsilon}
+  \hspace{2cm}\mbox{Wolfram Alpha}\\
+\end{align*}
+
+\textbf{e.}
+The limit of our energy correction as $\epsilon\rightarrow\infty$ is just $2V_0$,
+as expected, since this case is identical to the delta function case.\\
+
 \section{10.24: Degenerate Perturbation}
@@
   E_3^{(2)} &=
-  \frac{|H'_{3,1}|^2}{E_1 - E_3} +
-  \frac{|H'_{3,2}|^2}{E_2 - E_3} +
-  \frac{|H'_{3,4}|^2}{E_4 - E_3}\\
-  &= \frac{4\epsilon^2}{-2V_0} + \frac{\epsilon^2}{-4V_0}\\
-  &= \frac{-9\epsilon^2}{4V_0}
+  \frac{|H'_{3,1}|^2}{E_3 - E_1} +
+  \frac{|H'_{3,2}|^2}{E_3 - E_2} +
+  \frac{|H'_{3,4}|^2}{E_3 - E_4}\\
+  &= \frac{4\epsilon^2}{2V_0} - \frac{\epsilon^2}{2V_0}\\
+  &= \frac{3\epsilon^2}{2V_0}
 \end{align*}
@@
   E_4^{(2)} &=
-  \frac{|H'_{4,1}|^2}{E_1 - E_4} +
-  \frac{|H'_{4,2}|^2}{E_2 - E_4} +
-  \frac{|H'_{4,3}|^2}{E_3 - E_4}\\
-  &= \frac{-\epsilon^2}{2V_0}
+  \frac{|H'_{4,1}|^2}{E_4 - E_1} +
+  \frac{|H'_{4,2}|^2}{E_4 - E_2} +
+  \frac{|H'_{4,3}|^2}{E_4 - E_3}\\
+  &= \frac{\epsilon^2}{2V_0}
 \end{align*}
@@
 
+This means that the final corrected energies are:
+
+\begin{align*}
+  E_1 &= E_2 = 3 \pm \epsilon\\
+\end{align*}
+
 \end{document}

### 2016-04-05 solid_state/latex/a9.tex
Add SS stuff.

--- 2016-03-09 quantum/latex/a9.tex
@@
 
+\section{Problem three: p-state transitions}
+Here we use McIntyre (14.84) to find which quantum states can be reached from the p-state.
+From 14.84 we grab the following statement:\\
+
+\begin{align*}
+  \sqrt{\frac{3(2\ell_i+1)}{4\pi(2\ell_f+1)}}\bra{\ell_i,1,m_i,m}\ket{\ell_f,m_f}
+  \bra{\ell_i,1,0,0}\ket{\ell_f,0}\\
+\end{align*}
+
+We then plug in values for $\ell_i$ and find the CB coefficients that lead to nonzero
+values for this expression.\\
+
+By looking at the CB table, we can see that for the $m_1=0,m_2=0$ RHS braket, the only
+possible coupled basis J values are 0,2. This means that $\ell_f=0,2$, so by this we
+can see that $\ell_i\neq\ell_f$. This is the only selection rule imposed by the RHS
+braket, so we ignore it and look instead at the left braket:\\
+
+\begin{align*}
+  \bra{1,1,1,1}\ket{\ell_f,m_f}  \rightarrow\bra{\ell_f,m_f}=\braket{2,2}\\
+  \bra{1,1,1,0}\ket{\ell_f,m_f}  \rightarrow\bra{\ell_f,m_f}=\braket{2,1}\\
+  \bra{1,1,1,-1}\ket{\ell_f,m_f} \rightarrow\bra{\ell_f,m_f}=\braket{2,0},\braket{0,0}\\
+  \bra{1,1,0,1}\ket{\ell_f,m_f}  \rightarrow\bra{\ell_f,m_f}=\braket{2,1}\\
+  \bra{1,1,0,0}\ket{\ell_f,m_f}  \rightarrow\bra{\ell_f,m_f}=\braket{2,0},\braket{0,0}\\
+  \bra{1,1,0,-1}\ket{\ell_f,m_f} \rightarrow\bra{\ell_f,m_f}=\braket{2,-1}\\
+  \bra{1,1,-1,1}\ket{\ell_f,m_f} \rightarrow\bra{\ell_f,m_f}=\braket{2,0},\braket{0,0}\\
+  \bra{1,1,-1,0}\ket{\ell_f,m_f} \rightarrow\bra{\ell_f,m_f}=\braket{2,-1}\\
+  \bra{1,1,-1,-1}\ket{\ell_f,m_f}\rightarrow\bra{\ell_f,m_f}=\braket{2,-2}\\
+\end{align*}
+
+From this we can see that the allowed transitions from the p-state are either to the s state
+or any of the d-state orbitals.\\
+
+This was a weird problem. It is odd that you can treat the final electron state as
+represented in the coupled basis and the original state in the uncoupled basis and get
+meaningful results. This is really cool, but an odd usage of the CB table.\\
+
 \end{document}

### 2016-04-05 bioinformatics/latex/final.tex
last minute adds.

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{listings}

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

\allowdisplaybreaks

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Bioinformatics Final}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem one: Explaining difference}
\textbf{Smith-Waterman vs Needleman-Wunsch}
Needleman-Wunsch is  global alignment algorithm that allows negative
values in its dynamic programming table and always has path traceback
begin at the bottom-right square of the table.\\

Smith-Waterman is a local alignment algorithm that doesn't allow negative
values in its dynamic programming tbale and allows the path traceback to
begin at the highest-scoring square, not necessarily the bottom-right one.\\

\textbf{Paralogs vs Orthologs}
Paralogs: Two genes which come from the same ancestral gene via gene duplication.
Usually in the same species.\\

Orthologs: Two genes which come from the same ancestral gene via speciation.
Therefore in different species.\\

\textbf{RefSeq vs GenBank}
GenBank is a very large database with redundant entries that seeks to have
the most entries possible. It is not curated, and has tons of information. Contains
most of the world's known DNA and RNA sequences.\\

RefSeq is also a large (but less large) database which is curated. It doesn't have
redundant entries, and the entries have a threshold they must meet. Entries for DNA
and protein link together nicely.\\

\textbf{E vs P values}
These are in reference to the BLAST program.\\

The P-value is the probability of a BLAST alignment randomly having a score equal to
or greater than the given score. The lower the better.\\

The E-value is the expected number of hits at the current score due to chance. It is
related to the P-value via an exponential.\\

\section{Problem two: Alignment Scores}
\textbf{A.}
Number of matching aligned BP: 21\\
Number of nonmatching aligned BP: 2\\
Number of gaps: 6\\

Score: 21 - 2 - 6 = 13\\

\textbf{B.}
Number of matching aligned BP: 12\\
Number of nonmatching aligned BP: 4\\
Number of gaps: 2\\

Score: 12 - 2*4 - 2*2 = 0\\

\section{Problem three: BLOSUM matrix}
Here we calculate BLOSUM matrix elements for an alignment with
\textbf{4Cs, 5Vs, 2Ms}.\\

\begin{align*}
  f_{CC} &= \begin{pmatrix} n_C\\ 2\\ \end{pmatrix} = 6\\
  f_{VV} &= \begin{pmatrix} n_V\\ 2\\ \end{pmatrix} = 10\\
  f_{MM} &= \begin{pmatrix} n_M\\ 2\\ \end{pmatrix} = 1\\
  f_{CV} &= n_C \times n_V = 20\\
  f_{CM} &= n_C \times n_M = 8\\
  f_{VM} &= n_V \times n_M = 10\\
  q_{CC} &= \frac{f_{CC}}{f_{CC} + f_{VV} + f_{MM} + f_{CV} + f_{CM} + f_{VM}}
  = \frac{6}{55} = 0.1\\
  q_{VV} &= \frac{f_{VV}}{f_{CC} + f_{VV} + f_{MM} + f_{CV} + f_{CM} + f_{VM}}
  = \frac{10}{55} = 0.18\\
  q_{MM} &= \frac{f_{MM}}{f_{CC} + f_{VV} + f_{MM} + f_{CV} + f_{CM} + f_{VM}}
  = \frac{1}{55} = 0.018\\
  q_{CV} &= \frac{f_{CV}}{f_{CC} + f_{VV} + f_{MM} + f_{CV} + f_{CM} + f_{VM}}
  = \frac{20}{55} = 0.36\\
  q_{CM} &= \frac{f_{CM}}{f_{CC} + f_{VV} + f_{MM} + f_{CV} + f_{CM} + f_{VM}}
  = \frac{8}{55} = 0.145\\
  q_{VM} &= \frac{f_{VM}}{f_{CC} + f_{VV} + f_{MM} + f_{CV} + f_{CM} + f_{VM}}
  = \frac{10}{55} = 0.18\\
  p_C &= q_{CC} + \frac{1}{2}\left(q_{CV} + q_{CM}\right) = 0.35\\
  p_V &= q_{VV} + \frac{1}{2}\left(q_{VC} + q_{VM}\right) = 0.37\\
  p_M &= q_{MM} + \frac{1}{2}\left(q_{MC} + q_{MV}\right) = 0.18\\
\end{align*}

I might have made a mistake in my calculations here since the final probabilities
don't add up to one. My theory should be good though. I'll come back if I have time.\\

\section{Problem four: BLAST stats}
\textbf{A.)}
n = 200\\
m = $10^9$\\
E = 0.1\\

\begin{align*}
  S' &= -\log_2\left(\frac{E}{nm}\right)\\
  &= -\log_2\left(\frac{0.1}{200*10^9}\right)\\
  &= 40.8\\
\end{align*}

\textbf{B.)}
\begin{align*}
  P &= 1 - e^{-E}\\
  &= 1 - e^{-0.1}\\
  &= 0.095\\
\end{align*}

\section{Problem five: Duplex energy}
We put the 5’ and 3’ microRNA sequences into a fasta file and use the
command RNAduplex to find their duplex energies. Example command:
cat duplex.fa | RNAduplex\\

\begin{align*}
  \mbox{hsa-mir-17}&: -22.7\\
  \mbox{hsa-mir-17}&: -18.5\\
  \mbox{hsa-mir-17}&: -21.9\\
\end{align*}

\section{Problem six: BH correction}
First we rank the values:

\begin{align*}
  p_1, p_7, p_2, p_3, p_5, p_4, p_6, p_{10}, p_8, p_9\\
\end{alig*}

Maximum PrN/r under FDR of 0.05: p6, so everything below that is good.\\

\section{Problem seven: }

\section{Problem eight: MEME}
\textbf{A.)}meme family.fasta -protein -mod zoops -minw 12\\
\textbf{A.)}meme proteins.fasta -dna -mod anr -nmotifs 3 -minw 10 -maxw 20 --minites 5\\

\section{Problem nine: RNA structure}
\textbf{A.)} Using the command cat finalRNA.fa | RNAfold -T 27, we get a free energy
of -20.38.\\

\textbf{B.)}I counted a total of 47 BP in the sequence. Of this, 30 are involved
in bonds in the predicted secondary structure, leaving 17 unpaired nucleotides.\\

\section{Problem 10: BLAST formatDB tasks}
\textbf{A.)} formatdb -p T -t ``Protein Database'' -n ProteinDB\\
\textbf{B.)} blastall -p blastn -i transcripts.fa -d ProteinDB -g 5 -E 1\\
-o transcripts.txt

\section{Problem 11: Matching}
\textbf{A.)} Small RNA-Seq\\
\textbf{B.)} RNA-Seq\\
\textbf{C.)} ChIP-Seq\\
\textbf{D.)} Gro-Seq\\
\textbf{E.)} Pol-II ChIP-Seq\\

\section{Problem 12: More matchign}
\textbf{A.)} MEME \\
\textbf{B.)} cuffdiff\\
\textbf{C.)} BLAST\\
\textbf{D.)} TopHat\\
\textbf{E.)} MACS\\
\textbf{F.)} Trinity\\
\textbf{G.)} cufflinks\\

\section{Problem 13: Complexity}
\begin{align*}
  \frac{1}{10}\log_4\left(\frac{10!}{10!}\right) &= 0\\
  \frac{1}{10}\log_4\left(\frac{10!}{5!*5!}\right) &= 0.398\\
\end{align*}

\section{Problem 14: transcriptoin}
I did the program on another computer so I can't paste it in here. I followed
exactly the program outlined in Section 1.2.2. I got: MGIYL* as my protein.\\

\end{document}

### 2016-04-05 quantum/latex/a9.tex
last minute adds.

@@
 
+\section{Problem three: p-state transitions}
+Here we use McIntyre (14.84) to find which quantum states can be reached from the p-state.
+From 14.84 we grab the following statement:\\
+
+\begin{align*}
+  \sqrt{\frac{3(2\ell_i+1)}{4\pi(2\ell_f+1)}}\bra{\ell_i,1,m_i,m}\ket{\ell_f,m_f}
+  \bra{\ell_i,1,0,0}\ket{\ell_f,0}\\
+\end{align*}
+
+We then plug in values for $\ell_i$ and find the CB coefficients that lead to nonzero
+values for this expression.\\
+
+By looking at the CB table, we can see that for the $m_1=0,m_2=0$ RHS braket, the only
+possible coupled basis J values are 0,2. This means that $\ell_f=0,2$, so by this we
+can see that $\ell_i\neq\ell_f$. This is the only selection rule imposed by the RHS
+braket, so we ignore it and look instead at the left braket:\\
+
+\begin{align*}
+  \bra{1,1,1,1}\ket{\ell_f,m_f}  \rightarrow\bra{\ell_f,m_f}=\braket{2,2}\\
+  \bra{1,1,1,0}\ket{\ell_f,m_f}  \rightarrow\bra{\ell_f,m_f}=\braket{2,1}\\
+  \bra{1,1,1,-1}\ket{\ell_f,m_f} \rightarrow\bra{\ell_f,m_f}=\braket{2,0},\braket{0,0}\\
+  \bra{1,1,0,1}\ket{\ell_f,m_f}  \rightarrow\bra{\ell_f,m_f}=\braket{2,1}\\
+  \bra{1,1,0,0}\ket{\ell_f,m_f}  \rightarrow\bra{\ell_f,m_f}=\braket{2,0},\braket{0,0}\\
+  \bra{1,1,0,-1}\ket{\ell_f,m_f} \rightarrow\bra{\ell_f,m_f}=\braket{2,-1}\\
+  \bra{1,1,-1,1}\ket{\ell_f,m_f} \rightarrow\bra{\ell_f,m_f}=\braket{2,0},\braket{0,0}\\
+  \bra{1,1,-1,0}\ket{\ell_f,m_f} \rightarrow\bra{\ell_f,m_f}=\braket{2,-1}\\
+  \bra{1,1,-1,-1}\ket{\ell_f,m_f}\rightarrow\bra{\ell_f,m_f}=\braket{2,-2}\\
+\end{align*}
+
+From this we can see that the allowed transitions from the p-state are either to the s state
+or any of the d-state orbitals.\\
+
+This was a weird problem. It is odd that you can treat the final electron state as
+represented in the coupled basis and the original state in the uncoupled basis and get
+meaningful results. This is really cool, but an odd usage of the CB table.\\
+
+\section{Problem four: d-state transition}
+Here we are given a particular quantum state, $n=4,\ell=3,m_\ell=3$, and asked to find
+the possible transitions that can be made from this state. We use the basic selection
+rules to answer this question: $\Delta \ell = \pm 1$ and $\Delta m_\ell = 0$.\\
+
+By applying these we see that the following state can be transitioned to:\\
+
+\begin{align*}
+  \ket{n, \ell_f,m_{\ell f}} \rightarrow
+  &\ket{3, 2, 2}\hspace{1cm}\sigma^-\mbox{ transition}\\
+\end{align*}
+
+From this state we can transition to:\\
+
+\begin{align*}
+  \ket{3, 2, 2}\rightarrow
+  \ket{2, 1, 1}\hspace{1cm}\sigma^-\mbox{ transition}\\
+\end{align*}
+
+And then
+
+\begin{align*}
+  \ket{2, 1, 1}\rightarrow
+  \ket{1, 0, 0}\hspace{1cm}\sigma^-\mbox{ transition}\\
+\end{align*}
+
+So we will always be decreasing our z-momentum, so all of our light has $\sigma^-$
+polarization. It is interesting that after such complicated arguments we get such
+simple selection rules for quantum state transitions.\\
+
 \end{document}

### 2016-04-05 solid_state/latex/a2.tex

--- 2016-01-20 quantum/latex/a2.tex
@@
 
-\section{9.13: }
-
 \section{9.14: Unknown QHO wavefunction}
@@
   2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/(2\hbar)}dx\\
-  &= 2\\
+  &= 2A\\
   c_1 &= A\int_{-\infty}^{\infty} \frac{2\sqrt{\frac{m\omega}{\hbar}x}}{\sqrt{2}}
@@
     2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
- &= -\frac{3}{\sqrt{2}}\\
+ &= -\frac{3}{\sqrt{2}}A\\
  c_2 &= A\int_{-\infty}^{\infty} \frac{\frac{4m\omega}{\hbar}x^2 - 2}{\sqrt{8}}
@@
     2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
- &= \sqrt{2}\\
+ &= \sqrt{2}A\\
 \end{align*}
 
-From this we can see that $\psi(x,0) = \sqrt{\frac{21}{2}}
-\left(2\ket{0} + -\frac{3}{\sqrt{2}}\ket{1} + \sqrt{2}\ket{2}\right)$. It is then easy to compute
+From this we can see that $\psi(x,0) = \left(\frac{2\sqrt{2}}{\sqrt{21}}\ket{0}
++ \frac{-3}{\sqrt{21}}\ket{1} + \frac{sqrt{2}}{\sqrt{21}}\ket{2}\right)$. It is then easy to compute
 the expectation value: \\
@@
 \begin{align*}
-  <E> &= \sigma \hbar\omega\left(n+\frac{1}{2}\right)|c_n|^2\\
-  &= \frac{2 + \frac{27}{4} + 5}{\frac{21}{2}} = 1.3 eV\\
+  <E> &= \sum \hbar\omega\left(n+\frac{1}{2}\right)|c_n|^2\\
+  &= \frac{2 + \frac{27}{4} + 5}{\frac{21}{2}} = 1.3\hbar\omega\\
 \end{align*}
+
+\section{9.13: Half-QHO well}
+(Inspired by http://www.physicspages.com/2012/08/18/half-harmonic-oscillator/)
+
+This system is identical to the 1D QHO, except all positive x values have infinite
+potential. So the system is a mixture of the infinite well and SHO potentials. We
+thus expect the wave function to have characteristics of both systems. For positive
+x values we expect the wave function to be zero, as it is in the infinite well.
+Because the negative x-values have a potential identical to QHO, we expect the wave
+function to be identical as well. So our total wave function is a Hermite/Gaussian
+for negative x values and zero for positive x values.\\
+
+However, we're not done. We notice that the x=0 boundary condition is not obeyed
+for all Hermite/Gaussian wave functions. We require the wave function to be zero
+at x=0, which is only true for odd QHO wave functions. Thus our energy spectrum
+is identical to the QHO spectru, but only for odd values of n:
+
+\begin{align*}
+  E_n &= \left(n + \frac{1}{2}\right)\hbar\omega\hspace{2cm}\mbox{n=1,3,5,...}\\
+\end{align*}
+
+For this problem we just applied the basic logic that the value of a wave function
+is only based on its current location, not the overall shape of the potential. The
+wave function for a system that is identical to the QHO will have a wave function
+identical to QHO in the similar regions. This is a powerful result.\\
+
+\section{5: 2D QHO}
+(Derivation inspired by
+$\mbox{nyu.edu/classes/tuckerman/adv.chem/lectures/lecture\_18/node6.html)}$\\
+To take the QHO to two dimensions, we use the following Hamiltonian:
+
+\begin{align*}
+  H &= \frac{\hat{p}_x^2}{2m} + \frac{\hat{p}_y^2}{2m}
+  + \frac{1}{2}m\omega^2\left(x^2+y^2\right)\\
+\end{align*}
+
+We notice that none of the cartesian terms in this Hamiltonian are coupled, so
+our energy eigenvalue equation then takes the form:
+
+\begin{align*}
+  H_{xy}X(x)Y(y) &= (E_x + E_y)X(x)Y(y)\\
+\end{align*}
+
+This equation is basically the sum of two uncoupled QHOs. Because of this, we
+can calculate the energy fairly easily:
+
+\begin{align*}
+  E_{xy} &= E_x + E_y = \left(n_x+n_y + 1\right)\hbar\omega\\
+\end{align*}
+
+From this we can see that the system is very degenerate. For the $n^{th}$-highest
+energy level there are $\frac{n}{2}$ possible states that have that energy: for
+the tenth energy level, the possible n-numbers are $1,9$, $2,8$, $3,7$, $4,6$ and
+$5,5$.\\
+
+We can construct the ground level energy state by taking the product of the two
+uncoupled QHOs:
+
+\begin{align*}
+  \psi_0(x,y) &= \left(\frac{m\omega}{\pi\hbar}\right)^{1/4}
+  e^{-m\omega\left(x^2+y^2\right)/2\hbar}\\
+\end{align*}
+
+The ease with which we go from the 1D to 2D harmonic oscillators comes from the fact
+that the x and y coordinates of the oscillator are not coupled. The x and y cross
+sections of the wave can take on any allowed 1D state. It is not intuitive to me
+why this method exactly works. Why are the energies of the x and y components
+just summed together to get total energy, as opposed to being multiplied or
+something? The degeneracy of the system is interesting. Because a sum is used,
+for large energies there are a huge number of ways to configure the system.\\
+
+\section{6: Position, momentum commutator}
+Here we find the value of the position, momentum commutator:
+\begin{align*}
+  [\hat{x}, \hat{p}] &= \hat{x}\hat{p} - \hat{p}\hat{x}\\
+  &= \left(\hat{x}\hat{p}\psi(x) - \hat{p}\hat{x}\psi(x)\right)\frac{1}{\psi(x)}\\
+  &= \left(\hat{x}\left(-i\hbar\frac{d}{dx}\right)\psi(x) - \hat{p}x\psi(x)\right)\frac{1}{\psi(x)}\\
+  &= \left(-i\hbar x \frac{d\psi}{dx} + i\hbar\left(\psi(x)
+  + x\frac{d\psi}{dx}\right)\right)\frac{1}{\psi(x)}\\
+  &= \left(i\hbar\psi(x)\right)\frac{1}{\psi(x)}\\
+  &= i\hbar\\
+\end{align*}
+Here we use the definitions of the position and momentum operators to find the value of their
+commutator. We need to operate on a generic wave function in order to do algebra on the
+operators, since the differential in the momentum operator is meaningless without a function
+to act on. This result is critical, since it is at the heart of the Heisenberg Uncertainty
+Principle: position and momentum don't commute, so they do not share a basis and they are not
+co-measurable. This is the mathematical representation of the popular saying ``you cannot know
+the position and momentum of a particle simultaneously.''
+
+\section{7: Commutator algebra}
+
+\begin{align*}
+  \left[\hat{A},\hat{A}^n\right] &= AA^n - A^nA = A^{n+1} - A^{n+1} = 0\\
+  \left[\hat{A},\hat{B}\right] &= \hat{A}\hat{B} - \hat{B}\hat{A}
+  = -\left(\hat{B}\hat{A} - \hat{A}\hat{B}\right) = -\left[\hat{B},\hat{A}\right]\\
+  \left[\hat{A},c\hat{B}\right] &= \hat{A}c\hat{B} - c\hat{B}\hat{A}
+  = c\left(\hat{A}\hat{B} - \hat{B}\hat{A}\right) = c\left[\hat{A},\hat{B}\right]\\
+  \left[\hat{A},\left(\hat{B}+\hat{C}\right)\right] &= \hat{A}\left(\hat{B} + \hat{C}\right)
+  - \left(\hat{B} + \hat{C}\right)\hat{A} = \hat{A}\hat{B} - \hat{B}\hat{A}
+  + \hat{A}\hat{C} + \hat{C}\hat{A} = \left[\hat{A},\hat{B}\right] + \left[\hat{A},\hat{C}\right]\\
+  \left[\hat{A},\hat{B}\hat{C}\right] &= \hat{A}\hat{B}\hat{C} - \hat{B}\hat{C}\hat{A}
+  = \hat{A}\hat{B}\hat{C} - \hat{B}\hat{A}\hat{C} + \hat{B}\hat{A}\hat{C} - \hat{B}\hat{C}\hat{A}
+  = \left[\hat{A},\hat{B}\right]\hat{C} + \hat{B}\left[\hat{A},\hat{C}\right]\\
+  \left[\hat{A},\left[\hat{B},\hat{C}\right]\right]
+  + \left[\hat{B},\left[\hat{C},\hat{A}\right]\right]
+  &+ \left[\hat{C},\left[\hat{A},\hat{B}\right]\right] =
+  \left[\hat{A},\hat{B}\hat{C} - \hat{C}\hat{B}\right]
+  + \left[\hat{B},\hat{C}\hat{A} - \hat{A}\hat{C}\right]
+  + \left[\hat{C},\hat{A}\hat{B} - \hat{B}\hat{A}\right] = \\
+  &= \hat{A}\hat{B}\hat{C} - \hat{A}\hat{C}\hat{B}
+  + \hat{B}\hat{C}\hat{A} - \hat{B}\hat{A}\hat{C}
+  + \hat{C}\hat{A}\hat{B} - \hat{C}\hat{B}\hat{A}\\
+\end{align*}
+
 \end{document}

### 2016-04-05 solid_state/latex/a6.tex

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{marvosym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\usepackage{amsmath}
\usepackage{braket}
\usepackage{amssymb}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Quantum Mechanics Assignment Six}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{11.12: Spin-1, Spin-$\frac{1}{2}$}
This system is made up of two particles, one with spin-1 and the other with spin-$\frac{1}{2}$.\\

\textbf{a.)} The total states possible for the system are:\\

\begin{align*}
  \ket{1,\frac{1}{2},-\hbar,-\frac{\hbar}{2}}\\
  \ket{1,\frac{1}{2},0,-\frac{\hbar}{2}}\\
  \ket{1,\frac{1}{2},\hbar,-\frac{\hbar}{2}}\\
  \ket{1,\frac{1}{2},-\hbar,\frac{\hbar}{2}}\\
  \ket{1,\frac{1}{2},0,\frac{\hbar}{2}}\\
  \ket{1,\frac{1}{2},\hbar,\frac{\hbar}{2}}\\
\end{align*}

\textbf{b.)} The highest z-projection of these states, the stretched state, is
$\ket{1,\frac{1}{2},\hbar,\frac{\hbar}{2}}$.\\

\textbf{c.)} Here we just solve for the coupled state wavefunctions using raising
and lowering operators:\\

\begin{align*}
  \ket{\frac{3}{2},\frac{3}{2}} &= \ket{1,\frac{1}{2},1,\frac{1}{2}}\\
  S_-\ket{\frac{3}{2},\frac{3}{2}} &= \left(S_{1-}+S_{2-}\right)\ket{1,\frac{1}{2},1,\frac{1}{2}}\\
  \sqrt{3}\hbar\ket{\frac{3}{2},\frac{1}{2}} &= \sqrt{2}\hbar\ket{1,\frac{1}{2},0,\frac{1}{2}} + \hbar\ket{1,\frac{1}{2},1,-\frac{1}{2}}\\
  \ket{\frac{3}{2},\frac{1}{2}} &= \sqrt{\frac{2}{3}}\ket{1,\frac{1}{2},0,\frac{1}{2}} + \frac{1}{\sqrt{3}}\ket{1,\frac{1}{2},1,-\frac{1}{2}}\\
\end{align*}

\begin{align*}
  S_-\ket{\frac{3}{2},\frac{1}{2}} &=
  \left(S_{1-}+S_{2-}\right)\sqrt{\frac{2}{3}}\ket{1,\frac{1}{2},0,\frac{1}{2}} + \frac{1}{\sqrt{3}}\ket{1,\frac{1}{2},1,-\frac{1}{2}}\\
  \sqrt{2}\hbar{\frac{3}{2},\frac{1}{2}} &= \frac{2\hbar}{\sqrt{3}}\ket{1,1/2,-1,1/2}
  + \frac{\sqrt{2}\hbar}{\sqrt{3}}\ket{1,1/2,0,-1/2} + \frac{\hbar\sqrt{2}}{\sqrt{3}}\ket{1,1/2,0,-1/2}\\
  \ket{3/2,-1/2} &= \frac{1}{\sqrt{3}}\ket{1,1/2,-1,1/2} + \sqrt{\frac{2}{3}}\ket{1,1/2,0,-1/2}\\
\end{align*}

\begin{align*}
  S_-\ket{3/2,-1/2} &= \left(S_{1-}+S_{2-}\right)\left(\frac{1}{\sqrt{3}}\ket{1,1/2,-1,1/2} + \sqrt{\frac{2}{3}}\ket{1,1/2,0,-1/2}\right)\\
  \ket{3/2,-3/2} &= \ket{1,1/2,-1,-1/2}\\
\end{align*}

We then use the method described on pg 371 of McIntyre to find the 1/2-states: creating a linear equation and solving it.\\

\begin{align*}
  \ket{1/2,1/2} &= A\ket{1,1/2,0,1/2} + B\ket{1,1/2,1,-1/2} + C\ket{1,1/2,-1,1/2}\\
  \bra{3/2,1/2}\bra{1/2,1/2} = 0 \rightarrow c = 0\\
  \bra{3/2,1/2}\ket{1/2,1/2} = 0 \rightarrow a\sqrt{\frac{2}{3}} + b\sqrt{\frac{1}{3}} \rightarrow -\frac{b}{\sqrt{2}} = a\\
  3a^2 = 1 \rightarrow a = \frac{1}{\sqrt{3}}\\
  \ket{1/2,1/2} = -\frac{1}{\sqrt{3}}\ket{1,1/2,0,-1/2} + \sqrt{\frac{2}{3}}\ket{1,1/2,1,-1/2}\\
\end{align*}

\begin{align*}
  S_-\ket{1/2,1/2} &= \left(S_{1-}+S_{2-}\right)\left(-\frac{1}{\sqrt{3}}\ket{1,1/2,0,-1/2} + \sqrt{\frac{2}{3}}\ket{1,1/2,1,-1/2}\right)\\
  \ket{1/2,-1/2} = \frac{1}{\sqrt{3}}\ket{1,1/2,0,-1/2} - \sqrt{\frac{2}{3}}\ket{1,1/2,-1,1/2}\\
\end{align*}

We then construct a Clebsch-Gordan matrix out of our states: (sorry for the bad formatting,
the formatting is identical to that found in table 11.3 in McIntyre)

\begin{align*}
  \begin{pmatrix}
    1 & 0 & 0 & 0\\
    0 & \frac{1}{\sqrt{3}} & 0 & 0\\
    0 & \frac{\sqrt{2}}{\sqrt{3}} & 0 & 0\\
    0 & 0 & \frac{\sqrt{2}}{\sqrt{3}} & 0\\
    0 & 0 & \frac{1}{\sqrt{3}} & 0\\
    0 & 0 & 0 & 1\\
  \end{pmatrix}
\end{align*}

\section{11.15: Coupled / Uncoupled bases}
This system consists of two particles, one spin-$\frac{3}{2}$ and the other
spin-$\frac{1}{2}$.\\

\textbf{a.)} First we list all the possible states this system can take on:

\begin{align*}
  \ket{\frac{3}{2},\frac{1}{2},\frac{3\hbar}{2},\frac{\hbar}{2}}\hspace{1cm}&
  \ket{\frac{3}{2},\frac{1}{2},\frac{1\hbar}{2},\frac{\hbar}{2}}\\
  \ket{\frac{3}{2},\frac{1}{2},\frac{-1\hbar}{2},\frac{\hbar}{2}}\hspace{1cm}&
  \ket{\frac{3}{2},\frac{1}{2},\frac{-3\hbar}{2},\frac{\hbar}{2}}\\
  \ket{\frac{3}{2},\frac{1}{2},\frac{3\hbar}{2},\frac{-\hbar}{2}}\hspace{1cm}&
  \ket{\frac{3}{2},\frac{1}{2},\frac{1\hbar}{2},\frac{-\hbar}{2}}\\
  \ket{\frac{3}{2},\frac{1}{2},\frac{-1\hbar}{2},\frac{-\hbar}{2}}\hspace{1cm}&
  \ket{\frac{3}{2},\frac{1}{2},\frac{-3\hbar}{2},\frac{-\hbar}{2}}\\
\end{align*}

\textbf{b.)} J can take on two values: $J = j_1 \pm j_2 = 2, 1$. $J=2$ corresponds to
$M = -2, -1, 0, 1, 2$, while $J=1$ corresponds to $M=-1, 0, 1$.\\

\textbf{c.)} Here we enumerate all the possible combinations of S and M:

\begin{align*}
  \ket{2,-2}, \ket{2,-1}, \ket{2,0}&, \ket{2,1}, \ket{2,2}\\
  \ket{1,-1}, \ket{1,0}&, \ket{1,-1}\\
\end{align*}

\textbf{d.)} Then we use a Clebsch-Gordan table to find how the coupled basis states
relate to the uncoupled basis:

\begin{align*}
  \ket{2,-2} &= \ket{\frac{3}{2},\frac{1}{2},-\frac{3}{2}, -\frac{1}{2}}\\
  \ket{2,-1} &=
  \frac{\sqrt{3}}{2}\ket{\frac{3}{2},\frac{1}{2},-\frac{1}{2},-\frac{1}{2}}
  + \frac{1}{2}\ket{\frac{3}{2},\frac{1}{2},-\frac{3}{2},\frac{1}{2}}\\
  \ket{2,0} &=
  \frac{1}{\sqrt{2}}\ket{\frac{3}{2},\frac{1}{2},\frac{1}{2},-\frac{1}{2}}
  + \frac{1}{\sqrt{2}}\ket{\frac{3}{2},\frac{1}{2},-\frac{1}{2},\frac{1}{2}}\\
  \ket{2,1} &=
  \frac{1}{2}\ket{\frac{3}{2},\frac{1}{2},\frac{3}{2},-\frac{1}{2}}
  + \frac{\sqrt{3}}{2}\ket{\frac{3}{2},\frac{1}{2},\frac{1}{2},\frac{1}{2}}
  \ket{2,2} &= \ket{\frac{3}{2},\frac{1}{2},\frac{3}{2},\frac{1}{2}}\\
  \ket{1,-1} &=
  \frac{1}{2}\ket{\frac{3}{2},\frac{1}{2},-\frac{1}{2},-\frac{1}{2}}
  + -\frac{\sqrt{3}}{2}\ket{\frac{3}{2},\frac{1}{2},-\frac{3}{2},\frac{1}{2}}\\
  \ket{1,0} &=
  \frac{1}{\sqrt{2}}\ket{\frac{3}{2},\frac{1}{2},\frac{1}{2},-\frac{1}{2}}
    - \frac{1}{\sqrt{2}}\ket{\frac{3}{2},\frac{1}{2},-\frac{1}{2},\frac{1}{2}}\\
  \ket{1,-1} &=
  \frac{\sqrt{3}}{2}\ket{\frac{3}{2},\frac{1}{2},\frac{3}{2},-\frac{1}{2}}
  - \frac{1}{2} \ket{\frac{3}{2},\frac{1}{2},\frac{1}{2},\frac{1}{2}}\\
\end{align*}

This was a neat problem that showed how easy it is to related the coupled and
uncoupled states when one is given a CG table.\\

\section{11.16: Hyperfine Deuterium}
We can either add or subtract the electron spin: $\frac{1}{2}$ from the nuclear
spin: 1. Thus we get total system angular momenta of either $\frac{3}{2}$ or
$\frac{1}{2}$. The former momentum supports 4 z-measurement states, while the latter
supports 2. This means we will be dealing with a 6x6 matrix in our calculations.\\

We use the definition $F = \left(S+I\right)^2$ to create an expression for $S\cdot I$:
$S\cdot I = \frac{1}{2}\left(F^2 - S^2 - I^2\right)$. Our calculation of the
perturbation energies then becomes a simple matter of coming up with the matrices for
$F^2, S^2$ and $I^2$:\\

\begin{align*}
  H'_{hf} &= \frac{A}{\hbar^2}S\cdot I\\
  S\cdot I &= \frac{1}{2}\left(F^2 - S^2 - I^2\right)\\
  &=
  \frac{\hbar^2}{2}
  \begin{pmatrix}
    \frac{15}{4} & 0 & 0 & 0 & 0 & 0\\
    0 & \frac{15}{4} & 0 & 0 & 0 & 0\\
    0 & 0 & \frac{15}{4} & 0 & 0 & 0\\
    0 & 0 & 0 & \frac{15}{4} & 0 & 0\\
    0 & 0 & 0 & 0 & \frac{3}{2} & 0\\
    0 & 0 & 0 & 0 & 0 & \frac{3}{2}\\
  \end{pmatrix}
  -
  \frac{\hbar^2}{2}
  \begin{pmatrix}
    \frac{3}{2} & 0 & 0 & 0 & 0 & 0\\
    0 & \frac{3}{2} & 0 & 0 & 0 & 0\\
    0 & 0 & \frac{3}{2} & 0 & 0 & 0\\
    0 & 0 & 0 & \frac{3}{2} & 0 & 0\\
    0 & 0 & 0 & 0 & \frac{3}{2} & 0\\
    0 & 0 & 0 & 0 & 0 & \frac{3}{2}\\
  \end{pmatrix}
  -
  \frac{\hbar^2}{2}
  \begin{pmatrix}
    2 & 0 & 0 & 0 & 0 & 0\\
    0 & 2 & 0 & 0 & 0 & 0\\
    0 & 0 & 2 & 0 & 0 & 0\\
    0 & 0 & 0 & 2 & 0 & 0\\
    0 & 0 & 0 & 0 & 2 & 0\\
    0 & 0 & 0 & 0 & 0 & 2\\
  \end{pmatrix}\\
  &= \frac{\hbar^2}{2}
  \begin{pmatrix}
    \frac{1}{4} & 0 & 0 & 0 & 0 & 0\\
    0 & \frac{1}{4} & 0 & 0 & 0 & 0\\
    0 & 0 & \frac{1}{4} & 0 & 0 & 0\\
    0 & 0 & 0 & \frac{1}{4} & 0 & 0\\
    0 & 0 & 0 & 0 & -\frac{11}{4} & 0\\
    0 & 0 & 0 & 0 & 0 & -\frac{11}{4}\\
  \end{pmatrix}\\
  H'_{hf} &= \frac{A}{2}
  \begin{pmatrix}
    \frac{1}{4} & 0 & 0 & 0 & 0 & 0\\
    0 & \frac{1}{4} & 0 & 0 & 0 & 0\\
    0 & 0 & \frac{1}{4} & 0 & 0 & 0\\
    0 & 0 & 0 & \frac{1}{4} & 0 & 0\\
    0 & 0 & 0 & 0 & -\frac{11}{4} & 0\\
    0 & 0 & 0 & 0 & 0 & -\frac{11}{4}\\
  \end{pmatrix}\\
\end{align*}

Where the above matrices are indexed as:
$\ket{\frac{3}{2}, \frac{3}{2}},\ket{\frac{3}{2}, \frac{1}{2}},
\ket{\frac{3}{2}, -\frac{1}{2}},\ket{\frac{3}{2}, -\frac{3}{2}},
\ket{\frac{1}{2}, \frac{1}{2}},\ket{\frac{1}{2}, -\frac{1}{2}}$. This corresponds
to energies of $\frac{A}{16}$ and $\frac{-11}{4}A$. The first energy has a
degeneracy of five and the latter has a degeneracy of two. This is interestingly
different from the hydrogen atom, since the neutron is contributing to the nuclear
spin! Though we see that the nucleus-electron spin coupling only splits the energies
in two.\\

\vspace{5cm}

\end{document}

### 2016-04-14 stat_mech/latex/a1.tex
Add new statmech stuff.

--- 2015-11-28 classical/latex/a9.tex
@@
 
-\title{Classical Mechanics Assignment 9}
+\title{Statistical Mechanics Assignment One}
 \author{Elliott Capek}
@@
 
-\section{Problem One: }
+\section{Problem One: Adiabatic Ideal Gas}
+\textbf{a.)}
+First we calculate the final temperature of the gas, given $V_i, V_f$ and $T_i$:
+
+\begin{align*}
+  P_iV_i^\gamma &= nRT_iV_i^{\gamma-1}\\
+  nRT_iV_i^{\gamma-1} &= nRT_fV_f^{\gamma-1}\\
+  T_f &= T_i\left(\frac{V_i}{V_f}\right)^{\gamma-1}
+\end{align*}
+
+\textbf{b.)}
+If $dS=0$, then $dQ/T=0$ and so $dU = -pdV$. We use this as another route for finding $T_f$ in
+terms of $V_i, V_f$ and $T_i$:
+
+\begin{align*}
+  dU &= -pdV = C_vdT\\
+  -\frac{RT}{V}dV &= C_vdT \hspace{2cm}\mbox{Molar specific heat, so n=1}\\
+  -R\int_{V_i}^{V_f}\frac{1}{V}dV &= C_v\int_{T_i}^{T_f}\frac{1}{T}dT\\
+  -R\ln\left(\frac{V_f}{V_i}\right) &= C_v\ln\left(\frac{T_f}{T_i}\right)\\
+  -\frac{C_p-C_v}{C_v}\ln\left(\frac{V_f}{V_i}\right) &= \ln\left(\frac{T_f}{T_i}\right)\\
+  \ln\left(\frac{V_f}{V_i}\right)^{1-\gamma} &= \ln\left(\frac{T_f}{T_i}\right)\\
+  \left(\frac{V_f}{V_i}\right)^{1-\gamma} &= \frac{T_f}{T_i}\\
+  T_f &= T_i\left(\frac{V_f}{V_i}\right)^{1-\gamma} = T_i\left(\frac{V_i}{V_f}\right)^{\gamma-1}\\
+\end{align*}
 

### 2016-04-15 stat_mech/latex/a1.tex
add new a1 statmech stuff.

@@
 
+\section{Problem Four: Gas Engine}
+Here we use the ideal gas property $dU = C_v$ to express all changes in internal energy
+as changes in temperature. We also use the adiabatic $PV^\gamma = k$ relation to convert
+between variables during the insulated work-doing steps:\\
+
+We first find the work done:
+\begin{align*}
+  W_{ab} &= -dU \hspace{2cm}\mbox{Q = 0, First law of Thermo}\\
+  &= -C_vdT \hspace{2cm}dU_{ideal} = CvdT\\
+  &= -C_v\left(T_b-T_a\right)\\
+  &= -C_vT_b\left(1-\frac{T_a}{T_b}\right)\\
+  &= -C_vT_b\left(1-\left(\frac{V_2}{V_1}\right)^{\gamma-1}\right)
+             \hspace{2cm}P_aV_2^\gamma = P_bV_1^\gamma \rightarrow
+             \frac{T_a}{T_b} = \left(\frac{V_2}{V_1}\right)^{\gamma-1}\\
+  W_{cd} &= C_vT_c\left(1 - \left(\frac{V_2}{V_1}\right)^{\gamma-1}\right)
+\end{align*}
+
+Then we find the heat added during the heating step:
+
+\begin{align*}
+  dQ_1 &= dU_V = C_v\left(T_c-T_b\right)\\
+\end{align*}
+
+Now we find the efficiency:
+
+\begin{align*}
+  \eta &= \frac{W}{Q_1} = \frac{C_vT_c\left(1 - \left(\frac{V_2}{V_1}\right)^{\gamma-1}\right)
+    - C_vT_b\left(1-\left(\frac{V_2}{V_1}\right)^{\gamma-1}\right)}{C_v\left(T_c-T_b\right)}\\
+  &= 1 - \left(\frac{V_2}{V_1}\right)^{\gamma-1}\\
+\end{align*}
+
 \end{document}

### 2016-04-15 solid_state/latex/a1.tex
Add new work for SS

@@
 \usepackage{fancyhdr}
-\usepackage{physics}
+\usepackage{amsmath}
 \usepackage{amssymb}
@@
 
-\subsection{d
+First, we use Mathematica to plot the wave functions:
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.2\textwidth]{../figures/dxy.png}
+  \includegraphics[width=0.2\textwidth]{../figures/dyz.png}
+  \includegraphics[width=0.2\textwidth]{../figures/dzx.png}
+  \includegraphics[width=0.2\textwidth]{../figures/dz2.png}
+  \includegraphics[width=0.2\textwidth]{../figures/dx2y2.png}
+  \caption{From first to last: $d_{xy}$, $d_{yz}$, $d_{zx}$, $d_{z^2}$ and $d_{x^2-y^2}$, all on the same XY plane}
+  \label{fig:dxy}
+\end{figure}
+
+We then convert the Cartesian d wavefunctions from Stross Eq 1.22 into spherical
+coordinates, so we can more easily see which spherical harmonics they are made
+up of.\\
+
+\begin{align*}
+  x &= r\cos(\phi)\sin(\theta)\\
+  y &= r\sin(\phi)\sin(\theta)\\
+  z &= r\cos(\theta)\\
+\end{align*}
+
+\begin{align*}
+  d_{xy} &= \left(\frac{15}{4\pi}\right)^{1/2}\frac{xy}{r^2}
+  = \left(\frac{15}{4\pi}\right)^{1/2}\cos(\phi)\sin(\phi)\sin^2(\theta)\\
+  d_{yz} &= \left(\frac{15}{4\pi}\right)^{1/2}\frac{yz}{r^2}
+  = \left(\frac{15}{4\pi}\right)^{1/2}\sin(\phi)\sin(\theta)\cos(\theta)\\
+  d_{zx} &= \left(\frac{15}{4\pi}\right)^{1/2}\frac{zx}{r^2}
+  = \left(\frac{15}{4\pi}\right)^{1/2}\cos(\phi)\sin(\theta)\cos(\theta)\\
+  d_{z^2} &= \left(\frac{15}{4\pi}\right)^{1/2}
+  \left(\frac{3z^2-r^2}{2\sqrt{3}r^2}\right)
+  = \left(\frac{15}{4\pi}\right)^{1/2}
+  \left(\frac{3\cos^2(\theta)-1}{2\sqrt{3}}\right)\\
+  d_{x^2-y^2} &= \left(\frac{15}{4\pi}\right)^{1/2}
+  \left(\frac{x^2-y^2}{2r^2}\right)
+  = \left(\frac{15}{4\pi}\right)^{1/2}
+  \frac{\cos^2(\phi)\sin^2(\theta)-\sin^2(\phi)\sin^2(\theta)}{2}\\
+\end{align*}
+
+We then write these real-basis functions as linear combinations of spherical harmonics:
+
+\begin{align*}
+  d_{xy} &= \frac{i}{\sqrt{2}}\left(Y_{2-2} - Y_{22}\right)
+  = \frac{i}{\sqrt{2}}\left(\frac{15}{32\pi}\right)^{1/2}
+  \sin^2\theta\left(e^{2i\phi}-e^{-2i\phi}\right)
+  = \left(\frac{15}{4\pi}\right)^{1/2}\sin^2\theta\sin\theta\cos\theta\\
+  d_{yz} &= \frac{i}{\sqrt{2}}\left(Y_{2-1}+Y_{21}\right)\\
+  d_{zx} &= \frac{1}{\sqrt{2}}\left(Y_{2-1}-Y_{21}\right)\\
+  d_{z^2} &= Y_{20}\\
+  d_{x^2-y^2} &= \frac{1}{\sqrt{2}}\left(Y_{22} + Y_{2-2}\right)\\
+\end{align*}
+
+\section{2: Matrix representation of operators}
+First, we write the matrix representation of the two angular momentum operators. Here
+the spherical harmonics for $\ell=2$ are orthogonal unit vectors, so $Y_{2-2}=(1 0 0 0 0)^T$, etc...\\
+
+\begin{align*}
+  L^2 =
+  6\hbar^2\begin{pmatrix}
+    1 & 0 & 0 & 0 & 0\\
+    0 & 1 & 0 & 0 & 0\\
+    0 & 0 & 1 & 0 & 0\\
+    0 & 0 & 0 & 1 & 0\\
+    0 & 0 & 0 & 0 & 1\\
+  \end{pmatrix}&
+  \hspace{2cm}
+  L_z =
+  \hbar\begin{pmatrix}
+    -2 & 0 & 0 & 0 & 0\\
+    0 & -1 & 0 & 0 & 0\\
+    0 & 0 & 0 & 0 & 0\\
+    0 & 0 & 0 & 1 & 0\\
+    0 & 0 & 0 & 0 & 2\\
+  \end{pmatrix}
+\end{align*}
+
+Now we compute the same matrices for the ``real'' vectors in their basis. In this representation:
+\begin{align*}
+  d_{xy} =
+  \begin{pmatrix}
+    1\\
+    0\\
+    0\\
+    0\\
+    0\\
+  \end{pmatrix},
+  \hspace{1cm}
+  d_{yz} =
+  \begin{pmatrix}
+    0\\
+    1\\
+    0\\
+    0\\
+    0\\
+  \end{pmatrix},
+  \hspace{1cm}
+  d_{zx} =
+  \begin{pmatrix}
+    0\\
+    0\\
+    1\\
+    0\\
+    0\\
+  \end{pmatrix},
+  \hspace{1cm}
+  d_{z^2} =
+  \begin{pmatrix}
+    0\\
+    0\\
+    0\\
+    1\\
+    0\\
+  \end{pmatrix},
+  \hspace{1cm}
+  d_{x^2-y^2} =
+  \begin{pmatrix}
+    0\\
+    0\\
+    0\\
+    0\\
+    1\\
+  \end{pmatrix},
+  \hspace{1cm}
+\end{align*}
+
+\begin{align*}
+  L^2 =
+  6\hbar^2\begin{pmatrix}
+    1 & 0 & 0 & 0 & 0\\
+    0 & 1 & 0 & 0 & 0\\
+    0 & 0 & 1 & 0 & 0\\
+    0 & 0 & 0 & 1 & 0\\
+    0 & 0 & 0 & 0 & 1\\
+  \end{pmatrix}&
+  \hspace{2cm}
+  L_z =
+  \begin{pmatrix}
+   2 & 0 & 0 & 0 & -2i\\
+   0 & 0 & -i & 0 & 0\\
+   0 & -i & -1 & 0 & 0\\
+   0 & 0 & 0 & 0 & 0\\
+   -2i & 0 & 0 & 0 & 0\\
+  \end{pmatrix}
+\end{align*}
+
+\section{Problem 3: Paper}
+Mohnish, Pandey, Filip Rasmussen, and Korina Kuhar. "Defect-Tolerant Monolayer
+Transition Metal Dichalcogenides." Nano Letters (2015). 29 Mar. 2016.
+
+This paper is on TMDCs, which are described in problem 4 as well. TMDCs are
+monolayers with a unit cell of one transition metal and two chalcogens: S, Se
+or Te. They are special because they are two-dimensional like graphene, have
+interesting band-gaps partially due to spin-spin coupling, and have mobile
+charge carriers. This paper studies the crystal defects of TMDCs, which are local
+scatterings of charge-carriers which impact the conductivity by changing the
+local band structure. This paper examines how defects add energies between
+the band gap, which probably gives the material interesting properties. It then
+goes on to explain why some TMDCs are defect-tolerant, while others are less
+defect-tolerant.\\
+
+\section{Problem 4: S, P and D-dominated materials}
+Lithium is a very soft, nonreactive metal whose chemical properties are largely
+decided by its 2s orbital. It lacks any $\ell=1$ electrons, so $\ell=0$ will of
+course dominate. The 2s orbital is close to the nucleus, so it is fairly low
+in energy and difficult to bond with. This might be why lithium metal has such
+a low melting point: its intermolecular bonds are weak.\\
+
+Graphene: Graphene has two types of bonds: $sp^2$ hybrid orbitals forming $\sigma$
+bonds with the other carbons, and a $p_z$ orbital forming the $\pi$ bond in a
+delocalized layer around graphene. Both orbitals are important, but lots of the
+properties of graphene, such as its conductivity, come from the delocalized
+$\pi$ orbital, which has low ionization energy and so can be used easily for
+conduction.\\
+
+Transition Metal Dichalcogenide Monolayer: TMDCs are monolayers of form $MX_2$, where
+M is a transition metal and X is S, Se or Te. The d-orbital is the outermost, and
+makes up both the valence and conduction bands. There is a high degree of spin-spin
+coupling among these d-orbitals, meaning the energy bands are very wide and have
+spacings between them. This seems like it would make these materials bad at
+conducting, since the conducting-band electrons have holes spaced far apart in energy,
+meaning electrons have to amass lots of energy before they can join the conduction
+band. This is just a guess, though.\\
+
+All information from respective Wikipedia articles.\\
+
+\section{Problem 5: All traces for all bases are identical}
+For any basis, the trace of an operator H is given by:
+
+\begin{align*}
+  H_{11} + H_{22} + H_{33} + ... + H_{NN}\\
+\end{align*}
+
+This is equivalent to:
+
+\begin{align*}
+  <1|H|1> + <2|H|2> + ... + <N|H|N> &= E_1 + E_2 + ... + E_N\\
+\end{align*}
+
+where the Es are the eigenvalues of the H operator. Eigenvalues are independent
+of basis since they represent the physical measurables of the system. The trace is the
+sum of all the possible measurements, and so is thus independent of basis.\\
 

### 2016-04-15 solid_state/latex/a2.tex
all

--- 2016-01-20 quantum/latex/a2.tex
@@
 \usepackage{marvosym} % Required for symbols in the colored box
+\usepackage{ifsym} % Required for symbols in the colored box
 
@@
 \usepackage{amsmath}
-% \usepackage{physics}
 \usepackage{amssymb}
@@
 
-\title{Quantum Mechanics Assignment Two}
+\title{Solid State Assignment Two}
 \author{Elliott Capek}
@@
 
-\section{9.4: Forbidden region probability for QHO}
-For this problem we want to calculate the probability of measuring a QHO ground-state object outside
-of the classically allowed region. We do this by calculating the turning points of a classical HO,
-ie finding the x-values where energy equals potential, and then integrating $\psi^*\psi$ over the
-forbidden region to find the chance of forbidden observation.\\
+\section{Problem 2: O-C-O Anti/Bonding energies}
+This problem asks us to come up with the anti/bonding energies for an O-C-O molecule, given its
+on-site and hopping energies.\\
 
-We first find the turning point for the ground state $n=0$:
+First we construct the Hamiltonian for this system, using the rules defined in the problem:\\
+
 \begin{align*}
-  V_{HO} &= E_0\\
-  \frac{1}{2}m\omega^2x^2 &= \hbar\omega\left(0+\frac{1}{2}\right)\\
-  x_0 &= \sqrt{\frac{\hbar}{m\omega}}\\
+  H &=
+  \begin{pmatrix}
+    E_O & \beta & 0\\
+    \beta & E_C & \beta\\
+    0 & \beta & E_O\\
+  \end{pmatrix}
 \end{align*}
 
-We then find the probability of finding the particle in this region:
+We then find the eigenvalues of this equation:\\
 
 \begin{align*}
-  P_{forb} &= \int_{\infty}^{-x_0} \psi_0^*\psi_0 dx + \int_{x_0}^{\infty} \psi_0^*\psi_0 dx\\
-  &= \int_{\infty}^{-x_0} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar}dx
-  + \int_{x_0}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar} dx\\
-  &= \sqrt{\frac{m\omega}{\pi\hbar}} \left(\frac{-\hbar}{m\omega}\right)\left(
-    \frac{1}{x}e^{-m\omega x^2 / \hbar} \Bigg|_{-\infty}^{-\sqrt{\frac{\hbar}{m\omega}}} +
-    \frac{1}{x}e^{-m\omega x^2 / \hbar} \Bigg|_{\sqrt{\frac{\hbar}{m\omega}}}^{\infty}\right)\\
-  &= \frac{2}{\sqrt{\pi}} = 0.41\\
+  det(H) &=
+  det\begin{bmatrix}
+    E_O-\lambda & \beta & 0\\
+    \beta & E_C-\lambda & \beta\\
+    0 & \beta & E_O-\lambda\\
+  \end{bmatrix}\\
+  \mbox{Mathematica} \rightarrow
+  &= E_O, \frac{1}{2}\left(\pm\sqrt{8\beta^2 + E_c^2 - 2E_cE_O + E_O^2} + E_c + E_O\right)\\
 \end{align*}
 
-%% \section{9.9: QHO expectation values via integration, operator method}
-%% In this problem we use both wave function integration and the operator method to find the
-%% expectation values of various observables: x, p, $x^2$, $p^2$.
+This is an interesting result. As expected, there are three eigenvalues for a 3x3 matrix. One
+corresponds to the original atomic oxygen energy level, which seems odd. This seems to be saying
+that the oxygen atomic orbitals are retained in some way in the final molecule. The bonding
+(lower energy) orbital has energy
+$\frac{1}{2}\left(-\sqrt{8\beta^2 + E_c^2 - 2E_cE_O + E_O^2} + E_c + E_O\right)$, while the
+antibonding (higher energy) orbital has
+$\frac{1}{2}\left(\sqrt{8\beta^2 + E_c^2 - 2E_cE_O + E_O^2} + E_c + E_O\right)$.
 
-%% \textbf{a.) Ground state wave function calculations}
-%% \begin{align*}
-%%   <x> &= \int_{-\infty}^{\infty} \psi^*x\psi dx\\
-%%   &= \int_{-\infty}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar}x dx = 0\\
-%% \end{align*}
-%% We don't need to perform the $<x>$ integral because it involves integrating an odd function over
-%% an even domain, which always results in zero. This is to be expected, because the wave ground-state
-%% wave function is symmetric across zero.\\
+\section{Problem Three: Types of Bonds}
+Here we discuss different types of bonds and what they're made up of.\\
 
-%% \begin{align*}
-%%   <p> &= \int_{-\infty}^{\infty} \psi^* -i\hbar \frac{d}{dx}\psi dx\\
-%%   &= \int_{-\infty}^{\infty}
-%%   -i\sqrt{\frac{m\omega}{\pi\hbar}}(-m\omega x)e^{\frac{-m\omega x^2}{2\hbar}}dx = 0\\
-%% \end{align*}
-%% This integral is zero since we are again integrating an odd function over an even domain. This
-%% makes sense because, since the spacial distribution of the wave function is equal on both sides
-%% of the x-axis, the velocities (and hence momenta) should be equal as well.\\
+\textbf{sigma between $p_z$ orbitals: } Here two $p_z$ orbitals subtract from
+each other (bonding) or add (antibonding) on the bonding axis to form new orbitals.\\
+\vspace{3cm}
 
-%% \begin{align*}
-%%   <x^2> &= \int_{-\infty}^{\infty} \psi^*x^2\psi dx\\
-%%   &= \int_{-\infty}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar}x^2 dx\\
-%%   &= 2\sqrt{\frac{m\omega}{\pi\hbar}}\int_{0}^{\infty} e^{-m\omega x^2 / \hbar}x^2 dx\\
-%%   &= 2\sqrt{\frac{m\omega}{\pi\hbar}} *
-%%   \frac{\sqrt{\pi}}{4(\frac{m\omega}{\hbar})^{3/2}} \hspace{2cm}\mbox{Wolfram}\\
-%%   &= \frac{\hbar}{2m\omega}\\
-%% \end{align*}
-%% Here we exploit the fact that integrating an even function over an even domain is equivalent to
-%% doubling the integral of the function over half the domain. The answer makes sense, since increasing
-%% $m\omega$ corresponds to making the quadratic potential steeper, hence making the wave function
-%% tighter.\\
+\textbf{pi between $p_z$ orbitals: } Here two $p_z$ orbitals add to each other (bonding)
+or subtract (antibonding) perpendicular to the bonding axis to form new orbitals.\\
+\vspace{3cm}
 
-%% \begin{align*}
-%%   <p^2> &= \int_{-\infty}^{\infty} \psi^* \left(-\hbar^2 \frac{d^2}{dx^2}\right)\psi dx\\
-%%   &= \sqrt{\frac{m\omega}{\pi\hbar}}\int_{-\infty}^{\infty} -\hbar^2e^{\frac{-m\omega x^2}{\hbar}}
-%%   \left(\frac{m^2\omega^2x^2}{\hbar^2} - \frac{m\omega}{\hbar}\right)dx\\
-%%   &= -\hbar^2\sqrt{\frac{m\omega}{\pi\hbar}}\left(
-%%   \int_{-\infty}^{\infty} e^{\frac{-m\omega x^2}{\hbar}}\frac{m^2\omega^2x^2}{\hbar^2}dx -
-%%   \int_{-\infty}^{\infty} e^{\frac{-m\omega x^2}{\hbar}}\frac{m\omega}{\hbar}\right)dx\\
-%%   &= -\hbar^2\sqrt{\frac{m\omega}{\pi\hbar}}
-%%   \left(\frac{1}{2}\sqrt{\frac{m\pi\omega}{\hbar}} -
-%%   \sqrt{\frac{m\pi\omega}{\hbar}}\right) \hspace{2cm}\mbox{Wolfram}\\
-%%   &= \frac{\hbar m\omega}{2}\\
-%% \end{align*}
-%% This answer is intuitive, since increasing mass as well as oscillatory speed should both increase
-%% the expected momentum.
+\textbf{sigma between $d_{z^2}$ orbitals: } Here two $d_{z^2}$ orbitals subtract from each other
+(bonding) or add (antibonding) on the bonding axis to form new orbitals.\\
+\vspace{3cm}
 
-%% \textbf{b.) Any-state operator method calculations}
-%% Here we define the position and momentum operators in terms of the raising and lowering operators:
-%% \begin{align*}
-%%   \hat{x} &= \sqrt{\frac{\hbar}{2m\omega}}\left(a + a^\dagger\right)\\
-%%   \hat{p} &= i\sqrt{\frac{m\omega\hbar}{2}}\left(a^\dagger - a\right)\\
-%% \end{align*}
-%% We then solve for the various expectation values using these definitions:
+\textbf{pi between $d_{yz}$ orbitals: } Here two $d_{yz}$ orbitals subtract from each other
+(bonding) or add (antibonding) on the bonding plane such that each of the two lobes
+connect to form new orbitals.\\
+\vspace{3cm}
 
-%% \begin{align*}
-%%   <x> &= \bra{n}x\ket{n}\\
-%%   &= \sqrt{\frac{\hbar}{2m\omega}}\bra{n}\left(a + a^\dagger\right)\ket{n}\\
-%%   &= \sqrt{\frac{\hbar}{2m\omega}}
-%%   \left(\sqrt{n}\bra{n}\ket{n-1} + \sqrt{n+1}\bra{n}\ket{n+1}\right) = 0\\
-%% \end{align*}
-%% This answer indicates that all pure QHO wave functions are centered around zero. This makes
-%% sense since the QHO wave functions are all made up of either even or odd functions. When squared,
-%% as is done when finding the probability, the equation becomes entirely even, meaning it is
-%% symmetric across the y-axis. From this we can see that all QHO wave functions have an expectation
-%% value of zero for x-measurement.
-
-%% \begin{align*}
-%%   <p> &= \bra{n}\hat{p}\ket{n}\\
-%%   &= i\sqrt{\frac{m\hbar\omega}{2}}\bra{n}\left(a^\dagger-a\right)\ket{n}\\
-%%   &= i\sqrt{\frac{m\hbar\omega}{2}}
-%%   \left(\sqrt{n+1}\bra{n}\ket{n+1} - \sqrt{n}\bra{n}\ket{n-1}\right) = 0\\
-%% \end{align*}
-%% This is an interesting result. It is intuitive that any wave function should have an expectation of
-%% zero for momentum, else the wave function would move through space due to a nonzero average
-%% velocity.
-
-%% \begin{align*}
-%%   <x^2> &= \bra{n}x^2\ket{n}\\
-%%   &= \frac{\hbar}{2m\omega}\bra{n}\left(a + a^\dagger\right)\left(a + a^\dagger\right)\ket{n}\\
-%%   &= \frac{\hbar}{2m\omega}\bra{n}
-%%   \left(aa + aa^\dagger + a^\dagger a + a^\dagger a^\dagger\right)\ket{n}\\
-%%   &= \frac{\hbar}{2m\omega}
-%%   \left(aa + aa^\dagger + a^\dagger a + a^\dagger a^\dagger\right)\ket{n}\\
-%%   &= \frac{\hbar}{2m\omega} \left(\sqrt{n-1}\sqrt{n-2}\bra{n}\ket{n-2} + (n+1)\bra{n}\ket{n} +
-%%   (n)\bra{n}\ket{n} + \sqrt{n+1}\sqrt{n+2}\bra{n}\ket{n+2}\right)\\
-%%   &= \frac{\hbar}{2m\omega} \left(2n+1\right)\\
-%% \end{align*}
-%% This is an interesting result which indicates that for each extra rung on the ladder, the
-%% expectation of the squared position increases linearly, meaning the expectation of the magnitude
-%% of x increases as $\sqrt{n}$. This makes sense, since increasing energy levels have less and less
-%% room to expand in the quadratic potential well. The $n=0$ state corresponds to the wave function
-%% solution for the ground state found above.\\
-
-%% \begin{align*}
-%%   <p^2> &= \bra{n}p^2\ket{n}\\
-%%   &= -\frac{m\hbar\omega}{2}\bra{n}\left(a^\dagger-a\right)\left(a^\dagger-a\right)\ket{n}\\
-%%   &= -\frac{m\hbar\omega}{2}\bra{n}
-%%   \left(a^\dagger a^\dagger - a^\dagger a - aa^\dagger + aa\right)\ket{n}\\
-%%   &= -\frac{m\hbar\omega}{2}\bra{n}\left(\sqrt{n-1}\sqrt{n-2}\bra{n}\ket{n-2}
-%%   - (n)\bra{n}\ket{n} - (n+1)\bra{n}\ket{n} + \sqrt{n+1}\sqrt{n+2}\bra{n}\ket{n+2}\right)\\
-%%   &= \frac{\hbar m\omega}{2}\left(2n + 1\right)
-%% \end{align*}
-%% This result is similar to the above one for $x^2$ with different units. It is still a $\sqrt{n}$
-%% relationship for momentum, which makes sense.\\
-
-%% \textbf{c.) Uncertainty principle}\\
-%% Since the above two cases agree with each other, we check that the more general case of part
-%% \textbf{b.} agrees with the uncertainty principle.
-
-%% \begin{align*}
-%%   \Delta x &= \sqrt{<x^2>-<x>^2}\\
-%%   &= \sqrt{\frac{\hbar}{2m\omega} \left(2n+1\right)}\\
-%%   \Delta p &= \sqrt{<p^2>-<p>^2}\\
-%%   &= \sqrt{\frac{\hbar m\omega}{2}\left(2n + 1\right)}\\
-%%   \Delta x \Delta p &= \sqrt{\frac{\hbar}{2m\omega} \left(2n+1\right)}
-%%   \sqrt{\frac{\hbar m\omega}{2}\left(2n + 1\right)}\\
-%%   &= \frac{\hbar}{2}(2n+1)\\
-%% \end{align*}
-%% This is an interesting result which indicates that uncertainty increases as the pure wave function
-%% gets more energetic. Ground state is the state with maximum certainty. The higher states get
-%% progressively more uncertain, which makes sense since both the velocity and the well get larger
-%% for higher states.\\
-
-%% \section{9.14: Unknown QHO wavefunction}
-%% We are given the wave function $\psi(x,0) = A\left[1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
-%%   2\frac{m\omega}{\hbar}x^2\right]$. Because its highest-order polynomial term is $x^2$, we know
-%% that it must be made up of the $n = 0, 1, 2$ energy states. We now figure out the coefficients for
-%% each:
-
-%% \begin{align*}
-%%   c_0 &= \int_{-\infty}^{\infty} \phi_0^*\psi dx\\
-%%   &= A\int_{-\infty}^{\infty} \left(\frac{m\omega}{\pi\hbar}\right)^{1/2}e^{-m\omega x^2/\hbar}
-%%   \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
-%%   2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/(2\hbar)}dx\\
-%%   &= 2\\
-%%   c_1 &= A\int_{-\infty}^{\infty} \frac{2\sqrt{\frac{m\omega}{\hbar}x}}{\sqrt{2}}
-%%     \left(\frac{m\omega}{\pi\hbar}\right)^{1/2} \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
-%%     2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
-%%  &= -\frac{3}{\sqrt{2}}\\
-%%  c_2 &= A\int_{-\infty}^{\infty} \frac{\frac{4m\omega}{\hbar}x^2 - 2}{\sqrt{8}}
-%%     \left(\frac{m\omega}{\pi\hbar}\right)^{1/2} \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
-%%     2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
-%%  &= \sqrt{2}\\
-%% \end{align*}
-
-%% From this we can see that $\psi(x,0) = \sqrt{\frac{21}{2}}
-%% \left(2\ket{0} + -\frac{3}{\sqrt{2}}\ket{1} + \sqrt{2}\ket{2}\right)$. It is then easy to compute
-%% the expectation value: \\
-
-%% \begin{align*}
-%%   <E> &= \sigma \hbar\omega\left(n+\frac{1}{2}\right)|c_n|^2\\
-%%   &= \frac{2 + \frac{27}{4} + 5}{\frac{21}{2}} = 1.3 eV\\
-%% \end{align*}
-\section{5: 2D QHO}
-To take the QHO to two dimensions, we use the following Hamiltonian:
-
-\begin{align*}
-  H &= \frac{\hat{p}_x^2}{2m} + \frac{\hat{p}_y^2}{2m}
-  + \frac{1}{2}m\omega^2\left(x^2+y^2\right)\\
-\end{align*}
-
-We notice that none of the terms in this Hamiltonian are coupled, so we break it
-up into two Hamiltonians:
-
-\begin{align*}
-  H_x &= \frac{\hat{p}_x^2}{2m} \frac{1}{2}m\omega^2x^2)\\
-  H_y &= \frac{\hat{p}_y^2}{2m} \frac{1}{2}m\omega^2y^2)\\
-  H &= H_x + H_y\\
-\end{align*}
-
-our energy eigenvalue equation then takes the form:
-
-\begin{align*}
-  \left(\hat{H}_x + \hat{H}_y\right)X(x)Y(y) &= (E_x + E_y)X(x)Y(y)\\
-\end{align*}
-
-This equation is basically the sum of two uncoupled QHOs. Because of this, we
-can calculate the energy fairly easily:
-
-\begin{align*}
-  E_{xy} &= E_x + E_y = \left(n_x+n_y + 1\right)\hbar\omega\\
-\end{align*}
-
-From this we can see that the system is very degenerate. For the $n^{th}$-highest
-energy level there are $\frac{n}{2}$ possible states that have that energy: for
-the tenth energy level, the possible n-numbers are $1,9$, $2,8$, $3,7$, $4,6$ and
-$5,5$.\\
-
-We can construct the ground level energy state by taking the product of the two
-uncoupled QHOs:
-
-\begin{align*}
-  \psi_0(x,y) &= \left(\frac{m\omega}{\pi\hbar}\right)
-  e^{-m\omega\left(x^2+y^2\right)/2\hbar}\\
-\end{align*}
-
-\section{6: Position, momentum commutator}
-Here we find the value of the position, momentum commutator:
-\begin{align*}
-  [\hat{x}, \hat{p}] &= \hat{x}\hat{p} - \hat{p}\hat{x}\\
-  &= \left(\hat{x}\hat{p}\psi(x) - \hat{p}\hat{x}\psi(x)\right)\frac{1}{\psi(x)}\\
-  &= \left(\hat{x}\left(-i\hbar\frac{d}{dx}\right)\psi(x) - \hat{p}x\psi(x)\right)\frac{1}{\psi(x)}\\
-  &= \left(-i\hbar x \frac{d\psi}{dx} + i\hbar\left(\psi(x)
-  + x\frac{d\psi}{dx}\right)\right)\frac{1}{\psi(x)}\\
-  &= \left(i\hbar\psi(x)\right)\frac{1}{\psi(x)}\\
-  &= i\hbar\\
-\end{align*}
-Here we use the definitions of the position and momentum operators to find the value of their
-commutator. We need to operate on a generic wave function in order to do algebra on the
-operators, since the differential in the momentum operator is meaningless without a function
-to act on. This result is critical, since it is at the heart of the Heisenberg Uncertainty
-Principle: position and momentum don't commute, so they do not share a basis and they are not
-co-measurable. This is the mathematical representation of the popular saying ``you cannot know
-the position and momentum of a particle simultaneously.''
-
-\section{7: Commutator algebra}
-
-\begin{align*}
-  \left[\hat{A},\hat{A}^n\right] &= AA^n - A^nA = A^{n+1} - A^{n+1} = 0\\
-  \left[\hat{A},\hat{B}\right] &= \hat{A}\hat{B} - \hat{B}\hat{A}
-  = -\left(\hat{B}\hat{A} - \hat{A}\hat{B}\right) = -\left[\hat{B},\hat{A}\right]\\
-  \left[\hat{A},c\hat{B}\right] &= \hat{A}c\hat{B} - c\hat{B}\hat{A}
-  = c\left(\hat{A}\hat{B} - \hat{B}\hat{A}\right) = c\left[\hat{A},\hat{B}\right]\\
-  \left[\hat{A},\left(\hat{B}+\hat{C}\right)\right] &= \hat{A}\left(\hat{B} + \hat{C}\right)
-  - \left(\hat{B} + \hat{C}\right)\hat{A} = \hat{A}\hat{B} - \hat{B}\hat{A}
-  + \hat{A}\hat{C} + \hat{C}\hat{A} = \left[\hat{A},\hat{B}\right] + \left[\hat{A},\hat{C}\right]\\
-  \left[\hat{A},\hat{B}\hat{C}\right] &= \hat{A}\hat{B}\hat{C} - \hat{B}\hat{C}\hat{A}
-  = \hat{A}\hat{B}\hat{C} - \hat{B}\hat{A}\hat{C} + \hat{B}\hat{A}\hat{C} - \hat{B}\hat{C}\hat{A}
-  = \left[\hat{A},\hat{B}\right]\hat{C} + \hat{B}\left[\hat{A},\hat{C}\right]\\
-  \left[\hat{A},\left[\hat{B},\hat{C}\right]\right]
-  + \left[\hat{B},\left[\hat{C},\hat{A}\right]\right]
-  &+ \left[\hat{C},\left[\hat{A},\hat{B}\right]\right] =
-  \left[\hat{A},\hat{B}\hat{C} - \hat{C}\hat{B}\right]
-  + \left[\hat{B},\hat{C}\hat{A} - \hat{A}\hat{C}\right]
-  + \left[\hat{C},\hat{A}\hat{B} - \hat{B}\hat{A}\right] = \\
-  &= \hat{A}\hat{B}\hat{C} - \hat{A}\hat{C}\hat{B}
-  + \hat{B}\hat{C}\hat{A} - \hat{B}\hat{A}\hat{C}
-  + \hat{C}\hat{A}\hat{B} - \hat{C}\hat{B}\hat{A}\\
-\end{align*}
-
+\section{Problem Five: Silicon Crystal Structure}
+Silicon forms a face-centered cubic diamond crystal lattice, where the center of each face, in
+addition to each corner, has an atom at it. The length of a side of the cube is $5.431$ \AA
+$= 5.431 * 10^{-10}m$. This is the lattice constant of Silicon crystal. Correspondingly the
+bond lengths in silicon crystal are $5.431$ \AA (between corners) and
+$5.431 * \frac{\sqrt{2}}{2} = 3.84$ \AA.
+\hspace{2cm}
+Silicon is a diamond FCC, so its space group is Fd$\overline{3}$m. This space group is a
+Bravais lattice, meaning a lattice that is generated by adding any integer multiple of one
+of three basis vectors to a starting point to generate an arbitrary member of the lattice.\\
 \end{document}

### 2016-04-15 solid_state/latex/a2.tex
a2.tex

@@
 
+\section{Problem Four: Heteronuclear Diatomic Molecule}
+Here we follow the same process for finding the energies of a multiatomic system that we used
+in problem Two: defining which atomic wave functions overlap on-site and off-site, deciding
+the energies of these overlaps, creating a Hamiltonian and finding the energies of the
+Hamiltonian. We are given that for this heteroatomic system atom A's on-site interaction
+energy is $E_A$ and B's is $E_B$. Similarly B's is $E_B$, and $E_A > E_B$. We then let out
+hopping integral take the value $H_{12} = H_{21} = \beta$. Thus our Hamiltonian is:
+
+\begin{align*}
+  H &=
+  \begin{pmatrix}
+    E_A & \beta\\
+    \beta & E_B\\
+  \end{pmatrix}
+\end{align*}
+
+Our next job is to find the eigenvalues of this matrix:
+
+\begin{align*}
+  \det\begin{pmatrix}
+    E_A-\lambda & \beta\\
+    \beta & E_B-\lambda\\
+  \end{pmatrix}
+  &= 0\\
+  (E_A-\lambda)(E_B-\lambda) - \beta^2 &= 0\\
+  \lambda^2 - \lambda(E_A + E_B )+ E_AE_B - \beta^2 &= 0\\
+  \lambda &= \frac{(E_A + E_B) \pm \sqrt{(E_A + E_B)^2 - 4(E_AE_B - \beta^2)}}{2}\\
+  &= \frac{(E_A + E_B) \pm \sqrt{E_A^2 + E_B^2 - 2E_AE_B + 4\beta^2}}{2}\\
+  &= \frac{(E_A + E_B) \pm \sqrt{(E_A - E_B)^2 + 4\beta^2}}{2}\\
+  &= \frac{(E_A + E_B)}{2} \pm \sqrt{\frac{(E_A - E_B)^2}{4} + \beta^2}\\
+\end{align*}
+
 \section{Problem Five: Silicon Crystal Structure}

### 2016-04-15 stat_mech/latex/a1.tex
blablabl

@@
 
+\section{Problem Two: Sound compression of a liquid}
+We use the following variable scheme: $\Delta x$ is the initial length of our 1D
+liquid. $\epsilon$ is the change in our 1D liquid length due to rarefaction
+on the left side. $\epsilon + \Delta \epsilon$ is the change in length to the right
+side.
+
+\begin{align*}
+  K_S &= -\frac{1}{V}\left(\frac{dV}{dP}\right)_S
+  = \frac{-1}{V}\frac{\Delta V}{\Delta P}& \hspace{2cm}&\mbox{P-P0: displacement from eq pressure}\\
+  &= \frac{-1}{A\Delta x}\frac{A\Delta\epsilon}{P-P_0}& \hspace{2cm}&\mbox{A: Area}\\
+  \Delta P &= -\frac{1}{K_s}\frac{d\epsilon}{dx}\\
+\end{align*}
+
+We then need to find the time derivative of $\epsilon$, which is the force on our
+1D cell, or how fast it is growing. The pressure wave is moving in the negative
+x direction, so the force on the right of the cell is $P_0 + \Delta P$. On the
+left, our cell is being pushed by the normal pressure $P_0$. Thus by Newton's First
+Law we see that:\\
+
+\begin{align*}
+  m\left(\frac{d^2\epsilon}{dt^2}\right) &= P_0A - (P_0+\Delta P)A = -\Delta PA\\
+  \rho A \Delta x \frac{d^2\epsilon}{dt^2} &= -A\frac{dP}{dx}\Delta x\\
+  \rho \frac{d^2\epsilon}{dt^2} &= -\frac{dP}{dx}\\
+  \rho \frac{d^2\epsilon}{dt^2} &= -\frac{d \Delta P}{dx}\\
+  \rho \frac{d^2}{dt^2}\frac{d\epsilon}{dx} &= -\frac{d^2P}{dx^2}\\
+  \rho \frac{d^2}{dt^2}\left(-PK_s\right) &= -\frac{d^2P}{dx^2}\\
+  \rho K_s \frac{d^2P}{dt^2}\left(P\right) &= \frac{d^2P}{dx^2}\\
+  \frac{d^2P}{dt^2} &= \frac{1}{\rho K_s }\frac{d^2P}{dx^2}\\
+\end{align*}
+
+Solution inspired by several people from SPS.\\
+
+\section{Problem Three: Solid $\Delta T, \Delta P$ relation}
+
+\begin{align*}
+  dS &= \left(\frac{dS}{dT}_P\right)dT + \left(\frac{dS}{dP}_T\right)dP = 0\\
+  \left(\frac{dS}{dT}_P\right)dT &= -\left(\frac{dS}{dP}_T\right)dP\\
+  \frac{C_P}{T}dT &= \left(\frac{dV}{dT}_P\right)dP\\
+  \frac{C_P}{T}dT &= V\alpha dP\\
+  \frac{C_P}{TV\alpha} &= -\left(\frac{dP}{dT}_S\right)\\
+  \frac{C_P}{TV\alpha} &= \frac{\Delta P}{\Delta T}\\
+  \Delta T &= \frac{TV\alpha}{C_P} = \frac{\alpha T}{\rho c_p} \hspace{2cm}
+  c_p \mbox{ is the volumetric specific heat}
+\end{align*}
+
 \section{Problem Four: Gas Engine}

### 2016-04-26 stat_mech/latex/a2.tex
Add statmech A2

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{marvosym} % Required for symbols in the colored box
\usepackage{ifsym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

\allowdisplaybreaks

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Statistical Mechanics Assignment Two}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem One: Harmonic Oscillator Energy}
\textbf{a.)} First we find the ratio of probabilities for the first and
second energies using the Botzmann distribution:\\
\begin{align*}
  \frac{p_1}{p_0} &= \left(\frac{e^{-E_1/kT}}{Z}\right)
  \left(\frac{Z}{e^{-E_0/kT}}\right)\\
  &= e^{-(E_1-E_0)/kT} = e^{-\hbar\omega/kT}\\
\end{align*}

\textbf{b.)} Then we assume only the first two energy states are occupied, use the
Boltzmann distribution to calculate our partition function, and use the energy formula
$E = \sum p_iE_i$ to find the energy:\\

\begin{align*}
  <E> &= p_1E_1 + p_0E_0\\
  &= \frac{E_1e^{-E_1/kT} + E_0e^{-E_0/kT}}{e^{-E_1/kT} + e^{-E_0/kT}}\\
  &= \left(\frac{\hbar\omega}{2}\right)
  \left(\frac{e^{-3\hbar\omega/2kT} + 3e^{-\hbar\omega/2kT}}{e^{-3\hbar\omega/2kT}
    + e^{-\hbar\omega/2kT}}\right)\\
\end{align*}

We then take some limits to see behavior at low and high temperature:

\begin{align*}
  \lim_{T\rightarrow\infty} <E> &= \frac{\hbar\omega}{2}\left(\frac{1+3}{1+2}\right)
  = \hbar\omega\\
  \lim_{T\rightarrow 0} <E> &= \frac{\hbar\omega}{2}
  \left(\frac{e^{-3\hbar\omega/2kT}}{e^{-3\hbar\omega/2kT}}\right)\\
  &= \frac{\hbar\omega}{2}\\
\end{align*}

These are the expected results. At high temperature, either state is equally likely
to be occupied, so the energy is the average of the two possible states. For the low
temperature state, only the ground-state is occupied, so the average energy is just
the energy of the ground state.\\

\section{Problem Two: Two-energy system}

\textbf{a.) and b.)} The energy as a function of time is given by:

\begin{align*}
  <E> &= \frac{\epsilon_1e^{-\epsilon_1/kT} + \epsilon_2e^{-\epsilon_2/kT}}
  {e^{-\epsilon_1/kT} + e^{-\epsilon_2/kT}}
\end{align*}

By this, at low temperatures the average energy will be $\epsilon_1$, since the
1-labeled exponent will go to zero slower. At high temperature, the energy will
be the average of $\epsilon_1$ and $\epsilon_2$. Plot:

\vspace{3cm}

\textbf{c.)}We now explicitly calculate these values:

\begin{align*}
  <E>
  &= \frac{\epsilon_1e^{-\epsilon_1/kT} + \epsilon_2e^{-\epsilon_2/kT}}
  {e^{-\epsilon_1/kT} + e^{-\epsilon_2/kT}}
\end{align*}

\begin{align*}
  C_V &= \frac{d<E>}{dT}_V\\
  &= \left(\frac{d}{dT}\right)_V
  \frac{\epsilon_1e^{-\epsilon_1/kT} + \epsilon_2e^{-\epsilon_2/kT}}
       {e^{-\epsilon_1/kT} + e^{-\epsilon_2/kT}}\\
  &= \frac{1}{kT^2}\frac{\epsilon_1^2e^{-\epsilon_1/kT} + \epsilon_2^2e^{-\epsilon_2/kT}}
  {e^{-\epsilon_1/kT} + e^{-\epsilon_2/kT}}
  + \frac{-1}{kT^2}\frac{\epsilon_1^2e^{-\epsilon_1/kT} + \epsilon_2^2e^{-\epsilon_2/kT}}
  {e^{-2\epsilon_1/kT} + e^{-2\epsilon_2/kT} + 2e^{-(\epsilon_1+\epsilon_2)/kT}}
  \left(\epsilon_1e^{-\epsilon_1/kT} + \epsilon_2e^{-\epsilon_2/kT}\right)\\
  &= \frac{(\epsilon_2-\epsilon_1)^2e^{-(\epsilon_1+\epsilon_2)/kT}}
       {\left(e^{-\epsilon_1/kT} + e^{-\epsilon_2/kT}\right)^2}\\
\end{align*}

\end{document}

### 2016-05-14 bioinformatics/g_sampling/gsample.py
Add gsample.py

@@ -1,0 +1,52 @@
+#! /usr/bin/python2.7
+
+from __future__ import division
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+I = []
+
+data = np.genfromtxt('seq.txt', dtype='string')
+K = data[0]
+seq = data[1:]
+n = len(seq)
+l = len(seq[0])
+pa = 0
+pt = 0
+pc = 0
+pg = 0
+A = 0
+T = 1
+G = 2
+C = 3
+
+for s in seq:
+    for i in range(len(s)):
+        pa += (s[i] == 'A') * 1 / (n*l)
+        pt += (s[i] == 'T') * 1 / (n*l)
+        pg += (s[i] == 'G') * 1 / (n*l)
+        pc += (s[i] == 'C') * 1 / (n*l)
+
+W = np.zeros([4, l])
+for s in seq:
+    for i in range(len(s)):
+        W[A][i] += (s[i] == 'A') * 1 / (n * pa) # initialize W via definition
+        W[T][i] += (s[i] == 'T') * 1 / (n * pt)
+        W[G][i] += (s[i] == 'G') * 1 / (n * pg)
+        W[C][i] += (s[i] == 'C') * 1 / (n * pc)
+
+print W
+W = np.log(W)
+print W
+
+def S(x, W):
+    score = 0
+    for i in range(len(x)):
+        score += (x[i] == 'A') * W[A][i]
+        score += (x[i] == 'T') * W[T][i]
+        score += (x[i] == 'G') * W[G][i]
+        score += (x[i] == 'C') * W[C][i]
+
+# while (S(x) - s) / S(x) > 0.1:
+#     dostuff

### 2016-05-14 stat_mech/latex/a2.tex
Add changes to gsample.py

@@
 
+\section{Problem Three: Nuclear Spin Energy}
+For this problem we only consider energies due to the spin-1 nuclei. Nonzero spins
+$\pm \hbar$ have energy $E=\epsilon$ and the zero-spin has energy $E=0$.\\
+
+\textbf{a.)}We first find the total nuclear energy of the system.
+Our first task is to find the single-particle partition function:\\
+
+\begin{align*}
+  z &= \sum e^{-\beta E_i} = e^{-\beta E_{-1}} + e^{-\beta E_0} + e^{-\beta E_1}\\
+  &= 2e^{-\beta\epsilon}+1\\
+\end{align*}
+
+We then plug this into our equation for total system energy, using N as Avogadro's
+Number:\\
+
+\begin{align*}
+  E &= N\sum p_iE_i\\
+  &= N\frac{2\epsilon e^{-\beta\epsilon}}{2e^{-\beta\epsilon} + 1}\\
+\end{align*}
+
+\textbf{b.)}Next, we find the system's entropy, using $z^N$ as the total partition
+function:\\
+
+\begin{align*}
+  S &= -\left(\frac{\partial A}{\partial T}\right)_V =
+  \frac{\partial}{\partial T} k_BT\ln(z^N)\\
+  &= k_BN\ln(Z) + \frac{Nk_BT}{Z}\frac{\partial Z}{\partial T}\\
+  &= k_BN\ln\left(2e^{-\beta\epsilon}+1\right)
+  + \frac{k_BNT}{2e^{-\beta\epsilon}+1}
+  \left(\frac{2\epsilon e^{-\beta\epsilon}}{k_BT^2}\right)\\
+  &= k_BN\ln\left(2e^{-\beta\epsilon}+1\right)
+  + \frac{2N\epsilon e^{-\beta\epsilon}}{T\left(2e^{-\beta\epsilon}+1\right)}\\
+\end{align*}
+
+We then take the limits of this function at low and high T to find extreme
+behavior:
+
+\begin{align*}
+  \lim_{T\rightarrow\infty} k_BN\ln\left(2e^{-\beta\epsilon}+1\right)
+  + \frac{2N\epsilon e^{-\beta\epsilon}}{T\left(2e^{-\beta\epsilon}+1\right)}
+  &= Nk_B\ln\left(3\right) = k_B\ln\left(3^N\right)\\
+  \lim_{T\rightarrow0} k_BN\ln\left(2e^{-\beta\epsilon}+1\right)
+  + \frac{2N\epsilon e^{-\beta\epsilon}}{T\left(2e^{-\beta\epsilon}+1\right)}
+  &= 0\\
+\end{align*}
+
+Both of these results make sense when we use the $S=k_B\ln\left(\omega\right)$
+equation for entropy. At very high temperatures there are $3^N$ possible
+equivalent states the system can take on, and at low temperatures the system will
+have only one possible state: all in the spin-0 configuration, and hence zero
+entropy.\\
+
+\textbf{c.)} At very high temperatures, the denominator of the exponential
+in $p_i = e^{-E_i/kT}/Z$ dwarfs the energy in the numerator, so all probabilities
+are the same, ie all states are equally accessible. This means the single-particle
+partition function is $3$, and hence the system's molar partition function is
+$3^N$, where N is Avogadro's number. Then the entropy is
+$S = k_B\ln\left(3^N\right)$, which is what we calculated for the limit in part b.\\
+
+For very low temperatures, the exponentials in $p_i = e^{-E_i/kT}/Z$ decay to zero,
+so the only nonzero probability is the zero-energy spin-0 state. Thus the only
+reachable state is where all particles are in their ground state, so the entropy
+is $S = k_B\ln\left(1\right) = 0$.\\
+
+\textbf{d.)} We then calculate the constant-volume heat capacity for the
+system:\\
+
+%% \begin{align*}
+%%   C_V = T\left(\frac{\partial S}{\partial T}\right)_T
+%%   &= \frac{-2N\epsilon}{T\left(2e^{-\beta\epsilon}+1\right)}
+%%   + \frac{-2N\epsilon e^{-\beta\epsilon}}{k_BT^2\left(2e^{-\beta\epsilon}+1\right)}
+%%   + \frac{-4N\epsilon^2e^{-2\beta\epsilon}}{k_BT^2\left(2e^{-\beta\epsilon}+1\right)^2}\\
+%%   &= \left(\frac{-2N\epsilon}{T\left(2e^{-\beta\epsilon}+1\right)}\right)\left(1
+%%   + \frac{e^{-\beta\epsilon}}{k_BT}
+%%   + \frac{2\epsilon e^{-2\beta\epsilon}}{k_BT\left(2e^{-\beta\epsilon}+1\right)}\right)
+%% \end{align*}
+
+\begin{align*}
+  U &= \left(\frac{\partial U}{\partial T}\right)_V\\
+  &= \left(\frac{\partial}{\partial T}\right)_V
+  N\frac{2\epsilon e^{-\beta\epsilon}}{2e^{-\beta\epsilon} + 1}\\
+  &= N\frac{2\epsilon^2 e^{-\beta\epsilon}}{k_BT^2\left(2e^{-\beta\epsilon} + 1\right)}
+  + N\frac{4\epsilon^2e^{-2\beta\epsilon}}{k_BT^2(2e^{-\beta\epsilon} + 1)^2}\\
+  &= N\frac{2\epsilon^2 e^{-\beta\epsilon}}{k_BT^2\left(2e^{-\beta\epsilon} + 1\right)}
+  \left(1 + \frac{2e^{-\beta\epsilon}}{2e^{-\beta\epsilon} + 1}\right)\\
+  &= N\frac{2\epsilon^2 e^{-\beta\epsilon}}{k_BT^2\left(2e^{-\beta\epsilon} + 1\right)^2}
+\end{align*}
+
+This is proportional to $T^2$ for very high T. For very high T, this number goes to
+zero due to the T in the denominator and lack of any growing terms anywhere. This is
+what we expect, since eventually heating up a system will not increase its energy
+anymore, since the partition function changes less and less as the system reaches a
+state where all microstates are equally probable.\\
+
 \end{document}

### 2016-05-14 bioinformatics/g_sampling/gsample.py

@@ -4,11 +4,14 @@ from __future__ import division
 
 import numpy as np
 import matplotlib.pyplot as plt
+import random
 
-I = []
+random.seed()
+
+N = 100
 
 data = np.genfromtxt('seq.txt', dtype='string')
-K = data[0]
+K = int(data[0])
 seq = data[1:]
 n = len(seq)
 l = len(seq[0])
@@ -21,6 +24,66 @@ T = 1
 G = 2
 C = 3
 
+def get_W(sequence):
+    f = np.zeros([4, K])
+    for s in sequence:
+        for i in range(K):
+            f[A][i] += (s[i] == 'A')
+            f[T][i] += (s[i] == 'T')
+            f[G][i] += (s[i] == 'G')
+            f[C][i] += (s[i] == 'C')
+    f /= n
+    for i in range(len(f[A])):
+        if f[A][i] == 0:
+            f[A][i] = -1e-100
+        else:
+            f[A][i] = np.log(f[A][i]/pa)
+    for i in range(len(f[T])):
+        if f[T][i] == 0:
+            f[T][i] = -1e-100
+        else:
+            f[T][i] = np.log(f[T][i]/pt)
+    for i in range(len(f[G])):
+        if f[G][i] == 0:
+            f[G][i] = -1e-100
+        else:
+            f[G] = np.log(f[G][i]/pg)
+    for i in range(len(f[C])):
+        if f[C][i] == 0:
+            f[C][i] = -1e-100
+        else:
+            f[C][i] = np.log(f[C][i]/pc)
+    return f
+
+def get_HM(sequence):
+    f = np.zeros([4, K])
+    H_M = np.zeros([K])
+    for s in sequence:
+        for i in range(K):
+            f[A][i] += (s[i] == 'A') # initialize W via definition
+            f[T][i] += (s[i] == 'T')
+            f[G][i] += (s[i] == 'G')
+            f[C][i] += (s[i] == 'C')
+    for i in range(K):
+        if f[A][i] != 0:
+            H_M[i] += -f[A][i]*np.log(f[A][i])
+        if f[T][i] != 0:
+            H_M[i] += -f[T][i]*np.log(f[T][i])
+        if f[G][i] != 0:
+            H_M[i] += -f[G][i]*np.log(f[G][i])
+        if f[C][i] != 0:
+            H_M[i] += -f[C][i]*np.log(f[C][i])
+    return H_M
+
+def S(s, W):
[…]

### 2016-05-15 bioinformatics/g_sampling/gsample.py
Add it all.

@@ -8,7 +8,7 @@ import random
 
 random.seed()
 
-N = 100
+N = 20
 
 data = np.genfromtxt('seq.txt', dtype='string')
 K = int(data[0])
@@ -93,36 +93,50 @@ for s in seq:
 
 H_bg = -pa*np.log(pa) - pt*np.log(pt) - pg*np.log(pg) - pc*np.log(pc)
 
-j = np.empty([n]).astype(int)
-for i in range(n):
-    j[i] = random.randint(0,l-K)
+scores = []
+motifs = []
+infos  = []
 
-info = []
-
-for x in range(N):
-    guess_motif = []
-    for i in range(n):     # create new weight matrix
-        guess_motif.append(seq[i][j[i]:j[i]+K])
-    W = get_W(guess_motif)
-    H_M = get_HM(guess_motif)
-
-    S0 = np.empty([n])     # get new scores
+for a in range(200):
+    S_best = 0
+    j = np.empty([n]).astype(int)
     for i in range(n):
-        S0[i] = S(guess_motif[i], W)
+        j[i] = random.randint(0,l-K)
 
-    for i, s in enumerate(seq):   # update guess motif by weight matrix
-        for k in range(l-K):
-            if S(s[k], W) < S0[i]:
-                S0[i] = S(s[i], W)
-                j[i] = k
+    info = []
 
-    I = H_bg - sum(H_M)/K
-    print guess_motif
-    print I
-    info.append(I)
-    #print np.sum(S0)
+    for x in range(N):
+        guess_motif = []
+        for i in range(n):     # create new weight matrix
+            guess_motif.append(seq[i][j[i]:j[i]+K])
+            W = get_W(guess_motif)
+            H_M = get_HM(guess_motif)
 
-plt.plot(info)
+        S0 = np.empty([n])     # get new scores
+        for i in range(n):
+            S0[i] = S(guess_motif[i], W)
+
+        for i, s in enumerate(seq):   # update guess motif by weight matrix
+            for k in range(l-K):
+                if S(s[k], W) < S0[i]:
+                    S0[i] = S(s[i], W)
+                    j[i] = k
+
+        I = H_bg - sum(H_M)/K
+        info.append(I)
+        if S_best < np.sum(S0):
+            S_best = np.sum(S0)
+    scores.append(S_best)
+    motifs.append(guess_motif)
+    infos.append(info)
+
+bests = [i for i, j in enumerate(scores) if j==max(scores)]
+
[…]

### 2016-05-16 stat_mech/latex/a3.tex
Add it all

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{graphicx}
\usepackage{amsmath}
%% \usepackage{marvosym} % Required for symbols in the colored box
%% \usepackage{ifsym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

\allowdisplaybreaks

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Statistical Mechanics Assignment Three}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem One: Minimizing a 4-edged box}
\textbf{a.) Calculus}

\begin{align*}
  A &= 2xy + 2yz + xz\\
  2y(x+z) + xz &= A\\
  y = \frac{\left(A - xz\right)}{2\left(x+z\right)}\\
  V = xyz = xz\frac{\left(A - xz\right)}{2\left(x+z\right)}\\
  \frac{dV}{dx} &= 0 = z\frac{\left(A - xz\right)}{2\left(x+z\right)}
  - xz\frac{z}{2\left(x+z\right)} + xz\frac{\left(A - xz\right)}{\left(x+z\right)^2}\\
  \frac{2xz^2 - Az}{2\left(x+z\right)} =& xz\frac{\left(A - xz\right)}{\left(x+z\right)^2}\\
\end{align*}

\textbf{b.)}

\begin{align*}
  g(x,y,z) &= 2xy + 2yz + xz\\
  \nabla f &= \lambda \nabla g\\
  yz &= \lambda(2y+z)\\
  xz &= \lambda(2x+2z)\\
  xy &= \lambda(x+2y)\\
\end{align*}

Then...

\begin{align*}
  xyz &= x\lambda(2y+z) = 2\lambda xy + \lambda xz\\
  xyz &= y\lambda(2x+2z)= 2\lambda xy + 2\lambda yz\\
  xyz &= z\lambda(x+2y) = \lambda xz + 2\lambda yz\\
  \lambda xz + 2\lambda yz &= 2\lambda xy + 2\lambda yz\\
  \lambda xz &= 2\lambda xy\\
  z &= y\\
  2\lambda xy + 2\lambda y^2 &= 2\lambda xy + \lambda xy\\
  2\lambda y^2 &= \lambda xy\\
  y^2 &= xy \rightarrow x = y\\
  x = y = z\\
\end{align*}

\section{Problem Two: Two-system entropy}
\textbf{a.)} The probability of two unrelated events occuring is the product of the
two individual probabilities. Because the two systems are weakly interacting,
they are ``unrelated'' and so $P_{rs} = P_rP_s$.\\

\textbf{b.)}
\begin{align*}
  S &= -k_B\sum_r \sum_s P_{rs} \ln\left(P_{rs}\right)\\
  &= -k_B\sum_r \sum_s P_rP_s \ln\left(P_rP_s\right)\\
  &= -k_B\sum_r \sum_s P_rP_s\left(\ln P_r + \ln P_s\right)\\
  &= -k_B\sum_r \sum_s P_rP_s\ln P_r + -k_B\sum_r \sum_s P_rP_s\ln P_s\\
  &= -k_B\sum_r P_r\ln P_r + -k_B\sum_s P_s\ln P_s
  \hspace{2cm}\mbox{since }\sum_rP_r = 1\\
  &= S_1 + S_2\\
\end{align*}

\section{Problem Three: Two-system probability}
\textbf{a.)}
\begin{align*}
  S - S_0 &= -k_B\sum_rP_r^0\ln P_r^0 + k_B\sum_rP_r\ln P_r\\
  &= -k_B\sum_rP_r^0\ln P_r^0 + P_r\ln P_r + P_r\ln P_r^0 - P_r\ln P_r^0
  \hspace{2cm}P_r\ln P_r^0 - P_r\ln P_r^0=0\\
\end{align*}

We need to get rid of $P_r^0\ln P_r^0 - P_r\ln P_r^0$, so we do some fancy math. By taking
the logarithm of our probability definition, we see that:

\begin{align*}
  \ln P_r^0 &= -\beta E_r - \ln Z\\
  \sum_r P_r\ln P_r^0 &= -\sum_r P_r\beta E_r - \sum_r P_r\ln Z
  \hspace{2cm}\sum_r P_r=1\\
  \sum_r P_r\ln P_r^0 &= -\beta E - \ln Z\\
  -\sum_r P_r\ln P_r^0 &= \beta E + \ln Z\\
\end{align*}

\begin{align*}
  \ln P_r^0 &= -\beta E_r - \ln Z\\
  \sum_r P_r^0\ln P_r^0 &= -\sum_r P_r^0\beta E_r - \sum_r P_r^0\ln Z
  \hspace{2cm}\sum_r P_r^0=1\\
  \sum_r P_r^0\ln P_r^0 &= -\beta E - \ln Z\\
\end{align*}

Combining these two final equations, we see that:

\begin{align*}
  \sum_r P_r^0\ln P_r^0 - \sum_r P_r\ln P_r^0 = 0\\
\end{align*}

which is what we wanted to show. We plug this into our original equation:

\begin{align*}
  S - S_0 &= -k_B\sum_rP_r\ln P_r + P_r\ln P_r^0\\
  &= -k_B\sum_r P_r \ln\left(\frac{P_r^0}{P_r}\right)\\
\end{align*}

\textbf{b.)}
\begin{align*}
  S - S_0 &= -k_B\sum_r P_r \ln\left(\frac{P_r^0}{P_r}\right)\\
  S - S_0 &\leq -k_B\sum_r P_r \frac{P_r^0}{P_r} - 1
  \hspace{2cm}\mbox{By the relation in the problem}\\
  \sum_r P_r \frac{P_r^0}{P_r} -1 = 0 \rightarrow S \leq S_0\\
\end{align*}

\end{document}

### 2016-05-16 stat_mech/latex/a3.tex
a3 statmech

@@
 \usepackage{amsmath}
-%% \usepackage{marvosym} % Required for symbols in the colored box
-%% \usepackage{ifsym} % Required for symbols in the colored box
+\usepackage{marvosym} % Required for symbols in the colored box
+\usepackage{ifsym} % Required for symbols in the colored box
 
@@
   A &= 2xy + 2yz + xz\\
-  2y(x+z) + xz &= A\\
-  y = \frac{\left(A - xz\right)}{2\left(x+z\right)}\\
-  V = xyz = xz\frac{\left(A - xz\right)}{2\left(x+z\right)}\\
-  \frac{dV}{dx} &= 0 = z\frac{\left(A - xz\right)}{2\left(x+z\right)}
-  - xz\frac{z}{2\left(x+z\right)} + xz\frac{\left(A - xz\right)}{\left(x+z\right)^2}\\
-  \frac{2xz^2 - Az}{2\left(x+z\right)} =& xz\frac{\left(A - xz\right)}{\left(x+z\right)^2}\\
-\end{align*}
+  V = xyz\\
+  \grad V &= \left(\frac{d}{dx} + \frac{d}{dy} + \frac{d}{dz} \right)xyz = 0\\
+  xy + yz + xz = 0\\
+  A_{max} &= 2xy + 2yz - xy - yz = xy + yz\\
 
-\textbf{b.)}
-
-\begin{align*}
-  g(x,y,z) &= 2xy + 2yz + xz\\
-  \nabla f &= \lambda \nabla g\\
-  yz &= \lambda(2y+z)\\
-  xz &= \lambda(2x+2z)\\
-  xy &= \lambda(x+2y)\\
-\end{align*}
-
-Then...
-
-\begin{align*}
-  xyz &= x\lambda(2y+z) = 2\lambda xy + \lambda xz\\
-  xyz &= y\lambda(2x+2z)= 2\lambda xy + 2\lambda yz\\
-  xyz &= z\lambda(x+2y) = \lambda xz + 2\lambda yz\\
-  \lambda xz + 2\lambda yz &= 2\lambda xy + 2\lambda yz\\
-  \lambda xz &= 2\lambda xy\\
-  z &= y\\
-  2\lambda xy + 2\lambda y^2 &= 2\lambda xy + \lambda xy\\
-  2\lambda y^2 &= \lambda xy\\
-  y^2 &= xy \rightarrow x = y\\
-  x = y = z\\
 \end{align*}
@@
   &= -k_B\sum_r \sum_s P_rP_s\left(\ln P_r + \ln P_s\right)\\
-  &= -k_B\sum_r \sum_s P_rP_s\ln P_r + -k_B\sum_r \sum_s P_rP_s\ln P_s\\
-  &= -k_B\sum_r P_r\ln P_r + -k_B\sum_s P_s\ln P_s
-  \hspace{2cm}\mbox{since }\sum_rP_r = 1\\
-  &= S_1 + S_2\\
-\end{align*}
-
-\section{Problem Three: Two-system probability}
-\textbf{a.)}
-\begin{align*}
-  S - S_0 &= -k_B\sum_rP_r^0\ln P_r^0 + k_B\sum_rP_r\ln P_r\\
-  &= -k_B\sum_rP_r^0\ln P_r^0 + P_r\ln P_r + P_r\ln P_r^0 - P_r\ln P_r^0
-  \hspace{2cm}P_r\ln P_r^0 - P_r\ln P_r^0=0\\
-\end{align*}
-
-We need to get rid of $P_r^0\ln P_r^0 - P_r\ln P_r^0$, so we do some fancy math. By taking
-the logarithm of our probability definition, we see that:
-
-\begin{align*}
-  \ln P_r^0 &= -\beta E_r - \ln Z\\
-  \sum_r P_r\ln P_r^0 &= -\sum_r P_r\beta E_r - \sum_r P_r\ln Z
-  \hspace{2cm}\sum_r P_r=1\\
-  \sum_r P_r\ln P_r^0 &= -\beta E - \ln Z\\
-  -\sum_r P_r\ln P_r^0 &= \beta E + \ln Z\\
-\end{align*}
-
-\begin{align*}
-  \ln P_r^0 &= -\beta E_r - \ln Z\\
-  \sum_r P_r^0\ln P_r^0 &= -\sum_r P_r^0\beta E_r - \sum_r P_r^0\ln Z
-  \hspace{2cm}\sum_r P_r^0=1\\
-  \sum_r P_r^0\ln P_r^0 &= -\beta E - \ln Z\\
-\end{align*}
-
-Combining these two final equations, we see that:
-
-\begin{align*}
-  \sum_r P_r^0\ln P_r^0 - \sum_r P_r\ln P_r^0 = 0\\
-\end{align*}
-
-which is what we wanted to show. We plug this into our original equation:
-
-\begin{align*}
-  S - S_0 &= -k_B\sum_rP_r\ln P_r + P_r\ln P_r^0\\
-  &= -k_B\sum_r P_r \ln\left(\frac{P_r^0}{P_r}\right)\\
-\end{align*}
-
-\textbf{b.)}
-\begin{align*}
-  S - S_0 &= -k_B\sum_r P_r \ln\left(\frac{P_r^0}{P_r}\right)\\
-  S - S_0 &\leq -k_B\sum_r P_r \frac{P_r^0}{P_r} - 1
-  \hspace{2cm}\mbox{By the relation in the problem}\\
-  \sum_r P_r \frac{P_r^0}{P_r} -1 = 0 \rightarrow S \leq S_0\\
 \end{align*}

### 2016-05-16 stat_mech/latex/a3.tex
Add finishing stuff to a3.

--- 2016-05-16 stat_mech/latex/a3.tex
@@
   A &= 2xy + 2yz + xz\\
-  2y(x+z) + xz &= A\\
-  y = \frac{\left(A - xz\right)}{2\left(x+z\right)}\\
-  V = xyz = xz\frac{\left(A - xz\right)}{2\left(x+z\right)}\\
-  \frac{dV}{dx} &= 0 = z\frac{\left(A - xz\right)}{2\left(x+z\right)}
-  - xz\frac{z}{2\left(x+z\right)} + xz\frac{\left(A - xz\right)}{\left(x+z\right)^2}\\
-  \frac{2xz^2 - Az}{2\left(x+z\right)} =& xz\frac{\left(A - xz\right)}{\left(x+z\right)^2}\\
+  A - 2yz &= x(2y + z)\\
+  x &= \frac{A-2yz}{2y+z}\\
+  V &= xyz\\
+  &= yz\frac{A-2yz}{2y+z}\\
+  \frac{dV}{dy} &= z\frac{A-2yz}{y+2z} - yz\frac{2z}{y+2z}
+  - 2yz\frac{A-2yz}{(y+2z)^2} = 0\\
+  (A-2yz)(y+2z) - y(2z)(y+2z) - y(A-2yz) &= 0\\
+  \frac{dV}{dz} &= (A - 2yz)(y+2z) - z(2y)(y+2z) - 2z(A-2yz) = 0\\
+\end{align*}
+
+Subtracting our z derivative from our y derivative, we get that $y = 2z$, which
+is the condition for maximizing velocity given area of this box. Thus:\\
+
+\begin{align*}
+  y &= \frac{\sqrt{A}}{\sqrt{3}}\\
+  z &= \frac{\sqrt{A}}{2\sqrt{3}}\\
+  x &= \frac{\sqrt{A}}{\sqrt{3}}\\
 \end{align*}
@@
 \begin{align*}
-  g(x,y,z) &= 2xy + 2yz + xz\\
+  g(x,y,z) &= xy + 2yz + 2xz\\
   \nabla f &= \lambda \nabla g\\
-  yz &= \lambda(2y+z)\\
-  xz &= \lambda(2x+2z)\\
-  xy &= \lambda(x+2y)\\
+  yz &= \lambda(y+2z)\\
+  xz &= \lambda(x+2z)\\
+  xy &= \lambda(2x+2y)\\
 \end{align*}
@@
 \begin{align*}
-  xyz &= x\lambda(2y+z) = 2\lambda xy + \lambda xz\\
-  xyz &= y\lambda(2x+2z)= 2\lambda xy + 2\lambda yz\\
-  xyz &= z\lambda(x+2y) = \lambda xz + 2\lambda yz\\
-  \lambda xz + 2\lambda yz &= 2\lambda xy + 2\lambda yz\\
-  \lambda xz &= 2\lambda xy\\
-  z &= y\\
-  2\lambda xy + 2\lambda y^2 &= 2\lambda xy + \lambda xy\\
-  2\lambda y^2 &= \lambda xy\\
-  y^2 &= xy \rightarrow x = y\\
-  x = y = z\\
+  xyz &= x\lambda(y+2z) = \lambda xy + 2\lambda xz\\
+  xyz &= y\lambda(x+2z)= \lambda xy + 2\lambda yz\\
+  xyz &= z\lambda(2x+2y) = 2\lambda xz + 2\lambda yz\\
+  3xyz &= \lambda(2xy + 4xz + 4yz) \rightarrow xyz = \frac{3xyz}{2\lambda}\\
 \end{align*}
+
+We then substitute these equations into our yz, xz and xy equations to get:
+
+\begin{align*}
+  1 - \frac{3x}{2\lambda}(y+2z) &= 0\\
+  1 - \frac{3y}{2\lambda}(x+2z) &= 0\\
+  1 - \frac{3z}{2\lambda}(2x+2y) &= 0\\
+\end{align*}
+
+Thus we see that:
+
+\begin{align*}
+  y &= \frac{\sqrt{A}}{\sqrt{3}}\\
+  z &= \frac{\sqrt{A}}{2\sqrt{3}}\\
+  x &= \frac{\sqrt{A}}{\sqrt{3}}\\
+\end{align*}
+
+as desired.\\
 

### 2016-05-24 stat_mech/latex/a4.tex
Add statmech 4

--- 2016-05-16 stat_mech/latex/a3.tex
@@
 \usepackage{amsmath}
-\usepackage{marvosym} % Required for symbols in the colored box
-\usepackage{ifsym} % Required for symbols in the colored box
+%% \usepackage{marvosym} % Required for symbols in the colored box
+%% \usepackage{ifsym} % Required for symbols in the colored box
 
@@
 
-\title{Statistical Mechanics Assignment Three}
+\title{Statistical Mechanics Assignment Four}
 \author{Elliott Capek}
@@
 
-\section{Problem One: Minimizing a 4-edged box}
-\textbf{a.) Calculus}
+\section{Problem Two: Two-Dimensional Gas}
+To tackle this problem, we start by finding the 1-D ideal gas partition function, use it to get the
+2-D partition function, use this new partition function to find the energy, then use the energy
+to find the heat capacity of a 2-D ideal gas.\\
+
+1-D Partition function:
+\begin{align*}
+  Z_x &= \int_{-\infty}^{\infty} e^{-mv_x^2/2kT} dv_x\\
+  &= \sqrt{\frac{2kT\pi}{m}}\\
+\end{align*}
+
+2-D Partition function:
+\begin{align*}
+  Z &= Z_xZ_y = Z_x^2 = \frac{2kT\pi}{m}\\
+\end{align*}
+
+We then find the energy of the gas:
 
 \begin{align*}
-  A &= 2xy + 2yz + xz\\
-  V = xyz\\
-  \grad V &= \left(\frac{d}{dx} + \frac{d}{dy} + \frac{d}{dz} \right)xyz = 0\\
-  xy + yz + xz = 0\\
-  A_{max} &= 2xy + 2yz - xy - yz = xy + yz\\
-
-\end{align*}
-
-\section{Problem Two: Two-system entropy}
-\textbf{a.)} The probability of two unrelated events occuring is the product of the
-two individual probabilities. Because the two systems are weakly interacting,
-they are ``unrelated'' and so $P_{rs} = P_rP_s$.\\
-
-\textbf{b.)}
-\begin{align*}
-  S &= -k_B\sum_r \sum_s P_{rs} \ln\left(P_{rs}\right)\\
-  &= -k_B\sum_r \sum_s P_rP_s \ln\left(P_rP_s\right)\\
-  &= -k_B\sum_r \sum_s P_rP_s\left(\ln P_r + \ln P_s\right)\\
+  <E> &= N\sum_{i} \epsilon_i p_i\\
+  &= \frac{N}{Z}int_{-\infty}^{\infty}int_{-\infty}^{\infty}
+  \frac{m(v_x^2+v_y^2)}{2}e^{-m(v_x^2+v_y^2)/2kT} dv_xdv_y\\
+  &= \frac{Nm}{2kT\pi}
+  \int_{0}^{\infty}\int_0^{2\pi}\frac{mv_r^2}{2}e^{-mv_r^2/2kT}}v\sin\theta dvd\theta\\
+  &= \frac{Nm}{2kT\pi}
+  \int_{0}^{\infty}\frac{mv_r^2}{2}e^{-mv_r^2/2kT}}v\sin\theta dvd\theta\\
 \end{align*}

### 2016-05-24 bioinformatics/g_sampling/gsample.py

@@ -8,7 +8,8 @@ import random
 
 random.seed()
 
-N = 20
+num_iterations = 20
+num_attempts = 1
 
 data = np.genfromtxt('seq.txt', dtype='string')
 K = int(data[0])
@@ -93,51 +94,77 @@ for s in seq:
 
 H_bg = -pa*np.log(pa) - pt*np.log(pt) - pg*np.log(pg) - pc*np.log(pc)
 
-scores = []
-motifs = []
 infos  = []
+scores = []
+datas = []
 
-for a in range(200):
-    S_best = 0
+for a in range(num_attempts):
     j = np.empty([n]).astype(int)
     for i in range(n):
         j[i] = random.randint(0,l-K)
 
-    info = []
+    iter_infos = [] # scalar of the guess motif's information per iteration
+    iter_scores = [] # n-tuple of each sequence's score per iteration
+    iter_datas = []
 
-    for x in range(N):
+    for x in range(num_iterations):
         guess_motif = []
+        guess_data = []
+        guess_score = [0] * n
+
         for i in range(n):     # create new weight matrix
             guess_motif.append(seq[i][j[i]:j[i]+K])
-            W = get_W(guess_motif)
+            W = get_W(guess_motif) # why not outside the loop?
             H_M = get_HM(guess_motif)
 
-        S0 = np.empty([n])     # get new scores
-        for i in range(n):
-            S0[i] = S(guess_motif[i], W)
+        for i in range(n):      # get new scores
+            guess_score[i] = S(guess_motif[i], W)
+            guess_data.append(str(guess_motif[i]))
+            guess_data.append(str(guess_score[i])[0:3])
+            print str(guess_score[i])[0:3]
+            print guess_data
+        guess_data.append(str(sum(guess_score))[0:3])
+
+        iter_infos.append(H_bg - sum(H_M)/K)
+        iter_scores.append(guess_score)
+        iter_datas.append(guess_data)
+
+        print "guess_score: %s" % guess_score
+        print "iter_scores: %s" % iter_scores
 
         for i, s in enumerate(seq):   # update guess motif by weight matrix
-            for k in range(l-K):
-                if S(s[k], W) < S0[i]:
-                    S0[i] = S(s[i], W)
+            for k in range(l-K): # check if this is right
+                if S(s[k], W) < guess_score[i]:
+                    guess_score[i] = S(s[i], W)
[…]

### 2016-05-24 stat_mech/latex/a4.tex
Changes to SM4

@@
   <E> &= N\sum_{i} \epsilon_i p_i\\
-  &= \frac{N}{Z}int_{-\infty}^{\infty}int_{-\infty}^{\infty}
+  &= \frac{N}{Z}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}
   \frac{m(v_x^2+v_y^2)}{2}e^{-m(v_x^2+v_y^2)/2kT} dv_xdv_y\\
   &= \frac{Nm}{2kT\pi}
-  \int_{0}^{\infty}\int_0^{2\pi}\frac{mv_r^2}{2}e^{-mv_r^2/2kT}}v\sin\theta dvd\theta\\
-  &= \frac{Nm}{2kT\pi}
-  \int_{0}^{\infty}\frac{mv_r^2}{2}e^{-mv_r^2/2kT}}v\sin\theta dvd\theta\\
+  \int_{0}^{\infty}\int_0^{2\pi}\frac{mv_r^2}{2}e^{-mv_r^2/2kT}vdv_rd\theta\\
+  &= \frac{Nm^2}{2kT} \int_{0}^{\infty} v_r^3e^{-mv_r^2/2kT}dv\\
+  &= \frac{Nm^2}{2kT}\frac{2k^2T^2}{m^2}\\
+  &= NkT\\
 \end{align*}
 
+\begin{align*}
+  C_v &= \frac{d<E>}{dT} = Nk\\
+\end{align*}
+
+This is the number we expect. Because vector decomposition works on squared velocities and the
+energy of an ideal gas only depends on the squared velocities of its particles, we can separate
+the ideal gas energy into dimensional components and just add them up to get the total energy.
+We found that each dimension has energy $\frac{1}{2}NkT$, so we get double that for our final
+energy.\\
+
 \end{document}

### 2016-05-26 stat_mech/latex/a4.tex
all not done yeat
Add sterrf yadig.
all

@@
 
+\section{Problem One: Thermodynamics problem}
+Here we are solving for the behavior of a fairly simple thermodynamic system.\\
+
+\textbf{a.) Final pressure}
+
+\begin{align*}
+  \bar{p}V' &= \nu RT\\
+  p_f &= \frac{\nu_f RT}{V_f} = \frac{2\nu RT}{V'(1+b)}\\
+  &= \frac{2\nu RT\bar{p}}{(1+b)\nu RT} = \frac{2\bar{p}}{1+b}\\
+\end{align*}
+
 \section{Problem Two: Two-Dimensional Gas}
@@
 
+\section{Problem Three: Restoring-force systems}
+
+\textbf{a.) Linear restoring force}
+\begin{align*}
+  Z = \int_0^\infty e^{-\alpha x^2/2kT}dx\\
+  &= \left(\frac{2\pi kT}{\alpha}\right)^{1/2}\\
+  <E>_r &= \frac{N}{Z}\int_0^\infty \frac{\alpha x^2}{2}
+  &= \frac{NkT}{2}\\
+  <E> &= <E>_{ideal} + <E>_r = NkT\\
+\end{align*}
+
+\textbf{b.) Cubic restoring force}
+\begin{align*}
+  Z &= \int_{-\infty}^{\infty}e^{\alpha x^4/4kT}
+  = \frac{1}{2}\left(\frac{4kT}{\alpha}\right)^{1/4}\Gamma\left(\frac{1}{4}\right)\\
+  <E> &= 2N\left(\frac{\alpha}{4kT}\right)^{1/4}\Gamma\left(\frac{1}{4}\right)^{-1}
+  \int_{-\infty}^\infty \frac{\alpha x^4}{4}e^{-\alpha x^4 / 4kT}\\
+\end{align*}
+
+\textbf{b.) Cubic restoring force}
+\begin{align*}
+  Z &= \int_{-\infty}^{\infty}e^{\alpha x^4/4kT}
+  = \frac{1}{2}\left(\frac{\alpha}{4kT}\right)^{1/4}\Gamma\left(\frac{1}{4}\right)\\
+  <E> &= N\sum E_ip_i =
+  2N\left(\frac{\alpha}{4kT}\right)^{1/4}\Gamma\left(\frac{1}{4}\right)^{-1}
+  \int_{-\infty}^{\infty}\frac{\alpha x^4}{4}e^{\alpha x^4/4kT}\\
+  &= \frac{N\alpha}{2}\left(\frac{\alpha}{4kT}\right)^{1/4}\Gamma\left(\frac{1}{4}\right)^{-1}
+  \frac{5}{16}\left(\frac{2kT}{\alpha}\right)\left(\frac{2kT}{\alpha}\right)^{1/4}
+  \Gamma\left(\frac{1}{4}\right)\\
+  &= N\alpha\left(\frac{\alpha}{4kT}\right)^{1/4}
+  \frac{5}{16}\left(\frac{4kT}{\alpha}\right)\left(\frac{4kT}{\alpha}\right)^{1/4}\\
+  &= \frac{5}{4}NkT\\
+\end{align*}
+
 \end{document}

### 2016-10-02 em/latex/HW1.tex
Add HW1.tex.

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\nonstopmode

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{amsmath}

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\usepackage{amssymb}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Quantum Mechanics Assignment One}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem One}
\textbf{Boundaries:}

\begin{align*}
  \phi(\infty) &= 0\\
  \phi(x,0) &= 0\\
  \phi(0,y) &= 0\\
\end{align*}

Guess the solution is four charges in a rectangle. If this solution works, by the Uniqueness Theorem it is the only solution.

\begin{align*}
  \phi &= \phi_1 + \phi_2 + \phi_3 + \phi_4\\
\end{align*}

The potential due to the real point charge is:

\begin{align*}
  \phi_1(x,y) &= \frac{q}{4\pi\epsilon\sqrt{(x-d_1)^2+(y-d_2)^2}}\\
\end{align*}

We let the second and fourth-quadrant points have charge -q, so potential:

\begin{align*}
  \phi_2(x,y) &= \frac{-q}{4\pi\epsilon\sqrt{(x+d_1)^2+(y-d_2)^2}}\\
  \phi_4(x,y) &= \frac{-q}{4\pi\epsilon\sqrt{(x-d_1)^2+(y+d_2)^2}}\\
\end{align*}

and the third-quadrant point have charge +q, so potential:

\begin{align*}
  \phi_3(x,y) &= \frac{q}{4\pi\epsilon\sqrt{(x+d_1)^2+(y+d_2)^2}}\\
\end{align*}

Thus the total potential is given by:

\begin{align*}
  \phi(x,y) &= \frac{q}{4\pi\epsilon}\Big(
  \left((x-d_1)^2+(y-d_2)^2\right)^{-1/2} + \left((x+d_1)^2+(y+d_2)^2\right)^{-1/2}\\
  &-\left((x+d_1)^2+(y-d_2)^2\right)^{-1/2} - \left((x-d_1)^2+(y+d_2)^2\right)^{-1/2}\Big)
\end{align*}

This satisfies our two boundary conditions:

\begin{align*}
  \phi(0,y) &= \frac{q}{4\pi\epsilon}\Big(
  \left(d_1^2+(y-d_2)^2\right)^{-1/2} + \left(d_1^2+(y+d_2)^2\right)^{-1/2}
  -\left(d_1^2+(y-d_2)^2\right)^{-1/2} - \left(d_1^2+(y+d_2)^2\right)^{-1/2}\Big)
  = 0\\
\end{align*}

\begin{align*}
  \phi(x,0) &= \frac{q}{4\pi\epsilon}\Big(
  \left((x-d_1)^2+d_2^2\right)^{-1/2} + \left((x+d_1)^2+d_2^2\right)^{-1/2}
  -\left((x+d_1)^2+d_2^2\right)^{-1/2} - \left((x-d_1)^2+d_2^2\right)^{-1/2}\Big)
  = 0\\
\end{align*}

To find the force on the first-quadrant charge, we use Coulomb's Equation:

%% \begin{align*}
%%   F &= -q\left(\frac{d}{dx} + \frac{d}{dy}\right) \frac{q}{4\pi\epsilon}\Big(
%%   \left((x-d_1)^2+(y-d_2)^2\right)^{-1/2} + \left((x+d_1)^2+(y+d_2)^2\right)^{-1/2}\\
%%   &-\left((x+d_1)^2+(y-d_2)^2\right)^{-1/2} - \left((x-d_1)^2+(y+d_2)^2\right)^{-1/2}\Big)\\
%%   &= \frac{-q^2}{4\pi\epsilon} \Big(
%%   (x-d_1 + y-d_2)\left((x-d_1)^2+(y-d_2)^2\right)^{-3/2} - (x+d_1 + y+d_2)\left((x+d_1)^2+(y+d_2)^2\right)^{-3/2}\\
%%   &+ (x+d_1 + y-d_2)\left((x+d_1)^2+(y-d_2)^2\right)^{-3/2} - (x-d_1 + y+d_2)\left((x-d_1)^2+(y+d_2)^2\right)^{-3/2}\Big)\\
%%   F(d_1,d_2) &= \frac{-q^2}{4\pi\epsilon} \Big(
%%   - (2d_1 + 2d_2)\left(4d_1^2 + 4d_2^2\right)^{-3/2} + 2d_1\left(4d_1^2\right)^{-3/2} - 2d_2\left(4d_2^2\right)^{-3/2}\Big)\\
%%   &= \frac{q^2}{4\pi\epsilon} \left(\frac{(2d_1 + 2d_2)}{\sqrt{\left(4d_1^2 + 4d_2^2\right)}^3} - \frac{1}{4d_1^2} - \frac{1}{4d_2^2}\right)\\
%% \end{align*}

\begin{align*}
  \phi(\vec{r}) &= \frac{q}{\sqrt{2}4\pi\epsilon_0\sqrt{4d_1^2 + 4d_2^2}}\left(\hat{i} + \hat{j}\right) + \frac{-q}{4\pi\epsilon_0\sqrt{4d_1^2}}\hat{i} + \frac{-q}{4\pi\epsilon_0\sqrt{4d_2^2}}\hat{j}\\
  &= \frac{q}{4\pi\epsilon_0}\left(\frac{1}{\sqrt{8d_1^2 + 8d_2^2}} - \frac{1}{\sqrt{4d_1^2}}\right)\hat{i}
  + \frac{q}{4\pi\epsilon_0}\left(\frac{1}{\sqrt{8d_1^2 + 8d_2^2}} - \frac{1}{\sqrt{4d_2^2}}\right)\hat{j}
\end{align*}

If expanded, at large r, $\phi$ should look like a quadrupole scaling with $\frac{1}{r^3}$.\\

\section{Problem Three}
The solution to the full sphere in an external field $E_0$ was shown in class to be:

\begin{align*}
  \phi(r,\theta) &= \left(-E_0r + \frac{E_0a^3}{r^2}\right)\cos\theta
\end{align*}

given boundaries $\phi(r=a) = 0$ and $\phi(r\rightarrow\infty) = 0$. We notice that our boundaries for the half sphere problem are the same, plus one addition: $\phi(r<a, z=0) = 0$. The above solution satisfies this condition as well. Then by the Uniqueness Theorem this solution must also be valid for our half sphere.\\

The first term $-E_0r\cos\theta$ is the potential from the external field. The second term $\frac{E_0a^3\cos\theta}{r^2}$ is the term due
to the sphere. This term can be rewritten as the dipole term in the multipole expansion:\\

\begin{align*}
  \frac{E_0a\cos\theta}{r^2} &= \frac{1}{4\pi\epsilon_0}\frac{1}{r^2} \hat{r} \cdot \vec{p}\\
\end{align*}

where $\vec{p} = 4\pi\epsilon_0E_0 a^3 \hat{z}$. Thus the half sphere can be exactly expressed as a dipole.\\

\end{document}

### 2016-10-03 stat_mech/latex/a4.tex
old statmech stuff.

--- 2016-05-24 stat_mech/latex/a4.tex
@@
 
+\section{Problem One: Thermodynamics problem}
+Here we are solving for the behavior of a fairly simple thermodynamic system.\\
+
+\textbf{a.) Final pressure}
+
+\begin{align*}
+  \bar{p}V' &= \nu RT\\
+  p_f &= \frac{\nu_f RT}{V_f} = \frac{2\nu RT}{V'(1+b)}\\
+  &= \frac{2\nu RT\bar{p}}{(1+b)\nu RT} = \frac{2\bar{p}}{1+b}\\
+\end{align*}
+
+\textbf{b.) Different gas entropies}
+To find the entropies if the gases are different, we use the entropy equation to find
+the final and initial entropies:
+
+\begin{align*}
+  \Delta S &= S_f - S_i\\
+  &= S_{fl} + S_{fr} - S_{il} - S_{ir}\\
+  &= -Nk\left(\ln V + \frac{3}{2}\ln T + C_1\right)
+  + -Nk\left(\ln Vb + \frac{3}{2}\ln T + C_2\right)\\
+  &+ Nk\left(\ln V(b+1) + \frac{3}{2}\ln T + C_1\right)
+  + Nk\left(\ln V(b+1) + \frac{3}{2}\ln T + C_2\right)\\
+  &= Nk\ln\frac{(b+1)^2}{b}
+\end{align*}
+
+\textbf{c.) Identical gas entropies}
+The only difference here is the gas becomes one, because the particles are identical. We
+also need to use the molar volume:
+
+\begin{align*}
+  \Delta S &= S_f - S_i\\
+  &= S_{fl} + S_{fr} - S_{il} - S_{ir}\\
+  &= -Nk\left(\ln V + \frac{3}{2}\ln T + C\right)
+  + -Nk\left(\ln Vb + \frac{3}{2}\ln T + C\right)\\
+  &+ Nk\left(2\ln V(b+1) + 2\frac{3}{2}\ln T + C\right)\\
+  &= Nk\ln\frac{(b+1)^2}{b}\\
+\end{align*}
+
 \section{Problem Two: Two-Dimensional Gas}
@@
 
+\section{Problem Three: Restoring-force systems}
+
+\textbf{a.) Linear restoring force}
+\begin{align*}
+  Z = \int_0^\infty e^{-\alpha x^2/2kT}dx\\
+  &= \left(\frac{2\pi kT}{\alpha}\right)^{1/2}\\
+  <E>_r &= \frac{N}{Z}\int_0^\infty \frac{\alpha x^2}{2}
+  = \frac{NkT}{2}\\
+  <E> &= <E>_{ideal} + <E>_r = NkT\\
+ \rightarrow C_v &= Nk
+\end{align*}
+
+\textbf{b.) Cubic restoring force}
+\begin{align*}
+  Z &= \left(\frac{2\pi kT}{m}\right)^{3/2}\int_{-\infty}^{\infty}e^{\alpha x^4/4kT}
+  = \frac{1}{2}\left(\frac{4kT}{\alpha}\right)^{1/4}\Gamma\left(\frac{1}{4}\right)\\
+  <E> &= -\frac{d\ln(Z^N)}{d\beta}\\
+  &= \left(\frac{\beta m}{2\pi}\right)^{3/2}2\left(\frac{\alpha\beta}{4}\right)^{1/4}
+  \Gamma\left(\frac{1}{4}\right)\frac{1}{2}\left(\frac{2\pi}{m}\right)^{3/2}
+  \left(\frac{4}{\alpha}\right)^{1/4}\frac{-6}{4}\beta^{-11/4}\Gamma\left(\frac{1}{4}\right)^{-1}\\
+  &= \frac{3N}{4}kT\\
+  \rightarrow C_v &= \frac{3N}{4}k
+\end{align*}
+
 \end{document}

### 2016-10-03 stat_mech/latex/a5.tex
old statmech stuff.

--- 2016-05-24 stat_mech/latex/a4.tex
@@
 
-\title{Statistical Mechanics Assignment Four}
+\title{Statistical Mechanics Assignment Five}
 \author{Elliott Capek}
@@
 
-\section{Problem Two: Two-Dimensional Gas}
-To tackle this problem, we start by finding the 1-D ideal gas partition function, use it to get the
-2-D partition function, use this new partition function to find the energy, then use the energy
-to find the heat capacity of a 2-D ideal gas.\\
+\section{Problem One: Thermodynamics problem}
+Here we are solving for the behavior of a fairly simple thermodynamic system.\\
 
-1-D Partition function:
+\textbf{a.) Isobaric Coefficient}
+
 \begin{align*}
-  Z_x &= \int_{-\infty}^{\infty} e^{-mv_x^2/2kT} dv_x\\
-  &= \sqrt{\frac{2kT\pi}{m}}\\
+  p &= \frac{RT}{V-b} - \frac{a}{V^2} \rightarrow T = \left(\frac{V-b}{R}\right)\left(p + \frac{a}{V^2}\right)\\
+  \left(\frac{\partial T}{\partial V}\right)_p &=
+  \left(\frac{1}{R}\right)\left(p + \frac{a}{V^2}\right)
+  -2\left(\frac{V-b}{R}\right)\frac{a}{V^3}\\
+  &= \frac{pV^3}{RV^3} + \frac{aV}{RV^3} + \frac{-2aV}{RV^3} + \frac{2ab}{RV^3}
+  = \frac{pV^3 - aV + 2ab}{RV^3}\\
+  \mbox{Thus...} &\frac{1}{V}\left(\frac{\partial V}{\partial T}\right)_p
+  = \frac{RV^2}{pV^3 - aV + 2ab}\\
 \end{align*}
 
-2-D Partition function:
+\textbf{b.) Heat capacity only depends on T}
+To show this, we show that if we express Cv as Cv(V,T), then its volume derivative is zero.
+
 \begin{align*}
-  Z &= Z_xZ_y = Z_x^2 = \frac{2kT\pi}{m}\\
+  \left(\partial\frac{C_V}{\partial V}\right)_T &=
+  \left(\frac{\partial}{\partial T}\right)_T T\left(\frac{\partial V}{\partial T}\right)_V\\
 \end{align*}
 
-We then find the energy of the gas:
-
-\begin{align*}
-  <E> &= N\sum_{i} \epsilon_i p_i\\
-  &= \frac{N}{Z}\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}
-  \frac{m(v_x^2+v_y^2)}{2}e^{-m(v_x^2+v_y^2)/2kT} dv_xdv_y\\
-  &= \frac{Nm}{2kT\pi}
-  \int_{0}^{\infty}\int_0^{2\pi}\frac{mv_r^2}{2}e^{-mv_r^2/2kT}vdv_rd\theta\\
-  &= \frac{Nm^2}{2kT} \int_{0}^{\infty} v_r^3e^{-mv_r^2/2kT}dv\\
-  &= \frac{Nm^2}{2kT}\frac{2k^2T^2}{m^2}\\
-  &= NkT\\
-\end{align*}
-
-\begin{align*}
-  C_v &= \frac{d<E>}{dT} = Nk\\
-\end{align*}
-
-This is the number we expect. Because vector decomposition works on squared velocities and the
-energy of an ideal gas only depends on the squared velocities of its particles, we can separate
-the ideal gas energy into dimensional components and just add them up to get the total energy.
-We found that each dimension has energy $\frac{1}{2}NkT$, so we get double that for our final
-energy.\\
-
 \end{document}

### 2016-10-03 em/latex/HW1.tex
Finish HW1? Probably needs checking.

@@
 
+\section{Problem Two}
+\textbf{Part A}\\
+We want to show the following equality is always true:
+
+\begin{align*}
+  \int\rho_1\phi_2d^3\vec{r} &= \int\rho_2\phi_1d^3\vec{r}
+\end{align*}
+
+We first apply the Laplace Equation:
+
+\begin{align*}
+  -\epsilon_0\int\nabla^2\phi_1\phi_2d^3\vec{r} &= -\epsilon_0\int\nabla^2\phi_2\phi_1d^3\vec{r}
+\end{align*}
+
+We then expand this equation into its cartesian representation:
+
+\begin{align*}
+  \int\int\int\sum_{n=x,y,z}\frac{\partial^2}{\partial n^2} (\phi_1) \phi_2 dxdydz &=
+  \int\int\int\sum_{n=x,y,z}\frac{\partial^2}{\partial n^2} (\phi_2) \phi_1 dxdydz
+\end{align*}
+
+We then use Integration by Parts:
+
+\begin{align*}
+  \int\int\int\sum_{n=x,y,z}\left(\frac{d}{dn}\phi_1\phi_2\Big|_{-\infty}^{\infty} - \int\frac{\partial}{\partial n}\phi_1\frac{\partial}{\partial n}\phi_2dn \right) dxdydz &=
+  \int\int\int\sum_{n=x,y,z}\left(\frac{d}{dn}\phi_2\phi_1\Big|_{-\infty}^{\infty} - \int\frac{\partial}{\partial n}\phi_2\frac{\partial}{\partial n}\phi_1dn \right) dxdydz
+\end{align*}
+
+We define our voltages to be zero at infinity, so the UV term of the integration by parts goes to zero, leaving us with:
+
+\begin{align*}
+  -\int\int\int\sum_{n=x,y,z}\left(\int\frac{\partial}{\partial n}\phi_1\frac{\partial}{\partial n}\phi_2dn \right) dxdydz &=
+  -\int\int\int\sum_{n=x,y,z}\left(\int\frac{\partial}{\partial n}\phi_2\frac{\partial}{\partial n}\phi_1dn \right) dxdydz
+\end{align*}
+
+We see that these expressions are identical since multiplication is commutative, thus the initial statement must be true.
+\textbf{Part B}\\
+
+We define the voltage on conductor A as: $V_A = V_{AA} + V_{BA}$ where $V_{BA}$ is the potential due to conductor B.
+We define $V_{AB}$ as the potential on B due to A.
+We let $\rho_A$ and $\rho_B$ be the charge distributions on A and B respectively, where $\int\rho_AdV = \int\rho_BdV = Q$.
+We begin with Green's Reciprocity theorem:
+
+\begin{align*}
+  \int V_{AB}(\vec{R})\rho_A(\vec{r})dV &= \int V_{BA}(\vec{r})\rho_B(\vec{r})dV\\
+\end{align*}
+
+Because A and B are conductors, we know that their potentials $V_{AB}$ and $V{BA}$ are constant. Thus we can remove them from the integral:
+
+\begin{align*}
+  V_{AB}\int\rho_A(\vec{r})dV &= V_{BA}\int\rho_B(\vec{r})dV\\
+  V_{AB}Q &= V_{BA}Q\\
+\end{align*}
+
+Thus $V_{AB} = V_{BA}$. From this we can see that $V_A = V_{AA} + V_{AB}$.\\
+
+NOTE: $V_{A} = V_{AA} + V_{AB}$ is constant, but is $V_{AB}$ constant?\\
+
 \section{Problem Three}

### 2016-10-08 em/latex/HW1.tex
Make qm/Makefile work much better, clean up .gitignore as well.

@@
 
-\title{Quantum Mechanics Assignment One}
+\title{Electromagnetism Assignment One}
 \author{Elliott Capek}
@@
 
-We first apply the Laplace Equation:
+We first briefly solve the problem in one dimension to show how it is done:
+
+\begin{align*}
+  \int\rho_1\phi_2d^3\vec{r} &= \int\rho_2\phi_1d^3\vec{r}\\
+  -\epsilon\int \frac{\partial^2\phi_1}{\partial x^2}\phi_2dx &= -\epsilon\int \frac{\partial^2\phi_2}{\partial x^2}\phi_1dx\\
+  \int \frac{\partial^2\phi_1}{\partial x^2}\phi_2dx &= \int \frac{\partial^2\phi_2}{\partial x^2}\phi_1dx\\
+  \frac{\partial \phi_1}{\partial x} \phi_2 \Big|_{-\infty}^{\infty} -
+  \int \frac{\partial \phi_2}{\partial x}\frac{\partial \phi_1}{\partial x}dx&=
+  \frac{\partial \phi_2}{\partial x} \phi_1 \Big|_{-\infty}^{\infty} -
+  \int \frac{\partial \phi_1}{\partial x}\frac{\partial \phi_2}{\partial x}dx\\
+  \int \frac{\partial \phi_2}{\partial x}\frac{\partial \phi_1}{\partial x}dx&=
+  \int \frac{\partial \phi_2}{\partial x}\frac{\partial \phi_1}{\partial x}dx
+\end{align*}
+
+Now in 3D. We first apply the Laplace Equation:
 
@@
 \begin{align*}
-  \int\int\int\sum_{n=x,y,z}\frac{\partial^2}{\partial n^2} (\phi_1) \phi_2 dxdydz &=
-  \int\int\int\sum_{n=x,y,z}\frac{\partial^2}{\partial n^2} (\phi_2) \phi_1 dxdydz
+  \int\int\int\left(\sum_{n=x,y,z}\frac{\partial^2}{\partial n^2} (\phi_1) \phi_2\right) dxdydz &=
+  \int\int\int\left(\sum_{n=x,y,z}\frac{\partial^2}{\partial n^2} (\phi_2) \phi_1\right) dxdydz
 \end{align*}
@@
 \begin{align*}
-  \int\int\int\sum_{n=x,y,z}\left(\frac{d}{dn}\phi_1\phi_2\Big|_{-\infty}^{\infty} - \int\frac{\partial}{\partial n}\phi_1\frac{\partial}{\partial n}\phi_2dn \right) dxdydz &=
-  \int\int\int\sum_{n=x,y,z}\left(\frac{d}{dn}\phi_2\phi_1\Big|_{-\infty}^{\infty} - \int\frac{\partial}{\partial n}\phi_2\frac{\partial}{\partial n}\phi_1dn \right) dxdydz
+  \int\int\sum_{n=x,y,z}\left(\frac{d}{dn}\phi_1\phi_2\Big|_{-\infty}^{\infty} - \int\frac{\partial}{\partial n}\phi_1\frac{\partial}{\partial n}\phi_2dn \right) dxdydz &=
+  \int\int\sum_{n=x,y,z}\left(\frac{d}{dn}\phi_2\phi_1\Big|_{-\infty}^{\infty} - \int\frac{\partial}{\partial n}\phi_2\frac{\partial}{\partial n}\phi_1dn \right) dxdydz
 \end{align*}
@@
 \begin{align*}
-  -\int\int\int\sum_{n=x,y,z}\left(\int\frac{\partial}{\partial n}\phi_1\frac{\partial}{\partial n}\phi_2dn \right) dxdydz &=
-  -\int\int\int\sum_{n=x,y,z}\left(\int\frac{\partial}{\partial n}\phi_2\frac{\partial}{\partial n}\phi_1dn \right) dxdydz
+  \int\int\sum_{n=x,y,z}\left(\int\frac{\partial}{\partial n}\phi_1\frac{\partial}{\partial n}\phi_2dn \right) dxdydz &=
+  \int\int\sum_{n=x,y,z}\left(\int\frac{\partial}{\partial n}\phi_2\frac{\partial}{\partial n}\phi_1dn \right) dxdydz
 \end{align*}
@@
 \begin{align*}
-  \int V_{AB}(\vec{R})\rho_A(\vec{r})dV &= \int V_{BA}(\vec{r})\rho_B(\vec{r})dV\\
+  \int V_{AB}(\vec{r})\rho_A(\vec{r})dV &= \int V_{BA}(\vec{r})\rho_B(\vec{r})dV\\
 \end{align*}
 
-Because A and B are conductors, we know that their potentials $V_{AB}$ and $V{BA}$ are constant. Thus we can remove them from the integral:
+Because A and B are conductors, we know that their potentials $V_{AB}$ and $V_{BA}$ are constant. Thus we can remove them from the integral:
 
@@
 Thus $V_{AB} = V_{BA}$. From this we can see that $V_A = V_{AA} + V_{AB}$.\\
-
-NOTE: $V_{A} = V_{AA} + V_{AB}$ is constant, but is $V_{AB}$ constant?\\
 
@@
 The first term $-E_0r\cos\theta$ is the potential from the external field. The second term $\frac{E_0a^3\cos\theta}{r^2}$ is the term due
-to the sphere. This term can be rewritten as the dipole term in the multipole expansion:\\
-
+to the sphere. This term can be rewritten as the dipole term in the multipole expansion:
 \begin{align*}
-  \frac{E_0a\cos\theta}{r^2} &= \frac{1}{4\pi\epsilon_0}\frac{1}{r^2} \hat{r} \cdot \vec{p}\\
+  \frac{E_0a\cos\theta}{r^2} &= \frac{1}{4\pi\epsilon_0}\frac{1}{r^2} \hat{r} \cdot \vec{p}
 \end{align*}

### 2016-10-09 em/latex/HW2.tex
Add all.

--- 2016-10-02 em/latex/HW1.tex
@@
 
-\title{Quantum Mechanics Assignment One}
+\title{PH651 Assignment Five}
 \author{Elliott Capek}
 \titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles
+
+\newcommand{\bra}[1]{\big<#1\big|}
+\newcommand{\ket}[1]{\big|#1\big>}
+\newcommand{\braket}[2]{\big<#1\big|#2\big>}
 
@@
 \section{Problem One}
-\textbf{Boundaries:}
+To find the bound charges on this cross section, we use the following equations:
 
 \begin{align*}
-  \phi(\infty) &= 0\\
-  \phi(x,0) &= 0\\
-  \phi(0,y) &= 0\\
+  \rho_b &= -\nabla \vec{P} = 0\hspace{2cm} \mbox{P is constant}\\
+  \sigma_b &= \vec{P} \cdot \vec{n}\\
 \end{align*}
 
-Guess the solution is four charges in a rectangle. If this solution works, by the Uniqueness Theorem it is the only solution.
+The top two faces each make a $60^\circ$ angle with $\vec{P}$, and the bottom one is antiparallel. Thus we see:
 
 \begin{align*}
-  \phi &= \phi_1 + \phi_2 + \phi_3 + \phi_4\\
+  \sigma_{b,top} &= \vec{P} \cdot \hat{n} = P\cos\left(60^\circ\right) = \frac P2\\
+  \sigma_{b,bot} &= \vec{P} \cdot \hat{n} = P\cos\left(180^\circ\right) = -P\\
 \end{align*}
 
-The potential due to the real point charge is:
-
-\begin{align*}
-  \phi_1(x,y) &= \frac{q}{4\pi\epsilon\sqrt{(x-d_1)^2+(y-d_2)^2}}\\
-\end{align*}
-
-We let the second and fourth-quadrant points have charge -q, so potential:
-
-\begin{align*}
-  \phi_2(x,y) &= \frac{-q}{4\pi\epsilon\sqrt{(x+d_1)^2+(y-d_2)^2}}\\
-  \phi_4(x,y) &= \frac{-q}{4\pi\epsilon\sqrt{(x-d_1)^2+(y+d_2)^2}}\\
-\end{align*}
-
-and the third-quadrant point have charge +q, so potential:
-
-\begin{align*}
-  \phi_3(x,y) &= \frac{q}{4\pi\epsilon\sqrt{(x+d_1)^2+(y+d_2)^2}}\\
-\end{align*}
-
-Thus the total potential is given by:
-
-\begin{align*}
-  \phi(x,y) &= \frac{q}{4\pi\epsilon}\Big(
-  \left((x-d_1)^2+(y-d_2)^2\right)^{-1/2} + \left((x+d_1)^2+(y+d_2)^2\right)^{-1/2}\\
-  &-\left((x+d_1)^2+(y-d_2)^2\right)^{-1/2} - \left((x-d_1)^2+(y+d_2)^2\right)^{-1/2}\Big)
-\end{align*}
-
-This satisfies our two boundary conditions:
-
-\begin{align*}
-  \phi(0,y) &= \frac{q}{4\pi\epsilon}\Big(
-  \left(d_1^2+(y-d_2)^2\right)^{-1/2} + \left(d_1^2+(y+d_2)^2\right)^{-1/2}
-  -\left(d_1^2+(y-d_2)^2\right)^{-1/2} - \left(d_1^2+(y+d_2)^2\right)^{-1/2}\Big)
-  = 0\\
-\end{align*}
-
-\begin{align*}
-  \phi(x,0) &= \frac{q}{4\pi\epsilon}\Big(
-  \left((x-d_1)^2+d_2^2\right)^{-1/2} + \left((x+d_1)^2+d_2^2\right)^{-1/2}
-  -\left((x+d_1)^2+d_2^2\right)^{-1/2} - \left((x-d_1)^2+d_2^2\right)^{-1/2}\Big)
-  = 0\\
-\end{align*}
-
-To find the force on the first-quadrant charge, we use Coulomb's Equation:
-
-%% \begin{align*}
-%%   F &= -q\left(\frac{d}{dx} + \frac{d}{dy}\right) \frac{q}{4\pi\epsilon}\Big(
-%%   \left((x-d_1)^2+(y-d_2)^2\right)^{-1/2} + \left((x+d_1)^2+(y+d_2)^2\right)^{-1/2}\\
-%%   &-\left((x+d_1)^2+(y-d_2)^2\right)^{-1/2} - \left((x-d_1)^2+(y+d_2)^2\right)^{-1/2}\Big)\\
-%%   &= \frac{-q^2}{4\pi\epsilon} \Big(
-%%   (x-d_1 + y-d_2)\left((x-d_1)^2+(y-d_2)^2\right)^{-3/2} - (x+d_1 + y+d_2)\left((x+d_1)^2+(y+d_2)^2\right)^{-3/2}\\
-%%   &+ (x+d_1 + y-d_2)\left((x+d_1)^2+(y-d_2)^2\right)^{-3/2} - (x-d_1 + y+d_2)\left((x-d_1)^2+(y+d_2)^2\right)^{-3/2}\Big)\\
-%%   F(d_1,d_2) &= \frac{-q^2}{4\pi\epsilon} \Big(
-%%   - (2d_1 + 2d_2)\left(4d_1^2 + 4d_2^2\right)^{-3/2} + 2d_1\left(4d_1^2\right)^{-3/2} - 2d_2\left(4d_2^2\right)^{-3/2}\Big)\\
-%%   &= \frac{q^2}{4\pi\epsilon} \left(\frac{(2d_1 + 2d_2)}{\sqrt{\left(4d_1^2 + 4d_2^2\right)}^3} - \frac{1}{4d_1^2} - \frac{1}{4d_2^2}\right)\\
-%% \end{align*}
-
-\begin{align*}
-  \phi(\vec{r}) &= \frac{q}{\sqrt{2}4\pi\epsilon_0\sqrt{4d_1^2 + 4d_2^2}}\left(\hat{i} + \hat{j}\right) + \frac{-q}{4\pi\epsilon_0\sqrt{4d_1^2}}\hat{i} + \frac{-q}{4\pi\epsilon_0\sqrt{4d_2^2}}\hat{j}\\
-  &= \frac{q}{4\pi\epsilon_0}\left(\frac{1}{\sqrt{8d_1^2 + 8d_2^2}} - \frac{1}{\sqrt{4d_1^2}}\right)\hat{i}
-  + \frac{q}{4\pi\epsilon_0}\left(\frac{1}{\sqrt{8d_1^2 + 8d_2^2}} - \frac{1}{\sqrt{4d_2^2}}\right)\hat{j}
-\end{align*}
-
-If expanded, at large r, $\phi$ should look like a quadrupole scaling with $\frac{1}{r^3}$.\\
-
-\section{Problem Three}
-The solution to the full sphere in an external field $E_0$ was shown in class to be:
-
-\begin{align*}
-  \phi(r,\theta) &= \left(-E_0r + \frac{E_0a^3}{r^2}\right)\cos\theta
-\end{align*}
-
-given boundaries $\phi(r=a) = 0$ and $\phi(r\rightarrow\infty) = 0$.
-We notice that our boundaries for the half sphere problem are the same, plus one addition: $\phi(r<a, z=0) = 0$.
-The above solution satisfies this condition as well.
-Then by the Uniqueness Theorem this solution must also be valid for our half sphere.\\
-
-The first term $-E_0r\cos\theta$ is the potential from the external field. The second term $\frac{E_0a^3\cos\theta}{r^2}$ is the term due
-to the sphere. This term can be rewritten as the dipole term in the multipole expansion:\\
-
-\begin{align*}
-  \frac{E_0a\cos\theta}{r^2} &= \frac{1}{4\pi\epsilon_0}\frac{1}{r^2} \hat{r} \cdot \vec{p}\\
-\end{align*}
-
-where $\vec{p} = 4\pi\epsilon_0E_0 a^3 \hat{z}$. Thus the half sphere can be exactly expressed as a dipole.\\
+To find total charge we just sum these densities over the surface area of the prism. We notice that there are two top faces and one
+bottom face, thus the two surface charges cancel out and the net charge on the object is zero. We expect this for a dielectric with
+no free charge.
 

### 2016-10-11 em/latex/HW2.tex
Add all.

@@
 
+\section{Problem Two}
+We let the radius of the internal dielectric be \textbf{a} and the outer dielectric have radius \textbf{b}. We then list the
+boundary conditions for inner and outer potentials $\phi_1$ and $\phi_2$ and outer potential $\phi_o$.\\
+
+\textbf{Boundaries}
+\begin{align*}
+  \phi_1(a) &= \phi_2(a)\\
+  \phi_2(b) &= \phi_o(b)\\
+  \frac{\partial \phi_1(a)}{\partial r} &= \frac{\partial \phi_2(a)}{\partial r}\\
+  \frac{\partial \phi_2(b)}{\partial r} &= \frac{\partial \phi_o(b)}{\partial r}\\
+  \phi(r>>b) &= -E_or\cos\theta
+\end{align*}
+
+Our solutions will take the general form:
+
+\begin{align*}
+  \phi &= \sum_{\ell = 0}^\infty \left(A_\ell r^\ell + \frac{B_\ell}{r^{\ell+1}}\right)P_\ell(\cos\theta)
+\end{align*}
+
+For the inner potential $\phi_1$ we don't permit any B-terms, since the inverse r would blow up at zero.
+For the middle potential $\phi_2$ we can apply no constraints.
+For the outer potential $\phi_o$ we say that all A are zero, else the potential would blow up at infinity.
+Thus:
+
+\begin{align*}
+  \phi_1 &= \sum_{\ell = 0}^\infty A^1_\ell r^\ell P_\ell(\cos\theta)\\
+  \phi_2 &= \sum_{\ell = 0}^\infty \left(A^2_\ell r^\ell + \frac{B^2_\ell}{r^{\ell+1}}\right)P_\ell(\cos\theta)\\
+  \phi_o &= -E_0r\cos\theta + \sum_{\ell = 0}^\infty \frac{B^o_\ell}{r^{\ell+1}}P_\ell(\cos\theta)
+\end{align*}
+
+We then apply our first two boundary conditions, noting that only the individual $\ell$ terms need to equal, not the whole summation:
+
+\begin{align*}
+  A^1_\ell a^\ell &= A^2_\ell a^\ell + \frac{B^2_\ell}{a^{\ell+1}}\\
+  A^2_\ell b^\ell +& \frac{B^2_\ell}{b^{\ell+1}} = \frac{B^o_\ell}{b^{\ell+1}}\\
+  A^2_1 b^1 +& \frac{B^2_1}{b^2} = -E_ob^1
+\end{align*}
+
+We then apply our derivative boundary conditions:
+
+\begin{align*}
+  \epsilon_r^1\ell A^1_\ell a^{\ell-1} =& \epsilon_r^2\ell A^2_\ell a^{\ell-1} - \epsilon_r^2\frac{\left(\ell+1\right)B^2_\ell}{a^\ell}\\
+  \epsilon_r^2\ell A^2_\ell b^{\ell-1} -& \epsilon_r^2\frac{\left(\ell+1\right)B^2_\ell}{b^\ell}
+  = -\epsilon_0\frac{\left(\ell+1\right)B^o_\ell}{b^\ell}\\
+  \epsilon_r^2A^2_1 -2\frac{B_1^2}{b} &= -E_0
+\end{align*}
+
+If we solve for $A_1$ we get:
+
 \end{document}

### 2016-10-19 em/latex/HW2.tex
Add EM3 work so far.

--- 2016-10-09 em/latex/HW2.tex
@@
 
-\title{PH651 Assignment Five}
+\title{Elegromagnetism Assignment Two}
 \author{Elliott Capek}
@@
 To find total charge we just sum these densities over the surface area of the prism. We notice that there are two top faces and one
-bottom face, thus the two surface charges cancel out and the net charge on the object is zero. We expect this for a dielectric with
-no free charge.
+bottom face, thus the two surface charges cancel out and the net charge on the object is \textbf{zero}. We expect this for a dielectric with no free charge.\\
+
+We then find the far field by treating the triangle as a point dipole extending infinitely through the z axis:
+
+\begin{align*}
+  \vec{E}(\vec{r}, \theta) &= \int_\infty^\infty \frac{\vec{P}\cdot\vec{r}}{\sqrt{z^2+r^2}}dz
+  = P\cos(\theta)\int_\infty^\infty \frac{\sqrt{z^2+r^2}}{\sqrt{z^2+r^2}^3}dz
+  = P\cos(\theta)\int_\infty^\infty \frac{1}{z^2+r^2}dz\\
+  &= \frac{P\cos(\theta)\pi}{r}
+\end{align*}
+
+This result almost makes sense. Normally a dipole falls off by $\frac{1}{r^3}$, but this one is generated by an infinite line, so it
+falls off lots slower.\\
+
+\section{Problem Two}
+The example solved in Griffiths Ex 4.7 can be generalized to a two-dielectrics-in-all-space system by simply swapping the $\epsilon_0$ for vacuum with the dielectric constant of the outer dielectric, $\epsilon_r^2$.
+Thus the solution is identical to the book's plus this swap:
+
+\textbf{a.)}\\
+\begin{align*}
+	\epsilon &= \epsilon_r^1/\epsilon_r^2\\
+	V_{in}(r,\theta) &= -\frac{3E_0r\cos\theta}{\epsilon+2}\\
+	E_{in} &= \frac{3}{\epsilon+2}E_0\hat{z}\\
+        V_{out}(r,\theta) &= -E_0r\cos\theta + \frac{\epsilon-1}{\left(\epsilon+2\right)r^2}\cos(\theta)\\
+        E_{out}(r,\theta) &= \left(E_0 + 2\frac{\epsilon-1}{\left(\epsilon+2\right)r^3\right)\cos\left(\theta\right)\hat{r}
+        + \left(-E_0 + \frac{\epsilon-1}{\left(\epsilon+2\right)r^3}\right)\cos\left(\theta\right)\hat{\theta}\\
+\end{align*}\\
+
+\textbf{b.)}\\
+For linear dielectrics we can find $\sigma_b$ from the P vector, which comes from from $\vec{E}$:
+
+\begin{align*}
+  \vec{P}_{in} &= \left(\epsilon-\epsilon_0\right)\vec{E}_{in}\\
+  &= 3E_0\frac{\epsilon-\epsilon_0}{\epsilon+2}\hat{z}\\
+  \sigma_b &= \vec{P} \cdot \hat{r} = 3E_0\frac{\epsilon-\epsilon_0}{\epsilon+2}\cos(\theta)
+\end{align*}
+
+This charge is positive at the top of the sphere and negative at the bottom, as expected for a bound charge.\\
+
+\textbf{c.)}\\
+We now show that the tangential E-fields are continuous at the surface. We note that there are no discontinuities in either dot product for any value of $\theta$:
+
+\begin{align*}
+  E_{in}(R)\cdot\hat{\theta} &= \frac{3}{\epsilon+2}E_0\cos(\theta)\\
+  E_{out}(R)\cdot\hat{\theta} &= \frac{\epsilon-1}{\epsilon+2}E_0\sin\theta
+\end{align*}
+
+\textbf{d.)}
+$D_\theta$ can be arrived at from $E_{\theta}$ by multiplying by a factor of $\epsilon_r^1$ or $\epsilon_r^2$.
+
+\begin{align*}
+  D_{in}(R) &= \frac{3\epsilon_r^1}{\epsilon+2}E_0\cos(\theta)\\
+  D_{out}(R) &= \frac{\epsilon_r^1-\epsilon_r^2}{\epsilon+2}E_0\sin\theta\\
+  D_{in}(R) - D_{out}(R) &= \frac{3\epsilon_r^1}{\epsilon+2}E_0\cos(\theta) - \frac{\epsilon_r^1-\epsilon_r^2}{\epsilon+2}E_0\sin\theta\\
+  &= \frac{E_0}{\epsilon+2}\left(3\epsilon_r^1\cos(\theta) - \left(\epsilon_r^1 - \epsilon_r^2\right)\sin(\theta)\right)
+\end{align*}
 

### 2016-10-19 em/latex/HW3.tex
Add EM3 work so far.

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\nonstopmode

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{amsmath}

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\usepackage{amssymb}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Elegromagnetism Assignment Three}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\newcommand{\bra}[1]{\big<#1\big|}
\newcommand{\ket}[1]{\big|#1\big>}
\newcommand{\braket}[2]{\big<#1\big|#2\big>}

\newcommand{\b}[1]{\boldsymbol{#1}}
\newcommand{\bv}[1]{\boldsymbol{\vec{#1}}}

\begin{document}

\maketitle{}

\section{Problem One}

\section{Problem Two}
\textbf{Part I}

We begin with equation (2):

\begin{align*}
  m &= \frac{I}{2}\oint \boldsymbol{r} \times d\boldsymbol{l}
\end{align*}

This is a loop integral around a closed circle of radius R. $d\boldsymbol{l}$ always points tangential to the circle, so the inner
cross product becomes:

\begin{align*}
  m &= \frac{I}{2}\oint R\hat{n} dl
\end{align*}

where $\hat{n}$ is perpendicular to both $\boldsymbol{r}$ and $d\boldsymbol{l}$. We are now integrating a constant around a circle of circumference $2\pi R$, which turns to:

\begin{align*}
  m &= \frac{I}{2}2\pi R^2 = I\pi R^2\hat{n} = IS\hat{n}
\end{align*}

\textbf{Part II}

We begin with the definition of torque, plug in the definition of force on a current, apply the cross product triple product rule,
argue that one of the terms goes to zero, then apply the equation from Griffiths 4th ed 1.62e.

\begin{align*}
  \boldsymbol{N} &= \oint \bv{r} \times \bv{F}dl\\
  &= \oint \bv{r} \times \left(\bv{I} \times \bv{B}\right)dl\\
  &= \oint \bv{I}\left(\bv{r} \cdot \bv{B}\right) - \bv{B}\left(\bv{r} \cdot \bv{I}\right)dl \hspace{1.2cm}\mbox{cross triple product}\\
  &= \oint \bv{I}\left(\bv{r} \cdot \bv{B}\right)dl - \oint\bv{B}\left(\bv{r} \cdot \bv{I}\right)dl\\
  &= \oint \bv{I}\left(\bv{r} \cdot \bv{B}\right)dl \hspace{3cm}\mbox{since }\Big(\oint \bv{r} \cdot d\bv{l} = 0\Big)\\
  &= \oint \bv{I}\left(\bv{B} \cdot \bv{r}\right)dl\\
  &= I\oint \left(\bv{B} \cdot \bv{r}\right)d\boldsymbol{l}\\
  &= I\bv{a} \times \bv{B} = \bv{m} \times \bv{B}\\
\end{align*}

as desired.

\section{Problem Three}
The force F between two dipoles is given by:

\begin{align*}
  \vec{F} = \frac{3\mu_0}{4\pi |r|^4}((\hat{r} \times \vec{m}_1) \times \vec{m}_2 + (\hat{r} \times \vec{m_2}) \times \vec{m_1}
  - 2\hat{r}(\vec{m}_1 \cdot \vec{m}_2) + 5\hat{r}((\hat{r} \times \vec{m}_1) \cdot (\hat{r} \times \vec{m}_2)))
\end{align*}

We use the following coordinate system:

\begin{itemize}
\item $\vec{m}_1$ points in the $\hat{z}$ direction
\item $\vec{m}_2$ points $\phi$ away from $\hat{z}$ and is \textit{coplanar} to $\hat{z}$. This is a simplifying assumption.
\item $\vec{d}$ points from the position of ball 1 to ball 2, and is angle $\theta$ from $\hat{z}$
\end{itemize}

Thus the following are true:

\begin{align*}
  \hat{r} \times \vec{m}_1 &= m_1\sin\theta\hat{x}\\
  \hat{r} \times \vec{m}_2 &= m_2\sin\phi\hat{x}\\
  \vec{m}_1 \cdot \hat{r} &= m_1\cos\theta\\
  \vec{m}_1 \cdot \vec{m}_2 &= m_1m_2\cos\phi
\end{align*}

This allows us to simplify $\vec{F}$ to:

\begin{align*}
  \vec{F} &= \frac{3\mu_0}{4\pi d^4} \left(\left(m_1\sin\theta\hat{x} \times \vec{m}_2\right)
  + \left(m_2\sin\phi\hat{x} \times \vec{m}_1\right)
  - 2m_1m_2\left(\sin\theta + \cos\phi\right)\hat{r} + 5\left(m_1\sin\theta\hat{x} \cdot m_2\sin\phi\hat{x}\left)\hat{r}
  \right)\\
  &= \frac{3\mu_0}{4\pi d^4} \left(-2m_1m_2\sin\theta\hat{r} - 2m_1m_2\cos\phi\hat{r} + 5m_1m_2\sin\theta\sin\phi\hat{r}\right)\\
  &= \frac{3\mu_1m_1m_2}{4\pi d^4}\left(-2\left(\sin\theta+\cos\phi\right) + 5\sin\theta\sin\phi\right)\hat{r}
\end{align*}

This indicates that the dipole-dipole force will be of order $\frac{3\mu_0m_1m_2}{4\pi d^4}$ and will depend on $\theta$ and $\phi$.
We see that the force is maximized when vectors $\vec{m}_1$ and $\vec{m}_2$ are perpendicular to one another.\\

Now we need to find the values of $\vec{m}_1$ and $\vec{m}_2$. We say that at far distances, the dipole moment will be dominated
by the outer-rim charge.

\begin{align*}
  \vec{m} &= I\vec{s} = (2\pi R)Q_{rim} * \pi R^2\\
  Q_{rim} &= \frac{Q}{V} * 2\pi R = \frac{3Q}{4\pi R^3} * 2\pi R = \frac{3Q}{2R^2}\\
  \vec{m} &= I\vec{s} = 3\pi^2 RQ\\
\end{align*}

Thus the total magnetic force will be something like:

\begin{align*}
  F \approx \frac{27\mu_0\pi^3R^2Q^2}{4 d^4}
\end{align*}

\end{document}

### 2016-10-19 em/latex/HW2.tex

@@
 \section{Problem One}
+\textbf{Boundaries:}
+
+\begin{align*}
+  \phi(\infty) &= 0\\
+  \phi(x,0) &= 0\\
+  \phi(0,y) &= 0\\
+\end{align*}
+
+Guess the solution is four charges in a rectangle. If this solution works, by the Uniqueness Theorem it is the only solution.
+
+\begin{align*}
+  \phi &= \phi_1 + \phi_2 + \phi_3 + \phi_4\\
+\end{align*}
+
+The potential due to the real point charge is:
+
+\begin{align*}
+  \phi_1(x,y) &= \frac{q}{4\pi\epsilon\sqrt{(x-d_1)^2+(y-d_2)^2}}\\
+\end{align*}
+
+We let the second and fourth-quadrant points have charge -q, so potential:
+
+\begin{align*}
+  \phi_2(x,y) &= \frac{-q}{4\pi\epsilon\sqrt{(x+d_1)^2+(y-d_2)^2}}\\
+  \phi_4(x,y) &= \frac{-q}{4\pi\epsilon\sqrt{(x-d_1)^2+(y+d_2)^2}}\\
+\end{align*}
+
+and the third-quadrant point have charge +q, so potential:
+
+\begin{align*}
+  \phi_3(x,y) &= \frac{q}{4\pi\epsilon\sqrt{(x+d_1)^2+(y+d_2)^2}}\\
+\end{align*}
+
+Thus the total potential is given by:
+
+\begin{align*}
+  \phi(x,y) &= \frac{q}{4\pi\epsilon}\Big(
+  \left((x-d_1)^2+(y-d_2)^2\right)^{-1/2} + \left((x+d_1)^2+(y+d_2)^2\right)^{-1/2}\\
+  &-\left((x+d_1)^2+(y-d_2)^2\right)^{-1/2} - \left((x-d_1)^2+(y+d_2)^2\right)^{-1/2}\Big)
+\end{align*}
+
+This satisfies our two boundary conditions:
+
+\begin{align*}
+  \phi(0,y) &= \frac{q}{4\pi\epsilon}\Big(
+  \left(d_1^2+(y-d_2)^2\right)^{-1/2} + \left(d_1^2+(y+d_2)^2\right)^{-1/2}
+  -\left(d_1^2+(y-d_2)^2\right)^{-1/2} - \left(d_1^2+(y+d_2)^2\right)^{-1/2}\Big)
+  = 0\\
+\end{align*}
+
+\begin{align*}
+  \phi(x,0) &= \frac{q}{4\pi\epsilon}\Big(
+  \left((x-d_1)^2+d_2^2\right)^{-1/2} + \left((x+d_1)^2+d_2^2\right)^{-1/2}
+  -\left((x+d_1)^2+d_2^2\right)^{-1/2} - \left((x-d_1)^2+d_2^2\right)^{-1/2}\Big)
+  = 0\\
+\end{align*}
+
+To find the force on the first-quadrant charge, we use Coulomb's Equation:
+
+%% \begin{align*}
+%%   F &= -q\left(\frac{d}{dx} + \frac{d}{dy}\right) \frac{q}{4\pi\epsilon}\Big(
+%%   \left((x-d_1)^2+(y-d_2)^2\right)^{-1/2} + \left((x+d_1)^2+(y+d_2)^2\right)^{-1/2}\\
+%%   &-\left((x+d_1)^2+(y-d_2)^2\right)^{-1/2} - \left((x-d_1)^2+(y+d_2)^2\right)^{-1/2}\Big)\\
+%%   &= \frac{-q^2}{4\pi\epsilon} \Big(
+%%   (x-d_1 + y-d_2)\left((x-d_1)^2+(y-d_2)^2\right)^{-3/2} - (x+d_1 + y+d_2)\left((x+d_1)^2+(y+d_2)^2\right)^{-3/2}\\
+%%   &+ (x+d_1 + y-d_2)\left((x+d_1)^2+(y-d_2)^2\right)^{-3/2} - (x-d_1 + y+d_2)\left((x-d_1)^2+(y+d_2)^2\right)^{-3/2}\Big)\\
+%%   F(d_1,d_2) &= \frac{-q^2}{4\pi\epsilon} \Big(
+%%   - (2d_1 + 2d_2)\left(4d_1^2 + 4d_2^2\right)^{-3/2} + 2d_1\left(4d_1^2\right)^{-3/2} - 2d_2\left(4d_2^2\right)^{-3/2}\Big)\\
+%%   &= \frac{q^2}{4\pi\epsilon} \left(\frac{(2d_1 + 2d_2)}{\sqrt{\left(4d_1^2 + 4d_2^2\right)}^3} - \frac{1}{4d_1^2} - \frac{1}{4d_2^2}\right)\\
+%% \end{align*}
+
+\begin{align*}
+  \phi(\vec{r}) &= \frac{q}{\sqrt{2}4\pi\epsilon_0\sqrt{4d_1^2 + 4d_2^2}}\left(\hat{i} + \hat{j}\right) + \frac{-q}{4\pi\epsilon_0\sqrt{4d_1^2}}\hat{i} + \frac{-q}{4\pi\epsilon_0\sqrt{4d_2^2}}\hat{j}\\
+  &= \frac{q}{4\pi\epsilon_0}\left(\frac{1}{\sqrt{8d_1^2 + 8d_2^2}} - \frac{1}{\sqrt{4d_1^2}}\right)\hat{i}
+  + \frac{q}{4\pi\epsilon_0}\left(\frac{1}{\sqrt{8d_1^2 + 8d_2^2}} - \frac{1}{\sqrt{4d_2^2}}\right)\hat{j}
+\end{align*}
+
+If expanded, at large r, $\phi$ should look like a quadrupole scaling with $\frac{1}{r^3}$.\\
+
+\section{Problem Two}
+We want to show the following equality is always true:
+
+\begin{align*}
+  \int\rho_1\phi_2d^3\vec{r} &= \int\rho_2\phi_1d^3\vec{r}
+\end{align*}
+
+We first apply the Laplace Equation:
+
+\begin{align*}
+  -\epsilon_0\int\nabla^2\phi_1\phi_2d^3\vec{r} &= -\epsilon_0\int\nabla^2\phi_2\phi_1d^3\vec{r}
+\end{align*}
+
+We then expand this equation into its cartesian representation:
+
+\begin{align*}
+  \int\int\int\sum_{n=x,y,z}\frac{\partial^2}{\partial n^2} (\phi_1) \phi_2 dxdydz &=
+  \int\int\int\sum_{n=x,y,z}\frac{\partial^2}{\partial n^2} (\phi_2) \phi_1 dxdydz
+\end{align*}
+
+We then use Integration by Parts:
+
+\begin{align*}
+  \int\int\int\sum_{n=x,y,z}\left(\frac{d}{dn}\phi_1\phi_2\Big|_{-\infty}^{\infty} - \int\frac{\partial}{\partial n}\phi_1\frac{\partial}{\partial n}\phi_2dn \right) dxdydz &=
+  \int\int\int\sum_{n=x,y,z}\left(\frac{d}{dn}\phi_2\phi_1\Big|_{-\infty}^{\infty} - \int\frac{\partial}{\partial n}\phi_2\frac{\partial}{\partial n}\phi_1dn \right) dxdydz
+\end{align*}
+
+\section{Problem Three}
+The solution to the full sphere in an external field $E_0$ was shown in class to be:
+
+\begin{align*}
+  \phi(r,\theta) &= \left(-E_0r + \frac{E_0a^3}{r^2}\right)\cos\theta
+\end{align*}
+
+given boundaries $\phi(r=a) = 0$ and $\phi(r\rightarrow\infty) = 0$.
+We notice that our boundaries for the half sphere problem are the same, plus one addition: $\phi(r<a, z=0) = 0$.
+The above solution satisfies this condition as well.
+Then by the Uniqueness Theorem this solution must also be valid for our half sphere.\\
+
+The first term $-E_0r\cos\theta$ is the potential from the external field. The second term $\frac{E_0a^3\cos\theta}{r^2}$ is the term due
+to the sphere. This term can be rewritten as the dipole term in the multipole expansion:\\
+
+\begin{align*}
+  \frac{E_0a\cos\theta}{r^2} &= \frac{1}{4\pi\epsilon_0}\frac{1}{r^2} \hat{r} \cdot \vec{p}\\
+\end{align*}
+
+where $\vec{p} = 4\pi\epsilon_0E_0 a^3 \hat{z}$. Thus the half sphere can be exactly expressed as a dipole.\\
+
 To find the bound charges on this cross section, we use the following equations:

### 2016-10-19 em/latex/HW2.tex
Added HW2.tex.

@@
 
-\title{Elegromagnetism Assignment Two}
+\title{Quantum Mechanics Assignment One}
 \author{Elliott Capek}
 \titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles
-
-\newcommand{\bra}[1]{\big<#1\big|}
-\newcommand{\ket}[1]{\big|#1\big>}
-\newcommand{\braket}[2]{\big<#1\big|#2\big>}
 
@@
 
-To find the bound charges on this cross section, we use the following equations:
-
-\begin{align*}
-  \rho_b &= -\nabla \vec{P} = 0\hspace{2cm} \mbox{P is constant}\\
-  \sigma_b &= \vec{P} \cdot \vec{n}\\
-\end{align*}
-
-The top two faces each make a $60^\circ$ angle with $\vec{P}$, and the bottom one is antiparallel. Thus we see:
-
-\begin{align*}
-  \sigma_{b,top} &= \vec{P} \cdot \hat{n} = P\cos\left(60^\circ\right) = \frac P2\\
-  \sigma_{b,bot} &= \vec{P} \cdot \hat{n} = P\cos\left(180^\circ\right) = -P\\
-\end{align*}
-
-To find total charge we just sum these densities over the surface area of the prism. We notice that there are two top faces and one
-bottom face, thus the two surface charges cancel out and the net charge on the object is \textbf{zero}. We expect this for a dielectric with no free charge.\\
-
-We then find the far field by treating the triangle as a point dipole extending infinitely through the z axis:
-
-\begin{align*}
-  \vec{E}(\vec{r}, \theta) &= \int_\infty^\infty \frac{\vec{P}\cdot\vec{r}}{\sqrt{z^2+r^2}}dz
-  = P\cos(\theta)\int_\infty^\infty \frac{\sqrt{z^2+r^2}}{\sqrt{z^2+r^2}^3}dz
-  = P\cos(\theta)\int_\infty^\infty \frac{1}{z^2+r^2}dz\\
-  &= \frac{P\cos(\theta)\pi}{r}
-\end{align*}
-
-This result almost makes sense. Normally a dipole falls off by $\frac{1}{r^3}$, but this one is generated by an infinite line, so it
-falls off lots slower.\\
-
-\section{Problem Two}
-The example solved in Griffiths Ex 4.7 can be generalized to a two-dielectrics-in-all-space system by simply swapping the $\epsilon_0$ for vacuum with the dielectric constant of the outer dielectric, $\epsilon_r^2$.
-Thus the solution is identical to the book's plus this swap:
-
-\textbf{a.)}\\
-\begin{align*}
-	\epsilon &= \epsilon_r^1/\epsilon_r^2\\
-	V_{in}(r,\theta) &= -\frac{3E_0r\cos\theta}{\epsilon+2}\\
-	E_{in} &= \frac{3}{\epsilon+2}E_0\hat{z}\\
-        V_{out}(r,\theta) &= -E_0r\cos\theta + \frac{\epsilon-1}{\left(\epsilon+2\right)r^2}\cos(\theta)\\
-        E_{out}(r,\theta) &= \left(E_0 + 2\frac{\epsilon-1}{\left(\epsilon+2\right)r^3\right)\cos\left(\theta\right)\hat{r}
-        + \left(-E_0 + \frac{\epsilon-1}{\left(\epsilon+2\right)r^3}\right)\cos\left(\theta\right)\hat{\theta}\\
-\end{align*}\\
-
-\textbf{b.)}\\
-For linear dielectrics we can find $\sigma_b$ from the P vector, which comes from from $\vec{E}$:
-
-\begin{align*}
-  \vec{P}_{in} &= \left(\epsilon-\epsilon_0\right)\vec{E}_{in}\\
-  &= 3E_0\frac{\epsilon-\epsilon_0}{\epsilon+2}\hat{z}\\
-  \sigma_b &= \vec{P} \cdot \hat{r} = 3E_0\frac{\epsilon-\epsilon_0}{\epsilon+2}\cos(\theta)
-\end{align*}
-
-This charge is positive at the top of the sphere and negative at the bottom, as expected for a bound charge.\\
-
-\textbf{c.)}\\
-We now show that the tangential E-fields are continuous at the surface. We note that there are no discontinuities in either dot product for any value of $\theta$:
-
-\begin{align*}
-  E_{in}(R)\cdot\hat{\theta} &= \frac{3}{\epsilon+2}E_0\cos(\theta)\\
-  E_{out}(R)\cdot\hat{\theta} &= \frac{\epsilon-1}{\epsilon+2}E_0\sin\theta
-\end{align*}
-
-\textbf{d.)}
-$D_\theta$ can be arrived at from $E_{\theta}$ by multiplying by a factor of $\epsilon_r^1$ or $\epsilon_r^2$.
-
-\begin{align*}
-  D_{in}(R) &= \frac{3\epsilon_r^1}{\epsilon+2}E_0\cos(\theta)\\
-  D_{out}(R) &= \frac{\epsilon_r^1-\epsilon_r^2}{\epsilon+2}E_0\sin\theta\\
-  D_{in}(R) - D_{out}(R) &= \frac{3\epsilon_r^1}{\epsilon+2}E_0\cos(\theta) - \frac{\epsilon_r^1-\epsilon_r^2}{\epsilon+2}E_0\sin\theta\\
-  &= \frac{E_0}{\epsilon+2}\left(3\epsilon_r^1\cos(\theta) - \left(\epsilon_r^1 - \epsilon_r^2\right)\sin(\theta)\right)
-\end{align*}
-
 \end{document}

### 2016-10-19 em/latex/HW2.tex

@@
 
+To find the bound charges on this cross section, we use the following equations:
+
+\begin{align*}
+  \rho_b &= -\nabla \vec{P} = 0\hspace{2cm} \mbox{P is constant}\\
+  \sigma_b &= \vec{P} \cdot \vec{n}\\
+\end{align*}
+
+The top two faces each make a $60^\circ$ angle with $\vec{P}$, and the bottom one is antiparallel. Thus we see:
+
+\begin{align*}
+  \sigma_{b,top} &= \vec{P} \cdot \hat{n} = P\cos\left(60^\circ\right) = \frac P2\\
+  \sigma_{b,bot} &= \vec{P} \cdot \hat{n} = P\cos\left(180^\circ\right) = -P\\
+\end{align*}
+
+To find total charge we just sum these densities over the surface area of the prism. We notice that there are two top faces and one
+bottom face, thus the two surface charges cancel out and the net charge on the object is zero. We expect this for a dielectric with
+no free charge.
+
+\section{Problem Two}
+We let the radius of the internal dielectric be \textbf{a} and the outer dielectric have radius \textbf{b}. We then list the
+boundary conditions for inner and outer potentials $\phi_1$ and $\phi_2$ and outer potential $\phi_o$.\\
+
+\textbf{Boundaries}
+\begin{align*}
+  \phi_1(a) &= \phi_2(a)\\
+  \phi_2(b) &= \phi_o(b)\\
+  \frac{\partial \phi_1(a)}{\partial r} &= \frac{\partial \phi_2(a)}{\partial r}\\
+  \frac{\partial \phi_2(b)}{\partial r} &= \frac{\partial \phi_o(b)}{\partial r}\\
+  \phi(r>>b) &= -E_or\cos\theta
+\end{align*}
+
+Our solutions will take the general form:
+
+\begin{align*}
+  \phi &= \sum_{\ell = 0}^\infty \left(A_\ell r^\ell + \frac{B_\ell}{r^{\ell+1}}\right)P_\ell(\cos\theta)
+\end{align*}
+
+For the inner potential $\phi_1$ we don't permit any B-terms, since the inverse r would blow up at zero.
+For the middle potential $\phi_2$ we can apply no constraints.
+For the outer potential $\phi_o$ we say that all A are zero, else the potential would blow up at infinity.
+Thus:
+
+\begin{align*}
+  \phi_1 &= \sum_{\ell = 0}^\infty A^1_\ell r^\ell P_\ell(\cos\theta)\\
+  \phi_2 &= \sum_{\ell = 0}^\infty \left(A^2_\ell r^\ell + \frac{B^2_\ell}{r^{\ell+1}}\right)P_\ell(\cos\theta)\\
+  \phi_o &= -E_0r\cos\theta + \sum_{\ell = 0}^\infty \frac{B^o_\ell}{r^{\ell+1}}P_\ell(\cos\theta)
+\end{align*}
+
+We then apply our first two boundary conditions, noting that only the individual $\ell$ terms need to equal, not the whole summation:
+
+\begin{align*}
+  A^1_\ell a^\ell &= A^2_\ell a^\ell + \frac{B^2_\ell}{a^{\ell+1}}\\
+  A^2_\ell b^\ell +& \frac{B^2_\ell}{b^{\ell+1}} = \frac{B^o_\ell}{b^{\ell+1}}\\
+  A^2_1 b^1 +& \frac{B^2_1}{b^2} = -E_ob^1
+\end{align*}
+
+We then apply our derivative boundary conditions:
+
+\begin{align*}
+  \epsilon_r^1\ell A^1_\ell a^{\ell-1} =& \epsilon_r^2\ell A^2_\ell a^{\ell-1} - \epsilon_r^2\frac{\left(\ell+1\right)B^2_\ell}{a^\ell}\\
+  \epsilon_r^2\ell A^2_\ell b^{\ell-1} -& \epsilon_r^2\frac{\left(\ell+1\right)B^2_\ell}{b^\ell}
+  = -\epsilon_0\frac{\left(\ell+1\right)B^o_\ell}{b^\ell}\\
+  \epsilon_r^2A^2_1 -2\frac{B_1^2}{b} &= -E_0
+\end{align*}
+
+If we solve for $A_1$ we get:
+
 \end{document}

### 2016-10-19 em/latex/HW3.tex
bla

@@
 \section{Problem One}
+Worked with Jared Cayton.\\
+Here we use two image charges, one for when z > 0 and one for when z < 0. They are both located at (0, 0, $\pm$ d).
+
+Thus:
+
+\begin{align*}
+	V &= \frac{1}{4\pi\epsilon_1}\left(\frac{q_f}{r} + \frac{q_1}{r})
+\end{align*}
+
+We use the following boundary conditions:
+
+\begin{align*}
+	V_{z>0}(z=0) = V_{z<0}(z=0)\\
+	\epsilon_1\frac{dV_{top}}{dz}(z=0) = \epsilon_2\frac{dV_{bottom}}{dz}(z=0)\\
+\end{align*}
+
+We express our top voltage in Cartesian coordinates:
+
+\begin{align*}
+  V_{top} &= \frac{1}{4\pi\epsilon_1} \left(\frac{q_f}{\sqrt{x^2+y^2+(z-d)^2}} + \frac{q_1}{\sqrt{x^2+y^2+(z-d)^2}}\right)\\
+  \frac{dV_{top}}{dz} &= \frac{1}{4\pi\epsilon_1}\left(\frac{-q_f\sqrt{x^2+y^2+(z-d)^2}(z+d)}{x^2+y^2+(z-d)^2}
+  \frac{-q_1\sqrt{x^2+y^2+(z-d)^2}(z+d)}{x^2+y^2+(z+d)^2} - \right)\\
+  \epsilon_1 \frac{dV_{top}}{dz}(z=0) &= \frac{-1}{4\pi}\left(\frac{q_fd-q_1d}{\left(x^2+y^2+d^2\right)^{\frac32}}\right)\\
+\end{align*}
+
+We do the same for bottom voltage:
+\begin{align*}
+  V_{bot} &= \frac{1}{4\pi\epsilon_2} \left(\frac{q_2}{\sqrt{x^2+y^2+(z-d)^2}}\right)\\
+  \frac{dV_{bot}}{dz} &= \frac{1}{4\pi\epsilon_2}\left(\frac{-q_2\left(z-d\right)}{\left(x^2+y^2+(z-d)^2\right)^{\frac32}}\right)\\
+  \epsilon_2 \frac{dV_{bot}}{dz}(z=0) &= \frac{1}{4\pi}\left(\frac{-q_2d}{\left(x^2+y^2+d^2\right)^{\frac32}}\right)\\
+\end{align*}
+
+We then apply our boundary conditions due to charges:
+
+\begin{align*}
+  \frac{q_f+q_1}{\epsilon_1} &= \frac{q_2}{\epsilon_2}\\
+  q_f - q_1 &= q_2\\
+\end{align*}
+
+We now have enough information to find the value of the image charges:
+
+\begin{align*}
+  q_f - q_2 &= q_1\\
+  q_f - \frac{\epsilon_2}{\epsilon_1}q_f - \frac{\epsilon_2}{\epsilon_1}q_1 = q_1\\
+  q_f\left(1-\frac{\epsilon_2}{\epsilon_1}\right) &= q_1\left(1+\frac{\epsilon_2}{\epsilon_1}\right)\\
+  q_1 &= q_f\frac{\epsilon_1-\epsilon_2}{\epsilon_1+\epsilon_2}
+\end{align*}
+
+\begin{align*}
+  q_2 &= q_f-q_1 = q_f - q_f\frac{\epsilon_1-\epsilon_2}{\epsilon_1+\epsilon_2}\\
+  &= q_f\left(1-\frac{\epsilon_1-\epsilon_2}{\epsilon_1+\epsilon_2}\right) = q_f\frac{2\epsilon_2}{\epsilon_1+\epsilon_2}\\
+\end{align*}
+
+We then express our voltage in terms of these found charges:
+
+\begin{align*}
+  V_{top} &= \frac{q_f}{4\pi\epsilon_1}\left(\frac{1}{\sqrt{x^2+y^2+(z-d)^2}} + \frac{\left(\frac{\epsilon_1-\epsilon_2}{\epsilon_1+\epsilon_2}\right)}{\sqrt{x^2+y^2+(z+d)^2}}\right)
+\end{align*}
+
+Thus
+
+\begin{align*}
+  E_{top} &= -\nabla V_{top} = \frac{-1f}{4\pi\epsilon_1}\left(\frac{x\hat{x}+y\hat{y}+(z-d)\hat{z}}{\left(x^2+y^2+(z-d)^2\right)^{\frac32}} + \left(\frac{\epsilon_1-\epsilon_2}{\epsilon_1+\epsilon_2}\right)\left(\frac{x\hat{x}+y\hat{y}+(z+d)\hat{z}}{\left(x^2+y^2+(z+d)^2\right)^{\frac32}}\right)\right)\\
+\end{align*}
+
+And similarly for $E_{bot}$:
+
+\begin{align*}
+  E_{bot} = -\nabla V_{bot} &= \frac{q_f}{4\pi\left(\epsilon_1+\epsilon_2\right)}\left(\frac{x\hat{x}+y\hat{y}+\left(z-d\right)\hat{z}}{\left(x^2+y^2+(z-d)^2\right)^{\frac32}}\right)
+\end{align*}
 
@@
 \textbf{Part I}
+\textbf{Part I}\\
 
-We begin with equation (2):
+We begin with $\bv{m} = \frac{I}{2}\oint \bv{r} \times d\bv{l}$ and want to show it is equivalent to $\bv{m} = IS\hat{n}$ when integrating over a flat surface.\\
+
+We argue that Because $\bv{r}$ and $d\bv{l}$ are coplanar, the cross product always points in the upwards direction, $\hat{n}$.
+Each segment $d\bv{l}$ represents a small part of the loop which is a distance $\bv{r}$ from the origin.
+Thus each value $\bv{r} \times d\bv{l}$ will be a small rectangle with area $rdl$, or $2\hat{n}dA$, since dA is a triangle one half the size of $\bv{r}$-$d\bv{l}$.\\
+
+Thus we see that:
 
 \begin{align*}
-  m &= \frac{I}{2}\oint \boldsymbol{r} \times d\boldsymbol{l}
-\end{align*}
-
-This is a loop integral around a closed circle of radius R. $d\boldsymbol{l}$ always points tangential to the circle, so the inner
-cross product becomes:
-
-\begin{align*}
-  m &= \frac{I}{2}\oint R\hat{n} dl
-\end{align*}
-
-where $\hat{n}$ is perpendicular to both $\boldsymbol{r}$ and $d\boldsymbol{l}$. We are now integrating a constant around a circle of circumference $2\pi R$, which turns to:
-
-\begin{align*}
-  m &= \frac{I}{2}2\pi R^2 = I\pi R^2\hat{n} = IS\hat{n}
+	\bv{m} &= \frac{I}{2}\oint \bv{r} \times d\bv{l}\\
+	&= \frac{I}{2} \oint 2\hat{n}dA = IS\hat{n}
 \end{align*}
@@
 \begin{align*}
-  \boldsymbol{N} &= \oint \bv{r} \times \bv{F}dl\\
+  \boldsymbol{N} &= \oint d\bv{r} \times \bv{F}\\
   &= \oint \bv{r} \times \left(\bv{I} \times \bv{B}\right)dl\\
@@
 \section{Problem Three}
-The force F between two dipoles is given by:
+
+We solve this problem by guessing forces.\\
+
+For the magnetic force, we know several things:
+
+\begin{itemize}
+\item Force should have two magnetic dipole terms for each sphere
+\item Wikipedia says dipole-dipole interactions depend on distance with $r^{-4}$
+\item There should be a $\mu_0$ in our magnetic force
+\end{itemize}
+
+We know that magnetic dipoles look like current-areas and should depend on $\omega, Q, a$. We guess:
 
 \begin{align*}
-  \vec{F} = \frac{3\mu_0}{4\pi |r|^4}((\hat{r} \times \vec{m}_1) \times \vec{m}_2 + (\hat{r} \times \vec{m_2}) \times \vec{m_1}
-  - 2\hat{r}(\vec{m}_1 \cdot \vec{m}_2) + 5\hat{r}((\hat{r} \times \vec{m}_1) \cdot (\hat{r} \times \vec{m}_2)))
+	I \approx Q\omega\pi a^2
 \end{align*}
 
-We use the following coordinate system:
-
-\begin{itemize}
-\item $\vec{m}_1$ points in the $\hat{z}$ direction
-\item $\vec{m}_2$ points $\phi$ away from $\hat{z}$ and is \textit{coplanar} to $\hat{z}$. This is a simplifying assumption.
-\item $\vec{d}$ points from the position of ball 1 to ball 2, and is angle $\theta$ from $\hat{z}$
-\end{itemize}
-
-Thus the following are true:
+We know $\mu_0$ has units of force per current squared. We guess a solution which involves a magnetic dipole squared, $d^4$ and $\mu_0$ in a way which leads to force units:
 
 \begin{align*}
-  \hat{r} \times \vec{m}_1 &= m_1\sin\theta\hat{x}\\
-  \hat{r} \times \vec{m}_2 &= m_2\sin\phi\hat{x}\\
-  \vec{m}_1 \cdot \hat{r} &= m_1\cos\theta\\
-  \vec{m}_1 \cdot \vec{m}_2 &= m_1m_2\cos\phi
+	F_B \approx \frac{Q^2\omega^2\pi^2a^4\mu_0}{d^4}
 \end{align*}
 
-This allows us to simplify $\vec{F}$ to:
+We then use Coulomb's force equation to guess out electric force:
 
 \begin{align*}
-  \vec{F} &= \frac{3\mu_0}{4\pi d^4} \left(\left(m_1\sin\theta\hat{x} \times \vec{m}_2\right)
-  + \left(m_2\sin\phi\hat{x} \times \vec{m}_1\right)
-  - 2m_1m_2\left(\sin\theta + \cos\phi\right)\hat{r} + 5\left(m_1\sin\theta\hat{x} \cdot m_2\sin\phi\hat{x}\left)\hat{r}
-  \right)\\
-  &= \frac{3\mu_0}{4\pi d^4} \left(-2m_1m_2\sin\theta\hat{r} - 2m_1m_2\cos\phi\hat{r} + 5m_1m_2\sin\theta\sin\phi\hat{r}\right)\\
-  &= \frac{3\mu_1m_1m_2}{4\pi d^4}\left(-2\left(\sin\theta+\cos\phi\right) + 5\sin\theta\sin\phi\right)\hat{r}
+	F_E \approx \frac{Q^2}{4\pi\epsilon_0 d^2}
 \end{align*}
 
-This indicates that the dipole-dipole force will be of order $\frac{3\mu_0m_1m_2}{4\pi d^4}$ and will depend on $\theta$ and $\phi$.
-We see that the force is maximized when vectors $\vec{m}_1$ and $\vec{m}_2$ are perpendicular to one another.\\
-
-Now we need to find the values of $\vec{m}_1$ and $\vec{m}_2$. We say that at far distances, the dipole moment will be dominated
-by the outer-rim charge.
-
+Thus
 \begin{align*}
-  \vec{m} &= I\vec{s} = (2\pi R)Q_{rim} * \pi R^2\\
-  Q_{rim} &= \frac{Q}{V} * 2\pi R = \frac{3Q}{4\pi R^3} * 2\pi R = \frac{3Q}{2R^2}\\
-  \vec{m} &= I\vec{s} = 3\pi^2 RQ\\
-\end{align*}
-
-Thus the total magnetic force will be something like:
-
-\begin{align*}
-  F \approx \frac{27\mu_0\pi^3R^2Q^2}{4 d^4}
+	\frac{F_B}{F_E} \approx \frac{4\pi^3\omega^2a^4\mu_0\epsilon_0}{d^2}
 \end{align*}

### 2016-11-30 thesis/latex/thesis.tex
Add actual thesis skeleton.

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{titlesec}
\setcounter{secnumdepth}{4}
\titleformat{\paragraph}{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
\titlespacing*{\paragraph}
{0px}{3.25ex plus 1ex minux .2ex}{1.5ex plus .2ex}

\usepackage{fancyhdr}
\usepackage{amssymb}
\usepackage{amsmath}
\pagestyle{fancy}
\fancyhf{}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\begin{document}

\newpage
\tableofcontents

\newpage

Table of Contents

\section{Chapter: Introduction}
	\subsection{What is dynein?}
		\subsubsection{How does it behave?}
		\subsubsection{What role does it play in the cell?}
		\subsubsection{Why does the cell need motor transport?}
		\subsubsection{Why is dynein special and interesting?}
	\subsection{How does dynein walk?}
		\subsubsection{Experimental results on dynein’s step}
		\subsubsection{Experimental features of dynein (ATPase, multi-state…)}
		\subsubsection{Powerstroke Theory}
		    \paragraph{a The theory}
		    \paragraph{b Evidence of theory}
		\subsubsection{c What is still needed to support theory}
		\subsubsection{Other theories?}
	\subsection{Past Simulations}
		\subsubsection{Why simulations are necessary}
		\subsubsection{Sarlah Winch Model}
		\subsubsection{-x Other models}
	\subsection{Brownian motion}
                \subsubsection{Introduction/derivation}
                \subsubsection{Justification}
\section{Chapter: Methods}
	\subsection{Our model}
		\subsubsection{Rigid point model}
		\subsubsection{Harmonic energies}
		\subsubsection{Two-state model}
		\subsubsection{D model}
	\subsection{Simulating our model}
		\subsubsection{Motion equations}
			\paragraph{a Onebound solution}
			\paragraph{b Bothbound solution}
		\subsubsection{Code}
			\paragraph{a Running, compiling, language choice?}
		\subsubsection{Time evolution}
			\paragraph{Euler’s method}
			\paragraph{Vs other methods (Runge Kutta, etc)}
			\paragraph{Finding the right dt}
			\paragraph{State transitions}
		        \paragraph{Gibbs energy of transition}
		\subsection{Parameters}
			\subsubsection{a-n Lengths, angles, binding constants, etc…}
	\subsection{Tuning our model}
		\subsubsection{Searching for optimal parameters}
\section{Chapter: Results}
	\subsection{Histograms}
		\subsubsection{Step length histograms}
		\subsubsection{Step time histograms}
	\subsection{Other ways to represent data?}

\section{Chapter: Discussion}
	\subsection{Why our model worked, didn’t work}
	\subsection{Future work on this project}
	\subsection{Possible further projects}

\end{document}

### 2016-12-09 extra/bb497-dynein-info-content/get-info-content.py
Add relative entropy code.

@@ -1,0 +1,65 @@
+#! /usr/bin/python2.7
+
+import matplotlib.pyplot as plt
+import numpy as np
+import sys
+
+def get_entropy(column, text):
+    aminoacids = "GPAVLIMCFYWHKRQNEDST"
+    entropy = 0
+    for a in aminoacids:
+        background_p = float(text.count(a)) / len(text)
+        freq = float(column.count(a)) / len(column)
+        if background_p != 0 and freq != 0:
+            entropy += freq * np.log2(freq / background_p)
+            #print freq, background_p, freq/background_p, entropy
+    return entropy
+
+if len(sys.argv) != 2:
+    print "usage: ./script input_file"
+    exit(1)
+
+raw_file = open(sys.argv[1], 'r').read()
+chunks = raw_file.split('\n\n')
+
+numchunks = len(chunks)
+numseqs   = len(chunks[0].split('\n'))
+
+seqs = [""]*numseqs
+
+for c in range(numchunks):
+    chunk = chunks[c].split('\n')
+    for n in range(numseqs):
+        seqs[n] += chunk[n][-60:]
+
+human_position = 7
+        
+seqlen = len(seqs[0])
+human_seqlen = seqlen - seqs[human_position].count('-')
+
+entropy = np.zeros(human_seqlen)
+human_idx = 0
+
+background_text = "".join(seqs)
+
+for n in range(seqlen):
+    if (seqs[human_position][n] != '-'):
+        column = "".join([seq[n] for seq in seqs])
+        entropy[human_idx] = get_entropy(column, background_text)
+        human_idx += 1
+    else:
+        continue
+
+avenum = 300
+ave_entropy = np.empty(avenum)
+ave_width = np.floor(human_seqlen/avenum)
+
+for a in range(avenum):
+    print np.mean(entropy[a*ave_width:(a+1)*ave_width])
+    ave_entropy[a] = np.mean(entropy[a*ave_width:(a+1)*ave_width])
+
+plt.title("Relative entropy of DYNCH2")
+plt.xlabel("base")
+plt.ylabel("relative entropy (bits)")
+plt.plot(np.linspace(0,human_seqlen,avenum), ave_entropy)
+plt.show()

### 2016-12-09 thesis/latex/thesis.tex

@@
 \setcounter{secnumdepth}{4}
-\titleformat{\paragraph}{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
+
+\titleformat{\paragraph}
+{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
 \titlespacing*{\paragraph}
-{0px}{3.25ex plus 1ex minux .2ex}{1.5ex plus .2ex}
+{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
 
@@
 
+\tableofcontents
 \newpage
-\tableofcontents
-
-\newpage
-
-Table of Contents
 
@@
 \subsection{What is dynein?}
+\subsubsection{Basics}
+This section will describe everything people need to know about dynein to understand the project:
+what proteins are, what dynein does, the things it uses to do it, and why it is necessary.\\
+
+Possibly begin with a figure showing <x> vs T for a diffusing particle and a motor protein (2-3 weeks
+for a protein to diffuse 1 cm vs. 2-3 hours for a motor protein to move 1cm).\\
+
+Show a crystal structure as well as artist's rendition.\\
+
 \subsubsection{How does it behave?}
+Talk about its stepping behavior using Weihong's figures (as well as the new paper David found?)\\
+
+Also possibly talk about the other things Dynein has been found to do, such as hydrolyse ATP and
+move its subdomains around. This may be better in a different section than the introduction though,
+since its a lot of biochemistry for the second section in an intro? maybe not, it has to go somewhere.
+
 \subsubsection{What role does it play in the cell?}
+This'll require more reading...dynein transports cargo, such as proteins, vesicles, and other cellular structures.
+How does binding to structures work, is it random Brownian?
+Do dyneins bind to the MT then cargo, or vice versa?
+What does dynein actually bind to, since it has a single binding domain that
+can't recognize everything. Does it have different binding partners which bind to specific things?\\
+
+Other roles than transport: moving around microtubules to help with chromosome segregation, what else?
+
 \subsubsection{Why does the cell need motor transport?}
+This ties in a lot to the previous section. I originally thought of putting the <x> vs T diffusion/
+motor plot in the introduction, but maybe it would be better here. Not sure.\\
+
 \subsubsection{Why is dynein special and interesting?}
+Compare the processive motion of dynein to kinesin in \textbf{a.)} directionality, \textbf{b.)}
+stepping pattern and \textbf{c.)} size (like how it can step over roadblocks kinesin cannot).\\
+
+Talk about why a random stepping motor is interesting compared to a orderly-stepping motor: what
+is different about dynein's structure that makes it step differently? Compare the crystal structures
+and ask lots of questions. Build interest in the motors here.\\
+
 \subsection{How does dynein walk?}
+Show the \textit{Cianfrocco} annual review figure on the mechanochemical cycle of dynein. Talk about
+how each step works in some detail, according to the theory. Find out what other papers say about
+this cycle, and whether there is disagreement about how the cycle works.
+
 \subsubsection{Experimental results on dynein’s step}
+Weihong's figures go here.\\
 \subsubsection{Experimental features of dynein (ATPase, multi-state…)}
+Some history of studies which have determined what dynein has in it.\\
+
 \subsubsection{Powerstroke Theory}
-		    \paragraph{a The theory}
+Talk about the history of this idea, why it is supported and why it is not supported.\\
+\paragraph{The theory}
+Describe the theory.
 \paragraph{b Evidence of theory}
+Evidence supporting the powerstroke.
+
 \subsubsection{c What is still needed to support theory}
+Where our simulation fits in. Why we need this kind of information.
 \subsubsection{Other theories?}
@@
 \section{Chapter: Methods}
+\subsection{Simulation}
+Talk about why simulation is the only way to verify the mechanochemical cycle, since experiments
+aren't high resolution enough to make good videos of the walk, and even if they were, in some sense
+knowing that dynein behaves identically to a simple mathematical model is more informative than
+a video, since there is lots of information in a video and really precise info in a simulation.\\
+
+\textbf{Not sure where this goes:} The 'goal' of our experiment, at a very high level, is to
+understand how dynein works. We can know this if we can come up with the minimum set of characteristics a physical system needs to act like dynein. This information is the same thing as knowing
+``how dynein walks.'' We don't need to know all the complex dynamics of every single amino acid in
+dynein, we only need to know our rough model.\\
+
 \subsection{Our model}
+Talk about what the important parts of the mechanochemical cycle from the introduction are, and how
+we take those and model them.\\
 \subsubsection{Rigid point model}
+Talk about two things: reducing bulky circular domains to points (is this the same as unified atom
+model?) with a certain drag constant based on radius, and acting like the domains are rigid in
+distance from each other. Talk about why this is valid (is it? it's an assumption, David may know
+more).
 \subsubsection{Harmonic energies}
+How we're modeling the complicated folding and adjusting of a massive protein as a harmonic
+oscillator. Why this is valid. Also other approaches...what else could we have modelled it as?
 \subsubsection{Two-state model}
-		\subsubsection{D model}
+The mechanochemical cycle has eight states. We choose two as representative because \textbf{a.)}
+many of the states are identically conformationally or \textbf{b.)} the transitions happen so fast
+the intermediate states don't matter. Talk about rate constants here.
+\subsubsection{Limitations}
+Talk about the things we \textit{don't} take into account...motor-motor interaction, 2d information,
+other ATPase nucleotide state, saying certain mechanochemical steps occur ``fast,'' not taking into
+account all mechanochemical cycle states, making the coiled coil domain rigid even though it may be
+springy...
+
 \subsection{Simulating our model}
+Talk generally about what simulating stuff means.\\
 \subsubsection{Motion equations}
+
 \paragraph{a Onebound solution}
+Discuss how our model is represented by the following figure (maybe get one with springs on the
+hinges):\\
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.45\textwidth]{/home/elliotc12/dynein_walk/figures/equilibrium-onebound}
+  \caption{Definition of equilibrium angles.}
+  \label{fig:eq_angles}
+\end{figure}
+
+Show the derivation process.\\
+
 \paragraph{b Bothbound solution}
+Show how bothbound is derived and different from onebound. Put the actual equations in the appendix,
+or just leave them out 'cause they're ugly?\\
+
 \subsubsection{Code}
+Talk about using C++, Make, C, Python. Talk about the rough program flow, how we go from code to
+data.\\
+
 \paragraph{a Running, compiling, language choice?}
+Details on why we chose C++.\\
 \subsubsection{Time evolution}
+How we do time evolution.
 \paragraph{Euler’s method}
+Show how Euler's method is arrived at, how its error scales, why we use it.
+
 \paragraph{Vs other methods (Runge Kutta, etc)}
+Talk about why our differential equation is most conducive to Euler's method. Euler's is more
+efficient -- is there a way to show how our differential equation is smooth, not jagged?
+
 \paragraph{Finding the right dt}
+Figure on error rate vs. dt.
+
 \paragraph{State transitions}
+Talk about how we calculate transition rates based on differences in energies between states.
+Then how we use transition rates to transition our model accurately.\\
+
 \paragraph{Gibbs energy of transition}
+Do some stuff with Gibb's here? Why our energies are Gibb's and not internal energy?\\
+
+\subsection{Verifying our model}
+Talk about why we think our model accurately represents the system we are trying to model, and
+doesn't have bugs in it.\\
+
+Talk about this figure:
+
+\begin{figure}[h!]
+  \centering
+
+  \begin{minipage}[b]{0.49\textwidth}
+    \includegraphics[width=\textwidth]{/home/elliotc12/dynein_walk/figures/OB_Average_PE.pdf}
+    \caption{Onebound PE vs time.}
+  \end{minipage}
+  \begin{minipage}[b]{0.49\textwidth}
+    \includegraphics[width=\textwidth]{/home/elliotc12/dynein_walk/figures/BB_Average_PE.pdf}
+    \caption{Bothbound PE vs time.}
+  \end{minipage}
+  \label{fig:equipartition_agreement}
+\end{figure}
+
+and how it shows our model is physical, or at least, how obeying the equipartition theorem shows
+how our domains both feel harmonic energies and can fully explore their spaces?\\
+
 \subsection{Parameters}
-			\subsubsection{a-n Lengths, angles, binding constants, etc…}
+The following table shows our guess parameters from various papers/crystal structures.\\
+
+                        \begin{center}
+  \begin{tabular}{| l | l | l | l | l | l |}
+    \hline
+    & Winch (Sarlah) & Schmidt & Lin & PyMol 3VKH & PyMol 4RH7\\\hline
+    $L_s$ & 12nm & & & 21.0nm & 22.1nm\\ \hline
+    $L_t$ &  7nm & & & & 11.15nm\\ \hline
+    $R_b$ &  & & & 1.57nm & 1.45nm\\ \hline
+    $R_m$ &  7nm & & & 7.36nm & 6.3nm\\ \hline
+    $R_t$ &  & & & & 2.16nm \\ \hline
+    $\theta_{m}^{Pre}$ & 250$^{\circ}$ & & 171$^{\circ}$ & &\\ \hline
+    $\theta_{m}^{Post}$ & 330$^{\circ}$ & & 137.5$^{\circ}$ & &\\ \hline
+    $\theta_{b}$ & 56$^{\circ}$ & & 63.5$^{\circ}$ & &\\ \hline
+    $\theta_{t}$ & 0$^{\circ}$ & & & &\\ \hline
+    $k_{ub}$ & $180 s^{-1,a}$ & & & &\\ \hline
+    $k_b$ & $5000 s^{-1,b}$ & & & &\\ \hline
+  \end{tabular}
+\end{center}
+		\subsubsection{Lengths, angles, binding constants, etc…}
 	\subsection{Tuning our model}
@@
 \subsection{Why our model worked, didn’t work}
+Fill this out over winter when you know more.\\
 \subsection{Future work on this project}
+If setbacks happen, this'll be a big section. Or, further simulations which could be done.
+
+\subsubsection{Experiments to verify our model}
+Talk about how FRET could be used to verify our model. We have domain distance info over long
+timescales in our model - this is something we could compare to a FRET experiment exploring
+the distance between domains in dynein.\\
+
+Other biophysics experiments which could verify our model?\\
+
+Calirimetry could get Gibb's energy of binding if you froze the motor in onebound using
+vanadium? Maybe...\\
 \subsection{Possible further projects}
+\subsection{Difference between dynein and kinesin}
+Our project un/successfully predicted how dynein behaved. How would we modify it to describe kinesin?
+Is dynein's linker-motor connection less tense than kinesin's, leading to more random motion? This
+would likely be lots of speculation, but it could be cool to look up some papers on kinesin motility
+and get its crystal structure to do some comparisons.
 

### 2016-12-18 extra/bb497-dynein-info-content/get-info-content.py
Change to get-info-content.py

@@ -50,16 +50,63 @@ for n in range(seqlen):
     else:
         continue
 
-avenum = 300
-ave_entropy = np.empty(avenum)
-ave_width = np.floor(human_seqlen/avenum)
+# avenum = 300
+# ave_entropy = np.empty(avenum)
+# ave_width = np.floor(human_seqlen/avenum)
 
-for a in range(avenum):
-    print np.mean(entropy[a*ave_width:(a+1)*ave_width])
-    ave_entropy[a] = np.mean(entropy[a*ave_width:(a+1)*ave_width])
+# for a in range(avenum):
+#     print np.mean(entropy[a*ave_width:(a+1)*ave_width])
+#     ave_entropy[a] = np.mean(entropy[a*ave_width:(a+1)*ave_width])
 
-plt.title("Relative entropy of DYNCH2")
+plt.title("Relative entropy of AAA1")
 plt.xlabel("base")
 plt.ylabel("relative entropy (bits)")
-plt.plot(np.linspace(0,human_seqlen,avenum), ave_entropy)
+plt.plot(np.linspace(1651,1875,1875-1651), entropy[1651:1875])
 plt.show()
+
+stem_entropy = np.mean(entropy[1:1650])
+AAA1_entropy = np.mean(entropy[1651:1875])
+AAA2_entropy = np.mean(entropy[1938:2161])
+AAA3_entropy = np.mean(entropy[2251:2505])
+AAA4_entropy = np.mean(entropy[2617:2863])
+stalk_entropy = np.mean(entropy[2881:3169])
+AAA5_entropy = np.mean(entropy[3244:3473])
+AAA6_entropy = np.mean(entropy[3690:3905])
+
+CC1_entropy = np.mean(entropy[1074:1103])
+CC2_entropy = np.mean(entropy[2897:2982])
+CC3_entropy = np.mean(entropy[3109:3200])
+CC4_entropy = np.mean(entropy[3408:3442])
+
+print "stem_entropy: ", stem_entropy 
+print "AAA1_entropy: ", AAA1_entropy  
+print "AAA2_entropy: ", AAA2_entropy  
+print "AAA3_entropy: ", AAA3_entropy  
+print "AAA4_entropy: ", AAA4_entropy  
+print "stlk_entropy: ", stalk_entropy
+print "AAA5_entropy: ", AAA5_entropy  
+print "AAA6_entropy: ", AAA6_entropy
+
+print "CC1_entropy: ", CC1_entropy
[…]

### 2016-12-18 thesis/latex/thesis.tex
Add bibtex stuff.

@@
 
+\bibliographystyle{plain}
+
 \begin{document}
@@
 \subsubsection{Why does the cell need motor transport?}
+
+\cite{dewitt2012cytoplasmic}
+
 This ties in a lot to the previous section. I originally thought of putting the <x> vs T diffusion/
@@
   \centering
-  \includegraphics[width=.45\textwidth]{/home/elliotc12/dynein_walk/figures/equilibrium-onebound}
+  \includegraphics[width=.45\textwidth]{../figures/equilibrium-onebound}
   \caption{Definition of equilibrium angles.}
@@
   \begin{minipage}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{/home/elliotc12/dynein_walk/figures/OB_Average_PE.pdf}
+    \includegraphics[width=\textwidth]{../figures/OB_Average_PE.pdf}
     \caption{Onebound PE vs time.}
@@
   \begin{minipage}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{/home/elliotc12/dynein_walk/figures/BB_Average_PE.pdf}
+    \includegraphics[width=\textwidth]{../figures/BB_Average_PE.pdf}
     \caption{Bothbound PE vs time.}
@@
 
-Calirimetry could get Gibb's energy of binding if you froze the motor in onebound using
+Calorimetry could get Gibb's energy of binding if you froze the motor in onebound using
 vanadium? Maybe...\\
@@
 
+\bibliography{thesis}
+
 \end{document}

### 2016-12-22 thesis/latex/thesis.tex
Add work on Introduction.

@@
 \subsubsection{Basics}
-This section will describe everything people need to know about dynein to understand the project:
-what proteins are, what dynein does, the things it uses to do it, and why it is necessary.\\
 
-Possibly begin with a figure showing <x> vs T for a diffusing particle and a motor protein (2-3 weeks
-for a protein to diffuse 1 cm vs. 2-3 hours for a motor protein to move 1cm).\\
+Dynein is a cellular structure which converts chemical energy to mechanical energy.
+It does so by reacting with adenosine triphosphate (ATP) to take steps along long cellular tracks known as microtubules.
+Dynein's mechanical energy is used to transport molecules around the cell, pull chromosomes during division, move cellular propellers known as cilia
+and various other functions \cite{cianfroccoreview}.\\
 
-Show a crystal structure as well as artist's rendition.\\
+A depictions of dynein is shown in Figure (\ref{dynein-artist-rendition}).\\
 
-\subsubsection{How does it behave?}
-Talk about its stepping behavior using Weihong's figures (as well as the new paper David found?)\\
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/dynein-artist-rendition-2.jpg}
+  \caption{Artist rendition of the dynein motor. Modified. Source: Lander Lab, The Scripps Research Institute.}
+  \label{dynein-artist-rendition}
+\end{figure}
 
-Also possibly talk about the other things Dynein has been found to do, such as hydrolyse ATP and
-move its subdomains around. This may be better in a different section than the introduction though,
-since its a lot of biochemistry for the second section in an intro? maybe not, it has to go somewhere.
+Dynein is believed to move by using the energy of ATP to cycle through a ring of states, each with different spatial relations to the microtubule.
+By cycling forward through this ring, the molecule can end up in the same state it began, but moved forward a small amount.
+This ring of states is known as the mechanochemical cycle \cite{cianfroccoreview}.\\
+
+The goal of this project is to create a mathematical model of the mechanochemical cycle and verify that it can reproduce experimental dynein stepping data.\\
+
+\subsection{Biology review}
+
+\textit{This section will describe everything people need to know about dynein to understand the project: what proteins are, what dynein does, the things it uses to do it, and why it is necessary.
+Show a crystal structure as well as artist's rendition.}
+
+\subsubsection{Why do cells need motor transport?}
+Cells are structures with high organization in time and space. 
+Cells are organized, heretogenous structures which respond quickly to their environments.
+This means that cells require a mechanism for rapidly moving cargo, and moving cargo to precise locations within the cell.
+This can be a challenge, since cells are fairly large compared to the proteins which compose them \textbf{is it also a challenge because its hard to get precise positions within a cell?}.
+A human fibroblast cell has a volume of roughly 2000 $\mu m^3$\cite{fibroblastvolume}, corresponding to roughly $8 \mu m$ in diameter.
+In comparison, human hemoglobin has a diameter of roughly $5 nm$ (PDB id 5ME2) \textbf{((is this the best way to cite this?))}, a $10^3$ factor difference in size.\\
+
+Diffusion, or random motion of molecules due to collisions with solvent (i.e.
+Brownian motion), is one possible means cells could use to transport cargo.
+Diffusion has two problems: it is slow and nondirected.
+The expected distance covered for one-dimensional Brownian motion after time t was found by Einstein to be \textbf{cite here}:
+
+\begin{equation}
+  \label{diffusion-equation}
+  <x> = \sqrt{2Dt}
+\end{equation}
+
+where D is the diffusion constant (flesh this out). The $\sqrt{t}$ scaling of diffusion contrasts with the linear scaling of motor transport.
+
+This ties in a lot to the previous section. I originally thought of putting the <x> vs T diffusion/
+motor plot in the introduction, but maybe it would be better here. Not sure.\\
 
@@
 
-\subsubsection{Why does the cell need motor transport?}
+\subsubsection{How does it behave?}
+Talk about its stepping behavior using Weihong's figures (as well as the new paper David found?)\\
 
-\cite{dewitt2012cytoplasmic}
-
-This ties in a lot to the previous section. I originally thought of putting the <x> vs T diffusion/
-motor plot in the introduction, but maybe it would be better here. Not sure.\\
+Also possibly talk about the other things Dynein has been found to do, such as hydrolyse ATP and
+move its subdomains around. This may be better in a different section than the introduction though,
+since its a lot of biochemistry for the second section in an intro? maybe not, it has to go somewhere.
 
@@
   \centering
-
   \begin{minipage}[b]{0.49\textwidth}

### 2016-12-22 thesis/latex/thesis.tex
Add ATP energy of hydrolysis and new thesis-webpages bib file.

@@
 Dynein is a cellular structure which converts chemical energy to mechanical energy.
-It does so by reacting with adenosine triphosphate (ATP) to take steps along long cellular tracks known as microtubules.
-Dynein's mechanical energy is used to transport molecules around the cell, pull chromosomes during division, move cellular propellers known as cilia
-and various other functions \cite{cianfroccoreview}.\\
+It does so by reacting with adenosine triphosphate (ATP) to take steps along long cellular tracks known as microtubules (MTs).
+Dynein's mechanical energy is used to transport molecules around the cell, pull chromosomes during division, move cellular propellers known as cilia and various other functions \cite{cianfroccoreview}.\\
 
-A depictions of dynein is shown in Figure (\ref{dynein-artist-rendition}).\\
+Dynein is a protein with many smaller domains.
+A depiction of the protein and its labeled subdomains is shown in Figure (\ref{dynein-artist-rendition}).
+Each motor is a homodimer made up of two identical monomer proteins.
+Each monomer has a binding domain which binds to the microtubule, a motor domain which binds ATP, a tail domain which connects to the other monomer, a stalk which connects the binding and motor domains, and a linker which connects the motor and tail domains.
+The tail domain is also responsible for binding to cargo.\\
 
@@
   \centering
-  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/dynein-artist-rendition-2.jpg}
-  \caption{Artist rendition of the dynein motor. Modified. Source: Lander Lab, The Scripps Research Institute.}
+  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/dynein-artist-rendition.jpg}
+  \caption{Artist rendition of the dynein motor bound to a microtubule. One monomer is bound to the MT, and one is raised up. Modified. Source: Lander Lab, The Scripps Research Institute.}
   \label{dynein-artist-rendition}
@@
 
-\textit{This section will describe everything people need to know about dynein to understand the project: what proteins are, what dynein does, the things it uses to do it, and why it is necessary.
-Show a crystal structure as well as artist's rendition.}
+\textbf{ATP} Adenosine triphosphate is a small organic molecule which very spontaneously breaks apart into ADP and a phosphate.
+This reaction is used to power many energy-requiring mechanisms within the cell.
+This reaction has a $\Delta G^\circ \approx -36 kJ/mol$ at physiological conditions \ref{ATPenergy}.\\\\
+\textbf{Protein} Dynein is a protein, or long chain of amino acids linked together and folded into a particular shape.
 
@@
 
-\bibliography{thesis}
+\bibliography{thesis-articles}
+
+\bibliography{thesis-webpages}
 

### 2016-12-22 thesis/latex/thesis.tex
Add diffusion figure script and Why do cells need motor transport? section.

@@
 This reaction has a $\Delta G^\circ \approx -36 kJ/mol$ at physiological conditions \ref{ATPenergy}.\\\\
-\textbf{Protein} Dynein is a protein, or long chain of amino acids linked together and folded into a particular shape.
+\textbf{Protein} Dynein is a protein, or long chain of amino acids linked together and folded into a particular shape.\\
 
 \subsubsection{Why do cells need motor transport?}
-Cells are structures with high organization in time and space. 
 Cells are organized, heretogenous structures which respond quickly to their environments.
-This means that cells require a mechanism for rapidly moving cargo, and moving cargo to precise locations within the cell.
-This can be a challenge, since cells are fairly large compared to the proteins which compose them \textbf{is it also a challenge because its hard to get precise positions within a cell?}.
+This means that cells require a mechanism for rapidly moving components to precise locations within the cell.
+This can be a challenge, since cells are fairly large compared to the proteins which compose them.
 A human fibroblast cell has a volume of roughly 2000 $\mu m^3$\cite{fibroblastvolume}, corresponding to roughly $8 \mu m$ in diameter.
@@
 Diffusion, or random motion of molecules due to collisions with solvent (i.e.
-Brownian motion), is one possible means cells could use to transport cargo.
+Brownian motion), is one possible means cells could use to transport biomolecules.
 Diffusion has two problems: it is slow and nondirected.
@@
 
-where D is the diffusion constant (flesh this out). The $\sqrt{t}$ scaling of diffusion contrasts with the linear scaling of motor transport.
+where D is the diffusion constant describing the diffusability of the biomolecule.
+For a medium-sized (140kD, 1D = weight of H atom) protein with a diffusion constant $D = .2 \mu m^2/s$ \cite{diffusionconstants}, it would take about a month to travel across a millimeter-sized cell.
+In contrast, it would take dynein, which travels at roughly 100nm/s \cite{weihongpaper}, a much more reasonable three hours to do the same.
 
-This ties in a lot to the previous section. I originally thought of putting the <x> vs T diffusion/
-motor plot in the introduction, but maybe it would be better here. Not sure.\\
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/diffusion_vs_dynein.png}
+  \caption{Plot comparing the diffusion rate of a 140kD protein with a diffusion consant $D = .2 \mu m^2/s$ with a dynein motor travelling at 100nm/s.}
+  \label{diffusion_vs_dynein}
+\end{figure}
+
+Another advantage of motor transport is that microtubules can be highly specific in their location in the cell, as in chromosomal segregation, where dynein walks on MTs bound directly to chromosomes.
+Diffusion is an inherently random process, and so could never provide such spatial specificity.\\
 

### 2016-12-23 thesis/latex/thesis.tex
Add small changes to thesis.tex.

@@
 By cycling forward through this ring, the molecule can end up in the same state it began, but moved forward a small amount.
-This ring of states is known as the mechanochemical cycle \cite{cianfroccoreview}.\\
-
-The goal of this project is to create a mathematical model of the mechanochemical cycle and verify that it can reproduce experimental dynein stepping data.\\
+This ring of states is known as the mechanochemical cycle \cite{cianfroccoreview}.
+The goal of this project is to create a mathematical model of the mechanochemical cycle and verify that it can reproduce experimental dynein stepping data.
+There are several types of dynein which all behave differently.
+This project will focus on cytoplasmic dynein-1, hereby referred to as dynein.\\
 
@@
 Diffusion, or random motion of molecules due to collisions with solvent (i.e.
-Brownian motion), is one possible means cells could use to transport biomolecules.
+Brownian motion), is one possible process cells could use to transport biomolecules.
 Diffusion has two problems: it is slow and nondirected.
@@
 
-Another advantage of motor transport is that microtubules can be highly specific in their location in the cell, as in chromosomal segregation, where dynein walks on MTs bound directly to chromosomes.
-Diffusion is an inherently random process, and so could never provide such spatial specificity.\\
+\textit{Maybe use the more reasonable 2$\mu m / s$ number for dynein velocity, in-vitro speed may not be applicable since [ATP]-limited}
+
+Another advantage of motor transport is that microtubules can be highly specific in their location in the cell.
+During cell division chromosomes line up at the center of each cell in a plane.
+Such a geometrically precise shape requires some sort of directed motion which a random process could never provide.\\
+
+\textit{maybe decorate the above plot with pictures of cells of each size, or how long certain cellular events take? That could be cool and make it actually an interesting plot}
+\textit{this argument would be much better if you could find examples of things which need to be done quickly in the cell, like responding to foreign particles or signalling
+  molecules or cell division or neural events}
 
 \subsubsection{What role does it play in the cell?}
-This'll require more reading...dynein transports cargo, such as proteins, vesicles, and other cellular structures.
-How does binding to structures work, is it random Brownian?
-Do dyneins bind to the MT then cargo, or vice versa?
-What does dynein actually bind to, since it has a single binding domain that
-can't recognize everything. Does it have different binding partners which bind to specific things?\\
+Some of dynein's primary roles in the cell are to transport cellular cargos such as organelles, vesicles, mRNA, cytoskeletal filaments and certain proteins.
+Dynein also plays a role in positioning and breaking down the nucleus, cell death, spindle formation and the placement of other important cellular structures \cite{valetoolbox}.
 
-Other roles than transport: moving around microtubules to help with chromosome segregation, what else?
+Microtubules are long polymers of alternating $\alpha-$ and $\beta-$ tubulin subunits.
+MTs are polar, meaning they have a distinct directionality based on the orientation of thetubulin proteins which comprise them.
+One pole, the minus end, is where they initiate formation, typically at a MTOC (microtubule organizing center) around the center of the cell.
+The other pole, the plus end, is where they grow and shrink from.
+Kinesin and myosin, two other families of motor protein, typically walk towards the plus end of the MT.
+Dynein is unique in that it walks towards the minus end, making it a minus end-directed motor.
+This lends it to a particularly important function inside neurons known as retrograde transport.\\
+
+Neurons have cell bodies at their centers and axons which grow outwards.
+Axons are long, narrow structures extending up to a meter in length in humans.
+Cell bodies contain nuclei and important organelles for synthesizing proteins, but growth of axon tips is vital for neuron function.
+This means bidirectional motor transport between axon tips and cell bodies is very important in neurons, since diffusion would not be quick enough.
+Retrograde transport is transport from the axon tip to the cell body.
+Cargo includes vesicles full of proteins ready to be broken down in the cell body and microtubule fragments to be returned from the axon tip \cite{neuroanatomy}.
+Because microtubule minus ends are found in the cell body, dynein's minus end-directed nature makes it the only protein capable of retrograde transport.
+An interesting question is, if proteins like dynein are synthesized in the cell body but needed at the axon tip to perform retrograde transport, how do dynein get to the tip?
+The answer is shown in \ref{retrograde-transport}: kinesin, a plus end-directed motor, takes dynein to the tip from the cell body.\\
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/retrograde_transport.jpg}
+  \caption{Mechanism of dynein localization to axon tips, or growth cones, in retrograde transport mediated by kinesin anterograde transport. Modified from \cite{valetoolbox}.}
+  \label{retrograde-transport}
+\end{figure}
 
@@
 and ask lots of questions. Build interest in the motors here.\\
+
+%% \begin{figure}[h]
+%%   \centering
+%%   \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/diffusion_vs_dynein.png}
+%%   \caption{Crystal structures of cytoplasmic dynein 1 and kinesin KIF5B (PDB IDs 4RH7 and 1BG2).}
+%%   \label{kin_n_dyn}
+%% \end{figure}
+
+\textit{make this figure good -- find a full length kinesin crystal structure, not 1BG2. Use kindyn.pse as a template}
 
@@
 
+\section{Questions for Tate/Nicole}
+\begin{itemize}
+\item How responsible am I for the papers I cite? Should I have read them, know their methodologies?
+\item Is it appropriate to cite web pages in a thesis? In a separate citation section?
+\item When is a fact ``common knowledge'' enough that it doesn't need a citation?
+\end{itemize}
 \end{document}

### 2016-12-23 thesis/latex/thesis.tex
Add MTBD-complex figure and beginnings of dynein structure section.

@@
 
-\subsection{Biology review}
-
+\subsubsection{Glossary}
 \textbf{ATP} Adenosine triphosphate is a small organic molecule which very spontaneously breaks apart into ADP and a phosphate.
@@
 
+\subsubsection{Dynein structure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/dynein-artist-rendition.jpg}
+  \caption{Artist rendition of the dynein motor bound to a microtubule. One monomer is bound to the MT, and one is raised up. Modified. Source: Lander Lab, The Scripps Research Institute.}
+  \label{dynein-artist-rendition-2}
+\end{figure}
+
+\textit{eventually find a different figure}
+
+As shown in \ref{dynein-artist-rendition-2}, a dynein motor is a fusion of to identical subunits, or monomers, each with binding, stalk, motor and tail subdomains.
+Each of these subunits has a unique purpose for the motor.
+
+\textbf{Binding domain}
+Microtubule binding domains (MTBDs) allow dynein to bind to the microtubule.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/mtbd-complex.png}
+  \caption{Crystal structure of the MTBD-MT complex. Modified from Redwine \textit{et. al} \cite{redwineMTBDcomplex}.}
+  \label{dynein-artist-rendition-2}
+\end{figure}
+
 \subsubsection{How does it behave?}
@@
 
-\textit{make this figure good -- find a full length kinesin crystal structure, not 1BG2. Use kindyn.pse as a template}
+\textit{make this figure good -- find a full length kinesin crystal structure, not 1BG2. Use kindyn.pse as a template.}
 
@@
 \item When is a fact ``common knowledge'' enough that it doesn't need a citation?
+\item Some papers have 'Request permission to reuse' links - do I need to do this for copying figures?
 \end{itemize}

### 2016-12-31 thesis/latex/thesis.tex
Add work on dynein's role in the cell.
Add first bit of mechanochemical cycle words.

@@
 \subsection{How does dynein walk?}
-Show the \textit{Cianfrocco} annual review figure on the mechanochemical cycle of dynein. Talk about
-how each step works in some detail, according to the theory. Find out what other papers say about
-this cycle, and whether there is disagreement about how the cycle works.
+
+Cianfrocco \textit{et.
+al} \cite{cianfroccoreview} propose dynein achieves motion via transitioning through the cycle of states shown in Fig.
+\ref{mech-cycle}.
+This model will hereon be referred to as the Cianfrocco model.
+The model claims the primary events in the dynein step are as follows.
+First, starting with both binding domains attached to the MT, ATP binds to the motor domain, which causes it to change conformation.
+This change causes the binding domain to release the microtubule, meaning only one binding domain is MT-bound.
+The free binding domain then undergoes another conformation change known as the ``priming stroke,'' which changes the orientation of the binding domain with respect to the MT.
+Finally the binding domain rebinds to the MT, causing a final conformation change in the motor which returns it to the beginning of the cycle.
+The exact order of these events is not precisely known.
+More detail on each step of the cycle is given below.\\
+
+\textbf{Things which are known about dynein which support the Cianfrocco model}
+\begin{itemize}
+\item AAA1 only ATPase necessary for motility \cite{cianfroccoreview}
+\item AAA5 mediates stalk register changes Schmidt H, Zalyte R, Urnavicius L, Carter AP. 2014. Structure of human cytoplasmic dynein-2 primed for
+  its power stroke. Nature 518:435–38
+\item Linker moves as response to AAA1 nucleotide binding Burgess SA, Walker ML, Sakakibara H, Knight PJ, Oiwa K. 2003. Dynein structure and power stroke. Nature
+421:715–8
+\end{itemize}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/mechanochemical-cycle.png}
+  \caption{Cycle of states the dynein motor goes through during motion in the Cianfrocco model \cite{cianfroccoreview}.}
+  \label{mech-cycle}
+\end{figure}
+
+\subsubsection{ATP binding}
 
@@
 \subsubsection{Other theories?}
+Imamula model, how it is different from Cianfrocco model, and why we chose teh Cianfrocco model.\\
 	\subsection{Past Simulations}
@@
                 \subsubsection{Justification}
+
 \section{Chapter: Methods}

### 2017-01-08 thesis/latex/thesis.tex
Add more work to introduction.

@@
 First, starting with both binding domains attached to the MT, ATP binds to the motor domain, which causes it to change conformation.
-This change causes the binding domain to release the microtubule, meaning only one binding domain is MT-bound.
+ATP hydrolysis to ADP-Pi causes further conformation changes.
+These changes cause the binding domain to release the microtubule, meaning only one binding domain is MT-bound.
 The free binding domain then undergoes another conformation change known as the ``priming stroke,'' which changes the orientation of the binding domain with respect to the MT.
@@
 \begin{itemize}
-\item AAA1 only ATPase necessary for motility \cite{cianfroccoreview}
+\item AAA1 only ATPase necessary for motility, which is why we don't include the other ATPases in our model \cite{cianfroccoreview}
 \item AAA5 mediates stalk register changes Schmidt H, Zalyte R, Urnavicius L, Carter AP. 2014. Structure of human cytoplasmic dynein-2 primed for
@@
   421:715–8
+\item Crystal structures with Vanadium-Adp show dynein in the pre-powerstroke state, hinting that ADP-Pi-bound dynein takes on a different conformation than ADP-Pi free dynein
 \end{itemize}
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/mechanochemical-cycle.png}
-  \caption{Cycle of states the dynein motor goes through during motion in the Cianfrocco model \cite{cianfroccoreview}.}
-  \label{mech-cycle}
-\end{figure}
+Make a figure of pre/post-powerstroke dynein with the linker hilighted? Can you get what residues correspond to the linker somewhere?
 
-\subsubsection{ATP binding}
+%% \begin{figure}[h]
+%%   \centering
+%%   \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/mechanochemical-cycle.png}
+%%   \caption{Cycle of states the dynein motor goes through during motion in the Cianfrocco model \cite{cianfroccoreview}.}
+%%   \label{mech-cycle}
+%% \end{figure}
+
+\subsubsection{Step One: ATP binding}
+\subsubsection{Step Two: }
 
@@
 \subsubsection{Other theories?}
-Imamula model, how it is different from Cianfrocco model, and why we chose teh Cianfrocco model.\\
+Imamula model: how it is different from Cianfrocco model, and why we chose the Cianfrocco model.\\
 	\subsection{Past Simulations}
-		\subsubsection{Why simulations are necessary}
 		\subsubsection{Sarlah Winch Model}
@@
 springy...
+
+Mention the Burgess paper, which says spring constants might change between apo/ADP states.
 
@@
 
+Do our spring constants seem to indicate that the protein does about one ATP's worth of work through its cycle? I wonder if the motor cools down its surroundings when it walks...\\
+
 Calorimetry could get Gibb's energy of binding if you froze the motor in onebound using
@@
 \item When is a fact ``common knowledge'' enough that it doesn't need a citation?
-\item Some papers have 'Request permission to reuse' links - do I need to do this for copying figures?
+\item Some papers have 'Request permission to reuse' links for figures - do I need to do this for copying figures?
 \end{itemize}

### 2017-01-08 thesis/latex/thesis.tex

@@
 where D is the diffusion constant describing the diffusability of the biomolecule.
-For a medium-sized (140kD, 1D = weight of H atom) protein with a diffusion constant $D = .2 \mu m^2/s$ \cite{diffusionconstants}, it would take about a month to travel across a millimeter-sized cell.
+For a medium-sized (140kD, 1D = weight of H atom) protein with a diffusion constant $D = .2 \mu m^2/s$ %% \cite{diffusionconstants}
+, it would take about a month to travel across a millimeter-sized cell.
 In contrast, it would take dynein, which travels at roughly 100nm/s \cite{weihongpaper}, a much more reasonable three hours to do the same.
@@
 
+Maybe show the Cianfrocco model in a section \textit{after} talking about all the things which happen for sure in dynein, since it's more theoretical?
+Talk about the Cianfrocco model, but other interpretations?
+
 \textbf{Things which are known about dynein which support the Cianfrocco model}
@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/mechanochemical-cycle.jpeg}
+  \caption{Cycle of states the dynein motor goes through during motion in the Cianfrocco model \cite{cianfroccoreview}.}
+  \label{mech-cycle}
+\end{figure}
+
+\subsubsection{ATP binding / unbinding effects}
+\subsubsection{Microtubule un/rebinding}
+\subsubsection{Conformation change of motor (``priming stroke'')}
+A FRET (fluorescent resonance energy transfer) study, which probes the closeness of different labeled domains on a protein, showed that dynein is found in two different major conformations, depending on the nucleotide bound to its AAA1 ATPase site \cite{FRETstatepaper}.
+When no nucleotide, ATP or ADP is bound, the authors label the protein ``state I.'' When ADP-Pi bound, the motor is said to be in ``state II,'' as shown in Fig \ref{nucleotide-state}.
+In addition, motors in the presence of the unhydrolyzable ATP analog were found in state II, indicating the transition from state I to II occurs during ATP binding.
+Similar reasoning shows that the state II to I transition occurs when the protein is ADP bound.
+This allows the authors to create the state diagram shown in Fig \ref{nucleotide-state-diagram}.\\
+
 %% \begin{figure}[h]
 %%   \centering
-%%   \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/mechanochemical-cycle.png}
-%%   \caption{Cycle of states the dynein motor goes through during motion in the Cianfrocco model \cite{cianfroccoreview}.}
-%%   \label{mech-cycle}
+%%   \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/two-state-em-diagram}
+%%   \caption{}
+%%   \label{nuecleotide-state}
 %% \end{figure}
 
-\subsubsection{Step One: ATP binding}
-\subsubsection{Step Two: }
+%% \begin{figure}[h]
+%%   \centering
+%%   \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/nucleotide-state-cycle}
+%%   \caption{}
+%%   \label{nucleotide-state-diagram}
+%% \end{figure}
+
+This single FRET study illustrates how the conformational aspect of the mechanochemical cycle works.
+Dynein begins in state I and binds ATP into its AAA1 ATPase, changing into state II by changing its stalk-tail angle to move the tail outward.
+ATP is then hydrolized into ADP-Pi.
+Pi is then ejected from the AAA1 site, leaving ADP behind.
+This causes the reverse state transition, moving the tail inward.
+When combined with microtubule affinity changes, this primitive model of the dynein cycle could explain how dynein walks.\\
+
+\textit{It is interesting to note that the hydrolysis of ATP is not actually used to directly power the motor -- binding AMP-PMP, a nonhydrolyzable analog of ATP, can also cause the motor to change from state I to state II \cite{FRETstatepaper}.
+The specialness of ATP is that it can be broken down into ADP, which causes another conformation change back to state I.
+Thus, something something the transfer of energy from nucleotide to motor is in its shape, not in its chemistry.}\\
+
+\textbf{more on how ATP actually causes the conformation change, if that's available.
+The change from an open/closed ATPase is what changes the nucleotide affinity, go into depth on that, but actual info on how ATP binding causes conformation change would be good.
+Something about going from a planar to nonplanar conformation?}\\
+
+also how crystal structures are in different positions
 
@@
 Weihong's figures go here.\\
-\subsubsection{Experimental features of dynein (ATPase, multi-state…)}
-Some history of studies which have determined what dynein has in it.\\
-
-\subsubsection{Powerstroke Theory}
-Talk about the history of this idea, why it is supported and why it is not supported.\\
-\paragraph{The theory}
-Describe the theory.
-\paragraph{b Evidence of theory}
-Evidence supporting the powerstroke.
 
@@
 Where our simulation fits in. Why we need this kind of information.
+There are many different things which might be important for dynein's walk which we don't take into account: springiness of tail/stalk, changes in spring constants, repulsion between the heads, multiple pre/post powerstroke conformations.
+We are trying to see if these things are actually important.
+
 \subsubsection{Other theories?}
@@
 
-\bibliography{thesis-webpages}
+%% \bibliography{thesis-webpages}
 

### 2017-01-08 thesis/latex/thesis.tex
Add work to introduction.

@@
 First, starting with both binding domains attached to the MT, ATP binds to the motor domain, which causes it to change conformation.
-ATP hydrolysis to ADP-Pi causes further conformation changes.
-These changes cause the binding domain to release the microtubule, meaning only one binding domain is MT-bound.
+This change causes the binding domain to release the microtubule, meaning only one binding domain is MT-bound.
 The free binding domain then undergoes another conformation change known as the ``priming stroke,'' which changes the orientation of the binding domain with respect to the MT.
@@
 \begin{itemize}
-\item AAA1 only ATPase necessary for motility, which is why we don't include the other ATPases in our model \cite{cianfroccoreview}
+\item AAA1 only ATPase necessary for motility \cite{cianfroccoreview}
 \item AAA5 mediates stalk register changes Schmidt H, Zalyte R, Urnavicius L, Carter AP. 2014. Structure of human cytoplasmic dynein-2 primed for
@@
 421:715–8
-\item Crystal structures with Vanadium-Adp show dynein in the pre-powerstroke state, hinting that ADP-Pi-bound dynein takes on a different conformation than ADP-Pi free dynein
 \end{itemize}
-
-Make a figure of pre/post-powerstroke dynein with the linker hilighted? Can you get what residues correspond to the linker somewhere?
 
@@
 \subsubsection{Other theories?}
-Imamula model: how it is different from Cianfrocco model, and why we chose the Cianfrocco model.\\
+Imamula model, how it is different from Cianfrocco model, and why we chose teh Cianfrocco model.\\
 	\subsection{Past Simulations}
+		\subsubsection{Why simulations are necessary}
 		\subsubsection{Sarlah Winch Model}
@@
 springy...
-
-Mention the Burgess paper, which says spring constants might change between apo/ADP states.
 
@@
 
-Do our spring constants seem to indicate that the protein does about one ATP's worth of work through its cycle? I wonder if the motor cools down its surroundings when it walks...\\
-
 Calorimetry could get Gibb's energy of binding if you froze the motor in onebound using
@@
 \item When is a fact ``common knowledge'' enough that it doesn't need a citation?
-\item Some papers have 'Request permission to reuse' links for figures - do I need to do this for copying figures?
+\item Some papers have 'Request permission to reuse' links - do I need to do this for copying figures?
 \end{itemize}

### 2017-01-09 thesis/latex/thesis.tex
Add work to methods. Doesn't compile. Todo: make compile and finish methods.

--- 2017-01-08 thesis/latex/thesis.tex
@@
 \usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins
+\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors
 
-\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors
+\usepackage{listings}
+\lstset{escapechar=\&}
 
@@
 
-\section{Chapter: Methods}
 \subsection{Simulation}
@@
 
-\subsection{Simulating our model}
-Talk generally about what simulating stuff means.\\
-\subsubsection{Motion equations}
-
-\paragraph{a Onebound solution}
-Discuss how our model is represented by the following figure (maybe get one with springs on the
-hinges):\\
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.45\textwidth]{../figures/equilibrium-onebound}
-  \caption{Definition of equilibrium angles.}
-  \label{fig:eq_angles}
-\end{figure}
-
-Show the derivation process.\\
-
-\paragraph{b Bothbound solution}
-Show how bothbound is derived and different from onebound. Put the actual equations in the appendix,
-or just leave them out 'cause they're ugly?\\
-
-\subsubsection{Code}
-Talk about using C++, Make, C, Python. Talk about the rough program flow, how we go from code to
-data.\\
-
-\paragraph{a Running, compiling, language choice?}
-Details on why we chose C++.\\
-\subsubsection{Time evolution}
-How we do time evolution.
 \paragraph{Euler’s method}
@@
 
-\paragraph{Finding the right dt}
-Figure on error rate vs. dt.
-
-\paragraph{State transitions}
-Talk about how we calculate transition rates based on differences in energies between states.
-Then how we use transition rates to transition our model accurately.\\
-
 \paragraph{Gibbs energy of transition}
@@
 
-\subsection{Verifying our model}
-Talk about why we think our model accurately represents the system we are trying to model, and
-doesn't have bugs in it.\\
+\section{Chapter: Methods}
 
-Talk about this figure:
+\subsection{Dynein Models}
+In order to simulate a protein's passage through the above mechanochemical cycle, two mathematical models are needed: one for the prestroke, MTBD-detached state, and one for the poststroke, MTBD-bound state.
+The purpose of each model is to calculate the position and velocity of each domain at an arbitrary time.\\
+
+\subsubsection{Prestroke onebound model}
+Prestroke dynein is referred to as onebound, since one motor domain is bound to the MT in this state, and the other unbound.
+A spatial representation of onebound dynein is shown in Figure \ref{ob_fig}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.45\textwidth]{../figures/ob_fig.pdf}
+  \caption{Definition of equilibrium angles.}
+  \label{ob_fig}
+\end{figure}
+
+TODO: label domain names (MTBD motor tail), Ls, Lt, microtubule\\
+
+Each domain is represented by a point, and is connected to other domains via fixed-length rods.
+Angles represent the degrees of freedom of the model.
+Since the protein is a dimer, there are two copies of the MTB, motor and tail domains.
+The tails are fused together and so represented as a single point.
+The motor and MTB domains are given a ``bound'' and ``unbound'' designation to differentiate them.
+The ``bound'' domains are the MTBD attached to the MT and its corresponding motor domain.
+The unbound domains are the unbound MTBD and its corresponding motor domain.\\
+
+\paragraph{Angular and Cartesian coordinates}
+
+The four angles $\Theta_{bb}, \Theta_{bm}, \Theta_{um}$ and $\Theta_{ub}$, corresponding to bound binding, bound motor, unbound motor and unbound binding angle, together describe any possible conformation the system can take on.
+The domain of $\Theta_{bb}$ has a restricted domain of $[0,\pi]$ to prevent below-MT conformations, but the other angles have domains of $[0,2\pi]$.
+Each angle is defined relative to the horizontal axis.\\
+
+\begin{figure}[h]
+  \centering
+  \begin{tabular}{c}
+    \begin{lstlisting}[language=C++]
+      class Dynein_onebound{
+        double &$\Theta_{bb}, \Theta_{bb}, \Theta_{bb}, \Theta_{bb}$&
+        double &$\dot{\Theta}_{bb}, \dot{\Theta}_{bb}, \dot{\Theta}_{bb}, \dot{\Theta}_{bb}$&
+        double &$X_{bb}$&
+      }
+    \end{lstlisting}
+  \end{tabular}
+  \caption{Pseudocode of the onebound Dynein model.}
+  \label{ob_struct}
+\end{figure}
+
+Pseudocode for the numerical onebound model representation is shown in Figure \ref{ob_struct}.
+The position of the bound binding domain $X_{bb}$ and the four domain angles are all the values needed to determine the position of the protein at any time.
+The following equations allow recursive calculation of the positions of each domain in cartesian coordinates:
+
+\noindent\begin{minipage}{0.49\linewidth}
+\begin{align}
+  X_{bm} &= X_{bb}+L_{s}\cos(\Theta_{bb}) \\
+  X_{t}  &= X_{bm}+L_{t}\cos(\Theta_{bm}) \\
+  X_{fm} &= X_{t} - L_{t}\cos(\Theta_{fm}) \\
+  X_{fb} &= X_{fm} - L_{s}\cos(\Theta_{fb})
+\end{align}
+\end{minipage}
+\begin{minipage}{0.49\linewidth}
+\begin{align}
+  Y_{bb} &= 0 \\
+  Y_{bm} &= Y_{bb}+L_{s}\sin(\Theta_{bb}) \\
+  Y_{t}  &= Y_{bm}+L_{t}\sin(\Theta_{bm}) \\
+  Y_{fm} &= Y_{t} - L_{t}\sin(\Theta_{fm}) \\
+  Y_{fb} &= Y_{fm} - L_{s}\sin(\Theta_{fb})
+\end{align}
+\end{minipage}
+\vspace{.5cm}
+
+$L_s$ and $L_t$ correspond to lengths of each interdomain rod, and the $t$ subscripts refer to tail domain coordinates.\\
+
+\paragraph{Angular and Cartesian velocities}
+The goal is to express angular velocities $\dot{\Theta}_{bb}, \dot{\Theta}_{bb}, \dot{\Theta}_{bb}$ and $\dot{\Theta}_{bb}$ in terms of known quantities.
+To begin, the cartesian velocities of each domain can be calculated from the above position equations in a similar recursive manner:
+
+\noindent\begin{minipage}{0.49\linewidth}
+\begin{align}
+  \dot{X}_{bb} &= 0 \\
+  \dot{X}_{bm} &= \dot{X}_{bb} - L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} \label{cartesian-bmx}\\
+  \dot{X}_{t } &= \dot{X}_{bm} - L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} \\
+  \dot{X}_{fm} &= \dot{X}_{t } + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} \\
+  \dot{X}_{fb} &= \dot{X}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb}
+\end{align}
+\end{minipage}
+\begin{minipage}{0.49\linewidth}
+\begin{align}                                                                          
+  \dot{Y}_{bb} &= 0 \\                                                        
+  \dot{Y}_{bm} &= \dot{Y}_{bb} + L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} \\
+  \dot{Y}_{t}  &= \dot{Y}_{bm} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} \\
+  \dot{Y}_{fm} &= \dot{Y}_{t } - L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} \\
+  \dot{Y}_{fb} &= \dot{Y}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb}
+\end{align}
+\end{minipage}
+\vspace{.5cm}
+
+Another way to express these cartesian velocities is using the Brownian dynamics equation $\dot{X} = \frac1\gamma F_{net} + R$:
+
+\begin{align}  
+  \dot{X}_{bm} &= \frac{1}{\gamma_m} \Big(F_{xml} + - \lambda_{bs}(X_{bm} - X_{bb}) + \lambda_{bt}(X_{t } - X_{bm}) \Big) + R_{xml} \label{brownian-bmx}\\
+  \dot{X}_{t } &= \frac{1}{\gamma_t} \Big(F_{xt } + - \lambda_{bt}(X_{t } - X_{bm}) + \lambda_{ft}(X_{fm} - X_{t }) \Big) + R_{xt } \\
+  \dot{X}_{fm} &= \frac{1}{\gamma_m} \Big(F_{xmr} + - \lambda_{ft}(X_{fm} - X_{t }) + \lambda_{fs}(X_{fb} - X_{fm}) \Big) + R_{xmr} \\
+  \dot{X}_{fb} &= \frac{1}{\gamma_b} \Big(F_{xbr} + - \lambda_{fs}(X_{fb} - X_{fm}) \Big) + R_{xbr}
+\end{align}
+
+\begin{align}  
+  \dot{Y}_{bm} &= \frac{1}{\gamma_m} \Big(F_{yml} + - \lambda_{bs}(Y_{bm} - Y_{bb}) + \lambda_{bt}(Y_{t } - Y_{bm}) \Big) + R_{yml} \\
+  \dot{Y}_{t}  &= \frac{1}{\gamma_t} \Big(F_{yt } + - \lambda_{bt}(Y_{t } - Y_{bm}) + \lambda_{ft}(Y_{fm} - Y_{t }) \Big) + R_{yt } \\
+  \dot{Y}_{fm} &= \frac{1}{\gamma_m} \Big(F_{ymr} + - \lambda_{ft}(Y_{fm} - Y_{t }) + \lambda_{fs}(Y_{fb} - Y_{fm}) \Big) + R_{ymr} \\
+  \dot{Y}_{fb} &= \frac{1}{\gamma_b} \Big(F_{ybr} + - \lambda_{fs}(Y_{fb} - Y_{fm}) \Big) + R_{ybr}
+\end{align}
+
+where $\gamma_n$ is a drag coefficient for the binding, motor or tail domain with units of mass per second.
+$F_{xn}$ is the x component of external force on each domain $n$ due to various factors.
+$\lambda_{12}\left(X_1 - X_2\right)$ is the x component of internal force on domain $2$ due to domain $1$, where $\lambda$ is a tension coefficient with units of mass per second per second.
+$R_{xn}$ is the Brownian coefficient representing motion due to solvent collision, with units of velocity.\\
+
+These two sets of equations, cartesian and Brownian, can be equated to get more interesting equations.
+For example, Eq (\ref{cartesian-bmx}) and Eq (\ref{brownian-bmx}) can be equated.
+This equating, combined with expanding the recursive velocity definitions leads to:
+
+\begin{align*}
+  -L_s\sin(\theta_{bb})\dot{\theta}_{bb} &= \frac{1}{\gamma_m}F_{xml} + -\frac{1}{\gamma_m}\lambda_{bs}(X_{bm} - X_{bb}) + \frac{1}{\gamma_m}\lambda_{bt}(X_{t } - X_{bm}) + R_{bmx} \\
+  -L_s\sin(\theta_{bb})\dot{\theta}_{bb} - L_t\sin(\theta_{bm})\dot{\theta}_{bm} &= \frac{1}{\gamma_t}F_{xt } + -\frac{1}{\gamma_t}\lambda_{bt}(X_{t } - X_{bm}) + \frac{1}{\gamma_t}\lambda_{ft}(X_{um} - X_{t }) + R_{tx} \\
+  -L_s\sin(\theta_{bb})\dot{\theta}_{bb} - L_t\sin(\theta_{bm})\dot{\theta}_{bm} + L_t\sin(\theta_{um})\dot{\theta}_{um} &= \frac{1}{\gamma_m}F_{xmr} + -\frac{1}{\gamma_m}\lambda_{ft}(X_{um} - X_{t }) + \frac{1}{\gamma_m}\lambda_{fs}(X_{ub} - X_{um}) + R_{bmx} \\
+  -L_s\sin(\theta_{bb})\dot{\theta}_{bb} - L_t\sin(\theta_{bm})\dot{\theta}_{bm} + L_t\sin(\theta_{um})\dot{\theta}_{um} + L_s\sin(\theta_{ub})\dot{\theta}_{ub} &= \frac{1}{\gamma_b}F_{xbr} + -\frac{1}{\gamma_b}\lambda_{fs}(X_{ub} - X_{um}) + R_{bbx}
+\end{align*}
+
+\begin{align*}
+  L_s\cos(\theta_{bb})\dot{\theta}_{bm} &= \frac{1}{\gamma_m}F_{yml} + -\frac{1}{\gamma_m}\lambda_{bs}(Y_{bm} - Y_{bb}) + \frac{1}{\gamma_m}\lambda_{bt}(Y_{t } - Y_{bm}) + R_{umy} \\
+  L_s\cos(\theta_{bb})\dot{\theta}_{bb} + L_t\cos(\theta_{bm})\dot{\theta}_{bm} &= \frac{1}{\gamma_t}F_{yt } + -\frac{1}{\gamma_t}\lambda_{bt}(Y_{t } - Y_{bm}) + \frac{1}{\gamma_t}\lambda_{ft}(Y_{um} - Y_{t }) + R_{ty} \\
+  L_s\cos(\theta_{bb})\dot{\theta}_{bb} + L_t\cos(\theta_{bm})\dot{\theta}_{bm} - L_t\cos(\theta_{um})\dot{\theta}_{um} &= \frac{1}{\gamma_m}F_{ymr} + -\frac{1}{\gamma_m}\lambda_{ft}(Y_{um} - Y_{t }) + \frac{1}{\gamma_m}\lambda_{fs}(Y_{ub} - Y_{um}) + R_{bmy} \\
+  L_s\cos(\theta_{bb})\dot{\theta}_{bb} + L_t\cos(\theta_{bm})\dot{\theta}_{bm} - L_t\cos(\theta_{um})\dot{\theta}_{um} - L_s\cos(\theta_{ub})\dot{\theta}_{ub} &= \frac{1}{\gamma_b}F_{ybr} + -\frac{1}{\gamma_b}\lambda_{fs}(Y_{ub} - Y_{um}) + R_{bby}
+\end{align*}
+
+These eight equations form a system of equations with eight unknowns: $\dot{\Theta}_{bb}, \dot{\Theta}_{bb}, \dot{\Theta}_{bb}$ and $\dot{\Theta}_{bb}$, and the four tension coefficients $\lambda_{bs}, \lambda_{bt}, \lambda_{um}$ and $\lambda_{ub}$.
+This system is more compactly represented as:
+{\small
+  \[
+  \hspace{-1.0cm}
+  \resizebox{0.85\linewidth}{!}{%
+  $\begin{pmatrix}
+    L_s\sin\theta_{bb} & 0 & 0 & 0 & -\gamma_m (X_{bm} - X_{bb}) & \gamma_m (X_{t } - X_{bm}) & 0 & 0\\
+    L_s\sin(\theta_{bb}) & L_t\sin(\theta_{bm}) & 0 & 0 & 0 & -\gamma_t (X_{t } - X_{bm}) & \gamma_t (X_{um} - X_{t }) & 0\\
+    L_s\sin(\theta_{bb}) & L_t\sin(\theta_{bm}) & -L_t\sin(\theta_{um}) & 0 & 0 & 0 & -\gamma_m (X_{um} - X_{t }) & \gamma_m (X_{ub} - X_{um})\\
+    L_s\sin(\theta_{bb}) & L_t\sin(\theta_{bm}) & -L_t\sin(\theta_{um}) & -L_s\sin(\theta_{ub}) & 0 & 0 & 0 & -\gamma_b (X_{ub} - X_{um})\\
+    -L_s\cos(\theta_{bb}) & 0 & 0 & 0 & -\gamma_m (Y_{bm} - Y_{bb}) & \gamma_m (Y_{t } - Y_{bm}) & 0 & 0\\
+    -L_s\cos(\theta_{bb}) & -L_t\cos(\theta_{bm}) & 0 & 0 & 0 & -\gamma_t (Y_{t } - Y_{bm}) & \gamma_t (Y_{um} - Y_{t }) & 0\\
+    -L_s\cos(\theta_{bb}) & -L_t\cos(\theta_{bm}) & L_t\cos(\theta_{um}) & 0 & 0 & 0 & -\gamma_m (Y_{um} - Y_{t }) & \gamma_m (Y_{ub} - Y_{um})\\
+    -L_s\cos(\theta_{bb}) & -L_t\cos(\theta_{bm}) & L_t\cos(\theta_{um}) & L_s\cos(\theta_{ub}) & 0 & 0 & 0 & -\gamma_b (Y_{ub} - Y_{um})\\
+  \end{pmatrix}$%
+  }
+  \begin{pmatrix}
+    \dot{\theta}_{bb}\\
+    \dot{\theta}_{bm}\\
+    \dot{\theta}_{um}\\
+    \dot{\theta}_{ub}\\
+    \lambda_{bs}\\
+    \lambda_{bt}\\
+    \lambda_{ut}\\
+    \lambda_{us}\\
+  \end{pmatrix}
+  =
+  \begin{pmatrix}
+    -F_{bmx} + \gamma_m R_{bmx}\\
+    -F_{tx} + \gamma_t R_{tx}\\
+    -F_{umx} + \gamma_m R_{umx}\\
+    -F_{ubx} + \gamma_b R_{ubx}\\
+    -F_{bmy} + \gamma_m R_{bmy}\\
+    -F_{ty} + \gamma_t R_{ty}\\
+    -F_{umy} + \gamma_m R_{umy}\\
+    -F_{uby} + \gamma_b R_{uby}\\
+  \end{pmatrix}
+  \]
+}
+
+This matrix is then solved using the Mathematica algebra system, resulting in a set of equations shown in Appendix (\ref{ob-motion-equations]}).
+
+------------------------------------------
 
@@
 
-and how it shows our model is physical, or at least, how obeying the equipartition theorem shows
-how our domains both feel harmonic energies and can fully explore their spaces?\\
-
-\subsection{Parameters}
-The following table shows our guess parameters from various papers/crystal structures.\\
-
-                        \begin{center}
-  \begin{tabular}{| l | l | l | l | l | l |}
-    \hline
-    & Winch (Sarlah) & Schmidt & Lin & PyMol 3VKH & PyMol 4RH7\\\hline
-    $L_s$ & 12nm & & & 21.0nm & 22.1nm\\ \hline
-    $L_t$ &  7nm & & & & 11.15nm\\ \hline
-    $R_b$ &  & & & 1.57nm & 1.45nm\\ \hline
-    $R_m$ &  7nm & & & 7.36nm & 6.3nm\\ \hline
-    $R_t$ &  & & & & 2.16nm \\ \hline
-    $\theta_{m}^{Pre}$ & 250$^{\circ}$ & & 171$^{\circ}$ & &\\ \hline
-    $\theta_{m}^{Post}$ & 330$^{\circ}$ & & 137.5$^{\circ}$ & &\\ \hline
-    $\theta_{b}$ & 56$^{\circ}$ & & 63.5$^{\circ}$ & &\\ \hline
-    $\theta_{t}$ & 0$^{\circ}$ & & & &\\ \hline
-    $k_{ub}$ & $180 s^{-1,a}$ & & & &\\ \hline
-    $k_b$ & $5000 s^{-1,b}$ & & & &\\ \hline
-  \end{tabular}
-\end{center}
-		\subsubsection{Lengths, angles, binding constants, etc…}
-	\subsection{Tuning our model}
-	\subsubsection{Searching for optimal parameters}
 \section{Chapter: Results}
@@
 
+\section{Appendix}
+
+\subsection{Motion equations}
+
+\subsubsection{Onebound equations}
+\label{onebound-motion-equations}
+These are the full motion equations used to time-evolve the onebound model:
+
+\begin{align*}
+AA &= L_s\sin\theta_{bb}      &   LL &= -L_s\cos(\theta_{bb})\\
+BB &= L_t\sin(\theta_{bm})    &	 MM &= -L_t\cos(\theta_{bm})\\
+CC &= -L_t\sin(\theta_{um})   &	 NN &= L_t\cos(\theta_{um})\\
+DD &= -L_s\sin(\theta_{ub})   &	 OO &= L_s\cos(\theta_{ub})\\
+\\
+EE &= -\gamma_m (X_{bm} - X_{bb})       &  PP &= -\gamma_m (Y_{bm} - Y_{bb})\\
+FF &= \gamma_m (X_{t } - X_{bm})        &	 QQ &= \gamma_m (Y_{t } - Y_{bm})\\ 
+GG &= -\gamma_t (X_{t } - X_{bm})       &	 RR &= -\gamma_t (Y_{t } - Y_{bm})\\
+HH &= \gamma_t (X_{um} - X_{t })        &	 SS &= \gamma_t (Y_{um} - Y_{t })\\ 
+II &= -\gamma_m (X_{um} - X_{t })       &	 TT &= -\gamma_m (Y_{um} - Y_{t })\\
+JJ &= \gamma_m (X_{ub} - X_{um})        &	 UU &= \gamma_m (Y_{ub} - Y_{um})\\ 
+KK &= -\gamma_b (X_{ub} - X_{um})       &	 VV &= -\gamma_b (Y_{ub} - Y_{um})\\
+\end{align*}%
+%
+\[
+\begin{pmatrix}
+  AA & 0 & 0 & 0 & EE & FF & 0 & 0\\
+  AA & BB & 0 & 0 & 0 & GG & HH & 0\\
+  AA & BB & CC & 0 & 0 & 0 & II & JJ\\
+  AA & BB & CC & DD & 0 & 0 & 0 & KK\\
+  LL & 0 & 0 & 0 & PP & QQ & 0 & 0\\
+  LL & MM & 0 & 0 & 0 & RR & SS & 0\\
+  LL & MM & NN & 0 & 0 & 0 & TT & UU\\
+  LL & MM & NN & OO & 0 & 0 & 0 & VV\\
+\end{pmatrix}
+\begin{pmatrix}
+    \dot{\theta}_{bb}\\
+    \dot{\theta}_{bm}\\
+    \dot{\theta}_{um}\\
+    \dot{\theta}_{ub}\\
+    \lambda_{bs}\\
+    \lambda_{bt}\\
+    \lambda_{ut}\\
+    \lambda_{us}\\
+  \end{pmatrix}
+=
+\begin{pmatrix}
+  X1\\
+  X2\\
+  X3\\
+  X4\\
+  X5\\
+  X6\\
+  X7\\
+  X8\\
+\end{pmatrix}
+\]
+
+Solving the above system for $\dot{\Theta}_{bb}, \dot{\Theta}_{bb}, \dot{\Theta}_{bb}$ and $\dot{\Theta}_{bb}$ gives
+
+\begin{multline}
+d_bba = (-(GG*II*MM*NN*PP*X1) + CC*HH*MM*PP*RR*X1 -
+      BB*HH*NN*PP*RR*X1 + BB*II*NN*PP*RR*X1 - CC*GG*MM*PP*SS*X1 +
+      BB*GG*NN*PP*SS*X1 + CC*GG*MM*PP*TT*X1 - BB*CC*PP*RR*TT*X1 +
+      FF*II*MM*NN*PP*X2 - EE*II*MM*NN*QQ*X2 + CC*FF*MM*PP*SS*X2 -
+      BB*FF*NN*PP*SS*X2 - CC*EE*MM*QQ*SS*X2 + BB*EE*NN*QQ*SS*X2 -
+      CC*FF*MM*PP*TT*X2 + CC*EE*MM*QQ*TT*X2 - FF*HH*MM*NN*PP*X3 +
+      EE*HH*MM*NN*QQ*X3 + BB*FF*NN*PP*SS*X3 - BB*EE*NN*QQ*SS*X3 +
+      EE*GG*II*MM*NN*X5 - CC*EE*HH*MM*RR*X5 + BB*EE*HH*NN*RR*X5 -
+      BB*EE*II*NN*RR*X5 + CC*EE*GG*MM*SS*X5 - BB*EE*GG*NN*SS*X5 -
+      CC*EE*GG*MM*TT*X5 + BB*CC*EE*RR*TT*X5 - CC*FF*HH*MM*PP*X6 +
+      BB*FF*HH*NN*PP*X6 - BB*FF*II*NN*PP*X6 + CC*EE*HH*MM*QQ*X6 -
+      BB*EE*HH*NN*QQ*X6 + BB*EE*II*NN*QQ*X6 + BB*CC*FF*PP*TT*X6 -
+      BB*CC*EE*QQ*TT*X6 + CC*FF*HH*MM*PP*X7 - CC*EE*HH*MM*QQ*X7 -
+      BB*CC*FF*PP*SS*X7 + BB*CC*EE*QQ*SS*X7)/ (EE*GG*II*LL*MM*NN +
+      BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP +
+      AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ +
+      BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ -
+      CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR +
+      AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR +
+      CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS +
+      AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS +
+      BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT +
+      BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT -
+      BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT -
+      AA*BB*CC*PP*RR*TT) - ((FF*PP - EE*QQ)*(HH*MM - BB*SS)*(JJ*NN -
+      CC*UU)*(-(((Power(AA,6)*Power(BB,2)*CC*DD*HH*MM -
+      Power(AA,6)*Power(BB,3)*DD*HH*NN +
+      Power(AA,6)*Power(BB,3)*DD*II*NN -
+      Power(AA,6)*Power(BB,3)*CC*II*OO)* (-(EE*LL) +
+      AA*PP)*(-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+      Power(AA,2)*BB*RR))) - (-(EE*LL) +
+      AA*PP)*((Power(AA,5)*Power(BB,3)*CC*DD*FF*LL -
+      Power(AA,6)*Power(BB,2)*CC*DD*FF*MM +
+      Power(AA,6)*Power(BB,2)*CC*DD*GG*MM -
+      Power(AA,6)*Power(BB,3)*DD*GG*NN)* (-(EE*LL) + AA*PP) -
+      (Power(AA,5)*Power(BB,3)*CC*DD*EE*LL -
+      Power(AA,6)*Power(BB,2)*CC*DD*EE*MM)*(-(FF*LL) +
+      AA*QQ))*(Power(AA,4)*BB*CC*HH*MM -
+      Power(AA,4)*Power(BB,2)*CC*SS))*
+      (-(((Power(AA,3)*Power(BB,2)*CC*FF*LL - Power(AA,4)*BB*CC*FF*MM
+      + Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,4)*Power(BB,2)*GG*NN)*(-(EE*LL) + AA*PP) -
+      (Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ))*
+      (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(LL*X1) + AA*X5)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,3)*Power(BB,2)*CC*LL*X1 -
+      Power(AA,4)*BB*CC*MM*X1 + Power(AA,4)*BB*CC*MM*X2 -
+      Power(AA,4)*Power(BB,2)*CC*X6))) +
+      (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+      Power(AA,2)*BB*RR)))* (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(LL*X1) + AA*X5)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,3)*Power(BB,2)*CC*LL*X1 -
+      Power(AA,4)*BB*CC*MM*X1 + Power(AA,4)*BB*CC*MM*X2 -
+      Power(AA,4)*Power(BB,2)*NN*X2 + Power(AA,4)*Power(BB,2)*NN*X3 -
+      Power(AA,4)*Power(BB,2)*CC*X7)))) + (-((-(EE*LL) +
+      AA*PP)*((Power(AA,3)*Power(BB,2)*CC*FF*LL -
+      Power(AA,4)*BB*CC*FF*MM + Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,4)*Power(BB,2)*GG*NN)* (-(EE*LL) + AA*PP) -
+      (Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) +
+      AA*QQ))*(Power(AA,4)*BB*CC*HH*MM -
+      Power(AA,4)*Power(BB,2)*CC*SS)) + (-(EE*LL) +
+      AA*PP)*(-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+      Power(AA,2)*BB*RR)))* (Power(AA,4)*BB*CC*HH*MM -
+      Power(AA,4)*Power(BB,2)*HH*NN + Power(AA,4)*Power(BB,2)*II*NN -
+      Power(AA,4)*Power(BB,2)*CC*TT))*
+      (-(((Power(AA,5)*Power(BB,3)*CC*DD*FF*LL -
+      Power(AA,6)*Power(BB,2)*CC*DD*FF*MM +
+      Power(AA,6)*Power(BB,2)*CC*DD*GG*MM -
+      Power(AA,6)*Power(BB,3)*DD*GG*NN)*(-(EE*LL) + AA*PP) -
+      (Power(AA,5)*Power(BB,3)*CC*DD*EE*LL -
+      Power(AA,6)*Power(BB,2)*CC*DD*EE*MM)*(-(FF*LL) + AA*QQ))*
+      (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(LL*X1) + AA*X5)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,3)*Power(BB,2)*CC*LL*X1 -
+      Power(AA,4)*BB*CC*MM*X1 + Power(AA,4)*BB*CC*MM*X2 -
+      Power(AA,4)*Power(BB,2)*CC*X6))) +
+      (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+      Power(AA,2)*BB*RR)))* (-((Power(AA,5)*Power(BB,3)*CC*DD*EE*LL -
+      Power(AA,6)*Power(BB,2)*CC*DD*EE*MM)*(-(LL*X1) + AA*X5)) +
+      (-(EE*LL) + AA*PP)*(Power(AA,5)*Power(BB,3)*CC*DD*LL*X1 -
+      Power(AA,6)*Power(BB,2)*CC*DD*MM*X1 +
+      Power(AA,6)*Power(BB,2)*CC*DD*MM*X2 -
+      Power(AA,6)*Power(BB,3)*DD*NN*X2 +
+      Power(AA,6)*Power(BB,3)*DD*NN*X3 -
+      Power(AA,6)*Power(BB,3)*CC*OO*X3 +
+      Power(AA,6)*Power(BB,3)*CC*OO*X4 -
+      Power(AA,6)*Power(BB,3)*CC*DD*X8)))))/ ((-(EE*GG*II*LL*MM*NN) -
+      BB*FF*HH*LL*NN*PP + BB*FF*II*LL*NN*PP + AA*FF*HH*MM*NN*PP -
+      AA*FF*II*MM*NN*PP + AA*GG*II*MM*NN*PP + BB*EE*HH*LL*NN*QQ -
+      BB*EE*II*LL*NN*QQ - AA*EE*HH*MM*NN*QQ + AA*EE*II*MM*NN*QQ +
+      CC*EE*HH*LL*MM*RR - BB*EE*HH*LL*NN*RR + BB*EE*II*LL*NN*RR -
+      AA*CC*HH*MM*PP*RR + AA*BB*HH*NN*PP*RR - AA*BB*II*NN*PP*RR -
+      CC*EE*GG*LL*MM*SS + BB*EE*GG*LL*NN*SS + BB*CC*FF*LL*PP*SS -
+      AA*CC*FF*MM*PP*SS + AA*CC*GG*MM*PP*SS - AA*BB*GG*NN*PP*SS -
+      BB*CC*EE*LL*QQ*SS + AA*CC*EE*MM*QQ*SS + CC*EE*GG*LL*MM*TT -
+      BB*CC*FF*LL*PP*TT + AA*CC*FF*MM*PP*TT - AA*CC*GG*MM*PP*TT +
+      BB*CC*EE*LL*QQ*TT - AA*CC*EE*MM*QQ*TT - BB*CC*EE*LL*RR*TT +
+      AA*BB*CC*PP*RR*TT)* (-((-(EE*LL) +
+      AA*PP)*(-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+      Power(AA,2)*BB*RR)))* ((Power(AA,6)*Power(BB,2)*CC*DD*HH*MM -
+      Power(AA,6)*Power(BB,3)*DD*HH*NN +
+      Power(AA,6)*Power(BB,3)*DD*II*NN -
+      Power(AA,6)*Power(BB,3)*CC*II*OO)*(-(EE*LL) + AA*PP)*
+      (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+      Power(AA,2)*BB*RR))) - (-(EE*LL) +
+      AA*PP)*((Power(AA,5)*Power(BB,3)*CC*DD*FF*LL -
+      Power(AA,6)*Power(BB,2)*CC*DD*FF*MM +
+      Power(AA,6)*Power(BB,2)*CC*DD*GG*MM -
+      Power(AA,6)*Power(BB,3)*DD*GG*NN)* (-(EE*LL) + AA*PP) -
+      (Power(AA,5)*Power(BB,3)*CC*DD*EE*LL -
+      Power(AA,6)*Power(BB,2)*CC*DD*EE*MM)*(-(FF*LL) +
+      AA*QQ))*(Power(AA,4)*BB*CC*HH*MM -
+      Power(AA,4)*Power(BB,2)*CC*SS))* (Power(AA,4)*Power(BB,2)*JJ*NN
+      - Power(AA,4)*Power(BB,2)*CC*UU)) + (-(EE*LL) + AA*PP)*
+      (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+      Power(AA,2)*BB*RR)))* (-((-(EE*LL) +
+      AA*PP)*((Power(AA,3)*Power(BB,2)*CC*FF*LL -
+      Power(AA,4)*BB*CC*FF*MM + Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,4)*Power(BB,2)*GG*NN)*(-(EE*LL) + AA*PP) -
+      (Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) +
+      AA*QQ))*(Power(AA,4)*BB*CC*HH*MM -
+      Power(AA,4)*Power(BB,2)*CC*SS)) + (-(EE*LL) +
+      AA*PP)*(-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+      Power(AA,2)*BB*RR)))* (Power(AA,4)*BB*CC*HH*MM -
+      Power(AA,4)*Power(BB,2)*HH*NN + Power(AA,4)*Power(BB,2)*II*NN -
+      Power(AA,4)*Power(BB,2)*CC*TT))*
+      (Power(AA,6)*Power(BB,3)*DD*JJ*NN -
+      Power(AA,6)*Power(BB,3)*CC*JJ*OO +
+      Power(AA,6)*Power(BB,3)*CC*KK*OO -
+      Power(AA,6)*Power(BB,3)*CC*DD*VV)));
+\end{multline}
+
+\begin{multline}
+  d_bma = (GG*II*LL*NN*PP*X1 - CC*HH*LL*PP*RR*X1 + AA*HH*NN*PP*RR*X1 -
+      AA*II*NN*PP*RR*X1 + CC*GG*LL*PP*SS*X1 - AA*GG*NN*PP*SS*X1 -
+      CC*GG*LL*PP*TT*X1 + AA*CC*PP*RR*TT*X1 - FF*II*LL*NN*PP*X2 +
+      EE*II*LL*NN*QQ*X2 - EE*II*LL*NN*RR*X2 + AA*II*NN*PP*RR*X2 -
+      CC*FF*LL*PP*SS*X2 + AA*FF*NN*PP*SS*X2 + CC*EE*LL*QQ*SS*X2 -
+      AA*EE*NN*QQ*SS*X2 + CC*FF*LL*PP*TT*X2 - CC*EE*LL*QQ*TT*X2 +
+      CC*EE*LL*RR*TT*X2 - AA*CC*PP*RR*TT*X2 + FF*HH*LL*NN*PP*X3 -
+      EE*HH*LL*NN*QQ*X3 + EE*HH*LL*NN*RR*X3 - AA*HH*NN*PP*RR*X3 -
+      EE*GG*LL*NN*SS*X3 - AA*FF*NN*PP*SS*X3 + AA*GG*NN*PP*SS*X3 +
+      AA*EE*NN*QQ*SS*X3 - EE*GG*II*LL*NN*X5 + CC*EE*HH*LL*RR*X5 -
+      AA*EE*HH*NN*RR*X5 + AA*EE*II*NN*RR*X5 - CC*EE*GG*LL*SS*X5 +
+      AA*EE*GG*NN*SS*X5 + CC*EE*GG*LL*TT*X5 - AA*CC*EE*RR*TT*X5 +
+      EE*GG*II*LL*NN*X6 + CC*FF*HH*LL*PP*X6 - AA*FF*HH*NN*PP*X6 +
+      AA*FF*II*NN*PP*X6 - AA*GG*II*NN*PP*X6 - CC*EE*HH*LL*QQ*X6 +
+      AA*EE*HH*NN*QQ*X6 - AA*EE*II*NN*QQ*X6 - CC*EE*GG*LL*TT*X6 -
+      AA*CC*FF*PP*TT*X6 + AA*CC*GG*PP*TT*X6 + AA*CC*EE*QQ*TT*X6 -
+      CC*FF*HH*LL*PP*X7 + CC*EE*HH*LL*QQ*X7 - CC*EE*HH*LL*RR*X7 +
+      AA*CC*HH*PP*RR*X7 + CC*EE*GG*LL*SS*X7 + AA*CC*FF*PP*SS*X7 -
+      AA*CC*GG*PP*SS*X7 - AA*CC*EE*QQ*SS*X7)/ (EE*GG*II*LL*MM*NN +
+      BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP +
+      AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ +
+      BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ -
+      CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR +
+      AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR +
+      CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS +
+      AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS +
+      BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT +
+      BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT -
+      BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT -
+      AA*BB*CC*PP*RR*TT) + ((-(FF*HH*LL*PP) + EE*HH*LL*QQ -
+      EE*HH*LL*RR + AA*HH*PP*RR + EE*GG*LL*SS + AA*FF*PP*SS -
+      AA*GG*PP*SS - AA*EE*QQ*SS)*(JJ*NN - CC*UU)*
+      (-(((Power(AA,6)*Power(BB,2)*CC*DD*HH*MM -
+      Power(AA,6)*Power(BB,3)*DD*HH*NN +
+      Power(AA,6)*Power(BB,3)*DD*II*NN -
+      Power(AA,6)*Power(BB,3)*CC*II*OO)*(-(EE*LL) + AA*PP)*
+      (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+      Power(AA,2)*BB*RR))) - (-(EE*LL) +
+      AA*PP)*((Power(AA,5)*Power(BB,3)*CC*DD*FF*LL -
+      Power(AA,6)*Power(BB,2)*CC*DD*FF*MM +
+      Power(AA,6)*Power(BB,2)*CC*DD*GG*MM -
+      Power(AA,6)*Power(BB,3)*DD*GG*NN)* (-(EE*LL) + AA*PP) -
+      (Power(AA,5)*Power(BB,3)*CC*DD*EE*LL -
+      Power(AA,6)*Power(BB,2)*CC*DD*EE*MM)*(-(FF*LL) +
+      AA*QQ))*(Power(AA,4)*BB*CC*HH*MM -
+      Power(AA,4)*Power(BB,2)*CC*SS))*
+      (-(((Power(AA,3)*Power(BB,2)*CC*FF*LL - Power(AA,4)*BB*CC*FF*MM
+      + Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,4)*Power(BB,2)*GG*NN)*(-(EE*LL) + AA*PP) -
+      (Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ))*
+      (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(LL*X1) + AA*X5)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,3)*Power(BB,2)*CC*LL*X1 -
+      Power(AA,4)*BB*CC*MM*X1 + Power(AA,4)*BB*CC*MM*X2 -
+      Power(AA,4)*Power(BB,2)*CC*X6))) +
+      (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+      Power(AA,2)*BB*RR)))* (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(LL*X1) + AA*X5)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,3)*Power(BB,2)*CC*LL*X1 -
+      Power(AA,4)*BB*CC*MM*X1 + Power(AA,4)*BB*CC*MM*X2 -
+      Power(AA,4)*Power(BB,2)*NN*X2 + Power(AA,4)*Power(BB,2)*NN*X3 -
+      Power(AA,4)*Power(BB,2)*CC*X7)))) + (-((-(EE*LL) +
+      AA*PP)*((Power(AA,3)*Power(BB,2)*CC*FF*LL -
+      Power(AA,4)*BB*CC*FF*MM + Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,4)*Power(BB,2)*GG*NN)* (-(EE*LL) + AA*PP) -
+      (Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) +
+      AA*QQ))*(Power(AA,4)*BB*CC*HH*MM -
+      Power(AA,4)*Power(BB,2)*CC*SS)) + (-(EE*LL) +
+      AA*PP)*(-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+      Power(AA,2)*BB*RR)))* (Power(AA,4)*BB*CC*HH*MM -
+      Power(AA,4)*Power(BB,2)*HH*NN + Power(AA,4)*Power(BB,2)*II*NN -
+      Power(AA,4)*Power(BB,2)*CC*TT))*
+      (-(((Power(AA,5)*Power(BB,3)*CC*DD*FF*LL -
+      Power(AA,6)*Power(BB,2)*CC*DD*FF*MM +
+      Power(AA,6)*Power(BB,2)*CC*DD*GG*MM -
+      Power(AA,6)*Power(BB,3)*DD*GG*NN)*(-(EE*LL) + AA*PP) -
+      (Power(AA,5)*Power(BB,3)*CC*DD*EE*LL -
+      Power(AA,6)*Power(BB,2)*CC*DD*EE*MM)*(-(FF*LL) + AA*QQ))*
+      (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(LL*X1) + AA*X5)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,3)*Power(BB,2)*CC*LL*X1 -
+      Power(AA,4)*BB*CC*MM*X1 + Power(AA,4)*BB*CC*MM*X2 -
+      Power(AA,4)*Power(BB,2)*CC*X6))) +
+      (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+      Power(AA,2)*BB*RR)))* (-((Power(AA,5)*Power(BB,3)*CC*DD*EE*LL -
+      Power(AA,6)*Power(BB,2)*CC*DD*EE*MM)*(-(LL*X1) + AA*X5)) +
+      (-(EE*LL) + AA*PP)*(Power(AA,5)*Power(BB,3)*CC*DD*LL*X1 -
+      Power(AA,6)*Power(BB,2)*CC*DD*MM*X1 +
+      Power(AA,6)*Power(BB,2)*CC*DD*MM*X2 -
+      Power(AA,6)*Power(BB,3)*DD*NN*X2 +
+      Power(AA,6)*Power(BB,3)*DD*NN*X3 -
+      Power(AA,6)*Power(BB,3)*CC*OO*X3 +
+      Power(AA,6)*Power(BB,3)*CC*OO*X4 -
+      Power(AA,6)*Power(BB,3)*CC*DD*X8)))))/ ((EE*GG*II*LL*MM*NN +
+      BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP +
+      AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ +
+      BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ -
+      CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR +
+      AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR +
+      CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS +
+      AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS +
+      BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT +
+      BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT -
+      BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT -
+      AA*BB*CC*PP*RR*TT)* (-((-(EE*LL) +
+      AA*PP)*(-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+      Power(AA,2)*BB*RR)))* ((Power(AA,6)*Power(BB,2)*CC*DD*HH*MM -
+      Power(AA,6)*Power(BB,3)*DD*HH*NN +
+      Power(AA,6)*Power(BB,3)*DD*II*NN -
+      Power(AA,6)*Power(BB,3)*CC*II*OO)*(-(EE*LL) + AA*PP)*
+      (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+      Power(AA,2)*BB*RR))) - (-(EE*LL) +
+      AA*PP)*((Power(AA,5)*Power(BB,3)*CC*DD*FF*LL -
+      Power(AA,6)*Power(BB,2)*CC*DD*FF*MM +
+      Power(AA,6)*Power(BB,2)*CC*DD*GG*MM -
+      Power(AA,6)*Power(BB,3)*DD*GG*NN)* (-(EE*LL) + AA*PP) -
+      (Power(AA,5)*Power(BB,3)*CC*DD*EE*LL -
+      Power(AA,6)*Power(BB,2)*CC*DD*EE*MM)*(-(FF*LL) +
+      AA*QQ))*(Power(AA,4)*BB*CC*HH*MM -
+      Power(AA,4)*Power(BB,2)*CC*SS))* (Power(AA,4)*Power(BB,2)*JJ*NN
+      - Power(AA,4)*Power(BB,2)*CC*UU)) + (-(EE*LL) + AA*PP)*
+      (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+      Power(AA,2)*BB*RR)))* (-((-(EE*LL) +
+      AA*PP)*((Power(AA,3)*Power(BB,2)*CC*FF*LL -
+      Power(AA,4)*BB*CC*FF*MM + Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,4)*Power(BB,2)*GG*NN)*(-(EE*LL) + AA*PP) -
+      (Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) +
+      AA*QQ))*(Power(AA,4)*BB*CC*HH*MM -
+      Power(AA,4)*Power(BB,2)*CC*SS)) + (-(EE*LL) +
+      AA*PP)*(-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+      Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+      AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+      Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+      Power(AA,2)*BB*RR)))* (Power(AA,4)*BB*CC*HH*MM -
+      Power(AA,4)*Power(BB,2)*HH*NN + Power(AA,4)*Power(BB,2)*II*NN -
+      Power(AA,4)*Power(BB,2)*CC*TT))*
+      (Power(AA,6)*Power(BB,3)*DD*JJ*NN -
+      Power(AA,6)*Power(BB,3)*CC*JJ*OO +
+      Power(AA,6)*Power(BB,3)*CC*KK*OO -
+      Power(AA,6)*Power(BB,3)*CC*DD*VV)));
+\end{multline}
+
+\begin{multline}
+  d_uma = (-X2 + X3)/CC + (GG*(-(BB*LL*PP*X1) + AA*MM*PP*X1 +
+    EE*LL*MM*X2 - AA*MM*PP*X2 + BB*EE*LL*X5 - AA*EE*MM*X5 -
+    BB*EE*LL*X6 + AA*BB*PP*X6))/ (CC*(EE*GG*LL*MM - BB*FF*LL*PP +
+    AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ -
+    BB*EE*LL*RR + AA*BB*PP*RR)) - (((-HH + II)/CC - (GG*(EE*LL -
+    AA*PP)*(-(HH*MM) + BB*SS))/(CC*(EE*GG*LL*MM - BB*FF*LL*PP +
+    AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ -
+    BB*EE*LL*RR + AA*BB*PP*RR)))* (-(BB*GG*LL*NN*PP*X1) +
+    AA*GG*MM*NN*PP*X1 + BB*CC*LL*PP*RR*X1 - AA*CC*MM*PP*RR*X1 +
+    BB*FF*LL*NN*PP*X2 - AA*FF*MM*NN*PP*X2 - BB*EE*LL*NN*QQ*X2 +
+    AA*EE*MM*NN*QQ*X2 - CC*EE*LL*MM*RR*X2 + BB*EE*LL*NN*RR*X2 +
+    AA*CC*MM*PP*RR*X2 - AA*BB*NN*PP*RR*X2 + EE*GG*LL*MM*NN*X3 -
+    BB*FF*LL*NN*PP*X3 + AA*FF*MM*NN*PP*X3 - AA*GG*MM*NN*PP*X3 +
+    BB*EE*LL*NN*QQ*X3 - AA*EE*MM*NN*QQ*X3 - BB*EE*LL*NN*RR*X3 +
+    AA*BB*NN*PP*RR*X3 + BB*EE*GG*LL*NN*X5 - AA*EE*GG*MM*NN*X5 -
+    BB*CC*EE*LL*RR*X5 + AA*CC*EE*MM*RR*X5 + CC*EE*GG*LL*MM*X6 -
+    BB*EE*GG*LL*NN*X6 - BB*CC*FF*LL*PP*X6 + AA*CC*FF*MM*PP*X6 -
+    AA*CC*GG*MM*PP*X6 + AA*BB*GG*NN*PP*X6 + BB*CC*EE*LL*QQ*X6 -
+    AA*CC*EE*MM*QQ*X6 - CC*EE*GG*LL*MM*X7 + BB*CC*FF*LL*PP*X7 -
+    AA*CC*FF*MM*PP*X7 + AA*CC*GG*MM*PP*X7 - BB*CC*EE*LL*QQ*X7 +
+    AA*CC*EE*MM*QQ*X7 + BB*CC*EE*LL*RR*X7 - AA*BB*CC*PP*RR*X7))/
+    (EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP -
+    AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP -
+    BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ -
+    AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR -
+    BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR +
+    AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS -
+    BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS +
+    AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS -
+    CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT +
+    AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT +
+    BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT) - ((JJ/CC + ((EE*GG*LL*MM -
+    BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ -
+    AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)* ((-HH + II)/CC -
+    (GG*(EE*LL - AA*PP)*(-(HH*MM) + BB*SS))/(CC*(EE*GG*LL*MM -
+    BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ -
+    AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)))* (-(JJ*NN) +
+    CC*UU))/(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP
+    - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP -
+    BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ -
+    AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR -
+    BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR +
+    AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS -
+    BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS +
+    AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS -
+    CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT +
+    AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT +
+    BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT))*
+    (-(((Power(AA,6)*Power(BB,2)*CC*DD*HH*MM -
+    Power(AA,6)*Power(BB,3)*DD*HH*NN +
+    Power(AA,6)*Power(BB,3)*DD*II*NN -
+    Power(AA,6)*Power(BB,3)*CC*II*OO)*(-(EE*LL) + AA*PP)*
+    (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+    Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+    AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+    Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+    Power(AA,2)*BB*RR))) - (-(EE*LL) +
+    AA*PP)*((Power(AA,5)*Power(BB,3)*CC*DD*FF*LL -
+    Power(AA,6)*Power(BB,2)*CC*DD*FF*MM +
+    Power(AA,6)*Power(BB,2)*CC*DD*GG*MM -
+    Power(AA,6)*Power(BB,3)*DD*GG*NN)* (-(EE*LL) + AA*PP) -
+    (Power(AA,5)*Power(BB,3)*CC*DD*EE*LL -
+    Power(AA,6)*Power(BB,2)*CC*DD*EE*MM)*(-(FF*LL) +
+    AA*QQ))*(Power(AA,4)*BB*CC*HH*MM -
+    Power(AA,4)*Power(BB,2)*CC*SS))*
+    (-(((Power(AA,3)*Power(BB,2)*CC*FF*LL - Power(AA,4)*BB*CC*FF*MM +
+    Power(AA,4)*BB*CC*GG*MM - Power(AA,4)*Power(BB,2)*GG*NN)*(-(EE*LL)
+    + AA*PP) - (Power(AA,3)*Power(BB,2)*CC*EE*LL -
+    Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ))*
+    (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+    Power(AA,4)*BB*CC*EE*MM)*(-(LL*X1) + AA*X5)) + (-(EE*LL) +
+    AA*PP)*(Power(AA,3)*Power(BB,2)*CC*LL*X1 - Power(AA,4)*BB*CC*MM*X1
+    + Power(AA,4)*BB*CC*MM*X2 - Power(AA,4)*Power(BB,2)*CC*X6))) +
+    (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+    Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+    AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+    Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+    Power(AA,2)*BB*RR)))* (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+    Power(AA,4)*BB*CC*EE*MM)*(-(LL*X1) + AA*X5)) + (-(EE*LL) +
+    AA*PP)*(Power(AA,3)*Power(BB,2)*CC*LL*X1 - Power(AA,4)*BB*CC*MM*X1
+    + Power(AA,4)*BB*CC*MM*X2 - Power(AA,4)*Power(BB,2)*NN*X2 +
+    Power(AA,4)*Power(BB,2)*NN*X3 - Power(AA,4)*Power(BB,2)*CC*X7))))
+    + (-((-(EE*LL) + AA*PP)*((Power(AA,3)*Power(BB,2)*CC*FF*LL -
+    Power(AA,4)*BB*CC*FF*MM + Power(AA,4)*BB*CC*GG*MM -
+    Power(AA,4)*Power(BB,2)*GG*NN)* (-(EE*LL) + AA*PP) -
+    (Power(AA,3)*Power(BB,2)*CC*EE*LL -
+    Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) +
+    AA*QQ))*(Power(AA,4)*BB*CC*HH*MM - Power(AA,4)*Power(BB,2)*CC*SS))
+    + (-(EE*LL) + AA*PP)*(-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+    Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+    AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+    Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+    Power(AA,2)*BB*RR)))* (Power(AA,4)*BB*CC*HH*MM -
+    Power(AA,4)*Power(BB,2)*HH*NN + Power(AA,4)*Power(BB,2)*II*NN -
+    Power(AA,4)*Power(BB,2)*CC*TT))*
+    (-(((Power(AA,5)*Power(BB,3)*CC*DD*FF*LL -
+    Power(AA,6)*Power(BB,2)*CC*DD*FF*MM +
+    Power(AA,6)*Power(BB,2)*CC*DD*GG*MM -
+    Power(AA,6)*Power(BB,3)*DD*GG*NN)*(-(EE*LL) + AA*PP) -
+    (Power(AA,5)*Power(BB,3)*CC*DD*EE*LL -
+    Power(AA,6)*Power(BB,2)*CC*DD*EE*MM)*(-(FF*LL) + AA*QQ))*
+    (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+    Power(AA,4)*BB*CC*EE*MM)*(-(LL*X1) + AA*X5)) + (-(EE*LL) +
+    AA*PP)*(Power(AA,3)*Power(BB,2)*CC*LL*X1 - Power(AA,4)*BB*CC*MM*X1
+    + Power(AA,4)*BB*CC*MM*X2 - Power(AA,4)*Power(BB,2)*CC*X6))) +
+    (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+    Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+    AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+    Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+    Power(AA,2)*BB*RR)))* (-((Power(AA,5)*Power(BB,3)*CC*DD*EE*LL -
+    Power(AA,6)*Power(BB,2)*CC*DD*EE*MM)*(-(LL*X1) + AA*X5)) +
+    (-(EE*LL) + AA*PP)*(Power(AA,5)*Power(BB,3)*CC*DD*LL*X1 -
+    Power(AA,6)*Power(BB,2)*CC*DD*MM*X1 +
+    Power(AA,6)*Power(BB,2)*CC*DD*MM*X2 -
+    Power(AA,6)*Power(BB,3)*DD*NN*X2 +
+    Power(AA,6)*Power(BB,3)*DD*NN*X3 -
+    Power(AA,6)*Power(BB,3)*CC*OO*X3 +
+    Power(AA,6)*Power(BB,3)*CC*OO*X4 -
+    Power(AA,6)*Power(BB,3)*CC*DD*X8)))))/ (-((-(EE*LL) +
+    AA*PP)*(-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+    Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+    AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+    Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+    Power(AA,2)*BB*RR)))* ((Power(AA,6)*Power(BB,2)*CC*DD*HH*MM -
+    Power(AA,6)*Power(BB,3)*DD*HH*NN +
+    Power(AA,6)*Power(BB,3)*DD*II*NN -
+    Power(AA,6)*Power(BB,3)*CC*II*OO)*(-(EE*LL) + AA*PP)*
+    (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+    Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+    AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+    Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+    Power(AA,2)*BB*RR))) - (-(EE*LL) +
+    AA*PP)*((Power(AA,5)*Power(BB,3)*CC*DD*FF*LL -
+    Power(AA,6)*Power(BB,2)*CC*DD*FF*MM +
+    Power(AA,6)*Power(BB,2)*CC*DD*GG*MM -
+    Power(AA,6)*Power(BB,3)*DD*GG*NN)* (-(EE*LL) + AA*PP) -
+    (Power(AA,5)*Power(BB,3)*CC*DD*EE*LL -
+    Power(AA,6)*Power(BB,2)*CC*DD*EE*MM)*(-(FF*LL) +
+    AA*QQ))*(Power(AA,4)*BB*CC*HH*MM -
+    Power(AA,4)*Power(BB,2)*CC*SS))* (Power(AA,4)*Power(BB,2)*JJ*NN -
+    Power(AA,4)*Power(BB,2)*CC*UU)) + (-(EE*LL) + AA*PP)*
+    (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+    Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+    AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+    Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+    Power(AA,2)*BB*RR)))* (-((-(EE*LL) +
+    AA*PP)*((Power(AA,3)*Power(BB,2)*CC*FF*LL -
+    Power(AA,4)*BB*CC*FF*MM + Power(AA,4)*BB*CC*GG*MM -
+    Power(AA,4)*Power(BB,2)*GG*NN)*(-(EE*LL) + AA*PP) -
+    (Power(AA,3)*Power(BB,2)*CC*EE*LL -
+    Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) +
+    AA*QQ))*(Power(AA,4)*BB*CC*HH*MM - Power(AA,4)*Power(BB,2)*CC*SS))
+    + (-(EE*LL) + AA*PP)*(-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+    Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+    AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+    Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+    Power(AA,2)*BB*RR)))* (Power(AA,4)*BB*CC*HH*MM -
+    Power(AA,4)*Power(BB,2)*HH*NN + Power(AA,4)*Power(BB,2)*II*NN -
+    Power(AA,4)*Power(BB,2)*CC*TT))* (Power(AA,6)*Power(BB,3)*DD*JJ*NN
+    - Power(AA,6)*Power(BB,3)*CC*JJ*OO +
+    Power(AA,6)*Power(BB,3)*CC*KK*OO -
+    Power(AA,6)*Power(BB,3)*CC*DD*VV));
+\end{multline}
+
+\begin{multline}
+  d_uba = (-X3 + X4)/DD + (II*(-(BB*GG*LL*NN*PP*X1) +
+        AA*GG*MM*NN*PP*X1 + BB*CC*LL*PP*RR*X1 - AA*CC*MM*PP*RR*X1 +
+        BB*FF*LL*NN*PP*X2 - AA*FF*MM*NN*PP*X2 - BB*EE*LL*NN*QQ*X2 +
+        AA*EE*MM*NN*QQ*X2 - CC*EE*LL*MM*RR*X2 + BB*EE*LL*NN*RR*X2 +
+        AA*CC*MM*PP*RR*X2 - AA*BB*NN*PP*RR*X2 + EE*GG*LL*MM*NN*X3 -
+        BB*FF*LL*NN*PP*X3 + AA*FF*MM*NN*PP*X3 - AA*GG*MM*NN*PP*X3 +
+        BB*EE*LL*NN*QQ*X3 - AA*EE*MM*NN*QQ*X3 - BB*EE*LL*NN*RR*X3 +
+        AA*BB*NN*PP*RR*X3 + BB*EE*GG*LL*NN*X5 - AA*EE*GG*MM*NN*X5 -
+        BB*CC*EE*LL*RR*X5 + AA*CC*EE*MM*RR*X5 + CC*EE*GG*LL*MM*X6 -
+        BB*EE*GG*LL*NN*X6 - BB*CC*FF*LL*PP*X6 + AA*CC*FF*MM*PP*X6 -
+        AA*CC*GG*MM*PP*X6 + AA*BB*GG*NN*PP*X6 + BB*CC*EE*LL*QQ*X6 -
+        AA*CC*EE*MM*QQ*X6 - CC*EE*GG*LL*MM*X7 + BB*CC*FF*LL*PP*X7 -
+        AA*CC*FF*MM*PP*X7 + AA*CC*GG*MM*PP*X7 - BB*CC*EE*LL*QQ*X7 +
+        AA*CC*EE*MM*QQ*X7 + BB*CC*EE*LL*RR*X7 - AA*BB*CC*PP*RR*X7))/
+        (DD*(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP
+        - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP -
+        BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ -
+        AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR -
+        BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR +
+        AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS -
+        BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS +
+        AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS -
+        CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT +
+        AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT +
+        BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT)) - (((-JJ + KK)/DD -
+        (II*(EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP +
+        BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR +
+        AA*BB*PP*RR)*(-(JJ*NN) + CC*UU))/ (DD*(EE*GG*II*LL*MM*NN +
+        BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP +
+        AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ +
+        BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ -
+        CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR +
+        AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR +
+        CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS +
+        AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS +
+        BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT +
+        BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT -
+        BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT -
+        AA*BB*CC*PP*RR*TT)))* (-(((Power(AA,6)*Power(BB,2)*CC*DD*HH*MM
+        - Power(AA,6)*Power(BB,3)*DD*HH*NN +
+        Power(AA,6)*Power(BB,3)*DD*II*NN -
+        Power(AA,6)*Power(BB,3)*CC*II*OO)*(-(EE*LL) + AA*PP)*
+        (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+        Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+        AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+        Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+        Power(AA,2)*BB*RR))) - (-(EE*LL) +
+        AA*PP)*((Power(AA,5)*Power(BB,3)*CC*DD*FF*LL -
+        Power(AA,6)*Power(BB,2)*CC*DD*FF*MM +
+        Power(AA,6)*Power(BB,2)*CC*DD*GG*MM -
+        Power(AA,6)*Power(BB,3)*DD*GG*NN)* (-(EE*LL) + AA*PP) -
+        (Power(AA,5)*Power(BB,3)*CC*DD*EE*LL -
+        Power(AA,6)*Power(BB,2)*CC*DD*EE*MM)*(-(FF*LL) +
+        AA*QQ))*(Power(AA,4)*BB*CC*HH*MM -
+        Power(AA,4)*Power(BB,2)*CC*SS))*
+        (-(((Power(AA,3)*Power(BB,2)*CC*FF*LL -
+        Power(AA,4)*BB*CC*FF*MM + Power(AA,4)*BB*CC*GG*MM -
+        Power(AA,4)*Power(BB,2)*GG*NN)*(-(EE*LL) + AA*PP) -
+        (Power(AA,3)*Power(BB,2)*CC*EE*LL -
+        Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ))*
+        (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+        Power(AA,4)*BB*CC*EE*MM)*(-(LL*X1) + AA*X5)) + (-(EE*LL) +
+        AA*PP)*(Power(AA,3)*Power(BB,2)*CC*LL*X1 -
+        Power(AA,4)*BB*CC*MM*X1 + Power(AA,4)*BB*CC*MM*X2 -
+        Power(AA,4)*Power(BB,2)*CC*X6))) +
+        (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+        Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+        AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+        Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+        Power(AA,2)*BB*RR)))* (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+        Power(AA,4)*BB*CC*EE*MM)*(-(LL*X1) + AA*X5)) + (-(EE*LL) +
+        AA*PP)*(Power(AA,3)*Power(BB,2)*CC*LL*X1 -
+        Power(AA,4)*BB*CC*MM*X1 + Power(AA,4)*BB*CC*MM*X2 -
+        Power(AA,4)*Power(BB,2)*NN*X2 + Power(AA,4)*Power(BB,2)*NN*X3
+        - Power(AA,4)*Power(BB,2)*CC*X7)))) + (-((-(EE*LL) +
+        AA*PP)*((Power(AA,3)*Power(BB,2)*CC*FF*LL -
+        Power(AA,4)*BB*CC*FF*MM + Power(AA,4)*BB*CC*GG*MM -
+        Power(AA,4)*Power(BB,2)*GG*NN)* (-(EE*LL) + AA*PP) -
+        (Power(AA,3)*Power(BB,2)*CC*EE*LL -
+        Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) +
+        AA*QQ))*(Power(AA,4)*BB*CC*HH*MM -
+        Power(AA,4)*Power(BB,2)*CC*SS)) + (-(EE*LL) +
+        AA*PP)*(-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+        Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+        AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+        Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+        Power(AA,2)*BB*RR)))* (Power(AA,4)*BB*CC*HH*MM -
+        Power(AA,4)*Power(BB,2)*HH*NN + Power(AA,4)*Power(BB,2)*II*NN
+        - Power(AA,4)*Power(BB,2)*CC*TT))*
+        (-(((Power(AA,5)*Power(BB,3)*CC*DD*FF*LL -
+        Power(AA,6)*Power(BB,2)*CC*DD*FF*MM +
+        Power(AA,6)*Power(BB,2)*CC*DD*GG*MM -
+        Power(AA,6)*Power(BB,3)*DD*GG*NN)*(-(EE*LL) + AA*PP) -
+        (Power(AA,5)*Power(BB,3)*CC*DD*EE*LL -
+        Power(AA,6)*Power(BB,2)*CC*DD*EE*MM)*(-(FF*LL) + AA*QQ))*
+        (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+        Power(AA,4)*BB*CC*EE*MM)*(-(LL*X1) + AA*X5)) + (-(EE*LL) +
+        AA*PP)*(Power(AA,3)*Power(BB,2)*CC*LL*X1 -
+        Power(AA,4)*BB*CC*MM*X1 + Power(AA,4)*BB*CC*MM*X2 -
+        Power(AA,4)*Power(BB,2)*CC*X6))) +
+        (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+        Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+        AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+        Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+        Power(AA,2)*BB*RR)))* (-((Power(AA,5)*Power(BB,3)*CC*DD*EE*LL
+        - Power(AA,6)*Power(BB,2)*CC*DD*EE*MM)*(-(LL*X1) + AA*X5)) +
+        (-(EE*LL) + AA*PP)*(Power(AA,5)*Power(BB,3)*CC*DD*LL*X1 -
+        Power(AA,6)*Power(BB,2)*CC*DD*MM*X1 +
+        Power(AA,6)*Power(BB,2)*CC*DD*MM*X2 -
+        Power(AA,6)*Power(BB,3)*DD*NN*X2 +
+        Power(AA,6)*Power(BB,3)*DD*NN*X3 -
+        Power(AA,6)*Power(BB,3)*CC*OO*X3 +
+        Power(AA,6)*Power(BB,3)*CC*OO*X4 -
+        Power(AA,6)*Power(BB,3)*CC*DD*X8)))))/ (-((-(EE*LL) +
+        AA*PP)*(-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+        Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+        AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+        Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+        Power(AA,2)*BB*RR)))* ((Power(AA,6)*Power(BB,2)*CC*DD*HH*MM -
+        Power(AA,6)*Power(BB,3)*DD*HH*NN +
+        Power(AA,6)*Power(BB,3)*DD*II*NN -
+        Power(AA,6)*Power(BB,3)*CC*II*OO)*(-(EE*LL) + AA*PP)*
+        (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+        Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+        AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+        Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+        Power(AA,2)*BB*RR))) - (-(EE*LL) +
+        AA*PP)*((Power(AA,5)*Power(BB,3)*CC*DD*FF*LL -
+        Power(AA,6)*Power(BB,2)*CC*DD*FF*MM +
+        Power(AA,6)*Power(BB,2)*CC*DD*GG*MM -
+        Power(AA,6)*Power(BB,3)*DD*GG*NN)* (-(EE*LL) + AA*PP) -
+        (Power(AA,5)*Power(BB,3)*CC*DD*EE*LL -
+        Power(AA,6)*Power(BB,2)*CC*DD*EE*MM)*(-(FF*LL) +
+        AA*QQ))*(Power(AA,4)*BB*CC*HH*MM -
+        Power(AA,4)*Power(BB,2)*CC*SS))*
+        (Power(AA,4)*Power(BB,2)*JJ*NN -
+        Power(AA,4)*Power(BB,2)*CC*UU)) + (-(EE*LL) + AA*PP)*
+        (-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+        Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+        AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+        Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+        Power(AA,2)*BB*RR)))* (-((-(EE*LL) +
+        AA*PP)*((Power(AA,3)*Power(BB,2)*CC*FF*LL -
+        Power(AA,4)*BB*CC*FF*MM + Power(AA,4)*BB*CC*GG*MM -
+        Power(AA,4)*Power(BB,2)*GG*NN)*(-(EE*LL) + AA*PP) -
+        (Power(AA,3)*Power(BB,2)*CC*EE*LL -
+        Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) +
+        AA*QQ))*(Power(AA,4)*BB*CC*HH*MM -
+        Power(AA,4)*Power(BB,2)*CC*SS)) + (-(EE*LL) +
+        AA*PP)*(-((Power(AA,3)*Power(BB,2)*CC*EE*LL -
+        Power(AA,4)*BB*CC*EE*MM)*(-(FF*LL) + AA*QQ)) + (-(EE*LL) +
+        AA*PP)*(Power(AA,4)*BB*CC*GG*MM -
+        Power(AA,2)*BB*CC*(-(AA*BB*FF*LL) + Power(AA,2)*FF*MM +
+        Power(AA,2)*BB*RR)))* (Power(AA,4)*BB*CC*HH*MM -
+        Power(AA,4)*Power(BB,2)*HH*NN + Power(AA,4)*Power(BB,2)*II*NN
+        - Power(AA,4)*Power(BB,2)*CC*TT))*
+        (Power(AA,6)*Power(BB,3)*DD*JJ*NN -
+        Power(AA,6)*Power(BB,3)*CC*JJ*OO +
+        Power(AA,6)*Power(BB,3)*CC*KK*OO -
+        Power(AA,6)*Power(BB,3)*CC*DD*VV));
+\end{multline}
+
 \section{Questions for Tate/Nicole}

### 2017-01-10 thesis/latex/thesis.tex
Add work on methods. Add full skeleton.

--- 2017-01-08 thesis/latex/thesis.tex
@@
 \setcounter{secnumdepth}{4}
+
+\setlength{\emergencystretch}{3cm} % somehow helps prevent hboxes from overflowing
 
@@
 
-\subsection{Verifying our model}
-Talk about why we think our model accurately represents the system we are trying to model, and
-doesn't have bugs in it.\\
+\section{Chapter: Methods}
 
-Talk about this figure:
+\subsection{Dynein Models}
+In order to simulate a protein's passage through the above mechanochemical cycle, two mathematical models are needed: one for the prestroke, MTBD-detached state, and one for the poststroke, MTBD-bound state.
+The purpose of each model is to calculate the position and velocity of each domain at an arbitrary time.\\
 
-\begin{figure}[h!]
+\subsubsection{Prestroke onebound model}
+Prestroke dynein is referred to as onebound, since one motor domain is bound to the MT in this state, and the other unbound.
+A spatial representation of onebound dynein is shown in Figure \ref{ob_fig}.
+
+\begin{figure}[h]
   \centering
-  \begin{minipage}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{../figures/OB_Average_PE.pdf}
-    \caption{Onebound PE vs time.}
-  \end{minipage}
-  \begin{minipage}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{../figures/BB_Average_PE.pdf}
-    \caption{Bothbound PE vs time.}
-  \end{minipage}
-  \label{fig:equipartition_agreement}
+  \includegraphics[width=.45\textwidth]{../figures/ob_fig.pdf}
+  \caption{Spatial representation of onebound model.}
+  \label{ob_fig}
 \end{figure}
 
-and how it shows our model is physical, or at least, how obeying the equipartition theorem shows
-how our domains both feel harmonic energies and can fully explore their spaces?\\
+TODO: make figure, label domain names (MTBD motor tail), Ls, Lt, microtubule\\
 
-\subsection{Parameters}
-The following table shows our guess parameters from various papers/crystal structures.\\
+Each domain is represented by a point, and is connected to other domains via fixed-length rods.
+Angles represent the degrees of freedom of the model.
+Since the protein is a dimer, there are two copies of the MTB, motor and tail domains.
+The tails are fused together and so represented as a single point.
+The motor and MTB domains are given a ``bound'' and ``unbound'' designation to differentiate them.
+The ``bound'' domains are the MTBD attached to the MT and its corresponding motor domain.
+The unbound domains are the unbound MTBD and its corresponding motor domain.\\
 
-                        \begin{center}
-  \begin{tabular}{| l | l | l | l | l | l |}
-    \hline
-    & Winch (Sarlah) & Schmidt & Lin & PyMol 3VKH & PyMol 4RH7\\\hline
-    $L_s$ & 12nm & & & 21.0nm & 22.1nm\\ \hline
-    $L_t$ &  7nm & & & & 11.15nm\\ \hline
-    $R_b$ &  & & & 1.57nm & 1.45nm\\ \hline
-    $R_m$ &  7nm & & & 7.36nm & 6.3nm\\ \hline
-    $R_t$ &  & & & & 2.16nm \\ \hline
-    $\theta_{m}^{Pre}$ & 250$^{\circ}$ & & 171$^{\circ}$ & &\\ \hline
-    $\theta_{m}^{Post}$ & 330$^{\circ}$ & & 137.5$^{\circ}$ & &\\ \hline
-    $\theta_{b}$ & 56$^{\circ}$ & & 63.5$^{\circ}$ & &\\ \hline
-    $\theta_{t}$ & 0$^{\circ}$ & & & &\\ \hline
-    $k_{ub}$ & $180 s^{-1,a}$ & & & &\\ \hline
-    $k_b$ & $5000 s^{-1,b}$ & & & &\\ \hline
+\paragraph{Angular and Cartesian coordinates}
+
+The four angles $\Theta_{bb}, \Theta_{bm}, \Theta_{um}$ and $\Theta_{ub}$, corresponding to bound binding, bound motor, unbound motor and unbound binding angle, together describe any possible conformation the system can take on.
+The domain of $\Theta_{bb}$ has a restricted domain of $[0,\pi]$ to prevent below-MT conformations, but the other angles have domains of $[0,2\pi]$.
+Each angle is defined relative to the horizontal axis.\\
+
+\begin{figure}[h]
+  \centering
+  \begin{tabular}{c}
+    \begin{lstlisting}[language=C++]
+      class Dynein_onebound{
+        double &$\Theta_{bb}, \Theta_{bb}, \Theta_{bb}, \Theta_{bb}$&
+        double &$\dot{\Theta}_{bb}, \dot{\Theta}_{bb}, \dot{\Theta}_{bb}, \dot{\Theta}_{bb}$&
+        double &$X_{bb}$&
+      }
+    \end{lstlisting}
   \end{tabular}
-\end{center}
-		\subsubsection{Lengths, angles, binding constants, etc…}
-	\subsection{Tuning our model}
-	\subsubsection{Searching for optimal parameters}
+  \caption{Pseudocode of the onebound Dynein model.}
+  \label{ob_struct}
+\end{figure}
+
+Pseudocode for the numerical onebound model representation is shown in Figure \ref{ob_struct}.
+The position of the bound binding domain $X_{bb}$ and the four domain angles are all the values needed to determine the position of the protein at any time.\\
+
+\paragraph{$\dot{\Theta}_{ub}, \dot{\Theta}_{um}, \dot{\Theta}_{bm}, \dot{\Theta}_{bb}$ calculation}
+
+The following equations allow recursive calculation of the positions of each domain in cartesian coordinates:
+
+\noindent\begin{minipage}{0.49\linewidth}
+\begin{align}
+  X_{bm} &= X_{bb}+L_{s}\cos(\Theta_{bb}) \\
+  X_{t}  &= X_{bm}+L_{t}\cos(\Theta_{bm}) \\
+  X_{fm} &= X_{t} - L_{t}\cos(\Theta_{fm}) \\
+  X_{fb} &= X_{fm} - L_{s}\cos(\Theta_{fb})
+\end{align}
+\end{minipage}
+\begin{minipage}{0.49\linewidth}
+\begin{align}
+  Y_{bb} &= 0 \\
+  Y_{bm} &= Y_{bb}+L_{s}\sin(\Theta_{bb}) \\
+  Y_{t}  &= Y_{bm}+L_{t}\sin(\Theta_{bm}) \\
+  Y_{fm} &= Y_{t} - L_{t}\sin(\Theta_{fm}) \\
+  Y_{fb} &= Y_{fm} - L_{s}\sin(\Theta_{fb})
+\end{align}
+\end{minipage}
+\vspace{.5cm}
+
+$L_s$ and $L_t$ correspond to lengths of each interdomain rod, and the $t$ subscripts refer to tail domain coordinates.\\
+
+\paragraph{Angular and Cartesian velocities}
+The goal is to express angular velocities $\dot{\Theta}_{bb}, \dot{\Theta}_{bb}, \dot{\Theta}_{bb}$ and $\dot{\Theta}_{bb}$ in terms of known quantities.
+To begin, the cartesian velocities of each domain can be calculated from the above position equations in a similar recursive manner:
+
+\noindent\begin{minipage}{0.49\linewidth}
+\begin{align}
+  \dot{X}_{bb} &= 0 \\
+  \dot{X}_{bm} &= \dot{X}_{bb} - L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} \label{cartesian-bmx}\\
+  \dot{X}_{t } &= \dot{X}_{bm} - L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} \\
+  \dot{X}_{fm} &= \dot{X}_{t } + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} \\
+  \dot{X}_{fb} &= \dot{X}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb}
+\end{align}
+\end{minipage}
+\begin{minipage}{0.49\linewidth}
+\begin{align}                                                                          
+  \dot{Y}_{bb} &= 0 \\                                                        
+  \dot{Y}_{bm} &= \dot{Y}_{bb} + L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} \\
+  \dot{Y}_{t}  &= \dot{Y}_{bm} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} \\
+  \dot{Y}_{fm} &= \dot{Y}_{t } - L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} \\
+  \dot{Y}_{fb} &= \dot{Y}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb}
+\end{align}
+\end{minipage}
+\vspace{.5cm}
+
+Another way to express these cartesian velocities is using the Brownian dynamics equation $\dot{X} = \frac1\gamma F_{net} + R$:
+
+\begin{align}  
+  \dot{X}_{bm} &= \frac{1}{\gamma_m} \Big(F_{xml} + - \lambda_{bs}(X_{bm} - X_{bb}) + \lambda_{bt}(X_{t } - X_{bm}) \Big) + R_{xml} \label{brownian-bmx}\\
+  \dot{X}_{t } &= \frac{1}{\gamma_t} \Big(F_{xt } + - \lambda_{bt}(X_{t } - X_{bm}) + \lambda_{ft}(X_{fm} - X_{t }) \Big) + R_{xt } \\
+  \dot{X}_{fm} &= \frac{1}{\gamma_m} \Big(F_{xmr} + - \lambda_{ft}(X_{fm} - X_{t }) + \lambda_{fs}(X_{fb} - X_{fm}) \Big) + R_{xmr} \\
+  \dot{X}_{fb} &= \frac{1}{\gamma_b} \Big(F_{xbr} + - \lambda_{fs}(X_{fb} - X_{fm}) \Big) + R_{xbr}
+\end{align}
+
+\begin{align}  
+  \dot{Y}_{bm} &= \frac{1}{\gamma_m} \Big(F_{yml} + - \lambda_{bs}(Y_{bm} - Y_{bb}) + \lambda_{bt}(Y_{t } - Y_{bm}) \Big) + R_{yml} \\
+  \dot{Y}_{t}  &= \frac{1}{\gamma_t} \Big(F_{yt } + - \lambda_{bt}(Y_{t } - Y_{bm}) + \lambda_{ft}(Y_{fm} - Y_{t }) \Big) + R_{yt } \\
+  \dot{Y}_{fm} &= \frac{1}{\gamma_m} \Big(F_{ymr} + - \lambda_{ft}(Y_{fm} - Y_{t }) + \lambda_{fs}(Y_{fb} - Y_{fm}) \Big) + R_{ymr} \\
+  \dot{Y}_{fb} &= \frac{1}{\gamma_b} \Big(F_{ybr} + - \lambda_{fs}(Y_{fb} - Y_{fm}) \Big) + R_{ybr}
+\end{align}
+
+where $\gamma_n$ is a drag coefficient for the binding, motor or tail domain with units of mass per second.
+$F_{xn}$ is the x component of external force on each domain $n$ due to various factors.
+$\lambda_{12}\left(X_1 - X_2\right)$ is the x component of internal force on domain $2$ due to domain $1$, where $\lambda$ is a tension coefficient with units of mass per second per second.
+$R_{xn}$ is the Brownian coefficient representing motion due to solvent collision, with units of velocity.\\
+
+These two sets of equations, cartesian and Brownian, can be equated to get more interesting equations.
+For example, Eq (\ref{cartesian-bmx}) and Eq (\ref{brownian-bmx}) can be equated.
+This equating, combined with expanding the recursive velocity definitions leads to:
+
+\begin{align}
+  -L_s\sin(\theta_{bb})\dot{\theta}_{bb} &= \frac{1}{\gamma_m}F_{xml} + -\frac{1}{\gamma_m}\lambda_{bs}(X_{bm} - X_{bb}) + \frac{1}{\gamma_m}\lambda_{bt}(X_{t } - X_{bm}) + R_{bmx} \label{ob_system_first}\\
+  -L_s\sin(\theta_{bb})\dot{\theta}_{bb} - L_t\sin(\theta_{bm})\dot{\theta}_{bm} &= \frac{1}{\gamma_t}F_{xt } + -\frac{1}{\gamma_t}\lambda_{bt}(X_{t } - X_{bm}) + \frac{1}{\gamma_t}\lambda_{ft}(X_{um} - X_{t }) + R_{tx} \\
+  -L_s\sin(\theta_{bb})\dot{\theta}_{bb} - L_t\sin(\theta_{bm})\dot{\theta}_{bm} + L_t\sin(\theta_{um})\dot{\theta}_{um} &= \frac{1}{\gamma_m}F_{xmr} + -\frac{1}{\gamma_m}\lambda_{ft}(X_{um} - X_{t }) + \frac{1}{\gamma_m}\lambda_{fs}(X_{ub} - X_{um}) + R_{bmx} \\
+  -L_s\sin(\theta_{bb})\dot{\theta}_{bb} - L_t\sin(\theta_{bm})\dot{\theta}_{bm} + L_t\sin(\theta_{um})\dot{\theta}_{um} + L_s\sin(\theta_{ub})\dot{\theta}_{ub} &= \frac{1}{\gamma_b}F_{xbr} + -\frac{1}{\gamma_b}\lambda_{fs}(X_{ub} - X_{um}) + R_{bbx}
+\end{align}
+
+\begin{align}
+  L_s\cos(\theta_{bb})\dot{\theta}_{bm} &= \frac{1}{\gamma_m}F_{yml} + -\frac{1}{\gamma_m}\lambda_{bs}(Y_{bm} - Y_{bb}) + \frac{1}{\gamma_m}\lambda_{bt}(Y_{t } - Y_{bm}) + R_{umy} \\
+  L_s\cos(\theta_{bb})\dot{\theta}_{bb} + L_t\cos(\theta_{bm})\dot{\theta}_{bm} &= \frac{1}{\gamma_t}F_{yt } + -\frac{1}{\gamma_t}\lambda_{bt}(Y_{t } - Y_{bm}) + \frac{1}{\gamma_t}\lambda_{ft}(Y_{um} - Y_{t }) + R_{ty} \\
+  L_s\cos(\theta_{bb})\dot{\theta}_{bb} + L_t\cos(\theta_{bm})\dot{\theta}_{bm} - L_t\cos(\theta_{um})\dot{\theta}_{um} &= \frac{1}{\gamma_m}F_{ymr} + -\frac{1}{\gamma_m}\lambda_{ft}(Y_{um} - Y_{t }) + \frac{1}{\gamma_m}\lambda_{fs}(Y_{ub} - Y_{um}) + R_{bmy} \\
+  L_s\cos(\theta_{bb})\dot{\theta}_{bb} + L_t\cos(\theta_{bm})\dot{\theta}_{bm} - L_t\cos(\theta_{um})\dot{\theta}_{um} - L_s\cos(\theta_{ub})\dot{\theta}_{ub} &= \frac{1}{\gamma_b}F_{ybr} + -\frac{1}{\gamma_b}\lambda_{fs}(Y_{ub} - Y_{um}) + R_{bby} \label{ob_system_last}
+\end{align}
+
+These eight equations form a system of equations with eight unknowns: $\dot{\Theta}_{bb}, \dot{\Theta}_{bb}, \dot{\Theta}_{bb}$ and $\dot{\Theta}_{bb}$, and the four tension coefficients $\lambda_{bs}, \lambda_{bt}, \lambda_{um}$ and $\lambda_{ub}$.
+This system is more compactly represented as:
+{\small
+  \[
+  \hspace{-1.0cm}
+  \resizebox{0.85\linewidth}{!}{%
+  $\begin{pmatrix}
+    L_s\sin\theta_{bb} & 0 & 0 & 0 & -\gamma_m (X_{bm} - X_{bb}) & \gamma_m (X_{t } - X_{bm}) & 0 & 0\\
+    L_s\sin(\theta_{bb}) & L_t\sin(\theta_{bm}) & 0 & 0 & 0 & -\gamma_t (X_{t } - X_{bm}) & \gamma_t (X_{um} - X_{t }) & 0\\
+    L_s\sin(\theta_{bb}) & L_t\sin(\theta_{bm}) & -L_t\sin(\theta_{um}) & 0 & 0 & 0 & -\gamma_m (X_{um} - X_{t }) & \gamma_m (X_{ub} - X_{um})\\
+    L_s\sin(\theta_{bb}) & L_t\sin(\theta_{bm}) & -L_t\sin(\theta_{um}) & -L_s\sin(\theta_{ub}) & 0 & 0 & 0 & -\gamma_b (X_{ub} - X_{um})\\
+    -L_s\cos(\theta_{bb}) & 0 & 0 & 0 & -\gamma_m (Y_{bm} - Y_{bb}) & \gamma_m (Y_{t } - Y_{bm}) & 0 & 0\\
+    -L_s\cos(\theta_{bb}) & -L_t\cos(\theta_{bm}) & 0 & 0 & 0 & -\gamma_t (Y_{t } - Y_{bm}) & \gamma_t (Y_{um} - Y_{t }) & 0\\
+    -L_s\cos(\theta_{bb}) & -L_t\cos(\theta_{bm}) & L_t\cos(\theta_{um}) & 0 & 0 & 0 & -\gamma_m (Y_{um} - Y_{t }) & \gamma_m (Y_{ub} - Y_{um})\\
+    -L_s\cos(\theta_{bb}) & -L_t\cos(\theta_{bm}) & L_t\cos(\theta_{um}) & L_s\cos(\theta_{ub}) & 0 & 0 & 0 & -\gamma_b (Y_{ub} - Y_{um})\\
+  \end{pmatrix}$%
+  }
+  \begin{pmatrix}
+    \dot{\theta}_{bb}\\
+    \dot{\theta}_{bm}\\
+    \dot{\theta}_{um}\\
+    \dot{\theta}_{ub}\\
+    \lambda_{bs}\\
+    \lambda_{bt}\\
+    \lambda_{ut}\\
+    \lambda_{us}\\
+  \end{pmatrix}
+  =
+  \begin{pmatrix}
+    -F_{bmx} + \gamma_m R_{bmx}\\
+    -F_{tx} + \gamma_t R_{tx}\\
+    -F_{umx} + \gamma_m R_{umx}\\
+    -F_{ubx} + \gamma_b R_{ubx}\\
+    -F_{bmy} + \gamma_m R_{bmy}\\
+    -F_{ty} + \gamma_t R_{ty}\\
+    -F_{umy} + \gamma_m R_{umy}\\
+    -F_{uby} + \gamma_b R_{uby}\\
+  \end{pmatrix}
+  \]
+}
+
+This matrix is then solved using the Mathematica computer algebra system, resulting in a set of $\dot{\Theta}_n$ equations shown in Appendix (\ref{ob-motion-equations]}).
+
+\subsubsection{Poststroke bothbound model}
+The main difference between the prestroke and poststroke mathematical models is that the latter has a further constraint: both its MTB domains are attached to the MT.
+This has the effect of diminishing the degrees of freedom of the system from four to two, assuming the motor remains above the MT.
+A spatial representation of the bothbound model is shown in Figure (\ref{bb_fig}).\\
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.45\textwidth]{../figures/code-bothbound.pdf}
+  \caption{Definition of equilibrium angles.}
+  \label{bb_fig}
+\end{figure}
+
+The bothbound model is very similar in assumptions to the onebound: domains are represented as points connected by rigid rods, and the tails are still fused into a single point.
+The main differences are the naming scheme and spatial constraints.
+Dimers can no longer be differentiated by being un/bound, so are now differentiated by their z-position: being near or far from the ``camera.'' This is similar to labeling them as the right or left dimer.
+Thus there are the near binding (nb), near motor (nm), tail (t), far motor (fm) and far binding (fb) domains.
+Both the near and far binding domains have their x-coordinate fixed at zero.
+The system is described completely by two angles: $\Theta_{nm}$ and $\Theta_{fm}$.
+Angles $\Theta_t, \Theta_{nb}$ and $\Theta_{fb}$ are dependent on these two motor angles.\\
+
+The second MTBD constraint adds another required piece of information the model must track: the x position of the second MTBD.
+This position is described by $L$, the distance between the near and far binding domains.
+The pseudocode used to describe the bothbound model are shown in Figure (\ref{bb_struct}).\\
+
+\begin{figure}[h]
+  \centering
+  \begin{tabular}{c}
+    \begin{lstlisting}[language=C++]
+      class Dynein_bothbound{
+        double &$\Theta_{nm}, \Theta_{fm}$&
+        double &$\dot{\Theta}_{nm}, \dot{\Theta}_{fm}$&
+        double &$X_{nb, L}$&
+      }
+    \end{lstlisting}
+  \end{tabular}
+  \caption{Pseudocode of the bothbound Dynein model.}
+  \label{ob_struct}
+\end{figure}
+
+\paragraph{$\dot{\Theta}_{nm}, \dot{\Theta}_{fm}$ calculation}
+
+The first step in calculating angular velocities is to define new variables $L_n$ and $L_f$ using the Law of Cosines:
+
+\begin{align}
+  L_n &= \sqrt{L_s^2 + L_t^2 - 2L_sL_t\cos{\theta_{nm}}} \\
+  L_f &= \sqrt{L_s^2 + L_t^2 - 2L_sL_t\cos{\theta_{fm}}}
+\end{align}
+
+These new artificial stalks are not physically relevant, but allow the definition of new angles $\Theta_n, \Theta_{ns}, \Theta_{f}$ and $\Theta_{fs}$, which are useful for calculating the position of the motor domain.
+The treatment of the near and far versions of these angles is very similar, so only the near angles will be dealt with here.
+The angles themselves are never dealt with, only their $\sin$s and $\cos$s:
+
+\begin{align}
+  \cos\theta_n &= \frac{L^2 + L_n^2 - L_f^2}{2L L_n} \\
+  \sin\theta_{n} &= \sqrt{1 - \cos^2\theta_{n}} \\
+  \cos\theta_{ns} &= \frac{L_s^2 + L_n^2 - L_t^2}{2L_s L_n} \\
+  \sin\theta_{ns} &=
+  \begin{cases}
+    +\sqrt{1 - \cos^2\theta_{ns}} & \theta_{nm} < \pi \\
+    -\sqrt{1 - \cos^2\theta_{ns}} & \theta_{nm} > \pi
+  \end{cases}
+\end{align}
+
+These values are arrived at through the Law of Cosines and a trigonometric identity.
+The sign of $\sin\left(\Theta_n\right)$ is restricted to positive values due to the angle's $[0,\pi]$ domain.
+$\Theta_{ns}$'s domain is $[-\pi,\pi]$, meaning its sign is not restricted.
+The partial derivatives of these values are useful:
+
+\begin{align}
+  \frac{\partial \cos\theta_n}{\partial L_n} &= \frac{1}{L} - \frac{L^2 + L_n^2 - L_f^2}{2L L_n^2}\\
+  \frac{\partial \cos\theta_n}{\partial L_f} &= -\frac{L_f}{LL_n}\\
+  \frac{\partial \sin\theta_n}{\partial L_n} &= \frac{-\cos\theta_n}{\sqrt{1-\cos^2\theta_n}}
+  \left(\frac{1}{L} - \frac{L^2 + L_n^2 - L_f^2}{2L L_n^2}\right)\\
+  \frac{\partial \sin\theta_n}{\partial L_f} &= \frac{-\cos\theta_n}{\sqrt{1-\cos^2\theta_n}}
+  \left(\frac{-L_f}{LLn}\right)\\
+  \frac{\partial \cos\theta_{ns}}{\partial L_n} &= \frac{1}{L_s}
+  - \frac{L_s^2 + L_n^2 - L_t^2}{2L_sL_n^2}\\
+  \frac{\partial \cos\theta_{ns}}{\partial L_f} &= 0\\
+  \frac{\partial \sin\theta_{ns}}{\partial L_n} &=
+  \begin{cases}
+    \frac{-\cos\theta_{ns}}{\sqrt{1-\cos^2\theta_{ns}}}
+    \left(\frac{1}{L_s} - \frac{L_s^2+L_n^2-L_t^2}{2LL_n^2}\right) & \theta_{nm} < \pi \\
+    \frac{\cos\theta_{ns}}{\sqrt{1-\cos^2\theta_{ns}}}
+    \left(\frac{1}{L_s} - \frac{L_s^2+L_n^2-L_t^2}{2LL_n^2}\right) & \theta_{nm} > \pi
+  \end{cases}\\
+  \frac{\partial \sin\theta_{ns}}{\partial L_f} &= 0
+\end{align}
+
+The position of the motor and tail domains is calculated as follws, using the angle addition trigonometric identity:
+
+\begin{align}
+  X_{nm} &= L_s\left(
+  \cos\theta_n\cos\theta_{ns} - \sin\theta_n\sin\theta_{ns}
+  \right)
+  \\
+  Y_{nm} &= L_s\left(
+  \cos\theta_n\sin\theta_{ns} + \sin\theta_n\cos\theta_{ns}
+  \right)
+  \\
+  X_{t} &= L_n\cos\theta_n\\
+  Y_{t} &= L_n\sin\theta_n\\
+\end{align}
+
+The partial derivatives of these values are:
+
+\begin{align}
+  \frac{dX_{nm}}{dL_n} &= L_s\Big(\cos\theta_n\frac{d\cos\theta_{ns}}{dL_n} + \cos\theta_{ns}\frac{d\cos\theta_{n}}{dL_n} - \sin\theta_n\frac{d\sin\theta_{ns}}{dL_n} - \sin\theta_{ns}\frac{d\sin\theta_{n}}{dL_n}\Big)\\
+  \frac{dY_{nm}}{dL_n} &= L_s\Big(\cos\theta_n\frac{d\sin\theta_{ns}}{dL_n} + \sin\theta_{ns}\frac{d\cos\theta_{n}}{dL_n} + \sin\theta_n\frac{d\cos\theta_{ns}}{dL_n} + \cos\theta_{ns}\frac{d\sin\theta_{n}}{dL_n} \Big)\\
+  \frac{dX_{nm}}{dL_f} &= L_s\Big(\cos\theta_n\frac{d\cos\theta_{ns}}{dL_f} + \cos\theta_{ns}\frac{d\cos\theta_{n}}{dL_f} - \sin\theta_n\frac{d\sin\theta_{ns}}{dL_f} - \sin\theta_{ns}\frac{d\sin\theta_{n}}{dL_f} \Big)\\
+  \frac{dY_{nm}}{dL_f} &= L_s\Big(\cos\theta_n\frac{d\sin\theta_{ns}}{dL_f} + \sin\theta_{ns}\frac{d\cos\theta_{n}}{dL_f} + \sin\theta_n\frac{d\cos\theta_{ns}}{dL_f} + \cos\theta_{ns}\frac{d\sin\theta_{n}}{dL_f}\Big)\\
+  \frac{dX_{t}}{dL_n} &= \cos\theta_n + L_n\frac{\partial \cos\theta_n}{\partial L_n} = \frac{L_n}{L}\\
+  \frac{dY_{t}}{dL_n} &= \sin\theta_n + L_n\frac{\partial \sin\theta_n}{\partial L_n}\\
+  \frac{dX_{t}}{dL_f} &= L_n\frac{\partial \cos\theta_n}{\partial L_f} = -\frac{L_f}{L}\\
+  \frac{dY_{t}}{dL_f} &= L_n\frac{\partial \sin\theta_n}{\partial L_f}\\
+\end{align}
+
+These partial derivatives are then used in a system of equations very similar to that in Equations (\ref{ob_system_first}-\ref{ob_system_last}), combining Brownian and coordinate system definitions of domain velocities:
+
+\begin{align}
+  \dot{X}_{nm} &= \frac{1}{\gamma} \Big(F_{xml} - \lambda_{ns}(X_{bm} - X_{bb})
+  + \lambda_{nt}(X_{t } - X_{bm}) \Big) + R_{xml}
+  &=& \frac{\partial X_{nm}}{\partial L_n}\dot{L}_n + \frac{\partial X_{nm}}{\partial L_f}\dot{L}_f\\
+  \dot{X}_{t } &= \frac{1}{\gamma} \Big(F_{xt } - \lambda_{nt}(X_{t } - X_{bm})
+  + \lambda_{ft}(X_{fm} - X_{t }) \Big) + R_{xt }
+  &=& \frac{\partial X_{t}}{\partial L_n}\dot{L}_n + \frac{\partial X_{t}}{\partial L_f}\dot{L}_f\\
+  \dot{X}_{fm} &= \frac{1}{\gamma} \Big(F_{xmr} - \lambda_{ft}(X_{fm} - X_{t })
+  + \lambda_{fs}(X_{fb} - X_{fm}) \Big) + R_{xmr}
+  &=& \frac{\partial X_{fm}}{\partial L_n}\dot{L}_n + \frac{\partial X_{fm}}{\partial L_f}\dot{L}_f
+\end{align}
+
+\begin{align}
+  \dot{Y}_{nm} &= \hspace{1cm} \frac{1}{\gamma} \Big(F_{yml} - \lambda_{ns}(Y_{bm} - Y_{bb})
+  + \lambda_{nt}(Y_{t } - Y_{bm}) \Big) + R_{yml}
+  &= \frac{\partial Y_{nm}}{\partial L_n}\dot{L}_n + \frac{\partial Y_{nm}}{\partial L_f}\dot{L}_f\\
+  \dot{Y}_{t}  &= \hspace{1cm} \frac{1}{\gamma} \Big(F_{yt } - \lambda_{nt}(Y_{t } - Y_{bm})
+  + \lambda_{ft}(Y_{fm} - Y_{t }) \Big) + R_{yt }
+  &= \frac{\partial Y_{t}}{\partial L_n}\dot{L}_n + \frac{\partial Y_{t}}{\partial L_f}\dot{L}_f\\
+  \dot{Y}_{fm} &= \hspace{1cm} \frac{1}{\gamma} \Big(F_{ymr} - \lambda_{ft}(Y_{fm} - Y_{t })
+  + \lambda_{fs}(Y_{fb} - Y_{fm}) \Big) + R_{ymr}
+  &= \frac{\partial Y_{fm}}{\partial L_n}\dot{L}_n + \frac{\partial Y_{fm}}{\partial L_f}\dot{L}_f
+\end{align}
+
+The matrix version of this system looks like:
+
+\[
+\begin{pmatrix}
+  -\frac{X_{nm}}{Ln} & -\frac{X_{nm}}{Lf}
+  & -\frac{X_{nm}-X_{nb}}{\gamma} & \frac{X_{t}-X_{nm}}{\gamma} & 0 & 0\\
+  -\frac{X_{t}}{Ln} & -\frac{X_{t}}{Lf}
+  & 0 & -\frac{X_{t}-X_{nm}}{\gamma} & \frac{X_{fm}-X_{t}}{\gamma} & 0\\
+  -\frac{X_{fm}}{Ln} & -\frac{X_{fm}}{Lf}
+  & 0 & 0 & -\frac{X_{fm}-X_{t}}{\gamma} & \frac{X_{fb}-X_{fm}}{\gamma}\\
+  -\frac{Y_{nm}}{Ln} & -\frac{Y_{nm}}{Lf}
+  & -\frac{Y_{nm}-Y_{nb}}{\gamma} & \frac{Y_{t}-Y_{nm}}{\gamma} & 0 & 0\\
+  -\frac{Y_{t}}{Ln} & -\frac{Y_{t}}{Lf}
+  & 0 & -\frac{Y_{t}-Y_{nm}}{\gamma} & \frac{Y_{fm}-Y_{t}}{\gamma} & 0\\
+  -\frac{Y_{fm}}{Ln} & -\frac{Y_{fm}}{Lf}
+  & 0 & 0 & -\frac{Y_{fm}-Y_{t}}{\gamma} & \frac{Y_{fb}-Y_{fm}}{\gamma}\\
+\end{pmatrix}
+\begin{pmatrix}
+  \dot{L}_n\\
+  \dot{L}_f\\
+  \lambda_{ns}\\
+  \lambda_{nt}\\
+  \lambda_{ft}\\
+  \lambda_{fs}\\
+\end{pmatrix}
+=
+\begin{pmatrix}
+  -\frac{1}{\gamma}F_{nmx} - R_{nmx}\\
+  -\frac{1}{\gamma}F_{tx}  - R_{tx}\\
+  -\frac{1}{\gamma}F_{fmx} - R_{fmx}\\
+  -\frac{1}{\gamma}F_{nmy} - R_{nmy}\\
+  -\frac{1}{\gamma}F_{ty}  - R_{ty}\\
+  -\frac{1}{\gamma}F_{fmy} - R_{fmy}\\
+\end{pmatrix}
+\]
+
+This system of equations is then solved and used to calculate $\dot{L}_n$ and $\dot{L}_f$, as shown in Appendix (\ref{both-motion-equations}).
+
+\subsection{Simulation}
+To generate stepping data, the models were coded in \verb|C++| and time-evolved using Euler's method to generate reasonable dynamical behavior.
+When in a position deemed worthy of a state transition a conversion, the model was converted to the respective other model, representing part of a ``step'' through the mechanochemical cycle.
+Some unphysical but minor corrections at points in the simulation were needed to rescue the model from entering incalculable positions.\\
+
+Code is available at \verb|https://github.com/elliotc12/dynein_walk.git|.
+
+\textit{todo: come up with a better word than ``model'' to describe the state of the system at any time}
+\subsubsection{Time evolution}
+Euler's method for solving differential equations was used to calculate domain positions through time.
+Because each model is a set of multiple coupled differential equations, care was taken to update each domain position at the same time, and not in a cascade.
+A code snippet is shown in Figure (\ref{euler-code-snippet}) for the onebound updating scheme.
+
+\begin{figure}[h]
+  \centering
+  \begin{tabular}{c}
+    \begin{lstlisting}[language=C++]
+      double temp_bba = dyn_ob->get_bba() + dyn_ob->get_d_bba() * dt;
+      double temp_bma = dyn_ob->get_bma() + dyn_ob->get_d_bma() * dt;
+      double temp_uma = dyn_ob->get_uma() + dyn_ob->get_d_uma() * dt;
+      double temp_uba = dyn_ob->get_uba() + dyn_ob->get_d_uba() * dt;
+
+      dyn_ob->set_bba(temp_bba);
+      dyn_ob->set_bma(temp_bma);
+      dyn_ob->set_uma(temp_uma);
+      dyn_ob->set_uba(temp_uba);
+    \end{lstlisting}
+  \end{tabular}
+  \caption{Code snippet of Euler's method to update the positions of the onebound model simultaneously.}
+  \label{ob_struct}
+\end{figure}
+
+The \verb|dt| used to update was $10^{-11}$ seconds.
+
+\subsubsection{Transitioning states}
+Transitions from one state to another are accomplished by calculating the rate of transition in a certain position, then using this rate divided by \verb|dt| as a probability of transfer at each timestep.
+The rates of transition are calculated by assuming they follow the Arrhenius rate equation $r = Ae^{-\Delta G/k_BT}$, where the preexponential factor is a parameter of the model and $\Delta G$ the energy of transition between the repsective states.
+A code snippet of this rate calculation is shown in Figure (\ref{transition-rate-snippet}).
+
+\begin{figure}[h]
+  \begin{tabular}{c}
+    \begin{lstlisting}[language=C++]
+      double Dynein_bothbound::get_near_unbinding_rate() {
+        double dG_spring = Dynein_onebound(this, rand, NEARBOUND).get_PE() - get_PE();
+        double low_affinity_unbinding_preexponential_factor = low_affinity_unbinding_rate / exp(1.0);
+        return low_affinity_unbinding_preexponential_factor*exp(-dG_spring/kb/T);
+      }
+    \end{lstlisting}
+  \end{tabular}
+  \caption{Code snippet of transition rate calculation using the Arrhenius equation.}
+  \label{transition-rate-snippet}
+\end{figure}
+
+\textit{todo: include lots more on the derivation of preexponential factors from experimental rates and loss of degrees of freedom}
+
+\subsubsection{Corrections}
+Certain configurations of the model, such as in the bothbound model when $\Theta_{nm} = \pi$ or $\Theta_{fm} = \pi$, it becomes impossible to calculate intermediate variables due to values going to infinity (name for this??).
+For example when $\Theta_{nm} = \pi$, $Ln = Ls + Lt$ and so:
+
+\begin{align}
+  \cos\theta_{ns} &= 1 \\
+  \sin\theta_{ns} &= 0 \\
+  \mbox{some val?} &\rightarrow \infty
+\end{align}
+
+Alternative problem states are when $L = 0$ and ???. These states are corrected similarly.\\
+
+\subsubsection{Forces}
+
+\subsection{Verifying the model}
+\subsubsection{Equipartition theorem obeyed}
+%% \begin{figure}[h!]
+%%   \centering
+%%   \begin{minipage}[b]{0.49\textwidth}
+%%     \includegraphics[width=\textwidth]{../figures/OB_Average_PE.pdf}
+%%     \caption{Onebound PE vs time.}
+%%   \end{minipage}
+%%   \begin{minipage}[b]{0.49\textwidth}
+%%     \includegraphics[width=\textwidth]{../figures/BB_Average_PE.pdf}
+%%     \caption{Bothbound PE vs time.}
+%%   \end{minipage}
+%%   \label{fig:equipartition_agreement}
+%% \end{figure}
+
+\subsection{Data collection}
+
+\subsubsection{Outputting positions, step locations}
+\subsubsection{Reconstructing steps from step locations}
+\subsubsection{Job management}
+
+\subsection{Parameter fitting}
+\subsubsection{Fitting spring constants}
+\subsubsection{Finding parameters from PyMol}
+\subsubsection{Finding values from literature}
+\subsubsection{Final parameter guess table}
+\subsubsection{Finding the right timestep}
+
+------------------------------------------
+
 \section{Chapter: Results}
@@
 
+\section{Appendix}
+
+\subsection{Motion equations}
+
+\subsubsection{Onebound equations}
+\label{onebound-motion-equations}
+The following are the full motion equations used to time evolve the onebound model:
+
+\begin{align*}
+AA &= L_s\sin\theta_{bb}      &   LL &= -L_s\cos(\theta_{bb})\\
+BB &= L_t\sin(\theta_{bm})    &	 MM &= -L_t\cos(\theta_{bm})\\
+CC &= -L_t\sin(\theta_{um})   &	 NN &= L_t\cos(\theta_{um})\\
+DD &= -L_s\sin(\theta_{ub})   &	 OO &= L_s\cos(\theta_{ub})\\
+\\
+EE &= -\gamma_m (X_{bm} - X_{bb})       &  PP &= -\gamma_m (Y_{bm} - Y_{bb})\\
+FF &= \gamma_m (X_{t } - X_{bm})        &	 QQ &= \gamma_m (Y_{t } - Y_{bm})\\ 
+GG &= -\gamma_t (X_{t } - X_{bm})       &	 RR &= -\gamma_t (Y_{t } - Y_{bm})\\
+HH &= \gamma_t (X_{um} - X_{t })        &	 SS &= \gamma_t (Y_{um} - Y_{t })\\ 
+II &= -\gamma_m (X_{um} - X_{t })       &	 TT &= -\gamma_m (Y_{um} - Y_{t })\\
+JJ &= \gamma_m (X_{ub} - X_{um})        &	 UU &= \gamma_m (Y_{ub} - Y_{um})\\ 
+KK &= -\gamma_b (X_{ub} - X_{um})       &	 VV &= -\gamma_b (Y_{ub} - Y_{um})\\
+\end{align*}%
+%
+\[
+\begin{pmatrix}
+  AA & 0 & 0 & 0 & EE & FF & 0 & 0\\
+  AA & BB & 0 & 0 & 0 & GG & HH & 0\\
+  AA & BB & CC & 0 & 0 & 0 & II & JJ\\
+  AA & BB & CC & DD & 0 & 0 & 0 & KK\\
+  LL & 0 & 0 & 0 & PP & QQ & 0 & 0\\
+  LL & MM & 0 & 0 & 0 & RR & SS & 0\\
+  LL & MM & NN & 0 & 0 & 0 & TT & UU\\
+  LL & MM & NN & OO & 0 & 0 & 0 & VV\\
+\end{pmatrix}
+\begin{pmatrix}
+    \dot{\theta}_{bb}\\
+    \dot{\theta}_{bm}\\
+    \dot{\theta}_{um}\\
+    \dot{\theta}_{ub}\\
+    \lambda_{bs}\\
+    \lambda_{bt}\\
+    \lambda_{ut}\\
+    \lambda_{us}\\
+  \end{pmatrix}
+=
+\begin{pmatrix}
+  X1\\
+  X2\\
+  X3\\
+  X4\\
+  X5\\
+  X6\\
+  X7\\
+  X8\\
+\end{pmatrix}
+\]
+
+Solving the above system for $\dot{\Theta}_{bb}, \dot{\Theta}_{bm}, \dot{\Theta}_{um}$ and $\dot{\Theta}_{ub}$ gives
+
+$\dot{\Theta}_{bb}$ =
+{\tiny
+( - (GG*II*MM*NN*PP*X1) + CC*HH*MM*PP*RR*X1 - BB*HH*NN*PP*RR*X1 + BB*II*NN*PP*RR*X1 - CC*GG*MM*PP*SS*X1 + BB*GG*NN*PP*SS*X1 + CC*GG*MM*PP*TT*X1 - BB*CC*PP*RR*TT*X1 + FF*II*MM*NN*PP*X2 - EE*II*MM*NN*QQ*X2 + CC*FF*MM*PP*SS*X2 - BB*FF*NN*PP*SS*X2 - CC*EE*MM*QQ*SS*X2 + BB*EE*NN*QQ*SS*X2 - CC*FF*MM*PP*TT*X2 + CC*EE*MM*QQ*TT*X2 - FF*HH*MM*NN*PP*X3 + EE*HH*MM*NN*QQ*X3 + BB*FF*NN*PP*SS*X3 - BB*EE*NN*QQ*SS*X3 + EE*GG*II*MM*NN*X5 - CC*EE*HH*MM*RR*X5 + BB*EE*HH*NN*RR*X5 - BB*EE*II*NN*RR*X5 + CC*EE*GG*MM*SS*X5 - BB*EE*GG*NN*SS*X5 - CC*EE*GG*MM*TT*X5 + BB*CC*EE*RR*TT*X5 - CC*FF*HH*MM*PP*X6 + BB*FF*HH*NN*PP*X6 - BB*FF*II*NN*PP*X6 + CC*EE*HH*MM*QQ*X6 - BB*EE*HH*NN*QQ*X6 + BB*EE*II*NN*QQ*X6 + BB*CC*FF*PP*TT*X6 - BB*CC*EE*QQ*TT*X6 + CC*FF*HH*MM*PP*X7 - CC*EE*HH*MM*QQ*X7 - BB*CC*FF*PP*SS*X7 + BB*CC*EE*QQ*SS*X7)/(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT) - ((FF*PP - EE*QQ)*(HH*MM - BB*SS)*(JJ*NN - CC*UU)*( - ((( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( - ((( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*NN*X2 +  Power(AA,4)* Power(BB,2)*NN*X3 -  Power(AA,4)* Power(BB,2)*CC*X7)))) + ( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( - ((( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,5)* Power(BB,3)*CC*DD*LL*X1 -  Power(AA,6)* Power(BB,2)*CC*DD*MM*X1 +  Power(AA,6)* Power(BB,2)*CC*DD*MM*X2 -  Power(AA,6)* Power(BB,3)*DD*NN*X2 +  Power(AA,6)* Power(BB,3)*DD*NN*X3 -  Power(AA,6)* Power(BB,3)*CC*OO*X3 +  Power(AA,6)* Power(BB,3)*CC*OO*X4 -  Power(AA,6)* Power(BB,3)*CC*DD*X8)))))/(( - (EE*GG*II*LL*MM*NN) - BB*FF*HH*LL*NN*PP + BB*FF*II*LL*NN*PP + AA*FF*HH*MM*NN*PP - AA*FF*II*MM*NN*PP + AA*GG*II*MM*NN*PP + BB*EE*HH*LL*NN*QQ - BB*EE*II*LL*NN*QQ - AA*EE*HH*MM*NN*QQ + AA*EE*II*MM*NN*QQ + CC*EE*HH*LL*MM*RR - BB*EE*HH*LL*NN*RR + BB*EE*II*LL*NN*RR - AA*CC*HH*MM*PP*RR + AA*BB*HH*NN*PP*RR - AA*BB*II*NN*PP*RR - CC*EE*GG*LL*MM*SS + BB*EE*GG*LL*NN*SS + BB*CC*FF*LL*PP*SS - AA*CC*FF*MM*PP*SS + AA*CC*GG*MM*PP*SS - AA*BB*GG*NN*PP*SS - BB*CC*EE*LL*QQ*SS + AA*CC*EE*MM*QQ*SS + CC*EE*GG*LL*MM*TT - BB*CC*FF*LL*PP*TT + AA*CC*FF*MM*PP*TT - AA*CC*GG*MM*PP*TT + BB*CC*EE*LL*QQ*TT - AA*CC*EE*MM*QQ*TT - BB*CC*EE*LL*RR*TT + AA*BB*CC*PP*RR*TT)*( - (( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*(( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( Power(AA,4)* Power(BB,2)*JJ*NN -  Power(AA,4)* Power(BB,2)*CC*UU)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( Power(AA,6)* Power(BB,3)*DD*JJ*NN -  Power(AA,6)* Power(BB,3)*CC*JJ*OO +  Power(AA,6)* Power(BB,3)*CC*KK*OO -  Power(AA,6)* Power(BB,3)*CC*DD*VV)))
+} \break
+
+$\dot{\Theta}_{bm}$ =
+{\tiny
+  (GG*II*LL*NN*PP*X1 - CC*HH*LL*PP*RR*X1 + AA*HH*NN*PP*RR*X1 - AA*II*NN*PP*RR*X1 + CC*GG*LL*PP*SS*X1 - AA*GG*NN*PP*SS*X1 - CC*GG*LL*PP*TT*X1 + AA*CC*PP*RR*TT*X1 - FF*II*LL*NN*PP*X2 + EE*II*LL*NN*QQ*X2 - EE*II*LL*NN*RR*X2 + AA*II*NN*PP*RR*X2 - CC*FF*LL*PP*SS*X2 + AA*FF*NN*PP*SS*X2 + CC*EE*LL*QQ*SS*X2 - AA*EE*NN*QQ*SS*X2 + CC*FF*LL*PP*TT*X2 - CC*EE*LL*QQ*TT*X2 + CC*EE*LL*RR*TT*X2 - AA*CC*PP*RR*TT*X2 + FF*HH*LL*NN*PP*X3 - EE*HH*LL*NN*QQ*X3 + EE*HH*LL*NN*RR*X3 - AA*HH*NN*PP*RR*X3 - EE*GG*LL*NN*SS*X3 - AA*FF*NN*PP*SS*X3 + AA*GG*NN*PP*SS*X3 + AA*EE*NN*QQ*SS*X3 - EE*GG*II*LL*NN*X5 + CC*EE*HH*LL*RR*X5 - AA*EE*HH*NN*RR*X5 + AA*EE*II*NN*RR*X5 - CC*EE*GG*LL*SS*X5 + AA*EE*GG*NN*SS*X5 + CC*EE*GG*LL*TT*X5 - AA*CC*EE*RR*TT*X5 + EE*GG*II*LL*NN*X6 + CC*FF*HH*LL*PP*X6 - AA*FF*HH*NN*PP*X6 + AA*FF*II*NN*PP*X6 - AA*GG*II*NN*PP*X6 - CC*EE*HH*LL*QQ*X6 + AA*EE*HH*NN*QQ*X6 - AA*EE*II*NN*QQ*X6 - CC*EE*GG*LL*TT*X6 - AA*CC*FF*PP*TT*X6 + AA*CC*GG*PP*TT*X6 + AA*CC*EE*QQ*TT*X6 - CC*FF*HH*LL*PP*X7 + CC*EE*HH*LL*QQ*X7 - CC*EE*HH*LL*RR*X7 + AA*CC*HH*PP*RR*X7 + CC*EE*GG*LL*SS*X7 + AA*CC*FF*PP*SS*X7 - AA*CC*GG*PP*SS*X7 - AA*CC*EE*QQ*SS*X7)/(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT) + (( - (FF*HH*LL*PP) + EE*HH*LL*QQ - EE*HH*LL*RR + AA*HH*PP*RR + EE*GG*LL*SS + AA*FF*PP*SS - AA*GG*PP*SS - AA*EE*QQ*SS)*(JJ*NN - CC*UU)*( - ((( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( - ((( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*NN*X2 +  Power(AA,4)* Power(BB,2)*NN*X3 -  Power(AA,4)* Power(BB,2)*CC*X7)))) + ( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( - ((( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,5)* Power(BB,3)*CC*DD*LL*X1 -  Power(AA,6)* Power(BB,2)*CC*DD*MM*X1 +  Power(AA,6)* Power(BB,2)*CC*DD*MM*X2 -  Power(AA,6)* Power(BB,3)*DD*NN*X2 +  Power(AA,6)* Power(BB,3)*DD*NN*X3 -  Power(AA,6)* Power(BB,3)*CC*OO*X3 +  Power(AA,6)* Power(BB,3)*CC*OO*X4 -  Power(AA,6)* Power(BB,3)*CC*DD*X8)))))/((EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT)*( - (( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*(( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( Power(AA,4)* Power(BB,2)*JJ*NN -  Power(AA,4)* Power(BB,2)*CC*UU)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( Power(AA,6)* Power(BB,3)*DD*JJ*NN -  Power(AA,6)* Power(BB,3)*CC*JJ*OO +  Power(AA,6)* Power(BB,3)*CC*KK*OO -  Power(AA,6)* Power(BB,3)*CC*DD*VV)))
+}\break
+
+$\dot{\Theta}_{um}$ =
+{\tiny
+  ( - X2 + X3)/CC + (GG*( - (BB*LL*PP*X1) + AA*MM*PP*X1 + EE*LL*MM*X2 - AA*MM*PP*X2 + BB*EE*LL*X5 - AA*EE*MM*X5 - BB*EE*LL*X6 + AA*BB*PP*X6))/(CC*(EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)) - ((( - HH + II)/CC - (GG*(EE*LL - AA*PP)*( - (HH*MM) + BB*SS))/(CC*(EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)))*( - (BB*GG*LL*NN*PP*X1) + AA*GG*MM*NN*PP*X1 + BB*CC*LL*PP*RR*X1 - AA*CC*MM*PP*RR*X1 + BB*FF*LL*NN*PP*X2 - AA*FF*MM*NN*PP*X2 - BB*EE*LL*NN*QQ*X2 + AA*EE*MM*NN*QQ*X2 - CC*EE*LL*MM*RR*X2 + BB*EE*LL*NN*RR*X2 + AA*CC*MM*PP*RR*X2 - AA*BB*NN*PP*RR*X2 + EE*GG*LL*MM*NN*X3 - BB*FF*LL*NN*PP*X3 + AA*FF*MM*NN*PP*X3 - AA*GG*MM*NN*PP*X3 + BB*EE*LL*NN*QQ*X3 - AA*EE*MM*NN*QQ*X3 - BB*EE*LL*NN*RR*X3 + AA*BB*NN*PP*RR*X3 + BB*EE*GG*LL*NN*X5 - AA*EE*GG*MM*NN*X5 - BB*CC*EE*LL*RR*X5 + AA*CC*EE*MM*RR*X5 + CC*EE*GG*LL*MM*X6 - BB*EE*GG*LL*NN*X6 - BB*CC*FF*LL*PP*X6 + AA*CC*FF*MM*PP*X6 - AA*CC*GG*MM*PP*X6 + AA*BB*GG*NN*PP*X6 + BB*CC*EE*LL*QQ*X6 - AA*CC*EE*MM*QQ*X6 - CC*EE*GG*LL*MM*X7 + BB*CC*FF*LL*PP*X7 - AA*CC*FF*MM*PP*X7 + AA*CC*GG*MM*PP*X7 - BB*CC*EE*LL*QQ*X7 + AA*CC*EE*MM*QQ*X7 + BB*CC*EE*LL*RR*X7 - AA*BB*CC*PP*RR*X7))/(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT) - ((JJ/CC + ((EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)*(( - HH + II)/CC - (GG*(EE*LL - AA*PP)*( - (HH*MM) + BB*SS))/(CC*(EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)))*( - (JJ*NN) + CC*UU))/(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT))*( - ((( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( - ((( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*NN*X2 +  Power(AA,4)* Power(BB,2)*NN*X3 -  Power(AA,4)* Power(BB,2)*CC*X7)))) + ( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( - ((( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,5)* Power(BB,3)*CC*DD*LL*X1 -  Power(AA,6)* Power(BB,2)*CC*DD*MM*X1 +  Power(AA,6)* Power(BB,2)*CC*DD*MM*X2 -  Power(AA,6)* Power(BB,3)*DD*NN*X2 +  Power(AA,6)* Power(BB,3)*DD*NN*X3 -  Power(AA,6)* Power(BB,3)*CC*OO*X3 +  Power(AA,6)* Power(BB,3)*CC*OO*X4 -  Power(AA,6)* Power(BB,3)*CC*DD*X8)))))/( - (( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*(( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( Power(AA,4)* Power(BB,2)*JJ*NN -  Power(AA,4)* Power(BB,2)*CC*UU)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( Power(AA,6)* Power(BB,3)*DD*JJ*NN -  Power(AA,6)* Power(BB,3)*CC*JJ*OO +  Power(AA,6)* Power(BB,3)*CC*KK*OO -  Power(AA,6)* Power(BB,3)*CC*DD*VV))
+}\break
+
+$\dot{\Theta}_{ub}$ =
+{\tiny
+  ( - X3 + X4)/DD + (II*( - (BB*GG*LL*NN*PP*X1) + AA*GG*MM*NN*PP*X1 + BB*CC*LL*PP*RR*X1 - AA*CC*MM*PP*RR*X1 + BB*FF*LL*NN*PP*X2 - AA*FF*MM*NN*PP*X2 - BB*EE*LL*NN*QQ*X2 + AA*EE*MM*NN*QQ*X2 - CC*EE*LL*MM*RR*X2 + BB*EE*LL*NN*RR*X2 + AA*CC*MM*PP*RR*X2 - AA*BB*NN*PP*RR*X2 + EE*GG*LL*MM*NN*X3 - BB*FF*LL*NN*PP*X3 + AA*FF*MM*NN*PP*X3 - AA*GG*MM*NN*PP*X3 + BB*EE*LL*NN*QQ*X3 - AA*EE*MM*NN*QQ*X3 - BB*EE*LL*NN*RR*X3 + AA*BB*NN*PP*RR*X3 + BB*EE*GG*LL*NN*X5 - AA*EE*GG*MM*NN*X5 - BB*CC*EE*LL*RR*X5 + AA*CC*EE*MM*RR*X5 + CC*EE*GG*LL*MM*X6 - BB*EE*GG*LL*NN*X6 - BB*CC*FF*LL*PP*X6 + AA*CC*FF*MM*PP*X6 - AA*CC*GG*MM*PP*X6 + AA*BB*GG*NN*PP*X6 + BB*CC*EE*LL*QQ*X6 - AA*CC*EE*MM*QQ*X6 - CC*EE*GG*LL*MM*X7 + BB*CC*FF*LL*PP*X7 - AA*CC*FF*MM*PP*X7 + AA*CC*GG*MM*PP*X7 - BB*CC*EE*LL*QQ*X7 + AA*CC*EE*MM*QQ*X7 + BB*CC*EE*LL*RR*X7 - AA*BB*CC*PP*RR*X7))/(DD*(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT)) - ((( - JJ + KK)/DD - (II*(EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)*( - (JJ*NN) + CC*UU))/(DD*(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT)))*( - ((( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( - ((( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*NN*X2 +  Power(AA,4)* Power(BB,2)*NN*X3 -  Power(AA,4)* Power(BB,2)*CC*X7)))) + ( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( - ((( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,5)* Power(BB,3)*CC*DD*LL*X1 -  Power(AA,6)* Power(BB,2)*CC*DD*MM*X1 +  Power(AA,6)* Power(BB,2)*CC*DD*MM*X2 -  Power(AA,6)* Power(BB,3)*DD*NN*X2 +  Power(AA,6)* Power(BB,3)*DD*NN*X3 -  Power(AA,6)* Power(BB,3)*CC*OO*X3 +  Power(AA,6)* Power(BB,3)*CC*OO*X4 -  Power(AA,6)* Power(BB,3)*CC*DD*X8)))))/( - (( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*(( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( Power(AA,4)* Power(BB,2)*JJ*NN -  Power(AA,4)* Power(BB,2)*CC*UU)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( Power(AA,6)* Power(BB,3)*DD*JJ*NN -  Power(AA,6)* Power(BB,3)*CC*JJ*OO +  Power(AA,6)* Power(BB,3)*CC*KK*OO -  Power(AA,6)* Power(BB,3)*CC*DD*VV))
+}
+
+\subsubsection{Bothbound equations}
+\label{bothbound-motion-equations}
+The following are the full motion equations used to time evolve the bothbound model:
+
+\begin{center}
+  \begin{minipage}{0.3\textwidth}
+    \begin{align*}
+      a &= \frac{-dXnm}{dLn}\\
+       b &= \frac{-dXnm}{dLf}\\
+       c &= -(x_{nm} - x_{nb}) / \gamma_m\\
+       d &= (x_t - x_{nm}) / \gamma_m\\
+       e &= \frac{-dXt}{dLn}\\
+       f &= \frac{-dXt}{dLf}\\
+       g &= -(x_t - x_{nm}) / \gamma_t\\
+       h &= (x_{fm} - x_t) / \gamma_t\\
+       i &= \frac{-dXfm}{dLn}\\
+       j &= \frac{-dXfm}{dLf}\\
+       k &= -(x_{fm} - x_t) / \gamma_m\\
+       l &= (x_{fb} - x_{fm}) / \gamma_m\\
+    \end{align*}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+    \begin{align*}
+       m &= \frac{-dYnm}{dLn}\\
+       n &= \frac{-dYnm}{dLf}\\
+       p &= -(y_{nm} - y_{nb}) / \gamma_m\\
+       q &= (y_t - y_{nm}) / \gamma_m\\
+       r &= \frac{-dYt}{dLn}\\
+       s &= \frac{-dYt}{dLf}\\
+       t &= -(y_t - y_{nm}) / \gamma_t\\
+       u &= (y_{fm} - y_t) / \gamma_t\\
+       v &= \frac{-dYfm}{dLn}\\
+       w &= \frac{-dYfm}{dLf}\\
+       x &= -(y_{fm} - y_t) / \gamma_m\\
+       y &= (y_{fb} - y_{fm}) / \gamma_m
+    \end{align*}
+  \end{minipage}
+\end{center}
+
+$\dot{Ln}$ =
+{\tiny
+    (h*l*p*t*w*x1 - g*l*p*u*w*x1 + g*l*p*s*x*x1 - f*l*p*t*x*x1 +
+     d*l*p*u*w*x2 - c*l*q*u*w*x2 - d*l*p*s*x*x2 + c*l*q*s*x*x2 -
+     c*l*n*t*x*x2 + b*l*p*t*x*x2 - c*h*l*t*w*x4 + c*g*l*u*w*x4 -
+     c*g*l*s*x*x4 + c*f*l*t*x*x4 - d*h*l*p*w*x5 + c*h*l*q*w*x5 +
+     c*g*l*n*x*x5 + d*f*l*p*x*x5 - b*g*l*p*x*x5 - c*f*l*q*x*x5 +
+     d*h*l*p*s*x6 - c*h*l*q*s*x6 + c*h*l*n*t*x6 - b*h*l*p*t*x6 -
+     c*g*l*n*u*x6 - d*f*l*p*u*x6 + b*g*l*p*u*x6 + c*f*l*q*u*x6 -
+     g*k*p*s*x1*y - h*j*p*t*x1*y + f*k*p*t*x1*y + g*j*p*u*x1*y +
+     d*k*p*s*x2*y - c*k*q*s*x2*y + c*k*n*t*x2*y - b*k*p*t*x2*y -
+     d*j*p*u*x2*y + c*j*q*u*x2*y - d*h*p*s*x3*y + c*h*q*s*x3*y -
+     c*h*n*t*x3*y + b*h*p*t*x3*y + c*g*n*u*x3*y + d*f*p*u*x3*y -
+     b*g*p*u*x3*y - c*f*q*u*x3*y + c*g*k*s*x4*y + c*h*j*t*x4*y -
+     c*f*k*t*x4*y - c*g*j*u*x4*y - c*g*k*n*x5*y + d*h*j*p*x5*y -
+     d*f*k*p*x5*y + b*g*k*p*x5*y - c*h*j*q*x5*y + c*f*k*q*x5*y)/
+     (d*h*l*p*s*v - c*h*l*q*s*v + c*h*l*n*t*v - b*h*l*p*t*v -
+     c*g*l*n*u*v - d*f*l*p*u*v + b*g*l*p*u*v + c*f*l*q*u*v -
+     d*h*l*p*r*w + c*h*l*q*r*w - c*h*l*m*t*w + a*h*l*p*t*w +
+     c*g*l*m*u*w + d*e*l*p*u*w - a*g*l*p*u*w - c*e*l*q*u*w +
+     c*g*l*n*r*x + d*f*l*p*r*x - b*g*l*p*r*x - c*f*l*q*r*x -
+     c*g*l*m*s*x - d*e*l*p*s*x + a*g*l*p*s*x + c*e*l*q*s*x +
+     c*f*l*m*t*x - c*e*l*n*t*x + b*e*l*p*t*x - a*f*l*p*t*x -
+     c*g*k*n*r*y + d*h*j*p*r*y - d*f*k*p*r*y + b*g*k*p*r*y -
+     c*h*j*q*r*y + c*f*k*q*r*y + c*g*k*m*s*y - d*h*i*p*s*y +
+     d*e*k*p*s*y - a*g*k*p*s*y + c*h*i*q*s*y - c*e*k*q*s*y +
+     c*h*j*m*t*y - c*f*k*m*t*y - c*h*i*n*t*y + c*e*k*n*t*y +
+     b*h*i*p*t*y - a*h*j*p*t*y - b*e*k*p*t*y + a*f*k*p*t*y -
+     c*g*j*m*u*y + c*g*i*n*u*y + d*f*i*p*u*y - b*g*i*p*u*y -
+     d*e*j*p*u*y + a*g*j*p*u*y - c*f*i*q*u*y + c*e*j*q*u*y)
+}\break
+
+$\dot{Lf}$ =
+{\tiny
+    (-(h*l*p*t*v*x1) + g*l*p*u*v*x1 - g*l*p*r*x*x1 + e*l*p*t*x*x1 -
+     d*l*p*u*v*x2 + c*l*q*u*v*x2 + d*l*p*r*x*x2 - c*l*q*r*x*x2 +
+     c*l*m*t*x*x2 - a*l*p*t*x*x2 + c*h*l*t*v*x4 - c*g*l*u*v*x4 +
+     c*g*l*r*x*x4 - c*e*l*t*x*x4 + d*h*l*p*v*x5 - c*h*l*q*v*x5 -
+     c*g*l*m*x*x5 - d*e*l*p*x*x5 + a*g*l*p*x*x5 + c*e*l*q*x*x5 -
+     d*h*l*p*r*x6 + c*h*l*q*r*x6 - c*h*l*m*t*x6 + a*h*l*p*t*x6 +
+     c*g*l*m*u*x6 + d*e*l*p*u*x6 - a*g*l*p*u*x6 - c*e*l*q*u*x6 +
+     g*k*p*r*x1*y + h*i*p*t*x1*y - e*k*p*t*x1*y - g*i*p*u*x1*y -
+     d*k*p*r*x2*y + c*k*q*r*x2*y - c*k*m*t*x2*y + a*k*p*t*x2*y +
+     d*i*p*u*x2*y - c*i*q*u*x2*y + d*h*p*r*x3*y - c*h*q*r*x3*y +
+     c*h*m*t*x3*y - a*h*p*t*x3*y - c*g*m*u*x3*y - d*e*p*u*x3*y +
+     a*g*p*u*x3*y + c*e*q*u*x3*y - c*g*k*r*x4*y - c*h*i*t*x4*y +
+     c*e*k*t*x4*y + c*g*i*u*x4*y + c*g*k*m*x5*y - d*h*i*p*x5*y +
+     d*e*k*p*x5*y - a*g*k*p*x5*y + c*h*i*q*x5*y - c*e*k*q*x5*y)/
+     (d*h*l*p*s*v - c*h*l*q*s*v + c*h*l*n*t*v - b*h*l*p*t*v -
+     c*g*l*n*u*v - d*f*l*p*u*v + b*g*l*p*u*v + c*f*l*q*u*v -
+     d*h*l*p*r*w + c*h*l*q*r*w - c*h*l*m*t*w + a*h*l*p*t*w +
+     c*g*l*m*u*w + d*e*l*p*u*w - a*g*l*p*u*w - c*e*l*q*u*w +
+     c*g*l*n*r*x + d*f*l*p*r*x - b*g*l*p*r*x - c*f*l*q*r*x -
+     c*g*l*m*s*x - d*e*l*p*s*x + a*g*l*p*s*x + c*e*l*q*s*x +
+     c*f*l*m*t*x - c*e*l*n*t*x + b*e*l*p*t*x - a*f*l*p*t*x -
+     c*g*k*n*r*y + d*h*j*p*r*y - d*f*k*p*r*y + b*g*k*p*r*y -
+     c*h*j*q*r*y + c*f*k*q*r*y + c*g*k*m*s*y - d*h*i*p*s*y +
+     d*e*k*p*s*y - a*g*k*p*s*y + c*h*i*q*s*y - c*e*k*q*s*y +
+     c*h*j*m*t*y - c*f*k*m*t*y - c*h*i*n*t*y + c*e*k*n*t*y +
+     b*h*i*p*t*y - a*h*j*p*t*y - b*e*k*p*t*y + a*f*k*p*t*y -
+     c*g*j*m*u*y + c*g*i*n*u*y + d*f*i*p*u*y - b*g*i*p*u*y -
+     d*e*j*p*u*y + a*g*j*p*u*y - c*f*i*q*u*y + c*e*j*q*u*y)
+}
+
 \section{Questions for Tate/Nicole}

### 2017-01-10 thesis/latex/thesis.tex
Add internal and Brownian force calculations.

@@
 \subsubsection{Forces}
+Spring and Brownian forces are used to properly move the model through time.
+Appendices (\ref{onebound-motion-equatons}-\ref{bothbound-motion-equations}) show angular velocities are functions of the x- and y-components of both spring forces $F$ and Brownian forces $\gamma R$.
+We calculate these as shown:
+
+\paragraph{Brownian forces}
+Each Brownian force $R_{n,m}$ on domain $n$ in direction $m$ is sampled from a Gaussian with standard deviation $\sqrt{\frac{2k_BT\gamma_n}{dt}}$.\\
+
+\paragraph{Spring forces}
+Spring forces are calculated by first finding the torque $\tau$ on each angle, then converting this to a force.
+For example, the onebound near motor angle feels a force $\tau_{nm} = c_m(\Theta_{nm}-\Theta_{nm,eq})$, where $c_n$ is the spring constant of the motor domain and $\Theta_{nm,eq}$ the equilibrium angle.
+The magnitude of force on the adjacent binding and tail domains is thus $\tau/Ls$ and $\tau/Lt$, pointing in the direction adjacent the Ls and Lt rods, respectively.
+The force on the motor domain is found as the equal-and-opposite force to these two imposed forces.
+In code:
+
+\begin{figure}[h]
+  \centering
+  \begin{tabular}{c}
+    \begin{lstlisting}[language=C++]
+      T = cm*(nma - nma_eq);
+      f1 = T/Ls;
+      f2 = T/Lt;
+      f1x = f1 * sin(bba);
+      f1y = f1 * -cos(bba);
+      f2x = f2 * sin(bma);
+      f2y = f2 * -cos(bma);
+      f.bbx += f1x;
+      f.bby += f1y;
+      f.tx  += f2x;
+      f.ty  += f2y;
+      f.bmx += -(f1x + f2x); // equal and opposite forces
+      f.bmy += -(f1y + f2y);
+    \end{lstlisting}
+  \end{tabular}
+  \caption{Code snippet of internal force calculation.}
+  \label{transition-rate-snippet}
+\end{figure}
 

### 2017-01-11 thesis/latex/thesis.tex
Add work on tests.

@@
 \subsubsection{Corrections}
+\label{corrections-section}
 Certain configurations of the model, such as in the bothbound model when $\Theta_{nm} = \pi$ or $\Theta_{fm} = \pi$, it becomes impossible to calculate intermediate variables due to values going to infinity (name for this??).
@@
 \subsection{Verifying the model}
+The models were verified to follow physical laws by checking their behavior in various tests: general conformational tests, energy conservation and obeying the equipartition theorem.
+The purpose of these tests was to verify that equations had been entered properly (primarily the conformational and energy conservation tests), and that the model obeyed the laws of physics (all tests).
+
+\subsubsection{Conformational tests}
+The models were tested in the following different conformations to verify they behaved as expected:\\
+
+\textbf{Onebound tests}
+\begin{itemize}
+\item \textbf{Onebound:} \textit{Upwards line conformation with no forces.} Check that all domain x-values are zero$^*$ and all y-values are the proper addition of $L_s$ and $L_t$ lengths.
+Also checks that all angular velocities are zero.
+This test verifies that the angle scheme is defined properly, that the model can handle straight lines without generating NaNs and that there are no velocities (angular or cartesian) without forces.
+\item \textbf{Onebound:} \textit{Horizontal line conformation with no forces.} Check that all domain y-values are zero and all x-values are the proper addition of $L_s$ and $L_t$ lengths.
+Also check that all velocities are zero.
+This test double-verifies that the coordinate system behaves as expected and that velocities are zero in the absence of forces.
+\item \textbf{Onebound:} \textit{Upwards line conformation with +x forces} Check that cartesian x-velocities are positive and y-velocities are zero.
+Verifies that the cartesian velocities have the expected sign.
+\item \textbf{Onebound:} \textit{Upwards line conformation with +y forces} Check that all cartesian velocities are zero.
+Verifies that upwards forces in an already-vertical stalk have no effect on velocity.
+  \item \textbf{Onebound:} \textit{Prepowerstroke opposing internal/brownian forces} Check that, for two onebound models, one with +x spring forces and no Brownian forces and the other with -x Brownian forces and no spring forces, that the velocities of each unbound domain are equal and opposite.
+Verifies that equal Brownian and internal forces have the same magnitude and direction of effect on velocities.
+\end{itemize}
+*: checks for equality are within a margin of error of $\epsilon = 1e^{-5}$.\\
+
+These tests together verify that the onebound model's internal coordinate system's directions match the XY coordinate system directions, that the angular directions in the model properly align with the $\pm x$ and $\pm y$ directions of the XY coordinate system, that velocities have the proper magnitude at zero force and proper sign at nonzero force, and that Brownian and internal forces are in the same unit system.
+There are still errors which could exist in the model pertaining to magnitude of velocity, which will be tested for in the energy conservation section.\\
+
+\textbf{Bothbound tests}
+\begin{itemize}
+\item \textbf{Bothbound:} \textit{Upwards line conformation with no Brownian forces.} Both MTB domains are very close together ($L=10^(-25)$) and motor domains nearly vertical ($\Theta_{n/f} = \pi \pm 10^(-15)$).
+Equilibrium angles set to have a vertical line as the lowest energy point.
+Model is not exactly vertical for NaN concerns explained in section \ref{corrections-section}.
+Checks that domain x-values are zero$^*$, domain y-values are the correct combination of $L_s$ and $L_t$ lengths, all angles are at equilibrium and all forces are zero.
+Verifies that the domains are properly oriented by the angles, the angles are defined properly and forces are calculated from angles properly.
+\item \textbf{Bothbound:} \textit{Equilateral house conformation with outward x-forces.} The model forms the shape of a square with an equilateral triangle on top, with the tail domain as the triangle.
+Checks if all domains are at the desired locations and the derivatives are the proper sign and are equal and opposite to their counterparts on the opposing side of the house.
+This test verifies that the angular system behaves as expected and forces are symmetric about the two dimers.
+\item \textbf{Bothbound:} \textit{Two table conformations with near/far domains flipped.} Two models are put in the same table-like conformation, but the near/far domains are flipped.
+Checks if opposite domains have the same positions and forces.
+Re-verifies that the model is symmetric in position and force between the two dimers.
+\end{itemize}
+
+These tests verify the bothbound model's angle system behaves as expected, forces are zero when expected and have the proper signs, and that the whole system is symmetric with either dimer.
+
+\subsubsection{Energy conservation}
+To verify that the magnitudes of forces are calculated properly, a conservation of energy test was performed.
+A very slight angular nudge was applied to either model.
+This nudge caused a small change in energy.
+This change in energy was compared with the work the forces on each domain caused over the nudge.
+In order for energy to be conserved, these values were required to be very similar.
+Mathematically the test required the following:
+
+\begin{equation}
+  \sum_n \frac12 \left(\Theta_{nudge} - \Theta_{n} - \Theta_{n,eq}\right) = \sum_n \vec{F}_n \cdot \vec{r}_{n,nudge}
+\end{equation}
+
+where $\Theta_{nudge}$ was the angular nudge applied, $F_n$ the net force on domain $n$ and $\vec{r}_{n,nudge}$ the cartesian displacement of each domain $n$ due to the nudge.
+This test is powerful since it links the force-calculation system of either model to the energy calculation system.
+Energy is calculated by looking at displacements of angles from equilibrium, whereas forces are calculated using displacement of angles from equilibrium, but also orientation and position of each domain.
+Verifying these two systems produce the same result indicates...
+something.
+\textbf{make this more precise.}
+
+\textbf{is there a test to be run which could verify the velocity calculation is correct?}
+
+\subsubsection{Equipartition theorem}
+
 \subsubsection{Equipartition theorem obeyed}

### 2017-01-11 thesis/latex/thesis.tex
Add info on tests.

--- 2017-01-10 thesis/latex/thesis.tex
@@
 \subsection{Verifying the model}
+The models were verified to follow physical laws by checking their behavior in various tests: general conformational tests, energy conservation and obeying the equipartition theorem.
+The purpose of these tests was to verify that equations had been entered properly (primarily the conformational and energy conservation tests), and that the model obeyed the laws of physics (all tests).
+
+\subsubsection{Conformational tests}
+The models were tested in the following different conformations to verify they behaved as expected:\\
+
+\textbf{Onebound tests}
+\begin{itemize}
+\item \textbf{Onebound:} \textit{Upwards line conformation with no forces.} Check that all domain x-values are zero$^*$ and all y-values are the proper addition of $L_s$ and $L_t$ lengths.
+Also checks that all angular velocities are zero.
+This test verifies that the angle scheme is defined properly, that the model can handle straight lines without generating NaNs and that there are no velocities (angular or cartesian) without forces.
+\item \textbf{Onebound:} \textit{Horizontal line conformation with no forces.} Check that all domain y-values are zero and all x-values are the proper addition of $L_s$ and $L_t$ lengths.
+Also check that all velocities are zero.
+This test double-verifies that the coordinate system behaves as expected and that velocities are zero in the absence of forces.
+\item \textbf{Onebound:} \textit{Upwards line conformation with +x forces} Check that cartesian x-velocities are positive and y-velocities are zero.
+Verifies that the cartesian velocities have the expected sign.
+\item \textbf{Onebound:} \textit{Upwards line conformation with +y forces} Check that all cartesian velocities are zero.
+Verifies that upwards forces in an already-vertical stalk have no effect on velocity.
+  \item \textbf{Onebound:} \textit{Prepowerstroke opposing internal/brownian forces} Check that, for two onebound models, one with +x spring forces and no Brownian forces and the other with -x Brownian forces and no spring forces, that the velocities of each unbound domain are equal and opposite.
+Verifies that equal Brownian and internal forces have the same magnitude and direction of effect on velocities.
+\end{itemize}
+*: checks for equality are within a margin of error of $\epsilon = 1e^{-5}$.\\
+
+These tests together verify that the onebound model's internal coordinate system's directions match the XY coordinate system directions, that the angular directions in the model properly align with the $\pm x$ and $\pm y$ directions of the XY coordinate system, that velocities have the proper magnitude at zero force and proper sign at nonzero force, and that Brownian and internal forces are in the same unit system.
+There are still errors which could exist in the model pertaining to magnitude of velocity, which will be tested for in the energy conservation section.\\
+
+\textbf{Bothbound tests} go here.
+
+\subsubsection{Energy conservation}
+
+\subsubsection{Equipartition theorem}
+
 \subsubsection{Equipartition theorem obeyed}

### 2017-01-16 thesis/latex/tate-figures-1-17.tex
Add listoffigures and Tate assignment.

--- 2017-01-10 thesis/latex/thesis.tex
@@
 
+\usepackage[utf8]{inputenc}
+\usepackage{graphicx}
+\usepackage{array}
+
 \setlength{\emergencystretch}{3cm} % somehow helps prevent hboxes from overflowing
@@
 
-\tableofcontents
-\newpage
-
-\section{Chapter: Introduction}
-\subsection{What is dynein?}
-\subsubsection{Basics}
-
-Dynein is a cellular structure which converts chemical energy to mechanical energy.
-It does so by reacting with adenosine triphosphate (ATP) to take steps along long cellular tracks known as microtubules (MTs).
-Dynein's mechanical energy is used to transport molecules around the cell, pull chromosomes during division, move cellular propellers known as cilia and various other functions \cite{cianfroccoreview}.\\
-
-Dynein is a protein with many smaller domains.
-A depiction of the protein and its labeled subdomains is shown in Figure (\ref{dynein-artist-rendition}).
-Each motor is a homodimer made up of two identical monomer proteins.
-Each monomer has a binding domain which binds to the microtubule, a motor domain which binds ATP, a tail domain which connects to the other monomer, a stalk which connects the binding and motor domains, and a linker which connects the motor and tail domains.
-The tail domain is also responsible for binding to cargo.\\
-
-\begin{figure}[h]
+\begin{figure}
   \centering
-  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/dynein-artist-rendition.jpg}
+  \includegraphics[width=.45\textwidth,keepaspectratio]{../figures/dynein-artist-rendition.jpg}
   \caption{Artist rendition of the dynein motor bound to a microtubule. One monomer is bound to the MT, and one is raised up. Modified. Source: Lander Lab, The Scripps Research Institute.}
@@
 
-Dynein is believed to move by using the energy of ATP to cycle through a ring of states, each with different spatial relations to the microtubule.
-By cycling forward through this ring, the molecule can end up in the same state it began, but moved forward a small amount.
-This ring of states is known as the mechanochemical cycle \cite{cianfroccoreview}.
-The goal of this project is to create a mathematical model of the mechanochemical cycle and verify that it can reproduce experimental dynein stepping data.
-There are several types of dynein which all behave differently.
-This project will focus on cytoplasmic dynein-1, hereby referred to as dynein.\\
-
-\subsubsection{Glossary}
-\textbf{ATP} Adenosine triphosphate is a small organic molecule which very spontaneously breaks apart into ADP and a phosphate.
-This reaction is used to power many energy-requiring mechanisms within the cell.
-This reaction has a $\Delta G^\circ \approx -36 kJ/mol$ at physiological conditions \ref{ATPenergy}.\\\\
-\textbf{Protein} Dynein is a protein, or long chain of amino acids linked together and folded into a particular shape.\\
-
-\subsubsection{Why do cells need motor transport?}
-Cells are organized, heretogenous structures which respond quickly to their environments.
-This means that cells require a mechanism for rapidly moving components to precise locations within the cell.
-This can be a challenge, since cells are fairly large compared to the proteins which compose them.
-A human fibroblast cell has a volume of roughly 2000 $\mu m^3$\cite{fibroblastvolume}, corresponding to roughly $8 \mu m$ in diameter.
-In comparison, human hemoglobin has a diameter of roughly $5 nm$ (PDB id 5ME2) \textbf{((is this the best way to cite this?))}, a $10^3$ factor difference in size.\\
-
-Diffusion, or random motion of molecules due to collisions with solvent (i.e.
-Brownian motion), is one possible process cells could use to transport biomolecules.
-Diffusion has two problems: it is slow and nondirected.
-The expected distance covered for one-dimensional Brownian motion after time t was found by Einstein to be \textbf{cite here}:
-
-\begin{equation}
-  \label{diffusion-equation}
-  <x> = \sqrt{2Dt}
-\end{equation}
-
-where D is the diffusion constant describing the diffusability of the biomolecule.
-For a medium-sized (140kD, 1D = weight of H atom) protein with a diffusion constant $D = .2 \mu m^2/s$ %% \cite{diffusionconstants}
-, it would take about a month to travel across a millimeter-sized cell.
-In contrast, it would take dynein, which travels at roughly 100nm/s \cite{weihongpaper}, a much more reasonable three hours to do the same.
-
-\begin{figure}[h]
+\begin{figure}
   \centering
-  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/diffusion_vs_dynein.png}
+  \includegraphics[width=.45\textwidth,keepaspectratio]{../figures/diffusion_vs_dynein.png}
   \caption{Plot comparing the diffusion rate of a 140kD protein with a diffusion consant $D = .2 \mu m^2/s$ with a dynein motor travelling at 100nm/s.}
@@
 
-\textit{Maybe use the more reasonable 2$\mu m / s$ number for dynein velocity, in-vitro speed may not be applicable since [ATP]-limited}
-
-Another advantage of motor transport is that microtubules can be highly specific in their location in the cell.
-During cell division chromosomes line up at the center of each cell in a plane.
-Such a geometrically precise shape requires some sort of directed motion which a random process could never provide.\\
-
-\textit{maybe decorate the above plot with pictures of cells of each size, or how long certain cellular events take? That could be cool and make it actually an interesting plot}
-\textit{this argument would be much better if you could find examples of things which need to be done quickly in the cell, like responding to foreign particles or signalling
-  molecules or cell division or neural events}
-
-\subsubsection{What role does it play in the cell?}
-Some of dynein's primary roles in the cell are to transport cellular cargos such as organelles, vesicles, mRNA, cytoskeletal filaments and certain proteins.
-Dynein also plays a role in positioning and breaking down the nucleus, cell death, spindle formation and the placement of other important cellular structures \cite{valetoolbox}.
-
-Microtubules are long polymers of alternating $\alpha-$ and $\beta-$ tubulin subunits.
-MTs are polar, meaning they have a distinct directionality based on the orientation of thetubulin proteins which comprise them.
-One pole, the minus end, is where they initiate formation, typically at a MTOC (microtubule organizing center) around the center of the cell.
-The other pole, the plus end, is where they grow and shrink from.
-Kinesin and myosin, two other families of motor protein, typically walk towards the plus end of the MT.
-Dynein is unique in that it walks towards the minus end, making it a minus end-directed motor.
-This lends it to a particularly important function inside neurons known as retrograde transport.\\
-
-Neurons have cell bodies at their centers and axons which grow outwards.
-Axons are long, narrow structures extending up to a meter in length in humans.
-Cell bodies contain nuclei and important organelles for synthesizing proteins, but growth of axon tips is vital for neuron function.
-This means bidirectional motor transport between axon tips and cell bodies is very important in neurons, since diffusion would not be quick enough.
-Retrograde transport is transport from the axon tip to the cell body.
-Cargo includes vesicles full of proteins ready to be broken down in the cell body and microtubule fragments to be returned from the axon tip \cite{neuroanatomy}.
-Because microtubule minus ends are found in the cell body, dynein's minus end-directed nature makes it the only protein capable of retrograde transport.
-An interesting question is, if proteins like dynein are synthesized in the cell body but needed at the axon tip to perform retrograde transport, how do dynein get to the tip?
-The answer is shown in \ref{retrograde-transport}: kinesin, a plus end-directed motor, takes dynein to the tip from the cell body.\\
-
-\begin{figure}[h]
+\begin{figure}
   \centering
-  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/retrograde_transport.jpg}
+  \includegraphics[width=.45\textwidth,keepaspectratio]{../figures/retrograde_transport.jpg}
   \caption{Mechanism of dynein localization to axon tips, or growth cones, in retrograde transport mediated by kinesin anterograde transport. Modified from \cite{valetoolbox}.}
@@
 
-\subsubsection{Dynein structure}
-
-\begin{figure}[h]
+\begin{figure}
   \centering
-  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/dynein-artist-rendition.jpg}
-  \caption{Artist rendition of the dynein motor bound to a microtubule. One monomer is bound to the MT, and one is raised up. Modified. Source: Lander Lab, The Scripps Research Institute.}
-  \label{dynein-artist-rendition-2}
-\end{figure}
-
-\textit{eventually find a different figure}
-
-As shown in \ref{dynein-artist-rendition-2}, a dynein motor is a fusion of to identical subunits, or monomers, each with binding, stalk, motor and tail subdomains.
-Each of these subunits has a unique purpose for the motor.
-
-\textbf{Binding domain}
-Microtubule binding domains (MTBDs) allow dynein to bind to the microtubule.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/mtbd-complex.png}
+  \includegraphics[width=.45\textwidth,keepaspectratio]{../figures/mtbd-complex.png}
   \caption{Crystal structure of the MTBD-MT complex. Modified from Redwine \textit{et. al} \cite{redwineMTBDcomplex}.}
@@
 
-\subsubsection{How does it behave?}
-Talk about its stepping behavior using Weihong's figures (as well as the new paper David found?)\\
-
-Also possibly talk about the other things Dynein has been found to do, such as hydrolyse ATP and
-move its subdomains around. This may be better in a different section than the introduction though,
-since its a lot of biochemistry for the second section in an intro? maybe not, it has to go somewhere.
-
-\subsubsection{Why is dynein special and interesting?}
-Compare the processive motion of dynein to kinesin in \textbf{a.)} directionality, \textbf{b.)}
-stepping pattern and \textbf{c.)} size (like how it can step over roadblocks kinesin cannot).\\
-
-Talk about why a random stepping motor is interesting compared to a orderly-stepping motor: what
-is different about dynein's structure that makes it step differently? Compare the crystal structures
-and ask lots of questions. Build interest in the motors here.\\
-
-%% \begin{figure}[h]
-%%   \centering
-%%   \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/diffusion_vs_dynein.png}
-%%   \caption{Crystal structures of cytoplasmic dynein 1 and kinesin KIF5B (PDB IDs 4RH7 and 1BG2).}
-%%   \label{kin_n_dyn}
-%% \end{figure}
-
-\textit{make this figure good -- find a full length kinesin crystal structure, not 1BG2. Use kindyn.pse as a template.}
-
-\subsection{How does dynein walk?}
-
-Cianfrocco \textit{et.
-al} \cite{cianfroccoreview} propose dynein achieves motion via transitioning through the cycle of states shown in Fig.
-\ref{mech-cycle}.
-This model will hereon be referred to as the Cianfrocco model.
-The model claims the primary events in the dynein step are as follows.
-First, starting with both binding domains attached to the MT, ATP binds to the motor domain, which causes it to change conformation.
-ATP hydrolysis to ADP-Pi causes further conformation changes.
-These changes cause the binding domain to release the microtubule, meaning only one binding domain is MT-bound.
-The free binding domain then undergoes another conformation change known as the ``priming stroke,'' which changes the orientation of the binding domain with respect to the MT.
-Finally the binding domain rebinds to the MT, causing a final conformation change in the motor which returns it to the beginning of the cycle.
-The exact order of these events is not precisely known.
-More detail on each step of the cycle is given below.\\
-
-Maybe show the Cianfrocco model in a section \textit{after} talking about all the things which happen for sure in dynein, since it's more theoretical?
-Talk about the Cianfrocco model, but other interpretations?
-
-\textbf{Things which are known about dynein which support the Cianfrocco model}
-\begin{itemize}
-\item AAA1 only ATPase necessary for motility, which is why we don't include the other ATPases in our model \cite{cianfroccoreview}
-\item AAA5 mediates stalk register changes Schmidt H, Zalyte R, Urnavicius L, Carter AP. 2014. Structure of human cytoplasmic dynein-2 primed for
-  its power stroke. Nature 518:435–38
-\item Linker moves as response to AAA1 nucleotide binding Burgess SA, Walker ML, Sakakibara H, Knight PJ, Oiwa K. 2003. Dynein structure and power stroke. Nature
-  421:715–8
-\item Crystal structures with Vanadium-Adp show dynein in the pre-powerstroke state, hinting that ADP-Pi-bound dynein takes on a different conformation than ADP-Pi free dynein
-\end{itemize}
-
-Make a figure of pre/post-powerstroke dynein with the linker hilighted? Can you get what residues correspond to the linker somewhere?
-
-\begin{figure}[h]
+\begin{figure}
   \centering
-  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/mechanochemical-cycle.jpeg}
+  \includegraphics[width=.45\textwidth,keepaspectratio]{../figures/mechanochemical-cycle.jpeg}
   \caption{Cycle of states the dynein motor goes through during motion in the Cianfrocco model \cite{cianfroccoreview}.}
@@
 
-\subsubsection{ATP binding / unbinding effects}
-\subsubsection{Microtubule un/rebinding}
-\subsubsection{Conformation change of motor (``priming stroke'')}
-A FRET (fluorescent resonance energy transfer) study, which probes the closeness of different labeled domains on a protein, showed that dynein is found in two different major conformations, depending on the nucleotide bound to its AAA1 ATPase site \cite{FRETstatepaper}.
-When no nucleotide, ATP or ADP is bound, the authors label the protein ``state I.'' When ADP-Pi bound, the motor is said to be in ``state II,'' as shown in Fig \ref{nucleotide-state}.
-In addition, motors in the presence of the unhydrolyzable ATP analog were found in state II, indicating the transition from state I to II occurs during ATP binding.
-Similar reasoning shows that the state II to I transition occurs when the protein is ADP bound.
-This allows the authors to create the state diagram shown in Fig \ref{nucleotide-state-diagram}.\\
-
-%% \begin{figure}[h]
-%%   \centering
-%%   \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/two-state-em-diagram}
-%%   \caption{}
-%%   \label{nuecleotide-state}
-%% \end{figure}
-
-%% \begin{figure}[h]
-%%   \centering
-%%   \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/nucleotide-state-cycle}
-%%   \caption{}
-%%   \label{nucleotide-state-diagram}
-%% \end{figure}
-
-This single FRET study illustrates how the conformational aspect of the mechanochemical cycle works.
-Dynein begins in state I and binds ATP into its AAA1 ATPase, changing into state II by changing its stalk-tail angle to move the tail outward.
-ATP is then hydrolized into ADP-Pi.
-Pi is then ejected from the AAA1 site, leaving ADP behind.
-This causes the reverse state transition, moving the tail inward.
-When combined with microtubule affinity changes, this primitive model of the dynein cycle could explain how dynein walks.\\
-
-\textit{It is interesting to note that the hydrolysis of ATP is not actually used to directly power the motor -- binding AMP-PMP, a nonhydrolyzable analog of ATP, can also cause the motor to change from state I to state II \cite{FRETstatepaper}.
-The specialness of ATP is that it can be broken down into ADP, which causes another conformation change back to state I.
-Thus, something something the transfer of energy from nucleotide to motor is in its shape, not in its chemistry.}\\
-
-\textbf{more on how ATP actually causes the conformation change, if that's available.
-The change from an open/closed ATPase is what changes the nucleotide affinity, go into depth on that, but actual info on how ATP binding causes conformation change would be good.
-Something about going from a planar to nonplanar conformation?}\\
-
-also how crystal structures are in different positions
-
-\subsubsection{Experimental results on dynein’s step}
-Weihong's figures go here.\\
-
-\subsubsection{c What is still needed to support theory}
-Where our simulation fits in. Why we need this kind of information.
-There are many different things which might be important for dynein's walk which we don't take into account: springiness of tail/stalk, changes in spring constants, repulsion between the heads, multiple pre/post powerstroke conformations.
-We are trying to see if these things are actually important.
-
-\subsubsection{Other theories?}
-Imamula model: how it is different from Cianfrocco model, and why we chose the Cianfrocco model.\\
-	\subsection{Past Simulations}
-		\subsubsection{Sarlah Winch Model}
-		\subsubsection{-x Other models}
-	\subsection{Brownian motion}
-                \subsubsection{Introduction/derivation}
-                \subsubsection{Justification}
-
-\subsection{Simulation}
-Talk about why simulation is the only way to verify the mechanochemical cycle, since experiments
-aren't high resolution enough to make good videos of the walk, and even if they were, in some sense
-knowing that dynein behaves identically to a simple mathematical model is more informative than
-a video, since there is lots of information in a video and really precise info in a simulation.\\
-
-\textbf{Not sure where this goes:} The 'goal' of our experiment, at a very high level, is to
-understand how dynein works. We can know this if we can come up with the minimum set of characteristics a physical system needs to act like dynein. This information is the same thing as knowing
-``how dynein walks.'' We don't need to know all the complex dynamics of every single amino acid in
-dynein, we only need to know our rough model.\\
-
-\subsection{Our model}
-Talk about what the important parts of the mechanochemical cycle from the introduction are, and how
-we take those and model them.\\
-\subsubsection{Rigid point model}
-Talk about two things: reducing bulky circular domains to points (is this the same as unified atom
-model?) with a certain drag constant based on radius, and acting like the domains are rigid in
-distance from each other. Talk about why this is valid (is it? it's an assumption, David may know
-more).
-\subsubsection{Harmonic energies}
-How we're modeling the complicated folding and adjusting of a massive protein as a harmonic
-oscillator. Why this is valid. Also other approaches...what else could we have modelled it as?
-\subsubsection{Two-state model}
-The mechanochemical cycle has eight states. We choose two as representative because \textbf{a.)}
-many of the states are identically conformationally or \textbf{b.)} the transitions happen so fast
-the intermediate states don't matter. Talk about rate constants here.
-\subsubsection{Limitations}
-Talk about the things we \textit{don't} take into account...motor-motor interaction, 2d information,
-other ATPase nucleotide state, saying certain mechanochemical steps occur ``fast,'' not taking into
-account all mechanochemical cycle states, making the coiled coil domain rigid even though it may be
-springy...
-
-Mention the Burgess paper, which says spring constants might change between apo/ADP states.
-
-\paragraph{Euler’s method}
-Show how Euler's method is arrived at, how its error scales, why we use it.
-
-\paragraph{Vs other methods (Runge Kutta, etc)}
-Talk about why our differential equation is most conducive to Euler's method. Euler's is more
-efficient -- is there a way to show how our differential equation is smooth, not jagged?
-
-\paragraph{Gibbs energy of transition}
-Do some stuff with Gibb's here? Why our energies are Gibb's and not internal energy?\\
-
-\section{Chapter: Methods}
-
-\subsection{Dynein Models}
-In order to simulate a protein's passage through the above mechanochemical cycle, two mathematical models are needed: one for the prestroke, MTBD-detached state, and one for the poststroke, MTBD-bound state.
-The purpose of each model is to calculate the position and velocity of each domain at an arbitrary time.\\
-
-\subsubsection{Prestroke onebound model}
-Prestroke dynein is referred to as onebound, since one motor domain is bound to the MT in this state, and the other unbound.
-A spatial representation of onebound dynein is shown in Figure \ref{ob_fig}.
-
-\begin{figure}[h]
+\begin{figure}
   \centering
@@
 
-TODO: make figure, label domain names (MTBD motor tail), Ls, Lt, microtubule\\
-
-Each domain is represented by a point, and is connected to other domains via fixed-length rods.
-Angles represent the degrees of freedom of the model.
-Since the protein is a dimer, there are two copies of the MTB, motor and tail domains.
-The tails are fused together and so represented as a single point.
-The motor and MTB domains are given a ``bound'' and ``unbound'' designation to differentiate them.
-The ``bound'' domains are the MTBD attached to the MT and its corresponding motor domain.
-The unbound domains are the unbound MTBD and its corresponding motor domain.\\
-
-\paragraph{Angular and Cartesian coordinates}
-
-The four angles $\Theta_{bb}, \Theta_{bm}, \Theta_{um}$ and $\Theta_{ub}$, corresponding to bound binding, bound motor, unbound motor and unbound binding angle, together describe any possible conformation the system can take on.
-The domain of $\Theta_{bb}$ has a restricted domain of $[0,\pi]$ to prevent below-MT conformations, but the other angles have domains of $[0,2\pi]$.
-Each angle is defined relative to the horizontal axis.\\
-
-\begin{figure}[h]
-  \centering
-  \begin{tabular}{c}
-    \begin{lstlisting}[language=C++]
-      class Dynein_onebound{
-        double &$\Theta_{bb}, \Theta_{bb}, \Theta_{bb}, \Theta_{bb}$&
-        double &$\dot{\Theta}_{bb}, \dot{\Theta}_{bb}, \dot{\Theta}_{bb}, \dot{\Theta}_{bb}$&
-        double &$X_{bb}$&
-      }
-    \end{lstlisting}
-  \end{tabular}
-  \caption{Pseudocode of the onebound Dynein model.}
-  \label{ob_struct}
-\end{figure}
-
-Pseudocode for the numerical onebound model representation is shown in Figure \ref{ob_struct}.
-The position of the bound binding domain $X_{bb}$ and the four domain angles are all the values needed to determine the position of the protein at any time.\\
-
-\paragraph{$\dot{\Theta}_{ub}, \dot{\Theta}_{um}, \dot{\Theta}_{bm}, \dot{\Theta}_{bb}$ calculation}
-
-The following equations allow recursive calculation of the positions of each domain in cartesian coordinates:
-
-\noindent\begin{minipage}{0.49\linewidth}
-\begin{align}
-  X_{bm} &= X_{bb}+L_{s}\cos(\Theta_{bb}) \\
-  X_{t}  &= X_{bm}+L_{t}\cos(\Theta_{bm}) \\
-  X_{fm} &= X_{t} - L_{t}\cos(\Theta_{fm}) \\
-  X_{fb} &= X_{fm} - L_{s}\cos(\Theta_{fb})
-\end{align}
-\end{minipage}
-\begin{minipage}{0.49\linewidth}
-\begin{align}
-  Y_{bb} &= 0 \\
-  Y_{bm} &= Y_{bb}+L_{s}\sin(\Theta_{bb}) \\
-  Y_{t}  &= Y_{bm}+L_{t}\sin(\Theta_{bm}) \\
-  Y_{fm} &= Y_{t} - L_{t}\sin(\Theta_{fm}) \\
-  Y_{fb} &= Y_{fm} - L_{s}\sin(\Theta_{fb})
-\end{align}
-\end{minipage}
-\vspace{.5cm}
-
-$L_s$ and $L_t$ correspond to lengths of each interdomain rod, and the $t$ subscripts refer to tail domain coordinates.\\
-
-\paragraph{Angular and Cartesian velocities}
-The goal is to express angular velocities $\dot{\Theta}_{bb}, \dot{\Theta}_{bb}, \dot{\Theta}_{bb}$ and $\dot{\Theta}_{bb}$ in terms of known quantities.
-To begin, the cartesian velocities of each domain can be calculated from the above position equations in a similar recursive manner:
-
-\noindent\begin{minipage}{0.49\linewidth}
-\begin{align}
-  \dot{X}_{bb} &= 0 \\
-  \dot{X}_{bm} &= \dot{X}_{bb} - L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} \label{cartesian-bmx}\\
-  \dot{X}_{t } &= \dot{X}_{bm} - L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} \\
-  \dot{X}_{fm} &= \dot{X}_{t } + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} \\
-  \dot{X}_{fb} &= \dot{X}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb}
-\end{align}
-\end{minipage}
-\begin{minipage}{0.49\linewidth}
-\begin{align}                                                                          
-  \dot{Y}_{bb} &= 0 \\                                                        
-  \dot{Y}_{bm} &= \dot{Y}_{bb} + L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} \\
-  \dot{Y}_{t}  &= \dot{Y}_{bm} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} \\
-  \dot{Y}_{fm} &= \dot{Y}_{t } - L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} \\
-  \dot{Y}_{fb} &= \dot{Y}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb}
-\end{align}
-\end{minipage}
-\vspace{.5cm}
-
-Another way to express these cartesian velocities is using the Brownian dynamics equation $\dot{X} = \frac1\gamma F_{net} + R$:
-
-\begin{align}  
-  \dot{X}_{bm} &= \frac{1}{\gamma_m} \Big(F_{xml} + - \lambda_{bs}(X_{bm} - X_{bb}) + \lambda_{bt}(X_{t } - X_{bm}) \Big) + R_{xml} \label{brownian-bmx}\\
-  \dot{X}_{t } &= \frac{1}{\gamma_t} \Big(F_{xt } + - \lambda_{bt}(X_{t } - X_{bm}) + \lambda_{ft}(X_{fm} - X_{t }) \Big) + R_{xt } \\
-  \dot{X}_{fm} &= \frac{1}{\gamma_m} \Big(F_{xmr} + - \lambda_{ft}(X_{fm} - X_{t }) + \lambda_{fs}(X_{fb} - X_{fm}) \Big) + R_{xmr} \\
-  \dot{X}_{fb} &= \frac{1}{\gamma_b} \Big(F_{xbr} + - \lambda_{fs}(X_{fb} - X_{fm}) \Big) + R_{xbr}
-\end{align}
-
-\begin{align}  
-  \dot{Y}_{bm} &= \frac{1}{\gamma_m} \Big(F_{yml} + - \lambda_{bs}(Y_{bm} - Y_{bb}) + \lambda_{bt}(Y_{t } - Y_{bm}) \Big) + R_{yml} \\
-  \dot{Y}_{t}  &= \frac{1}{\gamma_t} \Big(F_{yt } + - \lambda_{bt}(Y_{t } - Y_{bm}) + \lambda_{ft}(Y_{fm} - Y_{t }) \Big) + R_{yt } \\
-  \dot{Y}_{fm} &= \frac{1}{\gamma_m} \Big(F_{ymr} + - \lambda_{ft}(Y_{fm} - Y_{t }) + \lambda_{fs}(Y_{fb} - Y_{fm}) \Big) + R_{ymr} \\
-  \dot{Y}_{fb} &= \frac{1}{\gamma_b} \Big(F_{ybr} + - \lambda_{fs}(Y_{fb} - Y_{fm}) \Big) + R_{ybr}
-\end{align}
-
-where $\gamma_n$ is a drag coefficient for the binding, motor or tail domain with units of mass per second.
-$F_{xn}$ is the x component of external force on each domain $n$ due to various factors.
-$\lambda_{12}\left(X_1 - X_2\right)$ is the x component of internal force on domain $2$ due to domain $1$, where $\lambda$ is a tension coefficient with units of mass per second per second.
-$R_{xn}$ is the Brownian coefficient representing motion due to solvent collision, with units of velocity.\\
-
-These two sets of equations, cartesian and Brownian, can be equated to get more interesting equations.
-For example, Eq (\ref{cartesian-bmx}) and Eq (\ref{brownian-bmx}) can be equated.
-This equating, combined with expanding the recursive velocity definitions leads to:
-
-\begin{align}
-  -L_s\sin(\theta_{bb})\dot{\theta}_{bb} &= \frac{1}{\gamma_m}F_{xml} + -\frac{1}{\gamma_m}\lambda_{bs}(X_{bm} - X_{bb}) + \frac{1}{\gamma_m}\lambda_{bt}(X_{t } - X_{bm}) + R_{bmx} \label{ob_system_first}\\
-  -L_s\sin(\theta_{bb})\dot{\theta}_{bb} - L_t\sin(\theta_{bm})\dot{\theta}_{bm} &= \frac{1}{\gamma_t}F_{xt } + -\frac{1}{\gamma_t}\lambda_{bt}(X_{t } - X_{bm}) + \frac{1}{\gamma_t}\lambda_{ft}(X_{um} - X_{t }) + R_{tx} \\
-  -L_s\sin(\theta_{bb})\dot{\theta}_{bb} - L_t\sin(\theta_{bm})\dot{\theta}_{bm} + L_t\sin(\theta_{um})\dot{\theta}_{um} &= \frac{1}{\gamma_m}F_{xmr} + -\frac{1}{\gamma_m}\lambda_{ft}(X_{um} - X_{t }) + \frac{1}{\gamma_m}\lambda_{fs}(X_{ub} - X_{um}) + R_{bmx} \\
-  -L_s\sin(\theta_{bb})\dot{\theta}_{bb} - L_t\sin(\theta_{bm})\dot{\theta}_{bm} + L_t\sin(\theta_{um})\dot{\theta}_{um} + L_s\sin(\theta_{ub})\dot{\theta}_{ub} &= \frac{1}{\gamma_b}F_{xbr} + -\frac{1}{\gamma_b}\lambda_{fs}(X_{ub} - X_{um}) + R_{bbx}
-\end{align}
-
-\begin{align}
-  L_s\cos(\theta_{bb})\dot{\theta}_{bm} &= \frac{1}{\gamma_m}F_{yml} + -\frac{1}{\gamma_m}\lambda_{bs}(Y_{bm} - Y_{bb}) + \frac{1}{\gamma_m}\lambda_{bt}(Y_{t } - Y_{bm}) + R_{umy} \\
-  L_s\cos(\theta_{bb})\dot{\theta}_{bb} + L_t\cos(\theta_{bm})\dot{\theta}_{bm} &= \frac{1}{\gamma_t}F_{yt } + -\frac{1}{\gamma_t}\lambda_{bt}(Y_{t } - Y_{bm}) + \frac{1}{\gamma_t}\lambda_{ft}(Y_{um} - Y_{t }) + R_{ty} \\
-  L_s\cos(\theta_{bb})\dot{\theta}_{bb} + L_t\cos(\theta_{bm})\dot{\theta}_{bm} - L_t\cos(\theta_{um})\dot{\theta}_{um} &= \frac{1}{\gamma_m}F_{ymr} + -\frac{1}{\gamma_m}\lambda_{ft}(Y_{um} - Y_{t }) + \frac{1}{\gamma_m}\lambda_{fs}(Y_{ub} - Y_{um}) + R_{bmy} \\
-  L_s\cos(\theta_{bb})\dot{\theta}_{bb} + L_t\cos(\theta_{bm})\dot{\theta}_{bm} - L_t\cos(\theta_{um})\dot{\theta}_{um} - L_s\cos(\theta_{ub})\dot{\theta}_{ub} &= \frac{1}{\gamma_b}F_{ybr} + -\frac{1}{\gamma_b}\lambda_{fs}(Y_{ub} - Y_{um}) + R_{bby} \label{ob_system_last}
-\end{align}
-
-These eight equations form a system of equations with eight unknowns: $\dot{\Theta}_{bb}, \dot{\Theta}_{bb}, \dot{\Theta}_{bb}$ and $\dot{\Theta}_{bb}$, and the four tension coefficients $\lambda_{bs}, \lambda_{bt}, \lambda_{um}$ and $\lambda_{ub}$.
-This system is more compactly represented as:
-{\small
-  \[
-  \hspace{-1.0cm}
-  \resizebox{0.85\linewidth}{!}{%
-  $\begin{pmatrix}
-    L_s\sin\theta_{bb} & 0 & 0 & 0 & -\gamma_m (X_{bm} - X_{bb}) & \gamma_m (X_{t } - X_{bm}) & 0 & 0\\
-    L_s\sin(\theta_{bb}) & L_t\sin(\theta_{bm}) & 0 & 0 & 0 & -\gamma_t (X_{t } - X_{bm}) & \gamma_t (X_{um} - X_{t }) & 0\\
-    L_s\sin(\theta_{bb}) & L_t\sin(\theta_{bm}) & -L_t\sin(\theta_{um}) & 0 & 0 & 0 & -\gamma_m (X_{um} - X_{t }) & \gamma_m (X_{ub} - X_{um})\\
-    L_s\sin(\theta_{bb}) & L_t\sin(\theta_{bm}) & -L_t\sin(\theta_{um}) & -L_s\sin(\theta_{ub}) & 0 & 0 & 0 & -\gamma_b (X_{ub} - X_{um})\\
-    -L_s\cos(\theta_{bb}) & 0 & 0 & 0 & -\gamma_m (Y_{bm} - Y_{bb}) & \gamma_m (Y_{t } - Y_{bm}) & 0 & 0\\
-    -L_s\cos(\theta_{bb}) & -L_t\cos(\theta_{bm}) & 0 & 0 & 0 & -\gamma_t (Y_{t } - Y_{bm}) & \gamma_t (Y_{um} - Y_{t }) & 0\\
-    -L_s\cos(\theta_{bb}) & -L_t\cos(\theta_{bm}) & L_t\cos(\theta_{um}) & 0 & 0 & 0 & -\gamma_m (Y_{um} - Y_{t }) & \gamma_m (Y_{ub} - Y_{um})\\
-    -L_s\cos(\theta_{bb}) & -L_t\cos(\theta_{bm}) & L_t\cos(\theta_{um}) & L_s\cos(\theta_{ub}) & 0 & 0 & 0 & -\gamma_b (Y_{ub} - Y_{um})\\
-  \end{pmatrix}$%
-  }
-  \begin{pmatrix}
-    \dot{\theta}_{bb}\\
-    \dot{\theta}_{bm}\\
-    \dot{\theta}_{um}\\
-    \dot{\theta}_{ub}\\
-    \lambda_{bs}\\
-    \lambda_{bt}\\
-    \lambda_{ut}\\
-    \lambda_{us}\\
-  \end{pmatrix}
-  =
-  \begin{pmatrix}
-    -F_{bmx} + \gamma_m R_{bmx}\\
-    -F_{tx} + \gamma_t R_{tx}\\
-    -F_{umx} + \gamma_m R_{umx}\\
-    -F_{ubx} + \gamma_b R_{ubx}\\
-    -F_{bmy} + \gamma_m R_{bmy}\\
-    -F_{ty} + \gamma_t R_{ty}\\
-    -F_{umy} + \gamma_m R_{umy}\\
-    -F_{uby} + \gamma_b R_{uby}\\
-  \end{pmatrix}
-  \]
-}
-
-This matrix is then solved using the Mathematica computer algebra system, resulting in a set of $\dot{\Theta}_n$ equations shown in Appendix (\ref{ob-motion-equations]}).
-
-\subsubsection{Poststroke bothbound model}
-The main difference between the prestroke and poststroke mathematical models is that the latter has a further constraint: both its MTB domains are attached to the MT.
-This has the effect of diminishing the degrees of freedom of the system from four to two, assuming the motor remains above the MT.
-A spatial representation of the bothbound model is shown in Figure (\ref{bb_fig}).\\
-
-\begin{figure}[h]
+\begin{figure}
   \centering
@@
 
-The bothbound model is very similar in assumptions to the onebound: domains are represented as points connected by rigid rods, and the tails are still fused into a single point.
-The main differences are the naming scheme and spatial constraints.
-Dimers can no longer be differentiated by being un/bound, so are now differentiated by their z-position: being near or far from the ``camera.'' This is similar to labeling them as the right or left dimer.
-Thus there are the near binding (nb), near motor (nm), tail (t), far motor (fm) and far binding (fb) domains.
-Both the near and far binding domains have their x-coordinate fixed at zero.
-The system is described completely by two angles: $\Theta_{nm}$ and $\Theta_{fm}$.
-Angles $\Theta_t, \Theta_{nb}$ and $\Theta_{fb}$ are dependent on these two motor angles.\\
-
-The second MTBD constraint adds another required piece of information the model must track: the x position of the second MTBD.
-This position is described by $L$, the distance between the near and far binding domains.
-The pseudocode used to describe the bothbound model are shown in Figure (\ref{bb_struct}).\\
-
-\begin{figure}[h]
+\begin{figure}[h!]
   \centering
-  \begin{tabular}{c}
-    \begin{lstlisting}[language=C++]
-      class Dynein_bothbound{
-        double &$\Theta_{nm}, \Theta_{fm}$&
-        double &$\dot{\Theta}_{nm}, \dot{\Theta}_{fm}$&
-        double &$X_{nb, L}$&
-      }
-    \end{lstlisting}
-  \end{tabular}
-  \caption{Pseudocode of the bothbound Dynein model.}
-  \label{ob_struct}
+  \begin{minipage}[b]{0.49\textwidth}
+    \includegraphics[width=\textwidth]{../figures/OB_Average_PE.pdf}
+    \caption{Onebound PE vs time.}
+  \end{minipage}
+  \begin{minipage}[b]{0.49\textwidth}
+    \includegraphics[width=\textwidth]{../figures/BB_Average_PE.pdf}
+    \caption{Bothbound PE vs time.}
+  \end{minipage}
+  \label{equipartition_agreement}
 \end{figure}
 
-\paragraph{$\dot{\Theta}_{nm}, \dot{\Theta}_{fm}$ calculation}
-
-The first step in calculating angular velocities is to define new variables $L_n$ and $L_f$ using the Law of Cosines:
-
-\begin{align}
-  L_n &= \sqrt{L_s^2 + L_t^2 - 2L_sL_t\cos{\theta_{nm}}} \\
-  L_f &= \sqrt{L_s^2 + L_t^2 - 2L_sL_t\cos{\theta_{fm}}}
-\end{align}
-
-These new artificial stalks are not physically relevant, but allow the definition of new angles $\Theta_n, \Theta_{ns}, \Theta_{f}$ and $\Theta_{fs}$, which are useful for calculating the position of the motor domain.
-The treatment of the near and far versions of these angles is very similar, so only the near angles will be dealt with here.
-The angles themselves are never dealt with, only their $\sin$s and $\cos$s:
-
-\begin{align}
-  \cos\theta_n &= \frac{L^2 + L_n^2 - L_f^2}{2L L_n} \\
-  \sin\theta_{n} &= \sqrt{1 - \cos^2\theta_{n}} \\
-  \cos\theta_{ns} &= \frac{L_s^2 + L_n^2 - L_t^2}{2L_s L_n} \\
-  \sin\theta_{ns} &=
-  \begin{cases}
-    +\sqrt{1 - \cos^2\theta_{ns}} & \theta_{nm} < \pi \\
-    -\sqrt{1 - \cos^2\theta_{ns}} & \theta_{nm} > \pi
-  \end{cases}
-\end{align}
-
-These values are arrived at through the Law of Cosines and a trigonometric identity.
-The sign of $\sin\left(\Theta_n\right)$ is restricted to positive values due to the angle's $[0,\pi]$ domain.
-$\Theta_{ns}$'s domain is $[-\pi,\pi]$, meaning its sign is not restricted.
-The partial derivatives of these values are useful:
-
-\begin{align}
-  \frac{\partial \cos\theta_n}{\partial L_n} &= \frac{1}{L} - \frac{L^2 + L_n^2 - L_f^2}{2L L_n^2}\\
-  \frac{\partial \cos\theta_n}{\partial L_f} &= -\frac{L_f}{LL_n}\\
-  \frac{\partial \sin\theta_n}{\partial L_n} &= \frac{-\cos\theta_n}{\sqrt{1-\cos^2\theta_n}}
-  \left(\frac{1}{L} - \frac{L^2 + L_n^2 - L_f^2}{2L L_n^2}\right)\\
-  \frac{\partial \sin\theta_n}{\partial L_f} &= \frac{-\cos\theta_n}{\sqrt{1-\cos^2\theta_n}}
-  \left(\frac{-L_f}{LLn}\right)\\
-  \frac{\partial \cos\theta_{ns}}{\partial L_n} &= \frac{1}{L_s}
-  - \frac{L_s^2 + L_n^2 - L_t^2}{2L_sL_n^2}\\
-  \frac{\partial \cos\theta_{ns}}{\partial L_f} &= 0\\
-  \frac{\partial \sin\theta_{ns}}{\partial L_n} &=
-  \begin{cases}
-    \frac{-\cos\theta_{ns}}{\sqrt{1-\cos^2\theta_{ns}}}
-    \left(\frac{1}{L_s} - \frac{L_s^2+L_n^2-L_t^2}{2LL_n^2}\right) & \theta_{nm} < \pi \\
-    \frac{\cos\theta_{ns}}{\sqrt{1-\cos^2\theta_{ns}}}
-    \left(\frac{1}{L_s} - \frac{L_s^2+L_n^2-L_t^2}{2LL_n^2}\right) & \theta_{nm} > \pi
-  \end{cases}\\
-  \frac{\partial \sin\theta_{ns}}{\partial L_f} &= 0
-\end{align}
-
-The position of the motor and tail domains is calculated as follws, using the angle addition trigonometric identity:
-
-\begin{align}
-  X_{nm} &= L_s\left(
-  \cos\theta_n\cos\theta_{ns} - \sin\theta_n\sin\theta_{ns}
-  \right)
-  \\
-  Y_{nm} &= L_s\left(
-  \cos\theta_n\sin\theta_{ns} + \sin\theta_n\cos\theta_{ns}
-  \right)
-  \\
-  X_{t} &= L_n\cos\theta_n\\
-  Y_{t} &= L_n\sin\theta_n\\
-\end{align}
-
-The partial derivatives of these values are:
-
-\begin{align}
-  \frac{dX_{nm}}{dL_n} &= L_s\Big(\cos\theta_n\frac{d\cos\theta_{ns}}{dL_n} + \cos\theta_{ns}\frac{d\cos\theta_{n}}{dL_n} - \sin\theta_n\frac{d\sin\theta_{ns}}{dL_n} - \sin\theta_{ns}\frac{d\sin\theta_{n}}{dL_n}\Big)\\
-  \frac{dY_{nm}}{dL_n} &= L_s\Big(\cos\theta_n\frac{d\sin\theta_{ns}}{dL_n} + \sin\theta_{ns}\frac{d\cos\theta_{n}}{dL_n} + \sin\theta_n\frac{d\cos\theta_{ns}}{dL_n} + \cos\theta_{ns}\frac{d\sin\theta_{n}}{dL_n} \Big)\\
-  \frac{dX_{nm}}{dL_f} &= L_s\Big(\cos\theta_n\frac{d\cos\theta_{ns}}{dL_f} + \cos\theta_{ns}\frac{d\cos\theta_{n}}{dL_f} - \sin\theta_n\frac{d\sin\theta_{ns}}{dL_f} - \sin\theta_{ns}\frac{d\sin\theta_{n}}{dL_f} \Big)\\
-  \frac{dY_{nm}}{dL_f} &= L_s\Big(\cos\theta_n\frac{d\sin\theta_{ns}}{dL_f} + \sin\theta_{ns}\frac{d\cos\theta_{n}}{dL_f} + \sin\theta_n\frac{d\cos\theta_{ns}}{dL_f} + \cos\theta_{ns}\frac{d\sin\theta_{n}}{dL_f}\Big)\\
-  \frac{dX_{t}}{dL_n} &= \cos\theta_n + L_n\frac{\partial \cos\theta_n}{\partial L_n} = \frac{L_n}{L}\\
-  \frac{dY_{t}}{dL_n} &= \sin\theta_n + L_n\frac{\partial \sin\theta_n}{\partial L_n}\\
-  \frac{dX_{t}}{dL_f} &= L_n\frac{\partial \cos\theta_n}{\partial L_f} = -\frac{L_f}{L}\\
-  \frac{dY_{t}}{dL_f} &= L_n\frac{\partial \sin\theta_n}{\partial L_f}\\
-\end{align}
-
-These partial derivatives are then used in a system of equations very similar to that in Equations (\ref{ob_system_first}-\ref{ob_system_last}), combining Brownian and coordinate system definitions of domain velocities:
-
-\begin{align}
-  \dot{X}_{nm} &= \frac{1}{\gamma} \Big(F_{xml} - \lambda_{ns}(X_{bm} - X_{bb})
-  + \lambda_{nt}(X_{t } - X_{bm}) \Big) + R_{xml}
-  &=& \frac{\partial X_{nm}}{\partial L_n}\dot{L}_n + \frac{\partial X_{nm}}{\partial L_f}\dot{L}_f\\
-  \dot{X}_{t } &= \frac{1}{\gamma} \Big(F_{xt } - \lambda_{nt}(X_{t } - X_{bm})
-  + \lambda_{ft}(X_{fm} - X_{t }) \Big) + R_{xt }
-  &=& \frac{\partial X_{t}}{\partial L_n}\dot{L}_n + \frac{\partial X_{t}}{\partial L_f}\dot{L}_f\\
-  \dot{X}_{fm} &= \frac{1}{\gamma} \Big(F_{xmr} - \lambda_{ft}(X_{fm} - X_{t })
-  + \lambda_{fs}(X_{fb} - X_{fm}) \Big) + R_{xmr}
-  &=& \frac{\partial X_{fm}}{\partial L_n}\dot{L}_n + \frac{\partial X_{fm}}{\partial L_f}\dot{L}_f
-\end{align}
-
-\begin{align}
-  \dot{Y}_{nm} &= \hspace{1cm} \frac{1}{\gamma} \Big(F_{yml} - \lambda_{ns}(Y_{bm} - Y_{bb})
-  + \lambda_{nt}(Y_{t } - Y_{bm}) \Big) + R_{yml}
-  &= \frac{\partial Y_{nm}}{\partial L_n}\dot{L}_n + \frac{\partial Y_{nm}}{\partial L_f}\dot{L}_f\\
-  \dot{Y}_{t}  &= \hspace{1cm} \frac{1}{\gamma} \Big(F_{yt } - \lambda_{nt}(Y_{t } - Y_{bm})
-  + \lambda_{ft}(Y_{fm} - Y_{t }) \Big) + R_{yt }
-  &= \frac{\partial Y_{t}}{\partial L_n}\dot{L}_n + \frac{\partial Y_{t}}{\partial L_f}\dot{L}_f\\
-  \dot{Y}_{fm} &= \hspace{1cm} \frac{1}{\gamma} \Big(F_{ymr} - \lambda_{ft}(Y_{fm} - Y_{t })
-  + \lambda_{fs}(Y_{fb} - Y_{fm}) \Big) + R_{ymr}
-  &= \frac{\partial Y_{fm}}{\partial L_n}\dot{L}_n + \frac{\partial Y_{fm}}{\partial L_f}\dot{L}_f
-\end{align}
-
-The matrix version of this system looks like:
-
-\[
-\begin{pmatrix}
-  -\frac{X_{nm}}{Ln} & -\frac{X_{nm}}{Lf}
-  & -\frac{X_{nm}-X_{nb}}{\gamma} & \frac{X_{t}-X_{nm}}{\gamma} & 0 & 0\\
-  -\frac{X_{t}}{Ln} & -\frac{X_{t}}{Lf}
-  & 0 & -\frac{X_{t}-X_{nm}}{\gamma} & \frac{X_{fm}-X_{t}}{\gamma} & 0\\
-  -\frac{X_{fm}}{Ln} & -\frac{X_{fm}}{Lf}
-  & 0 & 0 & -\frac{X_{fm}-X_{t}}{\gamma} & \frac{X_{fb}-X_{fm}}{\gamma}\\
-  -\frac{Y_{nm}}{Ln} & -\frac{Y_{nm}}{Lf}
-  & -\frac{Y_{nm}-Y_{nb}}{\gamma} & \frac{Y_{t}-Y_{nm}}{\gamma} & 0 & 0\\
-  -\frac{Y_{t}}{Ln} & -\frac{Y_{t}}{Lf}
-  & 0 & -\frac{Y_{t}-Y_{nm}}{\gamma} & \frac{Y_{fm}-Y_{t}}{\gamma} & 0\\
-  -\frac{Y_{fm}}{Ln} & -\frac{Y_{fm}}{Lf}
-  & 0 & 0 & -\frac{Y_{fm}-Y_{t}}{\gamma} & \frac{Y_{fb}-Y_{fm}}{\gamma}\\
-\end{pmatrix}
-\begin{pmatrix}
-  \dot{L}_n\\
-  \dot{L}_f\\
-  \lambda_{ns}\\
-  \lambda_{nt}\\
-  \lambda_{ft}\\
-  \lambda_{fs}\\
-\end{pmatrix}
-=
-\begin{pmatrix}
-  -\frac{1}{\gamma}F_{nmx} - R_{nmx}\\
-  -\frac{1}{\gamma}F_{tx}  - R_{tx}\\
-  -\frac{1}{\gamma}F_{fmx} - R_{fmx}\\
-  -\frac{1}{\gamma}F_{nmy} - R_{nmy}\\
-  -\frac{1}{\gamma}F_{ty}  - R_{ty}\\
-  -\frac{1}{\gamma}F_{fmy} - R_{fmy}\\
-\end{pmatrix}
-\]
-
-This system of equations is then solved and used to calculate $\dot{L}_n$ and $\dot{L}_f$, as shown in Appendix (\ref{both-motion-equations}).
-
-\subsection{Simulation}
-To generate stepping data, the models were coded in \verb|C++| and time-evolved using Euler's method to generate reasonable dynamical behavior.
-When in a position deemed worthy of a state transition a conversion, the model was converted to the respective other model, representing part of a ``step'' through the mechanochemical cycle.
-Some unphysical but minor corrections at points in the simulation were needed to rescue the model from entering incalculable positions.\\
-
-Code is available at \verb|https://github.com/elliotc12/dynein_walk.git|.
-
-\textit{todo: come up with a better word than ``model'' to describe the state of the system at any time}
-\subsubsection{Time evolution}
-Euler's method for solving differential equations was used to calculate domain positions through time.
-Because each model is a set of multiple coupled differential equations, care was taken to update each domain position at the same time, and not in a cascade.
-A code snippet is shown in Figure (\ref{euler-code-snippet}) for the onebound updating scheme.
-
-\begin{figure}[h]
-  \centering
-  \begin{tabular}{c}
-    \begin{lstlisting}[language=C++]
-      double temp_bba = dyn_ob->get_bba() + dyn_ob->get_d_bba() * dt;
-      double temp_bma = dyn_ob->get_bma() + dyn_ob->get_d_bma() * dt;
-      double temp_uma = dyn_ob->get_uma() + dyn_ob->get_d_uma() * dt;
-      double temp_uba = dyn_ob->get_uba() + dyn_ob->get_d_uba() * dt;
-
-      dyn_ob->set_bba(temp_bba);
-      dyn_ob->set_bma(temp_bma);
-      dyn_ob->set_uma(temp_uma);
-      dyn_ob->set_uba(temp_uba);
-    \end{lstlisting}
-  \end{tabular}
-  \caption{Code snippet of Euler's method to update the positions of the onebound model simultaneously.}
-  \label{ob_struct}
-\end{figure}
-
-The \verb|dt| used to update was $10^{-11}$ seconds.
-
-\subsubsection{Transitioning states}
-Transitions from one state to another are accomplished by calculating the rate of transition in a certain position, then using this rate divided by \verb|dt| as a probability of transfer at each timestep.
-The rates of transition are calculated by assuming they follow the Arrhenius rate equation $r = Ae^{-\Delta G/k_BT}$, where the preexponential factor is a parameter of the model and $\Delta G$ the energy of transition between the repsective states.
-A code snippet of this rate calculation is shown in Figure (\ref{transition-rate-snippet}).
-
-\begin{figure}[h]
-  \begin{tabular}{c}
-    \begin{lstlisting}[language=C++]
-      double Dynein_bothbound::get_near_unbinding_rate() {
-        double dG_spring = Dynein_onebound(this, rand, NEARBOUND).get_PE() - get_PE();
-        double low_affinity_unbinding_preexponential_factor = low_affinity_unbinding_rate / exp(1.0);
-        return low_affinity_unbinding_preexponential_factor*exp(-dG_spring/kb/T);
-      }
-    \end{lstlisting}
-  \end{tabular}
-  \caption{Code snippet of transition rate calculation using the Arrhenius equation.}
-  \label{transition-rate-snippet}
-\end{figure}
-
-\textit{todo: include lots more on the derivation of preexponential factors from experimental rates and loss of degrees of freedom}
-
-\subsubsection{Corrections}
-Certain configurations of the model, such as in the bothbound model when $\Theta_{nm} = \pi$ or $\Theta_{fm} = \pi$, it becomes impossible to calculate intermediate variables due to values going to infinity (name for this??).
-For example when $\Theta_{nm} = \pi$, $Ln = Ls + Lt$ and so:
-
-\begin{align}
-  \cos\theta_{ns} &= 1 \\
-  \sin\theta_{ns} &= 0 \\
-  \mbox{some val?} &\rightarrow \infty
-\end{align}
-
-Alternative problem states are when $L = 0$ and ???. These states are corrected similarly.\\
-
-\subsubsection{Forces}
-Spring and Brownian forces are used to properly move the model through time.
-Appendices (\ref{onebound-motion-equatons}-\ref{bothbound-motion-equations}) show angular velocities are functions of the x- and y-components of both spring forces $F$ and Brownian forces $\gamma R$.
-We calculate these as shown:
-
-\paragraph{Brownian forces}
-Each Brownian force $R_{n,m}$ on domain $n$ in direction $m$ is sampled from a Gaussian with standard deviation $\sqrt{\frac{2k_BT\gamma_n}{dt}}$.\\
-
-\paragraph{Spring forces}
-Spring forces are calculated by first finding the torque $\tau$ on each angle, then converting this to a force.
-For example, the onebound near motor angle feels a force $\tau_{nm} = c_m(\Theta_{nm}-\Theta_{nm,eq})$, where $c_n$ is the spring constant of the motor domain and $\Theta_{nm,eq}$ the equilibrium angle.
-The magnitude of force on the adjacent binding and tail domains is thus $\tau/Ls$ and $\tau/Lt$, pointing in the direction adjacent the Ls and Lt rods, respectively.
-The force on the motor domain is found as the equal-and-opposite force to these two imposed forces.
-In code:
-
-\begin{figure}[h]
-  \centering
-  \begin{tabular}{c}
-    \begin{lstlisting}[language=C++]
-      T = cm*(nma - nma_eq);
-      f1 = T/Ls;
-      f2 = T/Lt;
-      f1x = f1 * sin(bba);
-      f1y = f1 * -cos(bba);
-      f2x = f2 * sin(bma);
-      f2y = f2 * -cos(bma);
-      f.bbx += f1x;
-      f.bby += f1y;
-      f.tx  += f2x;
-      f.ty  += f2y;
-      f.bmx += -(f1x + f2x); // equal and opposite forces
-      f.bmy += -(f1y + f2y);
-    \end{lstlisting}
-  \end{tabular}
-  \caption{Code snippet of internal force calculation.}
-  \label{transition-rate-snippet}
-\end{figure}
-
-\subsection{Verifying the model}
-\subsubsection{Equipartition theorem obeyed}
-%% \begin{figure}[h!]
-%%   \centering
-%%   \begin{minipage}[b]{0.49\textwidth}
-%%     \includegraphics[width=\textwidth]{../figures/OB_Average_PE.pdf}
-%%     \caption{Onebound PE vs time.}
-%%   \end{minipage}
-%%   \begin{minipage}[b]{0.49\textwidth}
-%%     \includegraphics[width=\textwidth]{../figures/BB_Average_PE.pdf}
-%%     \caption{Bothbound PE vs time.}
-%%   \end{minipage}
-%%   \label{fig:equipartition_agreement}
-%% \end{figure}
-
-\subsection{Data collection}
-
-\subsubsection{Outputting positions, step locations}
-\subsubsection{Reconstructing steps from step locations}
-\subsubsection{Job management}
-
-\subsection{Parameter fitting}
-\subsubsection{Fitting spring constants}
-\subsubsection{Finding parameters from PyMol}
-\subsubsection{Finding values from literature}
-\subsubsection{Final parameter guess table}
-\subsubsection{Finding the right timestep}
-
-------------------------------------------
-
-\section{Chapter: Results}
-	\subsection{Histograms}
-		\subsubsection{Step length histograms}
-		\subsubsection{Step time histograms}
-	\subsection{Other ways to represent data?}
-
-\section{Chapter: Discussion}
-\subsection{Why our model worked, didn’t work}
-Fill this out over winter when you know more.\\
-\subsection{Future work on this project}
-If setbacks happen, this'll be a big section. Or, further simulations which could be done.
-
-\subsubsection{Experiments to verify our model}
-Talk about how FRET could be used to verify our model. We have domain distance info over long
-timescales in our model - this is something we could compare to a FRET experiment exploring
-the distance between domains in dynein.\\
-
-Other biophysics experiments which could verify our model?\\
-
-Do our spring constants seem to indicate that the protein does about one ATP's worth of work through its cycle? I wonder if the motor cools down its surroundings when it walks...\\
-
-Calorimetry could get Gibb's energy of binding if you froze the motor in onebound using
-vanadium? Maybe...\\
-\subsection{Possible further projects}
-\subsection{Difference between dynein and kinesin}
-Our project un/successfully predicted how dynein behaved. How would we modify it to describe kinesin?
-Is dynein's linker-motor connection less tense than kinesin's, leading to more random motion? This
-would likely be lots of speculation, but it could be cool to look up some papers on kinesin motility
-and get its crystal structure to do some comparisons.
-
-\bibliography{thesis-articles}
-
-%% \bibliography{thesis-webpages}
-
-\section{Appendix}
-
-\subsection{Motion equations}
-
-\subsubsection{Onebound equations}
-\label{onebound-motion-equations}
-The following are the full motion equations used to time evolve the onebound model:
-
-\begin{align*}
-AA &= L_s\sin\theta_{bb}      &   LL &= -L_s\cos(\theta_{bb})\\
-BB &= L_t\sin(\theta_{bm})    &	 MM &= -L_t\cos(\theta_{bm})\\
-CC &= -L_t\sin(\theta_{um})   &	 NN &= L_t\cos(\theta_{um})\\
-DD &= -L_s\sin(\theta_{ub})   &	 OO &= L_s\cos(\theta_{ub})\\
-\\
-EE &= -\gamma_m (X_{bm} - X_{bb})       &  PP &= -\gamma_m (Y_{bm} - Y_{bb})\\
-FF &= \gamma_m (X_{t } - X_{bm})        &	 QQ &= \gamma_m (Y_{t } - Y_{bm})\\ 
-GG &= -\gamma_t (X_{t } - X_{bm})       &	 RR &= -\gamma_t (Y_{t } - Y_{bm})\\
-HH &= \gamma_t (X_{um} - X_{t })        &	 SS &= \gamma_t (Y_{um} - Y_{t })\\ 
-II &= -\gamma_m (X_{um} - X_{t })       &	 TT &= -\gamma_m (Y_{um} - Y_{t })\\
-JJ &= \gamma_m (X_{ub} - X_{um})        &	 UU &= \gamma_m (Y_{ub} - Y_{um})\\ 
-KK &= -\gamma_b (X_{ub} - X_{um})       &	 VV &= -\gamma_b (Y_{ub} - Y_{um})\\
-\end{align*}%
-%
-\[
-\begin{pmatrix}
-  AA & 0 & 0 & 0 & EE & FF & 0 & 0\\
-  AA & BB & 0 & 0 & 0 & GG & HH & 0\\
-  AA & BB & CC & 0 & 0 & 0 & II & JJ\\
-  AA & BB & CC & DD & 0 & 0 & 0 & KK\\
-  LL & 0 & 0 & 0 & PP & QQ & 0 & 0\\
-  LL & MM & 0 & 0 & 0 & RR & SS & 0\\
-  LL & MM & NN & 0 & 0 & 0 & TT & UU\\
-  LL & MM & NN & OO & 0 & 0 & 0 & VV\\
-\end{pmatrix}
-\begin{pmatrix}
-    \dot{\theta}_{bb}\\
-    \dot{\theta}_{bm}\\
-    \dot{\theta}_{um}\\
-    \dot{\theta}_{ub}\\
-    \lambda_{bs}\\
-    \lambda_{bt}\\
-    \lambda_{ut}\\
-    \lambda_{us}\\
-  \end{pmatrix}
-=
-\begin{pmatrix}
-  X1\\
-  X2\\
-  X3\\
-  X4\\
-  X5\\
-  X6\\
-  X7\\
-  X8\\
-\end{pmatrix}
-\]
-
-Solving the above system for $\dot{\Theta}_{bb}, \dot{\Theta}_{bm}, \dot{\Theta}_{um}$ and $\dot{\Theta}_{ub}$ gives
-
-$\dot{\Theta}_{bb}$ =
-{\tiny
-( - (GG*II*MM*NN*PP*X1) + CC*HH*MM*PP*RR*X1 - BB*HH*NN*PP*RR*X1 + BB*II*NN*PP*RR*X1 - CC*GG*MM*PP*SS*X1 + BB*GG*NN*PP*SS*X1 + CC*GG*MM*PP*TT*X1 - BB*CC*PP*RR*TT*X1 + FF*II*MM*NN*PP*X2 - EE*II*MM*NN*QQ*X2 + CC*FF*MM*PP*SS*X2 - BB*FF*NN*PP*SS*X2 - CC*EE*MM*QQ*SS*X2 + BB*EE*NN*QQ*SS*X2 - CC*FF*MM*PP*TT*X2 + CC*EE*MM*QQ*TT*X2 - FF*HH*MM*NN*PP*X3 + EE*HH*MM*NN*QQ*X3 + BB*FF*NN*PP*SS*X3 - BB*EE*NN*QQ*SS*X3 + EE*GG*II*MM*NN*X5 - CC*EE*HH*MM*RR*X5 + BB*EE*HH*NN*RR*X5 - BB*EE*II*NN*RR*X5 + CC*EE*GG*MM*SS*X5 - BB*EE*GG*NN*SS*X5 - CC*EE*GG*MM*TT*X5 + BB*CC*EE*RR*TT*X5 - CC*FF*HH*MM*PP*X6 + BB*FF*HH*NN*PP*X6 - BB*FF*II*NN*PP*X6 + CC*EE*HH*MM*QQ*X6 - BB*EE*HH*NN*QQ*X6 + BB*EE*II*NN*QQ*X6 + BB*CC*FF*PP*TT*X6 - BB*CC*EE*QQ*TT*X6 + CC*FF*HH*MM*PP*X7 - CC*EE*HH*MM*QQ*X7 - BB*CC*FF*PP*SS*X7 + BB*CC*EE*QQ*SS*X7)/(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT) - ((FF*PP - EE*QQ)*(HH*MM - BB*SS)*(JJ*NN - CC*UU)*( - ((( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( - ((( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*NN*X2 +  Power(AA,4)* Power(BB,2)*NN*X3 -  Power(AA,4)* Power(BB,2)*CC*X7)))) + ( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( - ((( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,5)* Power(BB,3)*CC*DD*LL*X1 -  Power(AA,6)* Power(BB,2)*CC*DD*MM*X1 +  Power(AA,6)* Power(BB,2)*CC*DD*MM*X2 -  Power(AA,6)* Power(BB,3)*DD*NN*X2 +  Power(AA,6)* Power(BB,3)*DD*NN*X3 -  Power(AA,6)* Power(BB,3)*CC*OO*X3 +  Power(AA,6)* Power(BB,3)*CC*OO*X4 -  Power(AA,6)* Power(BB,3)*CC*DD*X8)))))/(( - (EE*GG*II*LL*MM*NN) - BB*FF*HH*LL*NN*PP + BB*FF*II*LL*NN*PP + AA*FF*HH*MM*NN*PP - AA*FF*II*MM*NN*PP + AA*GG*II*MM*NN*PP + BB*EE*HH*LL*NN*QQ - BB*EE*II*LL*NN*QQ - AA*EE*HH*MM*NN*QQ + AA*EE*II*MM*NN*QQ + CC*EE*HH*LL*MM*RR - BB*EE*HH*LL*NN*RR + BB*EE*II*LL*NN*RR - AA*CC*HH*MM*PP*RR + AA*BB*HH*NN*PP*RR - AA*BB*II*NN*PP*RR - CC*EE*GG*LL*MM*SS + BB*EE*GG*LL*NN*SS + BB*CC*FF*LL*PP*SS - AA*CC*FF*MM*PP*SS + AA*CC*GG*MM*PP*SS - AA*BB*GG*NN*PP*SS - BB*CC*EE*LL*QQ*SS + AA*CC*EE*MM*QQ*SS + CC*EE*GG*LL*MM*TT - BB*CC*FF*LL*PP*TT + AA*CC*FF*MM*PP*TT - AA*CC*GG*MM*PP*TT + BB*CC*EE*LL*QQ*TT - AA*CC*EE*MM*QQ*TT - BB*CC*EE*LL*RR*TT + AA*BB*CC*PP*RR*TT)*( - (( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*(( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( Power(AA,4)* Power(BB,2)*JJ*NN -  Power(AA,4)* Power(BB,2)*CC*UU)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( Power(AA,6)* Power(BB,3)*DD*JJ*NN -  Power(AA,6)* Power(BB,3)*CC*JJ*OO +  Power(AA,6)* Power(BB,3)*CC*KK*OO -  Power(AA,6)* Power(BB,3)*CC*DD*VV)))
-} \break
-
-$\dot{\Theta}_{bm}$ =
-{\tiny
-  (GG*II*LL*NN*PP*X1 - CC*HH*LL*PP*RR*X1 + AA*HH*NN*PP*RR*X1 - AA*II*NN*PP*RR*X1 + CC*GG*LL*PP*SS*X1 - AA*GG*NN*PP*SS*X1 - CC*GG*LL*PP*TT*X1 + AA*CC*PP*RR*TT*X1 - FF*II*LL*NN*PP*X2 + EE*II*LL*NN*QQ*X2 - EE*II*LL*NN*RR*X2 + AA*II*NN*PP*RR*X2 - CC*FF*LL*PP*SS*X2 + AA*FF*NN*PP*SS*X2 + CC*EE*LL*QQ*SS*X2 - AA*EE*NN*QQ*SS*X2 + CC*FF*LL*PP*TT*X2 - CC*EE*LL*QQ*TT*X2 + CC*EE*LL*RR*TT*X2 - AA*CC*PP*RR*TT*X2 + FF*HH*LL*NN*PP*X3 - EE*HH*LL*NN*QQ*X3 + EE*HH*LL*NN*RR*X3 - AA*HH*NN*PP*RR*X3 - EE*GG*LL*NN*SS*X3 - AA*FF*NN*PP*SS*X3 + AA*GG*NN*PP*SS*X3 + AA*EE*NN*QQ*SS*X3 - EE*GG*II*LL*NN*X5 + CC*EE*HH*LL*RR*X5 - AA*EE*HH*NN*RR*X5 + AA*EE*II*NN*RR*X5 - CC*EE*GG*LL*SS*X5 + AA*EE*GG*NN*SS*X5 + CC*EE*GG*LL*TT*X5 - AA*CC*EE*RR*TT*X5 + EE*GG*II*LL*NN*X6 + CC*FF*HH*LL*PP*X6 - AA*FF*HH*NN*PP*X6 + AA*FF*II*NN*PP*X6 - AA*GG*II*NN*PP*X6 - CC*EE*HH*LL*QQ*X6 + AA*EE*HH*NN*QQ*X6 - AA*EE*II*NN*QQ*X6 - CC*EE*GG*LL*TT*X6 - AA*CC*FF*PP*TT*X6 + AA*CC*GG*PP*TT*X6 + AA*CC*EE*QQ*TT*X6 - CC*FF*HH*LL*PP*X7 + CC*EE*HH*LL*QQ*X7 - CC*EE*HH*LL*RR*X7 + AA*CC*HH*PP*RR*X7 + CC*EE*GG*LL*SS*X7 + AA*CC*FF*PP*SS*X7 - AA*CC*GG*PP*SS*X7 - AA*CC*EE*QQ*SS*X7)/(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT) + (( - (FF*HH*LL*PP) + EE*HH*LL*QQ - EE*HH*LL*RR + AA*HH*PP*RR + EE*GG*LL*SS + AA*FF*PP*SS - AA*GG*PP*SS - AA*EE*QQ*SS)*(JJ*NN - CC*UU)*( - ((( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( - ((( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*NN*X2 +  Power(AA,4)* Power(BB,2)*NN*X3 -  Power(AA,4)* Power(BB,2)*CC*X7)))) + ( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( - ((( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,5)* Power(BB,3)*CC*DD*LL*X1 -  Power(AA,6)* Power(BB,2)*CC*DD*MM*X1 +  Power(AA,6)* Power(BB,2)*CC*DD*MM*X2 -  Power(AA,6)* Power(BB,3)*DD*NN*X2 +  Power(AA,6)* Power(BB,3)*DD*NN*X3 -  Power(AA,6)* Power(BB,3)*CC*OO*X3 +  Power(AA,6)* Power(BB,3)*CC*OO*X4 -  Power(AA,6)* Power(BB,3)*CC*DD*X8)))))/((EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT)*( - (( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*(( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( Power(AA,4)* Power(BB,2)*JJ*NN -  Power(AA,4)* Power(BB,2)*CC*UU)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( Power(AA,6)* Power(BB,3)*DD*JJ*NN -  Power(AA,6)* Power(BB,3)*CC*JJ*OO +  Power(AA,6)* Power(BB,3)*CC*KK*OO -  Power(AA,6)* Power(BB,3)*CC*DD*VV)))
-}\break
-
-$\dot{\Theta}_{um}$ =
-{\tiny
-  ( - X2 + X3)/CC + (GG*( - (BB*LL*PP*X1) + AA*MM*PP*X1 + EE*LL*MM*X2 - AA*MM*PP*X2 + BB*EE*LL*X5 - AA*EE*MM*X5 - BB*EE*LL*X6 + AA*BB*PP*X6))/(CC*(EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)) - ((( - HH + II)/CC - (GG*(EE*LL - AA*PP)*( - (HH*MM) + BB*SS))/(CC*(EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)))*( - (BB*GG*LL*NN*PP*X1) + AA*GG*MM*NN*PP*X1 + BB*CC*LL*PP*RR*X1 - AA*CC*MM*PP*RR*X1 + BB*FF*LL*NN*PP*X2 - AA*FF*MM*NN*PP*X2 - BB*EE*LL*NN*QQ*X2 + AA*EE*MM*NN*QQ*X2 - CC*EE*LL*MM*RR*X2 + BB*EE*LL*NN*RR*X2 + AA*CC*MM*PP*RR*X2 - AA*BB*NN*PP*RR*X2 + EE*GG*LL*MM*NN*X3 - BB*FF*LL*NN*PP*X3 + AA*FF*MM*NN*PP*X3 - AA*GG*MM*NN*PP*X3 + BB*EE*LL*NN*QQ*X3 - AA*EE*MM*NN*QQ*X3 - BB*EE*LL*NN*RR*X3 + AA*BB*NN*PP*RR*X3 + BB*EE*GG*LL*NN*X5 - AA*EE*GG*MM*NN*X5 - BB*CC*EE*LL*RR*X5 + AA*CC*EE*MM*RR*X5 + CC*EE*GG*LL*MM*X6 - BB*EE*GG*LL*NN*X6 - BB*CC*FF*LL*PP*X6 + AA*CC*FF*MM*PP*X6 - AA*CC*GG*MM*PP*X6 + AA*BB*GG*NN*PP*X6 + BB*CC*EE*LL*QQ*X6 - AA*CC*EE*MM*QQ*X6 - CC*EE*GG*LL*MM*X7 + BB*CC*FF*LL*PP*X7 - AA*CC*FF*MM*PP*X7 + AA*CC*GG*MM*PP*X7 - BB*CC*EE*LL*QQ*X7 + AA*CC*EE*MM*QQ*X7 + BB*CC*EE*LL*RR*X7 - AA*BB*CC*PP*RR*X7))/(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT) - ((JJ/CC + ((EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)*(( - HH + II)/CC - (GG*(EE*LL - AA*PP)*( - (HH*MM) + BB*SS))/(CC*(EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)))*( - (JJ*NN) + CC*UU))/(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT))*( - ((( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( - ((( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*NN*X2 +  Power(AA,4)* Power(BB,2)*NN*X3 -  Power(AA,4)* Power(BB,2)*CC*X7)))) + ( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( - ((( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,5)* Power(BB,3)*CC*DD*LL*X1 -  Power(AA,6)* Power(BB,2)*CC*DD*MM*X1 +  Power(AA,6)* Power(BB,2)*CC*DD*MM*X2 -  Power(AA,6)* Power(BB,3)*DD*NN*X2 +  Power(AA,6)* Power(BB,3)*DD*NN*X3 -  Power(AA,6)* Power(BB,3)*CC*OO*X3 +  Power(AA,6)* Power(BB,3)*CC*OO*X4 -  Power(AA,6)* Power(BB,3)*CC*DD*X8)))))/( - (( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*(( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( Power(AA,4)* Power(BB,2)*JJ*NN -  Power(AA,4)* Power(BB,2)*CC*UU)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( Power(AA,6)* Power(BB,3)*DD*JJ*NN -  Power(AA,6)* Power(BB,3)*CC*JJ*OO +  Power(AA,6)* Power(BB,3)*CC*KK*OO -  Power(AA,6)* Power(BB,3)*CC*DD*VV))
-}\break
-
-$\dot{\Theta}_{ub}$ =
-{\tiny
-  ( - X3 + X4)/DD + (II*( - (BB*GG*LL*NN*PP*X1) + AA*GG*MM*NN*PP*X1 + BB*CC*LL*PP*RR*X1 - AA*CC*MM*PP*RR*X1 + BB*FF*LL*NN*PP*X2 - AA*FF*MM*NN*PP*X2 - BB*EE*LL*NN*QQ*X2 + AA*EE*MM*NN*QQ*X2 - CC*EE*LL*MM*RR*X2 + BB*EE*LL*NN*RR*X2 + AA*CC*MM*PP*RR*X2 - AA*BB*NN*PP*RR*X2 + EE*GG*LL*MM*NN*X3 - BB*FF*LL*NN*PP*X3 + AA*FF*MM*NN*PP*X3 - AA*GG*MM*NN*PP*X3 + BB*EE*LL*NN*QQ*X3 - AA*EE*MM*NN*QQ*X3 - BB*EE*LL*NN*RR*X3 + AA*BB*NN*PP*RR*X3 + BB*EE*GG*LL*NN*X5 - AA*EE*GG*MM*NN*X5 - BB*CC*EE*LL*RR*X5 + AA*CC*EE*MM*RR*X5 + CC*EE*GG*LL*MM*X6 - BB*EE*GG*LL*NN*X6 - BB*CC*FF*LL*PP*X6 + AA*CC*FF*MM*PP*X6 - AA*CC*GG*MM*PP*X6 + AA*BB*GG*NN*PP*X6 + BB*CC*EE*LL*QQ*X6 - AA*CC*EE*MM*QQ*X6 - CC*EE*GG*LL*MM*X7 + BB*CC*FF*LL*PP*X7 - AA*CC*FF*MM*PP*X7 + AA*CC*GG*MM*PP*X7 - BB*CC*EE*LL*QQ*X7 + AA*CC*EE*MM*QQ*X7 + BB*CC*EE*LL*RR*X7 - AA*BB*CC*PP*RR*X7))/(DD*(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT)) - ((( - JJ + KK)/DD - (II*(EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)*( - (JJ*NN) + CC*UU))/(DD*(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT)))*( - ((( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( - ((( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*NN*X2 +  Power(AA,4)* Power(BB,2)*NN*X3 -  Power(AA,4)* Power(BB,2)*CC*X7)))) + ( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( - ((( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,5)* Power(BB,3)*CC*DD*LL*X1 -  Power(AA,6)* Power(BB,2)*CC*DD*MM*X1 +  Power(AA,6)* Power(BB,2)*CC*DD*MM*X2 -  Power(AA,6)* Power(BB,3)*DD*NN*X2 +  Power(AA,6)* Power(BB,3)*DD*NN*X3 -  Power(AA,6)* Power(BB,3)*CC*OO*X3 +  Power(AA,6)* Power(BB,3)*CC*OO*X4 -  Power(AA,6)* Power(BB,3)*CC*DD*X8)))))/( - (( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*(( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( Power(AA,4)* Power(BB,2)*JJ*NN -  Power(AA,4)* Power(BB,2)*CC*UU)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( Power(AA,6)* Power(BB,3)*DD*JJ*NN -  Power(AA,6)* Power(BB,3)*CC*JJ*OO +  Power(AA,6)* Power(BB,3)*CC*KK*OO -  Power(AA,6)* Power(BB,3)*CC*DD*VV))
-}
-
-\subsubsection{Bothbound equations}
-\label{bothbound-motion-equations}
-The following are the full motion equations used to time evolve the bothbound model:
-
-\begin{center}
-  \begin{minipage}{0.3\textwidth}
-    \begin{align*}
-      a &= \frac{-dXnm}{dLn}\\
-       b &= \frac{-dXnm}{dLf}\\
-       c &= -(x_{nm} - x_{nb}) / \gamma_m\\
-       d &= (x_t - x_{nm}) / \gamma_m\\
-       e &= \frac{-dXt}{dLn}\\
-       f &= \frac{-dXt}{dLf}\\
-       g &= -(x_t - x_{nm}) / \gamma_t\\
-       h &= (x_{fm} - x_t) / \gamma_t\\
-       i &= \frac{-dXfm}{dLn}\\
-       j &= \frac{-dXfm}{dLf}\\
-       k &= -(x_{fm} - x_t) / \gamma_m\\
-       l &= (x_{fb} - x_{fm}) / \gamma_m\\
-    \end{align*}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-    \begin{align*}
-       m &= \frac{-dYnm}{dLn}\\
-       n &= \frac{-dYnm}{dLf}\\
-       p &= -(y_{nm} - y_{nb}) / \gamma_m\\
-       q &= (y_t - y_{nm}) / \gamma_m\\
-       r &= \frac{-dYt}{dLn}\\
-       s &= \frac{-dYt}{dLf}\\
-       t &= -(y_t - y_{nm}) / \gamma_t\\
-       u &= (y_{fm} - y_t) / \gamma_t\\
-       v &= \frac{-dYfm}{dLn}\\
-       w &= \frac{-dYfm}{dLf}\\
-       x &= -(y_{fm} - y_t) / \gamma_m\\
-       y &= (y_{fb} - y_{fm}) / \gamma_m
-    \end{align*}
-  \end{minipage}
-\end{center}
-
-$\dot{Ln}$ =
-{\tiny
-    (h*l*p*t*w*x1 - g*l*p*u*w*x1 + g*l*p*s*x*x1 - f*l*p*t*x*x1 +
-     d*l*p*u*w*x2 - c*l*q*u*w*x2 - d*l*p*s*x*x2 + c*l*q*s*x*x2 -
-     c*l*n*t*x*x2 + b*l*p*t*x*x2 - c*h*l*t*w*x4 + c*g*l*u*w*x4 -
-     c*g*l*s*x*x4 + c*f*l*t*x*x4 - d*h*l*p*w*x5 + c*h*l*q*w*x5 +
-     c*g*l*n*x*x5 + d*f*l*p*x*x5 - b*g*l*p*x*x5 - c*f*l*q*x*x5 +
-     d*h*l*p*s*x6 - c*h*l*q*s*x6 + c*h*l*n*t*x6 - b*h*l*p*t*x6 -
-     c*g*l*n*u*x6 - d*f*l*p*u*x6 + b*g*l*p*u*x6 + c*f*l*q*u*x6 -
-     g*k*p*s*x1*y - h*j*p*t*x1*y + f*k*p*t*x1*y + g*j*p*u*x1*y +
-     d*k*p*s*x2*y - c*k*q*s*x2*y + c*k*n*t*x2*y - b*k*p*t*x2*y -
-     d*j*p*u*x2*y + c*j*q*u*x2*y - d*h*p*s*x3*y + c*h*q*s*x3*y -
-     c*h*n*t*x3*y + b*h*p*t*x3*y + c*g*n*u*x3*y + d*f*p*u*x3*y -
-     b*g*p*u*x3*y - c*f*q*u*x3*y + c*g*k*s*x4*y + c*h*j*t*x4*y -
-     c*f*k*t*x4*y - c*g*j*u*x4*y - c*g*k*n*x5*y + d*h*j*p*x5*y -
-     d*f*k*p*x5*y + b*g*k*p*x5*y - c*h*j*q*x5*y + c*f*k*q*x5*y)/
-     (d*h*l*p*s*v - c*h*l*q*s*v + c*h*l*n*t*v - b*h*l*p*t*v -
-     c*g*l*n*u*v - d*f*l*p*u*v + b*g*l*p*u*v + c*f*l*q*u*v -
-     d*h*l*p*r*w + c*h*l*q*r*w - c*h*l*m*t*w + a*h*l*p*t*w +
-     c*g*l*m*u*w + d*e*l*p*u*w - a*g*l*p*u*w - c*e*l*q*u*w +
-     c*g*l*n*r*x + d*f*l*p*r*x - b*g*l*p*r*x - c*f*l*q*r*x -
-     c*g*l*m*s*x - d*e*l*p*s*x + a*g*l*p*s*x + c*e*l*q*s*x +
-     c*f*l*m*t*x - c*e*l*n*t*x + b*e*l*p*t*x - a*f*l*p*t*x -
-     c*g*k*n*r*y + d*h*j*p*r*y - d*f*k*p*r*y + b*g*k*p*r*y -
-     c*h*j*q*r*y + c*f*k*q*r*y + c*g*k*m*s*y - d*h*i*p*s*y +
-     d*e*k*p*s*y - a*g*k*p*s*y + c*h*i*q*s*y - c*e*k*q*s*y +
-     c*h*j*m*t*y - c*f*k*m*t*y - c*h*i*n*t*y + c*e*k*n*t*y +
-     b*h*i*p*t*y - a*h*j*p*t*y - b*e*k*p*t*y + a*f*k*p*t*y -
-     c*g*j*m*u*y + c*g*i*n*u*y + d*f*i*p*u*y - b*g*i*p*u*y -
-     d*e*j*p*u*y + a*g*j*p*u*y - c*f*i*q*u*y + c*e*j*q*u*y)
-}\break
-
-$\dot{Lf}$ =
-{\tiny
-    (-(h*l*p*t*v*x1) + g*l*p*u*v*x1 - g*l*p*r*x*x1 + e*l*p*t*x*x1 -
-     d*l*p*u*v*x2 + c*l*q*u*v*x2 + d*l*p*r*x*x2 - c*l*q*r*x*x2 +
-     c*l*m*t*x*x2 - a*l*p*t*x*x2 + c*h*l*t*v*x4 - c*g*l*u*v*x4 +
-     c*g*l*r*x*x4 - c*e*l*t*x*x4 + d*h*l*p*v*x5 - c*h*l*q*v*x5 -
-     c*g*l*m*x*x5 - d*e*l*p*x*x5 + a*g*l*p*x*x5 + c*e*l*q*x*x5 -
-     d*h*l*p*r*x6 + c*h*l*q*r*x6 - c*h*l*m*t*x6 + a*h*l*p*t*x6 +
-     c*g*l*m*u*x6 + d*e*l*p*u*x6 - a*g*l*p*u*x6 - c*e*l*q*u*x6 +
-     g*k*p*r*x1*y + h*i*p*t*x1*y - e*k*p*t*x1*y - g*i*p*u*x1*y -
-     d*k*p*r*x2*y + c*k*q*r*x2*y - c*k*m*t*x2*y + a*k*p*t*x2*y +
-     d*i*p*u*x2*y - c*i*q*u*x2*y + d*h*p*r*x3*y - c*h*q*r*x3*y +
-     c*h*m*t*x3*y - a*h*p*t*x3*y - c*g*m*u*x3*y - d*e*p*u*x3*y +
-     a*g*p*u*x3*y + c*e*q*u*x3*y - c*g*k*r*x4*y - c*h*i*t*x4*y +
-     c*e*k*t*x4*y + c*g*i*u*x4*y + c*g*k*m*x5*y - d*h*i*p*x5*y +
-     d*e*k*p*x5*y - a*g*k*p*x5*y + c*h*i*q*x5*y - c*e*k*q*x5*y)/
-     (d*h*l*p*s*v - c*h*l*q*s*v + c*h*l*n*t*v - b*h*l*p*t*v -
-     c*g*l*n*u*v - d*f*l*p*u*v + b*g*l*p*u*v + c*f*l*q*u*v -
-     d*h*l*p*r*w + c*h*l*q*r*w - c*h*l*m*t*w + a*h*l*p*t*w +
-     c*g*l*m*u*w + d*e*l*p*u*w - a*g*l*p*u*w - c*e*l*q*u*w +
-     c*g*l*n*r*x + d*f*l*p*r*x - b*g*l*p*r*x - c*f*l*q*r*x -
-     c*g*l*m*s*x - d*e*l*p*s*x + a*g*l*p*s*x + c*e*l*q*s*x +
-     c*f*l*m*t*x - c*e*l*n*t*x + b*e*l*p*t*x - a*f*l*p*t*x -
-     c*g*k*n*r*y + d*h*j*p*r*y - d*f*k*p*r*y + b*g*k*p*r*y -
-     c*h*j*q*r*y + c*f*k*q*r*y + c*g*k*m*s*y - d*h*i*p*s*y +
-     d*e*k*p*s*y - a*g*k*p*s*y + c*h*i*q*s*y - c*e*k*q*s*y +
-     c*h*j*m*t*y - c*f*k*m*t*y - c*h*i*n*t*y + c*e*k*n*t*y +
-     b*h*i*p*t*y - a*h*j*p*t*y - b*e*k*p*t*y + a*f*k*p*t*y -
-     c*g*j*m*u*y + c*g*i*n*u*y + d*f*i*p*u*y - b*g*i*p*u*y -
-     d*e*j*p*u*y + a*g*j*p*u*y - c*f*i*q*u*y + c*e*j*q*u*y)
-}
-
-\section{Questions for Tate/Nicole}
-\begin{itemize}
-\item How responsible am I for the papers I cite? Should I have read them, know their methodologies?
-\item Is it appropriate to cite web pages in a thesis? In a separate citation section?
-\item When is a fact ``common knowledge'' enough that it doesn't need a citation?
-\item Some papers have 'Request permission to reuse' links for figures - do I need to do this for copying figures?
-\end{itemize}
 \end{document}

### 2017-01-16 thesis/latex/thesis.tex
Add listoffigures and Tate assignment.

--- 2017-01-11 thesis/latex/thesis.tex
@@
 \tableofcontents
+\newpage
+
+\listoffigures
 \newpage
@@
   \centering
-  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/mechanochemical-cycle.jpeg}
+  \includegraphics[width=.85\textwidth,keepaspectratio]{../figures/mechanochemical-cycle.jpeg}
   \caption{Cycle of states the dynein motor goes through during motion in the Cianfrocco model \cite{cianfroccoreview}.}
@@
 Energy is calculated by looking at displacements of angles from equilibrium, whereas forces are calculated using displacement of angles from equilibrium, but also orientation and position of each domain.
-Verifying these two systems produce the same result indicates...
-something.
+Having the force system and energy system calculate the same results indicates that they are both likely correct..
 \textbf{make this more precise.}
@@
 \subsubsection{Equipartition theorem}
+The equipartition theorem (ET) is a thermodynamic statement which says that all systems with solely quadratic energy dependence on $N$ degrees of freedom will have $<E> = \frac N2 k_bT$.
+In order for the dynein models to display ET behavior, they would have to both properly calculate forces and properly calculate velocities from forces.
+Brownian forces are also required to properly explore the different states of the system, but they need not be the correct magnitude for ET agreement.\\
+
+To test if the models obey the ET, the total potential energy of each model was logged over \textbf{ten million} (???) timesteps of simulation.
+The results are reported in Figure \ref{equipartition_agreement}, where the PE scale is in units of $\frac12 k_bT$.
+Energies are reported as a running average over the simulation from 0 to time $t$.
+As shown, both models have their average potential energy converge to the expected amount, where the onebound model has four degrees of freedom and the bothbound two.\\
+
+This test verifies both force and velocity are calculated properly for either model.
+If either force or velocity was wrong, then the model would not obey the Boltzmann hypothesis $P(s) = e^{-\Delta G_{s}/K_BT}$, where $s$ is a state of the system.
+The transition rate from $s_1$ to $s_2$ is $P(s_1\rightarrow s_2) = e^{-{\Delta G_2 - \Delta G_1}/K_BT}$.
+Wrong velocities are those that do not transition the system from $s_1$ to $s_2$ at this rate.\\
+
+\textbf{would it be a stronger test to also calculate the standard deviation of <PE>?
+It seems like <PE> would still equal N/2KbT as long as the forces/velocities were symmetrically wrong.
+Calculating the stdev would require the state occupation to actually the ET.
+Or, maybe just plot the angle occupation and show that it is what it should be?}
+
+\begin{figure}[h!]
+  \centering
+  \begin{minipage}[b]{0.49\textwidth}
+    \includegraphics[width=\textwidth]{../figures/OB_Average_PE.pdf}
+    \caption{Onebound PE vs time.}
+  \end{minipage}
+  \begin{minipage}[b]{0.49\textwidth}
+    \includegraphics[width=\textwidth]{../figures/BB_Average_PE.pdf}
+    \caption{Bothbound PE vs time.}
+  \end{minipage}
+  \label{equipartition_agreement}
+\end{figure}
+
+\textbf{Possibly due mutations on motion equations to show that tests are no longer behaved}
+\subsubsection{Finding proper $dt$}
 
@@
 \end{itemize}
+
+\section{Questions for David}
+\begin{itemize}
+\item Would it be possible to, for the ET test, also verify the standard deviation of energy is what it should be?
+Since just verifying the energy averages to N/2KbT doesn't verify the forces are correct, but the stdev would?
+\end{itemize}
+
+\section{Figures to make}
+\begin{itemize}
+\item Crystal structure of dynein with rigid rods imposed and or a cartoon model
+\item Brownian motion figure
+\item Crystal structure transitioning from pre/post stroke
+\item dt vs error plot
+\item ET angle histogram
+\item figure of finding lengths in PyMol
+\item stepping length/time histograms
+\end{itemize}
 \end{document}

### 2017-01-17 thesis/latex/thesis.tex
Add finishing touches on tests section of Methods.

@@
 
-\usepackage[utf8]{inputenc}
-\usepackage{graphicx}
-\usepackage{array}
-
 \setlength{\emergencystretch}{3cm} % somehow helps prevent hboxes from overflowing
@@
 \tableofcontents
-\newpage
-
-\listoffigures
 \newpage
@@
   \centering
-  \includegraphics[width=.85\textwidth,keepaspectratio]{../figures/mechanochemical-cycle.jpeg}
+  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/mechanochemical-cycle.jpeg}
   \caption{Cycle of states the dynein motor goes through during motion in the Cianfrocco model \cite{cianfroccoreview}.}
@@
 If either force or velocity was wrong, then the model would not obey the Boltzmann hypothesis $P(s) = e^{-\Delta G_{s}/K_BT}$, where $s$ is a state of the system.
-The transition rate from $s_1$ to $s_2$ is $P(s_1\rightarrow s_2) = e^{-{\Delta G_2 - \Delta G_1}/K_BT}$.
+The transition rate from $s_1$ to $s_2$ is $P(s_1\rightarrow_s_2) = e^{-{\Delta G_2 - \Delta G_1}/K_BT}$.
 Wrong velocities are those that do not transition the system from $s_1$ to $s_2$ at this rate.\\
@@
 
-\begin{figure}[h!]
-  \centering
-  \begin{minipage}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{../figures/OB_Average_PE.pdf}
-    \caption{Onebound PE vs time.}
-  \end{minipage}
-  \begin{minipage}[b]{0.49\textwidth}
-    \includegraphics[width=\textwidth]{../figures/BB_Average_PE.pdf}
-    \caption{Bothbound PE vs time.}
-  \end{minipage}
-  \label{equipartition_agreement}
-\end{figure}
+%% \begin{figure}[h!]
+%%   \centering
+%%   \begin{minipage}[b]{0.49\textwidth}
+%%     \includegraphics[width=\textwidth]{../figures/OB_Average_PE.pdf}
+%%     \caption{Onebound PE vs time.}
+%%   \end{minipage}
+%%   \begin{minipage}[b]{0.49\textwidth}
+%%     \includegraphics[width=\textwidth]{../figures/BB_Average_PE.pdf}
+%%     \caption{Bothbound PE vs time.}
+%%   \end{minipage}
+%%   \label{equipartition_agreement}
+%% \end{figure}
 
@@
 \end{itemize}
-
-\section{Figures to make}
-\begin{itemize}
-\item Crystal structure of dynein with rigid rods imposed and or a cartoon model
-\item Brownian motion figure
-\item Crystal structure transitioning from pre/post stroke
-\item dt vs error plot
-\item ET angle histogram
-\item figure of finding lengths in PyMol
-\item stepping length/time histograms
-\end{itemize}
 \end{document}

### 2017-01-20 thesis/latex/thesis.tex
Tweak one of the test descriptions.
Add work on rate constants.

@@
 \subsubsection{Transitioning states}
+Onebound and bothbound models transition between each other based on rate constants $k_b$, the binding constant, and $k_{ub}$, the unbinding constant, respectively.
+These values are calculated using the Arrhenius equation, and are given as:
+
+\begin{equation}
+  k_b = Ae^{\frac{-\Delta G}{k_BT}} = Ae^{\frac{-\Delta G_{conf}}{k_BT}}e^{\frac{-\Delta G_{bind}}{k_BT}}
+\end{equation}
+
+where $\Delta G_{conf}$ is the change in spring energy in going from bothbound to onebound, and $\Delta G_{bind}$ is the energy gained from binding to the microtubule, and A is an unknown constant known as the preexponential factor with units of $s^{-1}$.
+
 Transitions from one state to another are accomplished by calculating the rate of transition in a certain position, then using this rate divided by \verb|dt| as a probability of transfer at each timestep.
@@
 If either force or velocity was wrong, then the model would not obey the Boltzmann hypothesis $P(s) = e^{-\Delta G_{s}/K_BT}$, where $s$ is a state of the system.
-The transition rate from $s_1$ to $s_2$ is $P(s_1\rightarrow_s_2) = e^{-{\Delta G_2 - \Delta G_1}/K_BT}$.
+The transition rate from $s_1$ to $s_2$ is $P(s_1\rightarrow s_2) = e^{-{\Delta G_2 - \Delta G_1}/K_BT}$.
 Wrong velocities are those that do not transition the system from $s_1$ to $s_2$ at this rate.\\
@@
 
-\textbf{Possibly due mutations on motion equations to show that tests are no longer behaved}
+\textbf{Possibly do mutations on motion equations to show that tests are no longer behaved}
 \subsubsection{Finding proper $dt$}
@@
 
-\subsection{Data collection}
-
-\subsubsection{Outputting positions, step locations}
-\subsubsection{Reconstructing steps from step locations}
-\subsubsection{Job management}
-
 \subsection{Parameter fitting}
+There are 14 free variables in the dynein model: rigid rod lengths $L_s$ and $L_t$, domain radii $R_b$, $R_m$ and $R_t$, spring constants $c_b$, $c_m$ and $c_t$, equilibrium angles $\theta_{b,eq}$, $\theta_{t,eq}$, $\Theta_{m,eq}^{Pre}$, $\Theta_{m,eq}^{Post}$, onebound binding rate $k_b$, and bothbound unbinding rate $k_{ub}$.
+Some of these values are found easily from literature or by looking at crystal structures of the motor protein, as shown in Table \ref{parameter-table}.
+Others are more difficult to find.\\
+
+What's more, using exact literature values for each parameter does not necessarily create the simulation most similar to \textit{in vivo} dynein.
+Defects and simplifications in the model may cause slightly less physical values for individual parameters to lead to a more physical simulation.
+For this reason, fitting model parameters to experiment is necessary.
+This process is chronicled below.
+
+\begin{figure}[h]
+  \centering
+  \begin{tabular}{| l | l | l | l | l | l | l | l | l |}
+    \hline
+    & Winch (Sarlah) & Schmidt & Lin & PyMol 3VKH & PyMol 4RH7 & Redwine & Kon & Burgess \\\hline
+    $L_s$ & 12nm & & & 21.0nm & 22.1nm & & & \\ \hline
+    $L_t$ &  7nm & & & & 11.15nm & & & \\ \hline
+    $R_b$ &  & & & 1.57nm & 1.45nm & & & \\ \hline
+    $R_m$ &  7nm & & & 7.36nm & 6.3nm & & & \\ \hline
+    $R_t$ &  & & & & 2.16nm & & & \\ \hline
+    $\theta_{m}^{Pre}$ & 250$^{\circ}$ & & 171$^{\circ}$ & & & & & 160\\ \hline
+    $\theta_{m}^{Post}$ & 330$^{\circ}$ & & 137.5$^{\circ}$ & & & & & 136\\ \hline
+    $\theta_{b}$ & 56$^{\circ}$ & & 63.5$^{\circ}$ & & & & & \\ \hline
+    $\theta_{t}$ & 0$^{\circ}$ & & & & & & & \\ \hline
+    $k_{ub}$ & $180 s^{-1,a}$ & & & & & & $90.2 \pm 4.5$& \\ \hline
+    $k_b$ & $5000 s^{-1,b}$ & & & & & & & \\ \hline
+    $\Delta G_{bind, bb}$ & & & & & & -71 kJ/mol & & \\ \hline
+    $\Delta G_{bind, ob}$ & & & & & & -37 kJ/mol & & \\ \hline
+    $c_t$ & & & & & & & & \\ \hline
+    $c_m$ & & & & & & & & \\ \hline
+    $c_b$ & & & & & & 140 kJ/mol & & \\ \hline
+  \end{tabular}
+  \caption{Table of literature values for various free parameters in dynein model.}
+  \label{parameter-table}
+\end{figure}
+
+\subsubsection{Fitting rate constants}
+Rate constants govern the probability of transitioning between onebound and bothbound states. Rate constant
+
+\begin{equation}
+
+\end{equation}
+
 \subsubsection{Fitting spring constants}
@@
 \subsubsection{Finding the right timestep}
+
+\subsection{Data collection}
+
+\subsubsection{Outputting positions, step locations}
+\subsubsection{Reconstructing steps from step locations}
+\subsubsection{Job management}
 
@@
 \label{onebound-motion-equations}
-The following are the full motion equations used to time evolve the onebound model:
+The following is a continuation of the derivation for the full motion equations used to time evolve the onebound model:
 
@@
 
-Solving the above system for $\dot{\Theta}_{bb}, \dot{\Theta}_{bm}, \dot{\Theta}_{um}$ and $\dot{\Theta}_{ub}$ gives
-
-$\dot{\Theta}_{bb}$ =
-{\tiny
-( - (GG*II*MM*NN*PP*X1) + CC*HH*MM*PP*RR*X1 - BB*HH*NN*PP*RR*X1 + BB*II*NN*PP*RR*X1 - CC*GG*MM*PP*SS*X1 + BB*GG*NN*PP*SS*X1 + CC*GG*MM*PP*TT*X1 - BB*CC*PP*RR*TT*X1 + FF*II*MM*NN*PP*X2 - EE*II*MM*NN*QQ*X2 + CC*FF*MM*PP*SS*X2 - BB*FF*NN*PP*SS*X2 - CC*EE*MM*QQ*SS*X2 + BB*EE*NN*QQ*SS*X2 - CC*FF*MM*PP*TT*X2 + CC*EE*MM*QQ*TT*X2 - FF*HH*MM*NN*PP*X3 + EE*HH*MM*NN*QQ*X3 + BB*FF*NN*PP*SS*X3 - BB*EE*NN*QQ*SS*X3 + EE*GG*II*MM*NN*X5 - CC*EE*HH*MM*RR*X5 + BB*EE*HH*NN*RR*X5 - BB*EE*II*NN*RR*X5 + CC*EE*GG*MM*SS*X5 - BB*EE*GG*NN*SS*X5 - CC*EE*GG*MM*TT*X5 + BB*CC*EE*RR*TT*X5 - CC*FF*HH*MM*PP*X6 + BB*FF*HH*NN*PP*X6 - BB*FF*II*NN*PP*X6 + CC*EE*HH*MM*QQ*X6 - BB*EE*HH*NN*QQ*X6 + BB*EE*II*NN*QQ*X6 + BB*CC*FF*PP*TT*X6 - BB*CC*EE*QQ*TT*X6 + CC*FF*HH*MM*PP*X7 - CC*EE*HH*MM*QQ*X7 - BB*CC*FF*PP*SS*X7 + BB*CC*EE*QQ*SS*X7)/(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT) - ((FF*PP - EE*QQ)*(HH*MM - BB*SS)*(JJ*NN - CC*UU)*( - ((( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( - ((( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*NN*X2 +  Power(AA,4)* Power(BB,2)*NN*X3 -  Power(AA,4)* Power(BB,2)*CC*X7)))) + ( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( - ((( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,5)* Power(BB,3)*CC*DD*LL*X1 -  Power(AA,6)* Power(BB,2)*CC*DD*MM*X1 +  Power(AA,6)* Power(BB,2)*CC*DD*MM*X2 -  Power(AA,6)* Power(BB,3)*DD*NN*X2 +  Power(AA,6)* Power(BB,3)*DD*NN*X3 -  Power(AA,6)* Power(BB,3)*CC*OO*X3 +  Power(AA,6)* Power(BB,3)*CC*OO*X4 -  Power(AA,6)* Power(BB,3)*CC*DD*X8)))))/(( - (EE*GG*II*LL*MM*NN) - BB*FF*HH*LL*NN*PP + BB*FF*II*LL*NN*PP + AA*FF*HH*MM*NN*PP - AA*FF*II*MM*NN*PP + AA*GG*II*MM*NN*PP + BB*EE*HH*LL*NN*QQ - BB*EE*II*LL*NN*QQ - AA*EE*HH*MM*NN*QQ + AA*EE*II*MM*NN*QQ + CC*EE*HH*LL*MM*RR - BB*EE*HH*LL*NN*RR + BB*EE*II*LL*NN*RR - AA*CC*HH*MM*PP*RR + AA*BB*HH*NN*PP*RR - AA*BB*II*NN*PP*RR - CC*EE*GG*LL*MM*SS + BB*EE*GG*LL*NN*SS + BB*CC*FF*LL*PP*SS - AA*CC*FF*MM*PP*SS + AA*CC*GG*MM*PP*SS - AA*BB*GG*NN*PP*SS - BB*CC*EE*LL*QQ*SS + AA*CC*EE*MM*QQ*SS + CC*EE*GG*LL*MM*TT - BB*CC*FF*LL*PP*TT + AA*CC*FF*MM*PP*TT - AA*CC*GG*MM*PP*TT + BB*CC*EE*LL*QQ*TT - AA*CC*EE*MM*QQ*TT - BB*CC*EE*LL*RR*TT + AA*BB*CC*PP*RR*TT)*( - (( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*(( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( Power(AA,4)* Power(BB,2)*JJ*NN -  Power(AA,4)* Power(BB,2)*CC*UU)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( Power(AA,6)* Power(BB,3)*DD*JJ*NN -  Power(AA,6)* Power(BB,3)*CC*JJ*OO +  Power(AA,6)* Power(BB,3)*CC*KK*OO -  Power(AA,6)* Power(BB,3)*CC*DD*VV)))
-} \break
-
-$\dot{\Theta}_{bm}$ =
-{\tiny
-  (GG*II*LL*NN*PP*X1 - CC*HH*LL*PP*RR*X1 + AA*HH*NN*PP*RR*X1 - AA*II*NN*PP*RR*X1 + CC*GG*LL*PP*SS*X1 - AA*GG*NN*PP*SS*X1 - CC*GG*LL*PP*TT*X1 + AA*CC*PP*RR*TT*X1 - FF*II*LL*NN*PP*X2 + EE*II*LL*NN*QQ*X2 - EE*II*LL*NN*RR*X2 + AA*II*NN*PP*RR*X2 - CC*FF*LL*PP*SS*X2 + AA*FF*NN*PP*SS*X2 + CC*EE*LL*QQ*SS*X2 - AA*EE*NN*QQ*SS*X2 + CC*FF*LL*PP*TT*X2 - CC*EE*LL*QQ*TT*X2 + CC*EE*LL*RR*TT*X2 - AA*CC*PP*RR*TT*X2 + FF*HH*LL*NN*PP*X3 - EE*HH*LL*NN*QQ*X3 + EE*HH*LL*NN*RR*X3 - AA*HH*NN*PP*RR*X3 - EE*GG*LL*NN*SS*X3 - AA*FF*NN*PP*SS*X3 + AA*GG*NN*PP*SS*X3 + AA*EE*NN*QQ*SS*X3 - EE*GG*II*LL*NN*X5 + CC*EE*HH*LL*RR*X5 - AA*EE*HH*NN*RR*X5 + AA*EE*II*NN*RR*X5 - CC*EE*GG*LL*SS*X5 + AA*EE*GG*NN*SS*X5 + CC*EE*GG*LL*TT*X5 - AA*CC*EE*RR*TT*X5 + EE*GG*II*LL*NN*X6 + CC*FF*HH*LL*PP*X6 - AA*FF*HH*NN*PP*X6 + AA*FF*II*NN*PP*X6 - AA*GG*II*NN*PP*X6 - CC*EE*HH*LL*QQ*X6 + AA*EE*HH*NN*QQ*X6 - AA*EE*II*NN*QQ*X6 - CC*EE*GG*LL*TT*X6 - AA*CC*FF*PP*TT*X6 + AA*CC*GG*PP*TT*X6 + AA*CC*EE*QQ*TT*X6 - CC*FF*HH*LL*PP*X7 + CC*EE*HH*LL*QQ*X7 - CC*EE*HH*LL*RR*X7 + AA*CC*HH*PP*RR*X7 + CC*EE*GG*LL*SS*X7 + AA*CC*FF*PP*SS*X7 - AA*CC*GG*PP*SS*X7 - AA*CC*EE*QQ*SS*X7)/(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT) + (( - (FF*HH*LL*PP) + EE*HH*LL*QQ - EE*HH*LL*RR + AA*HH*PP*RR + EE*GG*LL*SS + AA*FF*PP*SS - AA*GG*PP*SS - AA*EE*QQ*SS)*(JJ*NN - CC*UU)*( - ((( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( - ((( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*NN*X2 +  Power(AA,4)* Power(BB,2)*NN*X3 -  Power(AA,4)* Power(BB,2)*CC*X7)))) + ( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( - ((( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,5)* Power(BB,3)*CC*DD*LL*X1 -  Power(AA,6)* Power(BB,2)*CC*DD*MM*X1 +  Power(AA,6)* Power(BB,2)*CC*DD*MM*X2 -  Power(AA,6)* Power(BB,3)*DD*NN*X2 +  Power(AA,6)* Power(BB,3)*DD*NN*X3 -  Power(AA,6)* Power(BB,3)*CC*OO*X3 +  Power(AA,6)* Power(BB,3)*CC*OO*X4 -  Power(AA,6)* Power(BB,3)*CC*DD*X8)))))/((EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT)*( - (( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*(( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( Power(AA,4)* Power(BB,2)*JJ*NN -  Power(AA,4)* Power(BB,2)*CC*UU)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( Power(AA,6)* Power(BB,3)*DD*JJ*NN -  Power(AA,6)* Power(BB,3)*CC*JJ*OO +  Power(AA,6)* Power(BB,3)*CC*KK*OO -  Power(AA,6)* Power(BB,3)*CC*DD*VV)))
-}\break
-
-$\dot{\Theta}_{um}$ =
-{\tiny
-  ( - X2 + X3)/CC + (GG*( - (BB*LL*PP*X1) + AA*MM*PP*X1 + EE*LL*MM*X2 - AA*MM*PP*X2 + BB*EE*LL*X5 - AA*EE*MM*X5 - BB*EE*LL*X6 + AA*BB*PP*X6))/(CC*(EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)) - ((( - HH + II)/CC - (GG*(EE*LL - AA*PP)*( - (HH*MM) + BB*SS))/(CC*(EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)))*( - (BB*GG*LL*NN*PP*X1) + AA*GG*MM*NN*PP*X1 + BB*CC*LL*PP*RR*X1 - AA*CC*MM*PP*RR*X1 + BB*FF*LL*NN*PP*X2 - AA*FF*MM*NN*PP*X2 - BB*EE*LL*NN*QQ*X2 + AA*EE*MM*NN*QQ*X2 - CC*EE*LL*MM*RR*X2 + BB*EE*LL*NN*RR*X2 + AA*CC*MM*PP*RR*X2 - AA*BB*NN*PP*RR*X2 + EE*GG*LL*MM*NN*X3 - BB*FF*LL*NN*PP*X3 + AA*FF*MM*NN*PP*X3 - AA*GG*MM*NN*PP*X3 + BB*EE*LL*NN*QQ*X3 - AA*EE*MM*NN*QQ*X3 - BB*EE*LL*NN*RR*X3 + AA*BB*NN*PP*RR*X3 + BB*EE*GG*LL*NN*X5 - AA*EE*GG*MM*NN*X5 - BB*CC*EE*LL*RR*X5 + AA*CC*EE*MM*RR*X5 + CC*EE*GG*LL*MM*X6 - BB*EE*GG*LL*NN*X6 - BB*CC*FF*LL*PP*X6 + AA*CC*FF*MM*PP*X6 - AA*CC*GG*MM*PP*X6 + AA*BB*GG*NN*PP*X6 + BB*CC*EE*LL*QQ*X6 - AA*CC*EE*MM*QQ*X6 - CC*EE*GG*LL*MM*X7 + BB*CC*FF*LL*PP*X7 - AA*CC*FF*MM*PP*X7 + AA*CC*GG*MM*PP*X7 - BB*CC*EE*LL*QQ*X7 + AA*CC*EE*MM*QQ*X7 + BB*CC*EE*LL*RR*X7 - AA*BB*CC*PP*RR*X7))/(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT) - ((JJ/CC + ((EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)*(( - HH + II)/CC - (GG*(EE*LL - AA*PP)*( - (HH*MM) + BB*SS))/(CC*(EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)))*( - (JJ*NN) + CC*UU))/(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT))*( - ((( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( - ((( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*NN*X2 +  Power(AA,4)* Power(BB,2)*NN*X3 -  Power(AA,4)* Power(BB,2)*CC*X7)))) + ( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( - ((( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,5)* Power(BB,3)*CC*DD*LL*X1 -  Power(AA,6)* Power(BB,2)*CC*DD*MM*X1 +  Power(AA,6)* Power(BB,2)*CC*DD*MM*X2 -  Power(AA,6)* Power(BB,3)*DD*NN*X2 +  Power(AA,6)* Power(BB,3)*DD*NN*X3 -  Power(AA,6)* Power(BB,3)*CC*OO*X3 +  Power(AA,6)* Power(BB,3)*CC*OO*X4 -  Power(AA,6)* Power(BB,3)*CC*DD*X8)))))/( - (( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*(( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( Power(AA,4)* Power(BB,2)*JJ*NN -  Power(AA,4)* Power(BB,2)*CC*UU)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( Power(AA,6)* Power(BB,3)*DD*JJ*NN -  Power(AA,6)* Power(BB,3)*CC*JJ*OO +  Power(AA,6)* Power(BB,3)*CC*KK*OO -  Power(AA,6)* Power(BB,3)*CC*DD*VV))
-}\break
-
-$\dot{\Theta}_{ub}$ =
-{\tiny
-  ( - X3 + X4)/DD + (II*( - (BB*GG*LL*NN*PP*X1) + AA*GG*MM*NN*PP*X1 + BB*CC*LL*PP*RR*X1 - AA*CC*MM*PP*RR*X1 + BB*FF*LL*NN*PP*X2 - AA*FF*MM*NN*PP*X2 - BB*EE*LL*NN*QQ*X2 + AA*EE*MM*NN*QQ*X2 - CC*EE*LL*MM*RR*X2 + BB*EE*LL*NN*RR*X2 + AA*CC*MM*PP*RR*X2 - AA*BB*NN*PP*RR*X2 + EE*GG*LL*MM*NN*X3 - BB*FF*LL*NN*PP*X3 + AA*FF*MM*NN*PP*X3 - AA*GG*MM*NN*PP*X3 + BB*EE*LL*NN*QQ*X3 - AA*EE*MM*NN*QQ*X3 - BB*EE*LL*NN*RR*X3 + AA*BB*NN*PP*RR*X3 + BB*EE*GG*LL*NN*X5 - AA*EE*GG*MM*NN*X5 - BB*CC*EE*LL*RR*X5 + AA*CC*EE*MM*RR*X5 + CC*EE*GG*LL*MM*X6 - BB*EE*GG*LL*NN*X6 - BB*CC*FF*LL*PP*X6 + AA*CC*FF*MM*PP*X6 - AA*CC*GG*MM*PP*X6 + AA*BB*GG*NN*PP*X6 + BB*CC*EE*LL*QQ*X6 - AA*CC*EE*MM*QQ*X6 - CC*EE*GG*LL*MM*X7 + BB*CC*FF*LL*PP*X7 - AA*CC*FF*MM*PP*X7 + AA*CC*GG*MM*PP*X7 - BB*CC*EE*LL*QQ*X7 + AA*CC*EE*MM*QQ*X7 + BB*CC*EE*LL*RR*X7 - AA*BB*CC*PP*RR*X7))/(DD*(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT)) - ((( - JJ + KK)/DD - (II*(EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)*( - (JJ*NN) + CC*UU))/(DD*(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT)))*( - ((( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( - ((( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*NN*X2 +  Power(AA,4)* Power(BB,2)*NN*X3 -  Power(AA,4)* Power(BB,2)*CC*X7)))) + ( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( - ((( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,5)* Power(BB,3)*CC*DD*LL*X1 -  Power(AA,6)* Power(BB,2)*CC*DD*MM*X1 +  Power(AA,6)* Power(BB,2)*CC*DD*MM*X2 -  Power(AA,6)* Power(BB,3)*DD*NN*X2 +  Power(AA,6)* Power(BB,3)*DD*NN*X3 -  Power(AA,6)* Power(BB,3)*CC*OO*X3 +  Power(AA,6)* Power(BB,3)*CC*OO*X4 -  Power(AA,6)* Power(BB,3)*CC*DD*X8)))))/( - (( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*(( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( Power(AA,4)* Power(BB,2)*JJ*NN -  Power(AA,4)* Power(BB,2)*CC*UU)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( Power(AA,6)* Power(BB,3)*DD*JJ*NN -  Power(AA,6)* Power(BB,3)*CC*JJ*OO +  Power(AA,6)* Power(BB,3)*CC*KK*OO -  Power(AA,6)* Power(BB,3)*CC*DD*VV))
-}
+Solving the above system for $\dot{\Theta}_{bb}, \dot{\Theta}_{bm}, \dot{\Theta}_{um}$ and $\dot{\Theta}_{ub}$ gives the desired velocities. Solution was done via Mathematica.
+
+%% $\dot{\Theta}_{bb}$ =
+%% {\tiny
+%% ( - (GG*II*MM*NN*PP*X1) + CC*HH*MM*PP*RR*X1 - BB*HH*NN*PP*RR*X1 + BB*II*NN*PP*RR*X1 - CC*GG*MM*PP*SS*X1 + BB*GG*NN*PP*SS*X1 + CC*GG*MM*PP*TT*X1 - BB*CC*PP*RR*TT*X1 + FF*II*MM*NN*PP*X2 - EE*II*MM*NN*QQ*X2 + CC*FF*MM*PP*SS*X2 - BB*FF*NN*PP*SS*X2 - CC*EE*MM*QQ*SS*X2 + BB*EE*NN*QQ*SS*X2 - CC*FF*MM*PP*TT*X2 + CC*EE*MM*QQ*TT*X2 - FF*HH*MM*NN*PP*X3 + EE*HH*MM*NN*QQ*X3 + BB*FF*NN*PP*SS*X3 - BB*EE*NN*QQ*SS*X3 + EE*GG*II*MM*NN*X5 - CC*EE*HH*MM*RR*X5 + BB*EE*HH*NN*RR*X5 - BB*EE*II*NN*RR*X5 + CC*EE*GG*MM*SS*X5 - BB*EE*GG*NN*SS*X5 - CC*EE*GG*MM*TT*X5 + BB*CC*EE*RR*TT*X5 - CC*FF*HH*MM*PP*X6 + BB*FF*HH*NN*PP*X6 - BB*FF*II*NN*PP*X6 + CC*EE*HH*MM*QQ*X6 - BB*EE*HH*NN*QQ*X6 + BB*EE*II*NN*QQ*X6 + BB*CC*FF*PP*TT*X6 - BB*CC*EE*QQ*TT*X6 + CC*FF*HH*MM*PP*X7 - CC*EE*HH*MM*QQ*X7 - BB*CC*FF*PP*SS*X7 + BB*CC*EE*QQ*SS*X7)/(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT) - ((FF*PP - EE*QQ)*(HH*MM - BB*SS)*(JJ*NN - CC*UU)*( - ((( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( - ((( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*NN*X2 +  Power(AA,4)* Power(BB,2)*NN*X3 -  Power(AA,4)* Power(BB,2)*CC*X7)))) + ( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( - ((( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,5)* Power(BB,3)*CC*DD*LL*X1 -  Power(AA,6)* Power(BB,2)*CC*DD*MM*X1 +  Power(AA,6)* Power(BB,2)*CC*DD*MM*X2 -  Power(AA,6)* Power(BB,3)*DD*NN*X2 +  Power(AA,6)* Power(BB,3)*DD*NN*X3 -  Power(AA,6)* Power(BB,3)*CC*OO*X3 +  Power(AA,6)* Power(BB,3)*CC*OO*X4 -  Power(AA,6)* Power(BB,3)*CC*DD*X8)))))/(( - (EE*GG*II*LL*MM*NN) - BB*FF*HH*LL*NN*PP + BB*FF*II*LL*NN*PP + AA*FF*HH*MM*NN*PP - AA*FF*II*MM*NN*PP + AA*GG*II*MM*NN*PP + BB*EE*HH*LL*NN*QQ - BB*EE*II*LL*NN*QQ - AA*EE*HH*MM*NN*QQ + AA*EE*II*MM*NN*QQ + CC*EE*HH*LL*MM*RR - BB*EE*HH*LL*NN*RR + BB*EE*II*LL*NN*RR - AA*CC*HH*MM*PP*RR + AA*BB*HH*NN*PP*RR - AA*BB*II*NN*PP*RR - CC*EE*GG*LL*MM*SS + BB*EE*GG*LL*NN*SS + BB*CC*FF*LL*PP*SS - AA*CC*FF*MM*PP*SS + AA*CC*GG*MM*PP*SS - AA*BB*GG*NN*PP*SS - BB*CC*EE*LL*QQ*SS + AA*CC*EE*MM*QQ*SS + CC*EE*GG*LL*MM*TT - BB*CC*FF*LL*PP*TT + AA*CC*FF*MM*PP*TT - AA*CC*GG*MM*PP*TT + BB*CC*EE*LL*QQ*TT - AA*CC*EE*MM*QQ*TT - BB*CC*EE*LL*RR*TT + AA*BB*CC*PP*RR*TT)*( - (( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*(( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( Power(AA,4)* Power(BB,2)*JJ*NN -  Power(AA,4)* Power(BB,2)*CC*UU)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( Power(AA,6)* Power(BB,3)*DD*JJ*NN -  Power(AA,6)* Power(BB,3)*CC*JJ*OO +  Power(AA,6)* Power(BB,3)*CC*KK*OO -  Power(AA,6)* Power(BB,3)*CC*DD*VV)))
+%% } \break
+
+%% $\dot{\Theta}_{bm}$ =
+%% {\tiny
+%%   (GG*II*LL*NN*PP*X1 - CC*HH*LL*PP*RR*X1 + AA*HH*NN*PP*RR*X1 - AA*II*NN*PP*RR*X1 + CC*GG*LL*PP*SS*X1 - AA*GG*NN*PP*SS*X1 - CC*GG*LL*PP*TT*X1 + AA*CC*PP*RR*TT*X1 - FF*II*LL*NN*PP*X2 + EE*II*LL*NN*QQ*X2 - EE*II*LL*NN*RR*X2 + AA*II*NN*PP*RR*X2 - CC*FF*LL*PP*SS*X2 + AA*FF*NN*PP*SS*X2 + CC*EE*LL*QQ*SS*X2 - AA*EE*NN*QQ*SS*X2 + CC*FF*LL*PP*TT*X2 - CC*EE*LL*QQ*TT*X2 + CC*EE*LL*RR*TT*X2 - AA*CC*PP*RR*TT*X2 + FF*HH*LL*NN*PP*X3 - EE*HH*LL*NN*QQ*X3 + EE*HH*LL*NN*RR*X3 - AA*HH*NN*PP*RR*X3 - EE*GG*LL*NN*SS*X3 - AA*FF*NN*PP*SS*X3 + AA*GG*NN*PP*SS*X3 + AA*EE*NN*QQ*SS*X3 - EE*GG*II*LL*NN*X5 + CC*EE*HH*LL*RR*X5 - AA*EE*HH*NN*RR*X5 + AA*EE*II*NN*RR*X5 - CC*EE*GG*LL*SS*X5 + AA*EE*GG*NN*SS*X5 + CC*EE*GG*LL*TT*X5 - AA*CC*EE*RR*TT*X5 + EE*GG*II*LL*NN*X6 + CC*FF*HH*LL*PP*X6 - AA*FF*HH*NN*PP*X6 + AA*FF*II*NN*PP*X6 - AA*GG*II*NN*PP*X6 - CC*EE*HH*LL*QQ*X6 + AA*EE*HH*NN*QQ*X6 - AA*EE*II*NN*QQ*X6 - CC*EE*GG*LL*TT*X6 - AA*CC*FF*PP*TT*X6 + AA*CC*GG*PP*TT*X6 + AA*CC*EE*QQ*TT*X6 - CC*FF*HH*LL*PP*X7 + CC*EE*HH*LL*QQ*X7 - CC*EE*HH*LL*RR*X7 + AA*CC*HH*PP*RR*X7 + CC*EE*GG*LL*SS*X7 + AA*CC*FF*PP*SS*X7 - AA*CC*GG*PP*SS*X7 - AA*CC*EE*QQ*SS*X7)/(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT) + (( - (FF*HH*LL*PP) + EE*HH*LL*QQ - EE*HH*LL*RR + AA*HH*PP*RR + EE*GG*LL*SS + AA*FF*PP*SS - AA*GG*PP*SS - AA*EE*QQ*SS)*(JJ*NN - CC*UU)*( - ((( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( - ((( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*NN*X2 +  Power(AA,4)* Power(BB,2)*NN*X3 -  Power(AA,4)* Power(BB,2)*CC*X7)))) + ( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( - ((( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,5)* Power(BB,3)*CC*DD*LL*X1 -  Power(AA,6)* Power(BB,2)*CC*DD*MM*X1 +  Power(AA,6)* Power(BB,2)*CC*DD*MM*X2 -  Power(AA,6)* Power(BB,3)*DD*NN*X2 +  Power(AA,6)* Power(BB,3)*DD*NN*X3 -  Power(AA,6)* Power(BB,3)*CC*OO*X3 +  Power(AA,6)* Power(BB,3)*CC*OO*X4 -  Power(AA,6)* Power(BB,3)*CC*DD*X8)))))/((EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT)*( - (( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*(( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( Power(AA,4)* Power(BB,2)*JJ*NN -  Power(AA,4)* Power(BB,2)*CC*UU)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( Power(AA,6)* Power(BB,3)*DD*JJ*NN -  Power(AA,6)* Power(BB,3)*CC*JJ*OO +  Power(AA,6)* Power(BB,3)*CC*KK*OO -  Power(AA,6)* Power(BB,3)*CC*DD*VV)))
+%% }\break
+
+%% $\dot{\Theta}_{um}$ =
+%% {\tiny
+%%   ( - X2 + X3)/CC + (GG*( - (BB*LL*PP*X1) + AA*MM*PP*X1 + EE*LL*MM*X2 - AA*MM*PP*X2 + BB*EE*LL*X5 - AA*EE*MM*X5 - BB*EE*LL*X6 + AA*BB*PP*X6))/(CC*(EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)) - ((( - HH + II)/CC - (GG*(EE*LL - AA*PP)*( - (HH*MM) + BB*SS))/(CC*(EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)))*( - (BB*GG*LL*NN*PP*X1) + AA*GG*MM*NN*PP*X1 + BB*CC*LL*PP*RR*X1 - AA*CC*MM*PP*RR*X1 + BB*FF*LL*NN*PP*X2 - AA*FF*MM*NN*PP*X2 - BB*EE*LL*NN*QQ*X2 + AA*EE*MM*NN*QQ*X2 - CC*EE*LL*MM*RR*X2 + BB*EE*LL*NN*RR*X2 + AA*CC*MM*PP*RR*X2 - AA*BB*NN*PP*RR*X2 + EE*GG*LL*MM*NN*X3 - BB*FF*LL*NN*PP*X3 + AA*FF*MM*NN*PP*X3 - AA*GG*MM*NN*PP*X3 + BB*EE*LL*NN*QQ*X3 - AA*EE*MM*NN*QQ*X3 - BB*EE*LL*NN*RR*X3 + AA*BB*NN*PP*RR*X3 + BB*EE*GG*LL*NN*X5 - AA*EE*GG*MM*NN*X5 - BB*CC*EE*LL*RR*X5 + AA*CC*EE*MM*RR*X5 + CC*EE*GG*LL*MM*X6 - BB*EE*GG*LL*NN*X6 - BB*CC*FF*LL*PP*X6 + AA*CC*FF*MM*PP*X6 - AA*CC*GG*MM*PP*X6 + AA*BB*GG*NN*PP*X6 + BB*CC*EE*LL*QQ*X6 - AA*CC*EE*MM*QQ*X6 - CC*EE*GG*LL*MM*X7 + BB*CC*FF*LL*PP*X7 - AA*CC*FF*MM*PP*X7 + AA*CC*GG*MM*PP*X7 - BB*CC*EE*LL*QQ*X7 + AA*CC*EE*MM*QQ*X7 + BB*CC*EE*LL*RR*X7 - AA*BB*CC*PP*RR*X7))/(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT) - ((JJ/CC + ((EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)*(( - HH + II)/CC - (GG*(EE*LL - AA*PP)*( - (HH*MM) + BB*SS))/(CC*(EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)))*( - (JJ*NN) + CC*UU))/(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT))*( - ((( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( - ((( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*NN*X2 +  Power(AA,4)* Power(BB,2)*NN*X3 -  Power(AA,4)* Power(BB,2)*CC*X7)))) + ( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( - ((( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,5)* Power(BB,3)*CC*DD*LL*X1 -  Power(AA,6)* Power(BB,2)*CC*DD*MM*X1 +  Power(AA,6)* Power(BB,2)*CC*DD*MM*X2 -  Power(AA,6)* Power(BB,3)*DD*NN*X2 +  Power(AA,6)* Power(BB,3)*DD*NN*X3 -  Power(AA,6)* Power(BB,3)*CC*OO*X3 +  Power(AA,6)* Power(BB,3)*CC*OO*X4 -  Power(AA,6)* Power(BB,3)*CC*DD*X8)))))/( - (( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*(( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( Power(AA,4)* Power(BB,2)*JJ*NN -  Power(AA,4)* Power(BB,2)*CC*UU)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( Power(AA,6)* Power(BB,3)*DD*JJ*NN -  Power(AA,6)* Power(BB,3)*CC*JJ*OO +  Power(AA,6)* Power(BB,3)*CC*KK*OO -  Power(AA,6)* Power(BB,3)*CC*DD*VV))
+%% }\break
+
+%% $\dot{\Theta}_{ub}$ =
+%% {\tiny
+%%   ( - X3 + X4)/DD + (II*( - (BB*GG*LL*NN*PP*X1) + AA*GG*MM*NN*PP*X1 + BB*CC*LL*PP*RR*X1 - AA*CC*MM*PP*RR*X1 + BB*FF*LL*NN*PP*X2 - AA*FF*MM*NN*PP*X2 - BB*EE*LL*NN*QQ*X2 + AA*EE*MM*NN*QQ*X2 - CC*EE*LL*MM*RR*X2 + BB*EE*LL*NN*RR*X2 + AA*CC*MM*PP*RR*X2 - AA*BB*NN*PP*RR*X2 + EE*GG*LL*MM*NN*X3 - BB*FF*LL*NN*PP*X3 + AA*FF*MM*NN*PP*X3 - AA*GG*MM*NN*PP*X3 + BB*EE*LL*NN*QQ*X3 - AA*EE*MM*NN*QQ*X3 - BB*EE*LL*NN*RR*X3 + AA*BB*NN*PP*RR*X3 + BB*EE*GG*LL*NN*X5 - AA*EE*GG*MM*NN*X5 - BB*CC*EE*LL*RR*X5 + AA*CC*EE*MM*RR*X5 + CC*EE*GG*LL*MM*X6 - BB*EE*GG*LL*NN*X6 - BB*CC*FF*LL*PP*X6 + AA*CC*FF*MM*PP*X6 - AA*CC*GG*MM*PP*X6 + AA*BB*GG*NN*PP*X6 + BB*CC*EE*LL*QQ*X6 - AA*CC*EE*MM*QQ*X6 - CC*EE*GG*LL*MM*X7 + BB*CC*FF*LL*PP*X7 - AA*CC*FF*MM*PP*X7 + AA*CC*GG*MM*PP*X7 - BB*CC*EE*LL*QQ*X7 + AA*CC*EE*MM*QQ*X7 + BB*CC*EE*LL*RR*X7 - AA*BB*CC*PP*RR*X7))/(DD*(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT)) - ((( - JJ + KK)/DD - (II*(EE*GG*LL*MM - BB*FF*LL*PP + AA*FF*MM*PP - AA*GG*MM*PP + BB*EE*LL*QQ - AA*EE*MM*QQ - BB*EE*LL*RR + AA*BB*PP*RR)*( - (JJ*NN) + CC*UU))/(DD*(EE*GG*II*LL*MM*NN + BB*FF*HH*LL*NN*PP - BB*FF*II*LL*NN*PP - AA*FF*HH*MM*NN*PP + AA*FF*II*MM*NN*PP - AA*GG*II*MM*NN*PP - BB*EE*HH*LL*NN*QQ + BB*EE*II*LL*NN*QQ + AA*EE*HH*MM*NN*QQ - AA*EE*II*MM*NN*QQ - CC*EE*HH*LL*MM*RR + BB*EE*HH*LL*NN*RR - BB*EE*II*LL*NN*RR + AA*CC*HH*MM*PP*RR - AA*BB*HH*NN*PP*RR + AA*BB*II*NN*PP*RR + CC*EE*GG*LL*MM*SS - BB*EE*GG*LL*NN*SS - BB*CC*FF*LL*PP*SS + AA*CC*FF*MM*PP*SS - AA*CC*GG*MM*PP*SS + AA*BB*GG*NN*PP*SS + BB*CC*EE*LL*QQ*SS - AA*CC*EE*MM*QQ*SS - CC*EE*GG*LL*MM*TT + BB*CC*FF*LL*PP*TT - AA*CC*FF*MM*PP*TT + AA*CC*GG*MM*PP*TT - BB*CC*EE*LL*QQ*TT + AA*CC*EE*MM*QQ*TT + BB*CC*EE*LL*RR*TT - AA*BB*CC*PP*RR*TT)))*( - ((( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( - ((( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*NN*X2 +  Power(AA,4)* Power(BB,2)*NN*X3 -  Power(AA,4)* Power(BB,2)*CC*X7)))) + ( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( - ((( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,3)* Power(BB,2)*CC*LL*X1 -  Power(AA,4)*BB*CC*MM*X1 +  Power(AA,4)*BB*CC*MM*X2 -  Power(AA,4)* Power(BB,2)*CC*X6))) + ( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (LL*X1) + AA*X5)) + ( - (EE*LL) + AA*PP)*( Power(AA,5)* Power(BB,3)*CC*DD*LL*X1 -  Power(AA,6)* Power(BB,2)*CC*DD*MM*X1 +  Power(AA,6)* Power(BB,2)*CC*DD*MM*X2 -  Power(AA,6)* Power(BB,3)*DD*NN*X2 +  Power(AA,6)* Power(BB,3)*DD*NN*X3 -  Power(AA,6)* Power(BB,3)*CC*OO*X3 +  Power(AA,6)* Power(BB,3)*CC*OO*X4 -  Power(AA,6)* Power(BB,3)*CC*DD*X8)))))/( - (( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*(( Power(AA,6)* Power(BB,2)*CC*DD*HH*MM -  Power(AA,6)* Power(BB,3)*DD*HH*NN +  Power(AA,6)* Power(BB,3)*DD*II*NN -  Power(AA,6)* Power(BB,3)*CC*II*OO)*( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR))) - ( - (EE*LL) + AA*PP)*(( Power(AA,5)* Power(BB,3)*CC*DD*FF*LL -  Power(AA,6)* Power(BB,2)*CC*DD*FF*MM +  Power(AA,6)* Power(BB,2)*CC*DD*GG*MM -  Power(AA,6)* Power(BB,3)*DD*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,5)* Power(BB,3)*CC*DD*EE*LL -  Power(AA,6)* Power(BB,2)*CC*DD*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS))*( Power(AA,4)* Power(BB,2)*JJ*NN -  Power(AA,4)* Power(BB,2)*CC*UU)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( - (( - (EE*LL) + AA*PP)*(( Power(AA,3)* Power(BB,2)*CC*FF*LL -  Power(AA,4)*BB*CC*FF*MM +  Power(AA,4)*BB*CC*GG*MM -  Power(AA,4)* Power(BB,2)*GG*NN)*( - (EE*LL) + AA*PP) - ( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*CC*SS)) + ( - (EE*LL) + AA*PP)*( - (( Power(AA,3)* Power(BB,2)*CC*EE*LL -  Power(AA,4)*BB*CC*EE*MM)*( - (FF*LL) + AA*QQ)) + ( - (EE*LL) + AA*PP)*( Power(AA,4)*BB*CC*GG*MM -  Power(AA,2)*BB*CC*( - (AA*BB*FF*LL) +  Power(AA,2)*FF*MM +  Power(AA,2)*BB*RR)))*( Power(AA,4)*BB*CC*HH*MM -  Power(AA,4)* Power(BB,2)*HH*NN +  Power(AA,4)* Power(BB,2)*II*NN -  Power(AA,4)* Power(BB,2)*CC*TT))*( Power(AA,6)* Power(BB,3)*DD*JJ*NN -  Power(AA,6)* Power(BB,3)*CC*JJ*OO +  Power(AA,6)* Power(BB,3)*CC*KK*OO -  Power(AA,6)* Power(BB,3)*CC*DD*VV))
+%% }
 
@@
 
-$\dot{Ln}$ =
-{\tiny
-    (h*l*p*t*w*x1 - g*l*p*u*w*x1 + g*l*p*s*x*x1 - f*l*p*t*x*x1 +
-     d*l*p*u*w*x2 - c*l*q*u*w*x2 - d*l*p*s*x*x2 + c*l*q*s*x*x2 -
-     c*l*n*t*x*x2 + b*l*p*t*x*x2 - c*h*l*t*w*x4 + c*g*l*u*w*x4 -
-     c*g*l*s*x*x4 + c*f*l*t*x*x4 - d*h*l*p*w*x5 + c*h*l*q*w*x5 +
-     c*g*l*n*x*x5 + d*f*l*p*x*x5 - b*g*l*p*x*x5 - c*f*l*q*x*x5 +
-     d*h*l*p*s*x6 - c*h*l*q*s*x6 + c*h*l*n*t*x6 - b*h*l*p*t*x6 -
-     c*g*l*n*u*x6 - d*f*l*p*u*x6 + b*g*l*p*u*x6 + c*f*l*q*u*x6 -
-     g*k*p*s*x1*y - h*j*p*t*x1*y + f*k*p*t*x1*y + g*j*p*u*x1*y +
-     d*k*p*s*x2*y - c*k*q*s*x2*y + c*k*n*t*x2*y - b*k*p*t*x2*y -
-     d*j*p*u*x2*y + c*j*q*u*x2*y - d*h*p*s*x3*y + c*h*q*s*x3*y -
-     c*h*n*t*x3*y + b*h*p*t*x3*y + c*g*n*u*x3*y + d*f*p*u*x3*y -
-     b*g*p*u*x3*y - c*f*q*u*x3*y + c*g*k*s*x4*y + c*h*j*t*x4*y -
-     c*f*k*t*x4*y - c*g*j*u*x4*y - c*g*k*n*x5*y + d*h*j*p*x5*y -
-     d*f*k*p*x5*y + b*g*k*p*x5*y - c*h*j*q*x5*y + c*f*k*q*x5*y)/
-     (d*h*l*p*s*v - c*h*l*q*s*v + c*h*l*n*t*v - b*h*l*p*t*v -
-     c*g*l*n*u*v - d*f*l*p*u*v + b*g*l*p*u*v + c*f*l*q*u*v -
-     d*h*l*p*r*w + c*h*l*q*r*w - c*h*l*m*t*w + a*h*l*p*t*w +
-     c*g*l*m*u*w + d*e*l*p*u*w - a*g*l*p*u*w - c*e*l*q*u*w +
-     c*g*l*n*r*x + d*f*l*p*r*x - b*g*l*p*r*x - c*f*l*q*r*x -
-     c*g*l*m*s*x - d*e*l*p*s*x + a*g*l*p*s*x + c*e*l*q*s*x +
-     c*f*l*m*t*x - c*e*l*n*t*x + b*e*l*p*t*x - a*f*l*p*t*x -
-     c*g*k*n*r*y + d*h*j*p*r*y - d*f*k*p*r*y + b*g*k*p*r*y -
-     c*h*j*q*r*y + c*f*k*q*r*y + c*g*k*m*s*y - d*h*i*p*s*y +
-     d*e*k*p*s*y - a*g*k*p*s*y + c*h*i*q*s*y - c*e*k*q*s*y +
-     c*h*j*m*t*y - c*f*k*m*t*y - c*h*i*n*t*y + c*e*k*n*t*y +
-     b*h*i*p*t*y - a*h*j*p*t*y - b*e*k*p*t*y + a*f*k*p*t*y -
-     c*g*j*m*u*y + c*g*i*n*u*y + d*f*i*p*u*y - b*g*i*p*u*y -
-     d*e*j*p*u*y + a*g*j*p*u*y - c*f*i*q*u*y + c*e*j*q*u*y)
-}\break
-
-$\dot{Lf}$ =
-{\tiny
-    (-(h*l*p*t*v*x1) + g*l*p*u*v*x1 - g*l*p*r*x*x1 + e*l*p*t*x*x1 -
-     d*l*p*u*v*x2 + c*l*q*u*v*x2 + d*l*p*r*x*x2 - c*l*q*r*x*x2 +
-     c*l*m*t*x*x2 - a*l*p*t*x*x2 + c*h*l*t*v*x4 - c*g*l*u*v*x4 +
-     c*g*l*r*x*x4 - c*e*l*t*x*x4 + d*h*l*p*v*x5 - c*h*l*q*v*x5 -
-     c*g*l*m*x*x5 - d*e*l*p*x*x5 + a*g*l*p*x*x5 + c*e*l*q*x*x5 -
-     d*h*l*p*r*x6 + c*h*l*q*r*x6 - c*h*l*m*t*x6 + a*h*l*p*t*x6 +
-     c*g*l*m*u*x6 + d*e*l*p*u*x6 - a*g*l*p*u*x6 - c*e*l*q*u*x6 +
-     g*k*p*r*x1*y + h*i*p*t*x1*y - e*k*p*t*x1*y - g*i*p*u*x1*y -
-     d*k*p*r*x2*y + c*k*q*r*x2*y - c*k*m*t*x2*y + a*k*p*t*x2*y +
-     d*i*p*u*x2*y - c*i*q*u*x2*y + d*h*p*r*x3*y - c*h*q*r*x3*y +
-     c*h*m*t*x3*y - a*h*p*t*x3*y - c*g*m*u*x3*y - d*e*p*u*x3*y +
-     a*g*p*u*x3*y + c*e*q*u*x3*y - c*g*k*r*x4*y - c*h*i*t*x4*y +
-     c*e*k*t*x4*y + c*g*i*u*x4*y + c*g*k*m*x5*y - d*h*i*p*x5*y +
-     d*e*k*p*x5*y - a*g*k*p*x5*y + c*h*i*q*x5*y - c*e*k*q*x5*y)/
-     (d*h*l*p*s*v - c*h*l*q*s*v + c*h*l*n*t*v - b*h*l*p*t*v -
-     c*g*l*n*u*v - d*f*l*p*u*v + b*g*l*p*u*v + c*f*l*q*u*v -
-     d*h*l*p*r*w + c*h*l*q*r*w - c*h*l*m*t*w + a*h*l*p*t*w +
-     c*g*l*m*u*w + d*e*l*p*u*w - a*g*l*p*u*w - c*e*l*q*u*w +
-     c*g*l*n*r*x + d*f*l*p*r*x - b*g*l*p*r*x - c*f*l*q*r*x -
-     c*g*l*m*s*x - d*e*l*p*s*x + a*g*l*p*s*x + c*e*l*q*s*x +
-     c*f*l*m*t*x - c*e*l*n*t*x + b*e*l*p*t*x - a*f*l*p*t*x -
-     c*g*k*n*r*y + d*h*j*p*r*y - d*f*k*p*r*y + b*g*k*p*r*y -
-     c*h*j*q*r*y + c*f*k*q*r*y + c*g*k*m*s*y - d*h*i*p*s*y +
-     d*e*k*p*s*y - a*g*k*p*s*y + c*h*i*q*s*y - c*e*k*q*s*y +
-     c*h*j*m*t*y - c*f*k*m*t*y - c*h*i*n*t*y + c*e*k*n*t*y +
-     b*h*i*p*t*y - a*h*j*p*t*y - b*e*k*p*t*y + a*f*k*p*t*y -
-     c*g*j*m*u*y + c*g*i*n*u*y + d*f*i*p*u*y - b*g*i*p*u*y -
-     d*e*j*p*u*y + a*g*j*p*u*y - c*f*i*q*u*y + c*e*j*q*u*y)
-}
+Solving the above system for $\dot{\Theta}_{nm}$ and $\dot{\Theta}_{fm}$ gives the desired velocities. Solution was done via Mathematica.
+
+%% $\dot{Ln}$ =
+%% {\tiny
+%%     (h*l*p*t*w*x1 - g*l*p*u*w*x1 + g*l*p*s*x*x1 - f*l*p*t*x*x1 +
+%%      d*l*p*u*w*x2 - c*l*q*u*w*x2 - d*l*p*s*x*x2 + c*l*q*s*x*x2 -
+%%      c*l*n*t*x*x2 + b*l*p*t*x*x2 - c*h*l*t*w*x4 + c*g*l*u*w*x4 -
+%%      c*g*l*s*x*x4 + c*f*l*t*x*x4 - d*h*l*p*w*x5 + c*h*l*q*w*x5 +
+%%      c*g*l*n*x*x5 + d*f*l*p*x*x5 - b*g*l*p*x*x5 - c*f*l*q*x*x5 +
+%%      d*h*l*p*s*x6 - c*h*l*q*s*x6 + c*h*l*n*t*x6 - b*h*l*p*t*x6 -
+%%      c*g*l*n*u*x6 - d*f*l*p*u*x6 + b*g*l*p*u*x6 + c*f*l*q*u*x6 -
+%%      g*k*p*s*x1*y - h*j*p*t*x1*y + f*k*p*t*x1*y + g*j*p*u*x1*y +
+%%      d*k*p*s*x2*y - c*k*q*s*x2*y + c*k*n*t*x2*y - b*k*p*t*x2*y -
+%%      d*j*p*u*x2*y + c*j*q*u*x2*y - d*h*p*s*x3*y + c*h*q*s*x3*y -
+%%      c*h*n*t*x3*y + b*h*p*t*x3*y + c*g*n*u*x3*y + d*f*p*u*x3*y -
+%%      b*g*p*u*x3*y - c*f*q*u*x3*y + c*g*k*s*x4*y + c*h*j*t*x4*y -
+%%      c*f*k*t*x4*y - c*g*j*u*x4*y - c*g*k*n*x5*y + d*h*j*p*x5*y -
+%%      d*f*k*p*x5*y + b*g*k*p*x5*y - c*h*j*q*x5*y + c*f*k*q*x5*y)/
+%%      (d*h*l*p*s*v - c*h*l*q*s*v + c*h*l*n*t*v - b*h*l*p*t*v -
+%%      c*g*l*n*u*v - d*f*l*p*u*v + b*g*l*p*u*v + c*f*l*q*u*v -
+%%      d*h*l*p*r*w + c*h*l*q*r*w - c*h*l*m*t*w + a*h*l*p*t*w +
+%%      c*g*l*m*u*w + d*e*l*p*u*w - a*g*l*p*u*w - c*e*l*q*u*w +
+%%      c*g*l*n*r*x + d*f*l*p*r*x - b*g*l*p*r*x - c*f*l*q*r*x -
+%%      c*g*l*m*s*x - d*e*l*p*s*x + a*g*l*p*s*x + c*e*l*q*s*x +
+%%      c*f*l*m*t*x - c*e*l*n*t*x + b*e*l*p*t*x - a*f*l*p*t*x -
+%%      c*g*k*n*r*y + d*h*j*p*r*y - d*f*k*p*r*y + b*g*k*p*r*y -
+%%      c*h*j*q*r*y + c*f*k*q*r*y + c*g*k*m*s*y - d*h*i*p*s*y +
+%%      d*e*k*p*s*y - a*g*k*p*s*y + c*h*i*q*s*y - c*e*k*q*s*y +
+%%      c*h*j*m*t*y - c*f*k*m*t*y - c*h*i*n*t*y + c*e*k*n*t*y +
+%%      b*h*i*p*t*y - a*h*j*p*t*y - b*e*k*p*t*y + a*f*k*p*t*y -
+%%      c*g*j*m*u*y + c*g*i*n*u*y + d*f*i*p*u*y - b*g*i*p*u*y -
+%%      d*e*j*p*u*y + a*g*j*p*u*y - c*f*i*q*u*y + c*e*j*q*u*y)
+%% }\break
+
+%% $\dot{Lf}$ =
+%% {\tiny
+%%     (-(h*l*p*t*v*x1) + g*l*p*u*v*x1 - g*l*p*r*x*x1 + e*l*p*t*x*x1 -
+%%      d*l*p*u*v*x2 + c*l*q*u*v*x2 + d*l*p*r*x*x2 - c*l*q*r*x*x2 +
+%%      c*l*m*t*x*x2 - a*l*p*t*x*x2 + c*h*l*t*v*x4 - c*g*l*u*v*x4 +
+%%      c*g*l*r*x*x4 - c*e*l*t*x*x4 + d*h*l*p*v*x5 - c*h*l*q*v*x5 -
+%%      c*g*l*m*x*x5 - d*e*l*p*x*x5 + a*g*l*p*x*x5 + c*e*l*q*x*x5 -
+%%      d*h*l*p*r*x6 + c*h*l*q*r*x6 - c*h*l*m*t*x6 + a*h*l*p*t*x6 +
+%%      c*g*l*m*u*x6 + d*e*l*p*u*x6 - a*g*l*p*u*x6 - c*e*l*q*u*x6 +
+%%      g*k*p*r*x1*y + h*i*p*t*x1*y - e*k*p*t*x1*y - g*i*p*u*x1*y -
+%%      d*k*p*r*x2*y + c*k*q*r*x2*y - c*k*m*t*x2*y + a*k*p*t*x2*y +
+%%      d*i*p*u*x2*y - c*i*q*u*x2*y + d*h*p*r*x3*y - c*h*q*r*x3*y +
+%%      c*h*m*t*x3*y - a*h*p*t*x3*y - c*g*m*u*x3*y - d*e*p*u*x3*y +
+%%      a*g*p*u*x3*y + c*e*q*u*x3*y - c*g*k*r*x4*y - c*h*i*t*x4*y +
+%%      c*e*k*t*x4*y + c*g*i*u*x4*y + c*g*k*m*x5*y - d*h*i*p*x5*y +
+%%      d*e*k*p*x5*y - a*g*k*p*x5*y + c*h*i*q*x5*y - c*e*k*q*x5*y)/
+%%      (d*h*l*p*s*v - c*h*l*q*s*v + c*h*l*n*t*v - b*h*l*p*t*v -
+%%      c*g*l*n*u*v - d*f*l*p*u*v + b*g*l*p*u*v + c*f*l*q*u*v -
+%%      d*h*l*p*r*w + c*h*l*q*r*w - c*h*l*m*t*w + a*h*l*p*t*w +
+%%      c*g*l*m*u*w + d*e*l*p*u*w - a*g*l*p*u*w - c*e*l*q*u*w +
+%%      c*g*l*n*r*x + d*f*l*p*r*x - b*g*l*p*r*x - c*f*l*q*r*x -
+%%      c*g*l*m*s*x - d*e*l*p*s*x + a*g*l*p*s*x + c*e*l*q*s*x +
+%%      c*f*l*m*t*x - c*e*l*n*t*x + b*e*l*p*t*x - a*f*l*p*t*x -
+%%      c*g*k*n*r*y + d*h*j*p*r*y - d*f*k*p*r*y + b*g*k*p*r*y -
+%%      c*h*j*q*r*y + c*f*k*q*r*y + c*g*k*m*s*y - d*h*i*p*s*y +
+%%      d*e*k*p*s*y - a*g*k*p*s*y + c*h*i*q*s*y - c*e*k*q*s*y +
+%%      c*h*j*m*t*y - c*f*k*m*t*y - c*h*i*n*t*y + c*e*k*n*t*y +
+%%      b*h*i*p*t*y - a*h*j*p*t*y - b*e*k*p*t*y + a*f*k*p*t*y -
+%%      c*g*j*m*u*y + c*g*i*n*u*y + d*f*i*p*u*y - b*g*i*p*u*y -
+%%      d*e*j*p*u*y + a*g*j*p*u*y - c*f*i*q*u*y + c*e*j*q*u*y)
+%% }
 

### 2017-01-27 311/latex/HW1.tex
Add lots of stuff.

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\nonstopmode

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{amsmath}

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\usepackage{amssymb}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Advanced Calc Homework \large 1: 1.1: 1,3,5,7,8,11,17, 1.2: 1,4}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem 1.1}

\section{Problem 1.1.3}
\textbf{WTS} For any natural number $n$:

\begin{equation*}
  \sum_{j=1}^{n} j^2 = \frac{n(n+1)(2n+1)}{6}
\end{equation*}

\textbf{Proof}
We will prove this using the Principle of Induction. We will first show that the statement S(k), where S(k) is the above equality with $n=k$, is true for $S(1)$, then show that $S(k)$ being true implies $S(k+1)$ is true.\\

\textbf{Base case k=1}
Notice that for $k=1$:

\begin{align*}
  &\sum_{j=1}^{1} j^2 = 1\\
  &\frac{1(1+1)(2+1)}{6} = \frac66 = 1\\
\end{align*}

Thus the base case holds, so $S(1)$ is true.\\

\textbf{Inductive step}
Now we let S(k) be true, so:

\begin{equation*}
  \sum_{j=1}^{k} j^2 = \frac{k(k+1)(2k+1)}{6}
\end{equation*}

We want to show S(m) is true, where $m=k+1$. Add $(k+1)^2$ to both sides and do algebra:

\begin{align*}
  \sum_{j=1}^{k} j^2 + (k+1)^2 &= \frac{k(k+1)(2k+1)}{6} + (k+1)^2\\
  \sum_{j=1}^{m} j^2 &= \frac{k(k+1)(2k+1)}{6} + k^2 + 2k + 1\\
  &= \frac{2k^3+3k^2+k}{6} + \frac{6k^2 + 12k + 6}{6}\\
  &= \frac{2k^3+8k^2+13k+6}{6}\\
  &= \frac{(k+1)(2k^2+7k+6)}{6}\\
  &= \frac{(k+1)(k+2)(2k+3)}{6}\\
  &= \frac{m(m+1)(2m+1)}{6}\\
\end{align*}

thus $S(k+1)$. We have shown the base case S(1) and inductive step $S(k)$ implies $S(k+1)$, so the statement $S(n)$ must be true for all $n\in \mathbb{N}$.\\

\section{Problem 1.1.5}

\section{Problem 1.1.7}
\textbf{WTS} $n \in \mathbb{N}$ implies $n-1 \in \mathbb{N}$\\

\textbf{Proof}

Consider set $S = \{n | n = 1 \mbox{ or for } n > 1, n, n-1 \in \mathbb{N}\}$. Note that the inductive property holds for this set, since for $s \in S$, $s+1 \in S$.

\section{Problem 1.1.8}
\textbf{WTS} For $n,m \in \mathbb{N}$ where $n > m$, $n-m \in \mathbb{N}$.

\textbf{Proof} We will prove this using the Principle of Induction. We will show that the statement $S(k)$ is true for all $k\in\mathbb{N}$, where S(k) is the statement $n-k\in\mathbb{N}$.

\textbf{Base case} The base case S(1) states that for $n\in\mathh{N}$, $n-1\in{mathbb}$. This is true by Problem 1.1.7.\\

\textbf{Inductive step} Statement S(k) says that for $n\in\mathbb{N}$, $n-k\in\mathbb{N}$. Note that, by Problem 1.1.7, $n-k-1\in\mathbb{N}$.\\

Thus both $S(1)$ and $S(k)\rightarrow S(k+1)$ are true. Thus $S(n)$ is true for all $n\in\mathbb{N}$, so for $n,m \in \mathbb{N}$ where $n > m$, $n-m \in \mathbb{N}$.\\

\section{Problem 1.1.11a}
\textbf{WTS} If $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $a+b \in \mathb{R}\ \mathbb{Q}$.

\textbf{Proof} In search of a contradiction, assume for $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, that $a+b \in \mathbb{Q}$. Then by the definition of rational numbers there exist $m,n \in \mathbb{Z}$ such that $a+b = \frac{m}{n}$. $a$ can be similarly expanded using $x,y \in \mathbb{Z}: $\frac{x}{y}+b = \frac{m}{n}$. We will perform algebra and show this leads to a contradiction:

\begin{align*}
  &\frac{x}{y}+b = \frac{m}{n}\\
  &b = \frac{m}{n} - \frac{x}{y}
  &b = \frac{my}{ny} - \frac{xn}{ny}
  &b = \frac{my-xn}{ny}\\
\end{align*}

Because the product and difference of two integers is also an integer, this means b can be expressed as a fraction of two integers. This is the desired contradiction. Thus, if $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $a+b \in \mathb{R}\ \mathbb{Q}$.

\section{Problem 1.1.11b}
\textbf{WTS} If $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $ab \in \mathb{R}\ \mathbb{Q}$.

\textbf{Proof} In search of a contradiction, assume for $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, that $ab \in \mathbb{Q}$. Then by the definition of rational numbers there exist $m,n \in \mathbb{Z}$ such that $ab = \frac{m}{n}$. $a$ can be similarly expanded using $x,y \in \mathbb{Z}: $\frac{x}{y}b = \frac{m}{n}$. This can be rewritten as $b = \frac{my}{nx}$. Because the product of two integers is itself an integer, we have expressed b as a fraction of two integers. This is the desired contradiction. Thus, if $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $ab \in \mathb{R}\ \mathbb{Q}$.

\section{Problem 1.1.17}
Consider set $S = \{x | x \in \mathbb{R}, x \geq 0, x^2 < c\}$.

\textbf{A}\\
We wish to show that $\forall s \in S$, $s \leq c + 1$. Consider $s \in S$. If $s > 1$ then, by multiplying by s, $s^2 > s$. Thus by the definition of S, $s < s^2 < c$ and so $s < c+1$. If $s \leq 1$, then since $c$ must be positive, $c + 1 > 1$ and so $s < c + 1$.\\

\textbf{B}\\

\end{document}

### 2017-01-27 311/latex/HW1.tex
Add 311 homework almost done.

@@
 
-\section{Problem 1.1}
+\section{Problem 1.1.1}
+\textbf{A} False, since $1$ isn't irrational.
+\textbf{B} False, since $\sqrt{2}$ is irrational, and thus $2$ cannot be inductive stepped to from $1$.
+\textbf{C} False. Consider $a = \sqrt{2}$ and $b = 2 - \sqrt{2}$. Both are irrational, since the sum of a rational and irrational is irrational, but their sum is a rational.
+\textbf{D} False for a similar reason: consider $\pi$ and $\frac{1}{\pi}$.
+Both irrational since rational and irrational numbers combine via multiplication to be irrational, but have a rational product.
+\textbf{E} True. ISOAC, let $n^2$ be odd but $n$ even. Thus $\exists k\in\mathbb{N} st n=2k$, so $n^2 = 4k^2 = 2*2k^2$. This is a contradiction, since $n^2$ was supposed to be odd.\\
 
@@
 
-\textbf{Base case} The base case S(1) states that for $n\in\mathh{N}$, $n-1\in{mathbb}$. This is true by Problem 1.1.7.\\
+\textbf{Base case} The base case S(1) states that for $n\in\mathbb{N}$, $n-1\in{mathbb}$. This is true by Problem 1.1.7.\\
 
@@
 \section{Problem 1.1.11a}
-\textbf{WTS} If $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $a+b \in \mathb{R}\ \mathbb{Q}$.
+\textbf{WTS} If $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $a+b \in \mathbb{R}\ \mathbb{Q}$.
 
@@
 Then by the definition of rational numbers there exist $m,n \in \mathbb{Z}$ such that $a+b = \frac{m}{n}$.
-$a$ can be similarly expanded using $x,y \in \mathbb{Z}: $\frac{x}{y}+b = \frac{m}{n}$.
+$a$ can be similarly expanded using $x,y \in \mathbb{Z}: \frac{x}{y}+b = \frac{m}{n}$.
 We will perform algebra and show this leads to a contradiction:
@@
 This is the desired contradiction.
-Thus, if $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $a+b \in \mathb{R}\ \mathbb{Q}$.
+Thus, if $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $a+b \in \mathbb{R}\ \mathbb{Q}$.
 
 \section{Problem 1.1.11b}
-\textbf{WTS} If $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $ab \in \mathb{R}\ \mathbb{Q}$.
+\textbf{WTS} If $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $ab \in \mathbb{R}\ \mathbb{Q}$.
 
@@
 Then by the definition of rational numbers there exist $m,n \in \mathbb{Z}$ such that $ab = \frac{m}{n}$.
-$a$ can be similarly expanded using $x,y \in \mathbb{Z}: $\frac{x}{y}b = \frac{m}{n}$.
+$a$ can be similarly expanded using $x,y \in \mathbb{Z}: \frac{x}{y}b = \frac{m}{n}$.
 This can be rewritten as $b = \frac{my}{nx}$.
@@
 This is the desired contradiction.
-Thus, if $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $ab \in \mathb{R}\ \mathbb{Q}$.
+Thus, if $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $ab \in \mathbb{R}\ \mathbb{Q}$.
 
@@
 
+Then, by the Completeness Axiom, there must exist some $b \in R$ such that $b = Sub(S)$.\\
+
 \textbf{B}\\
 
+\section{Problem 1.2.1}
+\textbf{A} False. Consider real numbers $1.6$ and $1.8$. There are no integers between these two, thus the integers are not dense.\\
+\textbf{B} False. Consider real numbers $-2.0 and -1.0$. There are no positive reals between these two, and thus the positive reals are not dense.\\
+\textbf{C} True.
+The rational numbers are dense in the reals.
+For reals a, b, where $a > b$, by the Archimedean principle there exists a number $n \in \mathbb{N}$ such that $\frac{1}{n} \leq a-b$.
+Then $b + \frac{1}{n}$ is between the two.
+If it is an integer, then $b + \frac{1}{n^2}$ shouldn't be, and is also between the two.
+
+\section{Problem 1.2.4}
+\textbf{A} Set $\{1/n | n \in \mathbb{N}\}$
+\textit{Maximum} - 1, all $n > 1$ are smaller.\\
+\textit{Minimum} - None. For any $1/n$ in the set, $1/(n+1)$ is smaller.\\
+\textit{Supremum} - 1. $1/1$ is the largest value in the set.\\
+\textit{Infimum} - 0. Any number $\epsilon > 0$ has a number $1\m$ smaller than it, by the Archimedean Principle, so 0 is the largest lower bound.\\
+
+\textbf{B} Set $\{x | x \in \mathbb{R} \mbox{ and } x^2 < 2\}$
+\textit{Maximum} - None. For any number $e < \sqrt{2}$, $e + \frac{\sqrt{2}-e}{2}$ is bigger
+\textit{Minimum} - Same argument with $-\sqrt{2}$.\\
+\textit{Supremum} - $\sqrt{2}$. Anything smaller is in the set, and it isn't in the set.
+\textit{Infimum} - $-\sqrt{2}$ by the same argument.\\
 \end{document}

### 2017-01-27 thesis/latex/tate-figures-1-17.tex
bla

@@
 
-\begin{figure}
+\begin{figure}[h!]
   \centering
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=.45\textwidth,keepaspectratio]{../figures/retrograde_transport.jpg}
-  \caption{Mechanism of dynein localization to axon tips, or growth cones, in retrograde transport mediated by kinesin anterograde transport. Modified from \cite{valetoolbox}.}
-  \label{retrograde-transport}
-\end{figure}
+%% \begin{figure}[h!]
+%%   \centering
+%%   \includegraphics[width=.45\textwidth,keepaspectratio]{../figures/retrograde_transport.jpg}
+%%   \caption{Mechanism of dynein localization to axon tips, or growth cones, in retrograde transport mediated by kinesin anterograde transport. Modified from \cite{valetoolbox}.}
+%%   \label{retrograde-transport}
+%% \end{figure}
 
-\begin{figure}
-  \centering
-  \includegraphics[width=.45\textwidth,keepaspectratio]{../figures/mtbd-complex.png}
-  \caption{Crystal structure of the MTBD-MT complex. Modified from Redwine \textit{et. al} \cite{redwineMTBDcomplex}.}
-  \label{dynein-artist-rendition-2}
-\end{figure}
+%% \begin{figure}[h!]
+%%   \centering
+%%   \includegraphics[width=.45\textwidth,keepaspectratio]{../figures/mtbd-complex.png}
+%%   \caption{Crystal structure of the MTBD-MT complex. Modified from Redwine \textit{et. al} \cite{redwineMTBDcomplex}.}
+%%   \label{dynein-artist-rendition-2}
+%% \end{figure}
 
-\begin{figure}
-  \centering
-  \includegraphics[width=.45\textwidth,keepaspectratio]{../figures/mechanochemical-cycle.jpeg}
-  \caption{Cycle of states the dynein motor goes through during motion in the Cianfrocco model \cite{cianfroccoreview}.}
-  \label{mech-cycle}
-\end{figure}
-
-\begin{figure}
+\begin{figure}[h!]
   \centering
@@
 
-\begin{figure}
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.85\textwidth,keepaspectratio]{../figures/mechanochemical-cycle.jpeg}
+  \caption{Cycle of states the dynein motor goes through during motion in the Cianfrocco model \cite{cianfroccoreview}.}
+  \label{mech-cycle}
+\end{figure}
+
+\begin{figure}[h!]
   \centering
@@
 
+\begin{figure}[h!]
+  \centering
+  \begin{minipage}[b]{0.49\textwidth}
+    \includegraphics[width=\textwidth]{../figures/length_histogram_sample}
+    \caption{Histogram of step lengths over 3 seconds of simulation.}
+  \end{minipage}
+  \begin{minipage}[b]{0.49\textwidth}
+    \includegraphics[width=\textwidth]{../figures/time_histogram_sample}
+    \caption{Histogram of step times over 3 seconds of simulation.}
+  \end{minipage}
+\end{figure}
+
 \end{document}

### 2017-02-02 311/latex/HW1.tex

@@
 \section{Problem 1.1.1}
-\textbf{A} False, since $1$ isn't irrational.
-\textbf{B} False, since $\sqrt{2}$ is irrational, and thus $2$ cannot be inductive stepped to from $1$.
-\textbf{C} False. Consider $a = \sqrt{2}$ and $b = 2 - \sqrt{2}$. Both are irrational, since the sum of a rational and irrational is irrational, but their sum is a rational.
+\textbf{A} False, since $1$ isn't irrational.\\
+\textbf{B} False, since $\sqrt{2}$ is irrational, and thus $2$ cannot be inductive stepped to from $1$.\\
+\textbf{C} False. Consider $a = \sqrt{2}$ and $b = 2 - \sqrt{2}$. Both are irrational, since the sum of a rational and irrational is irrational, but their sum is a rational.\\
 \textbf{D} False for a similar reason: consider $\pi$ and $\frac{1}{\pi}$.
-Both irrational since rational and irrational numbers combine via multiplication to be irrational, but have a rational product.
+Both irrational since rational and irrational numbers combine via multiplication to be irrational, but have a rational product.\\
 \textbf{E} True. ISOAC, let $n^2$ be odd but $n$ even. Thus $\exists k\in\mathbb{N} st n=2k$, so $n^2 = 4k^2 = 2*2k^2$. This is a contradiction, since $n^2$ was supposed to be odd.\\
@@
 \section{Problem 1.1.5}
+Before we do this proof, we'll need to prove a subclaim:\\
+\textbf{WTS} S(k): $(1 + 2 + ... + k) = \frac{k(k+1)}{2}$\\
+
+\textbf{Proof} We will use the Principle of Induction to prove the above claim for all S(k). First, the base case: S(1): $1 = \frac{1(2)}{2} = 1$. Then the inductive step:
+
+\begin{align*}
+  (1 + 2 + ... + k) &= \frac{k(k+1)}{2}\\
+  1 + 2 + ... + k + k + 1 &= \frac{k(k+1)}{2} + k + 1\\
+  &= \frac{k^2 + 3k + 2}{2}\\
+  &= \frac{(k+1)(k+2)}{2}\\
+\end{align*}
+
+thus $S(k) \rightarrow S(k+1)$, so S(n) is true for all $n \in \mathbb{N}$. Now to prove the problem claim.\\
+
+\textbf{WTS} S(k): $(1 + 2 + ... + k)^2 = 1^3 + 2^3 + ... + k^3$\\
+
+\textbf{Proof} We will use the Principle of Induction to prove that S(k) is true for all $k \in \mathbb{N}$.\\
+
+\textbf{Base case:} $1^2 = 1^3 = 1$.\\
+
+\textbf{Inductive step:} Let $(1 + 2 + ... + k)^2 = 1^3 + 2^3 + ... + k^3$. Then by algebra:
+
+\begin{align*}
+  (1 + 2 + ... + k + k + 1)^2 &= (1 + 2 + ... + k)(1 + 2 + ... + k + k + 1) + (k + 1)(1 + 2 + ... + k + k + 1)\\
+  &= (1 + 2 + ... + k)^2 + 2(k+1)(1 + 2 + ... + k) + (k + 1)^2\\
+  &= (1 + 2 + ... + k)^2 + 2(k+1)\frac{k(k+1)}{2} + (k + 1)^2\\
+  &= (1 + 2 + ... + k)^2 + k(k+1)^2 + (k + 1)^2\\
+  &= (1 + 2 + ... + k)^2 + (k+1)^3\\
+  &= 1^3 + 2^3 + ... + k^3 + (k+1)^3\\
+\end{align*}
+
+as desired. Thus since S(1) and $S(k) \rightarrow S(k+1)$, the claim S(n) must be true for all $n \in \mathbb{N}$.\\
 
@@
 
-\textbf{Proof}
+\textbf{Proof} First we will prove that set $S = {n | n = 1 \mbox{ or } n, n-1 \in \mathbb{N}}$ is inductive.
+First note that $S$ holds the base case property, since $1 \in S$.
+Also note that if $s = n - 1$, then $s \in S \rightarrow s + 1 \in S$ by the requirement that both $n$ and $n+1$ be in S simultaneously.
+Thus S is an inductive set.\\
 
-Consider set $S = \{n | n = 1 \mbox{ or for } n > 1, n, n-1 \in \mathbb{N}\}$. Note that the inductive property holds for this set, since for $s \in S$, $s+1 \in S$.
+Because set $S$ is inductive, if $n \in \mathbb{N}$ and $n > 1$ then $n \in S$, and also $n-1 \in S$ by definition of membership in $S$. Thus $n > 1 \in \mathbb{N} \rightarrow n-1\in\mathbb{N}$.
 
@@
 \section{Problem 1.1.11a}
-\textbf{WTS} If $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $a+b \in \mathbb{R}\ \mathbb{Q}$.
+\textbf{WTS} If $a \in \mathbb{Q}$ and $b \in \mathbb{R}\setminus \mathbb{Q}$, then $a+b \in \mathbb{R}\setminus \mathbb{Q}$.
 
-\textbf{Proof} In search of a contradiction, assume for $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, that $a+b \in \mathbb{Q}$.
+\textbf{Proof} In search of a contradiction, assume for $a \in \mathbb{Q}$ and $b \in \mathbb{R}\setminus \mathbb{Q}$, that $a+b \in \mathbb{Q}$.
 Then by the definition of rational numbers there exist $m,n \in \mathbb{Z}$ such that $a+b = \frac{m}{n}$.
@@
   &\frac{x}{y}+b = \frac{m}{n}\\
-  &b = \frac{m}{n} - \frac{x}{y}
-  &b = \frac{my}{ny} - \frac{xn}{ny}
+  &b = \frac{m}{n} - \frac{x}{y}\\
+  &b = \frac{my}{ny} - \frac{xn}{ny}\\
   &b = \frac{my-xn}{ny}\\
@@
 This is the desired contradiction.
-Thus, if $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $a+b \in \mathbb{R}\ \mathbb{Q}$.
+Thus, if $a \in \mathbb{Q}$ and $b \in \mathbb{R}\setminus \mathbb{Q}$, then $a+b \in \mathbb{R}\setminus \mathbb{Q}$.
 
 \section{Problem 1.1.11b}
-\textbf{WTS} If $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $ab \in \mathbb{R}\ \mathbb{Q}$.
+\textbf{WTS} If $a \in \mathbb{Q}$ and $b \in \mathbb{R}\setminus \mathbb{Q}$, then $ab \in \mathbb{R}\setminus \mathbb{Q}$.
 
-\textbf{Proof} In search of a contradiction, assume for $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, that $ab \in \mathbb{Q}$.
+\textbf{Proof} In search of a contradiction, assume for $a \in \mathbb{Q}$ and $b \in \mathbb{R}\setminus \mathbb{Q}$, that $ab \in \mathbb{Q}$.
 Then by the definition of rational numbers there exist $m,n \in \mathbb{Z}$ such that $ab = \frac{m}{n}$.
@@
 This is the desired contradiction.
-Thus, if $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $ab \in \mathbb{R}\ \mathbb{Q}$.
+Thus, if $a \in \mathbb{Q}$ and $b \in \mathbb{R}\setminus \mathbb{Q}$, then $ab \in \mathbb{R}\setminus \mathbb{Q}$.
 
@@
 \section{Problem 1.2.4}
-\textbf{A} Set $\{1/n | n \in \mathbb{N}\}$
+\textbf{A} Set $\{1/n | n \in \mathbb{N}\}$\\
 \textit{Maximum} - 1, all $n > 1$ are smaller.\\
@@
 
-\textbf{B} Set $\{x | x \in \mathbb{R} \mbox{ and } x^2 < 2\}$
-\textit{Maximum} - None. For any number $e < \sqrt{2}$, $e + \frac{\sqrt{2}-e}{2}$ is bigger
+\textbf{B} Set $\{x | x \in \mathbb{R} \mbox{ and } x^2 < 2\}$\\
+\textit{Maximum} - None. For any number $e < \sqrt{2}$, $e + \frac{\sqrt{2}-e}{2}$ is bigger\\
 \textit{Minimum} - Same argument with $-\sqrt{2}$.\\
-\textit{Supremum} - $\sqrt{2}$. Anything smaller is in the set, and it isn't in the set.
+\textit{Supremum} - $\sqrt{2}$. Anything smaller is in the set, and it isn't in the set.\\
 \textit{Infimum} - $-\sqrt{2}$ by the same argument.\\

### 2017-02-02 311/latex/HW2.tex

--- 2017-01-27 311/latex/HW1.tex
@@
 
-\title{Advanced Calc Homework \large 1: 1.1: 1,3,5,7,8,11,17, 1.2: 1,4}
+\title{Advanced Calc Homework 2 \large 1.3.4, 1.3.7, 1.3.14, 2.1.2a, 2.1.6, 2.1.7, 2.1.15a}
 \author{Elliott Capek}
@@
 
-\section{Problem 1.1}
+\section{Problem 2.1.2a}
+We want to show $\forall \epsilon,$ $\exists N$ such that $\forall n \geq N,$ $\frac{1}{\sqrt{n}} < \epsilon$ using the Archimedian principle.\\
 
-\section{Problem 1.1.3}
-\textbf{WTS} For any natural number $n$:
+By the Archimedian principle, $\exists c$ such that $\frac{1}{c} < c$.
+Pick a $c$ with this property, and let $N = c^2$.
+Then $\frac{1}{\sqrt{N}} < \epsilon$.
+Note that $\forall n \geq N$, $\frac{1}{\sqrt{n}} \leq \frac{1}{\sqrt{N}}$.
+Thus $\frac{1}{\sqrt{n}} < \epsilon$, as desired.\\
 
-\begin{equation*}
-  \sum_{j=1}^{n} j^2 = \frac{n(n+1)(2n+1)}{6}
-\end{equation*}
+\section{Problem 2.1.7}
+We want to show that $\exists N$ such that $\forall n \geq N$, $a_n > 0$.\\
 
-\textbf{Proof}
-We will prove this using the Principle of Induction.
-We will first show that the statement S(k), where S(k) is the above equality with $n=k$, is true for $S(1)$, then show that $S(k)$ being true implies $S(k+1)$ is true.\\
+We are given that $\{a_n\}$ converges to $a$.
+Consider the $N$ such that $\forall n \geq N$, $\left|a_n - a\right| < a$.
+By Proposition 1.2, $\left|x - a\right| < r$ and $a-r < x < a+r$ are equivalent, thus $0 < a_n < 2a$.
+Thus $a_n > 0$, as desired.\\
 
-\textbf{Base case k=1}
-Notice that for $k=1$:
+\section{Problem 2.1.7}
+We want to show that $\{b_n\}$ converges to $\ell$.\\
+
+We are given that $\exists N$ such that $\forall n \geq N$, $a_n = b_n$. Consider the following algebraic manipulation:
 
 \begin{align*}
-  &\sum_{j=1}^{1} j^2 = 1\\
-  &\frac{1(1+1)(2+1)}{6} = \frac66 = 1\\
+  a_n &= b_n\\
+  a_n - \ell &= b_n - \ell\\
+  \left|a_n - \ell\right| &= \left|b_n - \ell\right|\\
+  c\left|a_n - \ell\right| &= \left|b_n - \ell\right|\hspace{1cm}\mbox{c=1}\\
+  \left|b_n - \ell\right| &\leq c\left|a_n - \ell\right|\\
 \end{align*}
 
-Thus the base case holds, so $S(1)$ is true.\\
-
-\textbf{Inductive step}
-Now we let S(k) be true, so:
-
-\begin{equation*}
-  \sum_{j=1}^{k} j^2 = \frac{k(k+1)(2k+1)}{6}
-\end{equation*}
-
-We want to show S(m) is true, where $m=k+1$. Add $(k+1)^2$ to both sides and do algebra:
-
-\begin{align*}
-  \sum_{j=1}^{k} j^2 + (k+1)^2 &= \frac{k(k+1)(2k+1)}{6} + (k+1)^2\\
-  \sum_{j=1}^{m} j^2 &= \frac{k(k+1)(2k+1)}{6} + k^2 + 2k + 1\\
-  &= \frac{2k^3+3k^2+k}{6} + \frac{6k^2 + 12k + 6}{6}\\
-  &= \frac{2k^3+8k^2+13k+6}{6}\\
-  &= \frac{(k+1)(2k^2+7k+6)}{6}\\
-  &= \frac{(k+1)(k+2)(2k+3)}{6}\\
-  &= \frac{m(m+1)(2m+1)}{6}\\
-\end{align*}
-
-thus $S(k+1)$. We have shown the base case S(1) and inductive step $S(k)$ implies $S(k+1)$, so the statement $S(n)$ must be true for all $n\in \mathbb{N}$.\\
-
-\section{Problem 1.1.5}
-
-\section{Problem 1.1.7}
-\textbf{WTS} $n \in \mathbb{N}$ implies $n-1 \in \mathbb{N}$\\
-
-\textbf{Proof}
-
-Consider set $S = \{n | n = 1 \mbox{ or for } n > 1, n, n-1 \in \mathbb{N}\}$. Note that the inductive property holds for this set, since for $s \in S$, $s+1 \in S$.
-
-\section{Problem 1.1.8}
-\textbf{WTS} For $n,m \in \mathbb{N}$ where $n > m$, $n-m \in \mathbb{N}$.
-
-\textbf{Proof} We will prove this using the Principle of Induction. We will show that the statement $S(k)$ is true for all $k\in\mathbb{N}$, where S(k) is the statement $n-k\in\mathbb{N}$.
-
-\textbf{Base case} The base case S(1) states that for $n\in\mathh{N}$, $n-1\in{mathbb}$. This is true by Problem 1.1.7.\\
-
-\textbf{Inductive step} Statement S(k) says that for $n\in\mathbb{N}$, $n-k\in\mathbb{N}$. Note that, by Problem 1.1.7, $n-k-1\in\mathbb{N}$.\\
-
-Thus both $S(1)$ and $S(k)\rightarrow S(k+1)$ are true. Thus $S(n)$ is true for all $n\in\mathbb{N}$, so for $n,m \in \mathbb{N}$ where $n > m$, $n-m \in \mathbb{N}$.\\
-
-\section{Problem 1.1.11a}
-\textbf{WTS} If $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $a+b \in \mathb{R}\ \mathbb{Q}$.
-
-\textbf{Proof} In search of a contradiction, assume for $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, that $a+b \in \mathbb{Q}$.
-Then by the definition of rational numbers there exist $m,n \in \mathbb{Z}$ such that $a+b = \frac{m}{n}$.
-$a$ can be similarly expanded using $x,y \in \mathbb{Z}: $\frac{x}{y}+b = \frac{m}{n}$.
-We will perform algebra and show this leads to a contradiction:
-
-\begin{align*}
-  &\frac{x}{y}+b = \frac{m}{n}\\
-  &b = \frac{m}{n} - \frac{x}{y}
-  &b = \frac{my}{ny} - \frac{xn}{ny}
-  &b = \frac{my-xn}{ny}\\
-\end{align*}
-
-Because the product and difference of two integers is also an integer, this means b can be expressed as a fraction of two integers.
-This is the desired contradiction.
-Thus, if $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $a+b \in \mathb{R}\ \mathbb{Q}$.
-
-\section{Problem 1.1.11b}
-\textbf{WTS} If $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $ab \in \mathb{R}\ \mathbb{Q}$.
-
-\textbf{Proof} In search of a contradiction, assume for $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, that $ab \in \mathbb{Q}$.
-Then by the definition of rational numbers there exist $m,n \in \mathbb{Z}$ such that $ab = \frac{m}{n}$.
-$a$ can be similarly expanded using $x,y \in \mathbb{Z}: $\frac{x}{y}b = \frac{m}{n}$.
-This can be rewritten as $b = \frac{my}{nx}$.
-Because the product of two integers is itself an integer, we have expressed b as a fraction of two integers.
-This is the desired contradiction.
-Thus, if $a \in \mathbb{Q}$ and $b \in \mathbb{R}\ \mathbb{Q}$, then $ab \in \mathb{R}\ \mathbb{Q}$.
-
-\section{Problem 1.1.17}
-Consider set $S = \{x | x \in \mathbb{R}, x \geq 0, x^2 < c\}$.
-
-\textbf{A}\\
-We wish to show that $\forall s \in S$, $s \leq c + 1$.
-Consider $s \in S$.
-If $s > 1$ then, by multiplying by s, $s^2 > s$.
-Thus by the definition of S, $s < s^2 < c$ and so $s < c+1$.
-If $s \leq 1$, then since $c$ must be positive, $c + 1 > 1$ and so $s < c + 1$.\\
-
-\textbf{B}\\
+Thus, by the Comparison Lemma, $\{b_n\}$ converges to $\ell$ as desired.\\
 

### 2017-02-02 311/latex/HW2.tex

@@
 
+\section{Problem 1.3.14}
+We want to show the relation $ab \leq \frac12\left(a^2+b^2\right)$. for any $a,b\in\mathbb{R}$\\
+
+Note that since the square of any real number is nonnegative, $\left(a-b\right) \geq 0$. Squaring both sides yields $a^2 + b^2 - 2ab \geq 0$, and so $ab \leq \frac12\left(a^2+b^2\right$, as desired.\\
+
 \section{Problem 2.1.2a}
@@
 
-\section{Problem 2.1.7}
+\section{Problem 2.1.6}
 We want to show that $\exists N$ such that $\forall n \geq N$, $a_n > 0$.\\

### 2017-02-12 311/latex/HW2.tex
Add to resume.pdf.

@@
 
-\title{Advanced Calc Homework 2 \large 1.3.4, 1.3.7, 1.3.14, 2.1.2a, 2.1.6, 2.1.7, 2.1.15a}
+\title{Advanced Calc Homework 2\\ \large 1.3.4, 1.3.7, 1.3.14, 2.1.2a, 2.1.6, 2.1.7, 2.1.15a}
 \author{Elliott Capek}
@@
 
+\section{Problem 1.3.2}
+For $a > 0$, we want to show that $|x - a| < \frac a2$ then $x > \frac a2$.\\
+
+Let $|x - a| < \frac a2$. Then by Proposition 1.2 this is equivalent to statement $a - \frac a2 < x < a + \frac{3a}{2}$, thus $x > \frac a2$, as desired.\\
+
+\section{Problem 1.3.7}
+Starting with statement $(a+b) + (-b)$, apply the triangle inequality to get $|a+b| + |-b| \geq |a + b -b| = |a|$. By rearranging terms, we get $|a| - |b| \leq |a+b|$.\\
+
+Now interchanging $a$ and $b$ yields $|b| - |a| \leq |b+a|$. By algebra, $-\left(|a| - |b|\right) \leq |a+b|$. Then $\left||a| - |b|\right| \leq |a+b|$.\\
+
 \section{Problem 1.3.14}
@@
 
-Note that since the square of any real number is nonnegative, $\left(a-b\right) \geq 0$. Squaring both sides yields $a^2 + b^2 - 2ab \geq 0$, and so $ab \leq \frac12\left(a^2+b^2\right$, as desired.\\
+Note that since the square of any real number is nonnegative, $\left(a-b\right) \geq 0$.
+Squaring both sides yields $a^2 + b^2 - 2ab \geq 0$, and so $ab \leq \frac12\left(a^2+b^2\right)$, as desired.\\
 
@@
 
+\newpage
+
 \section{Problem 2.1.7}
@@
 
+\section{Problem 2.1.15a}
+We want to show the polynomial $a_n = n^3 - 4n^2 - 100n$ converge to infinity, or that $\forall c \in \mathbb{R}, c> 0$, $\exists N$ such that $\forall n \geq N$, $a_n > c$.\\
+
+Consider $a_n$ and its derivatives:\\
+
+\begin{align*}
+  a_n &= n^3 - 4n^2 - 100n\\
+  a_n' &= 3n^2 - 8n - 100\\
+  a_n'' &= 6n - 8\\
+  a_n''' &= 6\\
+  a_n'''' &= 0
+\end{align*}
+
+Notice that, at $n=100$, it is the case that $a_n >0$, $a_n' > 0$, $a_n'' > 0$, $a_n''' > 0$ and $a_n'''' = 0$.
+Because the third derivative is a constant positive number, $a_n$ where $n > 100$ will always increase at an increasing rate.
+Thus for any $c < 100$, $N = 100$ satisfies the claim.\\
+
+Now consider for any $c > 100$, because $a_n$ is increasing at a steadily increasing rate, there must exist some $n'$ such that $a_{n'} = c$.
+Now for all $n > n'$, $a_n > c$, so the claim is also satisfied for $c > 100$.
+Thus $a_n$ converges at infinity.\\
+
 \end{document}

### 2017-02-12 311/latex/HW3.tex
HW3.tex

--- 2017-02-02 311/latex/HW2.tex
@@
 
-\title{Advanced Calc Homework 2 \large 1.3.4, 1.3.7, 1.3.14, 2.1.2a, 2.1.6, 2.1.7, 2.1.15a}
+\title{Advanced Calc Homework 3}
 \author{Elliott Capek}
@@
 
-\section{Problem 2.1.2a}
-We want to show $\forall \epsilon,$ $\exists N$ such that $\forall n \geq N,$ $\frac{1}{\sqrt{n}} < \epsilon$ using the Archimedian principle.\\
+\section{Problem 2.2.2}
 
-By the Archimedian principle, $\exists c$ such that $\frac{1}{c} < c$.
-Pick a $c$ with this property, and let $N = c^2$.
-Then $\frac{1}{\sqrt{N}} < \epsilon$.
-Note that $\forall n \geq N$, $\frac{1}{\sqrt{n}} \leq \frac{1}{\sqrt{N}}$.
-Thus $\frac{1}{\sqrt{n}} < \epsilon$, as desired.\\
+\section{Problem 2.4.4}
+\textbf{a.} $\{\frac{1}{n}\}$\\
+Peak at $n=1$. $\forall m \in \mathbb{N}$, $\frac{1}{n} \geq \frac{1}{n+m}$. Thus $a_n \geq a_{n+m}$.\\
 
-\section{Problem 2.1.7}
-We want to show that $\exists N$ such that $\forall n \geq N$, $a_n > 0$.\\
+\textbf{b.} $\{\left(01\right)^n\}$\\
+Peak at even $n$s. For any even $n$, $a_n = 1$ and $\forall m \in \mathbb{N}$, $a_{n+m} = 1$ or $-1$. Thus, $a_n \geq a_{n+m}$.\\
 
-We are given that $\{a_n\}$ converges to $a$.
-Consider the $N$ such that $\forall n \geq N$, $\left|a_n - a\right| < a$.
-By Proposition 1.2, $\left|x - a\right| < r$ and $a-r < x < a+r$ are equivalent, thus $0 < a_n < 2a$.
-Thus $a_n > 0$, as desired.\\
+\textbf{c.} $\{(-1)^nn\}$\\
+No peaks. $\forall n \in \mathbb{N}$, if $n$ is even then $a_n = n < n+2 = a_{n+2}$. If $n$ is odd, then $a_n = -n < n+1 = a_{n+1}$.\\
 
-\section{Problem 2.1.7}
-We want to show that $\{b_n\}$ converges to $\ell$.\\
+\textbf{d.} $\{\frac{(-1)^n}{n}\}$\\
+Peak at $n=2$. $\forall m \in \mathbb{N}$, if $m$ is odd then $a_{2+m} < 0 < a_{2}$. If $m$ is even then $a_{n+m} = \frac{1}{2+m} < \frac{1}{2}$.\\
 
-We are given that $\exists N$ such that $\forall n \geq N$, $a_n = b_n$. Consider the following algebraic manipulation:
-
-\begin{align*}
-  a_n &= b_n\\
-  a_n - \ell &= b_n - \ell\\
-  \left|a_n - \ell\right| &= \left|b_n - \ell\right|\\
-  c\left|a_n - \ell\right| &= \left|b_n - \ell\right|\hspace{1cm}\mbox{c=1}\\
-  \left|b_n - \ell\right| &\leq c\left|a_n - \ell\right|\\
-\end{align*}
-
-Thus, by the Comparison Lemma, $\{b_n\}$ converges to $\ell$ as desired.\\
+\section{Problem 2.4.7}
 

### 2017-02-28 311/latex/HW4.tex
bla

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\nonstopmode

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{amsmath}

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\usepackage{amssymb}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Advanced Calc Homework 4}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem 3.1.1}
\textbf{a.)}\\
False. Consider non-continuous functions $f: \mathbb{R} \rightarrow \mathbb{R}$ and $g = -f$. Then $f(x) + g(x) = 0$, and so $f+g$ is continuous, but $f$ is not.\\

\textbf{b.)}\\
Skip

\textbf{c.)}\\
Skip

\textbf{d.)}\\
Skip

\section{Problem 3.1.7}
Consider the continuous function $f[0,1] \rightarrow \mathbb{R}$, where $f(x) \geq 2$ if $[0,1)$. In search of a contradiction, let $f(1) < 2$. By $f$'s continuity, $\forall \epsilon > 0$ and $x_0 \in \mathbb{R}$, $\exists \delta$ such that $|x - x_0| < \delta$ implies $|f(x) - f(x_0)| < \epsilon$. Consider $x_0 = 1$ and $\epsilon = 2 - f(x_0)$. By our assumption, $\epsilon > 0$. Consider $x' < 1$ such that $|x' - 1| < \delta$. Thus, $|f(x) - f(x_0)| < 2 - f(x_0)$. Thus $f(x_0) - 2 + f(x_0) < f(x) < f(x_0) + 2 - f(x_0) = 2$, so $f(x) < 2$. This is a contradiction, thus $f(1) \geq 2$.\\

\section{Problem 3.1.13}
Let $\{x_n\}$ be a sequence such that $x_n \in D$ and $\lim_{n\rightarrow\infty} x_n = x_0$ for $x_0 \in D$. Let $f: D \rightarrow \mathbb{R}$ be a Lipschitz function. Then by $f$'s Lipschitz-ness, $|f(x_n) - f(x_0)| \leq C|x_n - x_0|$ for some $C \geq 0$. Thus by the Comparison Lemma $\lim_{n\rightarrow\infty} f(x_n) = f(x_0)$. Thus we have $\lim_{n\rightarrow\infty} x_n = x_0$ implying $\lim_{n\rightarrow\infty} f(x_n) = f(x_0)$ for a generic $\{x_n\}$ and $x_0$, and so $f$ is continuous.\\

\section{Problem 3.5.3}
For each point $x_0$, consider $\delta = \min\left(1,\frac{\epsilon}{3|x_0|^2+3|x_0|+1}\right)$. Assume $|x - x_0| \leq \delta$. Then consider $|x^3-x_0^3| = |x-x_0||x^2+xx_0+x_0^2|$. By our $\delta$ assumption $|x-x_0| < 1$, which unwraps to $|x| < |x_0|+1$. Thus $|x^3-x_0^3| < \left(|x_0|^2 + 2|x_0| + 1 + |x_0|^2 + |x_0| + |x_0|^2\right)|x-x_0| = \left(3|x_0|^2 + 3|x_0| + 1\right)|x-x_0|$. Also by our assumption $|x-x_0| < \frac{\epsilon}{3|x_0|^2+3|x_0|+1}$, thus $|x^3-x_0^3| < \epsilon$. Thus $f$ is continuous for all $x_0$.\\

\section{Problem 3.5.4}
In search of a contradiction, let $f$ be continuous at $x_0 = \frac{3}{4}$. Then for $\epsilon = \frac{1}{8}$ there must exist some $\delta$ such that $|x - \frac{3}{4}| < \delta$ implies $|f(x)-f(\frac{3}{4})| < \frac{1}{8}$. Consider $\frac{3}{4} + \delta > x > \frac{3}{4}$. Then $|x-\frac{3}{4}|<\delta$, which implies $|2-\frac{7}{4}| = \frac{1}{4} < \frac{1}{8}$, which is a contradiction. Thus $f$ is not continuous at $x_0 = \frac{3}{4}$.\\

\section{Problem 3.5.7}
\textbf{a.)} Consider $\delta = \epsilon\sqrt{x_0}$. Then $|x-x_0| < \epsilon\sqrt{x_0}$, so $\frac{|x-x_0|}{\sqrt{x_0}} < \epsilon$. Because $x$ is constrained to be positive $|\sqrt{x_0}| < |\sqrt{x} + \sqrt{x_0}|$, and so $\frac{|x-x_0|}{|\sqrt{x}+\sqrt{x_0}|} = |\sqrt{x}-\sqrt{x_0}| < \epsilon$, as desired. Thus $f$ is continuous for positive $x$.\\
\textbf{b.)} Skip

\textbf{c.)} Skip

\end{document}

### 2017-03-06 311/latex/HW5.tex
Add 311 HW 5

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\nonstopmode

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{amsmath}

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\usepackage{amssymb}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Advanced Calc Homework 5}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem 3.2.1}
\textbf{a.) Every function $f:[0,1] \rightarrow \mathbb{R}$ has a maximum.}\\
False. Consider non-continuous function $f:[0,1]\rightarrow\mathbb{R}, f(x) = \frac{1}{x-0.5}$. This function has no maximum.\\

\textbf{b.) Every continuous function $f:[a,b] \rightarrow \mathbb{R}$ has a minimum}\\
True. Because $[a,b]$ is bounded, it is subsequentially compact. Thus by the extreme value theorem, f has a max and min.\\

\textbf{c.) Every continuous function $f:(0,1) \rightarrow \mathbb{R}$ has a maximum}\\
False. Consider function $f:(0,1) \rightarrow \mathbb{R}, f(x) = \frac{1}{x-1}$. This function has no max.\\

\textbf{d.) Every continuous function $f:(0,1) \rightarrow \mathbb{R}$ has a bounded range}\\
False. Consider function $f:(0,1) \rightarrow \mathbb{R}, f(x) = \frac{1}{x-1}$. This function has no upper bound, and is therefore not bounded.\\

\textbf{d.) If the image of the continuous function $f:(0,1) \rightarrow \mathbb{R}$ is bounded below, then the function has a minimum.}\\
False. Consider function $f:(0,1) \rightarrow \mathbb{R}, f(x) = e^x$. This function's image is bounded from below by $1$, but has no minimum.\\

\section{Problem 3.2.3}
\textbf{a.) Find a function $f:(a,b)\rightarrow\mathbb{R}$ which is continuous and whose image is unbounded above.}\\
Consider $f(x) = \frac{1}{x-a}$. This function is unbounded above and continuous on its domain.\\

\textbf{b.) Find a function $f:(a,b)\rightarrow\mathbb{R}$ which is continuous and bounded from above but which does not have a max.}
Consider $f(x) = a-x$. This function is bounded from above by zero and has no maximum.\\

\section{Problem 3.3.1}
\textbf{a.) If $f:\mathbb{R}\rightarrow\mathbb{R}$ is continuous, then $f(\mathbb{R}) = \mathbb{R}$.}\\
False. Consider $f(x) = x^2$. This function never reaches negative reals.\\

\textbf{b.) For any $f:[0,1]\rightarrow\mathbb{R}$, $f([0,1])$ is an interval.}\\
False. Consider\\

\[
f(x) =
\begin{cases} 
      -1 & x\leq 0 \\
      1  & x > 0 \\
   \end{cases}
\]

$\{0,1\}$ is not an interval, since $\frac12$ is not a member.\\

\textbf{c.) For any continuous $f:D\rightarrow\mathbb{R}$, $f(D)$ is an interval.}\\
True. By the intermediate value theorem, for any continuous function on a domain, any value between to elements in the image has a preimage within the domain. Thus any value is allowed within the bounds of the image. Thus the image is an interval.\\

\textbf{d.) For a continuous, strictly increasing function $f[0,1]\rightarrow\mathbb{R}$, its image is the interval $[f(0), f(1)]$.}\\
By the intermediate value theorem, any element $c \in (f(0), f(1))$ must have some element $c = f(x)$ where $x \in (0,1)$ Thus every value between f(0) and f(1) is hit by the function. Because of f's monotonicity, all $x$'s in the domain must be $f(x) > f(0)$, and all elements must be $f(x) < 1$. Thus the image equals the interval $[f(0), f(1)]$.\\

\section{Problem 3.3.2}
Note that $f(-2)=-504$ and $f(2)=520$. For the moment restrict $f$'s domain to $[-2,2]$. Thus by the Intermediate Value Theorem, there must exist some $x \in (-2, 2)$ such that $f(x) = 0$. Thus there must exist soem solution to the equation.\\

\section{Problem 3.3.4}
Consider point $c = f(x) \in \left(f(-1),f(1)\right)$. By the intermediate value theorem, because the function has closed bounds of $[-1,1]$ and a bounded image, there must exist some x such that $f(x) = y$. Thus there must exist a fixed point in the function.\\

\section{Problem 3.4.1}
\textbf{a.) All continuous functions are uniformly continuous.}
False. Consider $f(x) = x^1$. It is continuous, but not uniformly continuous.\\

\textbf{b.) All continuous functions on $[0,1)$ are uniformly continuous.}\\
False. Consider $f(x) = 1/x$. This is continuous on $[0,1)$, but not uniformly continuous. The function gets steeper and steeper towards 1, so there is no maximum slope.\\

\textbf{c.) All continuous functions on $[0,1]$ are uniformly continuous.}
False. Consider $f(x) = \sin(\frac{1}{x-\frac12})$. This function attains every possible slope, but is still continuous.\\

\textbf{d.) Are all uniformly continuous functions continuous?}
True, uniform continuity implies continuity. If there exists a $\delta$ such that for all $\epsilon>0$ and $x_0$, $|x-x_0|<\delta$ implies $|f(x)-f(x_0)|<\epsilon$, then this $\delta$ works for the normal continuity case for a specific $x_0$ as well.\\

\section{Problem 3.4.8}
Consider the function $f:(a,b)\rightarrow\mathbb{R}, f(x) = \frac{1}{x-b}$. This function is continuous, but not uniformly continuous.\\

\section{Problem 3.4.11}
Let $f$ be a Lipschitz function with constant $C$. Then consider the case where $|u-v|<\frac{\epsilon}{C}$ for some $\epsilon>0$ and arbitrary $u,v$. Then $C|u-v|<\epsilon$, and by the Lipschitz condition $|f(u)-f(v)|<\epsilon$. Thus $f$ is uniformly continuous.\\

\section{Problem 3.6.1}
\textbf{a.) Monotone functions are one-to-one.}\\
False. Consider function

\[
f(x) =
\begin{cases} 
      -1 & x\leq 0 \\
      1  & x > 0 \\
   \end{cases}
\]

notice that all $x\leq0$ map to $-1$, thus $f$ is not one-to-one.\\

\textbf{b.) A strictly increasing function is one-to-one.}\\
True. For any $x_1, x_2 \in \mathbb{R}$, if $x_1 \neq x_2$ then either $f(x_1) > f(x_2)$ or $f(x_1) < f(x_2)$. Thus the function is one-to-one.\\

\textbf{c.) A strictly increasing function is continuous.}\\
False. Consider the function

\[
f(x) =
\begin{cases} 
      x-1 & x\leq 0 \\
      x+1  & x > 0 \\
   \end{cases}
\]

note that the image $f(\mathbb{R})$ does not contain 0. Thus, $f$ is not one-to-one.\\

\textbf{d.) A one-to-one function is monotone}
False. Consider

\[
f(x) =
\begin{cases} 
      -x & |x|\leq 10 \\
      x  & |x| > 10 \\
   \end{cases}
\]

The preimage equals the real numbers, but the function is not monotone.\\

\section{Problem 3.6.2}
\textbf{a.) Find a continuous function $f:(0,1)\rightarrow\mathbb{R}$ with an image equal to $\mathbb{R}$.}\\
Consider function $f(x) = \tan((x-\frac{1}{2})*\pi)$. This hits all points in $\mathbb{R}$ when restricted to a domain $D = (0,1)$.\\

\textbf{b.) Find a continuous function $f:(0,1)\rightarrow\mathbb{R}$ with an image equal to $[0,1]$}\\
Consider function $f(x) = \sin(2*x*\pi)$. This hits all points in $[0,1]$ when restricted to a domain $D = (0,1)$.\\

\textbf{c.) Find a continuous function $f:\mathbb{R}\rightarrow\mathbb{R}$, strictly increasing, with an image equal to $(-1,1)$}\\
Consider function $f(x) = \frac{2}{\pi}\arctan(x)$. It is continuous and maps the real numbers to a value between $(-1,1)$.\\

\section{Problem 3.7.2}
\textbf{a.)} Consider the following manipulation:

\begin{align*}
  \frac{x^4}{x-1} &= \frac{(x-1)(x+1)(x^2+1)}{x-1}\\
  &= (x+1)(x^2+1)\\
  \lim_{x \rightarrow 1} (x+1)(x^2+1) = 4\\
\end{align*}

\textbf{a.)} Consider the following manipulation:

\begin{align*}
  \frac{\sqrt{x}-1}{x-1} &= \frac{(\sqrt{x}-1)(\sqrt{x}+1)}{(x-1)(\sqrt{x}+1)}
  &= \frac{x-1}{(x-1)(\sqrt{x}+1)}\\
  &= \frac{1}{\sqrt{x}+1} = \frac{1}{2}
  \lim_{x \rightarrow 1} \frac{\sqrt{x}-1}{x-1} = \frac{1}{2}\\
\end{align*}

\section{Problem 3.7.5}
In search of a contradiction, let $D$ have a limit point $s$. If $s = x_0$, then there can exist no sequence with elements in $x_n \in D, x_n \neq x_0$, since $x_0$ is the only element of $D$. This is a contradiction. If $s \neq x_0$, then $\{x_n\}, x_n = x_0$ is the only possible subsequence which meets the limit point criteria. Consider $\epsilon = \left|\frac{x_0 - s}{2}\right|$. Because $|x_0 - s| > \left|\frac{x_n - s}{2}\right|$, this sequence does not converge to $x_0$ either, a contradiction. Thus $D$ has no limit points.\\

Now consider the set of natural numbers $\mathbb{N}$. In search of a contradiction, consider arbitrary limit point $s \in \mathbb{R}$. First, consider the case where $s$ is not a natural number. Consider arbitrary natural number sequence $\{x_n\}$. Consider $\epsilon = \min\left(|\lceil{s}\rceil-s|, |\lfloor{s}\rfloor-s|\right)$. No integer sequence can get closer than this $\epsilon$ to $s$; a contradiction. Now consider $s$ a natural number. Now consider an integer sequence $\{x_n\}$ where $x_n \neq s \forall n\in\mathbb{N}$. Thus for $\epsilon=\frac12$, there are no $|x_n-s|<\epsilon$, a contradiction. Thus $\mathbb{N}$ has no limit points.\\

\section{Problem 3.7.6}
Because D is a nonempty bounded subset of $\mathbb{R}$, it has a supremum $s=Sup(D)$. Consider sequence $\{x_n\}, x_n = s-\frac{1}{n}$. Note that each element is in D, since it is less than the supremum. Note that each element does not equal $s$. Thus we have found a sequence which is a subset of D which converges to $s$, but never equals it. Thus $s$ is a limit point of $D$.\\

\end{document}

### 2017-03-06 311/latex/HW4-lastproblems.tex

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\nonstopmode

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{amsmath}

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\usepackage{amssymb}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Advanced Calc Homework 4}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem 3.1.1}
\textbf{a.)} No. Consider discontinuous $f$, and $g=-f$. Then $f+g$ is continuous, but $f$ isn't.\\

\textbf{b.)} No. Consider $f=\sqrt{x}$. Then $f^2$ is continuous, but on $\mathbb{R}$ f is not continuous.\\

\textbf{c.)} \\

\textbf{d.)} \\

\section{Problem 3.5.3}
We want to show that $x^3$ is a continuous function. We will do algebra on the $|x^3-x_0^3|$ statement and try to show that $|x-x_0|<\delta$ implies $|x^3-x_0^3|<\epsilon$ for some $\epsilon, x_0$. Consider the following.\\

$|x^3-x_0^3| = |x-x_0||x^2+xx_0+x_0^2|$. Then impose that $|x-x_0|<1$, thus $x<x_0+1$. Substituting this in:

\begin{align*}
  |x^3-x_0^3| &= |x-x_0||x^2+xx_0+x_0^2|\\
  &\leq |x-x_0||(x_0+1)^2+x_0(x_0+1)+x_0^2|\\
  &= |x-x_0||3x_0^2+3x_0+1|\\
\end{align*}

Thus if we let $\delta = \min(1,\frac{\epsilon}{|3x_0^2+3x_0+1|})$ then $|x-x_0|\leq\delta$ implies $|x^3-x_0^3|<\epsilon$. Thus $x^3$ is continuous.\\

\section{Problem 3.5.7}
\textbf{a.)} Want to show $\sqrt{x}$ is continuous. Consider arbitrary $x_0 \in \mathbb{R}$ and $\epsilon>0 \in \mathbb{R}$. Then

\begin{align*}
  |\sqrt{x} - \sqrt{x_0}| = \frac{|\sqrt{x} - \sqrt{x_0}||\sqrt{x} + \sqrt{x_0}|}{|\sqrt{x} + \sqrt{x_0}|}\\
  &= \frac{|x-x_0|}{|\sqrt{x} + \sqrt{x_0}|} < \frac{|x-x_0|}{|\sqrt{x_0}|}
\end{align*}

Thus if $|x-c|<\delta=\sqrt{x_0}\epsilon$, then $|\sqrt{x}-\sqrt{c}|$. Thus $\sqrt{x}$ is continuous.\\

\textbf{b.)} Want to show for some $\epsilon$ there exists a $\delta$ such that $\forall x,y >0 \in \mathbb{R}$, $|x-y|<\delta$ implies $\sqrt{x}-\sqrt{y}|<\epsilon$.\\

Consider $\delta = \epsilon^2$. Note that because $x,y>0$, $|\sqrt{x}-\sqrt{y}|\leq|\sqrt{x}+\sqrt{y}|$. Then

\begin{align*}
  |\sqrt{x}-\sqrt{y}|^2 \leq |\sqrt{x}-\sqrt{y}||\sqrt{x}+\sqrt{y}| &= |x-y| < \epsilon^2
\end{align*}

Because absolute values and $\epsilon$s are nonnegative, $|\sqrt{x}-\sqrt{y}|^2 < \epsilon^2$ implies $|\sqrt{x}-\sqrt{y}|<\epsilon$. Thus we have found a $\delta$ for each $\epsilon$, and so $\sqrt{x}$ is uniformly continuous in $[0,1]$.

\textbf{c.)} We want to show $\sqrt{x}:[0,1]\rightarrow\mathbb{R}$ is not a Lipschitz function. In search of a contradiction, let $\sqrt{x}$ be a Lipschitz function. Then there must exist some $C$ such that $\forall x,y \in D$, $|\sqrt{x}-\sqrt{y}|<C|x-y|$. Consider the following manipulation:

\begin{align*}
  |\sqrt{x}-\sqrt{y}|&<C|x-y|\\
  \frac{|\sqrt{x}-\sqrt{y}|}{|x-y|}&<C\\
  \frac{|\sqrt{x}-\sqrt{y}|}{|\sqrt{x}-\sqrt{y}||\sqrt{x}+\sqrt{y}|}&<C\\
  \frac{1}{|\sqrt{x}+\sqrt{y}|}&<C\\
\end{align*}

This is a contradiction, since for any $C$, there will always be some $x,y$ such that  $\frac{1}{\sqrt{x}+\sqrt{y}} > C$. Thus $\sqrt{x}:[0,1]\rightarrow\mathbb{R}$ is not a Lipschitz function.\\

\end{document}

### 2017-03-17 311/latex/HW6.tex
Add new hw.

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\nonstopmode

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{amsmath}

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\usepackage{amssymb}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Advanced Calc Homework 5}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem 6.4.2}
In search of a contradiction, say $f(x)$ is integrable. Then $\forall \epsilon > 0$ $\exists \delta>0$ such that for an arbitrary $\{x_i^1\}$ and $\{x_i^2\}$, $width(\{x_i^1\})<\delta$ and $width(\{x_i^2\})<\delta$ implies $|S_1-S_2|<\epsilon$.\\

Consider a regular partition $\{x_i\}$ such that $x_i = \frac{i}{N}$ for some $N$. Define sequences $\{x_i'\}$ and $\{x_i''\}$ such that $x_i \leq x_i' \leq x_{i+1}$ and $x_i \leq x_i'' \leq x_{i+1}$, and $S_1 = \sum_{i=1}^{i=N}f(x_i')\left(x_i-x_{i-1}\right)$ and $S_2 = \sum_{i=1}^{i=N}f(x_i'')\left(x_i-x_{i-1}\right)$. Note that:

\begin{align*}
  S_1 &= \sum_{i=1}^{i=N}f(x_i')(x_i-x_{i-1}) = \frac{1}{N}\sum_{i=1}^{i=N}x_i' \geq \frac{1}{N}\sum_{i=1}^{i=N}x_i\\
  &= \frac{1}{N^2}\sum_{i=1}^{i=N}i = \frac{N(N+1)}{2N^2} = \frac{1}{2} + \frac{1}{2N}\\
  S_2 &= \sum_{i=1}^{i=N}f(x_i'')(x_i-x_{i-1}) = \frac{-1}{N}\sum_{i=1}^{i=N}x_i'' \leq \frac{-1}{N}\sum_{i=1}^{i=N}x_i\\
  &= \frac{-1}{N^2}\sum_{i=1}^{i=N}i = \frac{-N(N+1)}{2N^2} = \frac{-1}{2} + \frac{-1}{2N}\\
\end{align*}

Thus $S_1 \geq \frac{1}{2} + \frac{1}{2N}$ and $S_2 \leq \frac{-1}{2} + \frac{-1}{2N}$, so $|S_1-S_2| = |1 + \frac{1}{N}|$. Note that the width of $\{x_i\}$ is given by $\frac{1}{N}$. Thus $\forall \epsilon>0$, $\exists \delta > 0$ such that $\frac{1}{N}<\delta$ implies $|S_1-S_2|<\epsilon$ for our chosen $S_1$ and $S_2$. Consider $\epsilon = \frac{1}{2}$, and the associated $\delta$. Consider $N$ such that $\frac{1}{N}<\delta$. Note that $|S_1-S_2|>1>\frac{1}{2}$, a contradiction. Thus $f(x)$ is not integrable.\\

\section{Problem 6.4.9}
%% Let $\delta=\frac{\epsilon}{2}$ and f, g be integrable functions. Then for all $\epsilon > 0$ there exist $\delta_f, \delta_g$ such that for partitions $\{x_i^f\}$ and $\{x_i^g\}$ where $S_f$ and $S_g$ are the Riemann sums using these partitions, then $|S_1 - \int_a^bf(x)dx|<\epsilon$ and $|S_2 - \int_a^bg(x)dx|<\epsilon$.

By Theorem 3.5 because $f:[a,b]\rightarrow\mathbb{R}$ and $g:[a,b]\rightarrow\mathbb{R}$ are continuous, they are both integrable. By the triangle inequality, $\forall x, |f(x) + g(x)| \leq |f(x) + |g(x)|$. By Corollary 2.3, $\int_a^b |f(x)+g(x)| \leq \int_a^b |f(x)|+|g(x)| = \int_a^b |f(x)| + \int_a^b |g(x)|$, as desired.\\

\section{Problem 3}
Because $f(x) = x$ is integrable, then for all $\epsilon > 0$, it is the case that for some $\delta > 0$, for all $\{x_i\}$, $max(x_i-x_{i-1})<\delta$, $|S - A| < \epsilon$. Consider $\{x_i\}$ such that $x_i = a + \frac{b-a}{N}i$ for an arbitrary $N$. Then $max(x_i-x_{i-1}) = \frac{b-a}{N}$. Consider the following algebraic manipulation, where $x_i' = x_i$ and $x_i$ is a regular partition and $a=0, b=1$:

\begin{align*}
  S &= \sum_{i=1}{i=N}f(x_i')(x_{i+1}-x_i)\\
  &= \sum_{i=1}{i=N}\left(a+\frac{b-a}{N}i\right)\left(a+\frac{b-a}{N}i - a -\frac{b-a}{N}(i-1)\right)\\
  &= \sum_{i=1}{i=N}\left(\frac{i}{N}\right)\left(\frac{1}{N}\right)\\
  &= \frac{1}{N^2}\sum_{i=1}{i=N}i\\
  &= \frac{1}{N^2}\frac{N(N+1)}{2} = \frac{1}{2} + \frac{1}{2N}\\
\end{align*}

Then $\lim_{N\rightarrow0} max(x_i-x_{i-1}) = 0$, so $\lim_{width(\{x_i\})} S = \frac{1}{2}$. Then $\int_0^1xdx = \frac{1}{2}$.\\

%% Now consider $A = \frac{1}{2}$. Then $|S - A| = |\frac{1}{2N}|$. Now for some arbitrary $\epsilon$ by the Archimedean Principle there exists some $N'$ such that $\frac{1}{2N'} < \epsilon$. Because $N$ is arbitrary, consider $N = N'$. Then for any $\delta > \frac{b - a}{N}$, $||$

\section{Problem 4}
We want to show that $f$ is integrable. That is, for all $\epsilon>0$ there exist some step functions $f_1:[a,b]\rightarrow\mathbb{R}$ and $f_2:[a,b]\rightarrow\mathbb{R}$ where $f_1(x) \leq f(x) \leq f_2(x)$ and $\int_a^b f_1(x) - f_2(x)dx \leq \epsilon$ for all $\epsilon$. We will do this by defining a partition and step functions such that the difference between $f_2$ and $f_1$ is always less than $\frac{\epsilon}{N}$, then showing that the integral of $f_1-f_2$ is zero. This proof works for a monotone increasing function, but would work equally well for a monotone decreasing function with minimal switching.\\

Define a partition $\{x_i\}$ on $[a,b]$. Define step functions $f_1$ and $f_2$ such that for each $x' \in [a,b]$ where $x_i \leq x' \leq x_{i+1}$, $f_1(x') = f(x_i)$ and $f_2(x') = f(x_{i+1})$. Define the partition $\{x_i\}$ such that $x_0 = a$ and $x_i$ is the value in the domain such that $f(x_i) = f(x_{i-1})+\frac{\epsilon}{N}$ for some $N$ and $\epsilon$, OR $b$, whichever is smaller. Note that this value must necessarily exist by the IVT, because $f$ is continuous on $[a,b]$ and $f(x_{i-1}) \leq f(x_i) \leq f(b)$ by f's positive monotonicity. Notice that $f(x_{i+1}) - f(x_i) \leq \frac{\epsilon}{N}$. Consider the following Riemann sum for $x_i \leq x'_i \leq x_{i+1}$:

\begin{align*}
  S &= \sum_{i=1}^{i=N} \left(f_1(x'_i) - f_2(x'_i)\right)\left(x_i-x_{i-1}\right)\\
  &\leq \sum_{i=1}^{i=N} \frac{\epsilon}{N}\left(x_i-x_{i-1}\right)\\
  &= \frac{\epsilon(b-a)}{N}\\
\end{align*}

Thus for arbitrary $\epsilon>0$ and $N > (b-a)$, $|S| < \epsilon$ and so $|S-0|<\epsilon$. Thus we can say $\int_a^bf_1(x)-f_2(x)dx = 0$. By extension $\int_a^bf_1(x)-f_2(x)dx < \epsilon$ for all $\epsilon>0$.

\end{document}

### 2017-03-17 311/latex/HW6.tex
more stuff.

@@
 
-\section{Problem 6.4.9}
-%% Let $\delta=\frac{\epsilon}{2}$ and f, g be integrable functions.
-Then for all $\epsilon > 0$ there exist $\delta_f, \delta_g$ such that for partitions $\{x_i^f\}$ and $\{x_i^g\}$ where $S_f$ and $S_g$ are the Riemann sums using these partitions, then $|S_1 - \int_a^bf(x)dx|<\epsilon$ and $|S_2 - \int_a^bg(x)dx|<\epsilon$.
-
-By Theorem 3.5 because $f:[a,b]\rightarrow\mathbb{R}$ and $g:[a,b]\rightarrow\mathbb{R}$ are continuous, they are both integrable.
-By the triangle inequality, $\forall x, |f(x) + g(x)| \leq |f(x) + |g(x)|$.
-By Corollary 2.3, $\int_a^b |f(x)+g(x)| \leq \int_a^b |f(x)|+|g(x)| = \int_a^b |f(x)| + \int_a^b |g(x)|$, as desired.\\
-
 \section{Problem 3}
@@
 \section{Problem 4}
-We want to show that $f$ is integrable.
-That is, for all $\epsilon>0$ there exist some step functions $f_1:[a,b]\rightarrow\mathbb{R}$ and $f_2:[a,b]\rightarrow\mathbb{R}$ where $f_1(x) \leq f(x) \leq f_2(x)$ and $\int_a^b f_1(x) - f_2(x)dx \leq \epsilon$ for all $\epsilon$.
-We will do this by defining a partition and step functions such that the difference between $f_2$ and $f_1$ is always less than $\frac{\epsilon}{N}$, then showing that the integral of $f_1-f_2$ is zero.
-This proof works for a monotone increasing function, but would work equally well for a monotone decreasing function with minimal switching.\\
-
-Define a partition $\{x_i\}$ on $[a,b]$.
-Define step functions $f_1$ and $f_2$ such that for each $x' \in [a,b]$ where $x_i \leq x' \leq x_{i+1}$, $f_1(x') = f(x_i)$ and $f_2(x') = f(x_{i+1})$.
-Define the partition $\{x_i\}$ such that $x_0 = a$ and $x_i$ is the value in the domain such that $f(x_i) = f(x_{i-1})+\frac{\epsilon}{N}$ for some $N$ and $\epsilon$, OR $b$, whichever is smaller.
-Note that this value must necessarily exist by the IVT, because $f$ is continuous on $[a,b]$ and $f(x_{i-1}) \leq f(x_i) \leq f(b)$ by f's positive monotonicity.
-Notice that $f(x_{i+1}) - f(x_i) \leq \frac{\epsilon}{N}$.
-Consider the following Riemann sum for $x_i \leq x'_i \leq x_{i+1}$:
-
-\begin{align*}
-  S &= \sum_{i=1}^{i=N} \left(f_1(x'_i) - f_2(x'_i)\right)\left(x_i-x_{i-1}\right)\\
-  &\leq \sum_{i=1}^{i=N} \frac{\epsilon}{N}\left(x_i-x_{i-1}\right)\\
-  &= \frac{\epsilon(b-a)}{N}\\
-\end{align*}
-
-Thus for arbitrary $\epsilon>0$ and $N > (b-a)$, $|S| < \epsilon$ and so $|S-0|<\epsilon$.
-Thus we can say $\int_a^bf_1(x)-f_2(x)dx = 0$.
-By extension $\int_a^bf_1(x)-f_2(x)dx < \epsilon$ for all $\epsilon>0$.
 

### 2017-03-19 311/latex/HW6.tex
bla

--- 2017-03-17 311/latex/HW6.tex
@@
 
-\title{Advanced Calc Homework 5}
+\title{Advanced Calc Homework 6}
 \author{Elliott Capek}
@@
 Consider a regular partition $\{x_i\}$ such that $x_i = \frac{i}{N}$ for some $N$.
-Define sequences $\{x_i'\}$ and $\{x_i''\}$ such that $x_i \leq x_i' \leq x_{i+1}$ and $x_i \leq x_i'' \leq x_{i+1}$, and $S_1 = \sum_{i=1}^{i=N}f(x_i')\left(x_i-x_{i-1}\right)$ and $S_2 = \sum_{i=1}^{i=N}f(x_i'')\left(x_i-x_{i-1}\right)$.
+Define sequences $\{x_i'\}$ and $\{x_i''\}$ such that $x_i \leq x_i' \leq x_{i+1}$ and $x_i \leq x_i'' \leq x_{i+1}$, so $S_1 = \sum_{i=1}^{i=N}f(x_i')\left(x_i-x_{i-1}\right)$ and $S_2 = \sum_{i=1}^{i=N}f(x_i'')\left(x_i-x_{i-1}\right)$.
 Note that:
@@
 
-Thus $S_1 \geq \frac{1}{2} + \frac{1}{2N}$ and $S_2 \leq \frac{-1}{2} + \frac{-1}{2N}$, so $|S_1-S_2| = |1 + \frac{1}{N}|$.
+Thus $S_1 \geq \frac{1}{2} + \frac{1}{2N}$ and $S_2 \leq \frac{-1}{2} + \frac{-1}{2N}$, so $|S_1-S_2| \geq |1 + \frac{1}{N}|$.
 Note that the width of $\{x_i\}$ is given by $\frac{1}{N}$.
@@
 Consider $N$ such that $\frac{1}{N}<\delta$.
-Note that $|S_1-S_2|>1>\frac{1}{2}$, a contradiction.
+For these parameters, $|S_1 - S_2| < \frac{1}{2}$.
+However, $|S_1-S_2|>1>\frac{1}{2}$, a contradiction.
 Thus $f(x)$ is not integrable.\\
@@
 By Theorem 3.5 because $f:[a,b]\rightarrow\mathbb{R}$ and $g:[a,b]\rightarrow\mathbb{R}$ are continuous, they are both integrable.
-By the triangle inequality, $\forall x, |f(x) + g(x)| \leq |f(x) + |g(x)|$.
+By the triangle inequality, $\forall x, |f(x) + g(x)| \leq |f(x)| + |g(x)|$.
 By Corollary 2.3, $\int_a^b |f(x)+g(x)| \leq \int_a^b |f(x)|+|g(x)| = \int_a^b |f(x)| + \int_a^b |g(x)|$, as desired.\\
@@
 
-Then $\lim_{N\rightarrow0} max(x_i-x_{i-1}) = 0$, so $\lim_{width(\{x_i\})} S = \frac{1}{2}$. Then $\int_0^1xdx = \frac{1}{2}$.\\
-
-%% Now consider $A = \frac{1}{2}$.
-Then $|S - A| = |\frac{1}{2N}|$.
-Now for some arbitrary $\epsilon$ by the Archimedean Principle there exists some $N'$ such that $\frac{1}{2N'} < \epsilon$.
-Because $N$ is arbitrary, consider $N = N'$.
-Then for any $\delta > \frac{b - a}{N}$, $||$
+Thus $S = \frac{1}{2} + \frac{1}{2N}$.
+Consider $A=\frac{1}{2}$.
+Then $|S - A| = \frac{1}{2N}$.
+Now for any $\epsilon > 0$, $\exists N$ such that $\frac{1}{2N} < \epsilon$ by the Archimedean property.
+Let $\delta \geq \frac{1}{N}$.
+Thus $max(x_i-x_{i-1}) = \frac{1}{N} < \delta$ implies $|S-A| < \epsilon$.
+We have found an A such that the integrability definition holds.
+Thus $\int_0^1xdx = \frac{1}{2}$.\\
 
@@
 We want to show that $f$ is integrable.
-That is, for all $\epsilon>0$ there exist some step functions $f_1:[a,b]\rightarrow\mathbb{R}$ and $f_2:[a,b]\rightarrow\mathbb{R}$ where $f_1(x) \leq f(x) \leq f_2(x)$ and $\int_a^b f_1(x) - f_2(x)dx \leq \epsilon$ for all $\epsilon$.
-We will do this by defining a partition and step functions such that the difference between $f_2$ and $f_1$ is always less than $\frac{\epsilon}{N}$, then showing that the integral of $f_1-f_2$ is zero.
-This proof works for a monotone increasing function, but would work equally well for a monotone decreasing function with minimal switching.\\
+That is, we want to show for all $\epsilon>0$ there exist some step functions $f_1:[a,b]\rightarrow\mathbb{R}$ and $f_2:[a,b]\rightarrow\mathbb{R}$ where $f_1(x) \leq f(x) \leq f_2(x)$ and $\int_a^b f_1(x) - f_2(x)dx \leq \epsilon$.
+We will do this by defining a partition and step functions such that the difference between $f_2$ and $f_1$ is always less than $\frac{\epsilon}{N}$, then showing that the integral of $f_1-f_2$ for these functions is zero.\\
 
+Let $f$ be monotone increasing.
 Define a partition $\{x_i\}$ on $[a,b]$.
@@
 
-Thus for arbitrary $\epsilon>0$ and $N > (b-a)$, $|S| < \epsilon$ and so $|S-0|<\epsilon$.
-Thus we can say $\int_a^bf_1(x)-f_2(x)dx = 0$.
-By extension $\int_a^bf_1(x)-f_2(x)dx < \epsilon$ for all $\epsilon>0$.
+Thus for arbitrary $\epsilon>0$ and $N > (b-a)$, $S = \int_a^bf_2(x)-f_1(x)dx < \epsilon$. Thus $f$ is integrable.\\
+
+A similar argument works if $f$ is monotone decreasing.\\
 

### 2017-03-22 ch362/latex/report3.tex
Add CH362 directory with report 3 part 1.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%This is the article LaTeX template for RSC journals
%Copyright The Royal Society of Chemistry 2010
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[8.5pt,twoside,twocolumn]{article}
\oddsidemargin -1.2cm
\evensidemargin -1.2cm
\textwidth 18cm
\headheight 1.0in
\topmargin -3.5cm
\textheight 22cm
\usepackage[super,sort&compress,comma]{natbib} 
\usepackage{mhchem}
\usepackage{times,mathptmx}
% \usepackage{times}
% feel free not to use mathptmx if it causes difficulties
\usepackage{sectsty}
\usepackage{balance}

\usepackage{graphicx} %eps figures can be used instead
\usepackage{lastpage}
\usepackage[format=plain,justification=raggedright,singlelinecheck=false,font=small,labelfont=bf,labelsep=space]{caption} 
\usepackage{fancyhdr}
\pagestyle{fancy}

\begin{document}

\thispagestyle{plain}
\fancypagestyle{plain}{
\fancyhead[L]{Communication}
\fancyhead[C]{\hspace{-1cm}\includegraphics[height=20pt]{headers/CH}}
\fancyhead[R]{Sience Chemistry\vspace{-0.2cm}}
\renewcommand{\headrulewidth}{1pt}}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand\footnoterule{\vspace*{1pt}% 
\hrule width 3.4in height 0.4pt \vspace*{5pt}} 
\setcounter{secnumdepth}{5}

\makeatletter 
\def\subsubsection{\@startsection{subsubsection}{3}{10pt}{-1.25ex plus -1ex minus -.1ex}{0ex plus 0ex}{\normalsize\bf}} 
\def\paragraph{\@startsection{paragraph}{4}{10pt}{-1.25ex plus -1ex minus -.1ex}{0ex plus 0ex}{\normalsize\textit}} 
\renewcommand\@biblabel[1]{#1}            
\renewcommand\@makefntext[1]% 
{\noindent\makebox[0pt][r]{\@thefnmark\,}#1}
\makeatother 
\renewcommand{\figurename}{\small{Fig.}~}
\sectionfont{\large}
\subsectionfont{\normalsize}

\fancyfoot{}
\fancyfoot[LO,RE]{\vspace{-7pt}Oregon State University 2017}
\fancyfoot[CO]{\vspace{-7.2pt}\hspace{12.2cm}Sience Chemistry}
\fancyfoot[CE]{\vspace{-7.5pt}\hspace{-14.5cm}\includegraphics{headers/RF}}
\fancyfoot[RO]{\footnotesize{\sffamily{1--\pageref{LastPage} ~\textbar  \hspace{2pt}\thepage}}}
\fancyfoot[LE]{\footnotesize{\sffamily{\thepage~\textbar\hspace{3.45cm} 1--\pageref{LastPage}}}}
\fancyhead{}
\renewcommand{\headrulewidth}{1pt} 
\renewcommand{\footrulewidth}{1pt}
\setlength{\arrayrulewidth}{1pt}
\setlength{\columnsep}{6.5mm}
\setlength\bibsep{1pt}

\twocolumn[
  \begin{@twocolumnfalse}
\noindent\LARGE{\textbf{Identification of an unknown disubstituted benzene derivative}}
\vspace{0.6cm}

\noindent\large{\textbf{Elliott Capek,\textit{$^{a}$} Steven Nguyen,\textit{$^{b\ddag}$} Kristin Ziebart,\textit{$^{b\ddag}$} Kevin Gable\textit{$^{b}$}}}\vspace{0.2cm}
%Please note that \ast indicates the corresponding author(s) but no footnote text is required.

\noindent\textit{\small{\textbf{Received 22\textit{$^{nd}$} March 2017, Accepted 27\textit{$^{th}$} March 2017}}}\newline

\vspace{0cm}
%Please do not change this text.

\noindent \normalsize{Small organic molecules give off distinct signals in mass spectrometry (MS) and NMR. MS of an unknown disubstituted benzene (codename \textit{Sample 8}) produces a spectra indicating a mass of $136$ atomic mass units (amu). $^{13}$C NMR suggests a total of 8 carbons present in the molecule, six being aromatic, one carboxylic and the other methyl. $^{1}$H-NMR and analysis of molecular mass confirm \textit{Sample 8} as a likely methyl benzoic acid. $^{1}$H-NMR splitting patterns, namely a lack of a singlet peak and presence of four unique aromatic hydrogens, suggest the \textit{meta} substitution pattern. $^{13}$C-$^{1}$H HMBC and HSQC experiments validate this designation. Thus \textit{Sample 8} is given the designation \textit{3-methylbenzoic acid}.}
\vspace{0.5cm}
 \end{@twocolumnfalse}
  ]

\section{Introduction}
%Footnotes

%Please use \dag to cite the ESI in the main text of the article.
%If you article does not have ESI please remove the the \dag symbol from the title and the above footnotetext.

\footnotetext{\textit{$^{a}$~Department of Physics, Oregon State University, Corvallis, Oregon, USA. E-mail: capeke@oregonstate.edu}}
\footnotetext{\textit{$^{b}$~Department of Chemistry, Oregon State University, Corvallis, Oregon, USA}}
\footnotetext{\textit{\ddag~ Authors contributed equally to this work.}}

%additional addresses can be cited as above using the lower-case letters, c, d, e... If all authors are from the same address, no letter is required

Mass spectrometry (MS) and NMR are invaluable techniques for structure elucidation of unknown organic molecules. When a molecule is sufficiently simple, knowledge of its molecular weight, carbon number, and rough identity of carbons and hydrogens can be enough to uniquely determine identity. For more complex molecules, or those with many isomers, further information from 2D experiments like $^{13}$C-$^{1}$H HSQC or $^{13}$C-$^{1}$H HMBC may be necessary.\\

This article follows the designation of a disubstituted benzene unknown using the above techniques. MS is used to determine weight, $^{13}$C NMR is used to come up with a candidate structure, $^{1}$H NMR is used to verify this candidate and guess its substitution, and $^{13}$C-$^{1}$H HSQC is used to verify the substitution.\\

\section{MS}
Molecular mass is first determined by examining the MS spectra shown in Figure \ref{fig:MS}. The largest peak in the highest M/Z ratio peak group is at 136 M/Z, indicating the molecular weight of the unknown is 135 g/mol.\\

The intensity ratio of the major 136 peak and the 137 peak is 0.89, which matches with the theoretical ratio of $0.011*8 = 0.88$ for an 8-carbon molecule.\\

\begin{figure}[h]
\centering
  \includegraphics[height=5cm]{figures/MS.png}
  \caption{Mass spectrometry spectra of \textit{Sample 8}.}
  \label{fig:MS}
\end{figure}

\section{$^{13}$C NMR}
The $^{13}$C NMR spectra shown in Figure \ref{fig:CNMR} has 9 peaks. As shown in Table \ref{table:MS}, one peak is due to a DMSO solvent, six to aromatic carbons, one to a methyl carbon and one to a carboxylic carbon. These results agree with the 8-carbon estimate from peak ratios in MS.\\

\begin{figure}[h]
\centering
  \includegraphics[height=5cm]{figures/CNMR.png}
  \caption{$^{13}$C NMR of \textit{Sample 8}.}
  \label{fig:CNMR}
\end{figure}

\begin{table}[h]
\small
  \caption{$^{13}$C NMR identified peaks for spectra shown in Figure \ref{fig:CNMR}.}
  \label{table:MS}
  \begin{tabular*}{0.5\textwidth}{@{\extracolsep{\fill}}lll}
    \hline
    Peak(s) shift(s) ppm & Designation \\
    \hline
    22 & Methyl carbon \\
    40 & DMSO solvent\\
    130-140 & 6x Aromatic carbons\\
    168 & Carboxylic carbon\\
    \hline
  \end{tabular*}
\end{table}

\section{$^{1}$H NMR}

\begin{figure}[h]
\centering
  \includegraphics[height=5cm]{figures/HNMR.png}
  \caption{$^{1}$H NMR of \textit{Sample 8}.}
  \label{fig:HNMR}
\end{figure}

\begin{table}[h]
\small
  \caption{$^{13}$C NMR identified peaks for spectra shown in Figure \ref{fig:HNMR}.}
  \label{table:HNMR}
  \begin{tabular*}{0.5\textwidth}{@{\extracolsep{\fill}}llll}
    \hline
    Peak(s) shift(s) ppm & Integration & Splitting & Designation \\
    \hline
    2.3 & 3.1 & Singlet & Ar-CH$_3$ hydrogens\\
    2.4 & 0.4 & Pentet & DMSO\\
    7.35 & 1.1 & Singlet & Aromatic H\\
    7.4 & 1.0 &  Meta Doublet & Aromatic H\\
    7.72 & 1.1 & Meta Doublet & Aromatic H\\
    7.75 & 1.1 & Singlet & Aromatic H\\
    12.8 & 1.0 & Singlet & Carboxylic H\\
    \hline
  \end{tabular*}
\end{table}

\section{Equations}

% an example of a two-column figure
%% \begin{figure*}
%%   \centering
%%   \includegraphics[height=3cm]{example.jpg}
%%   \caption{An example figure caption, an image from the \textit{Physical Chemistry Chemical Physics} cover gallery.}
%%   \label{fgr:example}
%% \end{figure*}

% an example of a two-column table
%\begin{table*}
%\small
  %\caption{\ An example of a caption to accompany a table, table captions do not end in a full point}
  %\label{tbl:example}
  %\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lllllll}
    %\hline
    %Header one & Header two & Header three & Header four & Header five & Header six  & Header seven\\
    %\hline
    %1 & 2 & 3 & 4 & 5 & 6  & 7\\
    %8 & 9 & 10 & 11 & 12 & 13 & 14 \\
    %15 & 16 & 17 & 18 & 19 & 20 & 21\\
    %\hline
  %\end{tabular*}
%\end{table*}

%The \balance command can be used to balance the columns on the final page if desired. It should be placed anywhere within the first column of the last page.

%\balance

%If notes are included in your references you can change the title from 'References' to 'Notes and references' using the following command:
%\renewcommand\refname{Notes and references}

\footnotesize{
\bibliography{rsc} %your .bib file
\bibliographystyle{rsc} %the RSC's .bst file
}

\end{document}

### 2017-03-22 ch362/latex/equilibration.tex
all

--- 2017-03-22 ch362/latex/report3.tex
@@
 \usepackage{balance}
+\usepackage{graphicx} %eps figures can be used instead
 
-\usepackage{graphicx} %eps figures can be used instead
 \usepackage{lastpage}
@@
 \fancyhead[C]{\hspace{-1cm}\includegraphics[height=20pt]{headers/CH}}
-\fancyhead[R]{Sience Chemistry\vspace{-0.2cm}}
+\fancyhead[R]{Science \vspace{-0.2cm}}
 \renewcommand{\headrulewidth}{1pt}}
@@
 \fancyfoot[LO,RE]{\vspace{-7pt}Oregon State University 2017}
-\fancyfoot[CO]{\vspace{-7.2pt}\hspace{12.2cm}Sience Chemistry}
-\fancyfoot[CE]{\vspace{-7.5pt}\hspace{-14.5cm}\includegraphics{headers/RF}}
+\fancyfoot[CO]{\vspace{-7.2pt}\hspace{12.2cm}Science}
 \fancyfoot[RO]{\footnotesize{\sffamily{1--\pageref{LastPage} ~\textbar  \hspace{2pt}\thepage}}}
@@
 
-\noindent \normalsize{Small organic molecules give off distinct signals in mass spectrometry (MS) and NMR.
-MS of an unknown disubstituted benzene (codename \textit{Sample 8}) produces a spectra indicating a mass of $136$ atomic mass units (amu).
-$^{13}$C NMR suggests a total of 8 carbons present in the molecule, six being aromatic, one carboxylic and the other methyl.
-$^{1}$H-NMR and analysis of molecular mass confirm \textit{Sample 8} as a likely methyl benzoic acid.
-$^{1}$H-NMR splitting patterns, namely a lack of a singlet peak and presence of four unique aromatic hydrogens, suggest the \textit{meta} substitution pattern.
-$^{13}$C-$^{1}$H HMBC and HSQC experiments validate this designation.
-Thus \textit{Sample 8} is given the designation \textit{3-methylbenzoic acid}.}
+\noindent \normalsize{}
 \vspace{0.5cm}
@@
 
-Mass spectrometry (MS) and NMR are invaluable techniques for structure elucidation of unknown organic molecules.
-When a molecule is sufficiently simple, knowledge of its molecular weight, carbon number, and rough identity of carbons and hydrogens can be enough to uniquely determine identity.
-For more complex molecules, or those with many isomers, further information from 2D experiments like $^{13}$C-$^{1}$H HSQC or $^{13}$C-$^{1}$H HMBC may be necessary.\\
-
-This article follows the designation of a disubstituted benzene unknown using the above techniques.
-MS is used to determine weight, $^{13}$C NMR is used to come up with a candidate structure, $^{1}$H NMR is used to verify this candidate and guess its substitution, and $^{13}$C-$^{1}$H HSQC is used to verify the substitution.\\
-
-\section{MS}
-Molecular mass is first determined by examining the MS spectra shown in Figure \ref{fig:MS}.
-The largest peak in the highest M/Z ratio peak group is at 136 M/Z, indicating the molecular weight of the unknown is 135 g/mol.\\
-
-The intensity ratio of the major 136 peak and the 137 peak is 0.89, which matches with the theoretical ratio of $0.011*8 = 0.88$ for an 8-carbon molecule.\\
-
-\begin{figure}[h]
-\centering
-  \includegraphics[height=5cm]{figures/MS.png}
-  \caption{Mass spectrometry spectra of \textit{Sample 8}.}
-  \label{fig:MS}
-\end{figure}
-
-\section{$^{13}$C NMR}
-The $^{13}$C NMR spectra shown in Figure \ref{fig:CNMR} has 9 peaks.
-As shown in Table \ref{table:MS}, one peak is due to a DMSO solvent, six to aromatic carbons, one to a methyl carbon and one to a carboxylic carbon.
-These results agree with the 8-carbon estimate from peak ratios in MS.\\
-
-\begin{figure}[h]
-\centering
-  \includegraphics[height=5cm]{figures/CNMR.png}
-  \caption{$^{13}$C NMR of \textit{Sample 8}.}
-  \label{fig:CNMR}
-\end{figure}
+\section{Results}
 
@@
   \small
-  \caption{$^{13}$C NMR identified peaks for spectra shown in Figure \ref{fig:CNMR}.}
-  \label{table:MS}
-  \begin{tabular*}{0.5\textwidth}{@{\extracolsep{\fill}}lll}
+  \center
+  \caption{Equilibrium constants of tautomerization of the keto and enol forms of 2,4-pentanedione in different solvents.}
+  \label{table:keqs}
+  \begin{tabular*}{0.3\textwidth}{@{\extracolsep{\fill}}cc}
     \hline
-    Peak(s) shift(s) ppm & Designation \\
+    Solvent & $K_{eq} = [enol]/[keto]$ \\
     \hline
-    22 & Methyl carbon \\
-    40 & DMSO solvent\\
-    130-140 & 6x Aromatic carbons\\
-    168 & Carboxylic carbon\\
+    Acetone-d6      & 3.25\\
+    Acetic acid-d4  & 3.32\\
+    Acetonitrile-d3 & 1.54\\
+    Benzene-d6      & 9.44\\
+    Chloroform      & 5.79\\
+    DMF             & 2.04\\
+    DMSO-d6         & 1.43\\
+    Methanol        & 3.65\\
+    Toluene         & 10.59\\
     \hline
@@
 
-\section{$^{1}$H NMR}
+\begin{figure*} \label{ fig7}
+  \begin{minipage}[b]{0.49\linewidth}
+    \centering
+    \includegraphics[width=.5\linewidth]{figures/image}
+    \caption{Initial condition} 
+  \end{minipage}%%
+  \begin{minipage}[b]{0.4\linewidth}
+    \includegraphics[width=.5\linewidth]{figures/image1}
+    \caption{Rupture} 
+  \end{minipage}
+  \begin{minipage}[b]{0.49\linewidth}
+    \includegraphics[width=.5\linewidth]{figures/image2}
+    \caption{DFT, Initial condition} 
+  \end{minipage}%%
+  \begin{minipage}[b]{0.49\linewidth}
+    \includegraphics[width=.5\linewidth]{figures/image3}
+    \caption{DFT, rupture} 
+  \end{minipage} 
+\end{figure*}
 
-\begin{figure}[h]
-\centering
-  \includegraphics[height=5cm]{figures/HNMR.png}
-  \caption{$^{1}$H NMR of \textit{Sample 8}.}
-  \label{fig:HNMR}
-\end{figure}
+\begin{@twocolumnfalse}
+\section{References}
+[1] Experimental Chemistry I: CH 362, OSU Department of Chemistry 2017\\
 
-\begin{table}[h]
-\small
-  \caption{$^{13}$C NMR identified peaks for spectra shown in Figure \ref{fig:HNMR}.}
-  \label{table:HNMR}
-  \begin{tabular*}{0.5\textwidth}{@{\extracolsep{\fill}}llll}
-    \hline
-    Peak(s) shift(s) ppm & Integration & Splitting & Designation \\
-    \hline
-    2.3 & 3.1 & Singlet & Ar-CH$_3$ hydrogens\\
-    2.4 & 0.4 & Pentet & DMSO\\
-    7.35 & 1.1 & Singlet & Aromatic H\\
-    7.4 & 1.0 &  Meta Doublet & Aromatic H\\
-    7.72 & 1.1 & Meta Doublet & Aromatic H\\
-    7.75 & 1.1 & Singlet & Aromatic H\\
-    12.8 & 1.0 & Singlet & Carboxylic H\\
-    \hline
-  \end{tabular*}
-\end{table}
+[2] OSU of Chemistry. 1H NMR Chemical Shifts. http://www.science.oregonstate.edu/~gablek/CH335/Chapter10/ChemicalShift.htm. (Accessed March 22, 2017)\\
 
-\section{Equations}
+[3] University of Washington Department of Chemistry. Dielectric constant table for common solvents. $https://depts.washington.edu/eooptic/linkfiles/dielectric_chart$\%$5B1$\%$5D.pdf$
 
-% an example of a two-column figure
-%% \begin{figure*}
-%%   \centering
-%%   \includegraphics[height=3cm]{example.jpg}
-%%   \caption{An example figure caption, an image from the \textit{Physical Chemistry Chemical Physics} cover gallery.}
-%%   \label{fgr:example}
-%% \end{figure*}
-
-% an example of a two-column table
-%\begin{table*}
-%\small
-  %\caption{\ An example of a caption to accompany a table, table captions do not end in a full point}
-  %\label{tbl:example}
-  %\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lllllll}
-    %\hline
-    %Header one & Header two & Header three & Header four & Header five & Header six  & Header seven\\
-    %\hline
-    %1 & 2 & 3 & 4 & 5 & 6  & 7\\
-    %8 & 9 & 10 & 11 & 12 & 13 & 14 \\
-    %15 & 16 & 17 & 18 & 19 & 20 & 21\\
-    %\hline
-  %\end{tabular*}
-%\end{table*}
-
-%The \balance command can be used to balance the columns on the final page if desired. It should be placed anywhere within the first column of the last page.
-
-%\balance
-
-%If notes are included in your references you can change the title from 'References' to 'Notes and references' using the following command:
-%\renewcommand\refname{Notes and references}
-
-\footnotesize{
-\bibliography{rsc} %your .bib file
-\bibliographystyle{rsc} %the RSC's .bst file
-}
-
+[4] Computational Chemistry Comparison and Benchmark Database. Experimental Dipoles. http://cccbdb.nist.gov/diplistx.asp
+\end{@twocolumnfalse}
 \end{document}

### 2017-03-23 311/latex/final.tex
all

--- 2017-03-17 311/latex/HW6.tex
@@
 
-\section{Problem 6.4.2}
-In search of a contradiction, say $f(x)$ is integrable.
-Then $\forall \epsilon > 0$ $\exists \delta>0$ such that for an arbitrary $\{x_i^1\}$ and $\{x_i^2\}$, $width(\{x_i^1\})<\delta$ and $width(\{x_i^2\})<\delta$ implies $|S_1-S_2|<\epsilon$.\\
+\section{Problem 2}
+\textbf{Prove that $\lim_{n\rightarrow\infty}a_n = a$ if and only if the sequence $a_1, a, a_2, a, a_3, a,...$ is convergent.}
 
-Consider a regular partition $\{x_i\}$ such that $x_i = \frac{i}{N}$ for some $N$.
-Define sequences $\{x_i'\}$ and $\{x_i''\}$ such that $x_i \leq x_i' \leq x_{i+1}$ and $x_i \leq x_i'' \leq x_{i+1}$, and $S_1 = \sum_{i=1}^{i=N}f(x_i')\left(x_i-x_{i-1}\right)$ and $S_2 = \sum_{i=1}^{i=N}f(x_i'')\left(x_i-x_{i-1}\right)$.
-Note that:
+To prove this, we will need to show \textbf{A.)} that $\lim_{n\rightarrow\infty}a_n = a$ implies $a_1, a, a_2, a, a_3, a,...$ is convergent and \textbf{B.)} that $a_1, a, a_2, a, a_3, a,...$ is convergent implies $\lim_{n\rightarrow\infty}a_n = a$.\\
 
-\begin{align*}
-  S_1 &= \sum_{i=1}^{i=N}f(x_i')(x_i-x_{i-1}) = \frac{1}{N}\sum_{i=1}^{i=N}x_i' \geq \frac{1}{N}\sum_{i=1}^{i=N}x_i\\
-  &= \frac{1}{N^2}\sum_{i=1}^{i=N}i = \frac{N(N+1)}{2N^2} = \frac{1}{2} + \frac{1}{2N}\\
-  S_2 &= \sum_{i=1}^{i=N}f(x_i'')(x_i-x_{i-1}) = \frac{-1}{N}\sum_{i=1}^{i=N}x_i'' \leq \frac{-1}{N}\sum_{i=1}^{i=N}x_i\\
-  &= \frac{-1}{N^2}\sum_{i=1}^{i=N}i = \frac{-N(N+1)}{2N^2} = \frac{-1}{2} + \frac{-1}{2N}\\
-\end{align*}
+\textbf{A.)}
+Let $\lim_{n\rightarrow\infty}a_n = a$.
+Then $\forall \epsilon>0$ $\exists N$ such that $\forall n \geq N$, $|a_n-a|<\epsilon$.
+Consider sequence $A = \{a_1, a, a_2, a, a_3, ...\}$.
+Note that $|a-a|<\epsilon$.
+Consider $J$ such that $a_N = A_J$, which necessarily must exist since $A \subseteq$.
+Note that $\forall j>J$, either $A_j = a$, in which case $|A_j-a|<\epsilon$, or $A_j \in \{a_n\}$ such that $A_j = a_n$ where $n > N$, thus $|A_j-a|<\epsilon$.
+Thus $A$ is convergent.\\
 
-Thus $S_1 \geq \frac{1}{2} + \frac{1}{2N}$ and $S_2 \leq \frac{-1}{2} + \frac{-1}{2N}$, so $|S_1-S_2| = |1 + \frac{1}{N}|$.
-Note that the width of $\{x_i\}$ is given by $\frac{1}{N}$.
-Thus $\forall \epsilon>0$, $\exists \delta > 0$ such that $\frac{1}{N}<\delta$ implies $|S_1-S_2|<\epsilon$ for our chosen $S_1$ and $S_2$.
-Consider $\epsilon = \frac{1}{2}$, and the associated $\delta$.
-Consider $N$ such that $\frac{1}{N}<\delta$.
-Note that $|S_1-S_2|>1>\frac{1}{2}$, a contradiction.
-Thus $f(x)$ is not integrable.\\
-
-\section{Problem 3}
-Because $f(x) = x$ is integrable, then for all $\epsilon > 0$, it is the case that for some $\delta > 0$, for all $\{x_i\}$, $max(x_i-x_{i-1})<\delta$, $|S - A| < \epsilon$.
-Consider $\{x_i\}$ such that $x_i = a + \frac{b-a}{N}i$ for an arbitrary $N$.
-Then $max(x_i-x_{i-1}) = \frac{b-a}{N}$.
-Consider the following algebraic manipulation, where $x_i' = x_i$ and $x_i$ is a regular partition and $a=0, b=1$:
-
-\begin{align*}
-  S &= \sum_{i=1}{i=N}f(x_i')(x_{i+1}-x_i)\\
-  &= \sum_{i=1}{i=N}\left(a+\frac{b-a}{N}i\right)\left(a+\frac{b-a}{N}i - a -\frac{b-a}{N}(i-1)\right)\\
-  &= \sum_{i=1}{i=N}\left(\frac{i}{N}\right)\left(\frac{1}{N}\right)\\
-  &= \frac{1}{N^2}\sum_{i=1}{i=N}i\\
-  &= \frac{1}{N^2}\frac{N(N+1)}{2} = \frac{1}{2} + \frac{1}{2N}\\
-\end{align*}
-
-Then $\lim_{N\rightarrow0} max(x_i-x_{i-1}) = 0$, so $\lim_{width(\{x_i\})} S = \frac{1}{2}$. Then $\int_0^1xdx = \frac{1}{2}$.\\
-
-%% Now consider $A = \frac{1}{2}$.
-Then $|S - A| = |\frac{1}{2N}|$.
-Now for some arbitrary $\epsilon$ by the Archimedean Principle there exists some $N'$ such that $\frac{1}{2N'} < \epsilon$.
-Because $N$ is arbitrary, consider $N = N'$.
-Then for any $\delta > \frac{b - a}{N}$, $||$
-
-\section{Problem 4}
+\textbf{B.)}
 

### 2017-03-24 311/latex/final.tex
all

@@
 
-\textbf{A.)}
+\textbf{A.)}\\
 Let $\lim_{n\rightarrow\infty}a_n = a$.
@@
 
-\textbf{B.)}
+\textbf{B.) $a_1, a, a_2, a, a_3, a,...$ is convergent implies $\lim_{n\rightarrow\infty}a_n = a$}\\
+Assume $A = \{a_1, a, a_2, a, a_3, ...\}$ converges.
+We want to show $\lim_{n\rightarrow\infty}A_n = a$.
+In search of a contradiction, let $A$ converge to $a + b$ for some $b \neq 0$.
+Then for $\epsilon > b$, $\exists N$ such that $\forall n > N$, $|A_N-a-b|<\epsilon$.
+However, note that $\exists n > N$ such that $a = A_n$, ths $|a - a - b| = |b| < \epsilon$.
+This is a contradiction.
+Thus, $\lim_{n\rightarrow\infty} A = a$.\\
+
+Now $\forall \epsilon>0$, $\exists J$ such that $\forall j > J$, $|A_J - a|<\epsilon$.
+Note that by construction, $\exists M$ such that $a_M = A_J$ or $a_M = A_{J+1}$.
+Then $\forall m > M$, $|a_M - a|<\epsilon$.
+Thus $\lim_{n\rightarrow\infty}\{a_n\} = a$.\\
+
+\section{Problem 3}
+We want to show
+
+\[
+f(x) =
+\begin{cases}
+  0 & x <    0 \\
+  x & x \geq 0 \\
+\end{cases}
+\]
+
+is continuous at $x_0 = 0$.
+This entails showing that $\forall \epsilon>0$, $\exists \delta>0$ such that $|x-x_0|<\delta$ implies $|f(x)-f(x_0)|<\epsilon$.
+Consider the following manipulation, where $x_0 = 0$:
+
+\begin{align*}
+  |f(x) - f(x_0)| &= |f(x)| \leq |x| = |x-x_0|
+\end{align*}
+
+Thus $|x-x_0|<\epsilon$ implies $|f(x)-f(x_0)|<\epsilon$. Thus $f(x)$ is continuous.\\
+
+\section{Problem 5} \textbf{Prove that if p(x) is a polynomial of degree $n$, then there are most $n$ solutions to the equation $p(x) = 0$.}\\
+We will do this by induction.
+First, we will show the $k=1$ case where $p^{n-1}(x) = 0$ has at most one solution.
+Then we will show the inductive step $k(i)$ implies $k(i+1)$, that $p^{i}(x) = 0$ has $n$ solutions implies $p^{i-1}(x) = 0$ has at most $n+1$ solutions.\\
+
+\textbf{Base case}\\
+Consider degree-n polynomial $p(x)$.
+Consider the $n-1$th derivative, $p^{n-1}(x)=n\!x+c$.
+$$p^{n-1}(x)=0$ has at least one solution; consider $x=\frac{-c}{n}$.
+We want to show this is a unique solution.
+In search of a contradiction, consider another solution $x_0 \neq \frac{-c}{n}$ such that $p^{n-1}(x_0) = 0$.
+$p^{n-1}(x)$ is a polynomial, and thus is everywhere differentiable and continuous.
+Then by Rolle's Theorem $\exists x'$ such that, if $x_0>\frac{-c}{n}$ $x'\in\(\frac{-c}{n},x_0\)$ or if $x_0<\frac{-c}{n}$ $x'\in\(x_0,\frac{-c}{n}\)$ and $p^{n}(x)=0$.
+However, $p^{n}(x)=n\!$.
+This is a contradiction.
+Thus, $p^{n-1}(x)$ has exactly one solution.\\
+
+\textbf{Inductive step}\\
 

### 2017-03-24 311/latex/final.tex
final final

@@
 
+\section{Problem 1}
+Because $b>1$, by the Archimedean Property, $\exists n \in \mathbb{N}$ such that $b > 1 - \frac{1}{n}$. Now consider:\\
+
+\begin{align*}
+  b^n &\geq \left(1 + \frac{1}{n}\right)^n\\
+  &= \sum_{i=0}^{n}{n \choose i}\frac{1}{n^i}\\
+  &= \sum_{i=2}^{n}{n \choose i}\frac{1}{n^i} + 2
+\end{align*}
+
+Thus, since all components of $\sum_{i=2}^{n}{n \choose i}\frac{1}{n^i}$ are positive, $b^n \geq 2$.
+Thus for any $x>2$, $\exists m \in \mathbb{N}$ such that $m > log_2(x)$ where $b^{n+m} \in \left(b, b^2, b^3, ...\right)$ and $b^{n+m}>x$.
+Thus, $\left(b, b^2, b^3, ...\right)$ is not bounded from above.\\
+
 \section{Problem 2}
@@
 This entails showing that $\forall \epsilon>0$, $\exists \delta>0$ such that $|x-x_0|<\delta$ implies $|f(x)-f(x_0)|<\epsilon$.
-Consider the following manipulation, where $x_0 = 0$:
+Consider the following manipulation, where $x_0 = 0$.
+Note that $|f(x)|\leq x$ for all $x$.\\
 
@@
 
+\section{Problem 4}
+Consider the following algebraic manipulation:
+
+\begin{align*}
+  \lim_{x\rightarrow0}\frac{f(x)}{g(x)} &= \lim_{x\rightarrow0}\frac{f(x)-f(0)}{g(x)-g(0)}\\
+  &= \lim_{x\rightarrow0}\frac{\frac{f(x)-f(0)}{x}}{\frac{g(x)-g(0)}{x}}\\
+  &= \frac{f'(0)}{g'(0)}\\
+\end{align*}
+
+Note that because $f'$ and $g'$ are continuous at zero, their limits can be taken:
+
+\begin{align*}
+  \lim_{x\rightarrow0}\frac{f(x)}{g(x)} &= \lim_{x\rightarrow0}\frac{f'(x)}{g'(x)}\\
+\end{align*}
+
 \section{Problem 5} \textbf{Prove that if p(x) is a polynomial of degree $n$, then there are most $n$ solutions to the equation $p(x) = 0$.}\\
@@
 First, we will show the $k=1$ case where $p^{n-1}(x) = 0$ has at most one solution.
-Then we will show the inductive step $k(i)$ implies $k(i+1)$, that $p^{i}(x) = 0$ has $n$ solutions implies $p^{i-1}(x) = 0$ has at most $n+1$ solutions.\\
+Then we will show the inductive step $k(i)$ implies $k(i+1)$, that $p^{i}(x) = 0$ has $j$ solutions implies $p^{i-1}(x) = 0$ has at most $j+1$ solutions.\\
 
@@
 Consider degree-n polynomial $p(x)$.
-Consider the $n-1$th derivative, $p^{n-1}(x)=n\!x+c$.
-$$p^{n-1}(x)=0$ has at least one solution; consider $x=\frac{-c}{n}$.
+Consider the $n-1$th derivative, $p^{n-1}(x)=n!x+c$.
+$p^{n-1}(x)=0$ has at least one solution; consider $x=\frac{-c}{n}$.
 We want to show this is a unique solution.
@@
 $p^{n-1}(x)$ is a polynomial, and thus is everywhere differentiable and continuous.
-Then by Rolle's Theorem $\exists x'$ such that, if $x_0>\frac{-c}{n}$ $x'\in\(\frac{-c}{n},x_0\)$ or if $x_0<\frac{-c}{n}$ $x'\in\(x_0,\frac{-c}{n}\)$ and $p^{n}(x)=0$.
-However, $p^{n}(x)=n\!$.
+Then by Rolle's Theorem $\exists x'$ such that, if $x_0>\frac{-c}{n}$ $x'\in(\frac{-c}{n},x_0)$ or if $x_0<\frac{-c}{n}$ $x'\in(x_0,\frac{-c}{n})$ and $p^{n}(x)=0$.
+However, $p^{n}(x)=n!$.
 This is a contradiction.
@@
 \textbf{Inductive step}\\
+We now want to show that if $p^{i}(x)=0$ has $n$ unique solutions, then $p^{i-1}(x)=0$ has at most $n+1$ unique solutions.
+Let $\{a_1, a_2, a_3, ..., a_n\}$ be the set of unique roots of $p^i(x)=0$.
+Consider interval $(a_m, a_{m+1})$.
+Note that, since there are no roots between $a_m$ and $a_{m+1}$, either $\forall x \in (a_m, a_{m+1})$, thus $p^i(x)>0$, or $\forall x \in (a_m, a_{m+1})$, thus $p^i(x)<0$.
+In the first case, $\lim_{x_0\rightarrow x} \frac{p^{i-1}(x_0)-p^{i-1}(x)}{x_0-x} > 0$, so $x_0>x$ implies $p^{i-1}(x_0)>p^{i-1}(x)$, and $x_0<x$ implies $p^{i-1}(x_0)<p^{i-1}(x)$.
+Thus if $\exists x' \in (a_m, a_{m+1})$ such that $p^{i-1}(x')=0$, then $\forall x \in (a_m, a_{m+1})$, $x>x'$ implies $p^{i-1}(x)>0$, or if $x<x'$ then $p^{i-1}(x)<0$.
+In the second case, similarly if $\exists x' \in (a_m, a_{m+1})$ such that $p^{i-1}(x')=0$, then $\forall x \in (a_m, a_{m+1})$, $x>x'$ implies $p^{i-1}(x)<0$, or if $x<x'$ then $p^{i-1}(x)>0$.
+Thus there can occur only one root in $(a_m,a_{m+1})$.\\
+
+Now consider interval $[a_m,a_{m+1}]$.
+Note that if either $f(a_m)=0$ or $f(a_{m+1})$, then $\forall x \in (a_m,a_{m+1})$ either $p^{i-1}(x)>0$ or $p^{i-1}(x)<0$ by the strict monotonicity of the closed interval inside the open one.
+Thus each open interval $[a_m,a_{m+1}]$ has at most one root inside it.
+There are $n-1$ such open intervals, and thus $n-1$ such roots.\\
+
+Also consider the side intervals $(-\infty,a_1]$ and $[a_n,\infty)$.
+By a similar reasoning to that above, if any element $x \in (-\infty,a_1]$ or $x \in [a_n,\infty)$ meets criteria $p^{i-1}(x)=0$, then for all other elements $x'\neq x$ in the interval, either $p^{i-1}(x)>0$ or $p^{i-1}(x)<0$.
+Thus there can be only one root in each of these intervals.
+Thus, there can be a maximum of $n-1+2=n+1$ roots for $p^{i-1}$.\\
+
+Thus, by induction, $p(x)$ can have at most n roots.\\
+
+\section{Problem Six}
+Consider continuous function $f$ and antiderivative $F$ defined as $F(x) = \int_0^{x}f(t)dt$. By the Second Fundamental Theorem of Calculus:\\
+
+\begin{align*}
+  f(0) &= F'(0) = \lim_{x_0\rightarrow0} \frac{F(x_0)-F(0)}{x_0}\\
+  &= \lim_{x_0\rightarrow0} \frac{\int_0^{x_0}f(t)dt-\int_0^0f(t)dt}{x_0}\\
+  &= \lim_{x_0\rightarrow0} \frac{\int_0^{x_0}f(t)dt}{x_0}
+\end{align*}
+
+We also know that $f(0) = \lim_{x\rightarrow0}f(0)=c$, thus:
+
+\begin{align*}
+  c &= \lim_{x_0\rightarrow0} \frac{\int_0^{x_0}f(t)dt}{x_0}
+\end{align*}
 

### 2017-10-12 em/latex/HW2.tex
EM

--- 2016-10-19 em/latex/HW2.tex
@@
 
-\title{Quantum Mechanics Assignment One}
+\title{Elegromagnetism Assignment Two}
 \author{Elliott Capek}
 \titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles
+
+\newcommand{\bra}[1]{\big<#1\big|}
+\newcommand{\ket}[1]{\big|#1\big>}
+\newcommand{\braket}[2]{\big<#1\big|#2\big>}
 
@@
 \section{Problem One}
-\textbf{Boundaries:}
+To find the bound charges on this cross section, we use the following equations:
 
 \begin{align*}
-  \phi(\infty) &= 0\\
-  \phi(x,0) &= 0\\
-  \phi(0,y) &= 0\\
+  \rho_b &= -\nabla \vec{P} = 0\hspace{2cm} \mbox{P is constant}\\
+  \sigma_b &= \vec{P} \cdot \vec{n}\\
 \end{align*}
 
-Guess the solution is four charges in a rectangle. If this solution works, by the Uniqueness Theorem it is the only solution.
+The top two faces each make a $60^\circ$ angle with $\vec{P}$, and the bottom one is antiparallel. Thus we see:
 
 \begin{align*}
-  \phi &= \phi_1 + \phi_2 + \phi_3 + \phi_4\\
+  \sigma_{b,top} &= \vec{P} \cdot \hat{n} = P\cos\left(60^\circ\right) = \frac P2\\
+  \sigma_{b,bot} &= \vec{P} \cdot \hat{n} = P\cos\left(180^\circ\right) = -P\\
 \end{align*}
 
-The potential due to the real point charge is:
+To find total charge we just sum these densities over the surface area of the prism. We notice that there are two top faces and one
+bottom face, thus the two surface charges cancel out and the net charge on the object is \textbf{zero}. We expect this for a dielectric with no free charge.\\
+
+We then find the far field by treating the triangle as a point dipole extending infinitely through the z axis:
 
 \begin{align*}
-  \phi_1(x,y) &= \frac{q}{4\pi\epsilon\sqrt{(x-d_1)^2+(y-d_2)^2}}\\
+  \vec{E}(\vec{r}, \theta) &= \int_\infty^\infty \frac{\vec{P}\cdot\vec{r}}{\sqrt{z^2+r^2}}dz
+  = P\cos(\theta)\int_\infty^\infty \frac{\sqrt{z^2+r^2}}{\sqrt{z^2+r^2}^3}dz
+  = P\cos(\theta)\int_\infty^\infty \frac{1}{z^2+r^2}dz\\
+  &= \frac{P\cos(\theta)\pi}{r}
 \end{align*}
 
-We let the second and fourth-quadrant points have charge -q, so potential:
+This result almost makes sense. Normally a dipole falls off by $\frac{1}{r^3}$, but this one is generated by an infinite line, so it
+falls off lots slower.\\
+
+\section{Problem Two}
+The example solved in Griffiths Ex 4.7 can be generalized to a two-dielectrics-in-all-space system by letting $\epsilon_r = \epsilon_1 / \epsilon_2$:
+
+\textbf{a.)}\\
+\begin{align*}
+	V_{in} &= -\frac{3E_0r\cos\theta}{\epsilon_r+2}\\
+	E_{in} &= \frac{3E_0\cos\theta}{\epsilon_r+2}\hat{r} + \frac{-3E_0\sin\theta}{\epsilon_r+2}\hat{\theta}\\
+        V_{out} &= -E_0r\cos\theta + \frac{(\epsilon_r-1)R^3E_0\cos\theta}{(\epsilon_r+2)r^2}\\
+        E_{out} &= \left(E_0\cos\theta+\frac{2(\epsilon_r-1)R^3E_0\cos\theta}{(\epsilon_r+2)r^3}\right)\hat{r} + \left(-E_0\sin\theta+\frac{-(\epsilon_r-1)R^3E_0\sin\theta}{(\epsilon_r+2)r^3}\right)\hat{\theta}\\
+        &= \left(\frac{E_0\cos\theta(3\epsilon_r-1)}{\epsilon_r+2}\right)\hat{r} + \left(\frac{-3E_0\sin\theta}{\epsilon_r+2}\right)\hat{\theta}\\
+\end{align*}
+
+\textbf{b.)}\\
+Solving for free and bound surface charge:
 
 \begin{align*}
-  \phi_2(x,y) &= \frac{-q}{4\pi\epsilon\sqrt{(x+d_1)^2+(y-d_2)^2}}\\
-  \phi_4(x,y) &= \frac{-q}{4\pi\epsilon\sqrt{(x-d_1)^2+(y+d_2)^2}}\\
+  P &= \epsilon_1E\\
+  \sigma_b &= P \cdot \hat{r} = \frac{3E_0\cos\theta}{\epsilon_r+2}\\
+  \sigma_f &= \epsilon_1E_{in,r} - \epsilon_2E_{out,r} = \frac{-E_0\cos\theta\epsilon_2}{\epsilon_r+2}\\
+  \sigma &= \frac{-E_0\cos\theta(3-\epsilon_2)}{\epsilon_r+2}\\
 \end{align*}
 
-and the third-quadrant point have charge +q, so potential:
+This charge is positive at the top of the sphere and negative at the bottom, as expected for a bound charge.\\
+
+\textbf{c.)}\\
+Tangential E-fields are equal at $r=R$:
 
 \begin{align*}
-  \phi_3(x,y) &= \frac{q}{4\pi\epsilon\sqrt{(x+d_1)^2+(y+d_2)^2}}\\
+  E_{in,\theta} - E_{out,\theta} =& \frac{-3E_0\sin\theta}{\epsilon_r+2} - \frac{-3E_0\sin\theta}{\epsilon_r+2} = 0\\
 \end{align*}
 
-Thus the total potential is given by:
+\textbf{d.)}
+E-fields are curl-free. Because the dielectrics are linear, D is proportional to E, and so it seems like the difference should be zero. However, we see that:\\
 
 \begin{align*}
-  \phi(x,y) &= \frac{q}{4\pi\epsilon}\Big(
-  \left((x-d_1)^2+(y-d_2)^2\right)^{-1/2} + \left((x+d_1)^2+(y+d_2)^2\right)^{-1/2}\\
-  &-\left((x+d_1)^2+(y-d_2)^2\right)^{-1/2} - \left((x-d_1)^2+(y+d_2)^2\right)^{-1/2}\Big)
+  D_{out} - D_{in} &= \left(\epsilon_0-\epsilon_1\right)\frac{-3E_0\sin\theta}{\epsilon_r+2}
 \end{align*}
 
-This satisfies our two boundary conditions:
-
-\begin{align*}
-  \phi(0,y) &= \frac{q}{4\pi\epsilon}\Big(
-  \left(d_1^2+(y-d_2)^2\right)^{-1/2} + \left(d_1^2+(y+d_2)^2\right)^{-1/2}
-  -\left(d_1^2+(y-d_2)^2\right)^{-1/2} - \left(d_1^2+(y+d_2)^2\right)^{-1/2}\Big)
-  = 0\\
-\end{align*}
-
-\begin{align*}
-  \phi(x,0) &= \frac{q}{4\pi\epsilon}\Big(
-  \left((x-d_1)^2+d_2^2\right)^{-1/2} + \left((x+d_1)^2+d_2^2\right)^{-1/2}
-  -\left((x+d_1)^2+d_2^2\right)^{-1/2} - \left((x-d_1)^2+d_2^2\right)^{-1/2}\Big)
-  = 0\\
-\end{align*}
-
-To find the force on the first-quadrant charge, we use Coulomb's Equation:
-
-%% \begin{align*}
-%%   F &= -q\left(\frac{d}{dx} + \frac{d}{dy}\right) \frac{q}{4\pi\epsilon}\Big(
-%%   \left((x-d_1)^2+(y-d_2)^2\right)^{-1/2} + \left((x+d_1)^2+(y+d_2)^2\right)^{-1/2}\\
-%%   &-\left((x+d_1)^2+(y-d_2)^2\right)^{-1/2} - \left((x-d_1)^2+(y+d_2)^2\right)^{-1/2}\Big)\\
-%%   &= \frac{-q^2}{4\pi\epsilon} \Big(
-%%   (x-d_1 + y-d_2)\left((x-d_1)^2+(y-d_2)^2\right)^{-3/2} - (x+d_1 + y+d_2)\left((x+d_1)^2+(y+d_2)^2\right)^{-3/2}\\
-%%   &+ (x+d_1 + y-d_2)\left((x+d_1)^2+(y-d_2)^2\right)^{-3/2} - (x-d_1 + y+d_2)\left((x-d_1)^2+(y+d_2)^2\right)^{-3/2}\Big)\\
-%%   F(d_1,d_2) &= \frac{-q^2}{4\pi\epsilon} \Big(
-%%   - (2d_1 + 2d_2)\left(4d_1^2 + 4d_2^2\right)^{-3/2} + 2d_1\left(4d_1^2\right)^{-3/2} - 2d_2\left(4d_2^2\right)^{-3/2}\Big)\\
-%%   &= \frac{q^2}{4\pi\epsilon} \left(\frac{(2d_1 + 2d_2)}{\sqrt{\left(4d_1^2 + 4d_2^2\right)}^3} - \frac{1}{4d_1^2} - \frac{1}{4d_2^2}\right)\\
-%% \end{align*}
-
-\begin{align*}
-  \phi(\vec{r}) &= \frac{q}{\sqrt{2}4\pi\epsilon_0\sqrt{4d_1^2 + 4d_2^2}}\left(\hat{i} + \hat{j}\right) + \frac{-q}{4\pi\epsilon_0\sqrt{4d_1^2}}\hat{i} + \frac{-q}{4\pi\epsilon_0\sqrt{4d_2^2}}\hat{j}\\
-  &= \frac{q}{4\pi\epsilon_0}\left(\frac{1}{\sqrt{8d_1^2 + 8d_2^2}} - \frac{1}{\sqrt{4d_1^2}}\right)\hat{i}
-  + \frac{q}{4\pi\epsilon_0}\left(\frac{1}{\sqrt{8d_1^2 + 8d_2^2}} - \frac{1}{\sqrt{4d_2^2}}\right)\hat{j}
-\end{align*}
-
-If expanded, at large r, $\phi$ should look like a quadrupole scaling with $\frac{1}{r^3}$.\\
-
-\section{Problem Two}
-We want to show the following equality is always true:
-
-\begin{align*}
-  \int\rho_1\phi_2d^3\vec{r} &= \int\rho_2\phi_1d^3\vec{r}
-\end{align*}
-
-We first apply the Laplace Equation:
-
-\begin{align*}
-  -\epsilon_0\int\nabla^2\phi_1\phi_2d^3\vec{r} &= -\epsilon_0\int\nabla^2\phi_2\phi_1d^3\vec{r}
-\end{align*}
-
-We then expand this equation into its cartesian representation:
-
-\begin{align*}
-  \int\int\int\sum_{n=x,y,z}\frac{\partial^2}{\partial n^2} (\phi_1) \phi_2 dxdydz &=
-  \int\int\int\sum_{n=x,y,z}\frac{\partial^2}{\partial n^2} (\phi_2) \phi_1 dxdydz
-\end{align*}
-
-We then use Integration by Parts:
-
-\begin{align*}
-  \int\int\int\sum_{n=x,y,z}\left(\frac{d}{dn}\phi_1\phi_2\Big|_{-\infty}^{\infty} - \int\frac{\partial}{\partial n}\phi_1\frac{\partial}{\partial n}\phi_2dn \right) dxdydz &=
-  \int\int\int\sum_{n=x,y,z}\left(\frac{d}{dn}\phi_2\phi_1\Big|_{-\infty}^{\infty} - \int\frac{\partial}{\partial n}\phi_2\frac{\partial}{\partial n}\phi_1dn \right) dxdydz
-\end{align*}
-
-\section{Problem Three}
-The solution to the full sphere in an external field $E_0$ was shown in class to be:
-
-\begin{align*}
-  \phi(r,\theta) &= \left(-E_0r + \frac{E_0a^3}{r^2}\right)\cos\theta
-\end{align*}
-
-given boundaries $\phi(r=a) = 0$ and $\phi(r\rightarrow\infty) = 0$.
-We notice that our boundaries for the half sphere problem are the same, plus one addition: $\phi(r<a, z=0) = 0$.
-The above solution satisfies this condition as well.
-Then by the Uniqueness Theorem this solution must also be valid for our half sphere.\\
-
-The first term $-E_0r\cos\theta$ is the potential from the external field. The second term $\frac{E_0a^3\cos\theta}{r^2}$ is the term due
-to the sphere. This term can be rewritten as the dipole term in the multipole expansion:\\
-
-\begin{align*}
-  \frac{E_0a\cos\theta}{r^2} &= \frac{1}{4\pi\epsilon_0}\frac{1}{r^2} \hat{r} \cdot \vec{p}\\
-\end{align*}
-
-where $\vec{p} = 4\pi\epsilon_0E_0 a^3 \hat{z}$. Thus the half sphere can be exactly expressed as a dipole.\\
+The change in permittivity constants causes a change in curl.\\
 

### 2017-10-15 ch362/latex/report3.tex

@@
 Mass spectrometry (MS) and NMR are invaluable techniques for structure elucidation of unknown organic molecules.
-When a molecule is sufficiently simple, knowledge of its molecular weight, carbon number, and rough identity of carbons and hydrogens can be enough to uniquely determine identity.
+When a molecule is sufficiently simple, knowledge of its molecular weight, carbon number, and the shifts of its carbons and hydrogens can be enough to uniquely determine identity.
 For more complex molecules, or those with many isomers, further information from 2D experiments like $^{13}$C-$^{1}$H HSQC or $^{13}$C-$^{1}$H HMBC may be necessary.\\
@@
 This article follows the designation of a disubstituted benzene unknown using the above techniques.
-MS is used to determine weight, $^{13}$C NMR is used to come up with a candidate structure, $^{1}$H NMR is used to verify this candidate and guess its substitution, and $^{13}$C-$^{1}$H HSQC is used to verify the substitution.\\
+MS is used to determine weight, $^{13}$C NMR is used to come up with a candidate structure, $^{1}$H NMR is used to verify this candidate and guess its substitution, and $^{13}$C-$^{1}$H HSQC/HMBC are used to verify the substitution.\\
 
@@
 The $^{13}$C NMR spectra shown in Figure \ref{fig:CNMR} has 9 peaks.
-As shown in Table \ref{table:MS}, one peak is due to a DMSO solvent, six to aromatic carbons, one to a methyl carbon and one to a carboxylic carbon.
+Comparison with online shift reports [2] indicate one peak is due to a DMSO solvent, six to aromatic carbons, one to a methyl carbon and one to a carboxylic carbon, as shown in Table \ref{table:MS}.
 These results agree with the 8-carbon estimate from peak ratios in MS.\\
@@
 \centering
-  \includegraphics[height=5cm]{figures/CNMR.png}
+  \includegraphics[height=4.5cm]{figures/CNMR.png}
   \caption{$^{13}$C NMR of \textit{Sample 8}.}
@@
 \section{$^{1}$H NMR}
+$^{1}$H NMR is then used to get more information on substitutuent placement.
+A spectra is shown in Figure \ref{fig:HNMR} with annotated peaks shown in Table \ref{table:HNMR} indicating the presence of a carboxylic hydrogen at 12.8ppm and methyl hydrogens at 2.3ppm [2], verifying the methyl benzoic acid hypothesis.
+The splitting pattern of the aromatic hydrogens shown in Figure \ref{fig:HNMR-closeup} is used to determine configuration.
+Because four unique aromatic peaks are visible, the \textit{para} configuration is impossible.
+The presence of a singlet aromatic peak rules out \textit{ortho} and suggests \textit{meta} as the substitution pattern.\\
+
+The \textit{meta} designation is further evidenced by the presence of very small, 1-Hz meta coupling of the two doublet hydrogens.
+This coupling is expected between the singlet hydrogen and the two hydrogens adjacent to only one substitute group.\\
 
@@
 \centering
-  \includegraphics[height=5cm]{figures/HNMR.png}
+  \includegraphics[height=4cm]{figures/HNMR.png}
   \caption{$^{1}$H NMR of \textit{Sample 8}.}
@@
 
+\begin{figure}[h]
+\centering
+  \includegraphics[height=1.2cm]{figures/HNMR-splitting.png}
+  \caption{Close-up of aromatic peaks in $^{1}$H NMR spectra of \textit{Sample 8}.}
+  \label{fig:HNMR-closeup}
+\end{figure}
+
 \begin{table}[h]
 \small
-  \caption{$^{13}$C NMR identified peaks for spectra shown in Figure \ref{fig:HNMR}.}
+  \caption{$^{1}$H NMR identified peaks for spectra shown in Figure \ref{fig:HNMR}.}
   \label{table:HNMR}
@@
     \hline
-    Peak(s) shift(s) ppm & Integration & Splitting & Designation \\
+    Shift(s) ppm & Integration & Splitting & Designation \\
     \hline
@@
 
-\section{Equations}
+\section{COSY}
+The COSY spectrum of \textit{Sample 8} is not terribly useful. It verifies that aromatic hydrogens couple to one another, but HMBC provides much more information.
+
+\begin{figure}[h]
+\centering
+  \includegraphics[height=7cm]{figures/cosy.png}
+  \caption{COSY spectrum of \textit{Sample 8}.}
+  \label{fig:COSY}
+\end{figure}
+
+\section{HSQC}
+HSQC is similarly barren of information.
+The spectrum shown in Figure \ref{fig:HSQC} shows expected coupling between the solvent carbon and hydrogen, between the methyl carbon and methyl protons, and between aromatic carbons and aromatic hydrogens, as expected.
+More conclusively demonstrated is the presence of four different aromatic hydrogens, further verifying the molecule is not in the para configuration.\\
+
+\begin{figure}[h]
+\centering
+  \includegraphics[height=7cm]{figures/HSQC.png}
+  \caption{HSQC spectrum of \textit{Sample 8}. A blowup of the aromatic region is shown in the middle of the spectrum.}
+  \label{fig:HSQC}
+\end{figure}
+
+\section{HMBC}
+Now that the structure has been determined, individual carbons and hydrogens can be aligned with their respective peaks.
+Figure \ref{fig:3MBA} shows the candidate molecule along with labels for carbons and hydrogens.\\
+
+\begin{figure}[h]
+\centering
+  \includegraphics[height=5cm]{figures/molecule.png}
+  \caption{3-methylbenzoic acid.}
+  \label{fig:3MBA}
+\end{figure}
+
+Hydrogen peak placement is easy.
+The singlet hydrogen must correspond to the carbon between the two substituents.
+The triplet hydrogen must correspond to the hydrogen diametrically opposed to this carbon.
+This leaves the two doublet carbons.
+They can be distinguished by looking at the HMBC spectra shown in Figure \ref{fig:hydromethyl}.
+Hydrogen H2, indicated in Table \ref{table:H-peaks}, or the upfield doublet, correlates well with the methyl carbon.
+Carbon H4 is identified by exclusion.\\
+
+An HMBC-HSQC overlay in the aromatic region provides much information about aromatic carbon peaks.
+Such a spectrum is shown in Figure \ref{fig:overlay}.
+Carbons C2, C4, C5 and C6 can be easily determined as the carbons which HSQC-correlate with the respective hydrogens H1, H2, H3 and H4, as shown in Table \ref{table:C-peaks}.
+This leaves the two substituent aromatic carbons to label.
+Figure \ref{fig:methylaromatic} indicates three carbons pair strongly with the methyl hydrogen: carbons C2 and C4, already labeled, and one of the two substituent carbons shifted furthest downfield of the aromatic carbons.
+This is likely carbon C1, since 4-bond coupling is much stronger in HMBC spectra than 2-bond coupling.
+Thus C1 must be the carboxylic substituent aromatic carbon, and thus C3 must be the methyl aromatic carbon.
+These findings are catalogued in the tables below.\\
+
+\begin{figure}[h]
+\centering
+  \includegraphics[height=5cm]{figures/HMBC-hydromethyl.png}
+  \caption{A closeup of methyl carbon-aromatic hydrogen HMBC coupling.}
+  \label{fig:hydromethyl}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+  \includegraphics[height=5cm]{figures/hmbc-methylaromatic.png}
+  \caption{A closeup, edited image hilighting HMBC coupling between aromatic carbons and methyl hydrogens.}
+  \label{fig:methylaromatic}
+\end{figure}
+
+\begin{figure*}
+\centering
+  \includegraphics[height=10cm]{figures/Overlay.png}
+  \caption{HSQC-HMBC overlay in the C, H aromatic region.}
+  \label{fig:overlay}
+\end{figure*}
+
+\begin{table}[h]
+\small
+  \caption{Aromatic hydrogen peak identification. In correspondence with the labels shown in Figure \ref{fig:3MBA}.}
+  \label{table:H-peaks}
+  \begin{tabular*}{0.5\textwidth}{@{\extracolsep{\fill}}llll}
+    \hline
+    Shift(s) ppm & Designation \\
+    \hline
+    7.35 & H3 triplet hydrogen\\
+    7.4  & H2 doublet hydrogen\\
+    7.72 & H4 doublet hydrogen\\
+    7.75 & H1 singlet hydrogen\\
+    \hline
+  \end{tabular*}
+\end{table}
+
+\begin{table}[h]
+\small
+  \caption{Aromatic carbon peak identification. In correspondence with the labels shown in Figure \ref{fig:3MBA}.}
+  \label{table:C-peaks}
+  \begin{tabular*}{0.5\textwidth}{@{\extracolsep{\fill}}llll}
+    \hline
+    Shift(s) ppm & Designation \\
+    \hline
+    127 & Carbon 6\\
+    129 & Carbon 5\\
+    130 & Carbon 2\\
+    131 & Carbon 3\\
+    134 & Carbon 4\\
+    139 & Carbon 1\\
+    \hline
+  \end{tabular*}
+\end{table}
 
@@
 
-\footnotesize{
-\bibliography{rsc} %your .bib file
-\bibliographystyle{rsc} %the RSC's .bst file
-}
+\begin{@twocolumnfalse}
+\section{References}
+[1] Experimental Chemistry I: CH 362, OSU Department of Chemistry 2017\\
 
+[2] OSU of Chemistry. 1H NMR Chemical Shifts. http://www.science.oregonstate.edu/~gablek/CH335/Chapter10/ChemicalShift.htm. (Accessed March 22, 2017)\\
+\end{@twocolumnfalse}
 \end{document}

### 2017-11-04 em/latex/HW4.tex
HW4

--- 2016-10-19 em/latex/HW3.tex
@@
 
-\newcommand{\b}[1]{\boldsymbol{#1}}
+\newcommand{\bb}[1]{\boldsymbol{#1}}
 \newcommand{\bv}[1]{\boldsymbol{\vec{#1}}}
@@
 
-\section{Problem One}
-
 \section{Problem Two}
-\textbf{Part I}
-
-We begin with equation (2):
+Use an Ampere loop to calculate the H-field in the two magnetic materials:\\
 
 \begin{align*}
-  m &= \frac{I}{2}\oint \boldsymbol{r} \times d\boldsymbol{l}
+  2\pi r H &= I\hat{z}\\
+  H &= \frac{I\hat{z}}{2\pi r}
 \end{align*}
 
-This is a loop integral around a closed circle of radius R. $d\boldsymbol{l}$ always points tangential to the circle, so the inner
-cross product becomes:
+We can use this field, along with $M=\chi H$, to find M:\\
 
 \begin{align*}
-  m &= \frac{I}{2}\oint R\hat{n} dl
+  M_1 &= \frac{\chi_1I\hat{z}}{2\pi r}\\
+  M_2 &= \frac{\chi_2I\hat{z}}{2\pi r}
 \end{align*}
 
-where $\hat{n}$ is perpendicular to both $\boldsymbol{r}$ and $d\boldsymbol{l}$. We are now integrating a constant around a circle of circumference $2\pi R$, which turns to:
+We then find the bound currents:
 
 \begin{align*}
-  m &= \frac{I}{2}2\pi R^2 = I\pi R^2\hat{n} = IS\hat{n}
+  J_{b} &= 0\\
+  K_{b,1,in} &= \frac{\chi_1I\hat{z}}{2\pi a_1}
+  K_{b,2,in} &= \frac{-\chi_2I\hat{z}}{2\pi a_1}
+  K_{b,2,out} &= \frac{\chi_2I\hat{z}}{2\pi a_2}
 \end{align*}
 
-\textbf{Part II}
-
-We begin with the definition of torque, plug in the definition of force on a current, apply the cross product triple product rule,
-argue that one of the terms goes to zero, then apply the equation from Griffiths 4th ed 1.62e.
+Use an Amperian loop to find the B-field:\\
 
 \begin{align*}
-  \boldsymbol{N} &= \oint \bv{r} \times \bv{F}dl\\
-  &= \oint \bv{r} \times \left(\bv{I} \times \bv{B}\right)dl\\
-  &= \oint \bv{I}\left(\bv{r} \cdot \bv{B}\right) - \bv{B}\left(\bv{r} \cdot \bv{I}\right)dl \hspace{1.2cm}\mbox{cross triple product}\\
-  &= \oint \bv{I}\left(\bv{r} \cdot \bv{B}\right)dl - \oint\bv{B}\left(\bv{r} \cdot \bv{I}\right)dl\\
-  &= \oint \bv{I}\left(\bv{r} \cdot \bv{B}\right)dl \hspace{3cm}\mbox{since }\Big(\oint \bv{r} \cdot d\bv{l} = 0\Big)\\
-  &= \oint \bv{I}\left(\bv{B} \cdot \bv{r}\right)dl\\
-  &= I\oint \left(\bv{B} \cdot \bv{r}\right)d\boldsymbol{l}\\
-  &= I\bv{a} \times \bv{B} = \bv{m} \times \bv{B}\\
-\end{align*}
-
-as desired.
-
-\section{Problem Three}
-The force F between two dipoles is given by:
-
-\begin{align*}
-  \vec{F} = \frac{3\mu_0}{4\pi |r|^4}((\hat{r} \times \vec{m}_1) \times \vec{m}_2 + (\hat{r} \times \vec{m_2}) \times \vec{m_1}
-  - 2\hat{r}(\vec{m}_1 \cdot \vec{m}_2) + 5\hat{r}((\hat{r} \times \vec{m}_1) \cdot (\hat{r} \times \vec{m}_2)))
-\end{align*}
-
-We use the following coordinate system:
-
-\begin{itemize}
-\item $\vec{m}_1$ points in the $\hat{z}$ direction
-\item $\vec{m}_2$ points $\phi$ away from $\hat{z}$ and is \textit{coplanar} to $\hat{z}$. This is a simplifying assumption.
-\item $\vec{d}$ points from the position of ball 1 to ball 2, and is angle $\theta$ from $\hat{z}$
-\end{itemize}
-
-Thus the following are true:
-
-\begin{align*}
-  \hat{r} \times \vec{m}_1 &= m_1\sin\theta\hat{x}\\
-  \hat{r} \times \vec{m}_2 &= m_2\sin\phi\hat{x}\\
-  \vec{m}_1 \cdot \hat{r} &= m_1\cos\theta\\
-  \vec{m}_1 \cdot \vec{m}_2 &= m_1m_2\cos\phi
-\end{align*}
-
-This allows us to simplify $\vec{F}$ to:
-
-\begin{align*}
-  \vec{F} &= \frac{3\mu_0}{4\pi d^4} \left(\left(m_1\sin\theta\hat{x} \times \vec{m}_2\right)
-  + \left(m_2\sin\phi\hat{x} \times \vec{m}_1\right)
-  - 2m_1m_2\left(\sin\theta + \cos\phi\right)\hat{r} + 5\left(m_1\sin\theta\hat{x} \cdot m_2\sin\phi\hat{x}\left)\hat{r}
-  \right)\\
-  &= \frac{3\mu_0}{4\pi d^4} \left(-2m_1m_2\sin\theta\hat{r} - 2m_1m_2\cos\phi\hat{r} + 5m_1m_2\sin\theta\sin\phi\hat{r}\right)\\
-  &= \frac{3\mu_1m_1m_2}{4\pi d^4}\left(-2\left(\sin\theta+\cos\phi\right) + 5\sin\theta\sin\phi\right)\hat{r}
-\end{align*}
-
-This indicates that the dipole-dipole force will be of order $\frac{3\mu_0m_1m_2}{4\pi d^4}$ and will depend on $\theta$ and $\phi$.
-We see that the force is maximized when vectors $\vec{m}_1$ and $\vec{m}_2$ are perpendicular to one another.\\
-
-Now we need to find the values of $\vec{m}_1$ and $\vec{m}_2$. We say that at far distances, the dipole moment will be dominated
-by the outer-rim charge.
-
-\begin{align*}
-  \vec{m} &= I\vec{s} = (2\pi R)Q_{rim} * \pi R^2\\
-  Q_{rim} &= \frac{Q}{V} * 2\pi R = \frac{3Q}{4\pi R^3} * 2\pi R = \frac{3Q}{2R^2}\\
-  \vec{m} &= I\vec{s} = 3\pi^2 RQ\\
-\end{align*}
-
-Thus the total magnetic force will be something like:
-
-\begin{align*}
-  F \approx \frac{27\mu_0\pi^3R^2Q^2}{4 d^4}
+  B_{1} &= 0\\
+  B_{2} &= \frac{I(1+\chi_1-\chi_2)}{2\pi r}\hat{\theta}\\
+  B_{2} &= \frac{I(1+\chi_1-\chi_2)}{2\pi r}\hat{\theta} + \frac{\chi_2I\hat{z}}{2\pi r}\\
 \end{align*}

### 2017-11-08 em/latex/HW1.tex
stuff

@@
 
-Guess the solution is four charges in a rectangle. If this solution works, by the Uniqueness Theorem it is the only solution.
+Guess the solution is four charges in a rectangle. If this solution works, by the uniqueness theorem it is the only solution.
 
@@
 
-given boundaries $\phi(r=a) = 0$ and $\phi(r\rightarrow\infty) = 0$.
+given boundaries $\phi(r=a) = 0$ and $\phi(r\rightarrow\infty) = E_0r\cos\theta$.
 We notice that our boundaries for the half sphere problem are the same, plus one addition: $\phi(r<a, z=0) = 0$.
 The above solution satisfies this condition as well.
-Then by the Uniqueness Theorem this solution must also be valid for our half sphere.\\
+Then by the uniqueness theorem this solution must also be valid for our half sphere.\\
 
 The first term $-E_0r\cos\theta$ is the potential from the external field. The second term $\frac{E_0a^3\cos\theta}{r^2}$ is the term due
-to the sphere. This term can be rewritten as the dipole term in the multipole expansion:
-\begin{align*}
-  \frac{E_0a\cos\theta}{r^2} &= \frac{1}{4\pi\epsilon_0}\frac{1}{r^2} \hat{r} \cdot \vec{p}
-\end{align*}
-
-where $\vec{p} = 4\pi\epsilon_0E_0 a^3 \hat{z}$. Thus the half sphere can be exactly expressed as a dipole.\\
+to the sphere. This term is the second term in a multipole expansion, the dipole term.\\
 

### 2017-11-08 em/latex/HW4.tex
stuff

@@
 
-\title{Elegromagnetism Assignment Three}
+\title{Elegromagnetism Assignment Four}
 \author{Elliott Capek}
@@
 
-\newcommand{\bb}[1]{\boldsymbol{#1}}
-\newcommand{\bv}[1]{\boldsymbol{\vec{#1}}}
-
 \begin{document}
@@
 
-\section{Problem Two}
-Use an Ampere loop to calculate the H-field in the two magnetic materials:\\
+\section{Problem One}
+We solve this problem through the method of images. Define a current $I_g\hat{y}$ which flows at $(0,0,-d)$. First find the bound surface current, assuminig linear magnetic material:\\
 
 \begin{align*}
-  2\pi r H &= I\hat{z}\\
-  H &= \frac{I\hat{z}}{2\pi r}
+  K_{1,b}(0,0,0) &= \vec{M} \mbox{x} \vec{n} = \frac{\chi_1}{\mu_1}\vec{B} \mbox{x} \vec{n} = \frac{\chi_1}{\mu_1}\frac{\mu_1I}{2\pi d} = \frac{\chi_1I}{2\pi d}\\
+  K_{2,b}(0,0,0) &= \frac{-\chi_2I}{2\pi d}\\
+  K_{b} &= \frac{(\chi_1-\chi_2)I}{2\pi d}\\
 \end{align*}
 
-We can use this field, along with $M=\chi H$, to find M:\\
+The A-field will be that due to infinite wires of constant current.
+The non-mirror magnetic material will see currents of $I$ in the material and $I_g$ in the other material.
+There cannot be any image charge in the mirror material, so we note that $I+I_g$ in the non-mirror material is the only current which satisfies the continuity boundary conditions for the mirror material's A-field:\\
 
 \begin{align*}
-  M_1 &= \frac{\chi_1I\hat{z}}{2\pi r}\\
-  M_2 &= \frac{\chi_2I\hat{z}}{2\pi r}
+  A_1 &= \frac{\mu_1I\ln{\sqrt{x^2+y^2+(z-d)^2}}}{4\pi} + \frac{\mu_1I_g\ln{\sqrt{x^2+y^2+(z+d)^2}}}{4\pi}\\
+  A_2 &= \frac{\mu_1(I+I_g)\ln{\sqrt{x^2+y^2+(z-d)^2}}}{4\pi}\\
+  &A_1(x,y,0) = A_2(x,y,0)\\
 \end{align*}
 
-We then find the bound currents:
+We can calculate the value of $I_g$ through the derivative continuity boundary condition of A:
 
 \begin{align*}
-  J_{b} &= 0\\
-  K_{b,1,in} &= \frac{\chi_1I\hat{z}}{2\pi a_1}
-  K_{b,2,in} &= \frac{-\chi_2I\hat{z}}{2\pi a_1}
-  K_{b,2,out} &= \frac{\chi_2I\hat{z}}{2\pi a_2}
+  &\frac{\partial A_1}{\partial n}_{z=0} - \frac{\partial A_2}{\partial n}_{z=0} = -\mu_1 K\\
+  &\frac{\partial A_1}{\partial n}(0,0,z) = \frac{\mu_1I}{4\pi(z-d)} + \frac{\mu_1I_g}{4\pi(z+d)}\\
+  &\frac{\partial A_2}{\partial n}(0,0,z) = \frac{\mu_1(I+I_g)}{4\pi(z-d)}\\
+  &\frac{-\mu_1I}{4\pi(d)} + \frac{\mu_1I_g}{4\pi(d)} + \frac{\mu_1(I+I_g)}{4\pi(d)} = -\mu_0K_b\\
+  &\frac{\mu_1I_g}{2\pi d} = \frac{(\chi_1-\chi_2)I}{2\pi d}\\
+  &I_g = \frac{-\mu_0(\chi_1-\chi_2)I}{\mu_1} = \frac{-\mu_0}{\mu_1}\left(\frac{\mu_1}{\mu_0}-1-\frac{\mu_2}{\mu_0}+1\right)I\\
+  &I_g = \frac{\mu_2-\mu_1}{\mu_1}I\\
 \end{align*}
 
-Use an Amperian loop to find the B-field:\\
+Thus the B-fields can be computed as:\\
 
 \begin{align*}
-  B_{1} &= 0\\
-  B_{2} &= \frac{I(1+\chi_1-\chi_2)}{2\pi r}\hat{\theta}\\
-  B_{2} &= \frac{I(1+\chi_1-\chi_2)}{2\pi r}\hat{\theta} + \frac{\chi_2I\hat{z}}{2\pi r}\\
+  B_1(x,y,z) &= \frac{\mu_1I}{2\pi\sqrt{x^2+y^2+(z-d)^2}} + \frac{(\mu_2-\mu_1)I}{2\pi\sqrt{x^2+y^2+(z+d)^2}}\\
+  B_2(x,y,z) &= \frac{\mu_2I}{2\pi\sqrt{x^2+y^2+(z-d)^2}}\\
+\end{align*}
+
+\section{Problem 2}
+Use an Amperian loop of radius $s$ to calculate the H-field in and out of the cable:\\
+
+\begin{align*}
+  2\pi sH &= \pi s^2 I\\
+  H &= \frac{sI}{2} \hspace{1cm} s<a_2\\
+  H &= \frac{a_2^2I}{2r} \hspace{1cm} r>a_2\\
+\end{align*}
+
+By the definition $M=\chi H$, we can find the M-field:\\
+
+\begin{align*}
+  M_1 &= \frac{\chi_1 sI}{2}\\
+  M_2 &= \frac{\chi_2 sI}{2}\\
+  M_{out} &= 0\\
+\end{align*}
+
+This field is curl-free, thus there is no bound volume current. The surface bound current is given by:\\
+
+\begin{align*}
+  K_{b1} &= \frac{\chi_1 a_1 I\hat{z}}{2}\\
+  K_{b2,in} &= \frac{-\chi_2 a_1 I\hat{z}}{2}\\
+  K_{b2,out} &= \frac{\chi_2 a_2 I\hat{z}}{2}\\
+\end{align*}
+
+We can use an Amperian loop to find the B-fields:
+
+\begin{align*}
+  B_{in} &= 0\\
+  B_{mid} &= \left(\frac{I}{2\pi s}\right)\left(K_{b1}*2\pi a_1 + K_{b2,in}*2\pi a_1\right)\\
+  &= \frac{1}{2\pi s}\left(\frac{\chi_1 a_1 I\hat{z}}{2} * 2\pi a_1 + \frac{-\chi_2 a_1 I \hat{z}}{2}*2\pi a_1\right)\\
+  &= \frac{a_1^2I}{2s}\left(\chi_1-\chi_2\right)\hat{\theta}\\
+  B_{out} &= \frac{I}{2s}\left(a_1^2\chi_1 - a_1^2\chi_2 + a_2^2\chi_2\right)\\
 \end{align*}

### 2017-11-08 sm/bacteria.py

@@ -1,0 +1,72 @@
+#!/usr/bin/env python
+
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+
+simulations = 10000
+
+length10dist = np.zeros(1001)
+length = 10
+for _ in range(simulations):
+    R = 500
+    B = 500
+    for _ in range(length):
+        R = R*2
+        B = B*2
+        r = np.random.hypergeometric(R, B, 1000)
+        R = R - r
+        B = B - (1000-r)
+        length10dist[R] += 1
+length10dist = length10dist / np.sum(length10dist)
+        
+length100dist = np.zeros(1001)
+length = 100
+for _ in range(simulations):
+    R = 500
+    B = 500
+    for _ in range(length):
+        R = R*2
+        B = B*2
+        r = np.random.hypergeometric(R, B, 1000)
+        R = R - r
+        B = B - (1000-r)
+        length100dist[R] += 1
+length100dist = length100dist / np.sum(length100dist)
+
+length1000dist = np.zeros(1001)
+length = 1000
+for _ in range(simulations):
+    R = 500
+    B = 500
+    for _ in range(length):
+        R = R*2
+        B = B*2
+        r = np.random.hypergeometric(R, B, 1000)
+        R = R - r
+        B = B - (1000-r)
+        length1000dist[R] += 1
+length1000dist = length1000dist / np.sum(length1000dist)
+
+# length10000dist = np.zeros(1001)
+# length = 10000
+# for _ in range(simulations):
+#     R = 500
+#     B = 500
+#     for _ in range(length):
+#         R = R*2
+#         B = B*2
+#         r = np.random.hypergeometric(R, B, 1000)
+#         R = R - r
+#         B = B - (1000-r)
+#         length10000dist[R] += 1
+# length10000dist = length10000dist / np.sum(length10000dist)
+
+plt.figure()
+plt.plot(list(range(1001)), length10dist, label="10h")
+plt.plot(list(range(1001)), length100dist, label="100h")
+plt.plot(list(range(1001)), length1000dist, label="1000h")
+# plt.plot(list(range(1001)), length10000dist, label="10000h")
+plt.legend()
+plt.title("Red bacteria probability distribution after n hours")
+plt.savefig("distribution.png")

### 2017-11-17 em/latex/HW6.tex
HW6

--- 2016-10-19 em/latex/HW3.tex
@@
 
-\title{Elegromagnetism Assignment Three}
+\title{Elegromagnetism Assignment Six}
 \author{Elliott Capek}
@@
 
-\newcommand{\b}[1]{\boldsymbol{#1}}
-\newcommand{\bv}[1]{\boldsymbol{\vec{#1}}}
-
 \begin{document}
@@
 
-\section{Problem One}
-
-\section{Problem Two}
-\textbf{Part I}
-
-We begin with equation (2):
+\section{Problem One: Plane wave solution using Lorentz gauge}
+The Lorenz gauge condition can be rephrased as:\\
 
 \begin{align*}
-  m &= \frac{I}{2}\oint \boldsymbol{r} \times d\boldsymbol{l}
+  \frac{1}{c^2}\frac{\partial^2\phi}{\partial t^2} - \nabla^2 \phi &= \frac{\rho}{\epsilon_0}\\
+  \frac{1}{c^2}\frac{\partial^2\vec{A}}{\partial t^2} - \nabla^2 \vec{A} &= \vec{J}\mu_0\\
 \end{align*}
 
-This is a loop integral around a closed circle of radius R. $d\boldsymbol{l}$ always points tangential to the circle, so the inner
-cross product becomes:
+In a vacuum, these equations turn into wave equations for $\phi$ and $\vec{A}$:\\
 
 \begin{align*}
-  m &= \frac{I}{2}\oint R\hat{n} dl
+  \frac{1}{c^2}\frac{\partial^2\phi}{\partial t^2} &= \nabla^2 \phi\\
+  \frac{1}{c^2}\frac{\partial^2\vec{A}}{\partial t^2} &= \nabla^2 \vec{A}\\
 \end{align*}
 
-where $\hat{n}$ is perpendicular to both $\boldsymbol{r}$ and $d\boldsymbol{l}$. We are now integrating a constant around a circle of circumference $2\pi R$, which turns to:
+Thus equations for $\phi$ and $\vec{A}$ can be written:\\
 
 \begin{align*}
-  m &= \frac{I}{2}2\pi R^2 = I\pi R^2\hat{n} = IS\hat{n}
+  \phi(\vec{r},t) &= \phi_0e^{i\left(\vec{k}\cdot\vec{r} -\omega t\right)}\\
+  \vec{A}(\vec{r},t) &= \vec{A_0}e^{i\left(\vec{k}\cdot\vec{r} -\omega t\right)}
 \end{align*}
 
-\textbf{Part II}
-
-We begin with the definition of torque, plug in the definition of force on a current, apply the cross product triple product rule,
-argue that one of the terms goes to zero, then apply the equation from Griffiths 4th ed 1.62e.
+where $c=\omega/|k|$. We can then find an expression for $\vec{E}$:\\
 
 \begin{align*}
-  \boldsymbol{N} &= \oint \bv{r} \times \bv{F}dl\\
-  &= \oint \bv{r} \times \left(\bv{I} \times \bv{B}\right)dl\\
-  &= \oint \bv{I}\left(\bv{r} \cdot \bv{B}\right) - \bv{B}\left(\bv{r} \cdot \bv{I}\right)dl \hspace{1.2cm}\mbox{cross triple product}\\
-  &= \oint \bv{I}\left(\bv{r} \cdot \bv{B}\right)dl - \oint\bv{B}\left(\bv{r} \cdot \bv{I}\right)dl\\
-  &= \oint \bv{I}\left(\bv{r} \cdot \bv{B}\right)dl \hspace{3cm}\mbox{since }\Big(\oint \bv{r} \cdot d\bv{l} = 0\Big)\\
-  &= \oint \bv{I}\left(\bv{B} \cdot \bv{r}\right)dl\\
-  &= I\oint \left(\bv{B} \cdot \bv{r}\right)d\boldsymbol{l}\\
-  &= I\bv{a} \times \bv{B} = \bv{m} \times \bv{B}\\
+  \vec{E} &= -\nabla\phi - \frac{d\vec{A}}{dt}\\
+  &= -\vec{k}\phi + i\omega\vec{A}\\
+  &= -k\phi_0\cos(\vec{k}\cdot\vec{r} - \omega t)\hat{k} - \omega A_0\sin(\vec{k}\cdot\vec{r}-\omega t)\hat{A}\\
 \end{align*}
 
-as desired.
-
-\section{Problem Three}
-The force F between two dipoles is given by:
+Thus, unlike for the non-Lorentz gauge solution, $\vec{E}$ and $k$ are not perpendicular. Next we find B:\\
 
 \begin{align*}
-  \vec{F} = \frac{3\mu_0}{4\pi |r|^4}((\hat{r} \times \vec{m}_1) \times \vec{m}_2 + (\hat{r} \times \vec{m_2}) \times \vec{m_1}
-  - 2\hat{r}(\vec{m}_1 \cdot \vec{m}_2) + 5\hat{r}((\hat{r} \times \vec{m}_1) \cdot (\hat{r} \times \vec{m}_2)))
+  B &= \nabla \times \vec{A}\\
+  &= kA_0\cos(\vec{k}\cdot\vec{r}-\omega t)(\hat{k}\times\hat{A_0})\\
 \end{align*}
 
-We use the following coordinate system:
-
-\begin{itemize}
-\item $\vec{m}_1$ points in the $\hat{z}$ direction
-\item $\vec{m}_2$ points $\phi$ away from $\hat{z}$ and is \textit{coplanar} to $\hat{z}$. This is a simplifying assumption.
-\item $\vec{d}$ points from the position of ball 1 to ball 2, and is angle $\theta$ from $\hat{z}$
-\end{itemize}
-
-Thus the following are true:
+Thus $\vec{E}$ and $\vec{B}$ are perpendicular, as expected. To find energy flux, we calculate the Poynting vector:\\
 
 \begin{align*}
-  \hat{r} \times \vec{m}_1 &= m_1\sin\theta\hat{x}\\
-  \hat{r} \times \vec{m}_2 &= m_2\sin\phi\hat{x}\\
-  \vec{m}_1 \cdot \hat{r} &= m_1\cos\theta\\
-  \vec{m}_1 \cdot \vec{m}_2 &= m_1m_2\cos\phi
-\end{align*}
-
-This allows us to simplify $\vec{F}$ to:
-
-\begin{align*}
-  \vec{F} &= \frac{3\mu_0}{4\pi d^4} \left(\left(m_1\sin\theta\hat{x} \times \vec{m}_2\right)
-  + \left(m_2\sin\phi\hat{x} \times \vec{m}_1\right)
-  - 2m_1m_2\left(\sin\theta + \cos\phi\right)\hat{r} + 5\left(m_1\sin\theta\hat{x} \cdot m_2\sin\phi\hat{x}\left)\hat{r}
-  \right)\\
-  &= \frac{3\mu_0}{4\pi d^4} \left(-2m_1m_2\sin\theta\hat{r} - 2m_1m_2\cos\phi\hat{r} + 5m_1m_2\sin\theta\sin\phi\hat{r}\right)\\
-  &= \frac{3\mu_1m_1m_2}{4\pi d^4}\left(-2\left(\sin\theta+\cos\phi\right) + 5\sin\theta\sin\phi\right)\hat{r}
-\end{align*}
-
-This indicates that the dipole-dipole force will be of order $\frac{3\mu_0m_1m_2}{4\pi d^4}$ and will depend on $\theta$ and $\phi$.
-We see that the force is maximized when vectors $\vec{m}_1$ and $\vec{m}_2$ are perpendicular to one another.\\
-
-Now we need to find the values of $\vec{m}_1$ and $\vec{m}_2$. We say that at far distances, the dipole moment will be dominated
-by the outer-rim charge.
-
-\begin{align*}
-  \vec{m} &= I\vec{s} = (2\pi R)Q_{rim} * \pi R^2\\
-  Q_{rim} &= \frac{Q}{V} * 2\pi R = \frac{3Q}{4\pi R^3} * 2\pi R = \frac{3Q}{2R^2}\\
-  \vec{m} &= I\vec{s} = 3\pi^2 RQ\\
-\end{align*}
-
-Thus the total magnetic force will be something like:
-
-\begin{align*}
-  F \approx \frac{27\mu_0\pi^3R^2Q^2}{4 d^4}
+  \vec{S} &= \vec{E}\times\vec{B}\\
+  &= -k^2A_0\phi_0\cos^2(\vec{k}\cdot\vec{r}-\omega t)\hat{A_0} + -\omega A_0k\cos(\vec{k}\cdot\vec{r}-\omega t)\sin(\vec{k}\cdot\vec{r}-\omega t)\hat{k}\\
 \end{align*}

### 2017-11-17 em/latex/HW5.tex

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\nonstopmode

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{amsmath}

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\usepackage{amssymb}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Elegromagnetism Assignment Five}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\newcommand{\bra}[1]{\big<#1\big|}
\newcommand{\ket}[1]{\big|#1\big>}
\newcommand{\braket}[2]{\big<#1\big|#2\big>}

\begin{document}

\maketitle{}

\section{Problem One}
To solve for the current in a capacitor/nested inductor circuit, derive a differential equation for current using rules about inductors and capacitors, then solve it. First the inductance. Flux for two antiparallel superimposed inductors with loop densities $n_a$ and $n_b$, length $\ell$ and radius $a$ is given by:\\

\begin{align*}
  \Phi &= BA = \frac{\mu \left(N_a^2-N_b^2\right) I}{\ell}\pi a^2 = \mu \left(n_a^2-n_b^2\right)I\ell\pi a^2\\
\end{align*}

From one of Maxwell's equations:\\

\begin{align*}
  V_L &= \frac{d\Phi}{dt} = \mu \left(n_a^2-n_b^2\right)\ell\pi a^2\frac{d^2Q}{dt^2}\\
\end{align*}

We also know the ovltage drop across a parallel plate capacitor:\\

\begin{align*}
  V_c &= \frac{Q}{C} = \frac{Qd}{\epsilon_0 A} = \frac{Qd}{\epsilon_0 \pi a^2}\\
\end{align*}

Using Kirchoff's Loop Law:\\

\begin{align*}
  V_c &= -V_L\\
  \frac{Qd}{\epsilon_0 \pi a^2} &= \mu \left(n_a^2-n_b^2\right)\ell\pi a^2\frac{d^2Q}{dt^2}\\
  -\frac{d}{\pi^2 a^4\mu\left(n_a^2-n_b^2\right)\ell}Q &= \frac{d^2Q}{dt^2}\\
\end{align*}

The solution to which is, noting that Q must be real:\\

\begin{align*}
  Q(t) &= Q_0\cos(\sqrt{\frac{d}{\epsilon_0\pi^2 a^4\mu\left(n_a^2-n_b^2\right)\ell}}t) = \cos(\omega t)\\
  \omega &= \sqrt{\frac{d}{\epsilon_0\pi^2 a^4\mu\left(n_a^2-n_b^2\right)\ell}} = \sqrt{\frac{1}{LC}}\\
\end{align*}

\textbf{b.) Energy in circuit}
The individual component energies are:\\

\begin{align*}
  U_L &= \frac{1}{2}LI^2 = \left(\mu(n_a^2-n_b^2)\ell\pi a^2\right)\left(\frac{d}{\pi^2 a^4\mu\left(n_a^2-n_b^2\right)\ell}\right)Q_0^2\sin^2(\omega t)\\
  &= \left(\frac{d}{2\epsilon_0\pi a^2}\right)Q_0^2\sin^2(\omega t)\\
  U_C &= \frac{1}{2}\frac{Q^2}{C} = \frac{d}{2\epsilon_0 \pi a^2}Q_0^2\cos^2(\omega t)\\
\end{align*}

Thus:\\

\begin{align*}
  U &= \frac{Q_0^2d}{\epsilon_0 \pi a^2}(\cos^2(\omega t)+\sin^2(\omega t))\\
  &= \frac{Q_0^2d}{\epsilon_0 \pi a^2}\\
\end{align*}

\textbf{c.) Poynting vector}

Find the capacitor Poynting vector using oscillating E-field:

\begin{align*}
  V &= \frac{Q}{C}Ed = \frac{Q_0\cos(\omega t)d}{\epsilon_0 A}\\
  \vec{E} &= \frac{Q_0\cos(\omega t)}{\epsilon_0 A}
\end{align*}

The B-field in the capacitor can be found using Maxwell's Equations:\\

\begin{align*}
  \nabla \times B &= \mu_0\epsilon_0 \frac{dE}{dt} = -\frac{\mu_0\omega Q\sin(\omega t)\hat{z}}{A}\\
  \frac{1}{r}(\frac{d}{dr}(rB_\phi)-\frac{dB_r}{d\phi} &= \frac{d}{dr}(rB_\phi)\\
  \vec{B} &= \frac{r\mu_0\omega Q_0 \sin(\omega t)\hat{\vec}}{2A}\\
\end{align*}

Thus the Poynting vector is:\\

\begin{align*}
  \vec{S} &= \frac{1}{\mu_0}EB = \frac{Q^2}{2\epsilon_0A^2}r\omega \cos(\omega t)\sin(\omega t)\vec{r}
\end{align*}

The vector points radially; thus light will radiate from the capacitor during its operation, leaking energy.\\

\end{document}

### 2017-11-17 em/latex/HW6.tex
HW62

@@
 \begin{align*}
-  \vec{E} &= -\nabla\phi - \frac{d\vec{A}}{dt}\\
-  &= -\vec{k}\phi + i\omega\vec{A}\\
-  &= -k\phi_0\cos(\vec{k}\cdot\vec{r} - \omega t)\hat{k} - \omega A_0\sin(\vec{k}\cdot\vec{r}-\omega t)\hat{A}\\
+  \vec{E} &= -\frac{d\vec{A}}{dt}\\
+  &= i\omega\vec{A}\\
+  &= -\omega A_0\sin(\vec{k}\cdot\vec{r}-\omega t)\hat{A}\\
 \end{align*}
 
-Thus, unlike for the non-Lorentz gauge solution, $\vec{E}$ and $k$ are not perpendicular. Next we find B:\\
+Thus E is parallel to $A$ and perpendicular to $k$. Next we find B:\\
 
 \begin{align*}
-  B &= \nabla \times \vec{A}\\
+  B &= \nabla \times \vec{A} = i\vec{k}\times\vec{A}\\
   &= kA_0\cos(\vec{k}\cdot\vec{r}-\omega t)(\hat{k}\times\hat{A_0})\\
@@
 \begin{align*}
-  \vec{S} &= \vec{E}\times\vec{B}\\
-  &= -k^2A_0\phi_0\cos^2(\vec{k}\cdot\vec{r}-\omega t)\hat{A_0} + -\omega A_0k\cos(\vec{k}\cdot\vec{r}-\omega t)\sin(\vec{k}\cdot\vec{r}-\omega t)\hat{k}\\
+  \vec{S} &= \frac{1}{\mu_0}\vec{E}\times\vec{B}\\
+  &= -\frac{\omega}{\mu_0} A_0k\cos(\vec{k}\cdot\vec{r}-\omega t)\sin(\vec{k}\cdot\vec{r}-\omega t)\hat{k}\\
+\end{align*}
+
+\section{Problem Two: Coulomb gauge solution for complex A}
+First, note that in the Coulomb gauge $\nabla \times \vec{A} = 0$; thus, for $\vec{A} = \vec{A_0}e^{\vec{k}\times\vec{r}-\omega t}$, we see that $k\cdot A_0=0$.
+Because $\vec{A_0}=A_x\hat{x}+iA_y\hat{y}$, we see that $\vec{k}=k\hat{z}$.\\
+
+Next, use the equation $E=-\frac{d\vec{A}}{dt}$ to calculate $\vec{E}$:\\
+
+\begin{align*}
+  \vec{E} &= -\frac{d\vec{A}}{dt} = Re\left(-i\omega(A_x\hat{x}+iA_y\hat{y})e^{i(\vec{k}\cdot\vec{r}-\omega t)}\right)\\
+  &= -\omega A_yRe(e^{i(\vec{k}\cdot\vec{r}-\omega t)}\hat{y})\\
+  &= -\omega A_y\cos(\vec{k}\cdot\vec{r}-\omega t)\hat{y}\\
+\end{align*}
+
+Then we calculate $\vec{B}$, using $\vec{B}=\nabla \times \vec{A}$:\\
+
+\begin{align*}
+  \vec{B} &= \nabla \times \vec{A} = Re\left(i\vec{k} \times \vec{A}\right) = Re\left(i(\vec{k} \times A_xe^{i(\vec{k}\cdot\vec{r}-\omega t)})\hat{x} - (\vec{k} \times A_ye^{i(\vec{k}\cdot\vec{r}-\omega t)}\hat{y})\right)\\
+  &= -kA_yRe(e^{i(\vec{k}\cdot\vec{r}-\omega t)})\hat{x}\\
+  &= -kA_y\cos(\vec{k}\cdot\vec{r}-\omega t)\hat{x}\\
+\end{align*}
+
+We then calculate the Poynting vector:\\
+
+\begin{align*}
+  \vec{S} &= \frac{1}{\mu_0}\vec{E}\times\vec{B} = \frac{k}{\mu_0}\omega A_y^2\cos^2(\vec{k}\cdot\vec{r}-\omega t)\hat{z}\\
+\end{align*}
+
+And the energy density:\\
+
+\begin{align*}
+  U &= \epsilon_0 E^2 = \epsilon_0\omega^2A_y^2\cos^2(\vec{k}\cdot\vec{r}-\omega t)\\
 \end{align*}

### 2018-02-02 bb494/manuscript.tex
add bblab manuscript

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jacsat,manuscript=article]{achemso}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later. Do NOT use any
%% packages which require e-TeX (for example etoolbox): the e-TeX
%% extensions are not currently available on the ACS conversion
%% servers.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional macros here.  Please use \newcommand* where
%% possible, and avoid layout-changing macros (which are not used
%% when typesetting).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command. Phone and fax numbers can be given
%% using \phone and \fax, respectively; this information is optional.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%% On the other hand, \alsoaffiliation is used on a per author basis
%% when authors are associated with multiple institutions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Andrew N. Other}
\altaffiliation{A shared footnote}
\author{Fred T. Secondauthor}
\altaffiliation{Current address: Some other place, Othert\"own,
Germany}
\author{I. Ken Groupleader}
\altaffiliation{A shared footnote}
\email{i.k.groupleader@unknown.uu}
\phone{+123 (0)123 4445556}
\fax{+123 (0)123 4445557}
\affiliation[Unknown University]
{Department of Chemistry, Unknown University, Unknown Town}
\alsoaffiliation[Second University]
{Department of Chemistry, Second University, Nearby Town}
\author{Susanne K. Laborator}
\email{s.k.laborator@bigpharma.co}
\affiliation[BigPharma]
{Lead Discovery, BigPharma, Big Town, USA}
\author{Kay T. Finally}
\affiliation[Unknown University]
{Department of Chemistry, Unknown University, Unknown Town}
\alsoaffiliation[Second University]
{Department of Chemistry, Second University, Nearby Town}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[An \textsf{achemso} demo]
  {A demonstration of the \textsf{achemso} \LaTeX\
   class\footnote{A footnote for the title}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{IR,NMR,UV}
\keywords{American Chemical Society, \LaTeX}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The manuscript does not need to include \maketitle, which is
%% executed automatically.  The document should begin with an
%% abstract, if appropriate.  If one is given and should not be, the
%% contents will be gobbled.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  This is an example document for the \textsf{achemso} document
  class, intended for submissions to the American Chemical Society
  for publication. The class is based on the standard \LaTeXe\
  \textsf{report} file, and does not seek to reproduce the appearance
  of a published paper.

  This is an abstract for the \textsf{achemso} document class
  demonstration document.  An abstract is only allowed for certain
  manuscript types.  The selection of \texttt{journal} and
  \texttt{manuscript} will determine if an abstract is valid.  If
  not, the class will issue an appropriate error.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
This is a paragraph of text to fill the introduction of the
demonstration file.  The demonstration file attempts to show the
modifications of the standard \LaTeX\ macros that are implemented by
the \textsf{achemso} class.  These are mainly concerned with content,
as opposed to appearance.

\section{Results and discussion}

\subsection{Outline}

The document layout should follow the style of the journal concerned.
Where appropriate, sections and subsections should be added in the
normal way. If the class options are set correctly, warnings will be
given if these should not be present.

\subsection{References}

The class makes various changes to the way that references are
handled.  The class loads \textsf{natbib}, and also the
appropriate bibliography style.  References can be made using
the normal method; the citation should be placed before any
punctuation, as the class will move it if using a superscript
citation style
\cite{Mena2000,Abernethy2003,Friedman-Hill2003,EuropeanCommission2008}.
The use of \textsf{natbib} allows the use of the various citation
commands of that package: \citeauthor{Abernethy2003} have shown
something, in \citeyear{Cotton1999}, or as given by
Ref.~\citenum{Mena2000}.  Long lists of authors will be
automatically truncated in most article formats, but not in
supplementary information or reviews \cite{Pople2003}. If you
encounter problems with the citation macros, please check that
your copy of \textsf{natbib} is up to date. The demonstration
database file \texttt{achemso-demo.bib} shows how to complete
entries correctly. Notice that ``\latin{et al.}'' is auto-formatted
using the \texttt{\textbackslash latin} command.

Multiple citations to be combined into a list can be given as
a single citation.  This uses the \textsf{mciteplus} package
\cite{Johnson1972,*Arduengo1992,*Eisenstein2005,*Arduengo1994}.
Citations other than the first of the list should be indicated
with a star. If the \textsf{mciteplus} package is not installed,
the standard bibliography tools will still work but starred
references will be ignored. Individual references can be referred
to using \texttt{\textbackslash mciteSubRef}:
``ref.~\mciteSubRef{Eisenstein2005}''.

The class also handles notes to be added to the bibliography.  These
should be given in place in the document \bibnote{This is a note.
The text will be moved the the references section.  The title of the
section will change to ``Notes and References''.}.  As with
citations, the text should be placed before punctuation.  A note is
also generated if a citation has an optional note.  This assumes that
the whole work has already been cited: odd numbering will result if
this is not the case \cite[p.~1]{Cotton1999}.

\subsection{Floats}

New float types are automatically set up by the class file.  The
means graphics are included as follows (Scheme~\ref{sch:example}).  As
illustrated, the float is ``here'' if possible.
\begin{scheme}
  Your scheme graphic would go here: \texttt{.eps} format\\
  for \LaTeX\, or \texttt{.pdf} (or \texttt{.png}) for pdf\LaTeX\\
  \textsc{ChemDraw} files are best saved as \texttt{.eps} files:\\
  these can be scaled without loss of quality, and can be\\
  converted to \texttt{.pdf} files easily using \texttt{eps2pdf}.\\
  %\includegraphics{graphic}
  \caption{An example scheme}
  \label{sch:example}
\end{scheme}

\begin{figure}
  As well as the standard float types \texttt{table}\\
  and \texttt{figure}, the class also recognises\\
  \texttt{scheme}, \texttt{chart} and \texttt{graph}.
  \caption{An example figure}
  \label{fgr:example}
\end{figure}

Charts, figures and schemes do not necessarily have to be labelled or
captioned.  However, tables should always have a title. It is
possible to include a number and label for a graphic without any
title, using an empty argument to the \texttt{\textbackslash caption}
macro.

The use of the different floating environments is not required, but
it is intended to make document preparation easier for authors. In
general, you should place your graphics where they make logical
sense; the production process will move them if needed.

\subsection{Math(s)}

The \textsf{achemso} class does not load any particular additional
support for mathematics.  If packages such as \textsf{amsmath} are
required, they should be loaded in the preamble.  However,
the basic \LaTeX\ math(s) input should work correctly without
this.  Some inline material \( y = mx + c \) or $ 1 + 1 = 2 $
followed by some display. \[ A = \pi r^2 \]

It is possible to label equations in the usual way (Eq.~\ref{eqn:example}).
\begin{equation}
  \frac{\mathrm{d}}{\mathrm{d}x} \, r^2 = 2r \label{eqn:example}
\end{equation}
This can also be used to have equations containing graphical
content. To align the equation number with the middle of the graphic,
rather than the bottom, a minipage may be used.
\begin{equation}
  \begin{minipage}[c]{0.80\linewidth}
    \centering
    As illustrated here, the width of \\
    the minipage needs to allow some  \\
    space for the number to fit in to.
    %\includegraphics{graphic}
  \end{minipage}
  \label{eqn:graphic}
\end{equation}

\section{Experimental}

The usual experimental details should appear here.  This could
include a table, which can be referenced as Table~\ref{tbl:example}.
Notice that the caption is positioned at the top of the table.
\begin{table}
  \caption{An example table}
  \label{tbl:example}
  \begin{tabular}{ll}
    \hline
    Header one  & Header two  \\
    \hline
    Entry one   & Entry two   \\
    Entry three & Entry four  \\
    Entry five  & Entry five  \\
    Entry seven & Entry eight \\
    \hline
  \end{tabular}
\end{table}

Adding notes to tables can be complicated.  Perhaps the easiest
method is to generate these using the basic
\texttt{\textbackslash textsuperscript} and
\texttt{\textbackslash emph} macros, as illustrated (Table~\ref{tbl:notes}).
\begin{table}
  \caption{A table with notes}
  \label{tbl:notes}
  \begin{tabular}{ll}
    \hline
    Header one                            & Header two \\
    \hline
    Entry one\textsuperscript{\emph{a}}   & Entry two  \\
    Entry three\textsuperscript{\emph{b}} & Entry four \\
    \hline
  \end{tabular}

  \textsuperscript{\emph{a}} Some text;
  \textsuperscript{\emph{b}} Some more text.
\end{table}

The example file also loads the optional \textsf{mhchem} package, so
that formulas are easy to input: \texttt{\textbackslash ce\{H2SO4\}}
gives \ce{H2SO4}.  See the use in the bibliography file (when using
titles in the references section).

The use of new commands should be limited to simple things which will
not interfere with the production process.  For example,
\texttt{\textbackslash mycommand} has been defined in this example,
to give italic, mono-spaced text: \mycommand{some text}.

\section{Extra information when writing JACS Communications}

When producing communications for \emph{J.~Am.\ Chem.\ Soc.}, the
class will automatically lay the text out in the style of the
journal. This gives a guide to the length of text that can be
accommodated in such a publication. There are some points to bear in
mind when preparing a JACS Communication in this way.  The layout
produced here is a \emph{model} for the published result, and the
outcome should be taken as a \emph{guide} to the final length. The
spacing and sizing of graphical content is an area where there is
some flexibility in the process.  You should not worry about the
space before and after graphics, which is set to give a guide to the
published size. This is very dependant on the final published layout.

You should be able to use the same source to produce a JACS
Communication and a normal article.  For example, this demonstration
file will work with both \texttt{type=article} and
\texttt{type=communication}. Sections and any abstract are
automatically ignored, although you will get warnings to this effect.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "Acknowledgement" section can be given in all manuscript
%% classes.  This should be given within the "acknowledgement"
%% environment, which will make the correct section or running title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}

Please use ``The authors thank \ldots'' rather than ``The
authors would like to thank \ldots''.

The author thanks Mats Dahlgren for version one of \textsf{achemso},
and Donald Arseneau for the code taken from \textsf{cite} to move
citations after punctuation. Many users have provided feedback on the
class, which is reflected in all of the different demonstrations
shown in this document.

\end{acknowledgement}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The same is true for Supporting Information, which should use the
%% suppinfo environment.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{suppinfo}

This will usually read something like: ``Experimental procedures and
characterization data for all new compounds. The class will
automatically add a sentence pointing to the information on-line:

\end{suppinfo}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The appropriate \bibliography command should be placed here.
%% Notice that the class file automatically sets \bibliographystyle
%% and also names the section correctly.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{tocentry}

Some journals require a graphical entry for the Table of Contents.
This should be laid out ``print ready'' so that the sizing of the
text is correct.

Inside the \texttt{tocentry} environment, the font used is Helvetica
8\,pt, as required by \emph{Journal of the American Chemical
Society}.

The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
permitted for  \emph{Journal of the American Chemical Society}
graphical table of content entries. The box will not resize if the
content is too big: instead it will overflow the edge of the box.

This box and the associated title will always be printed on a
separate page at the end of the document.

\end{tocentry}

\end{document}

results section mantra:
1 - Explain what you did and then describe what you expected as an outcome
2 - Describe the data and show the dta
3 - Interpret the data

### 2018-02-02 bb494/manuscript.tex
all

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author{Andrew N. Other}
-\altaffiliation{A shared footnote}
-\author{Fred T. Secondauthor}
-\altaffiliation{Current address: Some other place, Othert\"own,
-Germany}
-\author{I. Ken Groupleader}
-\altaffiliation{A shared footnote}
-\email{i.k.groupleader@unknown.uu}
-\phone{+123 (0)123 4445556}
-\fax{+123 (0)123 4445557}
-\affiliation[Unknown University]
-{Department of Chemistry, Unknown University, Unknown Town}
-\alsoaffiliation[Second University]
-{Department of Chemistry, Second University, Nearby Town}
-\author{Susanne K. Laborator}
-\email{s.k.laborator@bigpharma.co}
-\affiliation[BigPharma]
-{Lead Discovery, BigPharma, Big Town, USA}
-\author{Kay T. Finally}
-\affiliation[Unknown University]
-{Department of Chemistry, Unknown University, Unknown Town}
-\alsoaffiliation[Second University]
-{Department of Chemistry, Second University, Nearby Town}
+\author{Elliott Capek}
+\altaffiliation{Department of Biochemistry, Oregon State University}
+%% \author{}
+%% \altaffiliation{}
+%% \author{}
+%% \altaffiliation{}
+%% \email{}
+%% \phone{}
+%% \fax{}
+%% \affiliation[]{}
+%% \alsoaffiliation[]{}
+%% \author{}
+%% \email{}
+%% \affiliation[]{}
+\author{Aidan Estelle}
+\affiliation[]{}
+\alsoaffiliation[]{}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title[An \textsf{achemso} demo]
-  {A demonstration of the \textsf{achemso} \LaTeX\
-   class\footnote{A footnote for the title}}
+\title[]{Click chemistry reveals an alternate mechanism for HPII catalase}
 
@@
 \begin{abstract}
-  This is an example document for the \textsf{achemso} document
-  class, intended for submissions to the American Chemical Society
-  for publication. The class is based on the standard \LaTeXe\
-  \textsf{report} file, and does not seek to reproduce the appearance
-  of a published paper.
+  %% This is an example document for the \textsf{achemso} document
+  %% class, intended for submissions to the American Chemical Society
+  %% for publication. The class is based on the standard \LaTeXe\
+  %% \textsf{report} file, and does not seek to reproduce the appearance
+  %% of a published paper.
 
-  This is an abstract for the \textsf{achemso} document class
-  demonstration document.  An abstract is only allowed for certain
-  manuscript types.  The selection of \texttt{journal} and
-  \texttt{manuscript} will determine if an abstract is valid.  If
-  not, the class will issue an appropriate error.
+  %% This is an abstract for the \textsf{achemso} document class
+  %% demonstration document.  An abstract is only allowed for certain
+  %% manuscript types.  The selection of \texttt{journal} and
+  %% \texttt{manuscript} will determine if an abstract is valid.  If
+  %% not, the class will issue an appropriate error.
 \end{abstract}
@@
 \section{Introduction}
-This is a paragraph of text to fill the introduction of the
-demonstration file.  The demonstration file attempts to show the
-modifications of the standard \LaTeX\ macros that are implemented by
-the \textsf{achemso} class.  These are mainly concerned with content,
-as opposed to appearance.
+
+\section{Methods}
 

### 2018-02-04 bb494/manuscript.tex
add bb494 stuff

@@
 \section{Methods}
+\textbf{Protein expression}\\
+Wildtype protein was expressed by transforming the pBAD-HPII-WT plasmid in Figure \ref{fig:pbad-hpii-wt} into competent \textit{E.
+coli} DH10B cells via heat-shock.
+Plasmid includes ampicillin resistance and a 6xHis tag on the N-term of the HPII gene.
+Cells were then grown in autoinduction media with an arabinose inducer (0.1\% aspartate pH 7.5, 0.2\% glycerol, mineral salts, 0.05\% glucose, 20mM MgSO4, 0.05\% arabinose, trace metals and AA mix) for 48 hours, shaking, at 37C.
+Cells were then pelletized and stored at -80C.\\
+
+\textbf{Protein purification}\\
+Pelletized cells were reconstituted in TALON equilibration buffer (50mM sodium phosphate, 300mM NaCl) and microfluidized, also in equilibration buffer.
+Lysed cells were spun down at 15K RPM for 20 minutes at 4C.
+Supernatant was collected and combined with 300$\mu L$ bed volume TALON beads per 100mL cultured cells.
+TALON-supernatant mix was nutated for 30 minutes at 4C.
+Bound beads were then spun down at 500g for 5 minutes and the supernatant discarded.
+The beads were applied to a pre-washed gravity-flow column at RT and washed twice with 10mL equilibration buffer.
+Elution was done using 2mL elution buffer (50mM sodium phosphate, 300mM NaCl, 250mM imidazole) into four 500$\mu L$ aliquots.
+Protein was then stored at 4C.\\
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{figures/crude-wt-gel}
+\caption{Electrophoresis gel of wildtype expression.
+Serial dillution of WT protein purification shown in two lanes.
+Major band in purification lanes correspond to protein of approximately 75kD, as expected for catalse single subunit.}
+\label{fig:duration-length}
+\end{figure}
+
 \subsection{Outline}
-
-The document layout should follow the style of the journal concerned.
-Where appropriate, sections and subsections should be added in the
-normal way. If the class options are set correctly, warnings will be
-given if these should not be present.
 
@@
 
-The class makes various changes to the way that references are
-handled.  The class loads \textsf{natbib}, and also the
-appropriate bibliography style.  References can be made using
-the normal method; the citation should be placed before any
-punctuation, as the class will move it if using a superscript
-citation style
 \cite{Mena2000,Abernethy2003,Friedman-Hill2003,EuropeanCommission2008}.
-The use of \textsf{natbib} allows the use of the various citation
-commands of that package: \citeauthor{Abernethy2003} have shown
-something, in \citeyear{Cotton1999}, or as given by
-Ref.~\citenum{Mena2000}.  Long lists of authors will be
-automatically truncated in most article formats, but not in
-supplementary information or reviews \cite{Pople2003}. If you
-encounter problems with the citation macros, please check that
-your copy of \textsf{natbib} is up to date. The demonstration
-database file \texttt{achemso-demo.bib} shows how to complete
-entries correctly. Notice that ``\latin{et al.}'' is auto-formatted
-using the \texttt{\textbackslash latin} command.
 
-Multiple citations to be combined into a list can be given as
-a single citation.  This uses the \textsf{mciteplus} package
-\cite{Johnson1972,*Arduengo1992,*Eisenstein2005,*Arduengo1994}.
-Citations other than the first of the list should be indicated
-with a star. If the \textsf{mciteplus} package is not installed,
-the standard bibliography tools will still work but starred
-references will be ignored. Individual references can be referred
-to using \texttt{\textbackslash mciteSubRef}:
-``ref.~\mciteSubRef{Eisenstein2005}''.
+\bibnote{This is a note. The text will be moved the the references section.  The title of the section will change to ``Notes and References''.}
 
-The class also handles notes to be added to the bibliography.  These
-should be given in place in the document \bibnote{This is a note.
-The text will be moved the the references section.  The title of the
-section will change to ``Notes and References''.}.  As with
-citations, the text should be placed before punctuation.  A note is
-also generated if a citation has an optional note.  This assumes that
-the whole work has already been cited: odd numbering will result if
-this is not the case \cite[p.~1]{Cotton1999}.
+\textsf{natbib}
 
-\subsection{Floats}
-
-New float types are automatically set up by the class file.  The
-means graphics are included as follows (Scheme~\ref{sch:example}).  As
-illustrated, the float is ``here'' if possible.
-\begin{scheme}
-  Your scheme graphic would go here: \texttt{.eps} format\\
-  for \LaTeX\, or \texttt{.pdf} (or \texttt{.png}) for pdf\LaTeX\\
-  \textsc{ChemDraw} files are best saved as \texttt{.eps} files:\\
-  these can be scaled without loss of quality, and can be\\
-  converted to \texttt{.pdf} files easily using \texttt{eps2pdf}.\\
-  %\includegraphics{graphic}
-  \caption{An example scheme}
-  \label{sch:example}
-\end{scheme}
-
-\begin{figure}
-  As well as the standard float types \texttt{table}\\
-  and \texttt{figure}, the class also recognises\\
-  \texttt{scheme}, \texttt{chart} and \texttt{graph}.
-  \caption{An example figure}
-  \label{fgr:example}
-\end{figure}
-
-Charts, figures and schemes do not necessarily have to be labelled or
-captioned.  However, tables should always have a title. It is
-possible to include a number and label for a graphic without any
-title, using an empty argument to the \texttt{\textbackslash caption}
-macro.
-
-The use of the different floating environments is not required, but
-it is intended to make document preparation easier for authors. In
-general, you should place your graphics where they make logical
-sense; the production process will move them if needed.
-
-\subsection{Math(s)}
-
-The \textsf{achemso} class does not load any particular additional
-support for mathematics.  If packages such as \textsf{amsmath} are
-required, they should be loaded in the preamble.  However,
-the basic \LaTeX\ math(s) input should work correctly without
-this.  Some inline material \( y = mx + c \) or $ 1 + 1 = 2 $
-followed by some display. \[ A = \pi r^2 \]
-
-It is possible to label equations in the usual way (Eq.~\ref{eqn:example}).
-\begin{equation}
-  \frac{\mathrm{d}}{\mathrm{d}x} \, r^2 = 2r \label{eqn:example}
-\end{equation}
-This can also be used to have equations containing graphical
-content. To align the equation number with the middle of the graphic,
-rather than the bottom, a minipage may be used.
-\begin{equation}
-  \begin{minipage}[c]{0.80\linewidth}
-    \centering
-    As illustrated here, the width of \\
-    the minipage needs to allow some  \\
-    space for the number to fit in to.
-    %\includegraphics{graphic}
-  \end{minipage}
-  \label{eqn:graphic}
-\end{equation}
-
-\section{Experimental}
-
-The usual experimental details should appear here.  This could
-include a table, which can be referenced as Table~\ref{tbl:example}.
-Notice that the caption is positioned at the top of the table.
-\begin{table}
-  \caption{An example table}
-  \label{tbl:example}
-  \begin{tabular}{ll}
-    \hline
-    Header one  & Header two  \\
-    \hline
-    Entry one   & Entry two   \\
-    Entry three & Entry four  \\
-    Entry five  & Entry five  \\
-    Entry seven & Entry eight \\
-    \hline
-  \end{tabular}
-\end{table}
-
-Adding notes to tables can be complicated.  Perhaps the easiest
-method is to generate these using the basic
-\texttt{\textbackslash textsuperscript} and
-\texttt{\textbackslash emph} macros, as illustrated (Table~\ref{tbl:notes}).
-\begin{table}
-  \caption{A table with notes}
-  \label{tbl:notes}
-  \begin{tabular}{ll}
-    \hline
-    Header one                            & Header two \\
-    \hline
-    Entry one\textsuperscript{\emph{a}}   & Entry two  \\
-    Entry three\textsuperscript{\emph{b}} & Entry four \\
-    \hline
-  \end{tabular}
-
-  \textsuperscript{\emph{a}} Some text;
-  \textsuperscript{\emph{b}} Some more text.
-\end{table}
-
-The example file also loads the optional \textsf{mhchem} package, so
-that formulas are easy to input: \texttt{\textbackslash ce\{H2SO4\}}
-gives \ce{H2SO4}.  See the use in the bibliography file (when using
-titles in the references section).
-
-The use of new commands should be limited to simple things which will
-not interfere with the production process.  For example,
-\texttt{\textbackslash mycommand} has been defined in this example,
-to give italic, mono-spaced text: \mycommand{some text}.
-
-\section{Extra information when writing JACS Communications}
-
-When producing communications for \emph{J.~Am.\ Chem.\ Soc.}, the
-class will automatically lay the text out in the style of the
-journal. This gives a guide to the length of text that can be
-accommodated in such a publication. There are some points to bear in
-mind when preparing a JACS Communication in this way.  The layout
-produced here is a \emph{model} for the published result, and the
-outcome should be taken as a \emph{guide} to the final length. The
-spacing and sizing of graphical content is an area where there is
-some flexibility in the process.  You should not worry about the
-space before and after graphics, which is set to give a guide to the
-published size. This is very dependant on the final published layout.
-
-You should be able to use the same source to produce a JACS
-Communication and a normal article.  For example, this demonstration
-file will work with both \texttt{type=article} and
-\texttt{type=communication}. Sections and any abstract are
-automatically ignored, although you will get warnings to this effect.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The "Acknowledgement" section can be given in all manuscript
-%% classes.  This should be given within the "acknowledgement"
-%% environment, which will make the correct section or running title.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{acknowledgement}
-
-Please use ``The authors thank \ldots'' rather than ``The
-authors would like to thank \ldots''.
-
-The author thanks Mats Dahlgren for version one of \textsf{achemso},
-and Donald Arseneau for the code taken from \textsf{cite} to move
-citations after punctuation. Many users have provided feedback on the
-class, which is reflected in all of the different demonstrations
-shown in this document.
-
-\end{acknowledgement}
+\citeauthor{Abernethy2003}
+\textsf{natbib}
+\citeyear{Cotton1999}
+~\citenum{Mena2000}
+\cite{Pople2003}
+\textsf{natbib}
+\texttt{achemso-demo.bib}
+\latin{et al.}
+\texttt{\textbackslash latin}
+\textsf{mciteplus}
+\cite{Johnson1972,*Arduengo1992,*Eisenstein2005,*Arduengo1994}
+\texttt{\textbackslash mciteSubRef}
+\mciteSubRef{Eisenstein2005}
+\cite[p.~1]{Cotton1999}.
 
@@
 \begin{suppinfo}
-
-This will usually read something like: ``Experimental procedures and
-characterization data for all new compounds. The class will
-automatically add a sentence pointing to the information on-line:
 
@@
 
-\begin{tocentry}
+%% \begin{tocentry}
 
-Some journals require a graphical entry for the Table of Contents.
-This should be laid out ``print ready'' so that the sizing of the
-text is correct.
-
-Inside the \texttt{tocentry} environment, the font used is Helvetica
-8\,pt, as required by \emph{Journal of the American Chemical
-Society}.
-
-The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
-permitted for  \emph{Journal of the American Chemical Society}
-graphical table of content entries. The box will not resize if the
-content is too big: instead it will overflow the edge of the box.
-
-This box and the associated title will always be printed on a
-separate page at the end of the document.
-
-\end{tocentry}
+%% \end{tocentry}
 

### 2018-02-05 bb494/manuscript.tex
all

@@
 \usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
+\usepackage{amsmath}
 
@@
 \author{Elliott Capek}
-\altaffiliation{Department of Biochemistry, Oregon State University}
-%% \author{}
-%% \altaffiliation{}
-%% \author{}
-%% \altaffiliation{}
-%% \email{}
-%% \phone{}
-%% \fax{}
-%% \affiliation[]{}
+\affiliation{Department of Biochemistry, Oregon State University}
+\email{capeke@oregonstate.edu}
+\phone{(971) 207 8193}
 %% \alsoaffiliation[]{}
-%% \author{}
-%% \email{}
-%% \affiliation[]{}
 \author{Aidan Estelle}
-\affiliation[]{}
-\alsoaffiliation[]{}
+%% \affiliation{Department of Biochemistry, Oregon State University}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\abbreviations{IR,NMR,UV}
-\keywords{American Chemical Society, \LaTeX}
+\keywords{Catalase, noncanonical amino acid, active site, channel flow}
 
@@
   %% not, the class will issue an appropriate error.
+  Abstract abstract abstracts abstract.
+Abstract abstract abstracts abstract.
+Abstract abstract abstracts abstract.
+Abstract abstract abstracts abstract.
+Abstract abstract abstracts abstract.
+Abstract abstract abstracts abstract.
+Abstract abstract abstracts abstract.
+Abstract abstract abstracts abstract.\\
 \end{abstract}
@@
 \section{Introduction}
+Catalase is an enzyme which catalyzes the reaction $2H_2O_2 \rightarrow O_2 + 2H_2O$.
+This is an important detoxification reaction, since hydrogen peroxide is a reactive toxin generated by several biological processes.
+
+The molecular mechaninism of catalase proceeds in two steps.
+First, the heme reduces a hydrogen peroxide, producing water and the so-called Cpd I.
+Then Cpd I oxidizes a second hydrogen peroxide, producing water and oxygen gas.
+This second step is mediated by a histidine near the active site \cite{alfonso-prieto}.
+
+This bimolecular mechanism relies on the efficient transport of hydrogen peroxide into the deeply buried active site.
+This is believed to be aided by an electrical gradient over 12 \AA between an active site aspartic acid and the positive heme iron \cite{chelikani}.
+Mutagenisis of this Asp to positive residues decreases activity up to 90\% \cite{chelikani}.
+However, this study was accomplished using canonical amino acids, and provides relatively weak evidence for the electric field hypothesis.\\
+
+In this study we seek to use the technique of bioorthogonal ligation to study the electric field hypothesis of catalase substrate transport.
+We will incorporate the noncanonical amino acid 3-nitrotyrosine into the active site, then use click chemistry to insert a gradient of charged particles into the active site.
+By using a large number of particles with charges varying from positive to neutral to negative, we can average out the effects of steric bulk and isolate the electrical effects of charged amino acids in the catalase active site.
+The results of such a study would conclusively determine the importance of an electric field in the catalase active site.\\
 
@@
 \textbf{Protein expression}\\
-Wildtype protein was expressed by transforming the pBAD-HPII-WT plasmid in Figure \ref{fig:pbad-hpii-wt} into competent \textit{E.
+Wildtype protein was expressed by transforming the plasmid pBad-HPII (Addgene ID 105839), containing WT HPII gene, into competent \textit{E.
 coli} DH10B cells via heat-shock.
-Plasmid includes ampicillin resistance and a 6xHis tag on the N-term of the HPII gene.
-Cells were then grown in autoinduction media with an arabinose inducer (0.1\% aspartate pH 7.5, 0.2\% glycerol, mineral salts, 0.05\% glucose, 20mM MgSO4, 0.05\% arabinose, trace metals and AA mix) for 48 hours, shaking, at 37C.
+pBad-HPII plasmid includes ampicillin resistance, an imidazole promoter, and a 6xHis tag on the N-term of the HPII gene.
+Cells were then grown in autoinduction media with an arabinose inducer (0.1\% aspartate pH 7.5, 0.2\% glycerol, 100$\mu g / mL$ ampicillin, mineral salts, 0.05\% glucose, 20mM MgSO4, 0.05\% arabinose, trace metals and AA mix) for 48 hours, shaking, at 37C.
 Cells were then pelletized and stored at -80C.\\
+
+The same protocol was followed for the ncAA mutant catalases F206NY and F206pAzF, except protein plasmid pBad-HPII TAG206 (Addgene ID 105846) was used as a base.
+ncAA translation machinery plasmids pDule-3-nitrotyrosine (Addgene ID 85498) and pDule-pCNF (85494), which contain a tetracycline resistance gene, a TAG tRNA, and a compatible tRNA synthase, were co-transformed into DH10B along with pBAD-HPII TAG206 for respective F206NY and F206pAzF cell lines.
+ncAA autoinduction media was supplemented with $25\mu g / mL$ tetracycline to select for both plasmids.\\
+
+OD600 measurements of cultures at 48h growth time were 0.65, 0.58 and 0.45 arb. units for WT, F206NY and F206pAzF cells.\\
 
@@
 
-\section{Results and discussion}
+\textbf{Bradford assay}\\
+A Bio-Rad Bradford assay kit was used using the standard protocol.
+Bovine serum albumin was used as the reference protein to construct the standard curve.
+Concentrations from 125 $\mu g / mL$ to 2 mg/mL were explored.
+Samples were incubated with dye in cuvettes for 5 minutes at RT.
+Samples were then analyzed via spectrophotometer at 595 \textit{nm} and used to create the standard curve shown in Figure \ref{fig:bradford-assay}.
+Protein samples were then compared to this standard curve to estimate their concentration.\\
+
+\textbf{Qualitative activity assay}\\
+10 $\mu g$ purified WT catalase was added to a 5\% solution of hydrogen peroxide in water. The mixture was then examined for bubble formation.\\
+
+\section{Results}
+DH10B cell lines with appropriate plasmids were grown for 48h and pelletized.
+Wildtype HPII-containing cells were purified using TALON resin.
+Purified wildtype protein and ncAA cell lysates are shown in Figure \ref{fig:crude-wt-gel}.
+The expected band corresponding to an 84-kD protein is seen across all lanes, verifying the presence of HPII wildtype in all cultures.\\
+
+A Bradford assay of wildtype purified protein was then conducted. The results, shown in Figure \ref{fig:bradford-assay}, further verify the presence of protein in the purified sample.\\
+
+A qualitative, rudimentary assay testing the activity of wildtype purified HPII catalase was conducted. This assay verified HPII-WT was active by producing bubbles.\\
 
@@
 \caption{Electrophoresis gel of wildtype expression.
-Serial dillution of WT protein purification shown in two lanes.
-Major band in purification lanes correspond to protein of approximately 75kD, as expected for catalse single subunit.}
-\label{fig:duration-length}
+Serial dillution of WT protein purification shown in lanes three and four.
+Major band in purification lanes correspond to protein of approximately 80kD, as expected for a catalase single subunit.
+}
+\label{fig:crude-wt-gel}
 \end{figure}
 
-\subsection{Outline}
+\begin{figure}
+  \centering
+  \begin{minipage}{.4\textwidth}
+    \begin{tabular}{lll}
+      \hline
+      Sample  & OD590 & mg/mL  \\
+      \hline
+      WT        & 2.18 & 2.74\\
+      %% F206NY    & 0.71 & 0.83\\
+      %% F206pAzF  & 1.51 & 1.88\\
+      \hline
+    \end{tabular}
+  \end{minipage}
+  \begin{minipage}{.59\textwidth}
+    \includegraphics[width=\linewidth]{figures/bradford-standard-curve}
+  \end{minipage}
+  \caption{Bradford assay results. \textbf{Left:} Sample absorbance at 590 nm and extrapolated concentration. \textbf{Right:} Bradford standard curve.}
+  \label{fig:bradford-assay}
+\end{figure}
 
 \subsection{References}
-
-\cite{Mena2000,Abernethy2003,Friedman-Hill2003,EuropeanCommission2008}.
-
-\bibnote{This is a note. The text will be moved the the references section.  The title of the section will change to ``Notes and References''.}
-
-\textsf{natbib}
-
-\citeauthor{Abernethy2003}
-\textsf{natbib}
-\citeyear{Cotton1999}
-~\citenum{Mena2000}
-\cite{Pople2003}
-\textsf{natbib}
-\texttt{achemso-demo.bib}
-\latin{et al.}
-\texttt{\textbackslash latin}
-\textsf{mciteplus}
-\cite{Johnson1972,*Arduengo1992,*Eisenstein2005,*Arduengo1994}
-\texttt{\textbackslash mciteSubRef}
-\mciteSubRef{Eisenstein2005}
-\cite[p.~1]{Cotton1999}.
 

### 2018-02-05 bb494/manuscript.tex
more

@@
 \phone{(971) 207 8193}
+\author{Hassan Alnatah}
+\author{Laikana Ly}
+\author{Joey Orton}
 %% \alsoaffiliation[]{}
@@
 This is an important detoxification reaction, since hydrogen peroxide is a reactive toxin generated by several biological processes.
+Catalase is a tetrameric protein which uses a heme group in its active site to facilitate the breakdown of hydrogen peroxide.
+It is a unique protein in that its reaction evolves gas, and thus its activity can be seen with the bare eye.
+The protein has been studied as a target for tumor suppression drugs \cite{bauer}.
+The enzyme is also useful industrially, in particular in the bleaching of textiles \cite{amorim}.
 
@@
 
-This bimolecular mechanism relies on the efficient transport of hydrogen peroxide into the deeply buried active site.
+This bimolecular substrate mechanism relies on the efficient transport of hydrogen peroxide into the deeply buried active site.
 This is believed to be aided by an electrical gradient over 12 \AA between an active site aspartic acid and the positive heme iron \cite{chelikani}.
 Mutagenisis of this Asp to positive residues decreases activity up to 90\% \cite{chelikani}.
-However, this study was accomplished using canonical amino acids, and provides relatively weak evidence for the electric field hypothesis.\\
+However, this study was accomplished using only a few canonical amino acid substitutions, and provides relatively weak evidence for the electric field hypothesis.
 
@@
 We will incorporate the noncanonical amino acid 3-nitrotyrosine into the active site, then use click chemistry to insert a gradient of charged particles into the active site.
-By using a large number of particles with charges varying from positive to neutral to negative, we can average out the effects of steric bulk and isolate the electrical effects of charged amino acids in the catalase active site.
+By using a large number of particles with charges varying from positive to negative, we can average out the effects of steric bulk and isolate the effect of field strength.
 The results of such a study would conclusively determine the importance of an electric field in the catalase active site.\\
@@
 Purified wildtype protein and ncAA cell lysates are shown in Figure \ref{fig:crude-wt-gel}.
-The expected band corresponding to an 84-kD protein is seen across all lanes, verifying the presence of HPII wildtype in all cultures.\\
+The expected band corresponding to an 84-kD protein is seen across all lanes, verifying the presence of HPII wildtype in all cultures.
+Contaminant bands in the gel are very faint, indicating pure protein.\\
 

### 2018-02-18 bb494/manuscript.tex
commit

@@
 Wildtype HPII-containing cells were purified using TALON resin.
-Purified wildtype protein and ncAA cell lysates are shown in Figure \ref{fig:crude-wt-gel}.
-The expected band corresponding to an 84-kD protein is seen across all lanes, verifying the presence of HPII wildtype in all cultures.
+Purified wildtype protein and ncAA cell lysates are shown in Figure \ref{fig:pure-gel}.
+The expected band corresponding to an 84-kD protein is seen across all lanes, indicating the presence of a catalase single subunit in each cell line.
 Contaminant bands in the gel are very faint, indicating pure protein.\\
@@
 \centering
-\includegraphics[width=\linewidth]{figures/crude-wt-gel}
-\caption{Electrophoresis gel of wildtype expression.
-Serial dillution of WT protein purification shown in lanes three and four.
+\includegraphics[width=0.7\linewidth]{figures/pure-gel}
+\caption{Electrophoresis gel of crude lysate and purified protein expression.
 Major band in purification lanes correspond to protein of approximately 80kD, as expected for a catalase single subunit.
-}
-\label{fig:crude-wt-gel}
+Crude negative control lanes lack 80kD band, consistent with noncanonical amino acid being necessary for noncanonical protein expression.}
+\label{fig:pure-gel}
 \end{figure}

### 2018-02-19 bb494/manuscript.tex
sd

@@
 3 - Interpret the data
+
+Substrate Flow, Loewen
+There are H2O2 binding sites in the heme distal pocket and main channel
+There are several input channels; what are their purposes?
+There is a ``molecular ruler'' in the main channel which selects for H2O2 against water
+Active site mutations which allow more water in slow down the rxn rate
+This paper discusses unidirectional flow of H2O2 through the main channel
+
+Active and inhibited catalase structures, tainer
+This is the paper about the hydrophobic ``molecular ruler''
+A hydrophobic region distal to the heme cleft contains phenylalanines, tryptophan and two valines
+Authors propose that two waters (one h-bonded to human Asn148 and His75; other h-bonded to gln168 and asp128) form a ``ruler''
+This ``ruler'' is far enough apart that only H2O2 can stably occupy the distance betweeen, thus allowing only H2O2 to enter, not H2O
+This would explain why opening the perpendicular channel decreases the rate
+
+Role of the lateral channel in catalase, loewen
+This paper discusses the lateral channel as an exhaust channel
+Proposes that perpendicular channel is the entrance, lateral channel is exit

### 2018-02-19 bb494/manuscript.tex
sfd

--- 2018-02-05 bb494/manuscript.tex
@@
 \usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
-\usepackage{amsmath}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author{Elliott Capek}
-\affiliation{Department of Biochemistry, Oregon State University}
-\email{capeke@oregonstate.edu}
-\phone{(971) 207 8193}
-\author{Hassan Alnatah}
-\author{Laikana Ly}
-\author{Joey Orton}
-%% \alsoaffiliation[]{}
-\author{Aidan Estelle}
-%% \affiliation{Department of Biochemistry, Oregon State University}
+\author{Andrew N. Other}
+\altaffiliation{A shared footnote}
+\author{Fred T. Secondauthor}
+\altaffiliation{Current address: Some other place, Othert\"own,
+Germany}
+\author{I. Ken Groupleader}
+\altaffiliation{A shared footnote}
+\email{i.k.groupleader@unknown.uu}
+\phone{+123 (0)123 4445556}
+\fax{+123 (0)123 4445557}
+\affiliation[Unknown University]
+{Department of Chemistry, Unknown University, Unknown Town}
+\alsoaffiliation[Second University]
+{Department of Chemistry, Second University, Nearby Town}
+\author{Susanne K. Laborator}
+\email{s.k.laborator@bigpharma.co}
+\affiliation[BigPharma]
+{Lead Discovery, BigPharma, Big Town, USA}
+\author{Kay T. Finally}
+\affiliation[Unknown University]
+{Department of Chemistry, Unknown University, Unknown Town}
+\alsoaffiliation[Second University]
+{Department of Chemistry, Second University, Nearby Town}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title[]{Click chemistry reveals an alternate mechanism for HPII catalase}
+\title[An \textsf{achemso} demo]
+  {A demonstration of the \textsf{achemso} \LaTeX\
+   class\footnote{A footnote for the title}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\keywords{Catalase, noncanonical amino acid, active site, channel flow}
+\abbreviations{IR,NMR,UV}
+\keywords{American Chemical Society, \LaTeX}
 
@@
 \begin{abstract}
-  %% This is an example document for the \textsf{achemso} document
-  %% class, intended for submissions to the American Chemical Society
-  %% for publication. The class is based on the standard \LaTeXe\
-  %% \textsf{report} file, and does not seek to reproduce the appearance
-  %% of a published paper.
+  This is an example document for the \textsf{achemso} document
+  class, intended for submissions to the American Chemical Society
+  for publication. The class is based on the standard \LaTeXe\
+  \textsf{report} file, and does not seek to reproduce the appearance
+  of a published paper.
 
-  %% This is an abstract for the \textsf{achemso} document class
-  %% demonstration document.  An abstract is only allowed for certain
-  %% manuscript types.  The selection of \texttt{journal} and
-  %% \texttt{manuscript} will determine if an abstract is valid.  If
-  %% not, the class will issue an appropriate error.
-  Abstract abstract abstracts abstract.
-Abstract abstract abstracts abstract.
-Abstract abstract abstracts abstract.
-Abstract abstract abstracts abstract.
-Abstract abstract abstracts abstract.
-Abstract abstract abstracts abstract.
-Abstract abstract abstracts abstract.
-Abstract abstract abstracts abstract.\\
+  This is an abstract for the \textsf{achemso} document class
+  demonstration document.  An abstract is only allowed for certain
+  manuscript types.  The selection of \texttt{journal} and
+  \texttt{manuscript} will determine if an abstract is valid.  If
+  not, the class will issue an appropriate error.
 \end{abstract}
@@
 \section{Introduction}
-Catalase is an enzyme which catalyzes the reaction $2H_2O_2 \rightarrow O_2 + 2H_2O$.
-This is an important detoxification reaction, since hydrogen peroxide is a reactive toxin generated by several biological processes.
-Catalase is a tetrameric protein which uses a heme group in its active site to facilitate the breakdown of hydrogen peroxide.
-It is a unique protein in that its reaction evolves gas, and thus its activity can be seen with the bare eye.
-The protein has been studied as a target for tumor suppression drugs \cite{bauer}.
-The enzyme is also useful industrially, in particular in the bleaching of textiles \cite{amorim}.
+This is a paragraph of text to fill the introduction of the
+demonstration file.  The demonstration file attempts to show the
+modifications of the standard \LaTeX\ macros that are implemented by
+the \textsf{achemso} class.  These are mainly concerned with content,
+as opposed to appearance.
 
-The molecular mechaninism of catalase proceeds in two steps.
-First, the heme reduces a hydrogen peroxide, producing water and the so-called Cpd I.
-Then Cpd I oxidizes a second hydrogen peroxide, producing water and oxygen gas.
-This second step is mediated by a histidine near the active site \cite{alfonso-prieto}.
+\section{Results and discussion}
 
-This bimolecular substrate mechanism relies on the efficient transport of hydrogen peroxide into the deeply buried active site.
-This is believed to be aided by an electrical gradient over 12 \AA between an active site aspartic acid and the positive heme iron \cite{chelikani}.
-Mutagenisis of this Asp to positive residues decreases activity up to 90\% \cite{chelikani}.
-However, this study was accomplished using only a few canonical amino acid substitutions, and provides relatively weak evidence for the electric field hypothesis.
+\subsection{Outline}
 
-In this study we seek to use the technique of bioorthogonal ligation to study the electric field hypothesis of catalase substrate transport.
-We will incorporate the noncanonical amino acid 3-nitrotyrosine into the active site, then use click chemistry to insert a gradient of charged particles into the active site.
-By using a large number of particles with charges varying from positive to negative, we can average out the effects of steric bulk and isolate the effect of field strength.
-The results of such a study would conclusively determine the importance of an electric field in the catalase active site.\\
+The document layout should follow the style of the journal concerned.
+Where appropriate, sections and subsections should be added in the
+normal way. If the class options are set correctly, warnings will be
+given if these should not be present.
 
-\section{Methods}
-\textbf{Protein expression}\\
-Wildtype protein was expressed by transforming the plasmid pBad-HPII (Addgene ID 105839), containing WT HPII gene, into competent \textit{E.
-coli} DH10B cells via heat-shock.
-pBad-HPII plasmid includes ampicillin resistance, an imidazole promoter, and a 6xHis tag on the N-term of the HPII gene.
-Cells were then grown in autoinduction media with an arabinose inducer (0.1\% aspartate pH 7.5, 0.2\% glycerol, 100$\mu g / mL$ ampicillin, mineral salts, 0.05\% glucose, 20mM MgSO4, 0.05\% arabinose, trace metals and AA mix) for 48 hours, shaking, at 37C.
-Cells were then pelletized and stored at -80C.\\
+\subsection{References}
 
-The same protocol was followed for the ncAA mutant catalases F206NY and F206pAzF, except protein plasmid pBad-HPII TAG206 (Addgene ID 105846) was used as a base.
-ncAA translation machinery plasmids pDule-3-nitrotyrosine (Addgene ID 85498) and pDule-pCNF (85494), which contain a tetracycline resistance gene, a TAG tRNA, and a compatible tRNA synthase, were co-transformed into DH10B along with pBAD-HPII TAG206 for respective F206NY and F206pAzF cell lines.
-ncAA autoinduction media was supplemented with $25\mu g / mL$ tetracycline to select for both plasmids.\\
+The class makes various changes to the way that references are
+handled.  The class loads \textsf{natbib}, and also the
+appropriate bibliography style.  References can be made using
+the normal method; the citation should be placed before any
+punctuation, as the class will move it if using a superscript
+citation style
+\cite{Mena2000,Abernethy2003,Friedman-Hill2003,EuropeanCommission2008}.
+The use of \textsf{natbib} allows the use of the various citation
+commands of that package: \citeauthor{Abernethy2003} have shown
+something, in \citeyear{Cotton1999}, or as given by
+Ref.~\citenum{Mena2000}.  Long lists of authors will be
+automatically truncated in most article formats, but not in
+supplementary information or reviews \cite{Pople2003}. If you
+encounter problems with the citation macros, please check that
+your copy of \textsf{natbib} is up to date. The demonstration
+database file \texttt{achemso-demo.bib} shows how to complete
+entries correctly. Notice that ``\latin{et al.}'' is auto-formatted
+using the \texttt{\textbackslash latin} command.
 
-OD600 measurements of cultures at 48h growth time were 0.65, 0.58 and 0.45 arb. units for WT, F206NY and F206pAzF cells.\\
+Multiple citations to be combined into a list can be given as
+a single citation.  This uses the \textsf{mciteplus} package
+\cite{Johnson1972,*Arduengo1992,*Eisenstein2005,*Arduengo1994}.
+Citations other than the first of the list should be indicated
+with a star. If the \textsf{mciteplus} package is not installed,
+the standard bibliography tools will still work but starred
+references will be ignored. Individual references can be referred
+to using \texttt{\textbackslash mciteSubRef}:
+``ref.~\mciteSubRef{Eisenstein2005}''.
 
-\textbf{Protein purification}\\
-Pelletized cells were reconstituted in TALON equilibration buffer (50mM sodium phosphate, 300mM NaCl) and microfluidized, also in equilibration buffer.
-Lysed cells were spun down at 15K RPM for 20 minutes at 4C.
-Supernatant was collected and combined with 300$\mu L$ bed volume TALON beads per 100mL cultured cells.
-TALON-supernatant mix was nutated for 30 minutes at 4C.
-Bound beads were then spun down at 500g for 5 minutes and the supernatant discarded.
-The beads were applied to a pre-washed gravity-flow column at RT and washed twice with 10mL equilibration buffer.
-Elution was done using 2mL elution buffer (50mM sodium phosphate, 300mM NaCl, 250mM imidazole) into four 500$\mu L$ aliquots.
-Protein was then stored at 4C.\\
+The class also handles notes to be added to the bibliography.  These
+should be given in place in the document \bibnote{This is a note.
+The text will be moved the the references section.  The title of the
+section will change to ``Notes and References''.}.  As with
+citations, the text should be placed before punctuation.  A note is
+also generated if a citation has an optional note.  This assumes that
+the whole work has already been cited: odd numbering will result if
+this is not the case \cite[p.~1]{Cotton1999}.
 
-\textbf{Bradford assay}\\
-A Bio-Rad Bradford assay kit was used using the standard protocol.
-Bovine serum albumin was used as the reference protein to construct the standard curve.
-Concentrations from 125 $\mu g / mL$ to 2 mg/mL were explored.
-Samples were incubated with dye in cuvettes for 5 minutes at RT.
-Samples were then analyzed via spectrophotometer at 595 \textit{nm} and used to create the standard curve shown in Figure \ref{fig:bradford-assay}.
-Protein samples were then compared to this standard curve to estimate their concentration.\\
+\subsection{Floats}
 
-\textbf{Qualitative activity assay}\\
-10 $\mu g$ purified WT catalase was added to a 5\% solution of hydrogen peroxide in water. The mixture was then examined for bubble formation.\\
-
-\section{Results}
-DH10B cell lines with appropriate plasmids were grown for 48h and pelletized.
-Wildtype HPII-containing cells were purified using TALON resin.
-Purified wildtype protein and ncAA cell lysates are shown in Figure \ref{fig:crude-wt-gel}.
-The expected band corresponding to an 84-kD protein is seen across all lanes, verifying the presence of HPII wildtype in all cultures.
-Contaminant bands in the gel are very faint, indicating pure protein.\\
-
-A Bradford assay of wildtype purified protein was then conducted. The results, shown in Figure \ref{fig:bradford-assay}, further verify the presence of protein in the purified sample.\\
-
-A qualitative, rudimentary assay testing the activity of wildtype purified HPII catalase was conducted. This assay verified HPII-WT was active by producing bubbles.\\
+New float types are automatically set up by the class file.  The
+means graphics are included as follows (Scheme~\ref{sch:example}).  As
+illustrated, the float is ``here'' if possible.
+\begin{scheme}
+  Your scheme graphic would go here: \texttt{.eps} format\\
+  for \LaTeX\, or \texttt{.pdf} (or \texttt{.png}) for pdf\LaTeX\\
+  \textsc{ChemDraw} files are best saved as \texttt{.eps} files:\\
+  these can be scaled without loss of quality, and can be\\
+  converted to \texttt{.pdf} files easily using \texttt{eps2pdf}.\\
+  %\includegraphics{graphic}
+  \caption{An example scheme}
+  \label{sch:example}
+\end{scheme}
 
 \begin{figure}
-\centering
-\includegraphics[width=\linewidth]{figures/crude-wt-gel}
-\caption{Electrophoresis gel of wildtype expression.
-Serial dillution of WT protein purification shown in lanes three and four.
-Major band in purification lanes correspond to protein of approximately 80kD, as expected for a catalase single subunit.
-}
-\label{fig:crude-wt-gel}
+  As well as the standard float types \texttt{table}\\
+  and \texttt{figure}, the class also recognises\\
+  \texttt{scheme}, \texttt{chart} and \texttt{graph}.
+  \caption{An example figure}
+  \label{fgr:example}
 \end{figure}
 
-\begin{figure}
+Charts, figures and schemes do not necessarily have to be labelled or
+captioned.  However, tables should always have a title. It is
+possible to include a number and label for a graphic without any
+title, using an empty argument to the \texttt{\textbackslash caption}
+macro.
+
+The use of the different floating environments is not required, but
+it is intended to make document preparation easier for authors. In
+general, you should place your graphics where they make logical
+sense; the production process will move them if needed.
+
+\subsection{Math(s)}
+
+The \textsf{achemso} class does not load any particular additional
+support for mathematics.  If packages such as \textsf{amsmath} are
+required, they should be loaded in the preamble.  However,
+the basic \LaTeX\ math(s) input should work correctly without
+this.  Some inline material \( y = mx + c \) or $ 1 + 1 = 2 $
+followed by some display. \[ A = \pi r^2 \]
+
+It is possible to label equations in the usual way (Eq.~\ref{eqn:example}).
+\begin{equation}
+  \frac{\mathrm{d}}{\mathrm{d}x} \, r^2 = 2r \label{eqn:example}
+\end{equation}
+This can also be used to have equations containing graphical
+content. To align the equation number with the middle of the graphic,
+rather than the bottom, a minipage may be used.
+\begin{equation}
+  \begin{minipage}[c]{0.80\linewidth}
     \centering
-  \begin{minipage}{.4\textwidth}
-    \begin{tabular}{lll}
+    As illustrated here, the width of \\
+    the minipage needs to allow some  \\
+    space for the number to fit in to.
+    %\includegraphics{graphic}
+  \end{minipage}
+  \label{eqn:graphic}
+\end{equation}
+
+\section{Experimental}
+
+The usual experimental details should appear here.  This could
+include a table, which can be referenced as Table~\ref{tbl:example}.
+Notice that the caption is positioned at the top of the table.
+\begin{table}
+  \caption{An example table}
+  \label{tbl:example}
+  \begin{tabular}{ll}
     \hline
-      Sample  & OD590 & mg/mL  \\
+    Header one  & Header two  \\
     \hline
-      WT        & 2.18 & 2.74\\
-      %% F206NY    & 0.71 & 0.83\\
-      %% F206pAzF  & 1.51 & 1.88\\
+    Entry one   & Entry two   \\
+    Entry three & Entry four  \\
+    Entry five  & Entry five  \\
+    Entry seven & Entry eight \\
     \hline
   \end{tabular}
-  \end{minipage}
-  \begin{minipage}{.59\textwidth}
-    \includegraphics[width=\linewidth]{figures/bradford-standard-curve}
-  \end{minipage}
-  \caption{Bradford assay results. \textbf{Left:} Sample absorbance at 590 nm and extrapolated concentration. \textbf{Right:} Bradford standard curve.}
-  \label{fig:bradford-assay}
-\end{figure}
+\end{table}
 
-\subsection{References}
+Adding notes to tables can be complicated.  Perhaps the easiest
+method is to generate these using the basic
+\texttt{\textbackslash textsuperscript} and
+\texttt{\textbackslash emph} macros, as illustrated (Table~\ref{tbl:notes}).
+\begin{table}
+  \caption{A table with notes}
+  \label{tbl:notes}
+  \begin{tabular}{ll}
+    \hline
+    Header one                            & Header two \\
+    \hline
+    Entry one\textsuperscript{\emph{a}}   & Entry two  \\
+    Entry three\textsuperscript{\emph{b}} & Entry four \\
+    \hline
+  \end{tabular}
+
+  \textsuperscript{\emph{a}} Some text;
+  \textsuperscript{\emph{b}} Some more text.
+\end{table}
+
+The example file also loads the optional \textsf{mhchem} package, so
+that formulas are easy to input: \texttt{\textbackslash ce\{H2SO4\}}
+gives \ce{H2SO4}.  See the use in the bibliography file (when using
+titles in the references section).
+
+The use of new commands should be limited to simple things which will
+not interfere with the production process.  For example,
+\texttt{\textbackslash mycommand} has been defined in this example,
+to give italic, mono-spaced text: \mycommand{some text}.
+
+\section{Extra information when writing JACS Communications}
+
+When producing communications for \emph{J.~Am.\ Chem.\ Soc.}, the
+class will automatically lay the text out in the style of the
+journal. This gives a guide to the length of text that can be
+accommodated in such a publication. There are some points to bear in
+mind when preparing a JACS Communication in this way.  The layout
+produced here is a \emph{model} for the published result, and the
+outcome should be taken as a \emph{guide} to the final length. The
+spacing and sizing of graphical content is an area where there is
+some flexibility in the process.  You should not worry about the
+space before and after graphics, which is set to give a guide to the
+published size. This is very dependant on the final published layout.
+
+You should be able to use the same source to produce a JACS
+Communication and a normal article.  For example, this demonstration
+file will work with both \texttt{type=article} and
+\texttt{type=communication}. Sections and any abstract are
+automatically ignored, although you will get warnings to this effect.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% The "Acknowledgement" section can be given in all manuscript
+%% classes.  This should be given within the "acknowledgement"
+%% environment, which will make the correct section or running title.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{acknowledgement}
+
+Please use ``The authors thank \ldots'' rather than ``The
+authors would like to thank \ldots''.
+
+The author thanks Mats Dahlgren for version one of \textsf{achemso},
+and Donald Arseneau for the code taken from \textsf{cite} to move
+citations after punctuation. Many users have provided feedback on the
+class, which is reflected in all of the different demonstrations
+shown in this document.
+
+\end{acknowledgement}
 
@@
 \begin{suppinfo}
+
+This will usually read something like: ``Experimental procedures and
+characterization data for all new compounds. The class will
+automatically add a sentence pointing to the information on-line:
 
@@
 
-%% \begin{tocentry}
+\begin{tocentry}
 
-%% \end{tocentry}
+Some journals require a graphical entry for the Table of Contents.
+This should be laid out ``print ready'' so that the sizing of the
+text is correct.
+
+Inside the \texttt{tocentry} environment, the font used is Helvetica
+8\,pt, as required by \emph{Journal of the American Chemical
+Society}.
+
+The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
+permitted for  \emph{Journal of the American Chemical Society}
+graphical table of content entries. The box will not resize if the
+content is too big: instead it will overflow the edge of the box.
+
+This box and the associated title will always be printed on a
+separate page at the end of the document.
+
+\end{tocentry}
 
@@
 3 - Interpret the data
+
+Role of the lateral channel in catalase HPII of Escherichia coli, Loewen
+They suggest that the perpendicular channel is the entrance channel and the lateral channel is the exhaust channel, because:
+-enlarging the perpendicular channel increases peroxidatic activity
+-perpendicular seems optimized for bringin in H2O2
+-lateral channel is blocked in some catalases

### 2018-02-19 bb494/manuscript.tex

--- 2018-02-19 bb494/manuscript.tex
@@
 
+Role of the lateral channel in catalase HPII of Escherichia coli, Loewen
+They suggest that the perpendicular channel is the entrance channel and the lateral channel is the exhaust channel, because:
+-enlarging the perpendicular channel increases peroxidatic activity
+-perpendicular seems optimized for bringin in H2O2
+-lateral channel is blocked in some catalases
+
 Substrate Flow, Loewen
@@
 This would explain why opening the perpendicular channel decreases the rate
-
-Role of the lateral channel in catalase, loewen
-This paper discusses the lateral channel as an exhaust channel
-Proposes that perpendicular channel is the entrance, lateral channel is exit

### 2018-02-20 bb494/manuscript.tex
b
add intro layout

@@
 
+\section{New Introduction}
+Catalase is a
+
+Understanding the optimizations of ubiquitous enzymes tells us how nature optimizes enzymes, and perhaps how faster-than-diffusion enzymes work
+In particular, understanding how the long-range electric fields between two molecules might be used for biological purpose is interesting.
+Typically amino acids are quantified by their polarity, pka, bulk.
+E-field is a new mechanism of interest.
+Catalase's tight perpendicular channel has a variety of optimizations:
+   it evolved to concentrate H2O2 near its channel
+   it is small, only allowing a couple solvent molecules through at once
+   it uses two channels, one for reactant entrance and the other for product exit
+   it uses hydrophobicity to prevent water from getting in, optimizing the H2O2 concentration in its heme pocket
+   It uses an electric field to orient the chain of solvent molecules for optimal transmission and chemistry, even away from the heme
+
 \section{Methods}
@@
 This would explain why opening the perpendicular channel decreases the rate
+
+An Electrical Potential in the Access Channel of Catalases Enhances Catalysis, loewen
+An electrical gradient between Asp181 and the positive heme orients the waters in the pocket, promoting H-bonding and rotating H2O2 properly

### 2018-02-21 bb494/manuscript.tex
more to intro

@@
 \section{New Introduction}
-Catalase is a
-
 Understanding the optimizations of ubiquitous enzymes tells us how nature optimizes enzymes, and perhaps how faster-than-diffusion enzymes work
@@
    It uses an electric field to orient the chain of solvent molecules for optimal transmission and chemistry, even away from the heme
+
+First para
+- reaction, relevance, structure, heme, conservation in nature
+
+2nd para
+- (talking about HPII) mechanism, channel, active site
+
+3rd para
+- catalase is FAST
+- evolutionary optimizations
+- why they are interesting
+- what we intend to study, and why it is intersesting
+
+4th para
+- our intended experiment, wrt the dielectric constant work
+
+5th para
+- our ncAA experiment, and how if we can destroy the dielectric variance of catalase via AA addition
+- how ncAAs are incorporated
+
+Catalase is an enzyme which catalyzes the breakdown of hydrogen peroxide into water and oxygen gas.
+This is an important detoxification reaction, since hydrogen peroxide is a reactive toxin generated by several biological processes.
+Catalase is a tetrameric protein which uses a heme group in its active site to facilitate the breakdown of hydrogen peroxide.
 
@@
 \section{Results}
+to add: Chemical nature and location of your ncAAs in the structure. (This might fit best in results).
 DH10B cell lines with appropriate plasmids were grown for 48h and pelletized.
@@
 An electrical gradient between Asp181 and the positive heme orients the waters in the pocket, promoting H-bonding and rotating H2O2 properly
+
+Want: 6-8 substrate concentrations; 3 repeats
+
+Hypothesis: can we control catalase catalytic efficiency by modulating dielectric constant, thus dampening the effects of electric field between Asp and the heme?
+salt vs dielectric const: Dielectric Relaxation of Aqueous NaCl Solutions, Buchner, May, 10.1021/jp982977k
+temp vs dielectric const: Dielectric Constant of Water from 0 0 to 1000 C , Malmberg, Mariott, eps = 87. 740 - 0.400081t + 9.398(10^-4)t^2 + 1.410(10^-6)t^3
+
+Diffusion-Controlled Reactions of Enzymes, Wei-Zhu Zhong
+klim: 1.8*10^-9
+
+EFFECT OF pH UPON THE REACTION KINETICS OF THE ENZYME-SUBSTRATE COMPOUNDS OF CATALASE*
+Catalase reaction rate is constant at pH 5-9
+
+New hypothesis: Catalase supposedly benefits from an electric field pointing through its active channel to properly position its substrate.
+We hypothesize that by destroying this electric field by inserting negatively charged groups near the heme, we can lower the reaction rate, possibly below the diffusion limit.
+Catalase has constant activity between pH 5 through 9.
+We will insert a group with pKa of 7, which should be neutral at pH 6 and mostly positive at pH 8.
+If such an enzyme experiences a drop in catalytic efficiency between pH 7 and 9, this will be evidence of the importance of an electric field orienting substrate molecules.
+
+3NY pka: 7.230 on the hydroxy group, Conversion of 3-nitrotyrosine to 3-aminotyrosine in peptides and proteins, Vallee
+
+Site-Specific Incorporation of 3-Nitrotyrosine as a Probe of pKa Perturbation of Redox-Active Tyrosines in Ribonucleotide Reductase, Stubbe
+pKa of 3NY can vary lots, between 7.5-10

### 2018-02-21 bb494/manuscript.tex
add errthing

@@
 
-5th para
-- our ncAA experiment, and how if we can destroy the dielectric variance of catalase via AA addition
-- how ncAAs are incorporated
-
 Catalase is an enzyme which catalyzes the breakdown of hydrogen peroxide into water and oxygen gas.
@@
 Catalase is a tetrameric protein which uses a heme group in its active site to facilitate the breakdown of hydrogen peroxide.
+The protein is
+
+Catalase's activity is nearly diffusion-limited \cite{kcatkm, difflimit}.
+This rate suggests that nearly every step in the substrate-enzyme interaction is optimized for speed.
+Understanding these optimizations is good because __________.
+Catalase has a variety of interesting optimizations in funneling substrate into its active pocket.
+The enzyme has a number of high-H2O2-residency surface residues near its active channel, which are believed to concentrate the substrate near the entrance to the active channel \cite{concentrateh2o2}.
+The active channel itself seems optimized to maximize substrate flow, to the point that mutagenesis to open the channel decreases enzyme activity \cite{substrateflow}.
+The channel seems to be capable of selecting the molecules which enter it using a patch of hydrophobic residues which are situated to preferentially pass the bulkier H2O2 \cite{molecularruler}.
+Even the directionality of substrate within the channel seems to be controlled: one channel is used for substrate entrance, and the other for product exit \cite{lateralchannel}.
+An new finding is that catalase might control not just the presence of catalase, but even its orientation as it enters the active site.
+It is believed that an electric field from a negative Asn residue to the positive heme acts to orient the H2O2 dipole, positioning the molecule such that it enters the active site at an angle ideal for reacting \cite{electricpotential}.
+It is this orientation that we hope to explore in this research.
 
@@
 pKa of 3NY can vary lots, between 7.5-10
+
+The Catalase-Hydrogen Peroxide System, Suggett
+Km: 1.1M,
+kcat/Km: 6.6*10^7/M/s

### 2018-02-21 bb494/manuscript.tex
minor

@@
 - reaction, relevance, structure, heme, conservation in nature
+- (talking about HPII) mechanism, channel, active site
 
 2nd para
-- (talking about HPII) mechanism, channel, active site
-
-3rd para
 - catalase is FAST
@@
 
-4th para
-- our intended experiment, wrt the dielectric constant work
+3rd para
+- our intended experiment, wrt
 
@@
 It is this orientation that we hope to explore in this research.
+
+The study on the electric potential in the active channel \cite{electricpotential} hypothesized the existance of an orientation potential, but ___flaws_in_experiment__.
+We hope to more thoroughly probe this effect using genetic code expansion ____ need to figure out how this mutation should influence the active site.
 

### 2018-02-22 bb494/manuscript.tex
finish intro?

@@
 This is an important detoxification reaction, since hydrogen peroxide is a reactive toxin generated by several biological processes.
-Catalase is a tetrameric protein which uses a heme group in its active site to facilitate the breakdown of hydrogen peroxide.
-It is a unique protein in that its reaction evolves gas, and thus its activity can be seen with the bare eye.
-The protein has been studied as a target for tumor suppression drugs \cite{bauer}.
-The enzyme is also useful industrially, in particular in the bleaching of textiles \cite{amorim}.
-
-The molecular mechaninism of catalase proceeds in two steps.
-First, the heme reduces a hydrogen peroxide, producing water and the so-called Cpd I.
-Then Cpd I oxidizes a second hydrogen peroxide, producing water and oxygen gas.
-This second step is mediated by a histidine near the active site \cite{alfonso-prieto}.
 
@@
 Catalase is a tetrameric protein which uses a heme group in its active site to facilitate the breakdown of hydrogen peroxide.
-The protein is
+The molecular mechaninism of catalase proceeds in two steps.
+First, the heme reduces a hydrogen peroxide, producing water and the so-called Cpd I.
+Then Cpd I oxidizes a second hydrogen peroxide, producing water and oxygen gas.
+This second step is mediated by a histidine near the active site \cite{alfonso-prieto}.
 
@@
 This rate suggests that nearly every step in the substrate-enzyme interaction is optimized for speed.
-Understanding these optimizations is good because __________.
-Catalase has a variety of interesting optimizations in funneling substrate into its active pocket.
+Understanding the optimizations of diffusion-limited proteins can provide many insights into how proteins achieve their functions.
+Catalase has a variety of interesting quirks for funneling substrate into its active pocket.
 The enzyme has a number of high-H2O2-residency surface residues near its active channel, which are believed to concentrate the substrate near the entrance to the active channel \cite{concentrateh2o2}.
@@
 Even the directionality of substrate within the channel seems to be controlled: one channel is used for substrate entrance, and the other for product exit \cite{lateralchannel}.
-An new finding is that catalase might control not just the presence of catalase, but even its orientation as it enters the active site.
+An new finding suggests that catalase might control not just the presence of hydrogen peroxide, but even its orientation as it enters the active site.
 It is believed that an electric field from a negative Asn residue to the positive heme acts to orient the H2O2 dipole, positioning the molecule such that it enters the active site at an angle ideal for reacting \cite{electricpotential}.
-It is this orientation that we hope to explore in this research.
+This use of an electric field over a long range to create an ``orienting field'' is a new and exciting usage of residues to achieve a function; understanding its role in catalase may help to understand the mechanisms of other proteins.
 
-The study on the electric potential in the active channel \cite{electricpotential} hypothesized the existance of an orientation potential, but ___flaws_in_experiment__.
-We hope to more thoroughly probe this effect using genetic code expansion ____ need to figure out how this mutation should influence the active site.
+In this study we hope to study the proposed orienting field optimization and its influence on catalase activity.
+In particular, we hope to carefully disrupt the wildtype active site potential environment by introducing a new charged residue, then examine the effect this disruption has on activity.
+For meaningful results, we must be careful to not introduce structural modifications to the enzyme.
+To ensure this, we will use genetic code expansion to incorporate 3-nitrotyrosine in place of tyrosine 206.
+The T206NY mutation will add only 30 $\angstrom^3$ of steric bulk \cite{3ntsize}, so should have minimal influence on active site structure.
+The theorized results of this addition are diagrammed in Figure \ref{fig:hypothesis}.
+To study the influence of field disruption on activity, we will study the activity of T206NY above and below its pKa of 6.8 \cite{3ntsize}.
+Wildtype catalase activity is constant under pH values from 5 to 10 \cite{phdependence,kcatkm}, so pH-dependence should only be seen if electric environment in the active site is important for activity.\\
+
+\begin{figure}
+  \caption{Depiction of catalase active pocket.
+\textbf{Left:} wildtype pocket with proposed ligand-orienting E-field \cite{electricpotential}.
+The field should orient H2O2 for optimal reaction with the heme.
+\textbf{Right:} modified HPII-3NY with a deprotonated 3-nitrotyrosine residue at site 206, with hypothetical augmented E-field.
+The disrupted E-field should result in un- or mis-aligned substrate.}
+  \label{fig:hypothesis}
+\end{figure}
 
@@
 10 $\mu g$ purified WT catalase was added to a 5\% solution of hydrogen peroxide in water. The mixture was then examined for bubble formation.\\
+
+\textbf{}
 
@@
 
-An Electrical Potential in the Access Channel of Catalases Enhances Catalysis, loewen
-An electrical gradient between Asp181 and the positive heme orients the waters in the pocket, promoting H-bonding and rotating H2O2 properly
-
-Want: 6-8 substrate concentrations; 3 repeats
-
-Hypothesis: can we control catalase catalytic efficiency by modulating dielectric constant, thus dampening the effects of electric field between Asp and the heme?
-salt vs dielectric const: Dielectric Relaxation of Aqueous NaCl Solutions, Buchner, May, 10.1021/jp982977k
-temp vs dielectric const: Dielectric Constant of Water from 0 0 to 1000 C , Malmberg, Mariott, eps = 87. 740 - 0.400081t + 9.398(10^-4)t^2 + 1.410(10^-6)t^3
-
-Diffusion-Controlled Reactions of Enzymes, Wei-Zhu Zhong
-klim: 1.8*10^-9
-
-EFFECT OF pH UPON THE REACTION KINETICS OF THE ENZYME-SUBSTRATE COMPOUNDS OF CATALASE*
-Catalase reaction rate is constant at pH 5-9
-
-New hypothesis: Catalase supposedly benefits from an electric field pointing through its active channel to properly position its substrate.
-We hypothesize that by destroying this electric field by inserting negatively charged groups near the heme, we can lower the reaction rate, possibly below the diffusion limit.
-Catalase has constant activity between pH 5 through 9.
-We will insert a group with pKa of 7, which should be neutral at pH 6 and mostly positive at pH 8.
-If such an enzyme experiences a drop in catalytic efficiency between pH 7 and 9, this will be evidence of the importance of an electric field orienting substrate molecules.
-
-3NY pka: 7.230 on the hydroxy group, Conversion of 3-nitrotyrosine to 3-aminotyrosine in peptides and proteins, Vallee
-
 Site-Specific Incorporation of 3-Nitrotyrosine as a Probe of pKa Perturbation of Redox-Active Tyrosines in Ribonucleotide Reductase, Stubbe
@@
 
-The Catalase-Hydrogen Peroxide System, Suggett
+\cite{kcatkm}
 Km: 1.1M,
 kcat/Km: 6.6*10^7/M/s
+
+possibly make a figure with the h2o2 and water from the electric potential paper?
+
+---------------
+Arguments about protein production, validation, and ncAA validation:
+We know we have wildtype protein because our gels have bands, indicating protein of the proper size was made with His tags
+We know our WT protein was active by quantitative assays
+
+We know our noncanonical protein was expressed by similar

### 2018-02-22 bb494/manuscript.tex
more changes

@@
 \usepackage{amsmath}
+\usepackage{subcaption}
 
@@
 \section{Introduction}
-Catalase is an enzyme which catalyzes the reaction $2H_2O_2 \rightarrow O_2 + 2H_2O$.
-This is an important detoxification reaction, since hydrogen peroxide is a reactive toxin generated by several biological processes.
-
-This bimolecular substrate mechanism relies on the efficient transport of hydrogen peroxide into the deeply buried active site.
-This is believed to be aided by an electrical gradient over 12 \AA between an active site aspartic acid and the positive heme iron \cite{chelikani}.
-Mutagenisis of this Asp to positive residues decreases activity up to 90\% \cite{chelikani}.
-However, this study was accomplished using only a few canonical amino acid substitutions, and provides relatively weak evidence for the electric field hypothesis.
-
-In this study we seek to use the technique of bioorthogonal ligation to study the electric field hypothesis of catalase substrate transport.
-We will incorporate the noncanonical amino acid 3-nitrotyrosine into the active site, then use click chemistry to insert a gradient of charged particles into the active site.
-By using a large number of particles with charges varying from positive to negative, we can average out the effects of steric bulk and isolate the effect of field strength.
-The results of such a study would conclusively determine the importance of an electric field in the catalase active site.\\
-
-\section{New Introduction}
-Understanding the optimizations of ubiquitous enzymes tells us how nature optimizes enzymes, and perhaps how faster-than-diffusion enzymes work
-In particular, understanding how the long-range electric fields between two molecules might be used for biological purpose is interesting.
-Typically amino acids are quantified by their polarity, pka, bulk.
-E-field is a new mechanism of interest.
-Catalase's tight perpendicular channel has a variety of optimizations:
-   it evolved to concentrate H2O2 near its channel
-   it is small, only allowing a couple solvent molecules through at once
-   it uses two channels, one for reactant entrance and the other for product exit
-   it uses hydrophobicity to prevent water from getting in, optimizing the H2O2 concentration in its heme pocket
-   It uses an electric field to orient the chain of solvent molecules for optimal transmission and chemistry, even away from the heme
-
-First para
-- reaction, relevance, structure, heme, conservation in nature
-- (talking about HPII) mechanism, channel, active site
-
-2nd para
-- catalase is FAST
-- evolutionary optimizations
-- why they are interesting
-- what we intend to study, and why it is intersesting
-
-3rd para
-- our intended experiment, wrt
-
 Catalase is an enzyme which catalyzes the breakdown of hydrogen peroxide into water and oxygen gas.
@@
 To ensure this, we will use genetic code expansion to incorporate 3-nitrotyrosine in place of tyrosine 206.
-The T206NY mutation will add only 30 $\angstrom^3$ of steric bulk \cite{3ntsize}, so should have minimal influence on active site structure.
+The T206NY mutation will add only 30 $\AA^3$ of steric bulk \cite{3ntsize}, so should have minimal influence on active site structure.
 The theorized results of this addition are diagrammed in Figure \ref{fig:hypothesis}.
@@
   \centering
-\includegraphics[width=0.7\linewidth]{figures/pure-gel}
-\caption{Electrophoresis gel of crude lysate and purified protein expression.
+  \begin{subfigure}{0.49\textwidth}
+    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
+    \begin{minipage}{0.9\textwidth}\includegraphics[width=0.9\linewidth]{figures/gene-diagram}\end{minipage}
+  \end{subfigure}
+  \begin{subfigure}{0.49\textwidth}
+    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
+    \begin{minipage}{0.9\textwidth}\includegraphics[width=0.9\linewidth]{figures/plasmid-diagram}\end{minipage}
+  \end{subfigure}
+  \\
+  \vspace{2mm}
+  \begin{subfigure}{\textwidth}
+    \centering
+    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
+    \begin{minipage}{0.9\textwidth}\centering\includegraphics[width=0.7\linewidth]{figures/pure-gel}\end{minipage}
+  \end{subfigure}
+  \caption{Schematic of protein production.
+\textbf{a.)} Gene constructs used.
+\textbf{b.)} Plasmid schematics for relevant plasmids.
+Amp$^R$ and Tet$^R$ are antibiotic resistance genes, Nitro-Y-RS and pCNF-RS are tRNA synthases compatible with 3-nitrotyrosine and p-azido-l-phenylalanine, and tRNA$_{CUA}$ is a tRNA compatible with the respective synthase.
+\textbf{c.} Electrophoresis gel of crude lysate and purified protein expression.
 Major band in purification lanes correspond to protein of approximately 80kD, as expected for a catalase single subunit.

### 2018-03-05 bb494/manuscript.tex
all
minor changes
add new changes

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title[]{Click chemistry reveals an alternate mechanism for HPII catalase}
+\title[]{Exploring the role of electric field in the catalase active site}
 
@@
 \begin{abstract}
-  %% This is an example document for the \textsf{achemso} document
-  %% class, intended for submissions to the American Chemical Society
-  %% for publication. The class is based on the standard \LaTeXe\
-  %% \textsf{report} file, and does not seek to reproduce the appearance
-  %% of a published paper.
-
-  %% This is an abstract for the \textsf{achemso} document class
-  %% demonstration document.  An abstract is only allowed for certain
-  %% manuscript types.  The selection of \texttt{journal} and
-  %% \texttt{manuscript} will determine if an abstract is valid.  If
-  %% not, the class will issue an appropriate error.
-  Abstract abstract abstracts abstract.
-Abstract abstract abstracts abstract.
-Abstract abstract abstracts abstract.
-Abstract abstract abstracts abstract.
-Abstract abstract abstracts abstract.
-Abstract abstract abstracts abstract.
-Abstract abstract abstracts abstract.
-Abstract abstract abstracts abstract.\\
+  Catalase is a diffusion-limited enzyme, and thus has optimized the rates of each step in its interaction with substrate.
+Recent evidence suggests that one such optimization is the use of an electric field in the active site to orient substrates for ideal reaction with catalytic residues.
+Here we study the role of electric field in the catalytic active site by introducing minor electric purturbations through genetic code expansion.
+Initial results indicate that the incorporation of noncanonical polar residues into the active site does not destroy function.
+More work is needed to fully determine the influence of electric field in the active site.\\
 \end{abstract}
@@
 
-\textbf{}
+\textbf{Kinetic assays}
+Enzyme velocity was measured using a Vernier Gas Pressure Sensor.
+Reaction buffer (50mM sodium phosphate, 300mM NaCl, 5\% $H_2O_2$ pH modulated to values from 5 to 10 via addition of HCL or NaOH) was combined with enzyme (25 $\mu g/mL$, 50 $\mu g/mL$, 50$\mu g/mL$ for HPII, HPII-3NY, HPII-pAzF) to a total volume of $3mL$ in pressure sensor tube.
+Pressure was monitored at 1s intervals for 120 seconds.
+Data was collected using Vernier LoggerPro.\\
 
 \section{Results}
-to add: Chemical nature and location of your ncAAs in the structure. (This might fit best in results).
+%% to add: Chemical nature and location of your ncAAs in the structure. (This might fit best in results).
 DH10B cell lines with appropriate plasmids were grown for 48h and pelletized.
 Wildtype HPII-containing cells were purified using TALON resin.
-Purified wildtype protein and ncAA cell lysates are shown in Figure \ref{fig:pure-gel}.
+Purified wildtype protein and ncAA cell lysates are shown in Figure \ref{fig:pure-gel}.c.
 The expected band corresponding to an 84-kD protein is seen across all lanes, indicating the presence of a catalase single subunit in each cell line.
@@
 
-A Bradford assay of wildtype purified protein was then conducted. The results, shown in Figure \ref{fig:bradford-assay}, further verify the presence of protein in the purified sample.\\
-
-A qualitative, rudimentary assay testing the activity of wildtype purified HPII catalase was conducted. This assay verified HPII-WT was active by producing bubbles.\\
-
-\begin{figure}
+\begin{figure}[h!]
   \centering
@@
   \caption{Schematic of protein production.
-\textbf{a.)} Gene constructs used.
-\textbf{b.)} Plasmid schematics for relevant plasmids.
+\textbf{a.} Gene constructs used.
+\textbf{b.} Plasmid schematics for relevant plasmids.
 Amp$^R$ and Tet$^R$ are antibiotic resistance genes, Nitro-Y-RS and pCNF-RS are tRNA synthases compatible with 3-nitrotyrosine and p-azido-l-phenylalanine, and tRNA$_{CUA}$ is a tRNA compatible with the respective synthase.
@@
 
-\begin{figure}
+A Bradford assay of wildtype purified protein was then conducted. The results, shown in Figure \ref{fig:bradford-assay}, further verify the presence of protein in the purified sample.\\
+
+A qualitative, rudimentary assay testing the activity of wildtype purified HPII catalase was conducted. This assay verified HPII-WT was active by producing bubbles.\\
+
+\begin{figure}[h!]
   \centering
@@
 
+\begin{figure}[h!]
+  \centering
+  \begin{subfigure}{0.29\textwidth}
+    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
+  \begin{minipage}{.9\textwidth}
+    \begin{tabular}{lll}
+      \hline
+      Sample  & $K_M$ (M) & kcat ($s^{-1}$) \\
+      \hline
+      HPII        & $1.3$ & $2.3x10^6$\\
+      HPII-3NY    & $1.4$ & $6.9x10^6$\\
+      HPII-pAzF   & $1.7$ & $2.2x10^5$\\
+      \hline
+    \end{tabular}
+  \end{minipage}
+  \end{subfigure}
+  \begin{subfigure}{0.69\textwidth}
+    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
+  \begin{minipage}{.9\textwidth}
+    \includegraphics[width=\linewidth]{figures/fake-results.png}
+  \end{minipage}
+  \end{subfigure}
+  \caption{Initial kinetic assay results.
+\textbf{a.} Final kinetic data extrapolated from Michaelis-Menten plot via linear regression.
+\textbf{b.} Michaelis-Menten plot of enzyme catalytic ($\pm \sigma^2$).}
+  \label{fig:bradford-assay}
+\end{figure}
+
+\subsection{Discussion}
+
+\subsection{Conclusion}
+
 \subsection{References}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{suppinfo}
+%% \begin{suppinfo}
 
-\end{suppinfo}
+%% \end{suppinfo}
 

### 2018-03-06 bb494/manuscript.tex
Add pymol sessions

@@
 \begin{figure}
+  \begin{subfigure}{0.69\textwidth}
+    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
+    \begin{minipage}{.9\textwidth}
+      \includegraphics[width=0.5\linewidth]{figures/cartoon-visualization.png}%
+      \includegraphics[width=0.5\linewidth]{figures/channel-visualization.png}
+    \end{minipage}
+  \end{subfigure}
+  \begin{subfigure}{0.3\textwidth}
+    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
+    \begin{minipage}{.9\textwidth}
+      \includegraphics[width=\linewidth]{figures/active-site-visualization.png}%
+    \end{minipage}
+  \end{subfigure}
+
   \caption{Depiction of catalase active pocket.
@@
   \begin{minipage}{.9\textwidth}
-    \includegraphics[width=\linewidth]{figures/fake-results.png}
+    %% \includegraphics[width=\linewidth]{figures/fake-results.png}
   \end{minipage}
@@
 We know our noncanonical protein was expressed by similar
+
+new stuff
+---------------------------
+Structure of Catalase HPII From Escherichia coli, at 1.9 A˚ Resolution, Bravo + Fita
+Describes the channels in HPII
+``Channels might be long and narrow to promote the selectivity for H2O2 of HPII''
+
+--Find a way to get the 2 channels: main one (intake? distal?), 590-595 one (lateral?)

### 2018-03-09 bb494/manuscript.tex
stuff

--- 2018-02-22 bb494/manuscript.tex
@@
 \usepackage{amsmath}
-\usepackage{subcaption}
 
@@
 \section{Introduction}
+Catalase is an enzyme which catalyzes the reaction $2H_2O_2 \rightarrow O_2 + 2H_2O$.
+This is an important detoxification reaction, since hydrogen peroxide is a reactive toxin generated by several biological processes.
+
+This bimolecular substrate mechanism relies on the efficient transport of hydrogen peroxide into the deeply buried active site.
+This is believed to be aided by an electrical gradient over 12 \AA between an active site aspartic acid and the positive heme iron \cite{chelikani}.
+Mutagenisis of this Asp to positive residues decreases activity up to 90\% \cite{chelikani}.
+However, this study was accomplished using only a few canonical amino acid substitutions, and provides relatively weak evidence for the electric field hypothesis.
+
+In this study we seek to use the technique of bioorthogonal ligation to study the electric field hypothesis of catalase substrate transport.
+We will incorporate the noncanonical amino acid 3-nitrotyrosine into the active site, then use click chemistry to insert a gradient of charged particles into the active site.
+By using a large number of particles with charges varying from positive to negative, we can average out the effects of steric bulk and isolate the effect of field strength.
+The results of such a study would conclusively determine the importance of an electric field in the catalase active site.\\
+
+\section{New Introduction}
+Understanding the optimizations of ubiquitous enzymes tells us how nature optimizes enzymes, and perhaps how faster-than-diffusion enzymes work
+In particular, understanding how the long-range electric fields between two molecules might be used for biological purpose is interesting.
+Typically amino acids are quantified by their polarity, pka, bulk.
+E-field is a new mechanism of interest.
+Catalase's tight perpendicular channel has a variety of optimizations:
+   it evolved to concentrate H2O2 near its channel
+   it is small, only allowing a couple solvent molecules through at once
+   it uses two channels, one for reactant entrance and the other for product exit
+   it uses hydrophobicity to prevent water from getting in, optimizing the H2O2 concentration in its heme pocket
+   It uses an electric field to orient the chain of solvent molecules for optimal transmission and chemistry, even away from the heme
+
+First para
+- reaction, relevance, structure, heme, conservation in nature
+- (talking about HPII) mechanism, channel, active site
+
+2nd para
+- catalase is FAST
+- evolutionary optimizations
+- why they are interesting
+- what we intend to study, and why it is intersesting
+
+3rd para
+- our intended experiment, wrt
+
 Catalase is an enzyme which catalyzes the breakdown of hydrogen peroxide into water and oxygen gas.
@@
 To ensure this, we will use genetic code expansion to incorporate 3-nitrotyrosine in place of tyrosine 206.
-The T206NY mutation will add only 30 $\AA^3$ of steric bulk \cite{3ntsize}, so should have minimal influence on active site structure.
+The T206NY mutation will add only 30 $\angstrom^3$ of steric bulk \cite{3ntsize}, so should have minimal influence on active site structure.
 The theorized results of this addition are diagrammed in Figure \ref{fig:hypothesis}.
@@
 \centering
-  \begin{subfigure}{0.49\textwidth}
-    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
-    \begin{minipage}{0.9\textwidth}\includegraphics[width=0.9\linewidth]{figures/gene-diagram}\end{minipage}
-  \end{subfigure}
-  \begin{subfigure}{0.49\textwidth}
-    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
-    \begin{minipage}{0.9\textwidth}\includegraphics[width=0.9\linewidth]{figures/plasmid-diagram}\end{minipage}
-  \end{subfigure}
-  \\
-  \vspace{2mm}
-  \begin{subfigure}{\textwidth}
-    \centering
-    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
-    \begin{minipage}{0.9\textwidth}\centering\includegraphics[width=0.7\linewidth]{figures/pure-gel}\end{minipage}
-  \end{subfigure}
-  \caption{Schematic of protein production.
-\textbf{a.)} Gene constructs used.
-\textbf{b.)} Plasmid schematics for relevant plasmids.
-Amp$^R$ and Tet$^R$ are antibiotic resistance genes, Nitro-Y-RS and pCNF-RS are tRNA synthases compatible with 3-nitrotyrosine and p-azido-l-phenylalanine, and tRNA$_{CUA}$ is a tRNA compatible with the respective synthase.
-\textbf{c.} Electrophoresis gel of crude lysate and purified protein expression.
+\includegraphics[width=0.7\linewidth]{figures/pure-gel}
+\caption{Electrophoresis gel of crude lysate and purified protein expression.
 Major band in purification lanes correspond to protein of approximately 80kD, as expected for a catalase single subunit.
@@
 possibly make a figure with the h2o2 and water from the electric potential paper?
+make inkscape gene block diagram
 
@@
 
-We know our noncanonical protein was expressed by similar
+We know our noncanonical protein was expressed because we have lanes in the gel of the proper size.
+repurposed TAG codon competes with the TAG release factor, meaning we should get some truncated protein, because ours is N-terminally labelled.
+We don't, which is interesting.

### 2018-03-10 bb494/manuscript.tex

--- 2018-03-06 bb494/manuscript.tex
@@
 
+\textbf{pH-dependent assay}
+pH assays were conducted using the same kinetic method described above.
+Reactions occured in 3mL equilibration buffer (50mM sodium phosphate, 300mM NaCl, 5\% $H_2O_2$ pH modulated to values from 5 to 10 via addition of HCL or NaOH).
+On addition of 43$\mu g$ protein, the tubes were mixed and incubated at RT for five minutes, then brought to 5\% $H_2O_2$ and added to pressure assay mechanism.
+Trials were conducted in triplicate.\\
+
 \section{Results}
@@
 
+\begin{figure}[h!]
+  \centering
+  \begin{subfigure}{\textwidth}
+    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
+    \begin{minipage}{0.9\textwidth}\includegraphics[width=0.9\linewidth]{figures/kinetics.png}\end{minipage}
+  \end{subfigure}
+  \caption{\textbf{Michaelis-Menten plot of kinetic efficiencies of HPII variants.} Kinetic measurements conducted at 25C using the pressure-based assay defined above.}
+\label{fig:kinetics}
+\end{figure}
+
+\begin{figure}[h!]
+  \centering
+  \begin{subfigure}{0.49\textwidth}
+    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
+    \begin{minipage}{0.9\textwidth}\includegraphics[width=0.9\linewidth]{figures/wildtype-velocity.png}\end{minipage}
+  \end{subfigure}
+  \begin{subfigure}{0.49\textwidth}
+    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
+    \begin{minipage}{0.9\textwidth}\includegraphics[width=0.9\linewidth]{figures/3ny-velocity.png}\end{minipage}
+  \end{subfigure}
+  \caption{\textbf{pH dependent initial velocities of HPII variants.} Left: HPII catalase Right: HPII-3NY.}
+\label{fig:ph-dependent-velocities}
+\end{figure}
+
 \subsection{Discussion}
@@
 
-We know our noncanonical protein was expressed by similar
+We know our noncanonical protein was expressed because we have lanes in the gel of the proper size.
+repurposed TAG codon competes with the TAG release factor, meaning we should get some truncated protein, because ours is N-terminally labelled.
+We don't, which is interesting.
 

### 2018-03-16 bb494/manuscript.tex
more stuff

@@
 \usepackage{subcaption}
+\usepackage{gensymb}
 
@@
 \section{Methods}
-\textbf{Protein expression}\\
-Wildtype protein was expressed by transforming the plasmid pBad-HPII (Addgene ID 105839), containing WT HPII gene, into competent \textit{E.
+\textbf{Protein production}\\
+Wildtype protein was expressed by transforming the plasmid pBad-HPII (Fig \ref{fig:protein-design}.b, modified from Addgene ID 105839), containing WT HPII gene, into competent \textit{E.
 coli} DH10B cells via heat-shock.
-pBad-HPII plasmid includes ampicillin resistance, an imidazole promoter, and a 6xHis tag on the N-term of the HPII gene.
-Cells were then grown in autoinduction media with an arabinose inducer (0.1\% aspartate pH 7.5, 0.2\% glycerol, 100$\mu g / mL$ ampicillin, mineral salts, 0.05\% glucose, 20mM MgSO4, 0.05\% arabinose, trace metals and AA mix) for 48 hours, shaking, at 37C.
-Cells were then pelletized and stored at -80C.\\
+pBad-HPII plasmid includes ampicillin resistance, an arabinose-induced promoter, and a 6xHis codon sequence on the N-term of the HPII gene.
+Cells were then grown in autoinduction media with an arabinose inducer (0.1\% aspartate, 0.2\% glycerol, 100$\mu g / mL$ ampicillin, mineral salts, 0.05\% glucose, 20mM MgSO4, 0.05\% arabinose, trace metals and amino acid mix, pH 7.5) for 48 hours, shaking, at 37\degree C.
+Cells were then pelletized and stored at -80\degree C.\\
 
-The same protocol was followed for the ncAA mutant catalases F206NY and F206pAzF, except protein plasmid pBad-HPII TAG206 (Addgene ID 105846) was used as a base.
-ncAA translation machinery plasmids pDule-3-nitrotyrosine (Addgene ID 85498) and pDule-pCNF (85494), which contain a tetracycline resistance gene, a TAG tRNA, and a compatible tRNA synthase, were co-transformed into DH10B along with pBAD-HPII TAG206 for respective F206NY and F206pAzF cell lines.
+The same protocol was followed for the ncAA mutant catalases HPII-3NY and HPII-pAzF, except using protein plasmids pBad-HPII-3NY and pBad-HPII-pAzF (Addgene ID 105846) instead.
+ncAA tRNA synthase plasmids pDule-Nitro-RS and pDule-pCNF-RS (Fig \ref{fig:protein-design}.b, modified from Addgene ID 85498), which contain a tetracycline resistance gene, a TAG tRNA, and a compatible tRNA synthase, were also transformed into DH10B.
 ncAA autoinduction media was supplemented with $25\mu g / mL$ tetracycline to select for both plasmids.\\
 
-OD600 measurements of cultures at 48h growth time were 0.65, 0.58 and 0.45 arb. units for WT, F206NY and F206pAzF cells.\\
+$OD_{600}$ measurements of cultures at 48 hours growth time were 0.65, 0.58 and 0.45 optical density for HPII, HPII-3NY and HPII-pAzF expressions.\\
 
 \textbf{Protein purification}\\
-Pelletized cells were reconstituted in TALON equilibration buffer (50mM sodium phosphate, 300mM NaCl) and microfluidized, also in equilibration buffer.
-Lysed cells were spun down at 15K RPM for 20 minutes at 4C.
+Pelletized cells were reconstituted in TALON equilibration buffer (50mM sodium phosphate, 300mM NaCl, pH 7.0) and microfluidized, also in equilibration buffer.
+Lysed cells were spun down at 15K RPM for 20 minutes at 4\degree C.
 Supernatant was collected and combined with 300$\mu L$ bed volume TALON beads per 100mL cultured cells.
-TALON-supernatant mix was nutated for 30 minutes at 4C.
+TALON-supernatant mix was nutated for 30 minutes at 4\degree C.
 Bound beads were then spun down at 500g for 5 minutes and the supernatant discarded.
-The beads were applied to a pre-washed gravity-flow column at RT and washed twice with 10mL equilibration buffer.
-Elution was done using 2mL elution buffer (50mM sodium phosphate, 300mM NaCl, 250mM imidazole) into four 500$\mu L$ aliquots.
-Protein was then stored at 4C.\\
+The beads were applied to a pre-washed gravity-flow column at room temperature and washed twice with 10mL equilibration buffer.
+Elution was done using 2mL elution buffer (50mM sodium phosphate, 300mM NaCl, 250mM imidazole, pH 7.0) into four 500$\mu L$ aliquots.
+Protein was then stored at 4\degree C.\\
 
-\textbf{Bradford assay}\\
-A Bio-Rad Bradford assay kit was used using the standard protocol.
+\textbf{Concentration determination}\\
+A Bio-Rad Bradford assay kit was employed using the standard protocol.
 Bovine serum albumin was used as the reference protein to construct the standard curve.
 Concentrations from 125 $\mu g / mL$ to 2 mg/mL were explored.
-Samples were incubated with dye in cuvettes for 5 minutes at RT.
-Samples were then analyzed via spectrophotometer at 595 \textit{nm} and used to create the standard curve shown in Figure \ref{fig:bradford-assay}.
-Protein samples were then compared to this standard curve to estimate their concentration.\\
+Samples were incubated with dye in cuvettes for 5 minutes at room temperature.
+Samples were then analyzed via spectrophotometer at 595 \textit{nm} and used to create the standard curve.
+HPII protein samples were then compared to this standard curve to estimate their concentration.\\
 
-\textbf{Qualitative activity assay}\\
-10 $\mu g$ purified WT catalase was added to a 5\% solution of hydrogen peroxide in water. The mixture was then examined for bubble formation.\\
-
-\textbf{Kinetic assays}
+\textbf{Activity assessment}\\
 Enzyme velocity was measured using a Vernier Gas Pressure Sensor.
-Reaction buffer (50mM sodium phosphate, 300mM NaCl, 5\% $H_2O_2$ pH modulated to values from 5 to 10 via addition of HCL or NaOH) was combined with enzyme (25 $\mu g/mL$, 50 $\mu g/mL$, 50$\mu g/mL$ for HPII, HPII-3NY, HPII-pAzF) to a total volume of $3mL$ in pressure sensor tube.
-Pressure was monitored at 1s intervals for 120 seconds.
+Reaction buffer (50mM sodium phosphate, 300mM NaCl, 5\% $H_2O_2$ pH modulated to 5.5, 6.25, 7.0, 7.75, 8.5 via addition of HCL or NaOH) was combined with enzyme (25 $\mu g/mL$, 50 $\mu g/mL$, 50$\mu g/mL$ for HPII, HPII-3NY, HPII-pAzF) to a total volume of $3mL$ in pressure sensor tube.
+Tubes were sealed and pressure was monitored at 1s intervals for 120 seconds.
 Data was collected using Vernier LoggerPro.\\
 
-\textbf{pH-dependent assay}
+\textbf{pH-dependent assay}\\
 pH assays were conducted using the same kinetic method described above.
@@
 %% to add: Chemical nature and location of your ncAAs in the structure. (This might fit best in results).
-DH10B cell lines with appropriate plasmids were grown for 48h and pelletized.
-Wildtype HPII-containing cells were purified using TALON resin.
-Purified wildtype protein and ncAA cell lysates are shown in Figure \ref{fig:pure-gel}.c.
-The expected band corresponding to an 84-kD protein is seen across all lanes, indicating the presence of a catalase single subunit in each cell line.
-Contaminant bands in the gel are very faint, indicating pure protein.\\
+Noncanonical amino acids (ncAAs) were incorporated into protein constructs by mutating the desired codon to TAG, then adding TAG-recognizing ncAA-tRNA synthetases and compatible tRNAs to cells.
+His-tagged protein constructs were grown in DH10 \textit{E.
+coli} and purified using TALON metal affinity resin.
+An SDS-PAGE gel of purified protein is shown in Figure \ref{fig:pure-gel}.
+HPII is expected to be in its 84-kD monomeric state in the denaturing and reducing conditions of SDS-PAGE; an 80kD band is seen for all three purified HPII constructs.
+This verifies that HPII-ncAA cells properly produce noncanonical aminoacyl-tRNA, which is the only tRNA which recognizes the TAG codon and allows for full-length protein translation.
+Contaminant bands in the gel are faint but visible, suggesting protein of $\approx$ 90\% purity.
+ncAA-incorporated proteins were grown with and without ncAA to verify proper incorporation.
+Without ncAA, cells should lack any aminoacyl-tRNA for the TAG codon, and thus HPII-ncAA peptides should stall on the ribosome, resulting in truncated peptides of approximately 22kD.
+This is seen in Figure \ref{fig:pure-gel}, where HPII-ncAA lanes lacking ncAA lack a 80kD band but do have bands of approximately 25kD.\\
 
@@
   \end{subfigure}
-  \\
   \vspace{2mm}
@@
   \end{subfigure}
-  \caption{Schematic of protein production.
-\textbf{a.} Gene constructs used.
+  \caption{\textbf{Electrophoresis gel of crude lysate and purified protein expression.} \textbf{a.} Gene constructs used.
 \textbf{b.} Plasmid schematics for relevant plasmids.
 Amp$^R$ and Tet$^R$ are antibiotic resistance genes, Nitro-Y-RS and pCNF-RS are tRNA synthases compatible with 3-nitrotyrosine and p-azido-l-phenylalanine, and tRNA$_{CUA}$ is a tRNA compatible with the respective synthase.
-\textbf{c.} Electrophoresis gel of crude lysate and purified protein expression.
-Major band in purification lanes correspond to protein of approximately 80kD, as expected for a catalase single subunit.
-Crude negative control lanes lack 80kD band, consistent with noncanonical amino acid being necessary for noncanonical protein expression.}
+\textbf{c.}Major band in purification lanes correspond to protein of approximately 80kD, as expected for a catalase single subunit.
+Crude negative control lanes lack 80kD band, consistent with noncanonical amino acid being necessary for noncanonical protein expression.
+Crude negative control lanes have smaller protein bands, consistent with truncated protein formation.}
 \label{fig:pure-gel}
@@
 
-A Bradford assay of wildtype purified protein was then conducted. The results, shown in Figure \ref{fig:bradford-assay}, further verify the presence of protein in the purified sample.\\
+%% \begin{figure}[h!]
+%%   \centering
+%%   \begin{minipage}{.4\textwidth}
+%%     \begin{tabular}{lll}
+%%       \hline
+%%       Sample  & OD590 & mg/mL  \\
+%%       \hline
+%%       WT        & 2.18 & 2.74\\
+%%       %% F206NY    & 0.71 & 0.83\\
+%%       %% F206pAzF  & 1.51 & 1.88\\
+%%       \hline
+%%     \end{tabular}
+%%   \end{minipage}
+%%   \begin{minipage}{.59\textwidth}
+%%     \includegraphics[width=\linewidth]{figures/bradford-standard-curve}
+%%   \end{minipage}
+%%   \caption{Bradford assay results. \textbf{Left:} Sample absorbance at 590 nm and extrapolated concentration. \textbf{Right:} Bradford standard curve.}
+%%   \label{fig:bradford-assay}
+%% \end{figure}
 
-A qualitative, rudimentary assay testing the activity of wildtype purified HPII catalase was conducted. This assay verified HPII-WT was active by producing bubbles.\\
-
-\begin{figure}[h!]
-  \centering
-  \begin{minipage}{.4\textwidth}
-    \begin{tabular}{lll}
-      \hline
-      Sample  & OD590 & mg/mL  \\
-      \hline
-      WT        & 2.18 & 2.74\\
-      %% F206NY    & 0.71 & 0.83\\
-      %% F206pAzF  & 1.51 & 1.88\\
-      \hline
-    \end{tabular}
-  \end{minipage}
-  \begin{minipage}{.59\textwidth}
-    \includegraphics[width=\linewidth]{figures/bradford-standard-curve}
-  \end{minipage}
-  \caption{Bradford assay results. \textbf{Left:} Sample absorbance at 590 nm and extrapolated concentration. \textbf{Right:} Bradford standard curve.}
-  \label{fig:bradford-assay}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
-  \begin{subfigure}{0.29\textwidth}
-    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
-  \begin{minipage}{.9\textwidth}
-    \begin{tabular}{lll}
-      \hline
-      Sample  & $K_M$ (M) & kcat ($s^{-1}$) \\
-      \hline
-      HPII        & $1.3$ & $2.3x10^6$\\
-      HPII-3NY    & $1.4$ & $6.9x10^6$\\
-      HPII-pAzF   & $1.7$ & $2.2x10^5$\\
-      \hline
-    \end{tabular}
-  \end{minipage}
-  \end{subfigure}
-  \begin{subfigure}{0.69\textwidth}
-    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
-  \begin{minipage}{.9\textwidth}
-    %% \includegraphics[width=\linewidth]{figures/fake-results.png}
-  \end{minipage}
-  \end{subfigure}
-  \caption{Initial kinetic assay results.
+%% \begin{figure}[h!]
+%%   \centering
+%%   \begin{subfigure}{0.29\textwidth}
+%%     \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
+%%   \begin{minipage}{.9\textwidth}
+%%     \begin{tabular}{lll}
+%%       \hline
+%%       Sample  & $K_M$ (M) & kcat ($s^{-1}$) \\
+%%       \hline
+%%       HPII        & $1.3$ & $2.3x10^6$\\
+%%       HPII-3NY    & $1.4$ & $6.9x10^6$\\
+%%       HPII-pAzF   & $1.7$ & $2.2x10^5$\\
+%%       \hline
+%%     \end{tabular}
+%%   \end{minipage}
+%%   \end{subfigure}
+%%   \begin{subfigure}{0.69\textwidth}
+%%     \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
+%%   \begin{minipage}{.9\textwidth}
+%%     %% \includegraphics[width=\linewidth]{figures/fake-results.png}
+%%   \end{minipage}
+%%   \end{subfigure}
+%%   \caption{Initial kinetic assay results.
 \textbf{a.} Final kinetic data extrapolated from Michaelis-Menten plot via linear regression.
 \textbf{b.} Michaelis-Menten plot of enzyme catalytic ($\pm \sigma^2$).}
-  \label{fig:bradford-assay}
-\end{figure}
+%%   \label{fig:bradford-assay}
+%% \end{figure}
+
+%% \begin{figure}[h!]
+%%   \centering
+%%   \begin{subfigure}{\textwidth}
+%%     \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
+%%     \begin{minipage}{0.9\textwidth}\includegraphics[width=0.9\linewidth]{figures/kinetics.png}\end{minipage}
+%%   \end{subfigure}
+%%   \caption{\textbf{Michaelis-Menten plot of kinetic efficiencies of HPII variants.} Kinetic measurements conducted at 25\degree C using the pressure-based assay defined above.}
+%% \label{fig:kinetics}
+%% \end{figure}
 
@@
     \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
-    \begin{minipage}{0.9\textwidth}\includegraphics[width=0.9\linewidth]{figures/kinetics.png}\end{minipage}
-  \end{subfigure}
-  \caption{\textbf{Michaelis-Menten plot of kinetic efficiencies of HPII variants.} Kinetic measurements conducted at 25C using the pressure-based assay defined above.}
-\label{fig:kinetics}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
-  \begin{subfigure}{0.49\textwidth}
-    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
     \begin{minipage}{0.9\textwidth}\includegraphics[width=0.9\linewidth]{figures/wildtype-velocity.png}\end{minipage}
   \end{subfigure}
-  \begin{subfigure}{0.49\textwidth}
+  \begin{subfigure}{\textwidth}
     \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%

### 2018-03-16 bb494/manuscript.tex
all

@@
 \section{Introduction}
-Catalase is an enzyme which catalyzes the breakdown of hydrogen peroxide into water and oxygen gas.
+%% Catalase is an enzyme which catalyzes the breakdown of hydrogen peroxide into water and oxygen gas.
 This is an important detoxification reaction, since hydrogen peroxide is a reactive toxin generated by several biological processes.
@@
 
-Catalase's activity is nearly diffusion-limited \cite{kcatkm, difflimit}.
+%% Catalase's activity is nearly diffusion-limited \cite{kcatkm, difflimit}.
 This rate suggests that nearly every step in the substrate-enzyme interaction is optimized for speed.
@@
 
-In this study we hope to study the proposed orienting field optimization and its influence on catalase activity.
+%% In this study we hope to study the proposed orienting field optimization and its influence on catalase activity.
 In particular, we hope to carefully disrupt the wildtype active site potential environment by introducing a new charged residue, then examine the effect this disruption has on activity.
@@
 
+Catalase is an enzyme which catalyzes the breakdown of hydrogen peroxide into water and oxygen gas.
+This is an important detoxification reaction, since hydrogen peroxide is a reactive toxin generated by several biological processes.
+Catalase is a tetrameric protein which uses a heme group in its active site to facilitate the breakdown of hydrogen peroxide.
+The molecular mechaninism of catalase proceeds in two steps.
+First, the heme reduces a hydrogen peroxide, producing water and the so-called Cpd I.
+Then Cpd I oxidizes a second hydrogen peroxide, producing water and oxygen gas.
+This second step is mediated by a histidine near the active site \cite{alfonso-prieto}.\\
+
+One remaining question about catalase is how it achieves its remarkable speed given its buried active site.
+The enzyme is nearly diffusion-limited\cite{kcatkm, difflimit}, indicating it turns substrate to product at close to the optimal rate.
+However, in order to reach the active site, hydrogen peroxide must diffuse through a 30 $\AA$ tunnel\cite{substrateflow}.
+Thus, substrate must be transported very efficiently through the protein.
+Understanding how this transport process works could provide lots of insight into how enzymes traffic substrate into buried active sites.
+Catalase is one of the most studied proteins in existence, and so represents a good model protein for understanding substrate trafficking.\\
+
+Previous work on substrate trafficking has illuminated various stages in the process.
+The enzyme has a number of high-H2O2-residency surface residues, which are believed to concentrate the substrate near the entrance to the active channel \cite{concentrateh2o2}.
+The active channel itself seems optimized to maximize substrate flow, to the point that mutagenesis to open the channel decreases enzyme activity \cite{substrateflow}.
+The channel seems to be capable of selecting the molecules which enter it using a patch of hydrophobic residues which are situated to preferentially pass the bulkier H2O2 \cite{molecularruler}.
+Even the directionality of substrate within the channel seems to be controlled: one channel is used for substrate entrance, and the other for product exit \cite{lateralchannel}.
+However, how substrate is traficked once in the active site antechamber is still an open question.\\
+
+One tool proteins are known to use to speed up active site kinetics is the use of electric fields.
+Superoxide dismutase is known to have a strong E-field in its active site to stabilize dipolar intermediates \cite{conserved-as-efield-sod, concentrated-as-efield-sod}.
+Several other proteins are known to use a similar effect \cite{efield-review} (be more precise here).
+It has been suggested that catalase uses an electric field in its active site to orient and traffick hydrogen peroxide to its active heme \cite{electricpotential}.
+This model, proposed by Chelikani \textit{et.
+al}, theorizes that a strong electric field from the negative asparigine 181 at the top of the active site antechamber extends to the strongly positive heme iron, as shown in Figure \ref{fig:hypothesis}.d.
+This electric field would orient incoming hydrogen peroxides in the optimal rotation for reaction with heme.\\
+
+An issue with mutagenesis studies like Chelikani \textit{et.
+al.} is that catalase's active site is very sensitive to changes.
+Mutagenesis of many residues results in large changes in enzyme kinetics \cite{substrateflow}.
+Here we conduct a less invasive experiment to probe the role of electric field in the catalase active site.
+We use genetic code expansion \cite{hammill} to incorporate two functional groups onto F206 in the catalase active site, turning it into 3-nitrotyrosine (Figure \ref{fig:hypothesis}.e).
+This change is important because it a.) is tiny, adding only 30 $\AA^3$ of steric bulk \cite{3ntsize}, and b.) incorporates a hydroxy group of pKa 7.0 into the catalase active site \cite{3ntsize}.
+Catalase activity is pH-invariant in the range of 5 to 9 \cite{phdependence,kcatkm}; thus, by incorporating a group which gains charge at pH 7, we can precisely study the effect of charge in the catalase active site on enzyme kinetics without worrying about the undesired effects of mutagenesis.\\
+
 \begin{figure}
-  \begin{subfigure}{0.69\textwidth}
+  \begin{subfigure}{0.49\textwidth}
     \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
     \begin{minipage}{.9\textwidth}
-      \includegraphics[width=0.5\linewidth]{figures/cartoon-visualization.png}%
-      \includegraphics[width=0.5\linewidth]{figures/channel-visualization.png}
+      %% \includegraphics[width=0.5\linewidth]{figures/cartoon-visualization.png}%
+      \includegraphics[width=\linewidth]{figures/channel-visualization.png}
     \end{minipage}
   \end{subfigure}
-  \begin{subfigure}{0.3\textwidth}
+  \begin{subfigure}{0.5\textwidth}
     \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
     \begin{minipage}{.9\textwidth}
-      \includegraphics[width=\linewidth]{figures/active-site-visualization.png}%
+      \includegraphics[width=0.5\linewidth]{figures/active-site-visualization.png}%
+      \includegraphics[width=0.5\linewidth]{figures/active-site-visualization.png}%
     \end{minipage}
   \end{subfigure}
-
   \caption{Depiction of catalase active pocket.

### 2018-03-17 bb494/manuscript.tex
add efield figures
new stuff

@@
 al.} is that catalase's active site is very sensitive to changes.
-Mutagenesis of many residues results in large changes in enzyme kinetics \cite{substrateflow}.
-Here we conduct a less invasive experiment to probe the role of electric field in the catalase active site.
+Replacing residues in the active site can dramatically impact enzyme efficiency \cite{substrateflow}.
+A better technique would purturb the electric field without full amino acid mutagenesis.
+Here we propose a new method for studying the catalase active site: genetic code expansion.
+We demonstrate how noncanonical amino acids can be incorporated into the catalase active site, as described in \textit{Hamill et.
+al.}, and how these modifications allow a less drastic change in active site environment.
 We use genetic code expansion \cite{hammill} to incorporate two functional groups onto F206 in the catalase active site, turning it into 3-nitrotyrosine (Figure \ref{fig:hypothesis}.e).
@@
 \begin{figure}
-  \begin{subfigure}{0.49\textwidth}
+  \begin{subfigure}{0.4\textwidth}
     \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
@@
     \end{minipage}
+  \end{subfigure}%
+  \begin{subfigure}{0.6\textwidth}
+    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
+    \begin{minipage}{.9\textwidth}
+      \includegraphics[width=0.4\linewidth]{figures/efield-wt.pdf}%
+      \hspace{1cm}
+      \includegraphics[width=0.4\linewidth]{figures/efield-3ny.pdf}%
+    \end{minipage}
   \end{subfigure}
+  \vspace{3cm}
   \begin{subfigure}{0.5\textwidth}
@@
     \begin{minipage}{.9\textwidth}
-      \includegraphics[width=0.5\linewidth]{figures/active-site-visualization.png}%
-      \includegraphics[width=0.5\linewidth]{figures/active-site-visualization.png}%
+      \includegraphics[width=\linewidth]{figures/3ny.png}%
       \end{minipage}
@@
 
+ncAA substitution is a powerful technique which allows easy testing of hypotheses in regions which are very sensitive to structural change.
+This technique could be used to test -other-catalase-hypotheses-.\\
+
+Perhaps this technique could be used to study e-fields generally.\\
+
+We propose that the following residues be incorporated into the channel at sites - - -. Coupled with APBS simulations, these results could be a powerful test of the effects of catalase e-field.\\
+
 \subsection{References}

### 2018-03-18 bb494/manuscript.tex
add new template

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% This is a (brief) model paper using the achemso class
-%% The document class accepts keyval options, which should include
-%% the target journal and optionally the manuscript type.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[journal=jacsat,manuscript=article]{achemso}
+\documentclass[9pt,twocolumn,twoside]{pnas-new}
+% Use the lineno option to display guide line numbers if required.
+% Note that the use of elements such as single-column equations
+% may affect the guide line number alignment.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Place any additional packages needed here.  Only include packages
-%% which are essential, to avoid problems later. Do NOT use any
-%% packages which require e-TeX (for example etoolbox): the e-TeX
-%% extensions are not currently available on the ACS conversion
-%% servers.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
+\templatetype{pnasresearcharticle} % Choose template 
+% {pnasresearcharticle} = Template for a two-column research article
+% {pnasmathematics} = Template for a one-column mathematics article
+% {pnasinvited} = Template for a PNAS invited submission
+
 \usepackage{amsmath}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% If issues arise when submitting your manuscript, you may want to
-%% un-comment the next line.  This provides information on the
-%% version of every file you have used.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%\listfiles
+\setboolean{displaywatermark}{false}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Place any additional macros here.  Please use \newcommand* where
-%% possible, and avoid layout-changing macros (which are not used
-%% when typesetting).
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
+\title{Perturbing the catalase active site electric field via genetic code expansion}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Meta-data block
-%% ---------------
-%% Each author should be given as a separate \author command.
-%%
-%% Corresponding authors should have an e-mail given after the author
-%% name as an \email command. Phone and fax numbers can be given
-%% using \phone and \fax, respectively; this information is optional.
-%%
-%% The affiliation of authors is given after the authors; each
-%% \affiliation command applies to all preceding authors not already
-%% assigned an affiliation.
-%%
-%% The affiliation takes an option argument for the short name.  This
-%% will typically be something like "University of Somewhere".
-%%
-%% The \altaffiliation macro should be used for new address, etc.
-%% On the other hand, \alsoaffiliation is used on a per author basis
-%% when authors are associated with multiple institutions.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author{Elliott Capek}
-\affiliation{Department of Biochemistry, Oregon State University}
-\email{capeke@oregonstate.edu}
-\phone{(971) 207 8193}
-\author{Hassan Alnatah}
-\author{Laikana Ly}
-\author{Joey Orton}
-%% \alsoaffiliation[]{}
-\author{Aidan Estelle}
-%% \affiliation{Department of Biochemistry, Oregon State University}
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\author[a,1,2]{Elliott Capek}
+\author[a]{Laikana Ly}
+\author[a]{Hassan Alnatah}
+\author[a]{Joey Orton}
+\author[a]{Aiden Estelle}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The document title should be given as usual. Some journals require
-%% a running title from the author: this should be supplied as an
-%% optional argument to \title.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title[]{Exploring the role of electric field in the catalase active site}
+\affil[a]{Oregon State University Department of Biohchemistry and Biophysics}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Some journals require a list of abbreviations or keywords to be
-%% supplied. These should be set up here, and will be printed after
-%% the title and author information, if needed.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\keywords{Catalase, noncanonical amino acid, active site, channel flow}
+% Please give the surname of the lead author for the running footer
+\leadauthor{Capek}
+
+% Please include corresponding author, author contribution and author declaration information
+\equalauthors{\textsuperscript{1} Author wrote the manuscript.}
+\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: capeke\@ oregonstate.edu}
+
+% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
+\keywords{Noncanonical amino acid $|$ catalase $|$ electric field $|$ active site}
+
+\begin{abstract}
+  Catalase is a ubiquitous enzyme which breaks down hydrogen peroxide into oxygen gas and water.
+Numerous theories exist about how catalase achieves its high catalytic efficiency.
+One theory is that an electric field within catalase's active site orients the dipole moments of substrate optimally to react with heme.
+Here we demonstrate how genetic code expansion can be used to purturb the electric field within catalase.
+We do so by replacing Phe 206 with 3-nitrotyrosine (3NY).
+We find that this mutation doubles Michaelis constant $K_M$ to $571$ mmol, compared to wildtype $245$ mmol.
+Mutant Michaelis constant $kcat$ goes down by a factor of 100 to $300 s^{-1}$, compared to wildtype $36,100 s^{-1}$.
+By switching 3NY's protonation state by conducting kinetic assays at high and low pH, we find charge on 3NY significantly effects enzyme velocity.
+At high pH the initial velocity is roughly half the low-pH velocity.
+\end{abstract}
+
+%% \dates{This manuscript was compiled on \today}
+\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
 
 \begin{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The manuscript does not need to include \maketitle, which is
-%% executed automatically.  The document should begin with an
-%% abstract, if appropriate.  If one is given and should not be, the
-%% contents will be gobbled.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{abstract}
-  Catalase is a diffusion-limited enzyme, and thus has optimized the rates of each step in its interaction with substrate.
-Recent evidence suggests that one such optimization is the use of an electric field in the active site to orient substrates for ideal reaction with catalytic residues.
-Here we study the role of electric field in the catalytic active site by introducing minor electric purturbations through genetic code expansion.
-Initial results indicate that the incorporation of noncanonical polar residues into the active site does not destroy function.
-More work is needed to fully determine the influence of electric field in the active site.\\
-\end{abstract}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Start the main part of the manuscript here.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-%% Catalase is an enzyme which catalyzes the breakdown of hydrogen peroxide into water and oxygen gas.
-This is an important detoxification reaction, since hydrogen peroxide is a reactive toxin generated by several biological processes.
-Catalase is a tetrameric protein which uses a heme group in its active site to facilitate the breakdown of hydrogen peroxide.
-The molecular mechaninism of catalase proceeds in two steps.
-First, the heme reduces a hydrogen peroxide, producing water and the so-called Cpd I.
-Then Cpd I oxidizes a second hydrogen peroxide, producing water and oxygen gas.
-This second step is mediated by a histidine near the active site \cite{alfonso-prieto}.
+% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
+% You should only change this length when you've finalised the article contents.
+\verticaladjustment{-2pt}
 
-%% Catalase's activity is nearly diffusion-limited \cite{kcatkm, difflimit}.
-This rate suggests that nearly every step in the substrate-enzyme interaction is optimized for speed.
-Understanding the optimizations of diffusion-limited proteins can provide many insights into how proteins achieve their functions.
-Catalase has a variety of interesting quirks for funneling substrate into its active pocket.
-The enzyme has a number of high-H2O2-residency surface residues near its active channel, which are believed to concentrate the substrate near the entrance to the active channel \cite{concentrateh2o2}.
-The active channel itself seems optimized to maximize substrate flow, to the point that mutagenesis to open the channel decreases enzyme activity \cite{substrateflow}.
-The channel seems to be capable of selecting the molecules which enter it using a patch of hydrophobic residues which are situated to preferentially pass the bulkier H2O2 \cite{molecularruler}.
-Even the directionality of substrate within the channel seems to be controlled: one channel is used for substrate entrance, and the other for product exit \cite{lateralchannel}.
-An new finding suggests that catalase might control not just the presence of hydrogen peroxide, but even its orientation as it enters the active site.
-It is believed that an electric field from a negative Asn residue to the positive heme acts to orient the H2O2 dipole, positioning the molecule such that it enters the active site at an angle ideal for reacting \cite{electricpotential}.
-This use of an electric field over a long range to create an ``orienting field'' is a new and exciting usage of residues to achieve a function; understanding its role in catalase may help to understand the mechanisms of other proteins.
+\maketitle
+\thispagestyle{firststyle}
+\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
 
-%% In this study we hope to study the proposed orienting field optimization and its influence on catalase activity.
-In particular, we hope to carefully disrupt the wildtype active site potential environment by introducing a new charged residue, then examine the effect this disruption has on activity.
-For meaningful results, we must be careful to not introduce structural modifications to the enzyme.
-To ensure this, we will use genetic code expansion to incorporate 3-nitrotyrosine in place of tyrosine 206.
-The T206NY mutation will add only 30 $\AA^3$ of steric bulk \cite{3ntsize}, so should have minimal influence on active site structure.
-The theorized results of this addition are diagrammed in Figure \ref{fig:hypothesis}.
-To study the influence of field disruption on activity, we will study the activity of T206NY above and below its pKa of 6.8 \cite{3ntsize}.
-Wildtype catalase activity is constant under pH values from 5 to 10 \cite{phdependence,kcatkm}, so pH-dependence should only be seen if electric environment in the active site is important for activity.\\
-
-Catalase is an enzyme which catalyzes the breakdown of hydrogen peroxide into water and oxygen gas.
+% If your first paragraph (i.e.
+with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation.
+If this is the case, add \parshape=0 to the end of the list environment.
+\dropcap{C}atalase is an enzyme which catalyzes the breakdown of hydrogen peroxide into water and oxygen gas.
 This is an important detoxification reaction, since hydrogen peroxide is a reactive toxin generated by several biological processes.
@@
 
-\begin{figure}
-  \begin{subfigure}{0.4\textwidth}
-    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
-    \begin{minipage}{.9\textwidth}
-      %% \includegraphics[width=0.5\linewidth]{figures/cartoon-visualization.png}%
-      \includegraphics[width=\linewidth]{figures/channel-visualization.png}
-    \end{minipage}
-  \end{subfigure}%
-  \begin{subfigure}{0.6\textwidth}
-    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
-    \begin{minipage}{.9\textwidth}
-      \includegraphics[width=0.4\linewidth]{figures/efield-wt.pdf}%
-      \hspace{1cm}
-      \includegraphics[width=0.4\linewidth]{figures/efield-3ny.pdf}%
-    \end{minipage}
-  \end{subfigure}
-  \vspace{3cm}
-  \begin{subfigure}{0.5\textwidth}
-    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
-    \begin{minipage}{.9\textwidth}
-      \includegraphics[width=\linewidth]{figures/3ny.png}%
-      \end{minipage}
-  \end{subfigure}
+\begin{figure*}
+  \includegraphics[width=0.24\linewidth]{figures/cartoon-visualization.png}
+  \includegraphics[width=0.24\linewidth]{figures/channel-visualization.png}
+  \includegraphics[width=0.24\linewidth]{figures/efield-wt.pdf}
+  \includegraphics[width=0.24\linewidth]{figures/efield-3ny.pdf}%
   \caption{Depiction of catalase active pocket.
@@
   \label{fig:hypothesis}
+\end{figure*}
+
+\begin{figure}
+  \includegraphics[width=0.4\linewidth]{figures/3ny.png}
+  \hspace{0.1\linewidth}
+  \includegraphics[width=0.5\linewidth]{figures/pazf.png}
+  \caption{\textbf{Structures of two noncanonical amino acids used in study.} \textit{Left:} 3-nitrotyrosine, \textit{Right:} para-azidophenylalanine.}
 \end{figure}
-
-\section{Methods}
-\textbf{Protein production}\\
-Wildtype protein was expressed by transforming the plasmid pBad-HPII (Fig \ref{fig:protein-design}.b, modified from Addgene ID 105839), containing WT HPII gene, into competent \textit{E.
-coli} DH10B cells via heat-shock.
-pBad-HPII plasmid includes ampicillin resistance, an arabinose-induced promoter, and a 6xHis codon sequence on the N-term of the HPII gene.
-Cells were then grown in autoinduction media with an arabinose inducer (0.1\% aspartate, 0.2\% glycerol, 100$\mu g / mL$ ampicillin, mineral salts, 0.05\% glucose, 20mM MgSO4, 0.05\% arabinose, trace metals and amino acid mix, pH 7.5) for 48 hours, shaking, at 37\degree C.
-Cells were then pelletized and stored at -80\degree C.\\
-
-The same protocol was followed for the ncAA mutant catalases HPII-3NY and HPII-pAzF, except using protein plasmids pBad-HPII-3NY and pBad-HPII-pAzF (Addgene ID 105846) instead.
-ncAA tRNA synthase plasmids pDule-Nitro-RS and pDule-pCNF-RS (Fig \ref{fig:protein-design}.b, modified from Addgene ID 85498), which contain a tetracycline resistance gene, a TAG tRNA, and a compatible tRNA synthase, were also transformed into DH10B.
-ncAA autoinduction media was supplemented with $25\mu g / mL$ tetracycline to select for both plasmids.\\
-
-$OD_{600}$ measurements of cultures at 48 hours growth time were 0.65, 0.58 and 0.45 optical density for HPII, HPII-3NY and HPII-pAzF expressions.\\
-
-\textbf{Protein purification}\\
-Pelletized cells were reconstituted in TALON equilibration buffer (50mM sodium phosphate, 300mM NaCl, pH 7.0) and microfluidized, also in equilibration buffer.
-Lysed cells were spun down at 15K RPM for 20 minutes at 4\degree C.
-Supernatant was collected and combined with 300$\mu L$ bed volume TALON beads per 100mL cultured cells.
-TALON-supernatant mix was nutated for 30 minutes at 4\degree C.
-Bound beads were then spun down at 500g for 5 minutes and the supernatant discarded.
-The beads were applied to a pre-washed gravity-flow column at room temperature and washed twice with 10mL equilibration buffer.
-Elution was done using 2mL elution buffer (50mM sodium phosphate, 300mM NaCl, 250mM imidazole, pH 7.0) into four 500$\mu L$ aliquots.
-Protein was then stored at 4\degree C.\\
-
-\textbf{Concentration determination}\\
-A Bio-Rad Bradford assay kit was employed using the standard protocol.
-Bovine serum albumin was used as the reference protein to construct the standard curve.
-Concentrations from 125 $\mu g / mL$ to 2 mg/mL were explored.
-Samples were incubated with dye in cuvettes for 5 minutes at room temperature.
-Samples were then analyzed via spectrophotometer at 595 \textit{nm} and used to create the standard curve.
-HPII protein samples were then compared to this standard curve to estimate their concentration.\\
-
-\textbf{Activity assessment}\\
-Enzyme velocity was measured using a Vernier Gas Pressure Sensor.
-Reaction buffer (50mM sodium phosphate, 300mM NaCl, 5\% $H_2O_2$ pH modulated to 5.5, 6.25, 7.0, 7.75, 8.5 via addition of HCL or NaOH) was combined with enzyme (25 $\mu g/mL$, 50 $\mu g/mL$, 50$\mu g/mL$ for HPII, HPII-3NY, HPII-pAzF) to a total volume of $3mL$ in pressure sensor tube.
-Tubes were sealed and pressure was monitored at 1s intervals for 120 seconds.
-Data was collected using Vernier LoggerPro.\\
-
-\textbf{pH-dependent assay}\\
-pH assays were conducted using the same kinetic method described above.
-Reactions occured in 3mL equilibration buffer (50mM sodium phosphate, 300mM NaCl, 5\% $H_2O_2$ pH modulated to values from 5 to 10 via addition of HCL or NaOH).
-On addition of 43$\mu g$ protein, the tubes were mixed and incubated at RT for five minutes, then brought to 5\% $H_2O_2$ and added to pressure assay mechanism.
-Trials were conducted in triplicate.\\
 
@@
 
-\begin{figure}[h!]
+\begin{figure}
   \centering
-  \begin{subfigure}{0.49\textwidth}
-    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
-    \begin{minipage}{0.9\textwidth}\includegraphics[width=0.9\linewidth]{figures/gene-diagram}\end{minipage}
-  \end{subfigure}
-  \begin{subfigure}{0.49\textwidth}
-    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
-    \begin{minipage}{0.9\textwidth}\includegraphics[width=0.9\linewidth]{figures/plasmid-diagram}\end{minipage}
-  \end{subfigure}
-  \vspace{2mm}
-  \begin{subfigure}{\textwidth}
+  \includegraphics[width=\linewidth]{figures/gene-diagram}
+  \includegraphics[width=\linewidth]{figures/plasmid-diagram}
+  \caption{\textbf{Design of protein constructs and expression plasmids} \textit{top:} genes of three constructs used in study. \textit{bottom:} pBad and pDule plasmids.}
+  \label{fig:design}
+\end{figure}
+
+\begin{figure}
   \centering
-    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
-    \begin{minipage}{0.9\textwidth}\centering\includegraphics[width=0.7\linewidth]{figures/pure-gel}\end{minipage}
-  \end{subfigure}
+  \includegraphics[width=\linewidth]{figures/pure-gel}
   \caption{\textbf{Electrophoresis gel of crude lysate and purified protein expression.} \textbf{a.} Gene constructs used.
@@
 
-%% \begin{figure}[h!]
-%%   \centering
-%%   \begin{minipage}{.4\textwidth}
-%%     \begin{tabular}{lll}
-%%       \hline
-%%       Sample  & OD590 & mg/mL  \\
-%%       \hline
-%%       WT        & 2.18 & 2.74\\
-%%       %% F206NY    & 0.71 & 0.83\\
-%%       %% F206pAzF  & 1.51 & 1.88\\
-%%       \hline
-%%     \end{tabular}
-%%   \end{minipage}
-%%   \begin{minipage}{.59\textwidth}
-%%     \includegraphics[width=\linewidth]{figures/bradford-standard-curve}
-%%   \end{minipage}
-%%   \caption{Bradford assay results. \textbf{Left:} Sample absorbance at 590 nm and extrapolated concentration. \textbf{Right:} Bradford standard curve.}
-%%   \label{fig:bradford-assay}
-%% \end{figure}
+\begin{figure}
+  \centering
+  \includegraphics[width=0.9\linewidth]{figures/kinetics.png}
+  \begin{tabular}{lll}
+      \hline
+      Sample  & $K_M$ (M) & kcat ($s^{-1}$) \\
+      \hline
+      HPII        & $1.3$ & $2.3x10^6$\\
+      HPII-3NY    & $1.4$ & $6.9x10^6$\\
+      HPII-pAzF   & $1.7$ & $2.2x10^5$\\
+      \hline
+  \end{tabular}
+  \caption{\textbf{Michaelis-Menten plot of kinetic efficiencies of HPII variants.} Kinetic measurements conducted at 25\degree C using the pressure-based assay defined above.}
+  \label{fig:kinetics}
+\end{figure}
 
-%% \begin{figure}[h!]
-%%   \centering
-%%   \begin{subfigure}{0.29\textwidth}
-%%     \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
-%%   \begin{minipage}{.9\textwidth}
-%%     \begin{tabular}{lll}
-%%       \hline
-%%       Sample  & $K_M$ (M) & kcat ($s^{-1}$) \\
-%%       \hline
-%%       HPII        & $1.3$ & $2.3x10^6$\\
-%%       HPII-3NY    & $1.4$ & $6.9x10^6$\\
-%%       HPII-pAzF   & $1.7$ & $2.2x10^5$\\
-%%       \hline
-%%     \end{tabular}
-%%   \end{minipage}
-%%   \end{subfigure}
-%%   \begin{subfigure}{0.69\textwidth}
-%%     \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
-%%   \begin{minipage}{.9\textwidth}
-%%     %% \includegraphics[width=\linewidth]{figures/fake-results.png}
-%%   \end{minipage}
-%%   \end{subfigure}
-%%   \caption{Initial kinetic assay results.
-\textbf{a.} Final kinetic data extrapolated from Michaelis-Menten plot via linear regression.
-\textbf{b.} Michaelis-Menten plot of enzyme catalytic ($\pm \sigma^2$).}
-%%   \label{fig:bradford-assay}
-%% \end{figure}
-
-%% \begin{figure}[h!]
-%%   \centering
-%%   \begin{subfigure}{\textwidth}
-%%     \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
-%%     \begin{minipage}{0.9\textwidth}\includegraphics[width=0.9\linewidth]{figures/kinetics.png}\end{minipage}
-%%   \end{subfigure}
-%%   \caption{\textbf{Michaelis-Menten plot of kinetic efficiencies of HPII variants.} Kinetic measurements conducted at 25\degree C using the pressure-based assay defined above.}
-%% \label{fig:kinetics}
-%% \end{figure}
-
-\begin{figure}[h!]
+\begin{figure}
   \centering
-  \begin{subfigure}{\textwidth}
-    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
-    \begin{minipage}{0.9\textwidth}\includegraphics[width=0.9\linewidth]{figures/wildtype-velocity.png}\end{minipage}
-  \end{subfigure}
-  \begin{subfigure}{\textwidth}
-    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
-    \begin{minipage}{0.9\textwidth}\includegraphics[width=0.9\linewidth]{figures/3ny-velocity.png}\end{minipage}
-  \end{subfigure}
+  \includegraphics[width=0.9\linewidth]{figures/wildtype-velocity.png}
+  \includegraphics[width=0.9\linewidth]{figures/3ny-velocity.png}
   \caption{\textbf{pH dependent initial velocities of HPII variants.} Left: HPII catalase Right: HPII-3NY.}
@@
 
-\subsection{Discussion}
+\subsection*{Discussion}
 
-\subsection{Conclusion}
+\subsection*{Conclusion}
 
@@
 
-\subsection{References}
+\matmethods{
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The same is true for Supporting Information, which should use the
-%% suppinfo environment.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% \begin{suppinfo}
+\subsection*{Protein production}
+Wildtype protein was expressed by transforming the plasmid pBad-HPII (Fig \ref{fig:protein-design}.b, modified from Addgene ID 105839), containing WT HPII gene, into competent \textit{E.
+coli} DH10B cells via heat-shock.
+pBad-HPII plasmid includes ampicillin resistance, an arabinose-induced promoter, and a 6xHis codon sequence on the N-term of the HPII gene.
+Cells were then grown in autoinduction media with an arabinose inducer (0.1\% aspartate, 0.2\% glycerol, 100$\mu g / mL$ ampicillin, mineral salts, 0.05\% glucose, 20mM MgSO4, 0.05\% arabinose, trace metals and amino acid mix, pH 7.5) for 48 hours, shaking, at 37\degree C.
+Cells were then pelletized and stored at -80\degree C.\\
 
-%% \end{suppinfo}
+The same protocol was followed for the ncAA mutant catalases HPII-3NY and HPII-pAzF, except using protein plasmids pBad-HPII-3NY and pBad-HPII-pAzF (Addgene ID 105846) instead.
+ncAA tRNA synthase plasmids pDule-Nitro-RS and pDule-pCNF-RS (Fig \ref{fig:protein-design}.b, modified from Addgene ID 85498), which contain a tetracycline resistance gene, a TAG tRNA, and a compatible tRNA synthase, were also transformed into DH10B.
+ncAA autoinduction media was supplemented with $25\mu g / mL$ tetracycline to select for both plasmids.\\
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The appropriate \bibliography command should be placed here.
-%% Notice that the class file automatically sets \bibliographystyle
-%% and also names the section correctly.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+$OD_{600}$ measurements of cultures at 48 hours growth time were 0.65, 0.58 and 0.45 optical density for HPII, HPII-3NY and HPII-pAzF expressions.\\
+
+\subsection*{Protein purification}
+Pelletized cells were reconstituted in TALON equilibration buffer (50mM sodium phosphate, 300mM NaCl, pH 7.0) and microfluidized, also in equilibration buffer.
+Lysed cells were spun down at 15K RPM for 20 minutes at 4\degree C.
+Supernatant was collected and combined with 300$\mu L$ bed volume TALON beads per 100mL cultured cells.
+TALON-supernatant mix was nutated for 30 minutes at 4\degree C.
+Bound beads were then spun down at 500g for 5 minutes and the supernatant discarded.
+The beads were applied to a pre-washed gravity-flow column at room temperature and washed twice with 10mL equilibration buffer.
+Elution was done using 2mL elution buffer (50mM sodium phosphate, 300mM NaCl, 250mM imidazole, pH 7.0) into four 500$\mu L$ aliquots.
+Protein was not desalted.
+Protein was then stored at 4\degree C.\\
+
+\subsection*{Concentration determination}
+A Bio-Rad Bradford assay kit was employed using the standard protocol.
+Bovine serum albumin was used as the reference protein to construct the standard curve.
+Concentrations from 125 $\mu g / mL$ to 2 mg/mL were explored.
+Samples were incubated with dye in cuvettes for 5 minutes at room temperature.
+Samples were then analyzed via spectrophotometer at 595 \textit{nm} and used to create the standard curve.
+HPII protein samples were then compared to this standard curve to estimate their concentration.\\
+
+\subsection*{Activity assessment}
+Enzyme velocity was measured using a Vernier Gas Pressure Sensor.
+Reaction buffer (50mM sodium phosphate, 300mM NaCl, 5\% $H_2O_2$ pH modulated to 5.5, 6.25, 7.0, 7.75, 8.5 via addition of HCL or NaOH) was combined with enzyme (25 $\mu g/mL$, 50 $\mu g/mL$, 50$\mu g/mL$ for HPII, HPII-3NY, HPII-pAzF) to a total volume of $3mL$ in pressure sensor tube.
+Tubes were sealed and pressure was monitored at 1s intervals for 120 seconds.
+Data was collected using Vernier LoggerPro.\\
+
+\subsection*{pH dependent assay}
+pH assays were conducted using the same kinetic method described above.
+Reactions occured in 3mL equilibration buffer (50mM sodium phosphate, 300mM NaCl, 5\% $H_2O_2$ pH modulated to values from 5 to 10 via addition of HCL or NaOH).
+On addition of 43$\mu g$ protein, the tubes were mixed and incubated at RT for five minutes, then brought to 5\% $H_2O_2$ and added to pressure assay mechanism.
+Trials were conducted in triplicate.\\
+}
+
+\showmatmethods % Display the Materials and Methods section
+
+% \pnasbreak splits and balances the columns before the references.
+% If you see unexpected formatting errors, try commenting out this line
+% as it can run into problems with floats and footnotes on the final page.
+\pnasbreak
+
+% Bibliography
 \bibliography{references}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The "tocentry" environment can be used to create an entry for the
-%% graphical table of contents.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% \begin{tocentry}
-
-%% \end{tocentry}
-
 \end{document}
-
-results section mantra:
-1 - Explain what you did and then describe what you expected as an outcome
-2 - Describe the data and show the dta
-3 - Interpret the data
-
-Role of the lateral channel in catalase HPII of Escherichia coli, Loewen
-They suggest that the perpendicular channel is the entrance channel and the lateral channel is the exhaust channel, because:
--enlarging the perpendicular channel increases peroxidatic activity
--perpendicular seems optimized for bringin in H2O2
--lateral channel is blocked in some catalases
-
-Substrate Flow, Loewen
-There are H2O2 binding sites in the heme distal pocket and main channel
-There are several input channels; what are their purposes?
-There is a ``molecular ruler'' in the main channel which selects for H2O2 against water
-Active site mutations which allow more water in slow down the rxn rate
-This paper discusses unidirectional flow of H2O2 through the main channel
-
-Active and inhibited catalase structures, tainer
-This is the paper about the hydrophobic ``molecular ruler''
-A hydrophobic region distal to the heme cleft contains phenylalanines, tryptophan and two valines
-Authors propose that two waters (one h-bonded to human Asn148 and His75; other h-bonded to gln168 and asp128) form a ``ruler''
-This ``ruler'' is far enough apart that only H2O2 can stably occupy the distance betweeen, thus allowing only H2O2 to enter, not H2O
-This would explain why opening the perpendicular channel decreases the rate
-
-Site-Specific Incorporation of 3-Nitrotyrosine as a Probe of pKa Perturbation of Redox-Active Tyrosines in Ribonucleotide Reductase, Stubbe
-pKa of 3NY can vary lots, between 7.5-10
-
-\cite{kcatkm}
-Km: 1.1M,
-kcat/Km: 6.6*10^7/M/s
-
-possibly make a figure with the h2o2 and water from the electric potential paper?
-make inkscape gene block diagram
-
----------------
-Arguments about protein production, validation, and ncAA validation:
-We know we have wildtype protein because our gels have bands, indicating protein of the proper size was made with His tags
-We know our WT protein was active by quantitative assays
-
-We know our noncanonical protein was expressed because we have lanes in the gel of the proper size.
-repurposed TAG codon competes with the TAG release factor, meaning we should get some truncated protein, because ours is N-terminally labelled.
-We don't, which is interesting.
-
-new stuff
----------------------------
-Structure of Catalase HPII From Escherichia coli, at 1.9 A˚ Resolution, Bravo + Fita
-Describes the channels in HPII
-``Channels might be long and narrow to promote the selectivity for H2O2 of HPII''
-
---Find a way to get the 2 channels: main one (intake? distal?), 590-595 one (lateral?)

### 2018-03-18 bb494/manuscript.tex
add newest additions to methods section.

@@
 %% \dates{This manuscript was compiled on \today}
-\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
+\doi{BB494 Final Manuscript}
 
@@
 al.}, and how these modifications allow a less drastic change in active site environment.
-We use genetic code expansion \cite{hammill} to incorporate two functional groups onto F206 in the catalase active site, turning it into 3-nitrotyrosine (Figure \ref{fig:hypothesis}.e).
+We use genetic code expansion \cite{hammill} to incorporate two functional groups onto F206 in the catalase active site, turning it into 3-nitrotyrosine (3NY) (Figure \ref{fig:noncanonicals}).
 This change is important because it a.) is tiny, adding only 30 $\AA^3$ of steric bulk \cite{3ntsize}, and b.) incorporates a hydroxy group of pKa 7.0 into the catalase active site \cite{3ntsize}.
@@
   \caption{\textbf{Structures of two noncanonical amino acids used in study.} \textit{Left:} 3-nitrotyrosine, \textit{Right:} para-azidophenylalanine.}
+  \label{fig:noncanonicals}
 \end{figure}
@@
 \subsection*{Protein production}
-Wildtype protein was expressed by transforming the plasmid pBad-HPII (Fig \ref{fig:protein-design}.b, modified from Addgene ID 105839), containing WT HPII gene, into competent \textit{E.
-coli} DH10B cells via heat-shock.
+Wildtype protein was expressed in competent DH10B \textit{E.
+coli} transformed with th eplasmid pBad-HPII (Figure \ref{fig:design}.b, modified from Addgene ID 105839), containing WT HPII gene, These cells were acquired from Dr.
+Van Zee in the OSU teaching lab.
 pBad-HPII plasmid includes ampicillin resistance, an arabinose-induced promoter, and a 6xHis codon sequence on the N-term of the HPII gene.
-Cells were then grown in autoinduction media with an arabinose inducer (0.1\% aspartate, 0.2\% glycerol, 100$\mu g / mL$ ampicillin, mineral salts, 0.05\% glucose, 20mM MgSO4, 0.05\% arabinose, trace metals and amino acid mix, pH 7.5) for 48 hours, shaking, at 37\degree C.
-Cells were then pelletized and stored at -80\degree C.\\
+Cells were dilluted 1:150 into 250mL autoinduction media with an arabinose inducer (0.1\% aspartate, 0.2\% glycerol, 100$\mu g / mL$ ampicillin, mineral salts, 0.05\% glucose, 20mM MgSO4, 0.05\% arabinose, trace metals and amino acid mix, pH 7.5).
+A positive control was conducted using pBad-sfGFP (superfolder GFP) in a 1:100 dillution into 10mL of autoinduction media.
+HPII cells were grown for 48 hours, shaking, at 37\degree C, to an optical density of 0.65.
+Cells were then pelletized at 5000 rcf and stored at -80\degree C.\\
 
-The same protocol was followed for the ncAA mutant catalases HPII-3NY and HPII-pAzF, except using protein plasmids pBad-HPII-3NY and pBad-HPII-pAzF (Addgene ID 105846) instead.
+The same expression protocol was followed for the ncAA mutant catalase proteins HPII-3NY and HPII-pAzF, except using protein plasmids pBad-HPII-3NY and pBad-HPII-pAzF (Figure \ref{fig:design}.b, Addgene ID 105846) instead.
 ncAA tRNA synthase plasmids pDule-Nitro-RS and pDule-pCNF-RS (Fig \ref{fig:protein-design}.b, modified from Addgene ID 85498), which contain a tetracycline resistance gene, a TAG tRNA, and a compatible tRNA synthase, were also transformed into DH10B.
-ncAA autoinduction media was supplemented with $25\mu g / mL$ tetracycline to select for both plasmids.\\
-
-$OD_{600}$ measurements of cultures at 48 hours growth time were 0.65, 0.58 and 0.45 optical density for HPII, HPII-3NY and HPII-pAzF expressions.\\
+ncAA autoinduction media was supplemented with $25\mu g / mL$ tetracycline to select for both plasmids.
+DH10B cells transformed with both respective pBad and pDule plasmids, also obtained from Dr.
+Van Zee, were dilluted in a 1:150 dillution into 250mL autoinduction media.
+Cells were shaken for 30 minutes at 37\degree C, then were treated with respectively 3NY and pAzF to final concentrations of 1.5 mM each.
+Noncanonical amino acids were dissolved in weak NaOH.
+Mutant cell lines were then grown for 48 hours, shaking, at 37\degree C, to optical densities of 0.58 (HPII-3NY) and 0.45 (HPII-pAzF).
+Cells were then pelletized at 5000 rcf and stored at -80\degree C.\\
 
 \subsection*{Protein purification}
-Pelletized cells were reconstituted in TALON equilibration buffer (50mM sodium phosphate, 300mM NaCl, pH 7.0) and microfluidized, also in equilibration buffer.
+Each protein construct was purified the same way.
+Pelletized cells were reconstituted in 10mL TALON equilibration buffer (50mM sodium phosphate, 300mM NaCl, pH 7.0) and microfluidized, also in equilibration buffer.
 Lysed cells were spun down at 15K RPM for 20 minutes at 4\degree C.
-Supernatant was collected and combined with 300$\mu L$ bed volume TALON beads per 100mL cultured cells.
+Supernatant was collected and combined with 300 $\mu L$ bed volume TALON cobalt affinity resin beads, also stored in TALON equilibration buffer.
+TALON beads were pre-washed according to the manufacturer manual.
 TALON-supernatant mix was nutated for 30 minutes at 4\degree C.
-Bound beads were then spun down at 500g for 5 minutes and the supernatant discarded.
-The beads were applied to a pre-washed gravity-flow column at room temperature and washed twice with 10mL equilibration buffer.
+Bound beads were then spun down at 500g for 5 minutes and most of the supernatant discarded, leaving 10mL to keep the beads wet.
+The bead mix was applied to a pre-washed gravity-flow column at room temperature and washed twice with 10mL equilibration buffer.
 Elution was done using 2mL elution buffer (50mM sodium phosphate, 300mM NaCl, 250mM imidazole, pH 7.0) into four 500$\mu L$ aliquots.
@@
 Protein was then stored at 4\degree C.\\
+
+\subsection*{SDS-PAGE gel}
+SDS-PAGE gels (30\% polyacrylamide, 10\% ammonium persulfate, 10\% SDS solution, 1 mmol TEMED, 1.5M Tris ph8.8 for resolving, 0.5M Tris pH 6.8 for stacking) were cast at 4\% for stacking and 8\% for resolving.
+Protein was dyed using a solution of 125mM Tris-HCL, 4\% SDS, 20\% glycerol, 0.02\% bromophenol blue, 5\% beta mercaptoethanol, pH 6.8 and loaded in 20 $\mu L$ volume into lanes.
+Bio-Rad Precision Plus Dual Color at 7.5 $\mu L$ was used as ladder.
+Gels were run in a running buffer of 25mM Tris base, 0.19 M glycine, 0.1\% SDS at pH 8.3.
+Gels were run at 200V for 50 minutes.\\
 
@@
 
-\subsection*{Activity assessment}
-Enzyme velocity was measured using a Vernier Gas Pressure Sensor.
-Reaction buffer (50mM sodium phosphate, 300mM NaCl, 5\% $H_2O_2$ pH modulated to 5.5, 6.25, 7.0, 7.75, 8.5 via addition of HCL or NaOH) was combined with enzyme (25 $\mu g/mL$, 50 $\mu g/mL$, 50$\mu g/mL$ for HPII, HPII-3NY, HPII-pAzF) to a total volume of $3mL$ in pressure sensor tube.
-Tubes were sealed and pressure was monitored at 1s intervals for 120 seconds.
-Data was collected using Vernier LoggerPro.\\
+\subsection*{Qualitative kinetic assay}
+One hundred mg protein was combined with 5\% $H_2O_2$ in pH 7.0 water at 25 \degree C. Activity was assessed by noting gas evolution by eye.\\
 
-\subsection*{pH dependent assay}
-pH assays were conducted using the same kinetic method described above.
-Reactions occured in 3mL equilibration buffer (50mM sodium phosphate, 300mM NaCl, 5\% $H_2O_2$ pH modulated to values from 5 to 10 via addition of HCL or NaOH).
-On addition of 43$\mu g$ protein, the tubes were mixed and incubated at RT for five minutes, then brought to 5\% $H_2O_2$ and added to pressure assay mechanism.
-Trials were conducted in triplicate.\\
+\subsection*{pH dependent velocity assessment}
+Enzyme initial velocity was measured using a Vernier Gas Pressure Sensor.
+Reaction buffer (50mM sodium phosphate, 300mM NaCl, pH modulated to 5.5, 6.25, 7.0, 7.75, 8.5 via addition of HCl or NaOH) was combined with 43.3 $\mu g$ protein to a total volume of $3mL$ in pressure sensor tube.
+Tubes were incubated for 5 minutes at 25 \degree C.
+Tubes were then treated with 5\% $H_2O_2$ and immediately sealed.
+Pressure was monitored at 1s intervals for 120 seconds.
+Data was collected using Vernier LoggerPro.
+The initial five seconds of constant-slope data was used to calculate initial pressure per second data.
+Each pH trial was measured in triplicate.
+For each pH value, initial velocities were averaged and the standard deviation was used as the error bar.\\
+
+\subsection*{Kinetic constant determination}
+Reactions occured in 3mL equilibration buffer (50mM sodium phosphate, 300mM NaCl, pH 7.0).
+Protein was added (10 $\mu g$ HPII, 25 $\mu g$ HPII-pAzF, 100 $\mu g$ HPII-3NY) and incubated for 20 minutes.
+$H_2O_2$ was added in ceoncentrations of 0\%, 0.5\%, 1.0\%, 1.5\% 2.0\%, 2.5\%, 3.0\%, then the reaction flask was immediately capped.
+Pressure was monitored at 1s intervals for 120 seconds.
+Data was collected using Vernier LoggerPro.
+The initial five seconds of constant-slope data was used to calculate initial pressure per second data.
+Presures were converted to moles of gas via the ideal gas law.
+Each trial was measured in triplicate.\\
+
+Michaelis-Menten kinetics were assumed to calculate enzyme efficiency parameters.
+Each velocity vs time data point was divided by the respective enzyme concentration to normalize the data with respect to enzyme amount.
+Then enzymes were fit to the Michaelis Menten equation:
+
+\begin{equation*}
+  V_0 = k_{cat} \frac{[S]}{K_M+[S]}.
+\end{equation*}
+
+Using least squares linear regression through the Google Sheets Solver package, values for $k_{cat}$ and $K_M$ were acquired.
+The correlation coefficient was used as the fractional error of each value.\\
+
 }

### 2018-03-18 bb494/manuscript.tex
add apbs figures
add discussion section

@@
   \centering
+  \includegraphics[width=0.49\linewidth]{figures/apbs-ph5.png}%
+  \includegraphics[width=0.49\linewidth]{figures/apbs-ph8.png}
+  \includegraphics[width=0.49\linewidth]{figures/apbs-3ny-ph5.png}%
+  \includegraphics[width=0.49\linewidth]{figures/apbs-3ny-ph8.png}
+  \caption{\textbf{pH dependent potential fields in active site of HPII and HPII-3NY.} \textit{Top left:} HPII pH 5.5.
+\textit{Top right:} HPII pH 8.5.
+\textit{Bottom left:} HPII-3NY pH 5.5.
+\textit{Bottom right:} HPII-3NY pH 8.5.
+}
+  \label{fig:apbs}
+\end{figure}
+
+\begin{figure}
+  \centering
   \includegraphics[width=0.9\linewidth]{figures/wildtype-velocity.png}
@@
 \subsection*{Discussion}
+Here we demonstrate that genetic code expansion can be used to successfully perturb the active site of catalase.
+We introduced either an azido group or a hydroxy and nitro group by mutating Phe 206 to para-azidophenylalanine or 3-nitrotyrosine.
+We show that these mutants retain activity, indicating their addition does not cause misfolding of the protein.
+We then modulate pH to show that the protonation state of added groups has an impact on enzyme catalysis; this supports the idea that electric field plays a large role in catalase active site dynamics.\\
+
+The electric field hypothesis described in Chelikani \textit{et.
+al.} \cite{electricpotential} theorizes that an electric potential gradient running from the negative Asn 181 in the active site to the positive heme iron acts to orient the dipole moments of water and hydrogen peroxide in the active site.
+This orientation is said to speed up reaction with heme by optimizing $H_2O_2$ - heme iron angle.
+Catalase wildtype is pH-invariant in ranges from 5 to 9 \cite{phdependence}.
+Thus, by introducing a functional group with pKa of 7 through HPII-3NY, we can get a privileged look into the effects of charge in the catalase active site separate from other pH-dependent effects.
+We find that when HPII-3NY's 3NY hydroxy group is deprotonated and thus anionic, the enzyme experiences a statistically significant 50\% drop in activity compared to the fully protonated neutral state.
+Due to 3NY's proximity to the active site, this charge effect is likely to be disrupting the Asn181 - heme electric field and preventing orienting.
+To verify this intuition, we conducted electrostatic modelling of the mutant active site using the APBS tool \cite{apbs}.
+These models verified that the large gradient pointing from Asn 181 to heme iron is disrupted at high pH, but not low pH.
+This is an important step supporting the electric field hypothesis.\\
+
+Using the tools of electrostatic simulation and genetic code expansion, we propose a new study to conclusively test the role of electric field in the catalase active site.
+This study would entail doing a large number of polar noncanonical amino acid substitutions to residues in the catalase active site, both close to and far from the heme.
+APBS would allow the calculation of the expected impact on electric field of these substitutions.
+Using APBS data, a rough ``dipole orientation factor $D$'' could be calculated via\\
+
+\begin{equation*}
+  D = \int_S -\vec{\nabla} V(s) \cdot d\vec{s}.
+\end{equation*}
+
+Here V is the electrostatic potential field from APBS and the integration path is along the hydrogen-bonded substrate chain found from crystal structures in Chelikani \cite{electricpotential}.
+This $D$ factor would quantify the strength of the dipole orientation in the active site.
+The strength of correlation of this $D$ factor with enzymatic activity is indicative of the importance of electric field in catalase.
+This experiment would be especially powerful if coupled with crystallography to verify proper active site structure.\\
+
+Such a study would not only determine the role of electric field, but could answer more precise questions.
+Chelikani propose that the electric field mediates rate by influencing substrate dipole moments.
+However, other authors propose that the role of active site electric fields is not dynamical, but rather to stabilize dipolar transition states \cite{asdipolestabilization}.
+This latter theory would suggest that electric field is only relevant near the dipolar transition state, not in the whole active channel.
+Electrostatic simulation guiding genetic code expansion would allow the control of electric field necessary to test such hypotheses.\\
+
+Such techniques could also be used for the study of other active site electric fields, such as in superoxide dismuatase.
+Previous studies \cite{concentrated-as-efield-sod} also used mutagenesis, and so could benefit from the less invasive technique of genetic code expansion.\\
 

### 2018-03-18 bb494/manuscript.tex
Add conclusion section.

@@
 
-\subsection*{Conclusion}
-
-ncAA substitution is a powerful technique which allows easy testing of hypotheses in regions which are very sensitive to structural change.
-This technique could be used to test -other-catalase-hypotheses-.\\
-
-Perhaps this technique could be used to study e-fields generally.\\
-
-We propose that the following residues be incorporated into the channel at sites - - -. Coupled with APBS simulations, these results could be a powerful test of the effects of catalase e-field.\\
+\subsection*{Conclusions}
+Noncanonical amino acid substitution allows much more precise atom-level perturbation of protein structure than residue-level mutagenesis.
+This is an emerging tool which can be used to study phenomena not reachable in the past.
+Here we demonstrate how the tool can be used to study the electric field in catalase, and propose a larger scale study which could definitely show the importance of electric field.
+The use of electrostatic simulations provides a theoretical explanation for the results seen after mutagenesis.
+By combining theory with atom-level perturbations, much can be learned about protein electrostatics.
 

### 2018-09-27 test/report.tex
Add test files

\documentclass[a4paper, 12pt]{article}
\usepackage{fullpage} % changes the margin

\begin{document}
\noindent
\huge\textbf{Project report}
\normalsize

\section*{Introduction}

\section*{Objectives}

\begin{itemize}
\item Create a biologically realistic leaky integrate-and-fire (LIF) model of a cortical section which can reproduce single-neuron details from 2PI \cite{bellaypaper}, such as firing rate distribution and quiescent time, correlations (average cross-correlation, distance-dependent correlation and autocorrelation), branching parameter, and time-binned avalanche dynamics

\item Use biophysical hierarchy of neural connections \cite{rubinov} and represent the main pyramidal and interneuron types within cortex using realistic parameters \cite{objectworkingmemory}

\item Reproduce the theoretical synchronous-to-asynchronous phase transition \cite{munozlg} using a synaptic resources parameter tuned to avoid the bistable up-down regime and non-critical (neutral) avalanches. Other models have found a continuous (presumably not up/down) phase transition \cite{rubinov}, so it is likely we will be able to find one as well. Other models have remained within the bistable regime and thus see only an up/down transition with neutral avalanches \cite{neutraltheory}.

\item Use the 'exact' solution method \cite{exactsolution} to quickly simulate large networks, possibly making the model fast enough to explore the parameter space to find different network modes (eg oscillation, up/down bistability \cite{munozlg}), or to optimize via gradient descent

\item schizophrenia - something something either simulate PCP problems biophysically and try to reproduce the PCP phenotype OR manually just increase the repeat-firing of cortical neurons

\item Possibly introduce a spatial component to neurons to estimate LFP electrode measurements from the LIF network, allowing us to see if the model can be made to reproduce coherence potentials
\end{itemize}

\section*{Methods}

\subsection*{Modular hierarchy}
\subsection*{Connectivity}
\subsection*{Plasticity & synaptic resources}
\subsection*{Voltage equations}

\section*{TODO}
\begin{itemize}
\item Find papers on LIF cortical organization - how should we connect the modules? Should we do modules? What should they look like?
\end{itemize}

\cite{lippert}

\bibliographystyle{plain}
\bibliography{report}

\end{document}

### 2018-09-27 test/report.tex
add more stuff

@@
 \subsection*{Modular hierarchy}
+Recent models for self-organized criticality have used neural models with hundreds to thousands of tightly-connected neurons \cite{munozlg, rubinov}.
+We will use a similar scheme to compare with these results.\\
+
 \subsection*{Connectivity}
-\subsection*{Plasticity & synaptic resources}
+\subsection*{Plasticity \& synaptic resources}
 \subsection*{Voltage equations}
@@
 \item Find papers on LIF cortical organization - how should we connect the modules? Should we do modules? What should they look like?
+\item Think about what sorts of plasticity rules would accomodate the Munoz 'synaptic resources' property - a way to make modules have propensity to do module spikes if sufficiently excited, and also a refractory period where the whole module must recover
 \end{itemize}
-
-\cite{lippert}
 

### 2018-09-29 test/report.tex
bla

@@
 \item Think about what sorts of plasticity rules would accomodate the Munoz 'synaptic resources' property - a way to make modules have propensity to do module spikes if sufficiently excited, and also a refractory period where the whole module must recover
+\item Are there any 'more' biophysical models which could be used to get more interesting results?
+Anything which would reproduce interesting behavior which LIF cannot, eg realistic connectivity between pyramidal and interneurons?
+\end{itemize}
+
+\textbf{coherence potentials subproject}
+\begin{itemize}
+\item Could CPs be found in 2P data?
+\item Are there any patterns in MEA CP traces suggestive of how they arise?
+\item Could large-scale synchronous dynamics in the brain, eg oscillations or avalanches, \textit{create} hopfield networks in different places? Seems unlikely since CPs are \textit{really} similar
+\item Are CPs clustered in space, or are they a small set of repeating patterns which the brain likes to make in different places?
+Does injecting a CP into a network make it more likely for it to be seen at a later time?
+\item Are there any processes which could move a conceivably-CP-creating strong hopfield network \textit{spatially} through a network?
 \end{itemize}

### 2018-10-02 test/report.tex
mini commit

@@
 
-\item Use biophysical hierarchy of neural connections \cite{rubinov} and represent the main pyramidal and interneuron types within cortex using realistic parameters \cite{objectworkingmemory}
+\item Use biophysical hierarchy of neural connections \cite{rubinov} and represent the main excitatory pyramidal and inhibitory interneuron types within cortex using realistic parameters \cite{objectworkingmemory}.
 
@@
 \subsection*{Modular hierarchy}
-Recent models for self-organized criticality have used neural models with hundreds to thousands of tightly-connected neurons \cite{munozlg, rubinov}.
+Recent models for self-organized criticality have used networks of many neural modules with hundreds to thousands of tightly-connected neurons each \cite{munozlg, rubinov}.
 We will use a similar scheme to compare with these results.\\
+
+Each module will follow the 80/20 principle of excitatory and inhibitory neurons \cite{larremorerestrepo}.
+Brunel and Wang used a model which used biologically realistic values for membrane capacitance and channel time constant to describe neurons with excitatory, GABA or NMDA/AMPA channels \cite{objectworkingmemory}.
+Using realistic proportions of these neurons to organize a module would be a nice way to make the model biologically realistic.
 
@@
 \subsection*{Plasticity \& synaptic resources}
+It seems that many different papers which achieve self-organised criticality use ``dynamical variables'' of some sort (eg synapse weight, synaptic resource or gain) \cite{munozlg, rubinov, kinouchi}.
+These variables have lead to interesting dynamics by their ability to store and integrate information at a slower timescale than voltage.
+A recent report showed SOC in a model with a supposedly biologically realistic dynamic ``gain'' instead of dynamic synapses, and claimed it was much faster to simulate.
+Perhaps this would be a thing to look into \cite{kinouchi}.
+
 \subsection*{Voltage equations}
@@
 \begin{itemize}
-\item Could CPs be found in 2P data?
+\item Question: Could CPs be found in 2P data?
+\item Question: What do the neural populations which trigger LFP spikes look like?
 \item Are there any patterns in MEA CP traces suggestive of how they arise?
@@
 \item Are there any processes which could move a conceivably-CP-creating strong hopfield network \textit{spatially} through a network?
+\item Maybe a modular network with tightly-knit modules would be better at supporting CPs?
+Modules would then need to transport their gross dynamics to other modules, which could be an easier task than neurons transferring info to other neurons.
+Assuming homogenous spatially-distinct modules, if a module could enter a state where it A.) takes on some LFP-measurable trace and B.) is a stable attractor state and C.) connects in a way with other modules such that it can get them to enter into the state too, then the system could support CPs.
+\item If modules took on a basic oscillatory behavior which only lasts for a couple cycles, they could conceivably entrain other modules which happen to connect to oscillate at similar frequencies.
+Maybe the actual spatial location of modules is irrelevent for CPs, maybe the modules just need to be kinda close and randomly happen to ``move'' their activity?
+\item One thing seems clear: tightly-knit modules which can take on gross stable oscillatory states and send these behaviors to other modules could support CP dynamics
+\item Maybe small-world networks would be ideal for this transmission?
+Maybe only a single 'entrainment' between modules would be sufficient to 'send' the attractive signal between distant modules, and once sent, the new module could interact with its neighbors to cause them to go into (phase-lagged) oscillations.
+Maybe neighbors oscillating out-of-phase is enough to cause a CP, where a combo of their distance, phase pattern and amplitude could produce the right behavior?
+Something something fourier theorem about decomposing any signal into a bunch of sins and cosines?
+\item Requirements: (modules can enter transient oscillations) (modules can entrain their neighbors to oscillate at a different phase and amplitude through tight connections) (different modules transiently oscillating together can produce a unique LFP signal) (small-world connections allow these states to be sent over long distances) (somehow these several-module attractor modes are numerous to support numerous patterns, possibly by varying the oscillation phase or something)
 \end{itemize}

### 2018-10-03 test/report.tex
more

@@
 
-Each module will follow the 80/20 principle of excitatory and inhibitory neurons \cite{larremorerestrepo}.
+Each module will follow the 80/20 principle of excitatory and inhibitory neurons, which is important for information processing \cite{larremorerestrepo, entropyinhibition}.
 Brunel and Wang used a model which used biologically realistic values for membrane capacitance and channel time constant to describe neurons with excitatory, GABA or NMDA/AMPA channels \cite{objectworkingmemory}.
-Using realistic proportions of these neurons to organize a module would be a nice way to make the model biologically realistic.
+Using realistic proportions of these neurons to organize a module would be a nice way to make the model biologically realistic.\\
 
@@
 These variables have lead to interesting dynamics by their ability to store and integrate information at a slower timescale than voltage.
+In particular, Rubinov \textit{et al} describe how dynamic synapses allow slow mesoscopic oscillations which are abolished when synapses are frozen \cite{rubinov}.
 A recent report showed SOC in a model with a supposedly biologically realistic dynamic ``gain'' instead of dynamic synapses, and claimed it was much faster to simulate.
-Perhaps this would be a thing to look into \cite{kinouchi}.
+Perhaps this would be a thing to look into \cite{kinouchi}.\\
 
@@
 \item Are there any patterns in MEA CP traces suggestive of how they arise?
+Could you do a Fourier decomposition on each CP and see if they are similar to other non-CP LFP traces?
+Are the main FFT terms within a biophysical oscillation range?
+Do all the below-threshold LFP traces (including CPs) look like similar variations on each other which could be conceivably caused by an error-prone process?
 \item Could large-scale synchronous dynamics in the brain, eg oscillations or avalanches, \textit{create} hopfield networks in different places? Seems unlikely since CPs are \textit{really} similar
@@
 
+\textbf{Other ideas}
+\begin{itemize}
+\item Has anyone tried to use ML to decode information from avalanching cortex during an object recognition task? This could be a fun project for the ML final.
+\end{itemize}
+
 \bibliographystyle{plain}

### 2018-10-10 test/report.tex
add more

--- 2018-09-29 test/report.tex
@@
 \noindent
-\huge\textbf{Project report}
+\huge\textbf{Project proposal}\\
+
+\large Elliott Capek, \today
 \normalsize
 
-\section*{Introduction}
+\section*{Summary}
+Coherence potentials are activity patterns which propagate with high fidelity through space.
+Their transmission fidelity and information content suggests that they are not the result of a few neurons firing stochastically, but rather the synchronous dynamics of microcircuits.
+If microcircuit ``modules'' could be made to take on stable, synchronous, whole-module firing states, and if these firing states could be entrained onto distant modules through long-range connections, then this could be an explanation for coherence potentials.
+I propose to test this idea by creating a leaky integrate-and-fire network with modular connectivity, implementing plasticity rules known to produce modular synchrony, then studying the pseudo-LFP dynamics of the model using a heuristic method.
+By varying simulation conditions such as external currents, model parameters, connectivities between modules and plasticity rules, I hope to create a network with the following properties:
+
+\begin{itemize}
+\item Modules within the network can be entrained into stable synchronous states through external currents
+\item These stable states are varied in nature enough that their average activity is as complex as real coherence potential signals
+\item These states are ``contagious'' and can entrain other distantly-connected modules into the same state, perpetuating the coherence potential
+\end{itemize}
+
+If a network meets these three requirements, then it should demonstrate coherence potentials.
 
 \section*{Objectives}
-
 \begin{itemize}
-\item Create a biologically realistic leaky integrate-and-fire (LIF) model of a cortical section which can reproduce single-neuron details from 2PI \cite{bellaypaper}, such as firing rate distribution and quiescent time, correlations (average cross-correlation, distance-dependent correlation and autocorrelation), branching parameter, and time-binned avalanche dynamics
-
-\item Use biophysical hierarchy of neural connections \cite{rubinov} and represent the main pyramidal and interneuron types within cortex using realistic parameters \cite{objectworkingmemory}
-
+\item Create a biologically realistic \cite{objectworkingmemory, biologicalsoc} leaky integrate-and-fire (LIF) model of a cortical area which can reproduce basic single-neuron details from 2PI \cite{bellaypaper}, such as firing rate distribution, branching parameter, and time-binned avalanche dynamics.
+  \begin{itemize}
   \item Reproduce the theoretical synchronous-to-asynchronous phase transition \cite{munozlg} using a synaptic resources parameter tuned to avoid the bistable up-down regime and non-critical (neutral) avalanches.
@@
 Other models have remained within the bistable regime and thus see only an up/down transition with neutral avalanches \cite{neutraltheory}.
+  \end{itemize}
+\item Explore parameter values and model features, such as hierarchical modularity \cite{rubinov, mountcastle}, short-term plasticity \cite{dynamicsynapses} and spike time-dependent plasticity \cite{heterogeneousnetwork, biologicalsoc, rubinov} to achieve module (but not network)-wide synchronous behavior, such as oscillations \cite{munozlg} or module spikes \cite{rubinov}.
+  \begin{itemize}
+  \item Use the ``exact'' solution method to very rapidly simulate models at different parameter values, allowing a large parameter-space to be explored quickly \cite{exactsolution}
+  \item This is the step which will decide the project.
+If a module can be made to enter into stable, high-information states merely through input currents (externally or from other modules), then the hardest bridge will be crossed
+  \end{itemize}
+\item Implement a heuristic-LFP method for estimating the electrode measurements in the model network; perhaps starting with simply averaging local module activity.
+Yahya would probably know more about this.
+\item Devise a method for ``entraining'' desired modules to enter into stable states, for example by feeding the same custom waveform into all the neurons in a single module as an external current.
+It has been shown that ``tetanizing'' a network can modify it through plasticity to reproduce the tetanized signal in a phenomena known as ``ommitted stimulus potential'' \cite{heterogeneousnetwork}; perhaps this phenomena would work for other inputs
+\end{itemize}
 
-\item Use the 'exact' solution method \cite{exactsolution} to quickly simulate large networks, possibly making the model fast enough to explore the parameter space to find different network modes (eg oscillation, up/down bistability \cite{munozlg}), or to optimize via gradient descent
-
-\item schizophrenia - something something either simulate PCP problems biophysically and try to reproduce the PCP phenotype OR manually just increase the repeat-firing of cortical neurons
-
-\item Possibly introduce a spatial component to neurons to estimate LFP electrode measurements from the LIF network, allowing us to see if the model can be made to reproduce coherence potentials
-\end{itemize}
+%% \section*{Alternate Objectives}
+%% \begin{itemize}
+%% \item Create a biologically realistic leaky integrate-and-fire (LIF) model of a cortical section which can reproduce single-neuron details from 2PI \cite{bellaypaper}, such as firing rate distribution and quiescent time, correlations (average cross-correlation, distance-dependent correlation and autocorrelation), branching parameter, and time-binned avalanche dynamics
+%% \end{itemize}
 
@@
 \subsection*{Modular hierarchy}
-Recent models for self-organized criticality have used neural models with hundreds to thousands of tightly-connected neurons \cite{munozlg, rubinov}.
+
+Recent models for self-organized criticality have used networks of many neural modules with hundreds to thousands of tightly-connected neurons each \cite{munozlg, rubinov}.
 We will use a similar scheme to compare with these results.\\
 
+Each module will follow the 80/20 principle of excitatory and inhibitory neurons, which is important for information processing \cite{larremorerestrepo, entropyinhibition}.
+Brunel and Wang used a model which used biologically realistic values for membrane capacitance and channel time constant to describe neurons with excitatory, GABA or NMDA/AMPA channels \cite{objectworkingmemory}.
+Using realistic proportions of these neurons to organize a module would be a nice way to make the model biologically realistic.\\
+
 \subsection*{Connectivity}
+Rubinov \textit{et al} start with 100-neuron modules which are uniformly connected, then use different connection rules to link neurons inter-module \cite{rubinov}.
+This could be a good place to start.\\
+
 \subsection*{Plasticity \& synaptic resources}
+It seems that many different papers which achieve self-organised criticality use ``dynamical variables'' of some sort (eg synapse weight, synaptic resource or gain) \cite{munozlg, rubinov, kinouchi}.
+These variables have lead to interesting dynamics by their ability to store and integrate information at a slower timescale than voltage.
+In particular, Rubinov \textit{et al} describe how dynamic synapses allow slow mesoscopic oscillations which are abolished when synapses are frozen \cite{rubinov}.
+Dynamic synapses have also been shown to be important for self-organized criticality \cite{dynamicsynapses}.
+It has also been shown that a combination of short-term plasticity (STP) and spike-time-dependent plasticity (STDP) can have synergistic effects and create very robust networks with intersting dynamics \cite{biologicalsoc}, and STDP may be important for the formation of neural communities and nonperiodic synchronicity \cite{heterogeneousnetwork}.
+A recent report showed SOC in a model with a supposedly biologically realistic dynamic ``gain'' instead of dynamic synapses, and claimed it was much faster to simulate.
+Perhaps this would be a thing to look into \cite{kinouchi}.\\
+
+It may also be important to have non-identical neurons, since this creates nonperiodicity \cite{heterogeneousnetwork}.
+
 \subsection*{Voltage equations}
+TBD, many useful papers: \cite{objectworkingmemory, rubinov, kinouchi, larremorerestrepo, entropyinhibition, dynamicsynapses, biologicalsoc, heterogeneousnetwork}
 
-\section*{TODO}
+\section*{Todo}
 \begin{itemize}
-\item Find papers on LIF cortical organization - how should we connect the modules? Should we do modules? What should they look like?
-\item Think about what sorts of plasticity rules would accomodate the Munoz 'synaptic resources' property - a way to make modules have propensity to do module spikes if sufficiently excited, and also a refractory period where the whole module must recover
-\item Are there any 'more' biophysical models which could be used to get more interesting results?
-Anything which would reproduce interesting behavior which LIF cannot, eg realistic connectivity between pyramidal and interneurons?
-\end{itemize}
-
-\textbf{coherence potentials subproject}
-\begin{itemize}
-\item Could CPs be found in 2P data?
-\item Are there any patterns in MEA CP traces suggestive of how they arise?
-\item Could large-scale synchronous dynamics in the brain, eg oscillations or avalanches, \textit{create} hopfield networks in different places? Seems unlikely since CPs are \textit{really} similar
-\item Are CPs clustered in space, or are they a small set of repeating patterns which the brain likes to make in different places?
-Does injecting a CP into a network make it more likely for it to be seen at a later time?
-\item Are there any processes which could move a conceivably-CP-creating strong hopfield network \textit{spatially} through a network?
+\item Find more models which create synchronous network states
 \end{itemize}

### 2018-10-16 test/report.tex
add stuff on 2pi data analysis

@@
 
-\section*{Objectives}
+\section*{Model Objectives}
 \begin{itemize}
@@
 It has been shown that ``tetanizing'' a network can modify it through plasticity to reproduce the tetanized signal in a phenomena known as ``ommitted stimulus potential'' \cite{heterogeneousnetwork}; perhaps this phenomena would work for other inputs
+\end{itemize}
+
+\section*{Data analysis objectives}
+\begin{itemize}
+\item 2-photon data analysis: Although the 2PI field-of-view (500x500 $\mu m^2$) is much smaller than the brain surface area monitored by an MEA grid in the original study \cite{cps2010} (1.6 x 1.6 $mm^2$), it may be possible to approximate a 2x2 MEA grid from 2PI data by partitioning the field into four 250x250 $\mu m^2$ quadrants, then using a heuristic method to estimate the local potential at the center of each quadrant.
+\item Given estimated potential, the same techniques from \cite{cps2010} could be used to find coherence potentials.
+Given knowledge of the windows each coherence potential shows up in, we could then go back to the 2PI data and answer some interesting questions:
+  \begin{itemize}
+  \item Is individual neuron firing deterministic when a CP revisits the same quadrant?
+  \item Is the combined spike raster of the whole quadrant similar for the same CP at different sites?
+  \item Are there trends in the correlation between neurons in quadrants exhibiting the same CP during or before the CP starts, suggesting info transfer?
+  \end{itemize}
+\item If this setup is impossible, are there other ways to do CP analysis on 2PI data?
+Would a z-stack volume imaging approach give more neurons?
+Can the field-of-view be broadened somehow?
+Are there other labs that do broader 2PI imaging?
 \end{itemize}
